### 2014-01-16 manuscript.tex
now compiles to pdf. math symbols, bibliography added. title fixed. can now start editing.

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage[pdftex]{graphicx}  
\usepackage{caption}
\usepackage{natbib}
\usepackage{color}
\usepackage{authblk}
\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }

%----------------------------------------
%AUTHORS
%----------------------------------------
\title{Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)}

\author[1]{Laura Vann}
\author[1]{Thomas Kon\o}
\author[1]{Tanja Pyh\"aj\"arvi}
\author[1]{Matthew Hufford}
\author[1,2]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
\affil[1]{Department of Plant Sciences, University of California Davis}
\affil[2]{Center for Population Biology and Genome Center, University of California Davis}

\date{}

\begin{document}

\maketitle

\jri{Please add addresses for all authors.  All should have plant sciences, but Kono, Tanja, and Huff need their new addresses too. }

\subsection*{ABSTRACT}

The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, was identified as a major QTL involved in branching differences between maize and its wild progenitor, teosinte. Further studies have shown that the insertion of a transposable element (Hopscotch) upstream of tb1 enhances its expression, causing the reduction in branching observed in domesticated maize. Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the Hopscotch element, led us to investigate the role of \emph{tb1} in teosinte. Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{parviglumis} and \emph{mexicana}. Analysis of linkage disequilibrium (LD) between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, yet we find no correlation between the frequency of the Hopscotch insertion and environmental variables. Additionally, we find no evidence that the Hopscotch plays a similar role in branching architecture in teosinte as it does in domesticated maize.

\section{INTRODUCTION}

Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \cite{Ross-Ibarra et al 2007}. Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \cite{Olsen and Gross 2010}. Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{Kovach and McCouch 2008}. But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.

Maize (\emph{Zea mays} ssp. \emph{mays}) was domesticated from the teosinte \emph{Zea mays} ssp. \emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 10,000 B.P. in southwest Mexico \cite{Piperno et al 2009, Matsuoka et al 2002}. Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \cite{Hufford et al 2012a, Doebley 2004, Hufford et al 2012b}. Additionally, large naturally occurring populations of both \emph{Zea mays} ssp. \emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp. \emph{mexicana} (highland teosinte) can be found throughout Mexico \cite{Wilkes 1977, Hufford et al 2013}, and genetic diversity and effective population size (Ne) of these taxa is estimated to be high \cite{Ross-Ibarra et al 2009} relative to other plants \cite{Nybom 2004}.

Many morphological changes are associated with the domestication of maize, and understanding the genetic basis for these changes has been a focus of maize research for a number of years. One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches. The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in branching differences between maize and teosinte \cite{Doebley et al 1990, Doebley and Stec 1991, Lukens and Doebley 1999}. Further studies have shown that the insertion of a 4.9 kb retrotransposon (Hopscotch) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer et al 2011}. The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in teosinte \cite{Studer et al 2011}. Dating of this element has suggested that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \cite{Studer et al 2011}. Furthermore, Studer and Doebley (2012) \cite{Studer and Doebley 2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte. In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{mexicana}, \emph{parviglumis}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the Hopscotch element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the Hopscotch element and tillering phenotypes in a population of \emph{parviglumis}.

\section{METHODS}

\subsection{Sampling and Genotyping}

We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 Zea mays ssp. mexicana (hereafter mexicana) populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (Table S1, Table S2). DNA was extracted from leaf tissue using a modified CTAB approach \cite{Doyle and Doyle 1990, Maloof et al 1984}. We designed primers using PRIMER3 \cite{Rozen and Skaletsky 2000} implemented in Geneious \cite{Kearse et al 2012} to amplify the entire Hopscotch element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the Hopscotch insertion. Two PCRs were performed for each individual, one with primers flanking the Hopscotch (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, 5'-TCGTTGATGCTTTGATGGATGG-3'; 
Hop R, 5'-AACAGTATGATTTCATGGGACCG-3'; and HopIntR, 5'-CCTCCACCTCTCATGAGATCC-3' (Figure S1). Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (Figure S2). When only one PCR resolved well, we scored one allele for the individual, which explains the odd number of alleles included in our analyses. We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98°C for 3 min, 30 cycles of 98°C for 15 s, 65°C for 30 s, and 72°C for 3 min 30 s, with a final extension of 72°C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the Hopscotch based on band size.

\subsection{Sequencing}

In addition to genotyping, we chose a subset of \emph{parviglumis} individuals for sequencing. We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B). For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of start of the Hopscotch and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2). The 5' UTR (containing sequenced region 1) has been shown to have elevated diversity in ssp. \emph{parviglumis} with respect to maize, while the the area upstream from the Hopscotch (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \cite{Wang et al 1999, Clark et al 2006}. We designed the following primers using PRIMER3 \cite{Rozen and Skaletsky 2000}: for the 5' UTR, 5' GGATAATGTGCACCAGGTGT 3' and 5' GCGTGCTAGAGACACYTGTTGCT 3'; for the 50 kb upstream region, 5' TGTCCTCGCCGCAACTC 3' and 5' TGTACGCCCGCCCCTCATCA 3' (Figure S1). We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94°C for 3 min, 30 cycles of 92°C for 40 s, annealing for 1 min, 72°C for 40 s, and a final 10 min extension at 72°C. Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7°C and 58.8°C, respectively. To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 µL of amplification product. This mix was placed on a thermal cycler with the following program: 37°C for 30 min, 80°C for 15 min, and a final cool-down step to 4°C.

We cloned cleaned fragments into a TOPO-TA vector (Invitrogen, Carlsbad) using OneShot TOP10 chemically competent E. coli cells, with an extended ligation time of 30 min for a complex target fragment. We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96°C for 5 min; then 35 cycles at 96°C for 30 s, 53°C for 30 s, 72°C for two min; and a final extension at 72°C for 4 min. We visualized amplification products for incorporation of our insert on a 1\% agarose TAE gel.

Amplification products with successful incorporation of our insert were cleaned using Exonuclease 1 and Antarctic Phosphatase following the procedures detailed above, and sequenced with vector primers M13 Forward and M13 Reverse using Sanger sequencing at the College of Agriculture and Environmental Sciences (CAES) sequencing center at UC Davis. We aligned and trimmed primer sequences from resulting sequences using the software Geneious \cite{Kearse et al 2012}. Following alignment, we verified singleton SNPs by sequencing an additional one to four colonies from each clone. If the singleton was not present in these additional sequences it was considered an amplification or cloning error, and we replaced the base with the base of the additional sequences. If the singleton appeared in at least one of the additional sequences we considered it a real variant and kept it for further analyses.

\subsection{Genotyping Analysis}

We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE), and to calculate differentiation between populations (FST) and subspecies (FCT) we used HierFstat \cite{Goudet 2005}. These analyses only included populations in which 8 or more individuals were sampled. To test the hypothesis that the Hopscotch insertion may be adaptive under certain environmental conditions, we looked for significant associations between the Hopscotch frequency and environmental variables using BayEnv \cite{Coop et al 2010}. BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\cite{Coop et al 2010}. We used genotyping and covariance data from Pyhäjärvi et al., (2013) \cite{Pyhajarvi et al 2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (Table S3). Environmental data were obtained from www.worldclim.org, the Harmonized World Soil Database and www.harvestchoice.org, and summarized by principle component analysis Pyhäjärvi et al. (2013) \cite{Pyhajarvi et al 2013}.

\subsection{Sequence Analysis}

For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \cite{Thornton 2003} to calculate pairwise FST between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator$\theta_{W}$; pairwise nucleotide diversity $\theta_{\pi}$; and Tajima's D). To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees. For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the Hopscotch insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \cite{Chia et al 2012}).

In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \cite{Bradbury et al 2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A, Ejutla B, San Lorenzo, and La Mesa populations \cite{Pyhajarvi et al 2013}. We chose these populations because we had both genotyping data for the Hopscotch as well as chromosome-wide SNP data for chromosome 1. For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for San Lorenzo, Ejutla B, Ejutla A, and La Mesa, respectively. We then used Tassel \cite{Bradbury et al 2007} to calculate linkage disequilibrium ($r^{2}$) across chromosome 1 for each population.

We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \cite{Falush et al 2003} and the same phased 55K SNP data from \cite{Pyhajarvi et al 2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \cite{Cook et al 2012}. SNPs were anchored in a modified version of the IBM genetic map ( \cite{Gerke et al 2013}, \url{http://arxiv.org/abs/1307.7313}). We created haplotype blocks using a custom Perl script that grouped all SNPs within a 5 kb window into a haplotype. We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.

\subsection{Phenotyping of \emph{Zea mays.} ssp. \emph{parviglumis}}

To investigate the phenotypic effects of the Hopscotch insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1). We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \cite{Hufford 2010} where the Hopscotch is segregating at highest frequency (0.4375) in our initial genotyping sample set. In order to maximize the likelihood of finding the Hopscotch in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion. We chose between 10-13 seeds from each of 23 sampling sites. We treated seeds with fungicide and germinated them in petri dishes with filter paper. Successful germinations (206 individuals) were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches. Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.

Starting on day 15, we measured tillering index, the ratio of the sum of tiller lengths to the height of the plant \cite{Briggs et al 2007}. Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \cite{Briggs et al 2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Culm diameter is not believed to be correlated with tillering index, or variation at \emph{tb1} (e.g. Hopscotch genotype). Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (http://www.ars.usda.gov). We genotyped individuals for the Hopscotch insertion following the protocols listed above. Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the Hopscotch on tillering index.

We performed a second phenotyping experiment (phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the Hopscotch insertion. Seeds were germinated and planted on day 7 post fruit-case removal into 2 gallon pots. Plants were watered twice daily, alternating between fertilized and non-fertilized water. We began phenotyping successful germinations (302) for tillering index on day 15 post fruit case removal, and phenotyped every five days until day 50. At day 50 we measured culm diameter between the third and fourth nodes. We extracted DNA and genotyped plants following the same guidelines as in phenotyping 1.

Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \cite{Fox and Weisberg 2011} to improve the normality and homogeneity of variance among genotype classes.  We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA). Additionally, in order to compare any association between Hopscotch genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.

\section{Results}

\subsection{Genotyping}

We genotyped 247 accessions of maize landraces, and found the Hopscotch element fixed in all but 8 of these accessions (inbred lines X,Y, Z, and landraces in Chihuahua, Tapalpa, Tenancingo, and Altotonga) (Table S1, Table S2). For all analyses we only included individuals that clearly resolved for both PCRs, resulting in 837 individuals total. Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoacán in central-western Mexico in both subspecies (Figure 1). We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population. Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p=0.05).

Using our Hopscotch genotyping data, we calculated differentiation between populations (FST) and subspecies (FCT) for populations in which we sampled 8 or more alleles. FCT is 0 within our dataset, and we found similar levels of FST between populations within each subspecies (0.22) and between all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi et al 2013} (Table 1). Although we found large variation in Hopscotch allele frequency among our populations, our BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors < 1; Table S3).

\subsection{Sequencing}

To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the tb1 region, we sequenced two small (<1kb) regions upstream of the tb1 ORF in four populations. After alignment and singleton checking we recovered 40 and 48 segregating sites for the 50kb upstream region and the 5' UTR region, respectively. For region 1, Ejutla A, Ejutla B, and La Mesa have comparable values for haplotype diversity,$\theta_{W}$, $\theta_{\pi}$, and Tajima's D, while San Lorenzo has lower values for haplotype diversity,$\theta_{W}$, $\theta_{\pi}$, and a more positive value for Tajima's D (Table 2). For region 2, haplotype diversity,$\theta_{W}$, and $\theta_{\pi}$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table 2). Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this latter population (Table 2). Pairwise values of FST within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (Table 1). Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILS; data from Maize HapMapV2, \cite{Chia et al 2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (Figure S3). Furthermore, individuals within our sample that have the Hopscotch insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the Hopscotch insertion.

\subsection{Evidence of introgression}

The teosinte populations with the highest frequency of the Hopscotch insertion in this study were sympatric with cultivated maize. Our initial hypothesis was that the high frequency of the Hopscotch element in these populations could be attributed to introgression from maize into teosinte. To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one, and specifically in the \emph{tb1} region. If the Hopscotch is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element. We find no evidence of elevated linkage disequilibrium between the Hopscotch and SNPs surrounding the tb1 region in our resequenced populations (Figure 2), and $r^{2}$ in the \emph{tb1} region (positions 264,596,664-265,891,456; AGPv2) does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the Hopscotch genotype (average $r^{2}=0.082$). In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (Table 3).

The lack of clustering of Hopscotch genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the Hopscotch insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \cite{Cook et al 2012, Phyhajarvi et al 2013}. The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations. If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is lower (average 0.017) than the chromosome wide average (0.20). (Figure 3).

\subsection{Phenotyping}

\subsubsection{Phenotyping 1}

We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from 23 sampling locations within the San Lorenzo population, and genotyped plants for the Hopscotch insertion. We found the Hopscotch segregating at a frequency of 0.65; with a 0.4 frequency of homozygotes for the Hopscotch, a 0.5 frequency of heterozygotes, and a 0.1 frequency of homozygotes for the teosinte allele (no Hopscotch), showing no significant deviations from expected frequencies under Hardy-Weinberg equilibrium. After performing a repeated measures ANOVA between our transformed tillering index data and Hopscotch genotype we find a weak positive correlation between presence of the Hopscotch and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (Figure 4). Additionally we find no significant correlation between tiller number and Hopscotch genotype, or culm diameter and Hopscotch genotype in phenotyping 1.

\subsection{Phenotyping 2}

For phenotyping 2 we measured tillering index every five days through day 50 for 302 plants. We followed the same transformations and analyses as described in phenotyping 1. We find the Hopscotch allele segregating at a frequency of 0.69, with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele. We find similar patterns as in phenotyping 1, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p=0.0611), with no significant correlation on any day. Similarly, relationships between Hopscotch genotype and tiller number, and hopscotch genotype and culm diameter are not significant.

\section{DISCUSSION}

Adaptation occurs either due to selection on standing variation or on de novo mutations. Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection on lactase tolerance in humans \cite{Plantinga et al 2012, Tishkoff et al 2007}; variation at the Eda locus in three-spined stickleback \cite{Kitano et al 2008, Colosimo et al 2005}; and pupal diapause in the Apple Maggot fly \cite{Feder et al 2003}. Although the role of standing variation with respect to adaptation has been described in many systems, its importance to domestication is not as well studied.

In maize, alleles at important domestication loci (RAMOSA1, \cite{Sigmon and Vollbrecht 2010}; barren stalk1, \cite{Gallavotti et al 2004}; and grassy tillers1, \cite{Whipple et al 2011} have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize. The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \cite{Clark et al 2006, Studer et al 2011}. Studer et al. (2011) genotyped 90 accessions of teosinte (inbred and non-inbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \cite{Studer et al 2011}. Our findings suggest that the Hopscotch insertion is more widespread in populations of teosinte than previously thought and that the presence of the insertion is not due to recent introgression. However, we find no effect of the Hopscotch insertion on branching architecture in teosinte.

Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it is still segregating in populations of teosinte. However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the Hopscotch in teosinte. While our findings are consistent with a previous study by Studer et al. (2011) in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \cite{Studer et al 2011}. Crop to wild introgression has previously been observed at domestication loci (\cite{Ellstrand et al 1999, Zhang et al 2009, Thurber et al 2010, Baack et al 2008, Hubner et al 2012, Wilkes et al 1977, Van Heerwaarden et al 2011, Barrett 1983}), but our results are more consistent with Hufford et al. (2013) who found resistance to introgression from maize into teosinte \cite{Hufford et al 2013}. Furthermore, Hufford et al. (2013) showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).

We find no evidence of recent introgression in our analyses. Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the tb1 locus had introgressed into populations of teosinte. We do not have sympatric maize samples from many of our populations, however, so we cannot rule out the possibility that Hopscotch alleles in these teosinte populations are segregating in sympatric maize. On the other hand, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1. The similarity in LD in the \emph{tb1} region in populations with and without the Hopscotch provides further support for an explanation other than introgression. Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (Figure 3, Table 4).

Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the Hopscotch in these populations. In the ancient introgression scenario, recombination would have broken up linked variants surrounding the Hopscotch and we would find LD patterns similar to the ones we report here. However, it seems unlikely that there would have been ancient introgression between these populations of teosinte and their sympatric maize populations and not present-day introgression. Studies have demonstrated the presence of ongoing gene flow between domesticated maize and both \emph{Zea mays} ssp. \emph{mexicana} and \emph{Zea mays} ssp. \emph{parviglumis} in a number of sympatric populations \cite{Hufford et al 2013, Ellstrand et al 2007, van Heerwaarden et al 2011}.

The lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D in the San Lorenzo population are consistent with patterns expected under neutrality. Though, deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus. Similarly, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.

The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \cite{Clark et al 2006, Studer et al 2011}, and Weber et al. (2007) have described its effects in partial inbred lines of teosinte \cite{Weber et al 2007}. Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in a population of teosinte. However, we found no significant effect of the Hopscotch on tillering index or tiller number in our phenotyping experiments, and the effect of the Hopscotch insertion in teosinte is discordant with that of maize. The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize, and preliminary data collected on a grow-out of the TILs (teosinte inbred lines).

It is certainly possible that even though data from the TILs suggest an effect of the Hopscotch on tillering in teosinte, that it in fact has no effect. Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \cite{Clark et al 2006}, and perhaps we would have detected an effect of the Hopscotch on these additional phenotypes. A recent study by Studer and Doebley (2012) examined the possibility of an allelic series at the \emph{tb1} locus in teosinte by introgressing 9 separate teosinte segments spanning the \emph{tb1} locus into an isogenic maize background and investigating their effects on phenotype. However, they found evidence for an allelic series contributing to all \emph{tb1} associated phenotypes but tillering. They did find variation in tillering both between introgressed lines and between isogenic maize lines, suggesting that there may be additional loci that contribute to tillering in both maize and teosinte \cite{Studer and Doebley 2012}. It seems plausible that variation in some of these other loci could have affected tillering in our greenhouse population, and contributed to the lack of correlation we see between Hopscotch genotype and tillering.

Another explanation for the lack of correlation between Hopscotch genotype and tillering could be growing conditions in out greenhouse affected growth habit of our plants. The \emph{tb1} gene is part of the shade avoidance pathway in maize, and this involvement may have presented problems for our phenotyping design \citep{Kebrom and Brutnell 2007}. Broadly, shade avoidance syndrome (SAS), refers to the morphological and physiological response of plants experiencing an increase in far-red (FR) light, which occurs during shading \cite{Kebrom and Brutnell 2007}. Under increased FR light conditions plants experience both morphological and physiological changes including reduction in branching, increased height, and increased ethylene and auxin production \cite{Ballare 1999, Smith and Whitelam 1997}. In maize, increase in shading has been shown to be correlated with changes in \emph{tb1} expression and subsequent reduction in tillering, suggesting that \emph{tb1} is involved in the shade avoidance response in maize \cite{Doebley et al 1997, Lukens and Doebley 1999}. If plants in our phenotyping experiments were grown too close together, such that they shaded each other, it could have reduced \emph{tb1} expression and tillering, which would have confounded our results.

In summary, our findings demonstrate that the Hopscotch allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought. Analysis of linkage using SNPs from across chromosome 1 does not suggest that the Hopscotch allele is present in these populations due to recent introgression; however, it seems unlikely that the it would have drifted to high frequency in multiple populations, and there may be another explanation for the high frequency we observe in some of our populations. The Hopscotch insertion does not appear to have any affect on tillering phenotypes in teosinte previously associated with \emph{tb1} and it is possible that variation at other loci may play a larger role in branching architecture in teosinte than in domesticated maize.

\clearpage
\bibliography{refs}
\bibliographystyle{apalike}

\end{document}

### 2014-01-16 manuscript.tex
added figure example, edits to tex

@@
 \usepackage{natbib}
+\bibliographystyle{apalike}
 \usepackage{color}
@@
 
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, was identified as a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
+The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
 Further studies have shown that the insertion of a transposable element (Hopscotch) upstream of tb1 enhances its expression, causing the reduction in branching observed in domesticated maize.
@@
 Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{parviglumis} and \emph{mexicana}.
-Analysis of linkage disequilibrium (LD) between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, yet we find no correlation between the frequency of the Hopscotch insertion and environmental variables.
-Additionally, we find no evidence that the Hopscotch plays a similar role in branching architecture in teosinte as it does in domesticated maize.
+Analysis of linkage disequilibrium between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
+Finally, two greenhouse experiments fail to find an important role for \emph{tb1} in controlling tillering in natural populations of \emph{parvilgumis}.
+\jri{Needs a concluding sentence.
+"Thus, we think..." or something.}
 
-\section{INTRODUCTION}
+\section*{Introduction}
+\jri{plase redo sections not in all caps}
 
@@
 But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
+
+% SFS figure
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{SFS.png}
+    \caption{This is an example figure.} 
+\label{sfs_non_syn}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
+
+\jri{This is how you cite the figure in the text.
+"I want to show you all Figure \ref{sfs_non_syn}." Latex takes care of the numbering.
+Please add all the figures to the text, and add them in in the middle like this close to where they are referenced.
+Note that it puts the figure where it thinks best, which is not necessarily near this text.}
 
@@
 Dating of this element has suggested that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \cite{Studer et al 2011}.
-Furthermore, Studer and Doebley (2012) \cite{Studer and Doebley 2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
+Furthermore, \cite{Studer and Doebley 2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
 In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{mexicana}, \emph{parviglumis}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
@@
 
-\section{METHODS}
+\section*{METHODS}
 
-\subsection{Sampling and Genotyping}
+\subsection*{Sampling and Genotyping}
 
@@
 mexicana (hereafter mexicana) populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (Table S1, Table S2).
-DNA was extracted from leaf tissue using a modified CTAB approach \cite{Doyle and Doyle 1990, Maloof et al 1984}.
+\jri{please add tables and reference same as figs; see example in Sofiane's "deleterious 282" repo on lab github page} DNA was extracted from leaf tissue using a modified CTAB approach \cite{Doyle and Doyle 1990, Maloof et al 1984}.
 We designed primers using PRIMER3 \cite{Rozen and Skaletsky 2000} implemented in Geneious \cite{Kearse et al 2012} to amplify the entire Hopscotch element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the Hopscotch insertion.
@@
 Hop R, 5'-AACAGTATGATTTCATGGGACCG-3'; and HopIntR, 5'-CCTCCACCTCTCATGAGATCC-3' (Figure S1).
-Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (Figure S2).
+\jri{supplemental material should be added to this file too.} Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (Figure S2).
 When only one PCR resolved well, we scored one allele for the individual, which explains the odd number of alleles included in our analyses.
@@
 
-\subsection{Sequencing}
+\subsection*{Sequencing}
 
@@
 
-\subsection{Genotyping Analysis}
+\subsection*{Genotyping Analysis}
 
@@
 
-\subsection{Sequence Analysis}
+\subsection*{Sequence Analysis}
 
@@
 
-\subsection{Phenotyping of \emph{Zea mays.} ssp. \emph{parviglumis}}
+\subsection*{Phenotyping of \emph{Zea mays.} ssp. \emph{parviglumis}}
 
@@
 
-\section{Results}
+\section*{Results}
 
-\subsection{Genotyping}
+\subsection*{Genotyping}
 
@@
 
-\subsection{Evidence of introgression}
+\subsection*{Evidence of introgression}
 
@@
 
-\subsection{Phenotyping}
+\subsection*{Phenotyping}
 
-\subsubsection{Phenotyping 1}
+\subsubsection*{Phenotyping 1}
 
@@
 
-\subsection{Phenotyping 2}
+\subsection*{Phenotyping 2}
 
@@
 
-\section{DISCUSSION}
+\section*{DISCUSSION}
 
 Adaptation occurs either due to selection on standing variation or on de novo mutations.
-Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection on lactase tolerance in humans \cite{Plantinga et al 2012, Tishkoff et al 2007}; variation at the Eda locus in three-spined stickleback \cite{Kitano et al 2008, Colosimo et al 2005}; and pupal diapause in the Apple Maggot fly \cite{Feder et al 2003}.
+Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection on lactase tolerance in humans \cite{Plantinga et al 2012, Tishkoff et al 2007}; variation at the \jri{italics?} Eda locus in three-spined stickleback \cite{Kitano et al 2008, Colosimo et al 2005}; and pupal diapause in the Apple Maggot fly \cite{Feder et al 2003}.
 Although the role of standing variation with respect to adaptation has been described in many systems, its importance to domestication is not as well studied.
 
-In maize, alleles at important domestication loci (RAMOSA1, \cite{Sigmon and Vollbrecht 2010}; barren stalk1, \cite{Gallavotti et al 2004}; and grassy tillers1, \cite{Whipple et al 2011} have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
+In maize, alleles at important domestication loci (RAMOSA1, \cite{Sigmon and Vollbrecht 2010}; barren stalk1, \cite{Gallavotti et al 2004}; and \jri{do gene names need to be italicized?} grassy tillers1, \cite{Whipple et al 2011} have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
 The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \cite{Clark et al 2006, Studer et al 2011}.
@@
 (2011) genotyped 90 accessions of teosinte (inbred and non-inbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \cite{Studer et al 2011}.
-Our findings suggest that the Hopscotch insertion is more widespread in populations of teosinte than previously thought and that the presence of the insertion is not due to recent introgression.
-However, we find no effect of the Hopscotch insertion on branching architecture in teosinte.
 
@@
 (2013) who found resistance to introgression from maize into teosinte \cite{Hufford et al 2013}.
-Furthermore, Hufford et al.
+Furthermore, \jri{all citations, even citations like these need to reference the bibtex file.
+check out the difference betweetn cite, citet, and citep as commands.} Hufford et al.
 (2013) showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).
@@
 We find no evidence of recent introgression in our analyses.
-Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the tb1 locus had introgressed into populations of teosinte.
+Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
 We do not have sympatric maize samples from many of our populations, however, so we cannot rule out the possibility that Hopscotch alleles in these teosinte populations are segregating in sympatric maize.
@@
 Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (Figure 3, Table 4).
+\jri{figure references need to be as in the example using the ref command}
 
@@
 \emph{parviglumis} in a number of sympatric populations \cite{Hufford et al 2013, Ellstrand et al 2007, van Heerwaarden et al 2011}.
+\jri{check these refs.
+Hufford 2013 is mexicana, as is i think Ellstrand.
+do we show introgression with parv in van Heerwaarden?.
+do is how it in my 2009 paper?}
 
 The lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D in the San Lorenzo population are consistent with patterns expected under neutrality.
-Though, deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
-Similarly, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
+\jri{lower diversity is consistent with neutrality?} Though, deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
+\jri{shouldn't Tanja's data be able to argue one way or the other here?
+she looked at eveidence of inbreeding.
+what did she find in san lorenzo?} Similarly, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
 
 The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \cite{Clark et al 2006, Studer et al 2011}, and Weber et al.
-(2007) have described its effects in partial inbred lines of teosinte \cite{Weber et al 2007}.
-Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in a population of teosinte.
+(2007) have described its effects in partially inbred lines of teosinte \cite{Weber et al 2007}.
+Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.
 However, we found no significant effect of the Hopscotch on tillering index or tiller number in our phenotyping experiments, and the effect of the Hopscotch insertion in teosinte is discordant with that of maize.
-The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize, and preliminary data collected on a grow-out of the TILs (teosinte inbred lines).
+The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize, as well as our own preliminary data collected on a grow-out of teosinte inbred lines.
+\jri{Huff, think we should include the TIL data just to show we can see effects in our greenhouse?}
 
 It is certainly possible that even though data from the TILs suggest an effect of the Hopscotch on tillering in teosinte, that it in fact has no effect.
-Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \cite{Clark et al 2006}, and perhaps we would have detected an effect of the Hopscotch on these additional phenotypes.
-A recent study by Studer and Doebley (2012) examined the possibility of an allelic series at the \emph{tb1} locus in teosinte by introgressing 9 separate teosinte segments spanning the \emph{tb1} locus into an isogenic maize background and investigating their effects on phenotype.
-However, they found evidence for an allelic series contributing to all \emph{tb1} associated phenotypes but tillering.
-They did find variation in tillering both between introgressed lines and between isogenic maize lines, suggesting that there may be additional loci that contribute to tillering in both maize and teosinte \cite{Studer and Doebley 2012}.
+\jri{i don't think we want to argue that tb1 has no effect!!
+we have no evidence for this.
+instead, it may be that the effect is hard to see if there are numerous other loci and/or epistasis, etc.} Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \cite{Clark et al 2006}, and perhaps we would have detected an effect of the Hopscotch on these additional phenotypes.
+\jri{why say we might have detected these?
+we didn't look for them, not sure worth mentioning}  A recent study by Studer and Doebley (2012) examined the possibility of an allelic series at the \emph{tb1} locus in teosinte by introgressing 9 separate teosinte segments spanning the \emph{tb1} locus into an isogenic maize background and investigating their effects on phenotype.
+However, they found evidence for an allelic series contributing to all \emph{tb1} associated phenotypes but tillering .
+They did find variation in tillering both between introgressed lines and between isogenic maize lines \jri{don't follow.
+you mean the different alleles didn't have different tillering phenotypes?
+but they did show that all alleles affected tillering, right?}, suggesting that there may be additional loci that contribute to tillering in both maize and teosinte \cite{Studer and Doebley 2012}.
 It seems plausible that variation in some of these other loci could have affected tillering in our greenhouse population, and contributed to the lack of correlation we see between Hopscotch genotype and tillering.
+\jri{are there other studies on other loci?
+what about Briggs?
+original Doebley QTL papers?
+one of the weber papers maybe?
+seems like there should be explicit cites to other loci we know are involved.
+also we could cite \% variance explained by tb1 of tillering in the maize x teo crosses, arguing that there must be other loci.}
 
-Another explanation for the lack of correlation between Hopscotch genotype and tillering could be growing conditions in out greenhouse affected growth habit of our plants.
+Another explanation for the lack of correlation between Hopscotch genotype and tillering could be that growing conditions in our greenhouse affected growth habit of our plants.
 The \emph{tb1} gene is part of the shade avoidance pathway in maize, and this involvement may have presented problems for our phenotyping design \citep{Kebrom and Brutnell 2007}.
@@
 If plants in our phenotyping experiments were grown too close together, such that they shaded each other, it could have reduced \emph{tb1} expression and tillering, which would have confounded our results.
+\jri{maybe, but we saw variation in tillering, especially in growout number 1.
+shade would have to so effect expression that it didn't matter what tb1 genotype a plant had -- so G x E effects.
+is it known whether there are such G x E effects?}
 
@@
 The Hopscotch insertion does not appear to have any affect on tillering phenotypes in teosinte previously associated with \emph{tb1} and it is possible that variation at other loci may play a larger role in branching architecture in teosinte than in domesticated maize.
+\jri{seems like we should end with at least a hypothesis.
+as is we are saying we ruled out drift, introgression, and selection.
+what else is there?
+perhaps we should think about selection or introgression scenarios that could explain our data.}
 
@@
 \bibliography{refs}
-\bibliographystyle{apalike}
 

### 2014-01-16 manuscript.tex
edits to tex

@@
 \documentclass[12pt]{article}
+\usepackage{geometry} 
 \usepackage{hyperref}
+\usepackage{color}
 \usepackage[pdftex]{graphicx}  
-\usepackage{caption}
+%\usepackage{caption}
 \usepackage{natbib}
+\usepackage{authblk}
+\renewcommand{\baselinestretch}{1.5}
 \bibliographystyle{apalike}
-\usepackage{color}
-\usepackage{authblk}
+
 \newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
@@
 \author[1]{Laura Vann}
-\author[1]{Thomas Kon\o}
+\author[1]{Thomas Kono}
 \author[1]{Tanja Pyh\"aj\"arvi}
@@
 \affil[2]{Center for Population Biology and Genome Center, University of California Davis}
-
-\date{}
 
@@
 
-\subsection*{ABSTRACT}
+\clearpage
+
+\subsection*{Abstract}
 
@@
 
+\clearpage
+
 \section*{Introduction}
-\jri{plase redo sections not in all caps}
 
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \cite{Ross-Ibarra et al 2007}.
-Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \cite{Olsen and Gross 2010}.
-Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{Kovach and McCouch 2008}.
+Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \cite{OlsenGross2010}.
+\jri{I believe you need to remove spaces from the bibtex citation names.
+See how this OlsenGross2010 works but the others don't?
+should be able to work with whitespace but i can't figure out why not.}  Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{Kovach and McCouch 2008}.
 But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
@@
 
-\section*{METHODS}
+\section*{Methods}
 
@@
 
-\subsection*{Phenotyping 2}
+\subsubsection*{Phenotyping 2}
 
@@
 
-\section*{DISCUSSION}
+\section*{Discussion}
 
@@
 Furthermore, \jri{all citations, even citations like these need to reference the bibtex file.
-check out the difference betweetn cite, citet, and citep as commands.} Hufford et al.
+check out the difference betweetn cite, citet, and citep as commands.
+see \url{http://merkel.zoneo.net/Latex/natbib.php} for details} Hufford et al.
 (2013) showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).
@@
 \clearpage
-\bibliography{refs}
+\bibliography{refs,bob}
 

### 2014-02-16 manuscript.tex
edits throughout manuscript, comments for revamping discussion

@@
 \usepackage{authblk}
+\usepackage{subscript}
 \renewcommand{\baselinestretch}{1.5}
@@
 \newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
+\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
 
@@
 \author[1]{Tanja Pyh\"aj\"arvi}
-\author[1]{Matthew Hufford}
+\author[1]{Matthew B. Hufford}
 \author[1,2]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
@@
 \jri{Please add addresses for all authors.  All should have plant sciences, but Kono, Tanja, and Huff need their new addresses too. }
+
+\mbh{my new address is: Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, IA, 50011}
 
@@
 Furthermore, \cite{Studer and Doebley 2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
-In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{mexicana}, \emph{parviglumis}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
+\mbh{A sentence or two is needed here explaining why tillering may be an ecologically relevant trait to a wild plant} In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{mexicana}, \emph{parviglumis}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
 We use a combination of PCR genotyping for the Hopscotch element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus.
@@
 We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of start of the Hopscotch and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
+For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of the start of the Hopscotch and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
 The 5' UTR (containing sequenced region 1) has been shown to have elevated diversity in ssp.
@@
 
-We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE), and to calculate differentiation between populations (FST) and subspecies (FCT) we used HierFstat \cite{Goudet 2005}.
+We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE), and to calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \cite{Goudet 2005}.
 These analyses only included populations in which 8 or more individuals were sampled.
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \cite{Thornton 2003} to calculate pairwise FST between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator$\theta_{W}$; pairwise nucleotide diversity $\theta_{\pi}$; and Tajima's D).
+For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \cite{Thornton 2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_{W}$; pairwise nucleotide diversity $\theta_{\pi}$; and Tajima's D).
 To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
@@
 SNPs were anchored in a modified version of the IBM genetic map ( \cite{Gerke et al 2013}, \url{http://arxiv.org/abs/1307.7313}).
-We created haplotype blocks using a custom Perl script that grouped all SNPs within a 5 kb window into a haplotype.
+We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 For all analyses we only included individuals that clearly resolved for both PCRs, resulting in 837 individuals total.
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoacán in central-western Mexico in both subspecies (Figure 1).
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (Figure 1).
 We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
 Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p=0.05).
+\mbh{not sure what p=0.05 means here.
+Should it be p $<$ 0.05 or $\alpha$ = 0.05?}
 
-Using our Hopscotch genotyping data, we calculated differentiation between populations (FST) and subspecies (FCT) for populations in which we sampled 8 or more alleles.
-FCT is 0 within our dataset, and we found similar levels of FST between populations within each subspecies (0.22) and between all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi et al 2013} (Table 1).
-Although we found large variation in Hopscotch allele frequency among our populations, our BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors < 1; Table S3).
+Using our Hopscotch genotyping data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+F\textsubscript{CT} is 0 within our dataset, and we found similar levels of F\textsubscript{ST} between populations within each subspecies (0.22) and between all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi et al 2013} (Table 1).
+Although we found large variation in Hopscotch allele frequency among our populations, our BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors $<$ 1; Table S3).
 
@@
 
-To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the tb1 region, we sequenced two small (<1kb) regions upstream of the tb1 ORF in four populations.
+To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the tb1 region, we sequenced two small ($<$1kb) regions upstream of the tb1 ORF in four populations.
 After alignment and singleton checking we recovered 40 and 48 segregating sites for the 50kb upstream region and the 5' UTR region, respectively.
-For region 1, Ejutla A, Ejutla B, and La Mesa have comparable values for haplotype diversity,$\theta_{W}$, $\theta_{\pi}$, and Tajima's D, while San Lorenzo has lower values for haplotype diversity,$\theta_{W}$, $\theta_{\pi}$, and a more positive value for Tajima's D (Table 2).
-For region 2, haplotype diversity,$\theta_{W}$, and $\theta_{\pi}$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table 2).
+For region 1, Ejutla A, Ejutla B, and La Mesa have comparable values for haplotype diversity, $\theta_{W}$, $\theta_{\pi}$, and Tajima's D, while San Lorenzo has lower values for haplotype diversity, $\theta_{W}$, $\theta_{\pi}$, and a more positive value for Tajima's D (Table 2).
+For region 2, haplotype diversity, $\theta_{W}$, and $\theta_{\pi}$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table 2).
 Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this latter population (Table 2).
-Pairwise values of FST within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (Table 1).
+\mbh{this text hasn't changed since the last time I read the manuscript and previously did not agree with the results presented in Table 2.
+Was Table 2 incorrect?} Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (Table 1).
 Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILS; data from Maize HapMapV2, \cite{Chia et al 2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (Figure S3).
@@
 If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is lower (average 0.017) than the chromosome wide average (0.20).
-(Figure 3).
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is lower (average 0.017) than the chromosome wide average (0.20; Figure 3).
 
@@
 
-Adaptation occurs either due to selection on standing variation or on de novo mutations.
-Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection on lactase tolerance in humans \cite{Plantinga et al 2012, Tishkoff et al 2007}; variation at the \jri{italics?} Eda locus in three-spined stickleback \cite{Kitano et al 2008, Colosimo et al 2005}; and pupal diapause in the Apple Maggot fly \cite{Feder et al 2003}.
+Adaptation occurs either due to selection on standing variation or on \emph{de novo} mutations.
+Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection for lactose tolerance in humans \cite{Plantinga et al 2012, Tishkoff et al 2007}; variation at the \jri{italics?} Eda locus in three-spined stickleback \cite{Kitano et al 2008, Colosimo et al 2005}; and pupal diapause in the Apple Maggot fly \cite{Feder et al 2003}.
 Although the role of standing variation with respect to adaptation has been described in many systems, its importance to domestication is not as well studied.
 
-In maize, alleles at important domestication loci (RAMOSA1, \cite{Sigmon and Vollbrecht 2010}; barren stalk1, \cite{Gallavotti et al 2004}; and \jri{do gene names need to be italicized?} grassy tillers1, \cite{Whipple et al 2011} have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
+In maize, alleles at important domestication loci (RAMOSA1, \cite{Sigmon and Vollbrecht 2010}; barren stalk1, \cite{Gallavotti et al 2004}; and \jri{do gene names need to be italicized?} \mbh{yup, they need to be italicized} grassy tillers1, \cite{Whipple et al 2011}) have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
 The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \cite{Clark et al 2006, Studer et al 2011}.
 Studer et al.
-(2011) genotyped 90 accessions of teosinte (inbred and non-inbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \cite{Studer et al 2011}.
+(2011) genotyped 90 accessions of teosinte (inbred and outbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \cite{Studer et al 2011}.
 
-Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it is still segregating in populations of teosinte.
+Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
 However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the Hopscotch in teosinte.
@@
 We do not have sympatric maize samples from many of our populations, however, so we cannot rule out the possibility that Hopscotch alleles in these teosinte populations are segregating in sympatric maize.
-On the other hand, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1.
+Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1.
 The similarity in LD in the \emph{tb1} region in populations with and without the Hopscotch provides further support for an explanation other than introgression.
@@
 do we show introgression with parv in van Heerwaarden?.
-do is how it in my 2009 paper?}
+how does it look in my 2009 paper?} \mbh{there is a low level of introgression between maize and \emph{parviglumis} based on the results from van Heerwaarden, much less than between \emph{mexicana} and maize}
 
 The lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D in the San Lorenzo population are consistent with patterns expected under neutrality.
-\jri{lower diversity is consistent with neutrality?} Though, deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
+\jri{lower diversity is consistent with neutrality?} \mbh{this is probably coming from my last round of edits; LV, my comment was that the Taj D results in Table 2 were more consistent with neutrality in San Lorenzo rather than weak selection as it was previously written in the manuscript.
+Perhaps the diversity is low because this is just a low-diversity population.
+My SSR data showed evidence of a pop.
+bottleneck in SLO...think that's what we're seeing at tb1 too} Though, deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
 \jri{shouldn't Tanja's data be able to argue one way or the other here?
 she looked at eveidence of inbreeding.
-what did she find in san lorenzo?} Similarly, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
+what did she find in san lorenzo?} \mbh{she found evidence of inbreeding in SLO as did I; you can cite both her paper and my dissertation for this if you'd like} Similarly, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
+\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the hopscotch allele.
+Why?
+Only ancient introgression in Jalisco and not in the Balsas?
+If hopscotch came from standing variation, is this more consistent with a Jalisco domestication?
+I know, heresy!!
+LV, can you look at the \emph{parviglumis} populations with high hopscotch and see if they are all in the Jalisco cluster of Fukunaga?
+Might be a nice observation we could make here...could even go so far as saying the results from Matsuoka deserve another look.}
 
@@
 The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize, as well as our own preliminary data collected on a grow-out of teosinte inbred lines.
-\jri{Huff, think we should include the TIL data just to show we can see effects in our greenhouse?}
+\jri{Huff, think we should include the TIL data just to show we can see effects in our greenhouse?} \mbh{Not sure...we would need to explain the opposite trends; in the grow-out Tom and I did, the lines with the hopscotch had reduced tillering, in LV's grow-outs, lines with the hopscotch had increased tillering.
+Jeff, is there expression data for TIL03,09,17 maybe from John Doebley?
+Two of them are in the Springer data set, but tb1 is missing from that data set.
+Would be nice if we were able to say these TILs had maize-level expression at tb1 and we saw phenotypic differences; could suggest follow up studies are needed to assess expression of tb1 in San Lorenzo; what would be really awesome would be to extract RNA from five or ten SLO plants and the TILs with hopscotch and measure expression at tb1 and maybe zfl2 based on the results from Weber et al....probably more than we want to do for this pub.}
 
@@
 we have no evidence for this.
-instead, it may be that the effect is hard to see if there are numerous other loci and/or epistasis, etc.} Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \cite{Clark et al 2006}, and perhaps we would have detected an effect of the Hopscotch on these additional phenotypes.
+instead, it may be that the effect is hard to see if there are numerous other loci and/or epistasis, etc.} \mbh{I think we could say that \emph{tb1} is only one gene in a complex pathway affecting the quantitative branching and tillering traits and that based on alleles at other loci, the hopscotch might not consistently reduce tillering.
+Also, please be mindful that currently there is no discussion of the tillering results from the TILs in the manuscript.
+You'll need to report those results to include this discussion.} Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \cite{Clark et al 2006}, and perhaps we would have detected an effect of the Hopscotch on these additional phenotypes.
 \jri{why say we might have detected these?
-we didn't look for them, not sure worth mentioning}  A recent study by Studer and Doebley (2012) examined the possibility of an allelic series at the \emph{tb1} locus in teosinte by introgressing 9 separate teosinte segments spanning the \emph{tb1} locus into an isogenic maize background and investigating their effects on phenotype.
+we didn't look for them, not sure worth mentioning}  \mbh{has the hopscotch been shown to contribute to other phenotypes?
+If so, we could hypothesize it is relevant to teosinte's ecology via some of these less studied traits?} A recent study by Studer and Doebley (2012) examined the possibility of an allelic series at the \emph{tb1} locus in teosinte by introgressing 9 separate teosinte segments spanning the \emph{tb1} locus into an isogenic maize background and investigating their effects on phenotype.
 However, they found evidence for an allelic series contributing to all \emph{tb1} associated phenotypes but tillering .
@@
 Under increased FR light conditions plants experience both morphological and physiological changes including reduction in branching, increased height, and increased ethylene and auxin production \cite{Ballare 1999, Smith and Whitelam 1997}.
-In maize, increase in shading has been shown to be correlated with changes in \emph{tb1} expression and subsequent reduction in tillering, suggesting that \emph{tb1} is involved in the shade avoidance response in maize \cite{Doebley et al 1997, Lukens and Doebley 1999}.
+In maize, increase in shading has been shown to be correlated with changes in \emph{tb1} expression and subsequent reduction in tillering, suggesting that \emph{tb1} is involved in the shade avoidance response \cite{Doebley et al 1997, Lukens and Doebley 1999}.
 If plants in our phenotyping experiments were grown too close together, such that they shaded each other, it could have reduced \emph{tb1} expression and tillering, which would have confounded our results.
@@
 shade would have to so effect expression that it didn't matter what tb1 genotype a plant had -- so G x E effects.
-is it known whether there are such G x E effects?}
+is it known whether there are such G x E effects?} \mbh{Tom and I saw reduced tillering in the TILs under similar greenhouse conditions and wouldn't reduced \emph{tb1} expression result in increased tillering?
+I'd lean toward removing this paragraph}
 
@@
 Analysis of linkage using SNPs from across chromosome 1 does not suggest that the Hopscotch allele is present in these populations due to recent introgression; however, it seems unlikely that the it would have drifted to high frequency in multiple populations, and there may be another explanation for the high frequency we observe in some of our populations.
-The Hopscotch insertion does not appear to have any affect on tillering phenotypes in teosinte previously associated with \emph{tb1} and it is possible that variation at other loci may play a larger role in branching architecture in teosinte than in domesticated maize.
+The Hopscotch insertion does not appear to have any affect on the tillering phenotypes measured here in teosinte previously associated with \emph{tb1} and it is possible that variation at other loci may play a larger role in branching architecture in teosinte than in domesticated maize.
 \jri{seems like we should end with at least a hypothesis.
@@
 what else is there?
-perhaps we should think about selection or introgression scenarios that could explain our data.}
+perhaps we should think about selection or introgression scenarios that could explain our data.} \mbh{I'd lean toward saying that 1) hopscotch does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, hopscotch may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the hopscotch.}
 

### 2014-04-06 manuscript.tex
Updates to tb1 manuscript
Figures, tables, revamped discussion, info. on other loci involved in
tillering

--- 2014-01-16 manuscript.tex
@@
 \newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
+\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
+\newcommand{\lev}[1]{\textcolor{green}{\emph{\scriptsize #1}} }
 
@@
 \author[1]{Laura Vann}
-\author[1]{Thomas Kono}
-\author[1]{Tanja Pyh\"aj\"arvi}
-\author[1]{Matthew Hufford}
-\author[1,2]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
+\author[1,2]{Thomas Kono}
+\author[1,3]{Tanja Pyh\"aj\"arvi}
+\author[1,4]{Matthew B. Hufford}
+\author[1,5]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
 \affil[1]{Department of Plant Sciences, University of California Davis}
-\affil[2]{Center for Population Biology and Genome Center, University of California Davis}
+\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
+\affil[3]{Department of Biology, University of Oulu}
+\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
+\affil[5]{Center for Population Biology and Genome Center, University of California Davis}
 
@@
 \maketitle
-
-\jri{Please add addresses for all authors.  All should have plant sciences, but Kono, Tanja, and Huff need their new addresses too. }
 
@@
 The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
-Further studies have shown that the insertion of a transposable element (Hopscotch) upstream of tb1 enhances its expression, causing the reduction in branching observed in domesticated maize.
+Further studies have shown that the insertion of a transposable element (Hopscotch) upstream of \emph{tb1} enhances its expression, causing the reduction in branching observed in domesticated maize.
 Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the Hopscotch element, led us to investigate the role of \emph{tb1} in teosinte.
-Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{parviglumis} and \emph{mexicana}.
-Analysis of linkage disequilibrium between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
-Finally, two greenhouse experiments fail to find an important role for \emph{tb1} in controlling tillering in natural populations of \emph{parvilgumis}.
-\jri{Needs a concluding sentence.
-"Thus, we think..." or something.}
+Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{parviglumis} and subspecies \emph{mexicana}.
+Analysis of linkage disequilibrium between the Hopscotch element and variation (SNPs) in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
+Finally, two greenhouse experiments fail to find an important role for \emph{tb1} in controlling tillering in natural populations of \emph{parviglumis}.
+Our findings suggest that the role of the Hopscotch in tillering in teosinte is not as straightforward as it is in domesticated maize, and other loci known to affect branching architecture in maize may play a more important role in tillering in teosinte.
 
@@
 
-Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \cite{Ross-Ibarra et al 2007}.
-Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \cite{OlsenGross2010}.
-\jri{I believe you need to remove spaces from the bibtex citation names.
-See how this OlsenGross2010 works but the others don't?
-should be able to work with whitespace but i can't figure out why not.}  Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{Kovach and McCouch 2008}.
+Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
+Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
+Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{pKovachMcCouch 2008}.
 But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
-
-% SFS figure
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=150mm]{SFS.png}
-    \caption{This is an example figure.} 
-\label{sfs_non_syn}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-\jri{This is how you cite the figure in the text.
-"I want to show you all Figure \ref{sfs_non_syn}." Latex takes care of the numbering.
-Please add all the figures to the text, and add them in in the middle like this close to where they are referenced.
-Note that it puts the figure where it thinks best, which is not necessarily near this text.}
 
@@
 \emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 10,000 B.P.
-in southwest Mexico \cite{Piperno et al 2009, Matsuoka et al 2002}.
-Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \cite{Hufford et al 2012a, Doebley 2004, Hufford et al 2012b}.
+in southwest Mexico \citep{Piperno2009, Matsuoka2002}.
+Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
 Additionally, large naturally occurring populations of both \emph{Zea mays} ssp.
 \emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp.
-\emph{mexicana} (highland teosinte) can be found throughout Mexico \cite{Wilkes 1977, Hufford et al 2013}, and genetic diversity and effective population size (Ne) of these taxa is estimated to be high \cite{Ross-Ibarra et al 2009} relative to other plants \cite{Nybom 2004}.
+\emph{mexicana} (highland teosinte) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity and effective population size (Ne) of these taxa is estimated to be high \citep{Ross-Ibarra2009} relative to other plants \citep{Nybom2004}.
 
@@
 One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches.
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in branching differences between maize and teosinte \cite{Doebley et al 1990, Doebley and Stec 1991, Lukens and Doebley 1999}.
-Further studies have shown that the insertion of a 4.9 kb retrotransposon (Hopscotch) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer et al 2011}.
-The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in teosinte \cite{Studer et al 2011}.
-Dating of this element has suggested that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \cite{Studer et al 2011}.
-Furthermore, \cite{Studer and Doebley 2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
+The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in branching differences between maize and teosinte \citep{Doebley1990, DoebleyStec1991, LukensDoebley1999}.
+Further studies have shown that the insertion of a 4.9 kb retrotransposon (Hopscotch) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer2011}.
+The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in teosinte \citep{Studer2011}.
+Dating of this element has suggested that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
+Furthermore, \citep{StuderDoebley2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
+Tillering, or more specifically lack of tillers in teosinte, may provide an ecological advantage to plants.
+In species such as teosinte, where plants grow densely packed together, if plants invest more resources into growing taller and fewer resources into tillering, plants may be able to out-compete neighbors for limiting resources such as light.
+In natural populations of teosinte variation at the \emph(tb1) locus may then play a role in the ecology of teosinte.
 In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{mexicana}, \emph{parviglumis}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
@@
 
-We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 Zea mays ssp.
-mexicana (hereafter mexicana) populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (Table S1, Table S2).
-\jri{please add tables and reference same as figs; see example in Sofiane's "deleterious 282" repo on lab github page} DNA was extracted from leaf tissue using a modified CTAB approach \cite{Doyle and Doyle 1990, Maloof et al 1984}.
-We designed primers using PRIMER3 \cite{Rozen and Skaletsky 2000} implemented in Geneious \cite{Kearse et al 2012} to amplify the entire Hopscotch element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the Hopscotch insertion.
+We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{Zea mays} ssp.
+\emph{mexicana} (hereafter \emph{mexicana}) populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (\ref{TableS1Mays}, \ref{TableS2}).
+DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
+We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \cite{Kearse2012} to amplify the entire Hopscotch element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the Hopscotch insertion.
 Two PCRs were performed for each individual, one with primers flanking the Hopscotch (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
 Primer sequences are HopF, 5'-TCGTTGATGCTTTGATGGATGG-3';
-Hop R, 5'-AACAGTATGATTTCATGGGACCG-3'; and HopIntR, 5'-CCTCCACCTCTCATGAGATCC-3' (Figure S1).
-\jri{supplemental material should be added to this file too.} Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (Figure S2).
+Hop R, 5'-AACAGTATGATTTCATGGGACCG-3'; and HopIntR, 5'-CCTCCACCTCTCATGAGATCC-3' (\ref{FigS1Locus}).
+Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (\ref{FigS2Gel}).
 When only one PCR resolved well, we scored one allele for the individual, which explains the odd number of alleles included in our analyses.
@@
 We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of start of the Hopscotch and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
+For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of the start of the Hopscotch and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
 The 5' UTR (containing sequenced region 1) has been shown to have elevated diversity in ssp.
-\emph{parviglumis} with respect to maize, while the the area upstream from the Hopscotch (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \cite{Wang et al 1999, Clark et al 2006}.
-We designed the following primers using PRIMER3 \cite{Rozen and Skaletsky 2000}: for the 5' UTR, 5' GGATAATGTGCACCAGGTGT 3' and 5' GCGTGCTAGAGACACYTGTTGCT 3'; for the 50 kb upstream region, 5' TGTCCTCGCCGCAACTC 3' and 5' TGTACGCCCGCCCCTCATCA 3' (Figure S1).
+\emph{parviglumis} with respect to maize, while the the area upstream from the Hopscotch (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \citep{Wang1999, Clark2006}.
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, 5' GGATAATGTGCACCAGGTGT 3' and 5' GCGTGCTAGAGACACYTGTTGCT 3'; for the 50 kb upstream region, 5' TGTCCTCGCCGCAACTC 3' and 5' TGTACGCCCGCCCCTCATCA 3' (\ref{FigS1Locus}).
 We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94°C for 3 min, 30 cycles of 92°C for 40 s, annealing for 1 min, 72°C for 40 s, and a final 10 min extension at 72°C.
@@
 
-We cloned cleaned fragments into a TOPO-TA vector (Invitrogen, Carlsbad) using OneShot TOP10 chemically competent E.
-coli cells, with an extended ligation time of 30 min for a complex target fragment.
+We cloned cleaned fragments into a TOPO-TA vector (Invitrogen, Carlsbad) using OneShot TOP10 chemically competent \emph{E.
+coli} cells, with an extended ligation time of 30 min for a complex target fragment.
 We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96°C for 5 min; then 35 cycles at 96°C for 30 s, 53°C for 30 s, 72°C for two min; and a final extension at 72°C for 4 min.
@@
 Amplification products with successful incorporation of our insert were cleaned using Exonuclease 1 and Antarctic Phosphatase following the procedures detailed above, and sequenced with vector primers M13 Forward and M13 Reverse using Sanger sequencing at the College of Agriculture and Environmental Sciences (CAES) sequencing center at UC Davis.
-We aligned and trimmed primer sequences from resulting sequences using the software Geneious \cite{Kearse et al 2012}.
+We aligned and trimmed primer sequences from resulting sequences using the software Geneious \citep{Kearse2012}.
 Following alignment, we verified singleton SNPs by sequencing an additional one to four colonies from each clone.
@@
 
-We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE), and to calculate differentiation between populations (FST) and subspecies (FCT) we used HierFstat \cite{Goudet 2005}.
+We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE), and to calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
 These analyses only included populations in which 8 or more individuals were sampled.
-To test the hypothesis that the Hopscotch insertion may be adaptive under certain environmental conditions, we looked for significant associations between the Hopscotch frequency and environmental variables using BayEnv \cite{Coop et al 2010}.
-BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\cite{Coop et al 2010}.
-We used genotyping and covariance data from Pyhäjärvi et al., (2013) \cite{Pyhajarvi et al 2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (Table S3).
-Environmental data were obtained from www.worldclim.org, the Harmonized World Soil Database and www.harvestchoice.org, and summarized by principle component analysis Pyhäjärvi et al.
-(2013) \cite{Pyhajarvi et al 2013}.
+To test the hypothesis that the Hopscotch insertion may be adaptive under certain environmental conditions, we looked for significant associations between the Hopscotch frequency and environmental variables using BayEnv \citep{Coop2010}.
+BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\citep{Coop2010}.
+We used genotyping and covariance data from Pyh\"aj\"arvi et al., (2013) \citep{Pyh\"aj\"arvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
+Environmental data were obtained from www.worldclim.org, the Harmonized World Soil Database and www.harvestchoice.org, and summarized by principle component analysis \citet{Pyh\"aj\"arvi2013} \citep{Pyh\"aj\"arvi2013}.
 
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \cite{Thornton 2003} to calculate pairwise FST between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator$\theta_{W}$; pairwise nucleotide diversity $\theta_{\pi}$; and Tajima's D).
+For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_{\subscript{W}}$; pairwise nucleotide diversity $\theta_{\subscript{\pi}}$; and Tajima's D).
 To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
-For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the Hopscotch insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \cite{Chia et al 2012}).
+For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the Hopscotch insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citep{Chia2012}).
 
-In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \cite{Bradbury et al 2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A, Ejutla B, San Lorenzo, and La Mesa populations \cite{Pyhajarvi et al 2013}.
+In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A, Ejutla B, San Lorenzo, and La Mesa populations \citep{Pyhajarvi2013}.
 We chose these populations because we had both genotyping data for the Hopscotch as well as chromosome-wide SNP data for chromosome 1.
 For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for San Lorenzo, Ejutla B, Ejutla A, and La Mesa, respectively.
-We then used Tassel \cite{Bradbury et al 2007} to calculate linkage disequilibrium ($r^{2}$) across chromosome 1 for each population.
+We then used Tassel \citep{Bradbury2007} to calculate linkage disequilibrium ($r^{2}$) across chromosome 1 for each population.
 
-We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \cite{Falush et al 2003} and the same phased 55K SNP data from \cite{Pyhajarvi et al 2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \cite{Cook et al 2012}.
-SNPs were anchored in a modified version of the IBM genetic map ( \cite{Gerke et al 2013}, \url{http://arxiv.org/abs/1307.7313}).
-We created haplotype blocks using a custom Perl script that grouped all SNPs within a 5 kb window into a haplotype.
+We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and the same phased 55K SNP data from \citep{Pyh\"aj\"arvi2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
+SNPs were anchored in a modified version of the IBM genetic map (\citep{Gerke2013}, \url{http://arxiv.org/abs/1307.7313}).
+We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 To investigate the phenotypic effects of the Hopscotch insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \cite{Hufford 2010} where the Hopscotch is segregating at highest frequency (0.4375) in our initial genotyping sample set.
+We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.4375) in our initial genotyping sample set.
 In order to maximize the likelihood of finding the Hopscotch in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
@@
 
-Starting on day 15, we measured tillering index, the ratio of the sum of tiller lengths to the height of the plant \cite{Briggs et al 2007}.
-Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \cite{Briggs et al 2007}.
+Starting on day 15, we measured tillering index, the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
+Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Briggs2007}.
 Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
@@
 
-Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \cite{Fox and Weisberg 2011} to improve the normality and homogeneity of variance among genotype classes.
+Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
@@
 
-We genotyped 247 accessions of maize landraces, and found the Hopscotch element fixed in all but 8 of these accessions (inbred lines X,Y, Z, and landraces in Chihuahua, Tapalpa, Tenancingo, and Altotonga) (Table S1, Table S2).
+We genotyped 247 accessions of maize landraces, and found the Hopscotch element fixed in all but 8 of these accessions (\ref{TableS1Mays}, \ref{TableS2}).
 For all analyses we only included individuals that clearly resolved for both PCRs, resulting in 837 individuals total.
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoacán in central-western Mexico in both subspecies (Figure 1).
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (\ref{Fig1Map}).
 We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
-Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p=0.05).
+Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p<0.05).
 
-Using our Hopscotch genotyping data, we calculated differentiation between populations (FST) and subspecies (FCT) for populations in which we sampled 8 or more alleles.
-FCT is 0 within our dataset, and we found similar levels of FST between populations within each subspecies (0.22) and between all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi et al 2013} (Table 1).
-Although we found large variation in Hopscotch allele frequency among our populations, our BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors < 1; Table S3).
+%Figure 1 Map 
+%--------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{Fig1Map.pdf}
+    \caption{Map showing the frequency of the Hopscotch allele in populations of \emph{Zea mays} ssp.
+\emph{parviglumis} where we sampled more than 6 individuals.
+Size of circles reflects number of alleles sampled.
+}
+\label{Fig1Map}
+  \end{center}
+\end{figure*}
+%--------------------------------------------
+
+Using our Hopscotch genotyping data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+F\textsubscript{CT} is 0 within our dataset, and we found similar levels of F\textsubscript{ST} between populations within each subspecies (0.22) and between all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi2013} (\ref{Table1Fst}).
+% Table 1
+\begin{table}[htbp]
+  \centering
+  \caption{Pairwise F\textsubscript{CT} values from sequence data}
+    \begin{tabular}{rccc}
+    \toprule
+    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Sequenced Region 2} & \textbf{Sequenced Region 1} & \textbf{Hopscotch} \\
+    \midrule
+    EjuA \& EJuB & 0     & 0     & 0 \\
+    EjuA \& MSA & 0.328 & 0.326 & 0.186 \\
+    EjuA \& SLO & 0.258 & 0.416 & 0.28 \\
+    EjuB \& MSA & 0.365 & 0.397 & 0.188 \\
+    EjuB \& SLO & 0.29  & 0.512 & 0.28 \\
+    MSA \& SLO & 0     & 0.007 & 0.016 \\
+    \bottomrule
+    \end{tabular}
+  \label{Table1Fst}
+\end{table}
+Although we found large variation in Hopscotch allele frequency among our populations, our BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors $<$ 1; \ref{TableS3BayEnv}).
 
@@
 
-To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the tb1 region, we sequenced two small (<1kb) regions upstream of the tb1 ORF in four populations.
+To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the tb1 ORF in four populations.
 After alignment and singleton checking we recovered 40 and 48 segregating sites for the 50kb upstream region and the 5' UTR region, respectively.
-For region 1, Ejutla A, Ejutla B, and La Mesa have comparable values for haplotype diversity,$\theta_{W}$, $\theta_{\pi}$, and Tajima's D, while San Lorenzo has lower values for haplotype diversity,$\theta_{W}$, $\theta_{\pi}$, and a more positive value for Tajima's D (Table 2).
-For region 2, haplotype diversity,$\theta_{W}$, and $\theta_{\pi}$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table 2).
-Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this latter population (Table 2).
-Pairwise values of FST within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (Table 1).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILS; data from Maize HapMapV2, \cite{Chia et al 2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (Figure S3).
+For region 1, Ejutla A has the highest values of haplotype diversity, $\theta_{\textsubscript{W}}$, and $\theta_{\textsubscript{\pi}}$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
+Additionally, Tajima's D is similarly negative in the two Ejutla populations and La Mesa, but is more positive in San Lorenzo (\ref{Table2Diversity}).
+% Table 2 Diversity Statistics
+\begin{table}[htbp]
+  \centering
+  \caption{Add caption}
+    \begin{tabular}{rccccc}
+    \toprule
+    Population & Number of Haplotypes & Haplotype Diversity & $\theta_{\textsubscript{W}}$    & $\theta_{\textsubscript{\pi}}$    & Tajima?s D \\
+    \midrule
+    \multicolumn{6}{c}{\textit{Sequenced region 2 (50kb upstream)}} \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.89394 & 0.01548 & 0.01763 & 0.6231 \\
+    \multicolumn{1}{c}{EJUB} & 8     & 0.89394 & 0.01493 & 0.01591 & 0.29504 \\
+    \multicolumn{1}{c}{MSA} & 3     & 0.68182 & 0.01111 & 0.01055 & -0.22212 \\
+    \multicolumn{1}{c}{SLO} & 4     & 0.74242 & 0.01167 & 0.01413 & 0.93185 \\
+    \multicolumn{6}{c}{\textit{Sequenced region 1(5? UTR)}} \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.85897 & 0.00874 & 0.00527 & -1.64955 \\
+    \multicolumn{1}{c}{EJUB} & 5     & 0.70909 & 0.00663 & 0.00378 & -1.83123 \\
+    \multicolumn{1}{c}{MSA} & 6     & 0.68182 & 0.00646 & 0.00373 & -1.75506 \\
+    \multicolumn{1}{c}{SLO} & 3     & 0.31818 & 0.00176 & 0.00137 & -0.72873 \\
+    \bottomrule
+    \end{tabular}
+  \label{Table2Diversity}
+\end{table}
+
+For region 2, haplotype diversity, $\theta_{\textsubscript{W}}$, and $\theta_{\textsubscript{\pi}}$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (\ref{Table2Diversity}).
+Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this latter population (\ref{Table2Diversity}).
+Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (\ref{Table1Fst}).
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILS; data from Maize HapMapV2, \citep{Chia2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (\ref{FigureS3NJ}).
 Furthermore, individuals within our sample that have the Hopscotch insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the Hopscotch insertion.
@@
 If the Hopscotch is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the Hopscotch and SNPs surrounding the tb1 region in our resequenced populations (Figure 2), and $r^{2}$ in the \emph{tb1} region (positions 264,596,664-265,891,456; AGPv2) does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the Hopscotch genotype (average $r^{2}=0.082$).
-In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (Table 3).
+We find no evidence of elevated linkage disequilibrium between the Hopscotch and SNPs surrounding the \emph{tb1} region in our resequenced populations (\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region (positions 264,596,664-265,891,456; AGPv2) does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the Hopscotch genotype (average $r^{2}=0.082$).
+In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
+
+% Table 3 $r^{2}$ values
+\begin{table}[htbp]
+  \centering
+  \caption{$r^{2}$ values between SNPs in the \emph{tb1} region (positions 264,596,664-265,891,456) and the rest of chromosome 1, within the 5?
+UTR (Sequenced region 1), and within the 66,169 bp upstream region (Sequenced region 2).}
+    \begin{tabular}{cccrc}
+    \toprule
+    \multirow{2}[2]{*}{Population} & \multirow{2}[2]{*}{Chromosome 1} & \multirow{2}[2]{*}{tb1 region} & \multicolumn{1}{c}{\multirow{2}[2]{*}{Sequenced region 1}} & \multirow{2}[2]{*}{Sequenced region 1 } \\
+    \midrule
+          &       &       & \multicolumn{1}{c}{} &  \\
+    Ejutla A & 0.095426101 & 0.050304 & 0.747295 & 0.214933 \\
+    Ejutla B & 0.068681837 & 0.051295 & 0.660354 & 0.186395 \\
+    La Mesa & 0.069500533 & 0.053306 & 0.914286 & 0.766234 \\
+    San Lorenzo & 0.100536784 & 0.067251 & 0.912281 & 0.636364 \\
+    \bottomrule
+    \end{tabular}
+  \label{Table3R2}
+\end{table}
+ 
+
+% Figure 2. LD plots
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{Fig2LDPlots.pdf}
+    \caption{Linkage equilibrium plots for SNPs within bp 261,000,000-268,000,000 on chromosome 1.
+The yellow rectangle indicates the location of the Hopscotch insertion and the green represents the \emph{tb1} ORF.
+A) Ejutla A; B) Ejutla B; C) La Mesa; D).
+San Lorenzo}
+\label{Fig2LD}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
 
 The lack of clustering of Hopscotch genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the Hopscotch insertion in these populations of \emph{parviglumis} is the result of recent introgression.
-However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \cite{Cook et al 2012, Phyhajarvi et al 2013}.
+However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Phyh\"aj\"arvi2013}.
 The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations.
 If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is lower (average 0.017) than the chromosome wide average (0.20).
-(Figure 3).
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is lower (average 0.017) than the chromosome wide average (0.20; \ref{Fig3Structure}).
+
+% Figure 3 STRUCTURE Plot 4 Pops
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{Fig3Structure.pdf}
+    \caption{STRUCTURE assignment to maize across a section of chromosome 1.
+The dotted lines mark the beginning of the sequenced region 50kb upstream (Sequenced region 2) and the end of the \emph{tb1} ORF.
+}
+\label{Fig3Structure}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
 
@@
 We found the Hopscotch segregating at a frequency of 0.65; with a 0.4 frequency of homozygotes for the Hopscotch, a 0.5 frequency of heterozygotes, and a 0.1 frequency of homozygotes for the teosinte allele (no Hopscotch), showing no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and Hopscotch genotype we find a weak positive correlation between presence of the Hopscotch and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (Figure 4).
+After performing a repeated measures ANOVA between our transformed tillering index data and Hopscotch genotype we find a weak positive correlation between presence of the Hopscotch and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
 Additionally we find no significant correlation between tiller number and Hopscotch genotype, or culm diameter and Hopscotch genotype in phenotyping 1.
+
+% Figure 4 Phenotyping BoxPlots
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{Fig4Boxplots.pdf}
+    \caption{Box-plot showing tillering index in our greenhouse population from day 20-60.
+White indicates individuals homozygous for the Hopscotch, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No Hopscotch) allele.
+Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.}
+\label{Fig4Boxplots}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
 
@@
 We find the Hopscotch allele segregating at a frequency of 0.69, with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
-We find similar patterns as in phenotyping 1, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p=0.0611), with no significant correlation on any day.
+We find similar patterns as in phenotyping 1, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p<0.0611), with no significant correlation on any day.
 Similarly, relationships between Hopscotch genotype and tiller number, and hopscotch genotype and culm diameter are not significant.
@@
 
-Adaptation occurs either due to selection on standing variation or on de novo mutations.
-Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection on lactase tolerance in humans \cite{Plantinga et al 2012, Tishkoff et al 2007}; variation at the \jri{italics?} Eda locus in three-spined stickleback \cite{Kitano et al 2008, Colosimo et al 2005}; and pupal diapause in the Apple Maggot fly \cite{Feder et al 2003}.
+Adaptation occurs either due to selection on standing variation or on \emph{de novo} mutations.
+Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}; variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}; and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
 Although the role of standing variation with respect to adaptation has been described in many systems, its importance to domestication is not as well studied.
 
-In maize, alleles at important domestication loci (RAMOSA1, \cite{Sigmon and Vollbrecht 2010}; barren stalk1, \cite{Gallavotti et al 2004}; and \jri{do gene names need to be italicized?} grassy tillers1, \cite{Whipple et al 2011} have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
-The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \cite{Clark et al 2006, Studer et al 2011}.
-Studer et al.
-(2011) genotyped 90 accessions of teosinte (inbred and non-inbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \cite{Studer et al 2011}.
+In maize, alleles at important domestication loci (\emph{RAMOSA1}, \citep{SigmonVollbrecht2010}; \emph{barren stalk1}, \citep{Gallavotti2004}; and \emph{grassy tillers1}, \citep{Whipple2011}) have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
+The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \citep{Clark2006, Studer2011}.
+\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \citep{Studer2011}.
 
-Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it is still segregating in populations of teosinte.
+Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
 However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the Hopscotch in teosinte.
-While our findings are consistent with a previous study by Studer et al.
-(2011) in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \cite{Studer et al 2011}.
-Crop to wild introgression has previously been observed at domestication loci (\cite{Ellstrand et al 1999, Zhang et al 2009, Thurber et al 2010, Baack et al 2008, Hubner et al 2012, Wilkes et al 1977, Van Heerwaarden et al 2011, Barrett 1983}), but our results are more consistent with Hufford et al.
-(2013) who found resistance to introgression from maize into teosinte \cite{Hufford et al 2013}.
-Furthermore, \jri{all citations, even citations like these need to reference the bibtex file.
-check out the difference betweetn cite, citet, and citep as commands.
-see \url{http://merkel.zoneo.net/Latex/natbib.php} for details} Hufford et al.
-(2013) showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).
+While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
+Crop to wild introgression has previously been observed at domestication loci (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}), but our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
+Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.
 
@@
 We do not have sympatric maize samples from many of our populations, however, so we cannot rule out the possibility that Hopscotch alleles in these teosinte populations are segregating in sympatric maize.
-On the other hand, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1.
+Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1.
 The similarity in LD in the \emph{tb1} region in populations with and without the Hopscotch provides further support for an explanation other than introgression.
-Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (Figure 3, Table 4).
-\jri{figure references need to be as in the example using the ref command}
+Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{\Table4Q}).
+
+ % Table 4 Assigment to maize and teosinte in tb1 region and chr 1 
+\begin{table}[htbp]
+  \centering
+  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
+    \begin{tabular}{ccccc}
+    \toprule
+          & \multicolumn{2}{c}{\emph{tb1} region} & \multicolumn{2}{c}{Chr 1} \\
+    \midrule
+    Population & Maize assignment & Teosinte assignment & Maize assignment & Teosinte assignment \\
+    Ejutla A & 0.02158681 & 0.9784132 & 0.2026814 & 0.7973186 \\
+    Ejutla B & 0.01888194 & 0.9811181 & 0.1872131 & 0.8127869 \\
+    La Mesa & 0.9881333 & 0.0118675 & 0.8068998 & 0.1931017 \\
+    San Lorenzo & 0.01551389 & 0.9844861 & 0.2048252 & 0.7951748 \\
+    \bottomrule
+    \end{tabular}
+  \label{Table4Q}
+\end{table}
 
@@
 \emph{mexicana} and \emph{Zea mays} ssp.
-\emph{parviglumis} in a number of sympatric populations \cite{Hufford et al 2013, Ellstrand et al 2007, van Heerwaarden et al 2011}.
-\jri{check these refs.
-Hufford 2013 is mexicana, as is i think Ellstrand.
-do we show introgression with parv in van Heerwaarden?.
-do is how it in my 2009 paper?}
+\emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, vanHeerwaarden2011}.
 
-The lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D in the San Lorenzo population are consistent with patterns expected under neutrality.
-\jri{lower diversity is consistent with neutrality?} Though, deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
-\jri{shouldn't Tanja's data be able to argue one way or the other here?
-she looked at eveidence of inbreeding.
-what did she find in san lorenzo?} Similarly, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
+Deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
+Previous studies using both SSRs and genome-wide SNP data have provided evidence for a population bottleneck in the San Lorenzo population that has resulted in lower levels of diversity across the genome \citep{Hufford2010, Pyh\"aj\"arvi2013}.
+The lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D in the San Lorenzo population that we present here are consistent with these previous findings.
+Although, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
+Interestingly, many of our populations with high frequency of the Hopscotch allele in the Jalisco region overlap with the Jalisco cluster in \citet{Fukunaga2005}.
 
-The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \cite{Clark et al 2006, Studer et al 2011}, and Weber et al.
-(2007) have described its effects in partially inbred lines of teosinte \cite{Weber et al 2007}.
+\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the hopscotch allele.
+Why?
+Only ancient introgression in Jalisco and not in the Balsas?
+If hopscotch came from standing variation, is this more consistent with a Jalisco domestication?
+I know, heresy!!
+LV, can you look at the \emph{parviglumis} populations with high hopscotch and see if they are all in the Jalisco cluster of Fukunaga?
+Might be a nice observation we could make here...could even go so far as saying the results from Matsuoka deserve another look.} \lev{Matt, yes they are in the Jalisco cluster, I added in a sentence mentioning it, but am not sure if that's the direction you were wanting me to take with it or not}
+
+The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007}.
 Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.
 However, we found no significant effect of the Hopscotch on tillering index or tiller number in our phenotyping experiments, and the effect of the Hopscotch insertion in teosinte is discordant with that of maize.
-The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize, as well as our own preliminary data collected on a grow-out of teosinte inbred lines.
-\jri{Huff, think we should include the TIL data just to show we can see effects in our greenhouse?}
+The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize.
+\lev{i took out the information about the TILS -- seemed the consensus was more than we wanted to do for this pub.
+If qPCR is working for tb1 it would be something easy to check...} It is certainly possible that even though previous data demonstrate an effect of the Hopscotch on tillering in maize \citep{Studer2011}, that the effect of the Hopscotch in teosinte is more complicated and may be more difficult to observe.
+Moreover, \emph{tb1} is a single gene in a complex pathway that affects branching and tillering traits, and perhaps in combination with alleles at other loci the phenotypic effects of the Hopscotch on tillering may not be consistent.
 
-It is certainly possible that even though data from the TILs suggest an effect of the Hopscotch on tillering in teosinte, that it in fact has no effect.
-\jri{i don't think we want to argue that tb1 has no effect!!
-we have no evidence for this.
-instead, it may be that the effect is hard to see if there are numerous other loci and/or epistasis, etc.} Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \cite{Clark et al 2006}, and perhaps we would have detected an effect of the Hopscotch on these additional phenotypes.
-\jri{why say we might have detected these?
-we didn't look for them, not sure worth mentioning}  A recent study by Studer and Doebley (2012) examined the possibility of an allelic series at the \emph{tb1} locus in teosinte by introgressing 9 separate teosinte segments spanning the \emph{tb1} locus into an isogenic maize background and investigating their effects on phenotype.
-However, they found evidence for an allelic series contributing to all \emph{tb1} associated phenotypes but tillering .
-They did find variation in tillering both between introgressed lines and between isogenic maize lines \jri{don't follow.
-you mean the different alleles didn't have different tillering phenotypes?
-but they did show that all alleles affected tillering, right?}, suggesting that there may be additional loci that contribute to tillering in both maize and teosinte \cite{Studer and Doebley 2012}.
+Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \citep{Clark2006}, and a recent study by \citet{StuderDoebley2012} examined the possibility of an allelic series at the \emph{tb1} locus in teosinte.
+\citet{StuderDoebley2012} introgressed 9 separate teosinte segments (one from \emph{Zea diploperennis}, and four from both \emph{Zea mays} ssp.
+\emph{mexicana} and \emph{Zea mays} ssp.
+{\parviglumis}) spanning the \emph{tb1} locus into an isogenic maize background and investigated their effects on previously associated phenotypes.
+They found that plants with teosinte chromosomal segments had greater tillering than their maize isogenic lines, and that different chromosomal segments of \emph{tb1} confer different amounts of tillering, suggesting that there are multiple genetic factors in this region that affect tillering.
+However, in addition to elucidating variance in tillering among \emph{tb1} teosinte segments, \citet{StuderDoebley2012} found significant variance among W22 control lines, suggesting that there are other genetic factors aside from alleles at the \emph{tb1} locus that affect tillering in maize.
+\citet{DoebleyStec1991, DoebleyStec1993} first attempted to map QTL controlling many of the phenotypic differences between domesticated maize and teosinte, and demonstrated the existence of numerous QTL that contribute to the differences in branching architecture between the two.
+Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012, PAPERS}, and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
+\emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number) \citep{Li2012}.
 It seems plausible that variation in some of these other loci could have affected tillering in our greenhouse population, and contributed to the lack of correlation we see between Hopscotch genotype and tillering.
-\jri{are there other studies on other loci?
-what about Briggs?
-original Doebley QTL papers?
-one of the weber papers maybe?
-seems like there should be explicit cites to other loci we know are involved.
-also we could cite \% variance explained by tb1 of tillering in the maize x teo crosses, arguing that there must be other loci.}
-
-Another explanation for the lack of correlation between Hopscotch genotype and tillering could be that growing conditions in our greenhouse affected growth habit of our plants.
-The \emph{tb1} gene is part of the shade avoidance pathway in maize, and this involvement may have presented problems for our phenotyping design \citep{Kebrom and Brutnell 2007}.
-Broadly, shade avoidance syndrome (SAS), refers to the morphological and physiological response of plants experiencing an increase in far-red (FR) light, which occurs during shading \cite{Kebrom and Brutnell 2007}.
-Under increased FR light conditions plants experience both morphological and physiological changes including reduction in branching, increased height, and increased ethylene and auxin production \cite{Ballare 1999, Smith and Whitelam 1997}.
-In maize, increase in shading has been shown to be correlated with changes in \emph{tb1} expression and subsequent reduction in tillering, suggesting that \emph{tb1} is involved in the shade avoidance response in maize \cite{Doebley et al 1997, Lukens and Doebley 1999}.
-If plants in our phenotyping experiments were grown too close together, such that they shaded each other, it could have reduced \emph{tb1} expression and tillering, which would have confounded our results.
-\jri{maybe, but we saw variation in tillering, especially in growout number 1.
-shade would have to so effect expression that it didn't matter what tb1 genotype a plant had -- so G x E effects.
-is it known whether there are such G x E effects?}
 
@@
 Analysis of linkage using SNPs from across chromosome 1 does not suggest that the Hopscotch allele is present in these populations due to recent introgression; however, it seems unlikely that the it would have drifted to high frequency in multiple populations, and there may be another explanation for the high frequency we observe in some of our populations.
-The Hopscotch insertion does not appear to have any affect on tillering phenotypes in teosinte previously associated with \emph{tb1} and it is possible that variation at other loci may play a larger role in branching architecture in teosinte than in domesticated maize.
+The Hopscotch does not appear to uniformly reduce tillering in teosinte as it does in maize, and the other loci involved in branching architecture of maize may play key roles in the regulation of tillering in teosinte.
+Furthermore, the high frequency with which we find the Hopscotch in a number of populations may suggest that the Hopscotch may play an additional ecological role in teosinte.
+In the future, additional experiments will be needed to examine expression levels of \emph{tb1} and additional loci involved in branching architecture (e.g.
+\emph{gt1}, \emph{tru1}, and \emph{ter1}) in conjuction with a more exhaustive phenotyping and genotyping assay.
+
 \jri{seems like we should end with at least a hypothesis.
@@
 
+\mbh{I'd lean toward saying that 1) hopscotch does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, hopscotch may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the hopscotch.}
+
+\lev{I went with Matt's suggestions.
+I am happy to include such scenarios, but everything I have thought of seems rather unlikely; i.e.
+only introgression or selection in Jalisco.
+Jeff I can mention that maybe parv is more of a weed there than other locations, but i thought that maize pollen spread rather substantial distances.
+If so wouldn't a number of our parv and mex sites would be within pollen dispersal distance?}
+
 \clearpage
@@
 
+%-------------------------------------------------
+Supplementary Materials
+%-------------------------------------------------
+\clearpage
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\renewcommand{\figurename}{Sup.\Fig.}
+\renewcommand{\tablename}{Sup.\Table}
+
+% Figure S1. Cartoon representation of tb1 locus
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{FigS1LocusCartoon.pdf}
+    \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red).
+Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced regions 1; within the 5?
+UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.}
+\label{FigS1Locus}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
+
+% Figure S2. Gel image
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{FigS2Gel.png}
+    \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+\label{FigS2Gel}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
+
+% Figure S3. NJ Trees
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
+    \caption{Neighbor-Joining tree of the sequenced region in the 5?
+UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}.
+Individuals with genotype data are colored: Homozygous for the teosinte (no Hopscotch) allele (red), homozygous for the maize (Hopscotch) allele (blue), heterozygotes (purple).
+TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the Hopscotch insertion.
+Black indicates individuals not genotyped for the Hopscotch insertion.}
+\label{FigS3NJ}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
+
+% Table S1. Parv and mex samples
+\begin{table}[htbp]
+  \centering
+  \caption{Accessions of \emph{Zea mays} ssp.
+\emph{mexicana} (RIMME) and \emph{Zea mays} ssp.
+\emph{parviglumis} (RIMPA) sampled.
+RIHY is a \emph{Z.
+mays} ssp.
+\emph{parviglumis} and \emph{Zea mays} ssp.
+\emph{mays} hybrid.
+}
+    \begin{tabular}{rrrrrr}
+    \toprule
+    Accession & USDA Accession ID & Locality & Number alleles sampled & Hopscotch Frequency & No Hopscotch Frequency \\
+    \midrule
+    RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
+    RIMME0006 & 566673 & Durango, Mexico & 2     & 0     & 1 \\
+    RIMME0007 & 566680 & Guanajuato, Mexico & 2     & 0     & 1 \\
+    RIMME0008 & 566681 & Michoacan, Mexico & 2     & 0     & 1 \\
+    RIMME0009 & 566682 & Distrito Federal, Mexico & 2     & 0     & 1 \\
+    RIMME0011 & 566685 & Mexico, Mexico & 2     & 0     & 1 \\
+    RIMME0014 & 714151 & Breeders line; Puga: 11066 & 6     & 0     & 1 \\
+    RIMME0017 & 699874 & Ayotlan, Mexico & 8     & 0     & 1 \\
+    RIMME0021 & N/A   & El Porvenir, Mexico & 69    & 0.173913043 & 0.826086957 \\
+    RIMME0026 & N/A   & Opopeo, Mexico & 42    & 0.071428571 & 0.928571429 \\
+    RIMME0028 & N/A   & Puruandiro, Mexico & 28    & 0.035714286 & 0.964285714 \\
+    RIMME0029 & N/A   & Ixtlan, Mexico & 35    & 0     & 1 \\
+    RIMME0030 & N/A   & San Pedro, Mexico & 27    & 0     & 1 \\
+    RIMME0031 & N/A   & Tenango del Aire, Mexico & 25    & 0.08  & 0.92 \\
+    RIMME0032 & N/A   & Nabogame, Mexico & 24    & 0     & 1 \\
+    RIMME0033 & N/A   & Puerta Encantada, Mexico & 25    & 0     & 1 \\
+    RIMME0034 & N/A   & Santa Clara, Mexico & 23    & 0     & 1 \\
+    RIMME0035 & N/A   & Xochimilco, Mexico & 25    & 0     & 1 \\
+    RIMPA0001 & 87168 & El Salado, Mexico & 4     & 0     & 1 \\
+    RIMPA0003 & 87171 & Mazatlan, Mexico & 8     & 0.125 & 0.875 \\
+    RIMPA0017 & 87200 & N/A   & 4     & 0     & 1 \\
+    RIMPA0019 & 87213 & El Salado, Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0029 & 87244 & N/A   & 2     & 0.5   & 0.5 \\
+    RIMPA0031 & 87249 & N/A   & 2     & 0.5   & 0.5 \\
+    RIMPA0035 & 87288 & Jalisco, Mexico & 4     & 0     & 1 \\
+    RIMPA0040 & 288185 & Mexico, Mexico & 4     & 0     & 1 \\
+    RIMPA0042 & 288187 & Guerrero, Mexico & 4     & 0.25  & 0.75 \\
+    RIMPA0043 & 288188 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0045 & 288193 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0055 & 714152 & Breeders line & 2     & 0     & 1 \\
+    RIMPA0056 & 714153 & Breeders line & 2     & 0.5   & 0.5 \\
+    RIMPA0057 & 714154 & Breeders line & 2     & 0.5   & 0.5 \\
+    RIMPA0058 & N/A   & N/A   & 4     & 0.5   & 0.5 \\
+    RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
+    RIMPA0060 & 714157 & Breeders line: CIMMYT 11355 & 2     & 0     & 1 \\
+    RIMPA0061 & 714158 & Breeders line: USDA PI566686 & 4     & 0.5   & 0.5 \\
+    RIMPA0062 & 714159 & Breeders line & 4     & 0.5   & 0.5 \\
+    RIMPA0063 & 714160 & Breeders line & 4     & 0     & 1 \\
+    RIMPA0064 & 714161 & Breeders line & 3     & 0     & 1 \\
+    RIMPA0065 & 714162 & Breeders line & 4     & 0.25  & 0.75 \\
+    RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
+    RIMPA0069 & 699862 & Ixtlan, Mexico & 14    & 0.142857143 & 0.857142857 \\
+    RIMPA0070 & 699863 & Benito Jaurez, Mexico & 16    & 0     & 1 \\
+    RIMPA0071 & 699864 & Tuzantla, Mexico & 28    & 0     & 1 \\
+    RIMPA0072 & 699865 & Tiquicheo, Mexico & 16    & 0     & 1 \\
+    RIMPA0073 & 699866 & Tiquicheo, Mexico & 16    & 0.125 & 0.875 \\
+    RIMPA0074 & 699867 & Huetamo, Mexico & 12    & 0     & 1 \\
+    RIMPA0075 & 699868 & Huetamo, Mexico & 2     & 0     & 1 \\
+    RIMPA0076 & 699869 & Huetamo, Mexico & 4     & 0     & 1 \\
+    RIMPA0077 & 699870 & Caracuaro, Mexico & 2     & 0     & 1 \\
+    RIMPA0078 & 699871 & Caracuaro, Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0079 & 699872 & Villa Madero, Mexico & 14    & 0     & 1 \\
+    RIMPA0080 & 699873 & Guachinango, Mexico & 12    & 0     & 1 \\
+    RIMPA0081 & 699875 & Ameca, Mexico & 16    & 0     & 1 \\
+    RIMPA0083 & 699877 & Tepoztlan, Mexico & 14    & 0     & 1 \\
+    RIMPA0084 & 699878 & Tepoztlan, Mexico & 16    & 0     & 1 \\
+    RIMPA0085 & 699879 & Miahuatlan, Mexico & 16    & 0     & 1 \\
+    RIMPA0086 & 699880 & Miahuatlan, Mexico & 16    & 0.0625 & 0.9375 \\
+    RIMPA0087 & 699881 & Tecoanapa, Mexico & 24    & 0     & 1 \\
+    RIMPA0089 & 699883 & Guerrero, Mexico & 12    & 0     & 1 \\
+    RIMPA0090 & 699884 & Guerrero, Mexico & 10    & 0     & 1 \\
+    RIMPA0091 & 699885 & Guerrero, Mexico & 16    & 0     & 1 \\
+    RIMPA0092 & 699886 & Guerrero, Mexico & 10    & 0     & 1 \\
+    RIMPA0093 & 699887 & Guerrero, Mexico & 26    & 0.076923077 & 0.923076923 \\
+    RIMPA0094 & 699888 & Guerrero, Mexico & 2     & 0     & 1 \\
+    RIMPA0095 & 699889 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0096 & 699890 & Guerrero, Mexico & 26    & 0.038461538 & 0.961538462 \\
+    RIMPA0097 & 699891 & Guerrero, Mexico & 6     & 0     & 1 \\
+    RIMPA0098 & 699892 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0099 & 699893 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0100 & 699894 & Guerrero, Mexico & 6     & 0     & 1 \\
+    RIMPA0101 & 699895 & Guerrero, Mexico & 2     & 0     & 1 \\
+    RIMPA0103 & 699897 & Guerrero, Mexico & 2     & 0     & 1 \\
+    RIMPA0104 & 699898 & Guerrero, Mexico & 22    & 0.090909091 & 0.909090909 \\
+    RIMPA0105 & 699899 & Guerrero, Mexico & 6     & 0     & 1 \\
+    RIMPA0106 & 699900 & Guerrero, Mexico & 6     & 0.333333333 & 0.666666667 \\
+    RIMPA0107 & 699901 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0108 & 699902 & Guerrero, Mexico & 6     & 0     & 1 \\
+    RIMPA0109 & 699903 & Michoacan, Mexico & 4     & 0.25  & 0.75 \\
+    RIMPA0110 & 699904 & Michoacan, Mexico & 2     & 0     & 1 \\
+    RIMPA0111 & 699905 & Michoacan, Mexico & 4     & 0     & 1 \\
+    RIMPA0112 & 699906 & Michoacan, Mexico & 4     & 0.25  & 0.75 \\
+    RIMPA0114 & 699908 & Michoacan, Mexico & 6     & 0.166666667 & 0.833333333 \\
+    RIMPA0116 & 699910 & Mexico, Mexico & 2     & 0     & 1 \\
+    RIMPA0117 & 699911 & Mexico, Mexico & 4     & 0     & 1 \\
+    RIMPA0118 & 699912 & Mexico, Mexico & 6     & 0.166666667 & 0.833333333 \\
+    RIMPA0119 & 699913 & Mexico, Mexico & 2     & 0     & 1 \\
+    RIMPA0120 & 699914 & Mexico, Mexico & 1     & 1     & 0 \\
+    RIMPA0121 & 699915 & Mexico, Mexico & 2     & 0     & 1 \\
+    RIMPA0128 & 699922 & Mexico, Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0129 & 699923 & Michoacan, Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0135 & 699929 & Nayarit, Mexico & 24    & 0     & 1 \\
+    RIMPA0138 & 699932 & Jalisco, Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0139 & 699933 & Jalisco, Mexico & 1     & 1     & 0 \\
+    RIMPA0142 & 699936 & Colima, Mexico & 18    & 0.444444444 & 0.555555556 \\
+    RIMPA0144 & 699938 & Jalisco, Mexico & 2     & 1     & 0 \\
+    RIMPA0145 & 699939 & Michoacan, Mexico & 1     & 1     & 0 \\
+    RIMPA0147 & 699941 & Jalisco, Mexico & 1     & 1     & 0 \\
+    RIMPA0155 & N/A   & Jalisco, Mexico & 73    & 0.01369863 & 0.98630137 \\
+    RIMPA0156 & N/A   & Jalisco, Mexico & 20    & 0     & 1 \\
+    RIMPA0157 & N/A   & Jalisco, Mexico & 58    & 0.344827586 & 0.655172414 \\
+    RIMPA0158 & N/A   & Jalisco, Mexico & 64    & 0.53125 & 0.46875 \\
+    RIMPA0159 & N/A   & Jalisco, Mexico & 26    & 0     & 1 \\
+    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
+    \bottomrule
+    \end{tabular}
+  \label{TableS1}
+\end{table}
+
+% Table S2. Mays sampling
+\begin{table}[htbp]
+  \centering
+  \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
+    \begin{tabular}{ccc}
+    \toprule
+    Accession & Number of alleles sampled & Hopscotch Frequency \\
+    \midrule
+    RIMMA0066 & 2     & 1 \\
+    RIMMA0075 & 2     & 1 \\
+    RIMMA0077 & 2     & 1 \\
+    RIMMA0079 & 2     & 1 \\
+    RIMMA0081 & 2     & 1 \\
+    RIMMA0084 & 2     & 1 \\
+    RIMMA0086 & 2     & 1 \\
+    RIMMA0088 & 2     & 1 \\
+    RIMMA0089 & 2     & 1 \\
+    RIMMA0090 & 2     & 1 \\
+    RIMMA0092 & 4     & 1 \\
+    RIMMA0094 & 4     & 1 \\
+    RIMMA0097 & 2     & 1 \\
+    RIMMA0099 & 2     & 1 \\
+    RIMMA0100 & 2     & 1 \\
+    RIMMA0101 & 2     & 1 \\
+    RIMMA0104 & 2     & 1 \\
+    RIMMA0108 & 2     & 1 \\
+    RIMMA0111 & 6     & 1 \\
+    RIMMA0115 & 2     & 1 \\
+    RIMMA0117 & 2     & 1 \\
+    RIMMA0130 & 2     & 1 \\
+    RIMMA0133 & 2     & 1 \\
+    RIMMA0134 & 2     & 1 \\
+    RIMMA0135 & 2     & 1 \\
+    RIMMA0142 & 2     & 0.5 \\
+    RIMMA0143 & 4     & 1 \\
+    RIMMA0146 & 4     & 1 \\
+    RIMMA0149 & 2     & 1 \\
+    RIMMA0152 & 2     & 1 \\
+    RIMMA0153 & 2     & 1 \\
+    RIMMA0154 & 2     & 1 \\
+    RIMMA0155 & 2     & 1 \\
+    RIMMA0156 & 2     & 1 \\
+    RIMMA0157 & 2     & 1 \\
+    RIMMA0158 & 2     & 1 \\
+    RIMMA0159 & 2     & 1 \\
+    RIMMA0160 & 2     & 1 \\
+    RIMMA0162 & 2     & 1 \\
+    RIMMA0166 & 2     & 1 \\
+    RIMMA0167 & 2     & 1 \\
+    RIMMA0168 & 2     & 1 \\
+    RIMMA0169 & 2     & 1 \\
+    RIMMA0172 & 2     & 1 \\
+    RIMMA0174 & 4     & 1 \\
+    RIMMA0177 & 2     & 1 \\
+    RIMMA0178 & 2     & 1 \\
+    RIMMA0179 & 2     & 1 \\
+    RIMMA0181 & 2     & 1 \\
+    RIMMA0183 & 2     & 1 \\
+    RIMMA0184 & 2     & 1 \\
+    RIMMA0186 & 2     & 1 \\
+    RIMMA0187 & 1     & 1 \\
+    RIMMA0188 & 2     & 1 \\
+    RIMMA0195 & 2     & 1 \\
+    RIMMA0196 & 2     & 1 \\
+    RIMMA0197 & 2     & 1 \\
+    RIMMA0198 & 2     & 1 \\
+    RIMMA0199 & 2     & 1 \\
+    RIMMA0200 & 2     & 1 \\
+    RIMMA0202 & 2     & 1 \\
+    RIMMA0203 & 2     & 1 \\
+    RIMMA0206 & 2     & 1 \\
+    RIMMA0208 & 2     & 1 \\
+    RIMMA0209 & 2     & 1 \\
+    RIMMA0210 & 2     & 1 \\
+    RIMMA0212 & 2     & 1 \\
+    RIMMA0213 & 2     & 1 \\
+    RIMMA0214 & 2     & 1 \\
+    RIMMA0217 & 2     & 1 \\
+    RIMMA0218 & 2     & 1 \\
+    RIMMA0220 & 2     & 1 \\
+    RIMMA0221 & 2     & 1 \\
+    RIMMA0222 & 2     & 1 \\
+    RIMMA0223 & 2     & 1 \\
+    RIMMA0226 & 2     & 1 \\
+    RIMMA0227 & 2     & 1 \\
+    RIMMA0228 & 2     & 1 \\
+    RIMMA0229 & 2     & 1 \\
+    RIMMA0230 & 2     & 1 \\
+    RIMMA0232 & 2     & 1 \\
+    RIMMA0233 & 2     & 1 \\
+    RIMMA0235 & 2     & 0.5 \\
+    RIMMA0242 & 2     & 1 \\
+    RIMMA0243 & 2     & 1 \\
+    RIMMA0247 & 4     & 1 \\
+    RIMMA0248 & 2     & 1 \\
+    RIMMA0249 & 2     & 1 \\
+    RIMMA0252 & 2     & 1 \\
+    RIMMA0253 & 2     & 1 \\
+    RIMMA0254 & 2     & 1 \\
+    RIMMA0256 & 2     & 1 \\
+    RIMMA0257 & 2     & 1 \\
+    RIMMA0258 & 2     & 1 \\
+    RIMMA0259 & 2     & 1 \\
+    RIMMA0260 & 2     & 1 \\
+    RIMMA0262 & 2     & 1 \\
+    RIMMA0263 & 2     & 1 \\
+    RIMMA0264 & 2     & 1 \\
+    RIMMA0265 & 2     & 1 \\
+    RIMMA0268 & 2     & 1 \\
+    RIMMA0269 & 2     & 1 \\
+    RIMMA0270 & 2     & 1 \\
+    RIMMA0272 & 2     & 1 \\
+    RIMMA0275 & 2     & 1 \\
+    RIMMA0276 & 2     & 1 \\
+    RIMMA0277 & 2     & 1 \\
+    RIMMA0279 & 2     & 1 \\
+    RIMMA0280 & 2     & 1 \\
+    RIMMA0283 & 2     & 1 \\
+    RIMMA0285 & 2     & 1 \\
+    RIMMA0288 & 2     & 1 \\
+    RIMMA0290 & 2     & 1 \\
+    RIMMA0291 & 2     & 1 \\
+    RIMMA0292 & 2     & 1 \\
+    RIMMA0293 & 2     & 1 \\
+    RIMMA0298 & 2     & 1 \\
+    RIMMA0302 & 2     & 1 \\
+    RIMMA0305 & 2     & 1 \\
+    RIMMA0310 & 2     & 1 \\
+    RIMMA0312 & 2     & 1 \\
+    RIMMA0320 & 2     & 1 \\
+    RIMMA0322 & 2     & 1 \\
+    RIMMA0324 & 2     & 1 \\
+    RIMMA0334 & 2     & 1 \\
+    RIMMA0336 & 2     & 1 \\
+    RIMMA0337 & 2     & 1 \\
+    RIMMA0338 & 2     & 1 \\
+    RIMMA0339 & 2     & 1 \\
+    RIMMA0340 & 2     & 1 \\
+    RIMMA0341 & 2     & 1 \\
+    RIMMA0342 & 2     & 1 \\
+    RIMMA0344 & 2     & 1 \\
+    RIMMA0346 & 2     & 1 \\
+    RIMMA0347 & 2     & 1 \\
+    RIMMA0350 & 2     & 1 \\
+    RIMMA0351 & 2     & 1 \\
+    RIMMA0352 & 2     & 1 \\
+    RIMMA0357 & 2     & 1 \\
+    RIMMA0360 & 18    & 1 \\
+    RIMMA0362 & 8     & 1 \\
+    RIMMA0363 & 16    & 1 \\
+    RIMMA0370 & 17    & 1 \\
+    RIMMA0372 & 24    & 1 \\
+    RIMMA0373 & 22    & 1 \\
+    RIMMA0408 & 2     & 1 \\
+    RIMMA0411 & 1     & 1 \\
+    RIMMA0413 & 1     & 1 \\
+    RIMMA0414 & 1     & 1 \\
+    RIMMA0424 & 8     & 0.9 \\
+    RIMMA0432 & 1     & 1 \\
+    RIMMA0434 & 2     & 0.5 \\
+    RIMMA0435 & 1     & 1 \\
+    RIMMA0442 & 2     & 1 \\
+    RIMMA0443 & 2     & 1 \\
+    RIMMA0444 & 2     & 1 \\
+    RIMMA0445 & 2     & 1 \\
+    RIMMA0446 & 2     & 1 \\
+    RIMMA0447 & 2     & 1 \\
+    RIMMA0448 & 2     & 1 \\
+    RIMMA0449 & 2     & 1 \\
+    RIMMA0450 & 2     & 1 \\
+    RIMMA0451 & 2     & 1 \\
+    RIMMA0452 & 2     & 1 \\
+    RIMMA0453 & 2     & 1 \\
+    RIMMA0454 & 2     & 1 \\
+    RIMMA0455 & 2     & 1 \\
+    RIMMA0456 & 2     & 1 \\
+    RIMMA0457 & 2     & 1 \\
+    RIMMA0458 & 2     & 1 \\
+    RIMMA0459 & 2     & 1 \\
+    RIMMA0483 & 2     & 1 \\
+    RIMMA0490 & 2     & 1 \\
+    RIMMA0515 & 2     & 1 \\
+    RIMMA0537 & 2     & 1 \\
+    RIMMA0550 & 2     & 1 \\
+    RIMMA0553 & 2     & 1 \\
+    RIMMA0559 & 2     & 1 \\
+    RIMMA0561 & 2     & 1 \\
+    RIMMA0562 & 2     & 1 \\
+    RIMMA0571 & 2     & 1 \\
+    RIMMA0572 & 2     & 1 \\
+    RIMMA0577 & 2     & 1 \\
+    RIMMA0579 & 2     & 1 \\
+    RIMMA0590 & 2     & 1 \\
+    RIMMA0591 & 2     & 1 \\
+    RIMMA0592 & 2     & 1 \\
+    RIMMA0593 & 2     & 1 \\
+    RIMMA0594 & 2     & 1 \\
+    RIMMA0595 & 2     & 1 \\
+    RIMMA0596 & 2     & 1 \\
+    RIMMA0597 & 2     & 1 \\
+    RIMMA0598 & 2     & 1 \\
+    RIMMA0599 & 2     & 1 \\
+    RIMMA0600 & 2     & 1 \\
+    RIMMA0601 & 2     & 1 \\
+    RIMMA0602 & 2     & 1 \\
+    RIMMA0603 & 2     & 1 \\
+    RIMMA0604 & 2     & 1 \\
+    RIMMA0605 & 2     & 1 \\
+    RIMMA0606 & 2     & 1 \\
+    RIMMA0607 & 2     & 1 \\
+    RIMMA0608 & 2     & 1 \\
+    RIMMA0609 & 2     & 1 \\
+    RIMMA0610 & 2     & 1 \\
+    RIMMA0611 & 2     & 1 \\
+    RIMMA0612 & 2     & 1 \\
+    RIMMA0613 & 2     & 1 \\
+    RIMMA0622 & 2     & 0.5 \\
+    RIMMA0624 & 1     & 1 \\
+    RIMMA0629 & 2     & 0.5 \\
+    RIMMA0631 & 2     & 0.5 \\
+    RIMMA0659 & 2     & 1 \\
+    RIMMA0660 & 2     & 1 \\
+    RIMMA0678 & 1     & 1 \\
+    RIMMA0681 & 2     & 1 \\
+    RIMMA0683 & 2     & 1 \\
+    RIMMA0684 & 2     & 1 \\
+    RIMMA0685 & 4     & 1 \\
+    RIMMA0693 & 2     & 1 \\
+    RIMMA0694 & 2     & 1 \\
+    RIMMA0695 & 1     & 1 \\
+    RIMMA0699 & 2     & 1 \\
+    RIMMA0704 & 2     & 1 \\
+    RIMMA0706 & 2     & 1 \\
+    RIMMA0711 & 1     & 1 \\
+    RIMMA0713 & 2     & 1 \\
+    RIMMA0715 & 2     & 1 \\
+    RIMMA0717 & 2     & 1 \\
+    RIMMA0718 & 2     & 1 \\
+    RIMMA0723 & 2     & 1 \\
+    RIMMA0724 & 2     & 1 \\
+    RIMMA0725 & 1     & 1 \\
+    RIMMA0728 & 1     & 1 \\
+    RIMMA0732 & 2     & 1 \\
+    RIMMA0734 & 2     & 1 \\
+    RIMMA0738 & 2     & 1 \\
+    RIMMA0739 & 2     & 1 \\
+    RIMMA0744 & 2     & 1 \\
+    RIMMA0747 & 1     & 1 \\
+    RIMMA0748 & 2     & 1 \\
+    RIMMA0749 & 2     & 1 \\
+    RIMMA0750 & 2     & 0.5 \\
+    RIMMA0751 & 1     & 1 \\
+    RIMMA0752 & 1     & 1 \\
+    RIMMA0753 & 2     & 1 \\
+    RIMMA0755 & 2     & 1 \\
+    \bottomrule
+    \end{tabular}
+  \label{TableS2Mays}
+\end{table}
+ 
+
+% Table S3. PCs for BayEnv
+\begin{table}[htbp]
+  \centering
+  \caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyh\"aj\"arvi2013}.}
+    \begin{tabular}{rrrrrrrrrrrr}
+    \toprule
+    \textbf{PC1} & \textbf{} & \textbf{PC2} & \textbf{} & \textbf{PC3} & \textbf{} & \textbf{PC4} & \textbf{} & \textbf{PC5} & \textbf{} & \textbf{PC6} & \textbf{} \\
+    \midrule
+    \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} \\
+    bio1  & 0.146 & bio4  & 0.244 & prec7 & 0.287 & ts\_clay & 0.41  & bio2  & 0.38  & bio2  & 0.412 \\
+    tmean11 & 0.146 & bio3  & 0.241 & prec8 & 0.276 & v\_mod & 0.359 & sq4   & 0.328 & x\_mod & 0.365 \\
+    tmean12 & 0.145 & bio7  & 0.241 & prec11 & 0.262 & ts\_sand & 0.329 & ts\_loam & 0.289 & sq7   & 0.332 \\
+    bio11 & 0.145 & prec6 & 0.237 & bio13 & 0.247 & bio15 & 0.272 & ts\_sand & 0.266 & bio7  & 0.312 \\
+    tmax12 & 0.145 & sq7   & 0.218 & prec1 & 0.246 & prec4 & 0.259 & sq7   & 0.231 & v\_mod & 0.307 \\
+    tmin5 & 0.145 & prec9 & 0.217 & bio16 & 0.242 & x\_mod & 0.244 & bio18 & 0.213 & prec11 & 0.23 \\
+    tmean1 & 0.145 & sq3   & 0.207 & prec12 & 0.24  & prec3 & 0.226 & bio13 & 0.207 & bio18 & 0.22 \\
+    tmean2 & 0.145 & prec12 & 0.207 & bio19 & 0.238 & sq3   & 0.21  & prec11 & 0.183 & sq3   & 0.176 \\
+    tmin4 & 0.145 & bio12 & 0.204 & bio12 & 0.231 & prec5 & 0.21  & bio7  & 0.17  & sq4   & 0.153 \\
+    tmax1 & 0.145 & bio19 & 0.196 & prec2 & 0.222 & prec7 & 0.19  & bio16 & 0.163 & ts\_sand & 0.153 \\
+    tmean4 & 0.145 & prec2 & 0.188 & bio18 & 0.221 & sq4   & 0.186 & bio4  & 0.157 & bio4  & 0.148 \\
+    tmin11 & 0.144 & prec1 & 0.185 & sq4   & 0.2   & bio3  & 0.185 & bio12 & 0.156 & prec7 & 0.139 \\
+    tmax11 & 0.144 & prec10 & 0.184 & prec9 & 0.18  & bio18 & 0.178 & bio3  & 0.155 & tmax3 & 0.127 \\
+    tmin12 & 0.144 & bio16 & 0.183 & prec10 & 0.171 & sq7   & 0.132 & prec6 & 0.154 & tmax4 & 0.126 \\
+    tmin2 & 0.144 & prec8 & 0.17  & prec5 & 0.161 & bio14 & 0.116 & x\_mod & 0.152 & bio5  & 0.118 \\
+    tmean5 & 0.144 & prec5 & 0.165 & prec4 & 0.154 & bio13 & 0.099 & prec9 & 0.144 & tmax5 & 0.115 \\
+    tmean10 & 0.144 & bio14 & 0.158 & sq3   & 0.147 & bio16 & 0.095 & prec8 & 0.143 & bio3  & 0.107 \\
+    bio6  & 0.144 & bio13 & 0.151 & bio2  & 0.143 & prec8 & 0.09  & v\_mod & 0.142 & ts\_loam & 0.095 \\
+    tmax2 & 0.144 & bio17 & 0.149 & bio17 & 0.129 & bio7  & 0.077 & bio15 & 0.136 & ts\_clay & 0.078 \\
+    tmean3 & 0.144 & prec3 & 0.144 & ts\_loam & 0.127 & bio4  & 0.075 & prec7 & 0.112 & tmin9 & 0.072 \\
+    tmin1 & 0.143 & ts\_clay & 0.141 & v\_mod & 0.123 & bio2  & 0.074 & prec4 & 0.108 & tmin8 & 0.071 \\
+    tmin10 & 0.143 & bio2  & 0.129 & prec3 & 0.113 & prec2 & 0.074 & bio14 & 0.096 & prec9 & 0.07 \\
+    Altitude & 0.143 & prec7 & 0.108 & x\_mod & 0.111 & bio19 & 0.068 & tmax7 & 0.093 & tmin10 & 0.069 \\
+    bio9  & 0.143 & tmax6 & 0.107 & bio14 & 0.099 & prec12 & 0.056 & tmax8 & 0.092 & tmin12 & 0.069 \\
+    tmin3 & 0.143 & x\_mod & 0.106 & bio4  & 0.07  & ts\_loam & 0.053 & prec1 & 0.091 & tmin7 & 0.066 \\
+    bio10 & 0.142 & bio15 & 0.098 & tmax3 & 0.067 & tmax12 & 0.047 & prec2 & 0.086 & tmean4 & 0.066 \\
+    tmax10 & 0.142 & ts\_loam & 0.088 & ts\_clay & 0.065 & bio17 & 0.047 & tmin11 & 0.086 & tmax2 & 0.066 \\
+    tmax3 & 0.142 & tmean6 & 0.085 & bio15 & 0.056 & bio9  & 0.043 & prec5 & 0.082 & tmax6 & 0.064 \\
+    tmax4 & 0.142 & tmin7 & 0.082 & tmax2 & 0.055 & tmax8 & 0.042 & bio17 & 0.082 & tmean3 & 0.064 \\
+    tmin6 & 0.142 & bio5  & 0.082 & tmean3 & 0.052 & tmax1 & 0.041 & tmin12 & 0.08  & bio9  & 0.058 \\
+    tmean9 & 0.141 & tmean7 & 0.081 & ts\_sand & 0.05  & tmax5 & 0.039 & prec3 & 0.078 & tmin11 & 0.058 \\
+    tmin9 & 0.141 & prec4 & 0.08  & prec6 & 0.048 & tmax7 & 0.039 & tmax9 & 0.078 & prec3 & 0.056 \\
+    tmean8 & 0.141 & tmax7 & 0.079 & sq7   & 0.048 & prec10 & 0.038 & tmin1 & 0.077 & tmean5 & 0.056 \\
+    bio8  & 0.14  & bio8  & 0.079 & tmin7 & 0.046 & Altitude & 0.037 & tmin10 & 0.074 & bio10 & 0.051 \\
+    tmean6 & 0.14  & tmax9 & 0.077 & bio3  & 0.044 & tmax10 & 0.035 & bio6  & 0.071 & bio15 & 0.047 \\
+    tmean7 & 0.14  & tmean8 & 0.076 & tmax4 & 0.043 & tmax2 & 0.033 & prec12 & 0.067 & tmin1 & 0.045 \\
+    tmin8 & 0.14  & tmin8 & 0.076 & bio7  & 0.042 & tmax9 & 0.03  & tmin2 & 0.061 & prec12 & 0.045 \\
+    tmax5 & 0.14  & tmax5 & 0.074 & tmax1 & 0.036 & tmean12 & 0.029 & tmin6 & 0.061 & bio17 & 0.045 \\
+    tmax9 & 0.139 & tmax8 & 0.074 & tmin3 & 0.035 & tmax11 & 0.027 & tmax6 & 0.059 & tmean10 & 0.043 \\
+    tmax8 & 0.139 & tmean9 & 0.072 & bio9  & 0.035 & tmean1 & 0.027 & tmin3 & 0.052 & prec4 & 0.043 \\
+    bio5  & 0.139 & bio18 & 0.07  & tmin8 & 0.034 & bio5  & 0.026 & bio8  & 0.046 & Altitude & 0.041 \\
+    tmax7 & 0.139 & v\_mod & 0.066 & tmean4 & 0.031 & tmean5 & 0.026 & tmean11 & 0.041 & bio6  & 0.039 \\
+    tmin7 & 0.138 & tmin9 & 0.066 & tmean2 & 0.031 & bio11 & 0.025 & tmean1 & 0.04  & tmin6 & 0.038 \\
+    tmax6 & 0.135 & bio10 & 0.065 & tmax12 & 0.03  & tmean2 & 0.022 & tmin9 & 0.04  & tmean12 & 0.035 \\
+    bio17 & 0.107 & tmax10 & 0.063 & tmean7 & 0.028 & tmin6 & 0.022 & tmean12 & 0.04  & tmean9 & 0.035 \\
+    bio14 & 0.095 & tmin6 & 0.061 & tmin9 & 0.023 & prec1 & 0.022 & tmax10 & 0.038 & prec5 & 0.033 \\
+    prec3 & 0.094 & prec11 & 0.056 & Altitude & 0.021 & tmean7 & 0.02  & tmin5 & 0.038 & tmean8 & 0.03 \\
+    bio15 & 0.092 & ts\_sand & 0.054 & tmax11 & 0.019 & tmean8 & 0.02  & bio11 & 0.037 & tmin2 & 0.03 \\
+    prec4 & 0.088 & sq4   & 0.049 & tmin4 & 0.018 & tmean9 & 0.019 & tmean8 & 0.035 & tmax1 & 0.03 \\
+    prec10 & 0.084 & tmin3 & 0.048 & tmax5 & 0.017 & prec11 & 0.019 & tmean7 & 0.033 & prec1 & 0.025 \\
+    ts\_sand & 0.073 & tmean10 & 0.047 & tmin10 & 0.017 & tmean11 & 0.018 & tmean2 & 0.03  & bio13 & 0.024 \\
+    ts\_loam & 0.072 & tmean5 & 0.044 & tmin6 & 0.017 & tmax4 & 0.016 & bio19 & 0.029 & tmean11 & 0.023 \\
+    prec6 & 0.069 & tmax4 & 0.042 & tmean6 & 0.017 & tmean10 & 0.016 & bio9  & 0.027 & tmax7 & 0.023 \\
+    prec9 & 0.063 & tmax11 & 0.041 & tmax6 & 0.017 & tmin5 & 0.015 & tmin7 & 0.026 & tmean7 & 0.023 \\
+    prec2 & 0.063 & bio6  & 0.038 & tmin12 & 0.017 & tmean6 & 0.015 & tmin4 & 0.025 & prec10 & 0.022 \\
+    bio4  & 0.06  & tmin1 & 0.038 & bio11 & 0.016 & prec6 & 0.014 & tmean5 & 0.023 & prec2 & 0.02 \\
+    prec5 & 0.058 & tmin2 & 0.037 & tmin11 & 0.015 & bio6  & 0.012 & tmean3 & 0.022 & bio8  & 0.018 \\
+    bio13 & 0.055 & Altitude & 0.036 & tmin1 & 0.013 & tmin1 & 0.012 & tmin8 & 0.022 & tmean2 & 0.017 \\
+    x\_mod & 0.053 & tmin10 & 0.034 & tmean5 & 0.013 & tmin12 & 0.011 & tmean10 & 0.021 & bio16 & 0.016 \\
+    prec1 & 0.05  & bio9  & 0.032 & tmean8 & 0.013 & tmin2 & 0.011 & tmean9 & 0.019 & tmean6 & 0.014 \\
+    bio16 & 0.049 & tmean1 & 0.029 & bio6  & 0.012 & bio1  & 0.011 & tmean4 & 0.019 & bio19 & 0.012 \\
+    bio7  & 0.043 & bio1  & 0.028 & tmean9 & 0.01  & tmin4 & 0.009 & ts\_clay & 0.018 & tmax11 & 0.012 \\
+    bio19 & 0.042 & tmean3 & 0.026 & tmean1 & 0.01  & tmax6 & 0.008 & prec10 & 0.015 & tmax10 & 0.011 \\
+    bio12 & 0.042 & tmean2 & 0.024 & bio10 & 0.01  & tmin11 & 0.007 & bio1  & 0.014 & tmax8 & 0.01 \\
+    bio3  & 0.04  & bio11 & 0.022 & tmax8 & 0.009 & bio8  & 0.007 & tmax4 & 0.014 & tmean1 & 0.009 \\
+    prec7 & 0.034 & tmean11 & 0.019 & tmax7 & 0.009 & bio10 & 0.007 & bio10 & 0.011 & bio14 & 0.009 \\
+    prec12 & 0.033 & tmin12 & 0.019 & bio5  & 0.008 & tmin9 & 0.006 & tmax5 & 0.009 & prec8 & 0.007 \\
+    bio2  & 0.033 & tmax1 & 0.017 & tmin5 & 0.008 & prec9 & 0.005 & bio5  & 0.008 & bio11 & 0.007 \\
+    prec8 & 0.032 & tmean4 & 0.016 & tmin2 & 0.007 & bio12 & 0.004 & tmax11 & 0.008 & tmin5 & 0.006 \\
+    sq4   & 0.026 & tmin5 & 0.013 & bio1  & 0.007 & tmax3 & 0.003 & tmax12 & 0.008 & tmax9 & 0.005 \\
+    sq3   & 0.024 & tmean12 & 0.012 & tmax10 & 0.006 & tmean4 & 0.003 & tmax3 & 0.007 & bio12 & 0.004 \\
+    v\_mod & 0.015 & tmax2 & 0.011 & tmean10 & 0.006 & tmin3 & 0.002 & sq3   & 0.007 & tmin3 & 0.004 \\
+    ts\_clay & 0.012 & tmin4 & 0.01  & tmean12 & 0.005 & tmin7 & 0.002 & tmax2 & 0.006 & prec6 & 0.004 \\
+    sq7   & 0.004 & tmax3 & 0.003 & bio8  & 0.004 & tmean3 & 0.002 & Altitude & 0.004 & tmin4 & 0.003 \\
+    prec11 & 0.004 & tmax12 & 0.002 & tmax9 & 0.003 & tmin10 & 0.002 & tmean6 & 0.002 & bio1  & 0.002 \\
+    bio18 & 0.001 & tmin11 & 0.001 & tmean11 & 0.001 & tmin8 & 0     & tmax1 & 0.001 & tmax12 & 0.002 \\
+    \textbf{Bayes Factors} &       &       &       &       &       &       &       &       &       &       &  \\
+    \multicolumn{2}{c}{0.77165} & \multicolumn{2}{c}{0.51652} & \multicolumn{2}{c}{0.21120} & \multicolumn{2}{c}{0.24063} & \multicolumn{2}{c}{1.96010} & \multicolumn{2}{c}{0.24594} \\
+    \bottomrule
+    \end{tabular}%
+  \label{TableS3BayEnv}%
+\end{table}%
+ 
+% Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
+\begin{table}[htbp]
+  \centering
+  \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
+    \begin{tabular}{ccccc}
+    \toprule
+    \textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
+    \midrule
+    Ejutla A & 4     & 0.15217 & 0.11902 & 0.76191 \\
+    Ejutla B & 5     & 0.15258 & 0.14877 & 0.07412 \\
+    La Mesa & 3     & 0.12802 & 0.08926 & 1.09209 \\
+    San Lorenzo & 3     & 0.09098 & 0.08926 & 0.04845 \\
+    \bottomrule
+    \end{tabular}%
+  \label{TableS4DiversitySNP}%
+\end{table}%
+ 
 \end{document}

### 2014-04-06 manuscript.tex
minor changes

@@
 BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\citep{Coop2010}.
-We used genotyping and covariance data from Pyh\"aj\"arvi et al., (2013) \citep{Pyh\"aj\"arvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
-Environmental data were obtained from www.worldclim.org, the Harmonized World Soil Database and www.harvestchoice.org, and summarized by principle component analysis \citet{Pyh\"aj\"arvi2013} \citep{Pyh\"aj\"arvi2013}.
+We used genotyping and covariance data from Pyh\"aj\"arvi et al., (2013) \citep{Pyhajarvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
+Environmental data were obtained from www.worldclim.org, the Harmonized World Soil Database and www.harvestchoice.org, and summarized by principle component analysis \citet{Pyhajarvi2013} \citep{Pyhajarvi2013}.
 
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_{\subscript{W}}$; pairwise nucleotide diversity $\theta_{\subscript{\pi}}$; and Tajima's D).
+For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_W$; pairwise nucleotide diversity $\theta_\pi$; and Tajima's D).
 To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
@@
 
-We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and the same phased 55K SNP data from \citep{Pyh\"aj\"arvi2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
+We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and the same phased 55K SNP data from \citep{Pyhajarvi2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
 SNPs were anchored in a modified version of the IBM genetic map (\citep{Gerke2013}, \url{http://arxiv.org/abs/1307.7313}).
@@
     \begin{tabular}{rccc}
-    \toprule
+ %   \toprule
     \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Sequenced Region 2} & \textbf{Sequenced Region 1} & \textbf{Hopscotch} \\
-    \midrule
+ %   \midrule
     EjuA \& EJuB & 0     & 0     & 0 \\
@@
     MSA \& SLO & 0     & 0.007 & 0.016 \\
-    \bottomrule
+%    \bottomrule
     \end{tabular}
@@
 After alignment and singleton checking we recovered 40 and 48 segregating sites for the 50kb upstream region and the 5' UTR region, respectively.
-For region 1, Ejutla A has the highest values of haplotype diversity, $\theta_{\textsubscript{W}}$, and $\theta_{\textsubscript{\pi}}$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
+For region 1, Ejutla A has the highest values of haplotype diversity, $\theta_W$, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
 Additionally, Tajima's D is similarly negative in the two Ejutla populations and La Mesa, but is more positive in San Lorenzo (\ref{Table2Diversity}).
@@
     \begin{tabular}{rccccc}
-    \toprule
-    Population & Number of Haplotypes & Haplotype Diversity & $\theta_{\textsubscript{W}}$    & $\theta_{\textsubscript{\pi}}$    & Tajima?s D \\
-    \midrule
+%    \toprule
+    Population & Number of Haplotypes & Haplotype Diversity & $\theta_W$    & $\theta_\pi$    & Tajima's D \\
+%    \midrule
     \multicolumn{6}{c}{\textit{Sequenced region 2 (50kb upstream)}} \\
@@
     \begin{tabular}{cccrc}
-    \toprule
+ %   \toprule
     \multirow{2}[2]{*}{Population} & \multirow{2}[2]{*}{Chromosome 1} & \multirow{2}[2]{*}{tb1 region} & \multicolumn{1}{c}{\multirow{2}[2]{*}{Sequenced region 1}} & \multirow{2}[2]{*}{Sequenced region 1 } \\
-    \midrule
+ %   \midrule
           &       &       & \multicolumn{1}{c}{} &  \\
@@
 The similarity in LD in the \emph{tb1} region in populations with and without the Hopscotch provides further support for an explanation other than introgression.
-Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{\Table4Q}).
+Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{Table4Q}).
 
@@
     \begin{tabular}{ccccc}
-    \toprule
+  %  \toprule
           & \multicolumn{2}{c}{\emph{tb1} region} & \multicolumn{2}{c}{Chr 1} \\
-    \midrule
+ %   \midrule
     Population & Maize assignment & Teosinte assignment & Maize assignment & Teosinte assignment \\
@@
     San Lorenzo & 0.01551389 & 0.9844861 & 0.2048252 & 0.7951748 \\
-    \bottomrule
+ %   \bottomrule
     \end{tabular}
@@
 \emph{mexicana} and \emph{Zea mays} ssp.
-{\parviglumis}) spanning the \emph{tb1} locus into an isogenic maize background and investigated their effects on previously associated phenotypes.
+\emph{parviglumis}) spanning the \emph{tb1} locus into an isogenic maize background and investigated their effects on previously associated phenotypes.
 They found that plants with teosinte chromosomal segments had greater tillering than their maize isogenic lines, and that different chromosomal segments of \emph{tb1} confer different amounts of tillering, suggesting that there are multiple genetic factors in this region that affect tillering.
@@
     \begin{tabular}{rrrrrr}
-    \toprule
+%    \toprule
     Accession & USDA Accession ID & Locality & Number alleles sampled & Hopscotch Frequency & No Hopscotch Frequency \\
-    \midrule
+ %   \midrule
     RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
@@
     RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
-    \bottomrule
+ %   \bottomrule
     \end{tabular}
@@
     \begin{tabular}{ccc}
-    \toprule
+ %   \toprule
     Accession & Number of alleles sampled & Hopscotch Frequency \\
-    \midrule
+ %   \midrule
     RIMMA0066 & 2     & 1 \\

### 2014-04-09 manuscript.tex
intro and methods edits

@@
 \author[1,3]{Tanja Pyh\"aj\"arvi}
-\author[1,4]{Matthew B. Hufford}
+\author[1,4]{Matthew B. Hufford\thanks{mbhufford@iastate.edu}}
 \author[1,5]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
@@
 Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{parviglumis} and subspecies \emph{mexicana}.
-Analysis of linkage disequilibrium between the Hopscotch element and variation (SNPs) in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
+Analysis of linkage disequilibrium between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
 Finally, two greenhouse experiments fail to find an important role for \emph{tb1} in controlling tillering in natural populations of \emph{parviglumis}.
-Our findings suggest that the role of the Hopscotch in tillering in teosinte is not as straightforward as it is in domesticated maize, and other loci known to affect branching architecture in maize may play a more important role in tillering in teosinte.
+Our findings suggest that the role of the Hopscotch in tillering in teosinte is not as straightforward as is in domesticated maize, and that other loci may play a role in observed variation in tillering in teosinte.
+
+\jri{Check citation format for AJB. I don't think it's numbers. adjust bibliography style accordingly }
 
@@
 Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
-Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{pKovachMcCouch 2008}.
-But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
+Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{KovachMcCouch2008}.
+But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \jri{CITE}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
+\jri{CITE; perhaps Arias and Rieseberg 1994 TAG?}
 
@@
 in southwest Mexico \citep{Piperno2009, Matsuoka2002}.
-Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
+\jri{shouldn't this be 9K BP?
+what does Piperno 2009 say?} Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
 Additionally, large naturally occurring populations of both \emph{Zea mays} ssp.
 \emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp.
-\emph{mexicana} (highland teosinte) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity and effective population size (Ne) of these taxa is estimated to be high \citep{Ross-Ibarra2009} relative to other plants \citep{Nybom2004}.
-
-Many morphological changes are associated with the domestication of maize, and understanding the genetic basis for these changes has been a focus of maize research for a number of years.
+\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity and effective population size (Ne) of these taxa is estimated to be high \citep{Ross-Ibarra2009} relative to other plants \citep{Nybom2004}.
+\jri{what are nybom's estimates?
+i didn't think mine were that high and don't remember Nybom giving Ne estimates}
+
+Many morphological changes are associated with the domestication of maize, and understanding the genetic basis of these changes has been a focus of maize research for a number of years\citep{Doebley2004}.
 One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches.
 The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in branching differences between maize and teosinte \citep{Doebley1990, DoebleyStec1991, LukensDoebley1999}.
-Further studies have shown that the insertion of a 4.9 kb retrotransposon (Hopscotch) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer2011}.
+\jri{an odd assortment of cites.
+why not the 1997 nature paper?} Further work showed that the insertion of a 4.9 kb retrotransposon (Hopscotch) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer2011}.
 The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in teosinte \citep{Studer2011}.
@@
 Furthermore, \citep{StuderDoebley2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
+\jri{MBH todo}
+
 Tillering, or more specifically lack of tillers in teosinte, may provide an ecological advantage to plants.
 In species such as teosinte, where plants grow densely packed together, if plants invest more resources into growing taller and fewer resources into tillering, plants may be able to out-compete neighbors for limiting resources such as light.
-In natural populations of teosinte variation at the \emph(tb1) locus may then play a role in the ecology of teosinte.
-In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{mexicana}, \emph{parviglumis}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
+\jri{there must be some citations in ecology lit to support this?} In natural populations of teosinte, variation at the \emph(tb1) locus may then play a role in the ecology of teosinte.
+In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
 We use a combination of PCR genotyping for the Hopscotch element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus.
@@
 We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter \emph{mexicana}) populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (\ref{TableS1Mays}, \ref{TableS2}).
+\emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (\ref{TableS1Mays}, \ref{TableS2}).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 Two PCRs were performed for each individual, one with primers flanking the Hopscotch (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
-Primer sequences are HopF, 5'-TCGTTGATGCTTTGATGGATGG-3';
-Hop R, 5'-AACAGTATGATTTCATGGGACCG-3'; and HopIntR, 5'-CCTCCACCTCTCATGAGATCC-3' (\ref{FigS1Locus}).
+Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
+Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, 5'-{\small CCTCCACCTCTCATGAGATCC-3'} (\ref{FigS1Locus}).
 Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (\ref{FigS2Gel}).
+\jri{fig references still need to say "Fig" etc., all the ref function does is set the number automaticall}.
 When only one PCR resolved well, we scored one allele for the individual, which explains the odd number of alleles included in our analyses.
-We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98°C for 3 min, 30 cycles of 98°C for 15 s, 65°C for 30 s, and 72°C for 3 min 30 s, with a final extension of 72°C for 10 min.
+We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98$^{\circ}$ for 3 min, 30 cycles of 98$^{\circ}$ for 15 s, 65$^{\circ}$ for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$ for 10 min.
 PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the Hopscotch based on band size.
@@
 \emph{parviglumis} with respect to maize, while the the area upstream from the Hopscotch (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \citep{Wang1999, Clark2006}.
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, 5' GGATAATGTGCACCAGGTGT 3' and 5' GCGTGCTAGAGACACYTGTTGCT 3'; for the 50 kb upstream region, 5' TGTCCTCGCCGCAACTC 3' and 5' TGTACGCCCGCCCCTCATCA 3' (\ref{FigS1Locus}).
+\jri{is this true?
+the area upstream of the hopscotch?
+those papers just found that something upstream was the causal variant, no?} We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, 5' GGATAATGTGCACCAGGTGT 3' and 5' GCGTGCTAGAGACACYTGTTGCT 3'; for the 50 kb upstream region, 5' TGTCCTCGCCGCAACTC 3' and 5' TGTACGCCCGCCCCTCATCA 3' (\ref{FigS1Locus}).
 We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94°C for 3 min, 30 cycles of 92°C for 40 s, annealing for 1 min, 72°C for 40 s, and a final 10 min extension at 72°C.
@@
 This mix was placed on a thermal cycler with the following program: 37°C for 30 min, 80°C for 15 min, and a final cool-down step to 4°C.
+\jri{redo primers with small function as above, and degrees above and below as prior}
 
@@
 
-\subsection*{Genotyping Analysis}
-
-We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE), and to calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
+\subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into githu
+
+We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).
+To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
 These analyses only included populations in which 8 or more individuals were sampled.
@@
 We used genotyping and covariance data from Pyh\"aj\"arvi et al., (2013) \citep{Pyhajarvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
-Environmental data were obtained from www.worldclim.org, the Harmonized World Soil Database and www.harvestchoice.org, and summarized by principle component analysis \citet{Pyhajarvi2013} \citep{Pyhajarvi2013}.
+Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database and \url{www.harvestchoice.org}, and summarized by principle component analysis \citet{Pyhajarvi2013} \citep{Pyhajarvi2013}.
+\jri{fix tanja cite here}
 
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_W$; pairwise nucleotide diversity $\theta_\pi$; and Tajima's D).
-To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
+For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_W$; pairwise nucleotide diversity $\hat\theta_\pi$; and Tajima's D).
+\jri{we should redo these to as shown for hat symbol}  To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) \jri{please use the url function for all web addresses} to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
 For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the Hopscotch insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citep{Chia2012}).
 
+\jri{this paragraph below is not sequence data any more. perhaps a different categorry and heading?}
 In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A, Ejutla B, San Lorenzo, and La Mesa populations \citep{Pyhajarvi2013}.
@@
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
+\jri{i'd like this perl script on github, maybe in this repo or as a gist.
+also structure input file too.
+all the stuff we'd need to redo this.}
 
@@
 To investigate the phenotypic effects of the Hopscotch insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.4375) in our initial genotyping sample set.
+We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.4375) \jri{accuracy to four digits?
+this is how many hop and no hop?} in our initial genotyping sample set.
 In order to maximize the likelihood of finding the Hopscotch in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
@@
 Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.
+\jri{i thought they were hand watered?}
 
@@
 Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Briggs2007}.
-Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
+\jri{was it brgiggs that showed this?} Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
 On day 65 we measured culm diameter between the third and fourth nodes of each plant.
@@
 Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the Hopscotch on tillering index.
+\jri{do you still have these posthoc calculations?}
 
@@
 
-Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
+Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, \jri{why?
+what criteria?} so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
 Additionally, in order to compare any association between Hopscotch genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+\jri{please add SAS scripts/code to a gist or something.
+}
 

### 2014-04-09 manuscript.tex
almost there.

@@
 
-\subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into githu
+\subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into github}
 
@@
 
-We genotyped 247 accessions of maize landraces, and found the Hopscotch element fixed in all but 8 of these accessions (\ref{TableS1Mays}, \ref{TableS2}).
-For all analyses we only included individuals that clearly resolved for both PCRs, resulting in 837 individuals total.
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (\ref{Fig1Map}).
+Genotype of the Hopscotch insertion was confirmed with two PCRs for 837 individuals.
+Among the 247 maize landrace accessions genotyped, all but 8 were homozygous for the presence of the insertion (\ref{TableS1Mays}, \ref{TableS2}).
+\jri{please fix table/figure references to say table/figure} Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations \jri{this is confusing as the map shows <37 populations.
+we should be consistent about reporting for all pops, or all pops with n>8, etc.
+or at least be explicit what cutoff we are using for each result}, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (\ref{Fig1Map}) \jri{the map only shows parviglumis.
+can we add mexicana?
+the text makes it seem as if both should be on the map}.
 We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
 Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p<0.05).
-
+\jri{in what direction?
+too many hets?
+what's the F?
+}
+
+\jri{please fix whitespace and black border on figure}
 %Figure 1 Map 
@@
   \begin{center}
-   \includegraphics[width=150mm]{Fig1Map.pdf}
+   \includegraphics[width=250mm]{Fig1Map.pdf}
     \caption{Map showing the frequency of the Hopscotch allele in populations of \emph{Zea mays} ssp.
@@
 Using our Hopscotch genotyping data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-F\textsubscript{CT} is 0 within our dataset, and we found similar levels of F\textsubscript{ST} between populations within each subspecies (0.22) and between all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi2013} (\ref{Table1Fst}).
+F\textsubscript{CT} is 0 within our dataset, and we found similar levels of F\textsubscript{ST} among populations within each subspecies (0.22) \jri{is this an average?
+this comes from libsequence?} and among all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi2013} (\ref{Table1Fst}).
+\jri{the table caption says just sequence data, but it has hopscotch genotypes too}
 % Table 1
@@
   \caption{Pairwise F\textsubscript{CT} values from sequence data}
-    \begin{tabular}{rccc}
+    \begin{tabular}{lccc}
  %   \toprule
-    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Sequenced Region 2} & \textbf{Sequenced Region 1} & \textbf{Hopscotch} \\
+    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Seq. Region 2} & \textbf{Seq.  Region 1} & \textbf{Hopscotch} \\
  %   \midrule
@@
 \end{table}
-Although we found large variation in Hopscotch allele frequency among our populations, our BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors $<$ 1; \ref{TableS3BayEnv}).
+Although we found large variation in Hopscotch allele frequency among our populations, BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors $<$ 1; \ref{TableS3BayEnv}).
 
@@
 For region 1, Ejutla A has the highest values of haplotype diversity, $\theta_W$, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is similarly negative in the two Ejutla populations and La Mesa, but is more positive in San Lorenzo (\ref{Table2Diversity}).
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is more positive in San Lorenzo (\ref{Table2Diversity}).
+\jri{can drop theta W from table and text.
+we show pi and D, which is sufficient.
+fix caption in table.
+also add Hopscotch allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats }
 % Table 2 Diversity Statistics
@@
   \caption{Add caption}
-    \begin{tabular}{rccccc}
+    \begin{tabular}{lccccc}
 %    \toprule
-    Population & Number of Haplotypes & Haplotype Diversity & $\theta_W$    & $\theta_\pi$    & Tajima's D \\
+    Population & \# Haplotypes & Hap. Diversity & $\hat\theta_W$    & $\hat\theta_\pi$    & Tajima's D \\
 %    \midrule
-    \multicolumn{6}{c}{\textit{Sequenced region 2 (50kb upstream)}} \\
+    \multicolumn{6}{c}{\textit{Seq. region 2 (50kb upstream)}} \\
     \multicolumn{1}{c}{EJUA} & 8     & 0.89394 & 0.01548 & 0.01763 & 0.6231 \\
@@
     \multicolumn{1}{c}{SLO} & 4     & 0.74242 & 0.01167 & 0.01413 & 0.93185 \\
-    \multicolumn{6}{c}{\textit{Sequenced region 1(5? UTR)}} \\
+    \multicolumn{6}{c}{\textit{Seq.  region 1(5? UTR)}} \\
     \multicolumn{1}{c}{EJUA} & 8     & 0.85897 & 0.00874 & 0.00527 & -1.64955 \\
@@
     \multicolumn{1}{c}{SLO} & 3     & 0.31818 & 0.00176 & 0.00137 & -0.72873 \\
-    \bottomrule
+ %   \bottomrule
     \end{tabular}
@@
 
-For region 2, haplotype diversity, $\theta_{\textsubscript{W}}$, and $\theta_{\textsubscript{\pi}}$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (\ref{Table2Diversity}).
-Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this latter population (\ref{Table2Diversity}).
+For region 2, haplotype diversity, $\theta_W$, and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (\ref{Table2Diversity}).
+Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this population (\ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (\ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILS; data from Maize HapMapV2, \citep{Chia2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (\ref{FigureS3NJ}).
-Furthermore, individuals within our sample that have the Hopscotch insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the Hopscotch insertion.
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citep{Chia2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (\ref{FigureS3NJ}); individuals within our sample that have the Hopscotch insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the Hopscotch insertion.
 
@@
 
-The teosinte populations with the highest frequency of the Hopscotch insertion in this study were sympatric with cultivated maize.
+The teosinte populations \jri{which?} with the highest frequency of the Hopscotch insertion in this study were sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the Hopscotch element in these populations could be attributed to introgression from maize into teosinte.
@@
 If the Hopscotch is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the Hopscotch and SNPs surrounding the \emph{tb1} region in our resequenced populations (\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region (positions 264,596,664-265,891,456; AGPv2) does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the Hopscotch genotype (average $r^{2}=0.082$).
+We find no evidence of elevated linkage disequilibrium between the Hopscotch and SNPs surrounding the \emph{tb1} region in our resequenced populations (\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the Hopscotch genotype (average $r^{2}=0.082$).
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
+\jri{table is too wide, need to round numbers, and column headers are messed up.}
 
@@
   \centering
-  \caption{$r^{2}$ values between SNPs in the \emph{tb1} region (positions 264,596,664-265,891,456) and the rest of chromosome 1, within the 5?
-UTR (Sequenced region 1), and within the 66,169 bp upstream region (Sequenced region 2).}
-    \begin{tabular}{cccrc}
+  \caption{$r^{2}$ values between SNPs in the \emph{tb1} region (positions 264,596,664-265,891,456 on chromosome 1 of the maize AGPv2 genome) and the rest of chromosome 1, within the 5' UTR (Sequenced region 1), and within the 66,169 bp upstream region (Sequenced region 2).}
+    \begin{tabular}{lccrc}
  %   \toprule
-    \multirow{2}[2]{*}{Population} & \multirow{2}[2]{*}{Chromosome 1} & \multirow{2}[2]{*}{tb1 region} & \multicolumn{1}{c}{\multirow{2}[2]{*}{Sequenced region 1}} & \multirow{2}[2]{*}{Sequenced region 1 } \\
+    \multirow{2}[2]{*}{Population} & \multirow{2}[2]{*}{Chromosome 1} & \multirow{2}[2]{*}{tb1 region} & \multicolumn{1}{c}{\multirow{2}[2]{*}{Seq. region 1}} & \multirow{2}[2]{*}{Seq. region 1 } \\
  %   \midrule
@@
 \end{table}
- 
 
@@
    \includegraphics[width=150mm]{Fig2LDPlots.pdf}
-    \caption{Linkage equilibrium plots for SNPs within bp 261,000,000-268,000,000 on chromosome 1.
+    \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
 The yellow rectangle indicates the location of the Hopscotch insertion and the green represents the \emph{tb1} ORF.
@@
 If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is lower (average 0.017) than the chromosome wide average (0.20; \ref{Fig3Structure}).
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} and much below the chromosome-wide average (0.20; \ref{Fig3Structure}).
 
@@
 %-------------------------------------------------------------------
+\jri{please put figures in the text rather than at the end.}
 
@@
 
-\subsubsection*{Phenotyping 1}
-
+\subsubsection*{Phenotyping}
+
+To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency \jri{(X)} of the Hopscotch insertion from our initial genotyping.
 We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from 23 sampling locations within the San Lorenzo population, and genotyped plants for the Hopscotch insertion.
-We found the Hopscotch segregating at a frequency of 0.65; with a 0.4 frequency of homozygotes for the Hopscotch, a 0.5 frequency of heterozygotes, and a 0.1 frequency of homozygotes for the teosinte allele (no Hopscotch), showing no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
+\jri{the 23 sampling locales can go in methods and scratch from here}.
+We found the Hopscotch segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
 After performing a repeated measures ANOVA between our transformed tillering index data and Hopscotch genotype we find a weak positive correlation between presence of the Hopscotch and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
 Additionally we find no significant correlation between tiller number and Hopscotch genotype, or culm diameter and Hopscotch genotype in phenotyping 1.
+\jri{shouldn't we expect a negative correlation betweeh Hop and TI on day 40?
+need to have an A and B in the figure and explain one is for pheno1 and one is for pheno2.
+please explain whiskers and dots on figure too.}
 
@@
 
-\subsubsection*{Phenotyping 2}
-
-For phenotyping 2 we measured tillering index every five days through day 50 for 302 plants.
-We followed the same transformations and analyses as described in phenotyping 1.
-We find the Hopscotch allele segregating at a frequency of 0.69, with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
-We find similar patterns as in phenotyping 1, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p<0.0611), with no significant correlation on any day.
+We performed a second grow-out  of teosinte to assess whether lighting conditions or sample size may have affected our ability to detect and effect of \emph{tb1}.
+For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
+We found the Hopscotch allele segregating at a frequency of 0.69, \jri{does this violate HWE?
+didn't we choose numbers of each genotype here, so that HWE isn't appropriate?
+if so, probably should drop genotype frequency data here} with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+We found similar patterns, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p<0.0611), with no significant correlation on any day.
 Similarly, relationships between Hopscotch genotype and tiller number, and hopscotch genotype and culm diameter are not significant.
@@
 The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \citep{Clark2006, Studer2011}.
-\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), and provided the first evidence that the Hopscotch insertion is segregating in teosinte \citep{Studer2011}.
+\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the Hopscotch insertion is segregating in teosinte \citep{Studer2011}.
 
@@
 While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Crop to wild introgression has previously been observed at domestication loci (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}), but our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
+Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \jri{cite matsuoka} .
+While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}),\jri{do all these show evidence of introgression AT domestication loci, or just introgression in general?}  our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
 Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.
@@
 Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-We do not have sympatric maize samples from many of our populations, however, so we cannot rule out the possibility that Hopscotch alleles in these teosinte populations are segregating in sympatric maize.
-Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1.
-The similarity in LD in the \emph{tb1} region in populations with and without the Hopscotch provides further support for an explanation other than introgression.
-Additionally, assignment to maize in this region in our STRUCTURE analysis is in fact lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{Table4Q}).
+Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{Table4Q}).
+Together, these data point to an explanation other than recent introgression for the high observed frequency of Hopscotch in some of  our \emph{parviglumis} populations.
 
@@
 Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the Hopscotch in these populations.
-In the ancient introgression scenario, recombination would have broken up linked variants surrounding the Hopscotch and we would find LD patterns similar to the ones we report here.
-However, it seems unlikely that there would have been ancient introgression between these populations of teosinte and their sympatric maize populations and not present-day introgression.
-Studies have demonstrated the presence of ongoing gene flow between domesticated maize and both \emph{Zea mays} ssp.
+If the Hopscotch allele was introgressed in the distant past, they could have been sufficient recombination to break up any initial LD, leading to observations similar to those obtained here.
+We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
+In fact, early generation maize-teosinte hybrids are easy to find in these populations today (MB Hufford, pers.
+observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{Zea mays} ssp.
 \emph{mexicana} and \emph{Zea mays} ssp.
@@
 
-Deviations from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus.
-Previous studies using both SSRs and genome-wide SNP data have provided evidence for a population bottleneck in the San Lorenzo population that has resulted in lower levels of diversity across the genome \citep{Hufford2010, Pyh\"aj\"arvi2013}.
-The lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D in the San Lorenzo population that we present here are consistent with these previous findings.
-Although, if the Hopscotch is segregating as neutral variation in teosinte, the probability that it would have simultaneously drifted to high frequency in multiple populations is quite low; however, we find no evidence of selection or introgression at or near the \emph{tb1} locus in our sample set.
-Interestingly, many of our populations with high frequency of the Hopscotch allele in the Jalisco region overlap with the Jalisco cluster in \citet{Fukunaga2005}.
-
-\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the hopscotch allele.
+Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.
+Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyh\"aj\"arvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
+\jri{deviations from HWE may be consistent too if we see excess of homozygotes.
+do we?}.
+This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyh\"aj\"arvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.
+\jri{here we need a few sentences on selection.
+the 5' UTR has much more negative D than the upstream.
+do we know the Hop genotype for sequenced lines?
+can we separate the sequences into hop/no hop and look for differences?
+it wasn't until we did this that gt1 stuff really popped out.}
+%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..
+
+%\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the hopscotch allele.
 Why?
@@
 
-The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007}.
+The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, \jri{did clark do phenotyping?} and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007}.
 Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.
@@
 \emph{gt1}, \emph{tru1}, and \emph{ter1}) in conjuction with a more exhaustive phenotyping and genotyping assay.
-
-\jri{seems like we should end with at least a hypothesis.
+\jri{why not Phyb and phya?
+these genes need citations.}
+
+%\jri{seems like we should end with at least a hypothesis.
 as is we are saying we ruled out drift, introgression, and selection.
@@
 
-\mbh{I'd lean toward saying that 1) hopscotch does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, hopscotch may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the hopscotch.}
-
-\lev{I went with Matt's suggestions.
+%\mbh{I'd lean toward saying that 1) hopscotch does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, hopscotch may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the hopscotch.}
+
+%\lev{I went with Matt's suggestions.
 I am happy to include such scenarios, but everything I have thought of seems rather unlikely; i.e.
@@
     \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red).
-Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced regions 1; within the 5?
-UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.}
+Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.}
 \label{FigS1Locus}
@@
 \end{table}%
- 
+
 % Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
@@
 \end{table}%
- 
+
 \end{document}

### 2014-04-09 manuscript.tex
next round of edits.

@@
 %-------------------------------------------------------------------
-
+\jri{lots of white space in fig 4 and fig. s1 too.} 
 We performed a second grow-out  of teosinte to assess whether lighting conditions or sample size may have affected our ability to detect and effect of \emph{tb1}.
@@
 
-The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, \jri{did clark do phenotyping?} and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007}.
+The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, \jri{did clark do phenotyping?} and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
 Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.
-However, we found no significant effect of the Hopscotch on tillering index or tiller number in our phenotyping experiments, and the effect of the Hopscotch insertion in teosinte is discordant with that of maize.
+\jri{isn't this what weber did??
+for 70+ populations!?} However, we found no significant effect of the Hopscotch on tillering index or tiller number in our phenotyping experiments, and the effect of the Hopscotch insertion in teosinte is discordant with that of maize.
 The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize.
-\lev{i took out the information about the TILS -- seemed the consensus was more than we wanted to do for this pub.
-If qPCR is working for tb1 it would be something easy to check...} It is certainly possible that even though previous data demonstrate an effect of the Hopscotch on tillering in maize \citep{Studer2011}, that the effect of the Hopscotch in teosinte is more complicated and may be more difficult to observe.
+It is certainly possible that even though previous data demonstrate an effect of the Hopscotch on tillering in maize \citep{Studer2011}, that the effect of the Hopscotch in teosinte is more complicated and may be more difficult to observe.
 Moreover, \emph{tb1} is a single gene in a complex pathway that affects branching and tillering traits, and perhaps in combination with alleles at other loci the phenotypic effects of the Hopscotch on tillering may not be consistent.
-
-Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \citep{Clark2006}, and a recent study by \citet{StuderDoebley2012} examined the possibility of an allelic series at the \emph{tb1} locus in teosinte.
+\jri{thsi section still needs work.
+i think we have to do more here.
+weber shows an association between SNPs in tb1 and branch length.
+we need to discuss that!}
+
+\jri{MBH todo} Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \citep{Clark2006}, and a recent study by \citet{StuderDoebley2012} examined the possibility of an allelic series at the \emph{tb1} locus in teosinte.
 \citet{StuderDoebley2012} introgressed 9 separate teosinte segments (one from \emph{Zea diploperennis}, and four from both \emph{Zea mays} ssp.
@@
 \citet{DoebleyStec1991, DoebleyStec1993} first attempted to map QTL controlling many of the phenotypic differences between domesticated maize and teosinte, and demonstrated the existence of numerous QTL that contribute to the differences in branching architecture between the two.
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012, PAPERS}, and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
+Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012, PAPERS}, \jri{yup, cite some PAPERS} and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
 \emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number) \citep{Li2012}.
@@
 In summary, our findings demonstrate that the Hopscotch allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
-Analysis of linkage using SNPs from across chromosome 1 does not suggest that the Hopscotch allele is present in these populations due to recent introgression; however, it seems unlikely that the it would have drifted to high frequency in multiple populations, and there may be another explanation for the high frequency we observe in some of our populations.
-The Hopscotch does not appear to uniformly reduce tillering in teosinte as it does in maize, and the other loci involved in branching architecture of maize may play key roles in the regulation of tillering in teosinte.
-Furthermore, the high frequency with which we find the Hopscotch in a number of populations may suggest that the Hopscotch may play an additional ecological role in teosinte.
+Analysis of linkage using SNPs from across chromosome 1 does not suggest that the Hopscotch allele is present in these populations due to recent introgression; however, it seems unlikely that  it would have drifted to high frequency in multiple populations and there may be another explanation for the high frequency we observe in some of our populations.
+The Hopscotch does not appear to have a strong effect reducing tillering in teosinte as it does in maize, and  other loci involved in branching architecture  may play also play roles in the regulation of tillering in teosinte.
+Finally, although we see no clear evidence of recent strong selection, the high frequency of the Hopscotch insertion in a number of populations continues to suggest to us that it plays an ecological role in teosinte.
 In the future, additional experiments will be needed to examine expression levels of \emph{tb1} and additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{ter1}) in conjuction with a more exhaustive phenotyping and genotyping assay.
+\emph{gt1}, \emph{tru1}, and \emph{ter1}) in conjunction with a more exhaustive phenotyping and genotyping assay.
 \jri{why not Phyb and phya?
-these genes need citations.}
+}
+
+\jri{please check format of supp figs and tables; some are running off the page.
+check fig/table references, bibliography, etc.
+what does "rotation" mean in supp.
+table 3?
+it isn't mentioned in methods.
+please check that all the tables and figs (including supplement) are referenced in the text.}
 

### 2014-04-15 manuscript.tex
edits
intro

@@
 Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
-Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \cite{KovachMcCouch2008}.
+Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
 But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \jri{CITE}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
-\jri{CITE; perhaps Arias and Rieseberg 1994 TAG?}
+\citep{Whitton1997}.
 
@@
 \emph{mays}) was domesticated from the teosinte \emph{Zea mays} ssp.
-\emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 10,000 B.P.
+\emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 9,000 B.P.
 in southwest Mexico \citep{Piperno2009, Matsuoka2002}.
-\jri{shouldn't this be 9K BP?
-what does Piperno 2009 say?} Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
+Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
 Additionally, large naturally occurring populations of both \emph{Zea mays} ssp.
@@
 Furthermore, \citep{StuderDoebley2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
-\jri{MBH todo}
+\jri{MBH todo} \lev{hmm?}
 

### 2014-04-18 manuscript.tex
ongoing edits

@@
 Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \jri{CITE}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
+But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
 \citep{Whitton1997}.
@@
 \emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp.
-\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity and effective population size (Ne) of these taxa is estimated to be high \citep{Ross-Ibarra2009} relative to other plants \citep{Nybom2004}.
-\jri{what are nybom's estimates?
-i didn't think mine were that high and don't remember Nybom giving Ne estimates}
+\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity and of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
 

### 2014-04-18 manuscript.tex
Edits

@@
 When only one PCR resolved well, we scored one allele for the individual, which explains the odd number of alleles included in our analyses.
-We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98$^{\circ}$ for 3 min, 30 cycles of 98$^{\circ}$ for 15 s, 65$^{\circ}$ for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$ for 10 min.
+We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
 PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the Hopscotch based on band size.
@@
 the area upstream of the hopscotch?
-those papers just found that something upstream was the causal variant, no?} We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, 5' GGATAATGTGCACCAGGTGT 3' and 5' GCGTGCTAGAGACACYTGTTGCT 3'; for the 50 kb upstream region, 5' TGTCCTCGCCGCAACTC 3' and 5' TGTACGCCCGCCCCTCATCA 3' (\ref{FigS1Locus}).
-We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94°C for 3 min, 30 cycles of 92°C for 40 s, annealing for 1 min, 72°C for 40 s, and a final 10 min extension at 72°C.
-Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7°C and 58.8°C, respectively.
+those papers just found that something upstream was the causal variant, no?} We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5' GGATAATGTGCACCAGGTGT 3'} and {\small 5' GCGTGCTAGAGACACYTGTTGCT 3'}; for the 50 kb upstream region, {\small 5' TGTCCTCGCCGCAACTC 3'} and {5' TGTACGCCCGCCCCTCATCA 3'} (\ref{FigS1Locus}).
+We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$°C for 3 min, 30 cycles of 92°C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
+Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
 To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 µL of amplification product.
-This mix was placed on a thermal cycler with the following program: 37°C for 30 min, 80°C for 15 min, and a final cool-down step to 4°C.
-\jri{redo primers with small function as above, and degrees above and below as prior}
+This mix was placed on a thermal cycler with the following program: 37$^{\circ}$C for 30 min, 80$^{\circ}$C for 15 min, and a final cool-down step to 4$^{\circ}$C.
 
@@
 coli} cells, with an extended ligation time of 30 min for a complex target fragment.
-We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96°C for 5 min; then 35 cycles at 96°C for 30 s, 53°C for 30 s, 72°C for two min; and a final extension at 72°C for 4 min.
+We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96$^{\circ}$C for 5 min; then 35 cycles at 96$^{\circ}$C for 30 s, 53$^{\circ}$C for 30 s, 72$^{\circ}$C for two min; and a final extension at 72$^{\circ}$C for 4 min.
 We visualized amplification products for incorporation of our insert on a 1\% agarose TAE gel.
@@
 
-\subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into github}
+\subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into github} \lev{which scripts are you talking about?
+I used matt's for making structure chunks, and then just command line to run structure, same with BayEnv -- it was Tanjas covariance matrix and then just the basic bayenv commands}
 
@@
 BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\citep{Coop2010}.
-We used genotyping and covariance data from Pyh\"aj\"arvi et al., (2013) \citep{Pyhajarvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database and \url{www.harvestchoice.org}, and summarized by principle component analysis \citet{Pyhajarvi2013} \citep{Pyhajarvi2013}.
-\jri{fix tanja cite here}
+We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
+Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database and \url{www.harvestchoice.org}, and summarized by principle component analysis \citep{Pyhajarvi2013}.
 
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\theta_W$; pairwise nucleotide diversity $\hat\theta_\pi$; and Tajima's D).
-\jri{we should redo these to as shown for hat symbol}  To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (http://evolution.genetics.washington.edu/phylip.html) \jri{please use the url function for all web addresses} to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
+For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\hat\theta_W$; pairwise nucleotide diversity $\hat\theta_\pi$; and Tajima's D).
+To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
 For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the Hopscotch insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citep{Chia2012}).
 
+\subsection*{Introgression Analysis}
 \jri{this paragraph below is not sequence data any more. perhaps a different categorry and heading?}
@@
 also structure input file too.
-all the stuff we'd need to redo this.}
+all the stuff we'd need to redo this.} \lev{See above note..not all of this or the BayEnv was script'ified}
 
@@
 To investigate the phenotypic effects of the Hopscotch insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.4375) \jri{accuracy to four digits?
-this is how many hop and no hop?} in our initial genotyping sample set.
+We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.44) in our initial genotyping sample set.
 In order to maximize the likelihood of finding the Hopscotch in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
@@
 Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.
-\jri{i thought they were hand watered?}
+\lev{it ended up being a combination of drip and hand watering because they dried out so much and did better when they had water on the leaves as well as in the soil}
 
@@
 Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Briggs2007}.
-\jri{was it brgiggs that showed this?} Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
+\jri{was it brgiggs that showed this?} \lev{I believe so, was first to characterize TILL phenotype in backcross} Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
 On day 65 we measured culm diameter between the third and fourth nodes of each plant.
@@
 Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the Hopscotch on tillering index.
-\jri{do you still have these posthoc calculations?}
+\jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?}
 
@@
 Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, \jri{why?
-what criteria?} so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
+what criteria?} \lev{basically none of the criteria -- even sample sizes in classes, equal variances, normal distribution, etc.} so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).

### 2014-04-21 manuscript.tex
edits

@@
 what's the F?
-}
+} \lev{they weren't all in the same direction, do you still want me to list out?}
 
@@
 this comes from libsequence?} and among all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi2013} (\ref{Table1Fst}).
-\jri{the table caption says just sequence data, but it has hopscotch genotypes too}
+
 % Table 1
@@
   \centering
-  \caption{Pairwise F\textsubscript{CT} values from sequence data}
+  \caption{Pairwise F\textsubscript{CT} values from sequence and Hopscotch genotyping data}
     \begin{tabular}{lccc}

### 2014-04-21 manuscript.tex
Edits + Tom's comments on .pdf

@@
 After alignment and singleton checking we recovered 40 and 48 segregating sites for the 50kb upstream region and the 5' UTR region, respectively.
-For region 1, Ejutla A has the highest values of haplotype diversity, $\theta_W$, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
+For region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
 Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is more positive in San Lorenzo (\ref{Table2Diversity}).
@@
 fix caption in table.
-also add Hopscotch allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats }
+also add Hopscotch allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats } \lev{the frequencies are in supplemental table 1, do you want them put in this table too, or just me to refer people to supp table 1?}
 % Table 2 Diversity Statistics

### 2014-04-28 manuscript.tex
edits
continued text edits

@@
 \subsection*{Introgression Analysis}
-\jri{this paragraph below is not sequence data any more. perhaps a different categorry and heading?}
+
 In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A, Ejutla B, San Lorenzo, and La Mesa populations \citep{Pyhajarvi2013}.
@@
 need to have an A and B in the figure and explain one is for pheno1 and one is for pheno2.
-please explain whiskers and dots on figure too.}
+please explain whiskers and dots on figure too.} \lev{sure, I mean, presumably we would expect things with Hop to have a smaller TI}
 
@@
 didn't we choose numbers of each genotype here, so that HWE isn't appropriate?
-if so, probably should drop genotype frequency data here} with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+if so, probably should drop genotype frequency data here} \lev{I don't believe we ended up culling individuals because we had issues with the genotyping for phenotypine 2} with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
 We found similar patterns, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p<0.0611), with no significant correlation on any day.

### 2014-04-28 manuscript.tex
edits
text

@@
 While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \jri{cite matsuoka} .
-While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}),\jri{do all these show evidence of introgression AT domestication loci, or just introgression in general?}  our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
+Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}.
+While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}),\jri{do all these show evidence of introgression AT domestication loci, or just introgression in general?}  \lev{just introgression in general, I believe Thurber and Huff looked at specific domestication loci} our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
 Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.
@@
 can we separate the sequences into hop/no hop and look for differences?
-it wasn't until we did this that gt1 stuff really popped out.}
+it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
 %   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..

### 2014-04-30 manuscript.tex
edits + TP comments on .pdf

@@
 
-The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, \jri{did clark do phenotyping?} and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
+The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, \jri{did clark do phenotyping?} \lev{yes they phenotyped their bc1 i believe, 8 traits shown to correlate with tb1 variation} and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
 Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.

### 2014-05-01 manuscript.tex
edits

@@
 
+%TO DO
+%Fix figures and tables
+%add usepackage for tables
+%check citations in last paragraph for term
+%reformat for journal
+%finish fixing paragraph that jeff doesn't like
+
 \subsection*{Abstract}
@@
 
-\jri{Check citation format for AJB. I don't think it's numbers. adjust bibliography style accordingly }
+\jri{Check citation format for AJB.
+I don't think it's numbers.
+adjust bibliography style accordingly } \lev{I would rather send to an open access journal.
+I don't like that no everyone would be able to have access to paper}
 
@@
 \citet{DoebleyStec1991, DoebleyStec1993} first attempted to map QTL controlling many of the phenotypic differences between domesticated maize and teosinte, and demonstrated the existence of numerous QTL that contribute to the differences in branching architecture between the two.
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012, PAPERS}, \jri{yup, cite some PAPERS} and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
+Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012}, \jri{yup, cite some PAPERS} \lev{hah oops i missed taking out that note...
+:)} and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
 \emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number) \citep{Li2012}.
@@
 \jri{why not Phyb and phya?
-}
+} \lev{Are they necessary to include?
+I'd had them in before in a paragraph but had been voted out}
 

### 2014-05-02 manuscript.tex
edits
working edits from TK and TP

@@
 %finish fixing paragraph that jeff doesn't like
+%Tanja and rest of Yoshitaka comments
 
@@
 Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the Hopscotch element, led us to investigate the role of \emph{tb1} in teosinte.
-Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{parviglumis} and subspecies \emph{mexicana}.
+Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{Zea mays} ssp.
+\emph{parviglumis} and subspecies \emph{Zea mays} ssp.
+\emph{mexicana}.
 Analysis of linkage disequilibrium between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.

### 2014-05-02 manuscript.tex
working edits
edits from tp and yoshitaka

@@
 those papers just found that something upstream was the causal variant, no?} We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5' GGATAATGTGCACCAGGTGT 3'} and {\small 5' GCGTGCTAGAGACACYTGTTGCT 3'}; for the 50 kb upstream region, {\small 5' TGTCCTCGCCGCAACTC 3'} and {5' TGTACGCCCGCCCCTCATCA 3'} (\ref{FigS1Locus}).
-We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$°C for 3 min, 30 cycles of 92°C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
+We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
 Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.

### 2014-05-02 manuscript.tex
new comments, responses.

@@
 adjust bibliography style accordingly } \lev{I would rather send to an open access journal.
-I don't like that no everyone would be able to have access to paper}
+I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for $500 which is cheaper than PLoS ONE}
 
@@
 Furthermore, \citep{StuderDoebley2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
-\jri{MBH todo} \lev{hmm?}
 
@@
 Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
-To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 µL of amplification product.
+To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 $\mu$ L of amplification product.
 This mix was placed on a thermal cycler with the following program: 37$^{\circ}$C for 30 min, 80$^{\circ}$C for 15 min, and a final cool-down step to 4$^{\circ}$C.
@@
 \subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into github} \lev{which scripts are you talking about?
-I used matt's for making structure chunks, and then just command line to run structure, same with BayEnv -- it was Tanjas covariance matrix and then just the basic bayenv commands}
+I used matt's for making structure chunks, and then just command line to run structure, same with BayEnv -- it was Tanjas covariance matrix and then just the basic bayenv commands} \jri{matt's scritp for structure should be up there.
+and even a readme documenting the commandline for bayenv and including the covariance matrix.
+basically whatever someone would need to be able to literally repeat what you did.}
 
@@
 also structure input file too.
-all the stuff we'd need to redo this.} \lev{See above note..not all of this or the BayEnv was script'ified}
+all the stuff we'd need to redo this.} \lev{See above note..not all of this or the BayEnv was script'ified} \jri{sure, even command line info should be included where possible.
+idea is to maximize reproducibility -- either other people or subsequent students.
+for example, matt has a student who wants to work on tb1 in natural pops, and she might want to try/redo some of these analyses with the same or new data.}
 
@@
 Successful germinations (206 individuals) were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches.
-Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.
-\lev{it ended up being a combination of drip and hand watering because they dried out so much and did better when they had water on the leaves as well as in the soil}
+Plants were watered three times a day.
 
@@
 Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Briggs2007}.
-\jri{was it brgiggs that showed this?} \lev{I believe so, was first to characterize TILL phenotype in backcross} Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
+\jri{was it brgiggs that showed this?} \lev{I believe so, was first to characterize TILL phenotype in backcross} \jri{I think this cite should probably be one of the earlier Doebley papers mapping to tb1 QTL, as those showed the QTL had effect on plant architecture, no?} Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
 On day 65 we measured culm diameter between the third and fourth nodes of each plant.
@@
 Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the Hopscotch on tillering index.
-\jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?}
+\jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?} \jri{would be good to include in a github, again so we could go back and reassess how we do things in case, for example, we decide to give the greenhouse experiment a 3rd try (yes, i'm a masochist)}
 
@@
 
-Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, \jri{why?
-what criteria?} \lev{basically none of the criteria -- even sample sizes in classes, equal variances, normal distribution, etc.} so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
+Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
@@
 what's the F?
-} \lev{they weren't all in the same direction, do you still want me to list out?}
+} \lev{they weren't all in the same direction, do you still want me to list out?} \jri{no need to list all, but if there were sig.
+deviations in multiple directions in different pops, that is worth saying.
+maybe could list them all out in supp.
+table?}
 
@@
 fix caption in table.
-also add Hopscotch allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats } \lev{the frequencies are in supplemental table 1, do you want them put in this table too, or just me to refer people to supp table 1?}
+also add Hopscotch allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats } \lev{the frequencies are in supplemental table 1, do you want them put in this table too, or just me to refer people to supp table 1?} \jri{yeah refer to table is fine}
 % Table 2 Diversity Statistics
@@
 need to have an A and B in the figure and explain one is for pheno1 and one is for pheno2.
-please explain whiskers and dots on figure too.} \lev{sure, I mean, presumably we would expect things with Hop to have a smaller TI}
+please explain whiskers and dots on figure too.} \lev{sure, I mean, presumably we would expect things with Hop to have a smaller TI} \jri{yup, but we should mention that the expectation is negative/}
 
@@
 For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
-We found the Hopscotch allele segregating at a frequency of 0.69, \jri{does this violate HWE?
-didn't we choose numbers of each genotype here, so that HWE isn't appropriate?
-if so, probably should drop genotype frequency data here} \lev{I don't believe we ended up culling individuals because we had issues with the genotyping for phenotypine 2} with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+We found the Hopscotch allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
 We found similar patterns, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p<0.0611), with no significant correlation on any day.
@@
 Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}.
-While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}),\jri{do all these show evidence of introgression AT domestication loci, or just introgression in general?}  \lev{just introgression in general, I believe Thurber and Huff looked at specific domestication loci} our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
+While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}), our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
 Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.
@@
 
-The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, \jri{did clark do phenotyping?} \lev{yes they phenotyped their bc1 i believe, 8 traits shown to correlate with tb1 variation} and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
+The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
 Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.
@@
 \citet{DoebleyStec1991, DoebleyStec1993} first attempted to map QTL controlling many of the phenotypic differences between domesticated maize and teosinte, and demonstrated the existence of numerous QTL that contribute to the differences in branching architecture between the two.
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012}, \jri{yup, cite some PAPERS} \lev{hah oops i missed taking out that note...
-:)} and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
+Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
 \emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number) \citep{Li2012}.
@@
 } \lev{Are they necessary to include?
-I'd had them in before in a paragraph but had been voted out}
+I'd had them in before in a paragraph but had been voted out} \jri{I'd ditch gt1 tru1 ter1 and maybe just cite some people including phyb etc.}
 
 \jri{please check format of supp figs and tables; some are running off the page.
+you can use "longtable" to fix that (ask Paul for example).
 check fig/table references, bibliography, etc.

### 2014-05-11 manuscript.tex
edits
working through TP and TKs text edits; updated refs

--- 2014-05-02 manuscript.tex
@@
 
-In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A, Ejutla B, San Lorenzo, and La Mesa populations \citep{Pyhajarvi2013}.
+In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
 We chose these populations because we had both genotyping data for the Hopscotch as well as chromosome-wide SNP data for chromosome 1.
-For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for San Lorenzo, Ejutla B, Ejutla A, and La Mesa, respectively.
+For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for SLO, EjuB, EjuA, and MSA, respectively.
 We then used Tassel \citep{Bradbury2007} to calculate linkage disequilibrium ($r^{2}$) across chromosome 1 for each population.
@@
 We treated seeds with fungicide and germinated them in petri dishes with filter paper.
-Successful germinations (206 individuals) were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches.
+Following germination, 206 successful germinations were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches.
 Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.

### 2014-05-13 manuscript.tex
ongoing edits
edits

--- 2014-05-02 manuscript.tex
@@
 %finish fixing paragraph that jeff doesn't like
-%Tanja and rest of Yoshitaka comments
 
@@
 idea is to maximize reproducibility -- either other people or subsequent students.
-for example, matt has a student who wants to work on tb1 in natural pops, and she might want to try/redo some of these analyses with the same or new data.}
+for example, matt has a student who wants to work on tb1 in natural pops, and she might want to try/redo some of these analyses with the same or new data.} \lev{okay so I should just put my command line stuff with good commenting as to what is what in repository?}
 
 \subsection*{Phenotyping of \emph{Zea mays.} ssp. \emph{parviglumis}}
+
+To investigate the phenotypic effects of the Hopscotch insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
+We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.44) in our initial genotyping sample set.
+In order to maximize the likelihood of finding the Hopscotch in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
+We chose between 10-13 seeds from each of 23 sampling sites.
+We treated seeds with fungicide and germinated them in petri dishes with filter paper.
+Following germination, 206 successful germinations were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches.
+Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.
+\lev{it ended up being a combination of drip and hand watering because they dried out so much and did better when they had water on the leaves as well as in the soil}
 
@@
 Starting on day 15, we measured tillering index, the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
-Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Briggs2007}.
-\jri{was it brgiggs that showed this?} \lev{I believe so, was first to characterize TILL phenotype in backcross} \jri{I think this cite should probably be one of the earlier Doebley papers mapping to tb1 QTL, as those showed the QTL had effect on plant architecture, no?} Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
+Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Clark2006}.
+Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
 On day 65 we measured culm diameter between the third and fourth nodes of each plant.
@@
 Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the Hopscotch on tillering index.
-\jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?} \jri{would be good to include in a github, again so we could go back and reassess how we do things in case, for example, we decide to give the greenhouse experiment a 3rd try (yes, i'm a masochist)}
+\jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?} \jri{would be good to include in a github, again so we could go back and reassess how we do things in case, for example, we decide to give the greenhouse experiment a 3rd try (yes, i'm a masochist)} \lev{I would do it a 3rd time, I'm convinced something went wrong....third time with expression analyses :) }
 
@@
 If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} and much below the chromosome-wide average (0.20; \ref{Fig3Structure}).
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} \lev{Yes really 0.017 assignment to maize in the tb1 region, and avg assignment across chr1 is 0.2} \lev{ I also have a table of assignment values for SLO individuals based on genotype.
+Though we had decided this wasn't super informative because sample size was low} and much below the chromosome-wide average (0.20; \ref{Fig3Structure}).
 
@@
   \begin{center}
-   \includegraphics[width=150mm]{Fig3Structure.pdf}
+   \includegraphics[width=90mm]{Fig3Structure.pdf}
     \caption{STRUCTURE assignment to maize across a section of chromosome 1.
@@
 %-------------------------------------------------------------------
-\jri{please put figures in the text rather than at the end.}
+\jri{please put figures in the text rather than at the end.} \lev{I can't figure out why the figures are going at the end. I did them following the format in Sofiane's 282 paper}
 
@@
 
-To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency \jri{(X)} of the Hopscotch insertion from our initial genotyping.
-We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from 23 sampling locations within the San Lorenzo population, and genotyped plants for the Hopscotch insertion.
-\jri{the 23 sampling locales can go in methods and scratch from here}.
+To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the Hopscotch insertion from our initial genotyping.
+We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from within the San Lorenzo population, and genotyped plants for the Hopscotch insertion.
 We found the Hopscotch segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.

### 2014-05-13 manuscript.tex
edits
updated to do list for LV
fixed numbers in tables

@@
 %add usepackage for tables
-%check citations in last paragraph for term
-%reformat for journal
+%check citations in last paragraph for ter1
+%reformat for AJB
+%preprint somewhere
 %finish fixing paragraph that jeff doesn't like
+%check map details
+%make figs smaller so they don't go to end of document
+%Code for SAS and command line stuff on Git
+%Phenotyping and genotyping data on Git (it's on Figshare?)
 
@@
     Ejutla B & 0.01888194 & 0.9811181 & 0.1872131 & 0.8127869 \\
-    La Mesa & 0.9881333 & 0.0118675 & 0.8068998 & 0.1931017 \\
+    La Mesa & 0.0118675 & 0.9881333 & 0.8068998 & 0.1931017 \\
     San Lorenzo & 0.01551389 & 0.9844861 & 0.2048252 & 0.7951748 \\

### 2014-05-15 manuscript.tex
ongoing edits, problem with captions

@@
 \usepackage{color}
+\usepackage{multirow}
+\usepackage{booktabs}
 \usepackage[pdftex]{graphicx}  
@@
 %----------------------------------------
-\title{Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)}
+\title{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}
 
@@
 \author[1,3]{Tanja Pyh\"aj\"arvi}
-\author[1,4]{Matthew B. Hufford\thanks{mbhufford@iastate.edu}}
+\author[1,4]{Matthew B. Hufford\thanks{mhufford@iastate.edu}}
 \author[1,5]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
@@
 adjust bibliography style accordingly } \lev{I would rather send to an open access journal.
-I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for $500 which is cheaper than PLoS ONE}
+I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for \$500 which is cheaper than PLoS ONE}
 
@@
 The lack of clustering of Hopscotch genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the Hopscotch insertion in these populations of \emph{parviglumis} is the result of recent introgression.
-However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Phyh\"aj\"arvi2013}.
+However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}.
 The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations.
@@
 Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.
-Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyh\"aj\"arvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
+Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
 \jri{deviations from HWE may be consistent too if we see excess of homozygotes.
 do we?}.
-This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyh\"aj\"arvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.
+This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.
 \jri{here we need a few sentences on selection.
@@
    \includegraphics[width=150mm]{FigS1LocusCartoon.pdf}
-    \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red).
+   % \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red).
 Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.}
@@
    \includegraphics[width=150mm]{FigS2Gel.png}
-    \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+   % \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
 \label{FigS2Gel}
@@
    \includegraphics[width=150mm]{FigS3NJtrees.pdf}
-    \caption{Neighbor-Joining tree of the sequenced region in the 5?
+    %\caption{Neighbor-Joining tree of the sequenced region in the 5?
 UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}.
@@
   \centering
-  \caption{Accessions of \emph{Zea mays} ssp.
+  %\caption{Accessions of \emph{Zea mays} ssp.
 \emph{mexicana} (RIMME) and \emph{Zea mays} ssp.
@@
   \centering
-  \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
+ % \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
     \begin{tabular}{ccc}
@@
   \centering
-  \caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyh\"aj\"arvi2013}.}
+  %\caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}
     \begin{tabular}{rrrrrrrrrrrr}
@@
   \centering
-  \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
+  %\caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
     \begin{tabular}{ccccc}
     \toprule
-    \textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
+    %\textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
     \midrule

### 2014-05-16 manuscript.tex
additional edits, .bst file created for AJB

@@
 \renewcommand{\baselinestretch}{1.5}
-\bibliographystyle{apalike}
+\bibliographystyle{AJB}
 
@@
 adjust bibliography style accordingly } \lev{I would rather send to an open access journal.
-I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for \$500 which is cheaper than PLoS ONE}
+I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for \$500 which is cheaper than PLoS ONE} \mbh{I created a .bst file for AJB...still some problems with references in bibtex file that need to be fixed}
 

### 2014-05-16 manuscript.tex
figure placement, whitespace fixed for fig 1.
edits to abstract, Hopscotch formatting throughout

--- 2014-05-15 manuscript.tex
@@
 The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
-Further studies have shown that the insertion of a transposable element (Hopscotch) upstream of \emph{tb1} enhances its expression, causing the reduction in branching observed in domesticated maize.
-Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the Hopscotch element, led us to investigate the role of \emph{tb1} in teosinte.
-Results from genotyping across many natural populations suggest that the Hopscotch element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{Zea mays} ssp.
+Further studies have shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances its expression, causing the reduction in branching observed in domesticated maize.
+Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the \emph{Hopscotch} element, led us to investigate the prevalence and potential role of \emph{tb1} in teosinte.
+Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{Zea mays} ssp.
 \emph{parviglumis} and subspecies \emph{Zea mays} ssp.
 \emph{mexicana}.
-Analysis of linkage disequilibrium between the Hopscotch element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
+Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
 Finally, two greenhouse experiments fail to find an important role for \emph{tb1} in controlling tillering in natural populations of \emph{parviglumis}.
-Our findings suggest that the role of the Hopscotch in tillering in teosinte is not as straightforward as is in domesticated maize, and that other loci may play a role in observed variation in tillering in teosinte.
-
-\jri{Check citation format for AJB.
-I don't think it's numbers.
-adjust bibliography style accordingly } \lev{I would rather send to an open access journal.
-I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for \$500 which is cheaper than PLoS ONE}
+Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte is not as straightforward as is in domesticated maize, and that other loci may play a role in observed variation in tillering in teosinte.
 
@@
 \jri{an odd assortment of cites.
-why not the 1997 nature paper?} Further work showed that the insertion of a 4.9 kb retrotransposon (Hopscotch) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer2011}.
+why not the 1997 nature paper?} Further work showed that the insertion of a 4.9 kb retrotransposon (\emph{Hopscotch}) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer2011}.
 The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in teosinte \citep{Studer2011}.
@@
 \jri{there must be some citations in ecology lit to support this?} In natural populations of teosinte, variation at the \emph(tb1) locus may then play a role in the ecology of teosinte.
-In this study we aim to characterize the distribution of the Hopscotch insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
-We use a combination of PCR genotyping for the Hopscotch element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus.
-Finally, we test for an association between the Hopscotch element and tillering phenotypes in a population of \emph{parviglumis}.
+In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
+We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus.
+Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.
 
@@
 We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{Zea mays} ssp.
-\emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the Hopscotch insertion (\ref{TableS1Mays}, \ref{TableS2}).
+\emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (\ref{TableS1Mays}, \ref{TableS2}).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
-We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \cite{Kearse2012} to amplify the entire Hopscotch element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the Hopscotch insertion.
-Two PCRs were performed for each individual, one with primers flanking the Hopscotch (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
+We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \cite{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the \emph{Hopscotch} insertion.
+Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
 Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, 5'-{\small CCTCCACCTCTCATGAGATCC-3'} (\ref{FigS1Locus}).
-Homozygotes show a single band for either the Hopscotch element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the Hopscotch element as well as a band for the internal primer set (\ref{FigS2Gel}).
+Homozygotes show a single band for either the \emph{Hopscotch} element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the \emph{Hopscotch} element as well as a band for the internal primer set (\ref{FigS2Gel}).
 \jri{fig references still need to say "Fig" etc., all the ref function does is set the number automaticall}.
@@
 We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
-PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the Hopscotch based on band size.
+PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.
 
@@
 We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of the start of the Hopscotch and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
+For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of the start of the \emph{Hopscotch} and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
 The 5' UTR (containing sequenced region 1) has been shown to have elevated diversity in ssp.
-\emph{parviglumis} with respect to maize, while the the area upstream from the Hopscotch (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \citep{Wang1999, Clark2006}.
+\emph{parviglumis} with respect to maize, while the the area upstream from the \emph{Hopscotch} (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \citep{Wang1999, Clark2006}.
 \jri{is this true?
-the area upstream of the hopscotch?
+the area upstream of the \emph{Hopscotch}?
 those papers just found that something upstream was the causal variant, no?} We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5' GGATAATGTGCACCAGGTGT 3'} and {\small 5' GCGTGCTAGAGACACYTGTTGCT 3'}; for the 50 kb upstream region, {\small 5' TGTCCTCGCCGCAACTC 3'} and {5' TGTACGCCCGCCCCTCATCA 3'} (\ref{FigS1Locus}).
@@
 These analyses only included populations in which 8 or more individuals were sampled.
-To test the hypothesis that the Hopscotch insertion may be adaptive under certain environmental conditions, we looked for significant associations between the Hopscotch frequency and environmental variables using BayEnv \citep{Coop2010}.
+To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
 BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\citep{Coop2010}.
-We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the Hopscotch insertion coded as an additional SNP (\ref{TableS3BayEnv}).
+We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (\ref{TableS3BayEnv}).
 Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database and \url{www.harvestchoice.org}, and summarized by principle component analysis \citep{Pyhajarvi2013}.
@@
 For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\hat\theta_W$; pairwise nucleotide diversity $\hat\theta_\pi$; and Tajima's D).
-To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by Hopscotch genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
-For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the Hopscotch insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citep{Chia2012}).
+To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
+For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citep{Chia2012}).
 
@@
 
-In order to assess patterns of linkage disequilibrium (LD) around the Hopscotch element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
-We chose these populations because we had both genotyping data for the Hopscotch as well as chromosome-wide SNP data for chromosome 1.
+In order to assess patterns of linkage disequilibrium (LD) around the \emph{Hopscotch} element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
+We chose these populations because we had both genotyping data for the \emph{Hopscotch} as well as chromosome-wide SNP data for chromosome 1.
 For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for SLO, EjuB, EjuA, and MSA, respectively.
@@
 
-To investigate the phenotypic effects of the Hopscotch insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the Hopscotch is segregating at highest frequency (0.44) in our initial genotyping sample set.
-In order to maximize the likelihood of finding the Hopscotch in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
+To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
+We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
+In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
 We chose between 10-13 seeds from each of 23 sampling sites.
@@
 Starting on day 15, we measured tillering index, the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
-Tillering index has been shown to be the most effective way to observe the phenotypic effects of the Hopscotch insertion on plant architecture in maize \citep{Clark2006}.
+Tillering index has been shown to be the most effective way to observe the phenotypic effects of the \emph{Hopscotch} insertion on plant architecture in maize \citep{Clark2006}.
 Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
@@
 Culm diameter is not believed to be correlated with tillering index, or variation at \emph{tb1} (e.g.
-Hopscotch genotype).
+\emph{Hopscotch} genotype).
 Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (http://www.ars.usda.gov).
-We genotyped individuals for the Hopscotch insertion following the protocols listed above.
-Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the Hopscotch on tillering index.
+We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
+Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
 \jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?} \jri{would be good to include in a github, again so we could go back and reassess how we do things in case, for example, we decide to give the greenhouse experiment a 3rd try (yes, i'm a masochist)} \lev{I would do it a 3rd time, I'm convinced something went wrong....third time with expression analyses :) }
 
-We performed a second phenotyping experiment (phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the Hopscotch insertion.
+We performed a second phenotyping experiment (phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the \emph{Hopscotch} insertion.
 Seeds were germinated and planted on day 7 post fruit-case removal into 2 gallon pots.
@@
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
-Additionally, in order to compare any association between Hopscotch genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
 \jri{please add SAS scripts/code to a gist or something.
@@
 
-Genotype of the Hopscotch insertion was confirmed with two PCRs for 837 individuals.
+Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
 Among the 247 maize landrace accessions genotyped, all but 8 were homozygous for the presence of the insertion (\ref{TableS1Mays}, \ref{TableS2}).
-\jri{please fix table/figure references to say table/figure} Within our \emph{parviglumis} and \emph{mexicana} samples we found the Hopscotch insertion segregating in 37 and 4 populations \jri{this is confusing as the map shows <37 populations.
+\jri{please fix table/figure references to say table/figure} Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations \jri{this is confusing as the map shows <37 populations.
 we should be consistent about reporting for all pops, or all pops with n>8, etc.
@@
 %--------------------------------------------
-\begin{figure*}[!t]
+\begin{figure*}[]
   \begin{center}
-   \includegraphics[width=250mm]{Fig1Map.pdf}
-    \caption{Map showing the frequency of the Hopscotch allele in populations of \emph{Zea mays} ssp.
+   \includegraphics[width=0.5\textwidth]{Fig1Map.pdf}
+    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{Zea mays} ssp.
 \emph{parviglumis} where we sampled more than 6 individuals.
@@
 
-Using our Hopscotch genotyping data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+Using our \emph{Hopscotch} genotyping data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
 F\textsubscript{CT} is 0 within our dataset, and we found similar levels of F\textsubscript{ST} among populations within each subspecies (0.22) \jri{is this an average?
@@
   \centering
-  \caption{Pairwise F\textsubscript{CT} values from sequence and Hopscotch genotyping data}
+  \caption{Pairwise F\textsubscript{CT} values from sequence and \emph{Hopscotch} genotyping data}
     \begin{tabular}{lccc}
  %   \toprule
-    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Seq. Region 2} & \textbf{Seq.  Region 1} & \textbf{Hopscotch} \\
+    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Seq. Region 2} & \textbf{Seq.  Region 1} & \textbf{\emph{Hopscotch}} \\
  %   \midrule
@@
 \end{table}
-Although we found large variation in Hopscotch allele frequency among our populations, BayEnv analysis did not indicate a correlation between the Hopscotch insertion and environmental variables (all Bayes Factors $<$ 1; \ref{TableS3BayEnv}).
+Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; \ref{TableS3BayEnv}).
 
@@
 fix caption in table.
-also add Hopscotch allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats } \lev{the frequencies are in supplemental table 1, do you want them put in this table too, or just me to refer people to supp table 1?} \jri{yeah refer to table is fine}
+also add \emph{Hopscotch} allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats } \lev{the frequencies are in supplemental table 1, do you want them put in this table too, or just me to refer people to supp table 1?} \jri{yeah refer to table is fine}
 % Table 2 Diversity Statistics
@@
 Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this population (\ref{Table2Diversity}).
-Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the Hopscotch, while they are high for other population pairs (\ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citep{Chia2012}) do not reveal any clear clustering pattern with respect to population or Hopscotch genotype (\ref{FigureS3NJ}); individuals within our sample that have the Hopscotch insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the Hopscotch insertion.
+Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (\ref{Table1Fst}).
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citep{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (\ref{FigureS3NJ}); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the \emph{Hopscotch} insertion.
 
@@
 
-The teosinte populations \jri{which?} with the highest frequency of the Hopscotch insertion in this study were sympatric with cultivated maize.
-Our initial hypothesis was that the high frequency of the Hopscotch element in these populations could be attributed to introgression from maize into teosinte.
+The teosinte populations \jri{which?} with the highest frequency of the \emph{Hopscotch} insertion in this study were sympatric with cultivated maize.
+Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
 To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one, and specifically in the \emph{tb1} region.
-If the Hopscotch is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the Hopscotch and SNPs surrounding the \emph{tb1} region in our resequenced populations (\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the Hopscotch genotype (average $r^{2}=0.082$).
+If the \emph{Hopscotch} is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
+We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the \emph{Hopscotch} genotype (average $r^{2}=0.082$).
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
@@
     \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
-The yellow rectangle indicates the location of the Hopscotch insertion and the green represents the \emph{tb1} ORF.
+The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
 A) Ejutla A; B) Ejutla B; C) La Mesa; D).
@@
 
-The lack of clustering of Hopscotch genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the Hopscotch insertion in these populations of \emph{parviglumis} is the result of recent introgression.
+The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
 However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}.
 The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations.
-If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
+If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
 In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} \lev{Yes really 0.017 assignment to maize in the tb1 region, and avg assignment across chr1 is 0.2} \lev{ I also have a table of assignment values for SLO individuals based on genotype.
@@
 
-To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the Hopscotch insertion from our initial genotyping.
-We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from within the San Lorenzo population, and genotyped plants for the Hopscotch insertion.
-We found the Hopscotch segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and Hopscotch genotype we find a weak positive correlation between presence of the Hopscotch and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
-Additionally we find no significant correlation between tiller number and Hopscotch genotype, or culm diameter and Hopscotch genotype in phenotyping 1.
+To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion from our initial genotyping.
+We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
+We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
+After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find a weak positive correlation between presence of the \emph{Hopscotch} and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
+Additionally we find no significant correlation between tiller number and \emph{Hopscotch} genotype, or culm diameter and \emph{Hopscotch} genotype in phenotyping 1.
 \jri{shouldn't we expect a negative correlation betweeh Hop and TI on day 40?
@@
     \caption{Box-plot showing tillering index in our greenhouse population from day 20-60.
-White indicates individuals homozygous for the Hopscotch, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No Hopscotch) allele.
+White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.
 Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.}
@@
 For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
-We found the Hopscotch allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of Hopscotch homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
-We found similar patterns, with a weak positive correlation between tillering index and Hopscotch genotype at day 40 (p<0.0611), with no significant correlation on any day.
-Similarly, relationships between Hopscotch genotype and tiller number, and hopscotch genotype and culm diameter are not significant.
+We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p<0.0611), with no significant correlation on any day.
+Similarly, relationships between \emph{Hopscotch} genotype and tiller number, and \emph{Hopscotch} genotype and culm diameter are not significant.
 
@@
 The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \citep{Clark2006, Studer2011}.
-\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the Hopscotch insertion is segregating in teosinte \citep{Studer2011}.
-
-Given that the Hopscotch insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
-However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the Hopscotch in teosinte.
-While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the Hopscotch allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Many of our populations with high frequency of the Hopscotch allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}.
+\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte \citep{Studer2011}.
+
+Given that the \emph{Hopscotch} insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
+However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
+While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
+Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}.
 While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}), our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
@@
 Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{Table4Q}).
-Together, these data point to an explanation other than recent introgression for the high observed frequency of Hopscotch in some of  our \emph{parviglumis} populations.
+Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in some of  our \emph{parviglumis} populations.
 
@@
 
-Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the Hopscotch in these populations.
-If the Hopscotch allele was introgressed in the distant past, they could have been sufficient recombination to break up any initial LD, leading to observations similar to those obtained here.
+Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the \emph{Hopscotch} in these populations.
+If the \emph{Hopscotch} allele was introgressed in the distant past, they could have been sufficient recombination to break up any initial LD, leading to observations similar to those obtained here.
 We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
@@
 
-Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.
+Other explanation for differential frequencies of the \emph{Hopscotch} among teosinte populations include both drift and natural selection.
 Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
@@
 do we?}.
-This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.
+This bottleneck, however, does not explain differences in \emph{Hopscotch} allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.
 \jri{here we need a few sentences on selection.
@@
 it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
-%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..
-
-%\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the hopscotch allele.
+%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the \emph{Hopscotch} insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..
+
+%\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the \emph{Hopscotch} allele.
 Why?
 Only ancient introgression in Jalisco and not in the Balsas?
-If hopscotch came from standing variation, is this more consistent with a Jalisco domestication?
+If \emph{Hopscotch} came from standing variation, is this more consistent with a Jalisco domestication?
 I know, heresy!!
-LV, can you look at the \emph{parviglumis} populations with high hopscotch and see if they are all in the Jalisco cluster of Fukunaga?
+LV, can you look at the \emph{parviglumis} populations with high \emph{Hopscotch} and see if they are all in the Jalisco cluster of Fukunaga?
 Might be a nice observation we could make here...could even go so far as saying the results from Matsuoka deserve another look.} \lev{Matt, yes they are in the Jalisco cluster, I added in a sentence mentioning it, but am not sure if that's the direction you were wanting me to take with it or not}
 
-The phenotypic effects of the Hopscotch insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
-Our study is the first to explicitly examine the phenotypic effects of the Hopscotch insertion in individuals sampled from a natural population of teosinte.
+The phenotypic effects of the \emph{Hopscotch} insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
+Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion in individuals sampled from a natural population of teosinte.
 \jri{isn't this what weber did??
-for 70+ populations!?} However, we found no significant effect of the Hopscotch on tillering index or tiller number in our phenotyping experiments, and the effect of the Hopscotch insertion in teosinte is discordant with that of maize.
-The lack of correlation between Hopscotch genotype and tillering index or tiller number is surprising given its effects in maize.
-It is certainly possible that even though previous data demonstrate an effect of the Hopscotch on tillering in maize \citep{Studer2011}, that the effect of the Hopscotch in teosinte is more complicated and may be more difficult to observe.
-Moreover, \emph{tb1} is a single gene in a complex pathway that affects branching and tillering traits, and perhaps in combination with alleles at other loci the phenotypic effects of the Hopscotch on tillering may not be consistent.
+for 70+ populations!?} However, we found no significant effect of the \emph{Hopscotch} on tillering index or tiller number in our phenotyping experiments, and the effect of the \emph{Hopscotch} insertion in teosinte is discordant with that of maize.
+The lack of correlation between \emph{Hopscotch} genotype and tillering index or tiller number is surprising given its effects in maize.
+It is certainly possible that even though previous data demonstrate an effect of the \emph{Hopscotch} on tillering in maize \citep{Studer2011}, that the effect of the \emph{Hopscotch} in teosinte is more complicated and may be more difficult to observe.
+Moreover, \emph{tb1} is a single gene in a complex pathway that affects branching and tillering traits, and perhaps in combination with alleles at other loci the phenotypic effects of the \emph{Hopscotch} on tillering may not be consistent.
 \jri{thsi section still needs work.
@@
 \emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number) \citep{Li2012}.
-It seems plausible that variation in some of these other loci could have affected tillering in our greenhouse population, and contributed to the lack of correlation we see between Hopscotch genotype and tillering.
-
-In summary, our findings demonstrate that the Hopscotch allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
-Analysis of linkage using SNPs from across chromosome 1 does not suggest that the Hopscotch allele is present in these populations due to recent introgression; however, it seems unlikely that  it would have drifted to high frequency in multiple populations and there may be another explanation for the high frequency we observe in some of our populations.
-The Hopscotch does not appear to have a strong effect reducing tillering in teosinte as it does in maize, and  other loci involved in branching architecture  may play also play roles in the regulation of tillering in teosinte.
-Finally, although we see no clear evidence of recent strong selection, the high frequency of the Hopscotch insertion in a number of populations continues to suggest to us that it plays an ecological role in teosinte.
+It seems plausible that variation in some of these other loci could have affected tillering in our greenhouse population, and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
+
+In summary, our findings demonstrate that the \emph{Hopscotch} allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
+Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that  it would have drifted to high frequency in multiple populations and there may be another explanation for the high frequency we observe in some of our populations.
+The \emph{Hopscotch} does not appear to have a strong effect reducing tillering in teosinte as it does in maize, and  other loci involved in branching architecture  may play also play roles in the regulation of tillering in teosinte.
+Finally, although we see no clear evidence of recent strong selection, the high frequency of the \emph{Hopscotch} insertion in a number of populations continues to suggest to us that it plays an ecological role in teosinte.
 In the future, additional experiments will be needed to examine expression levels of \emph{tb1} and additional loci involved in branching architecture (e.g.
@@
 
-%\mbh{I'd lean toward saying that 1) hopscotch does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, hopscotch may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the hopscotch.}
+%\mbh{I'd lean toward saying that 1) \emph{Hopscotch} does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, \emph{Hopscotch} may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the \emph{Hopscotch}.}
 
@@
    \includegraphics[width=150mm]{FigS1LocusCartoon.pdf}
-   % \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red).
-Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.}
+   % \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
 \label{FigS1Locus}
@@
 UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}.
-Individuals with genotype data are colored: Homozygous for the teosinte (no Hopscotch) allele (red), homozygous for the maize (Hopscotch) allele (blue), heterozygotes (purple).
-TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the Hopscotch insertion.
-Black indicates individuals not genotyped for the Hopscotch insertion.}
+Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).
+TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion.
+Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
 \label{FigS3NJ}
@@
 %    \toprule
-    Accession & USDA Accession ID & Locality & Number alleles sampled & Hopscotch Frequency & No Hopscotch Frequency \\
+    Accession & USDA Accession ID & Locality & Number alleles sampled & \emph{Hopscotch} Frequency & No \emph{Hopscotch} Frequency \\
  %   \midrule
@@
  %   \toprule
-    Accession & Number of alleles sampled & Hopscotch Frequency \\
+    Accession & Number of alleles sampled & \emph{Hopscotch} Frequency \\
  %   \midrule

### 2014-05-16 manuscript.tex
edits to abstract

@@
 Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the \emph{Hopscotch} element, led us to investigate the prevalence and potential role of \emph{tb1} in teosinte.
-Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of both subspecies \emph{Zea mays} ssp.
-\emph{parviglumis} and subspecies \emph{Zea mays} ssp.
+Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
+\emph{parviglumis} and \emph{Zea mays} ssp.
 \emph{mexicana}.
 Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
-Finally, two greenhouse experiments fail to find an important role for \emph{tb1} in controlling tillering in natural populations of \emph{parviglumis}.
-Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte is not as straightforward as is in domesticated maize, and that other loci may play a role in observed variation in tillering in teosinte.
+Finally, two greenhouse experiments with \emph{Zea mays} ssp.
+\emph{parviglumis} do not suggest an important role for \emph{tb1} in controlling tillering in natural populations of this subspecies.
+Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte is not as straightforward as is in domesticated maize, and that other loci may play a role in observed variation in this trait.
 

### 2014-05-16 manuscript.tex
edits to introduction

@@
 \emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp.
-\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity and of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
-
-Many morphological changes are associated with the domestication of maize, and understanding the genetic basis of these changes has been a focus of maize research for a number of years\citep{Doebley2004}.
+\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
+
+Many morphological changes are associated with maize domestication, and understanding the genetic basis of these changes has been a focus of maize research for a number of years \citep{Doebley2004}.
 One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches.
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in branching differences between maize and teosinte \citep{Doebley1990, DoebleyStec1991, LukensDoebley1999}.
-\jri{an odd assortment of cites.
-why not the 1997 nature paper?} Further work showed that the insertion of a 4.9 kb retrotransposon (\emph{Hopscotch}) in the upstream control region of \emph{tb1} led to increased expression of this gene, causing the reduction in branching observed in domesticated maize \cite{Studer2011}.
-The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in teosinte \citep{Studer2011}.
+The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in domestication branching differences achieved through higher expression of the maize allele than the teosinte allele \citep{DoebleyStecGustus1995, Doebley1997}.
+Further work showed that the insertion of a 4.9 kb retrotransposon (\emph{Hopscotch}) in the upstream control region of \emph{tb1} caused the increased expression of this gene reduction in branching observed in domesticated maize \cite{Studer2011}.
+The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in natural populations of teosinte \citep{Studer2011}.
 Dating of this element has suggested that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-Furthermore, \citep{StuderDoebley2012} investigated the phenotypic effects of 9 teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
-
-Tillering, or more specifically lack of tillers in teosinte, may provide an ecological advantage to plants.
-In species such as teosinte, where plants grow densely packed together, if plants invest more resources into growing taller and fewer resources into tillering, plants may be able to out-compete neighbors for limiting resources such as light.
-\jri{there must be some citations in ecology lit to support this?} In natural populations of teosinte, variation at the \emph(tb1) locus may then play a role in the ecology of teosinte.
+Furthermore, \citet{StuderDoebley2012} investigated the phenotypic effects of nine teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
+
+In high-density species such as teosinte, plants can detect impending competition from their neighbors through detection of the ratio of red to far-red light.
+An increase in far-red relative to red light accompanies shading and triggers physiological and morphological changes such as reduced tillering, increased plant height and early flowering collectively known as the shade avoidance syndrome \citep{KebromBrutnell2007}.
+The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.
 In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.

### 2014-05-16 manuscript.tex
formatting for AJB

@@
 \usepackage{hyperref}
+\usepackage{lineno}
 \usepackage{color}
@@
 %----------------------------------------
+
 \title{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}
@@
 \begin{document}
+
+\begin{linenumbers}
+	\pagewiselinenumbers
+	\renewcommand\linenumberfont{\normalfont\bfseries\small}
 
@@
 
+\begin{flushleft}
+
+\parindent=20pt
+
 \subsection*{Abstract}
@@
 
-We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{Zea mays} ssp.
-\emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (\ref{TableS1Mays}, \ref{TableS2}).
+We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1, Table~S2).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
-We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \cite{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence (~5 kb amplification product) and absence (~300 bp amplification product) of the \emph{Hopscotch} insertion.
+We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion.
 Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, 5'-{\small CCTCCACCTCTCATGAGATCC-3'} (\ref{FigS1Locus}).
-Homozygotes show a single band for either the \emph{Hopscotch} element (~5 kb) or the absence of the element (~300 bp), while heterozygotes are three-banded, showing a band for both the presence and absence of the \emph{Hopscotch} element as well as a band for the internal primer set (\ref{FigS2Gel}).
-\jri{fig references still need to say "Fig" etc., all the ref function does is set the number automaticall}.
-When only one PCR resolved well, we scored one allele for the individual, which explains the odd number of alleles included in our analyses.
+Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, 5'-{\small CCTCCACCTCTCATGAGATCC-3'} (Figure~S1, Figure~S2) \mbh{Primers in Figure~S1 should be labeled}.
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and XX \mbh{LV, please add the size of the second band}), whereas heterozygotes are three-banded (Figure~S2).
+When only one PCR resolved well, we scored one allele for the individual.
 We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
@@
 For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of the start of the \emph{Hopscotch} and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
-The 5' UTR (containing sequenced region 1) has been shown to have elevated diversity in ssp.
-\emph{parviglumis} with respect to maize, while the the area upstream from the \emph{Hopscotch} (sequenced region 2) has been shown to be critical in determining basal branching and ear architecture \citep{Wang1999, Clark2006}.
-\jri{is this true?
-the area upstream of the \emph{Hopscotch}?
-those papers just found that something upstream was the causal variant, no?} We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5' GGATAATGTGCACCAGGTGT 3'} and {\small 5' GCGTGCTAGAGACACYTGTTGCT 3'}; for the 50 kb upstream region, {\small 5' TGTCCTCGCCGCAACTC 3'} and {5' TGTACGCCCGCCCCTCATCA 3'} (\ref{FigS1Locus}).
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5' GGATAATGTGCACCAGGTGT 3'} and {\small 5' GCGTGCTAGAGACACYTGTTGCT 3'}; for the 50 kb upstream region, {\small 5' TGTCCTCGCCGCAACTC 3'} and {5' TGTACGCCCGCCCCTCATCA 3'} (\ref{FigS1Locus}).
 We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
@@
 
+\end{flushleft}
+
+\end{linenumbers}
+
 \end{document}

### 2014-05-16 manuscript.tex
formatting for AJB cont.

@@
 
+\begin{document}
+
 \title{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}
@@
 \affil[5]{Center for Population Biology and Genome Center, University of California Davis}
-
-\begin{document}
 
@@
 
-\section*{Introduction}
+\begin{centering}
+\section*{INTRODUCTION}
+\end{centering}
 
@@
 
-\section*{Methods}
-
-\subsection*{Sampling and Genotyping}
+\begin{centering}
+\section*{MATERIALS AND METHODS}
+\end{centering}
+
+\mbh{from AJB formatting instructions: add name, city, spelled-out state (if in USA), and country of manufacturers/suppliers after brand names}
+
+\subsection*{Sampling and genotyping}
 
@@
 
-\subsection*{Genotyping Analysis} \jri{please drop scripts for bayenv and STRUCTURE into github} \lev{which scripts are you talking about?
+\subsection*{Genotyping analysis} \jri{please drop scripts for bayenv and STRUCTURE into github} \lev{which scripts are you talking about?
 I used matt's for making structure chunks, and then just command line to run structure, same with BayEnv -- it was Tanjas covariance matrix and then just the basic bayenv commands} \jri{matt's scritp for structure should be up there.
@@
 
-\subsection*{Sequence Analysis}
+\subsection*{Sequence analysis}
 
@@
 
-\subsection*{Introgression Analysis}
+\subsection*{Introgression analysis}
 
@@
 
-\section*{Results}
+\begin{centering}
+\section*{RESULTS}
+\end{centering}
 
@@
 
-\subsubsection*{Phenotyping}
-
 To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion from our initial genotyping.
@@
 
-\section*{Discussion}
+\begin{centering}
+\section*{DISCUSSION}
+\end{centering}
 

### 2014-05-16 manuscript.tex
formatting for AJB cont.

@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 \usepackage{geometry} 
@@
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, 5'-{\small CCTCCACCTCTCATGAGATCC-3'} (Figure~S1, Figure~S2) \mbh{Primers in Figure~S1 should be labeled}.
+Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Figure~S1, Figure~S2) \mbh{Primers in Figure~S1 should be labeled}.
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and XX \mbh{LV, please add the size of the second band}), whereas heterozygotes are three-banded (Figure~S2).
@@
 We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235 bp upstream of the start of the \emph{Hopscotch} and 66,169 bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5' GGATAATGTGCACCAGGTGT 3'} and {\small 5' GCGTGCTAGAGACACYTGTTGCT 3'}; for the 50 kb upstream region, {\small 5' TGTCCTCGCCGCAACTC 3'} and {5' TGTACGCCCGCCCCTCATCA 3'} (\ref{FigS1Locus}).
-We used Taq Polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
+For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} and 66,169bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 50kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Figure~S1).
+We used Taq polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
 Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
-To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 $\mu$ L of amplification product.
+To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 $\mu$L of amplification product.
 This mix was placed on a thermal cycler with the following program: 37$^{\circ}$C for 30 min, 80$^{\circ}$C for 15 min, and a final cool-down step to 4$^{\circ}$C.
@@
 coli} cells, with an extended ligation time of 30 min for a complex target fragment.
-We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96$^{\circ}$C for 5 min; then 35 cycles at 96$^{\circ}$C for 30 s, 53$^{\circ}$C for 30 s, 72$^{\circ}$C for two min; and a final extension at 72$^{\circ}$C for 4 min.
+We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96$^{\circ}$C for 5 min; then 35 cycles at 96$^{\circ}$C for 30 s, 53$^{\circ}$C for 30 s, 72$^{\circ}$C for 2 min; and a final extension at 72$^{\circ}$C for 4 min.
 We visualized amplification products for incorporation of our insert on a 1\% agarose TAE gel.
@@
 
-\subsection*{Genotyping analysis} \jri{please drop scripts for bayenv and STRUCTURE into github} \lev{which scripts are you talking about?
-I used matt's for making structure chunks, and then just command line to run structure, same with BayEnv -- it was Tanjas covariance matrix and then just the basic bayenv commands} \jri{matt's scritp for structure should be up there.
-and even a readme documenting the commandline for bayenv and including the covariance matrix.
-basically whatever someone would need to be able to literally repeat what you did.}
+\subsection*{Genotyping analysis}
 

### 2014-05-16 manuscript.tex
edits to methods

@@
 %Fix figures and tables
-%add usepackage for tables
 %check citations in last paragraph for ter1
@@
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
-BayEnv creates a covariance matrix of relatedness between populations, and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (Bayes Factor)\citep{Coop2010}.
-We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (\ref{TableS3BayEnv}).
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database and \url{www.harvestchoice.org}, and summarized by principle component analysis \citep{Pyhajarvi2013}.
+BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
+We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (Table~S3).
+Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis \citep{Pyhajarvi2013}.
 
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations, and to calculate standard diversity statistics (number of haplotypes; haplotype diversity; Watterson's estimator $\hat\theta_W$; pairwise nucleotide diversity $\hat\theta_\pi$; and Tajima's D).
+For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
 To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
-For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citep{Chia2012}).
+For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citealt{Chia2012}).
 
@@
 We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and the same phased 55K SNP data from \citep{Pyhajarvi2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
-SNPs were anchored in a modified version of the IBM genetic map (\citep{Gerke2013}, \url{http://arxiv.org/abs/1307.7313}).
+SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
 We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
-\jri{i'd like this perl script on github, maybe in this repo or as a gist.
-also structure input file too.
-all the stuff we'd need to redo this.} \lev{See above note..not all of this or the BayEnv was script'ified} \jri{sure, even command line info should be included where possible.
-idea is to maximize reproducibility -- either other people or subsequent students.
-for example, matt has a student who wants to work on tb1 in natural pops, and she might want to try/redo some of these analyses with the same or new data.} \lev{okay so I should just put my command line stuff with good commenting as to what is what in repository?}
-
-\subsection*{Phenotyping of \emph{Zea mays.} ssp. \emph{parviglumis}}
+
+\subsection*{Phenotyping of \emph{parviglumis}}
 
@@
 We treated seeds with fungicide and germinated them in petri dishes with filter paper.
-Following germination, 206 successful germinations were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches.
-Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer.
-\lev{it ended up being a combination of drip and hand watering because they dried out so much and did better when they had water on the leaves as well as in the soil}
-
-To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
-In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
-We chose between 10-13 seeds from each of 23 sampling sites.
-We treated seeds with fungicide and germinated them in petri dishes with filter paper.
-Successful germinations (206 individuals) were then planted into one gallon size pots with potting soil and randomly spaced one foot apart on greenhouse benches.
-Plants were watered three times a day.
-
-Starting on day 15, we measured tillering index, the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
-Tillering index has been shown to be the most effective way to observe the phenotypic effects of the \emph{Hopscotch} insertion on plant architecture in maize \citep{Clark2006}.
+Following germination, 206 successful germinations were then planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
+Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer.
+
+Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
 Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
 On day 65 we measured culm diameter between the third and fourth nodes of each plant.
-Culm diameter is not believed to be correlated with tillering index, or variation at \emph{tb1} (e.g.
-\emph{Hopscotch} genotype).
-Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (http://www.ars.usda.gov).
+Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1}.
+Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (\url{http://www.ars.usda.gov}).
 We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
-Based on these initial data, we conducted a post hoc power analysis using data from day 40 of phenotyping 1, indicating that a minimum of 71 individuals in each genotype class are needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
-\jri{do you still have these posthoc calculations?} \lev{I believe I do, otherwise they are likely in a lab meeting slide on Dropbox -- do you want them in here?} \jri{would be good to include in a github, again so we could go back and reassess how we do things in case, for example, we decide to give the greenhouse experiment a 3rd try (yes, i'm a masochist)} \lev{I would do it a 3rd time, I'm convinced something went wrong....third time with expression analyses :) }
-
-We performed a second phenotyping experiment (phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the \emph{Hopscotch} insertion.
-Seeds were germinated and planted on day 7 post fruit-case removal into 2 gallon pots.
+Based on these initial data, we conducted a post hoc power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class are needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
+
+We performed a second phenotyping experiment (Phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the \emph{Hopscotch} insertion.
+Seeds were germinated and planted on day 7 post fruit-case removal into two gallon pots.
 Plants were watered twice daily, alternating between fertilized and non-fertilized water.
-We began phenotyping successful germinations (302) for tillering index on day 15 post fruit case removal, and phenotyped every five days until day 50.
+We began phenotyping successful germinations (302) for tillering index on day 15 post fruit-case removal, and phenotyped every five days until day 50.
 At day 50 we measured culm diameter between the third and fourth nodes.
-We extracted DNA and genotyped plants following the same guidelines as in phenotyping 1.
-
-Resulting tillering index data for each genotype class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
+We extracted DNA and genotyped plants following the same guidelines as in Phenotyping 1.
+
+Resulting tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) \mbh{what is the alpha value here?} implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
 Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
-\jri{please add SAS scripts/code to a gist or something.
-}
 
@@
 After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find a weak positive correlation between presence of the \emph{Hopscotch} and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
-Additionally we find no significant correlation between tiller number and \emph{Hopscotch} genotype, or culm diameter and \emph{Hopscotch} genotype in phenotyping 1.
+Additionally we find no significant correlation between tiller number and \emph{Hopscotch} genotype, or culm diameter and \emph{Hopscotch} genotype in Phenotyping 1.
 \jri{shouldn't we expect a negative correlation betweeh Hop and TI on day 40?

### 2014-05-17 manuscript.tex
edits to results

@@
 
-We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1, Table~S2).
+We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1 and Table~S2).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Figure~S1, Figure~S2) \mbh{Primers in Figure~S1 should be labeled}.
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and XX \mbh{LV, please add the size of the second band}), whereas heterozygotes are three-banded (Figure~S2).
+Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and XX \mbh{LV, please add the size of the second band}), whereas heterozygotes are three-banded (Fig.~S2).
 When only one PCR resolved well, we scored one allele for the individual.
@@
 For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} and 66,169bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 50kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Figure~S1).
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 50kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Fig.~S1).
 We used Taq polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
@@
 Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
-Among the 247 maize landrace accessions genotyped, all but 8 were homozygous for the presence of the insertion (\ref{TableS1Mays}, \ref{TableS2}).
-\jri{please fix table/figure references to say table/figure} Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations \jri{this is confusing as the map shows <37 populations.
-we should be consistent about reporting for all pops, or all pops with n>8, etc.
-or at least be explicit what cutoff we are using for each result}, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (\ref{Fig1Map}) \jri{the map only shows parviglumis.
+Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2).
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (\ref{Fig1Map}) \jri{the map only shows parviglumis.
 can we add mexicana?

### 2014-05-17 manuscript.tex
edits to results, tables, figures

@@
 \usepackage{hyperref}
+\usepackage[leftcaption]{sidecap}
+\sidecaptionvpos{figure}{c} %position side caption
 \usepackage{lineno}
@@
 \usepackage{authblk}
+\usepackage[table]{xcolor} %for table line color alteration
 \usepackage{subscript}
@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico in both subspecies (\ref{Fig1Map}) \jri{the map only shows parviglumis.
-can we add mexicana?
-the text makes it seem as if both should be on the map}.
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.\ref{Fig1Map}).
 We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
-Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p<0.05).
-\jri{in what direction?
-too many hets?
-what's the F?
-} \lev{they weren't all in the same direction, do you still want me to list out?} \jri{no need to list all, but if there were sig.
-deviations in multiple directions in different pops, that is worth saying.
-maybe could list them all out in supp.
-table?}
-
-\jri{please fix whitespace and black border on figure}
+Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05).
+\mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
 %Figure 1 Map 
 %--------------------------------------------
-\begin{figure*}[]
-  \begin{center}
-   \includegraphics[width=0.5\textwidth]{Fig1Map.pdf}
-    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{Zea mays} ssp.
-\emph{parviglumis} where we sampled more than 6 individuals.
+\begin{SCfigure}[][t!]
+  \centering \label{Fig1Map}
+   \includegraphics[width=0.7\textwidth]{Fig1Map.pdf}
+    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
 Size of circles reflects number of alleles sampled.
 }
-\label{Fig1Map}
-  \end{center}
-\end{figure*}
+\end{SCfigure}
 %--------------------------------------------
 
-Using our \emph{Hopscotch} genotyping data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-F\textsubscript{CT} is 0 within our dataset, and we found similar levels of F\textsubscript{ST} among populations within each subspecies (0.22) \jri{is this an average?
-this comes from libsequence?} and among all populations (0.23), to those reported in genome-wide estimates from previous studies \cite{Pyhajarvi2013} (\ref{Table1Fst}).
+Using our \emph{Hopscotch} genotyping and \emph{tb1} resequencing data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+We found that F\textsubscript{CT} = 0 within our dataset and levels of F\textsubscript{ST} among populations within each subspecies (0.22) \jri{is this an average?
+this comes from libsequence?} and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 
@@
 \begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
   \centering
-  \caption{Pairwise F\textsubscript{CT} values from sequence and \emph{Hopscotch} genotyping data}
-    \begin{tabular}{lccc}
- %   \toprule
-    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Seq. Region 2} & \textbf{Seq.  Region 1} & \textbf{\emph{Hopscotch}} \\
- %   \midrule
+  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
+
+    \begin{tabular}{lccc}\\\toprule
+    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Seq. Region 2} & \textbf{Seq.  Region 1} & \textbf{\emph{Hopscotch}} \\\midrule
     EjuA \& EJuB & 0     & 0     & 0 \\
@@
     EjuB \& SLO & 0.29  & 0.512 & 0.28 \\
-    MSA \& SLO & 0     & 0.007 & 0.016 \\
-%    \bottomrule
+    MSA \& SLO & 0     & 0.007 & 0.016 \\\bottomrule
     \end{tabular}

### 2014-05-17 manuscript.tex
edits to results, tables, figures cont.

@@
 We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (sequenced region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} and 66,169bp upstream from the start of the \emph{tb1} ORF (sequenced region 2).
+For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (Region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} (66,169bp upstream from the start of the \emph{tb1} ORF; Region 2).
 We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 50kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Fig.~S1).
@@
 
-For population genetic analyses of sequenced region 1 and sequenced region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
+For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
 To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
@@
 
-Using our \emph{Hopscotch} genotyping and \emph{tb1} resequencing data, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
 We found that F\textsubscript{CT} = 0 within our dataset and levels of F\textsubscript{ST} among populations within each subspecies (0.22) \jri{is this an average?
 this comes from libsequence?} and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
 
@@
     \begin{tabular}{lccc}\\\toprule
-    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Seq. Region 2} & \textbf{Seq.  Region 1} & \textbf{\emph{Hopscotch}} \\\midrule
-    EjuA \& EJuB & 0     & 0     & 0 \\
-    EjuA \& MSA & 0.328 & 0.326 & 0.186 \\
-    EjuA \& SLO & 0.258 & 0.416 & 0.28 \\
-    EjuB \& MSA & 0.365 & 0.397 & 0.188 \\
-    EjuB \& SLO & 0.29  & 0.512 & 0.28 \\
-    MSA \& SLO & 0     & 0.007 & 0.016 \\\bottomrule
+    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
+    EjuA \& EjuB & 0     & 0     & 0 \\
+    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
+    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
+    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
+    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
+    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
     \end{tabular}
@@
 \end{table}
-Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; \ref{TableS3BayEnv}).
 
@@
 
-To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the tb1 ORF in four populations.
-After alignment and singleton checking we recovered 40 and 48 segregating sites for the 50kb upstream region and the 5' UTR region, respectively.
-For region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is more positive in San Lorenzo (\ref{Table2Diversity}).
-\jri{can drop theta W from table and text.
-we show pi and D, which is sufficient.
-fix caption in table.
-also add \emph{Hopscotch} allele frequencies to table or list somewhere -- i think it's useful for comparison of seq stats } \lev{the frequencies are in supplemental table 1, do you want them put in this table too, or just me to refer people to supp table 1?} \jri{yeah refer to table is fine}
+To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
+After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
+For Region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}).
+\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
+Tajima's D is positive in all populations except San Lorenzo, \mbh{is the table wrong?
+MSA is the only negative value in the table} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Figure~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the \emph{Hopscotch} insertion.
+
 % Table 2 Diversity Statistics
 \begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
   \centering
-  \caption{Add caption}
-    \begin{tabular}{lccccc}
-%    \toprule
-    Population & \# Haplotypes & Hap. Diversity & $\hat\theta_W$    & $\hat\theta_\pi$    & Tajima's D \\
-%    \midrule
-    \multicolumn{6}{c}{\textit{Seq. region 2 (50kb upstream)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.89394 & 0.01548 & 0.01763 & 0.6231 \\
-    \multicolumn{1}{c}{EJUB} & 8     & 0.89394 & 0.01493 & 0.01591 & 0.29504 \\
-    \multicolumn{1}{c}{MSA} & 3     & 0.68182 & 0.01111 & 0.01055 & -0.22212 \\
-    \multicolumn{1}{c}{SLO} & 4     & 0.74242 & 0.01167 & 0.01413 & 0.93185 \\
-    \multicolumn{6}{c}{\textit{Seq.  region 1(5? UTR)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.85897 & 0.00874 & 0.00527 & -1.64955 \\
-    \multicolumn{1}{c}{EJUB} & 5     & 0.70909 & 0.00663 & 0.00378 & -1.83123 \\
-    \multicolumn{1}{c}{MSA} & 6     & 0.68182 & 0.00646 & 0.00373 & -1.75506 \\
-    \multicolumn{1}{c}{SLO} & 3     & 0.31818 & 0.00176 & 0.00137 & -0.72873 \\
- %   \bottomrule
+  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
+    \begin{tabular}{lcccc}\\\toprule
+    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
+    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.85897 & 0.00527 & -1.64955 \\
+    \multicolumn{1}{c}{EJUB} & 5     & 0.70909 & 0.00378 & -1.83123 \\
+    \multicolumn{1}{c}{MSA} & 6     & 0.68182 & 0.00373 & -1.75506 \\
+    \multicolumn{1}{c}{SLO} & 3     & 0.31818 & 0.00137 & -0.72873 \\
+    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.89394 & 0.01763 & 0.6231 \\
+    \multicolumn{1}{c}{EJUB} & 8     & 0.89394 & 0.01591 & 0.29504 \\
+    \multicolumn{1}{c}{MSA} & 3     & 0.68182 & 0.01055 & -0.22212 \\
+    \multicolumn{1}{c}{SLO} & 4     & 0.74242 & 0.01413 & 0.93185 \\\bottomrule
     \end{tabular}
@@
 
-For region 2, haplotype diversity, $\theta_W$, and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (\ref{Table2Diversity}).
-Tajima's D is positive in all populations except San Lorenzo, indicating an excess of low frequency variants in this population (\ref{Table2Diversity}).
-Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (\ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citep{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (\ref{FigureS3NJ}); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the \emph{Hopscotch} insertion.
-
 \subsection*{Evidence of introgression}
 
-The teosinte populations \jri{which?} with the highest frequency of the \emph{Hopscotch} insertion in this study were sympatric with cultivated maize.
+Teosinte populations with the highest frequency of the \emph{Hopscotch} insertion in this study were \emph{parviglumis} sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
@@
 If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} \lev{Yes really 0.017 assignment to maize in the tb1 region, and avg assignment across chr1 is 0.2} \lev{ I also have a table of assignment values for SLO individuals based on genotype.
-Though we had decided this wasn't super informative because sample size was low} and much below the chromosome-wide average (0.20; \ref{Fig3Structure}).
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Fig.\ref{Fig3Structure}).
 
@@
 %-------------------------------------------------------------------
-\jri{please put figures in the text rather than at the end.} \lev{I can't figure out why the figures are going at the end. I did them following the format in Sofiane's 282 paper}
 

### 2014-05-17 manuscript.tex
edits to results, tables, figures cont.

@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.\ref{Fig1Map}).
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\ref{Fig1Map}).
 We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
@@
 
-    \begin{tabular}{lccc}\\\toprule
-    \multicolumn{1}{c}{\textbf{Comparison}} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
+    \begin{tabular}{cccc}\\\toprule
+    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
     EjuA \& EjuB & 0     & 0     & 0 \\
@@
   \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
-    \begin{tabular}{lcccc}\\\toprule
+    \begin{tabular}{ccccc}\\\toprule
     \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
     \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.85897 & 0.00527 & -1.64955 \\
-    \multicolumn{1}{c}{EJUB} & 5     & 0.70909 & 0.00378 & -1.83123 \\
-    \multicolumn{1}{c}{MSA} & 6     & 0.68182 & 0.00373 & -1.75506 \\
-    \multicolumn{1}{c}{SLO} & 3     & 0.31818 & 0.00137 & -0.72873 \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
+    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
+    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
+    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
     \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.89394 & 0.01763 & 0.6231 \\
-    \multicolumn{1}{c}{EJUB} & 8     & 0.89394 & 0.01591 & 0.29504 \\
-    \multicolumn{1}{c}{MSA} & 3     & 0.68182 & 0.01055 & -0.22212 \\
-    \multicolumn{1}{c}{SLO} & 4     & 0.74242 & 0.01413 & 0.93185 \\\bottomrule
+    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
+    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
+    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
+    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
     \end{tabular}
@@
 
-Teosinte populations with the highest frequency of the \emph{Hopscotch} insertion in this study were \emph{parviglumis} sympatric with cultivated maize.
+The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
-To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one, and specifically in the \emph{tb1} region.
+To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one and specifically in the \emph{tb1} region.
 If the \emph{Hopscotch} is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the \emph{Hopscotch} genotype (average $r^{2}=0.082$).
+We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.
+\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the \emph{Hopscotch} genotype (average $r^{2}=0.082$).
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
-\jri{table is too wide, need to round numbers, and column headers are messed up.}
+\mbh{LV, please go through and make sure the data entered into all the tables is correct.
+In Table3, both sequenced regions were labeled as "Region 1".
+I changed the second to Region 2 but don't know if the data in this column are really from Region 2}
 
@@
 \begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
   \centering
   \caption{$r^{2}$ values between SNPs in the \emph{tb1} region (positions 264,596,664-265,891,456 on chromosome 1 of the maize AGPv2 genome) and the rest of chromosome 1, within the 5' UTR (Sequenced region 1), and within the 66,169 bp upstream region (Sequenced region 2).}
-    \begin{tabular}{lccrc}
- %   \toprule
-    \multirow{2}[2]{*}{Population} & \multirow{2}[2]{*}{Chromosome 1} & \multirow{2}[2]{*}{tb1 region} & \multicolumn{1}{c}{\multirow{2}[2]{*}{Seq. region 1}} & \multirow{2}[2]{*}{Seq. region 1 } \\
- %   \midrule
-          &       &       & \multicolumn{1}{c}{} &  \\
-    Ejutla A & 0.095426101 & 0.050304 & 0.747295 & 0.214933 \\
-    Ejutla B & 0.068681837 & 0.051295 & 0.660354 & 0.186395 \\
-    La Mesa & 0.069500533 & 0.053306 & 0.914286 & 0.766234 \\
-    San Lorenzo & 0.100536784 & 0.067251 & 0.912281 & 0.636364 \\
+    \begin{tabular}{ccccc}\\\toprule
+    \textbf{Population} & \textbf{Chr. 1} & \textbf{\emph{tb1} region} & \textbf{Region 1} & \textbf{Region 2} \\\midrule
+    Ejutla A & 0.095 & 0.050 & 0.747 & 0.215 \\
+    Ejutla B & 0.069 & 0.051 & 0.660 & 0.186 \\
+    La Mesa & 0.070 & 0.053 & 0.914 & 0.766 \\
+    San Lorenzo & 0.101 & 0.067 & 0.912 & 0.636 \\
     \bottomrule
@@
 If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Fig.\ref{Fig3Structure}).
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Fig.
+\ref{Fig3Structure}).
 

### 2014-05-17 manuscript.tex
edits to results, tables, figures cont.

@@
 We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.
-\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without the \emph{Hopscotch} genotype (average $r^{2}=0.082$).
+\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} genotype.
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
@@
   \centering
-  \caption{$r^{2}$ values between SNPs in the \emph{tb1} region (positions 264,596,664-265,891,456 on chromosome 1 of the maize AGPv2 genome) and the rest of chromosome 1, within the 5' UTR (Sequenced region 1), and within the 66,169 bp upstream region (Sequenced region 2).}
+  \caption{$r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
     \begin{tabular}{ccccc}\\\toprule
@@
   \begin{center}
-   \includegraphics[width=150mm]{Fig2LDPlots.pdf}
+   \includegraphics[width=140mm]{Fig2LDPlots.pdf}
     \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
@@
 
-To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion from our initial genotyping.
-We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
+To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
+We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
 We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find a weak positive correlation between presence of the \emph{Hopscotch} and tillering index on day 40 (p=0.0848), but no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
+After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find a weak positive correlation between presence of the \emph{Hopscotch} and tillering index on day 40 (p=0.0848), a result indicating the \emph{Hopscotch} may actually increase tillering in \emph{parviglumis} in contrast to its phenotypic effect in maize.
+We find no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
 Additionally we find no significant correlation between tiller number and \emph{Hopscotch} genotype, or culm diameter and \emph{Hopscotch} genotype in Phenotyping 1.
-\jri{shouldn't we expect a negative correlation betweeh Hop and TI on day 40?
-need to have an A and B in the figure and explain one is for pheno1 and one is for pheno2.
-please explain whiskers and dots on figure too.} \lev{sure, I mean, presumably we would expect things with Hop to have a smaller TI} \jri{yup, but we should mention that the expectation is negative/}
 
@@
   \begin{center}
-   \includegraphics[width=150mm]{Fig4Boxplots.pdf}
-    \caption{Box-plot showing tillering index in our greenhouse population from day 20-60.
+   \includegraphics[width=95mm]{Fig4Boxplots.pdf}
+    \caption{Box-plots showing tillering index in greenhouse grow-outs for Phenotyping 1 (A) and Phenotyping 2 (B).
 White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.
-Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.}
+Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
+\jri{please explain whiskers and dots on figure too.}}
 \label{Fig4Boxplots}
@@
 %-------------------------------------------------------------------
-\jri{lots of white space in fig 4 and fig. s1 too.} 
+
 We performed a second grow-out  of teosinte to assess whether lighting conditions or sample size may have affected our ability to detect and effect of \emph{tb1}.

### 2014-05-17 manuscript.tex
edits to results, tables, figures cont.

@@
 If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Fig.
+In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig.
 \ref{Fig3Structure}).
+
+ % Table 4 Assigment to maize and teosinte in tb1 region and chr 1 
+\begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
+    \begin{tabular}{ccccc} \
+          & \multicolumn{2}{c}{\textbf{\emph{tb1} region}} & \multicolumn{2}{c}{\textbf{Chr 1}} \\\toprule
+    \bf{Population} & \bf{Maize} & \bf{Teosinte} & \bf{Maize} & \bf{Teosinte} \\\midrule
+    Ejutla A & 0.022 & 0.978 & 0.203 & 0.797 \\
+    Ejutla B & 0.019 & 0.981 & 0.187 & 0.813 \\
+    La Mesa & 0.012 & 0.988 & 0.193 & 0.807 \\
+    San Lorenzo & 0.016 & 0.984 & 0.205 & 0.795 \\\bottomrule
+    \end{tabular}
+  \label{Table4Q}
+\end{table}
 
@@
 
-We performed a second grow-out  of teosinte to assess whether lighting conditions or sample size may have affected our ability to detect and effect of \emph{tb1}.
+We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.
 For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
 We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
-We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p<0.0611), with no significant correlation on any day.
-Similarly, relationships between \emph{Hopscotch} genotype and tiller number, and \emph{Hopscotch} genotype and culm diameter are not significant.
+We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p$<$0.0611), with no significant correlation on any day.
+Similarly, relationships between \emph{Hopscotch} genotype and tiller number and \emph{Hopscotch} genotype and culm diameter were not significant.
 
@@
 Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, \ref{Table4Q}).
+Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, Table \ref{Table4Q}).
 Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in some of  our \emph{parviglumis} populations.
-
- % Table 4 Assigment to maize and teosinte in tb1 region and chr 1 
-\begin{table}[htbp]
-  \centering
-  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
-    \begin{tabular}{ccccc}
-  %  \toprule
-          & \multicolumn{2}{c}{\emph{tb1} region} & \multicolumn{2}{c}{Chr 1} \\
- %   \midrule
-    Population & Maize assignment & Teosinte assignment & Maize assignment & Teosinte assignment \\
-    Ejutla A & 0.02158681 & 0.9784132 & 0.2026814 & 0.7973186 \\
-    Ejutla B & 0.01888194 & 0.9811181 & 0.1872131 & 0.8127869 \\
-    La Mesa & 0.0118675 & 0.9881333 & 0.8068998 & 0.1931017 \\
-    San Lorenzo & 0.01551389 & 0.9844861 & 0.2048252 & 0.7951748 \\
- %   \bottomrule
-    \end{tabular}
-  \label{Table4Q}
-\end{table}
 

### 2014-05-17 manuscript.tex
edits to discussion

@@
 %TO DO
-%Fix figures and tables
 %check citations in last paragraph for ter1
@@
 %check map details
-%make figs smaller so they don't go to end of document
 %Code for SAS and command line stuff on Git
@@
 Adaptation occurs either due to selection on standing variation or on \emph{de novo} mutations.
-Adaptation as a result of selection on standing variation has been well-described in a number of systems, for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}; variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}; and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
+Adaptation as a result of selection on standing variation has been well-described in a number of systems; for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}, variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}, and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
 Although the role of standing variation with respect to adaptation has been described in many systems, its importance to domestication is not as well studied.
 
-In maize, alleles at important domestication loci (\emph{RAMOSA1}, \citep{SigmonVollbrecht2010}; \emph{barren stalk1}, \citep{Gallavotti2004}; and \emph{grassy tillers1}, \citep{Whipple2011}) have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
+In maize, alleles at important domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
 The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \citep{Clark2006, Studer2011}.
@@
 However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
-While our findings are consistent with a previous study by \citet{Studer2011} in that we identified the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
+While our findings are consistent with \citet{Studer2011} in that we identified the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
 Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}.
-While gene flow from crops into their wild relatives is well-known, (\citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}), our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
+While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
 Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.
@@
 Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (\ref{Fig3Structure}, Table \ref{Table4Q}).
+Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (Fig.
+\ref{Fig3Structure}, Table \ref{Table4Q}).
 Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in some of  our \emph{parviglumis} populations.
@@
 Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the \emph{Hopscotch} in these populations.
-If the \emph{Hopscotch} allele was introgressed in the distant past, they could have been sufficient recombination to break up any initial LD, leading to observations similar to those obtained here.
+If the \emph{Hopscotch} allele was introgressed in the distant past, recombination may have broken up LD, a pattern that would be consistent with our data.
 We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
@@
 \emph{mexicana} and \emph{Zea mays} ssp.
-\emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, vanHeerwaarden2011}.
-
-Other explanation for differential frequencies of the \emph{Hopscotch} among teosinte populations include both drift and natural selection.
+\emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011}.
+
+Additional explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both drift and natural selection.
 Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
@@
 do we?}.
-This bottleneck, however, does not explain differences in \emph{Hopscotch} allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.
+This bottleneck, however, does not explain differences in \emph{Hopscotch} allele frequency among multiple populations in the Jalisco group, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.
 \jri{here we need a few sentences on selection.
@@
 The phenotypic effects of the \emph{Hopscotch} insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
-Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion in individuals sampled from a natural population of teosinte.
+Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion in individuals sampled from natural populations of teosinte.
 \jri{isn't this what weber did??

### 2014-05-19 manuscript.tex
cleaning up .tex file
edits to discussion

@@
 While our findings are consistent with \citet{Studer2011} in that we identified the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus than in the Balsas region where maize was domesticated \citep{Matsuoka2002}.
+Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
+Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from Jalisco include gene flow from maize, genetic drift, and natural selection.
+
 While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
-Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize) \citep{Hufford2013}.
-
+Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize; \citealt{Hufford2013}).
 We find no evidence of recent introgression in our analyses.
@@
 
-Additional explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both drift and natural selection.
+Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both drift and natural selection.
 Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
@@
 %   from expected genotype frequencies under Hardy-Weinberg equilibrium at the \emph{Hopscotch} insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..
-
-%\mbh{many populations in the Jalisco cluster of Fukunaga seem to have high frequency of the \emph{Hopscotch} allele.
-Why?
-Only ancient introgression in Jalisco and not in the Balsas?
-If \emph{Hopscotch} came from standing variation, is this more consistent with a Jalisco domestication?
-I know, heresy!!
-LV, can you look at the \emph{parviglumis} populations with high \emph{Hopscotch} and see if they are all in the Jalisco cluster of Fukunaga?
-Might be a nice observation we could make here...could even go so far as saying the results from Matsuoka deserve another look.} \lev{Matt, yes they are in the Jalisco cluster, I added in a sentence mentioning it, but am not sure if that's the direction you were wanting me to take with it or not}
 

### 2014-05-19 manuscript.tex
edits to discussion

@@
 
-Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both drift and natural selection.
+Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.
+Drift may have played a role in the San Lorenzo \emph{parviglumis} population.
 Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
 \jri{deviations from HWE may be consistent too if we see excess of homozygotes.
-do we?}.
-This bottleneck, however, does not explain differences in \emph{Hopscotch} allele frequency among multiple populations in the Jalisco group, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.
+do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele at \emph{tb1} could have risen to high frequency entirely by chance.
+This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} allele in multiple populations in the Jalisco group.
+Moreover, available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.
+
 \jri{here we need a few sentences on selection.

### 2014-05-19 manuscript.tex
edits to discussion

@@
 This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} allele in multiple populations in the Jalisco group.
-Moreover, available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.
-
-\jri{here we need a few sentences on selection.
-the 5' UTR has much more negative D than the upstream.
-do we know the Hop genotype for sequenced lines?
+Moreover, available information on diversity and population structure among additional Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.
+Values of the Tajima's D statistic in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural population of \emph{parviglumis}.
+Whereas the genome-wide average of Tajima's D in genic regions of \emph{parviglumis} is 0.45 \citep{Hufford2012b}, the statistic is quite negative in the 5' UTR of \emph{tb1} (Table \ref{Table2Diversity}).
+This result is consistent with repeated selective sweeps near \emph{tb1} and an important ecological role for the gene in \emph{parviglumis}.
+
+\jri{do we know the Hop genotype for sequenced lines?
 can we separate the sequences into hop/no hop and look for differences?
 it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
-%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the \emph{Hopscotch} insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..
 

### 2014-05-19 manuscript.tex
edits to discussion

@@
 can we separate the sequences into hop/no hop and look for differences?
-it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
-
-The phenotypic effects of the \emph{Hopscotch} insertion in domesticated maize have been well documented \citep{Clark2006, Studer2011}, and \citet{Weber2007} have described its effects in partially inbred lines of teosinte \citep{Weber2007} \jri{i don't think these were inbred.please doublecheck.}.
-Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion in individuals sampled from natural populations of teosinte.
-\jri{isn't this what weber did??
-for 70+ populations!?} However, we found no significant effect of the \emph{Hopscotch} on tillering index or tiller number in our phenotyping experiments, and the effect of the \emph{Hopscotch} insertion in teosinte is discordant with that of maize.
-The lack of correlation between \emph{Hopscotch} genotype and tillering index or tiller number is surprising given its effects in maize.
-It is certainly possible that even though previous data demonstrate an effect of the \emph{Hopscotch} on tillering in maize \citep{Studer2011}, that the effect of the \emph{Hopscotch} in teosinte is more complicated and may be more difficult to observe.
-Moreover, \emph{tb1} is a single gene in a complex pathway that affects branching and tillering traits, and perhaps in combination with alleles at other loci the phenotypic effects of the \emph{Hopscotch} on tillering may not be consistent.
-\jri{thsi section still needs work.
-i think we have to do more here.
-weber shows an association between SNPs in tb1 and branch length.
-we need to discuss that!}
+it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check} \mbh{I've added a few sentences on selection.
+Do we still want to compare sequences with and without \emph{Hopscotch}?
+I agree its a good idea and could end up being really interesting.
+Perhaps something we could look at after submission and incorporate during revisions?}
+
+Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Clark2006, Studer2011}.
+\citet{Weber2007} have described significant phenotypic associations between genic markers in \emph{tb1} and lateral branch length and female ear length within a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
+Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte.
+We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number in our phenotyping experiments, a result that is discordant with its clear phenotypic effects in maize.
+One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
+In fact, \emph{tb1} is known to be a single gene in a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007}.
 

### 2014-05-20 manuscript.tex
JRI edits to intro, typos in refs

@@
 The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
-Further studies have shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances its expression, causing the reduction in branching observed in domesticated maize.
+Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances its expression, causing much of the reduction in tillering observed in domesticated maize.
 Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the \emph{Hopscotch} element, led us to investigate the prevalence and potential role of \emph{tb1} in teosinte.
@@
 Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors.
-\citep{Whitton1997}.
+But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
 
@@
 One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches.
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in domestication branching differences achieved through higher expression of the maize allele than the teosinte allele \citep{DoebleyStecGustus1995, Doebley1997}.
-Further work showed that the insertion of a 4.9 kb retrotransposon (\emph{Hopscotch}) in the upstream control region of \emph{tb1} caused the increased expression of this gene reduction in branching observed in domesticated maize \cite{Studer2011}.
-The effects of this insertion have been observed in tiller number in maize, but little is known about its role, if any, in natural populations of teosinte \citep{Studer2011}.
-Dating of this element has suggested that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-Furthermore, \citet{StuderDoebley2012} investigated the phenotypic effects of nine teosinte \emph{tb1} alleles in an isogenic maize background and found that the introgressions sort into three distinct phenotypic classes, suggesting that variation at the \emph{tb1} locus may play a functional role in teosinte.
-
-In high-density species such as teosinte, plants can detect impending competition from their neighbors through detection of the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers physiological and morphological changes such as reduced tillering, increased plant height and early flowering collectively known as the shade avoidance syndrome \citep{KebromBrutnell2007}.
+The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in  branching \citep{DoebleyStecGustus1995} and tillering \citep{DoebleyStec1991} differences between maize and teosinte.
+A 4.9 kb retrotransposon (\emph{Hopscotch}) insertion into the upstream control region of \emph{tb1} in maize acts to enhance expression of \emph{tb1}, thus repressing lateral organ growth \citep{Doebley1997, Studer2011}.
+Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
+The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citet{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+
+In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
+An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
 The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.

### 2014-05-20 manuscript.tex
edits to discussion

--- 2014-05-19 manuscript.tex
@@
 
-Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Clark2006, Studer2011}.
-\citet{Weber2007} have described significant phenotypic associations between genic markers in \emph{tb1} and lateral branch length and female ear length within a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
+Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
+\citet{Weber2007} have described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length within a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
 Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte.
-We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number in our phenotyping experiments, a result that is discordant with its clear phenotypic effects in maize.
+We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
 One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
-In fact, \emph{tb1} is known to be a single gene in a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007}.
-
-\jri{MBH todo} Variation at \emph{tb1} has also been shown to contribute to phenotypes other than tillering \citep{Clark2006}, and a recent study by \citet{StuderDoebley2012} examined the possibility of an allelic series at the \emph{tb1} locus in teosinte.
-\citet{StuderDoebley2012} introgressed 9 separate teosinte segments (one from \emph{Zea diploperennis}, and four from both \emph{Zea mays} ssp.
-\emph{mexicana} and \emph{Zea mays} ssp.
-\emph{parviglumis}) spanning the \emph{tb1} locus into an isogenic maize background and investigated their effects on previously associated phenotypes.
-They found that plants with teosinte chromosomal segments had greater tillering than their maize isogenic lines, and that different chromosomal segments of \emph{tb1} confer different amounts of tillering, suggesting that there are multiple genetic factors in this region that affect tillering.
-However, in addition to elucidating variance in tillering among \emph{tb1} teosinte segments, \citet{StuderDoebley2012} found significant variance among W22 control lines, suggesting that there are other genetic factors aside from alleles at the \emph{tb1} locus that affect tillering in maize.
-\citet{DoebleyStec1991, DoebleyStec1993} first attempted to map QTL controlling many of the phenotypic differences between domesticated maize and teosinte, and demonstrated the existence of numerous QTL that contribute to the differences in branching architecture between the two.
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} have been shown to affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
-\emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number) \citep{Li2012}.
-It seems plausible that variation in some of these other loci could have affected tillering in our greenhouse population, and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
+In fact, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.
+A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.
+\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated phenotypic effects.
+Phenotypes were shown to cluster by taxon, indicating \emph{tb1} potentially played a role in the morphological diversification of \emph{Zea}.
+Additional analysis suggested tillering index was controlled both by \emph{tb1} and loci elsewhere in the genome.
+Clues to the identity of these loci may be found in QTL studies that identified loci controlling branching architecture \citep{DoebleyStec1991, DoebleyStec1993}.
+Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
+\emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}).
+Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
 
 In summary, our findings demonstrate that the \emph{Hopscotch} allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
-Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that  it would have drifted to high frequency in multiple populations and there may be another explanation for the high frequency we observe in some of our populations.
-The \emph{Hopscotch} does not appear to have a strong effect reducing tillering in teosinte as it does in maize, and  other loci involved in branching architecture  may play also play roles in the regulation of tillering in teosinte.
-Finally, although we see no clear evidence of recent strong selection, the high frequency of the \emph{Hopscotch} insertion in a number of populations continues to suggest to us that it plays an ecological role in teosinte.
+Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
+The \emph{Hopscotch} does not appear to reduce tillering in \emph{parviglumis} as it does in maize.
+Other loci involved in branching architecture may regulate tillering in teosinte.
+Finally, we find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus plays an ecological role in teosinte.
 In the future, additional experiments will be needed to examine expression levels of \emph{tb1} and additional loci involved in branching architecture (e.g.

### 2014-05-20 manuscript.tex
methods JRI

--- 2014-05-20 manuscript.tex
@@
 %Code for SAS and command line stuff on Git
-%Phenotyping and genotyping data on Git (it's on Figshare?)
+%Phenotyping and genotyping data on Git (it's on Figshare?) <- http://dx.doi.org/10.6084/m9.figshare.776926
 
@@
 We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (Table~S3).
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis \citep{Pyhajarvi2013}.
+Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
 
@@
 
-For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the analysis package of Libsequence \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and to examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
-For creation of trees we also included homologous sequence data from teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize and landraces (data from Maize HapMapV2, \citealt{Chia2012}).
+For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
+To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
+For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.
 
@@
 Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (\url{http://www.ars.usda.gov}).
-We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
-Based on these initial data, we conducted a post hoc power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class are needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
+\jri{what is this url?} We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
+Based on these initial data, we conducted a post hoc power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
 

### 2014-05-20 manuscript.tex
JRI starting on results

@@
 \ref{Fig1Map}).
-We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
+%We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
 Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05).
 \mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
+
 %Figure 1 Map 
@@
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-We found that F\textsubscript{CT} = 0 within our dataset and levels of F\textsubscript{ST} among populations within each subspecies (0.22) \jri{is this an average?
-this comes from libsequence?} and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
@@
 MSA is the only negative value in the table} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
-Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Figure~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or the lines of domesticated maize that have the \emph{Hopscotch} insertion.
+Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch} \jri{table 1 shows 0.016 for hopscotch, not 0.
+which is right?}, while they are high for other population pairs (Table \ref{Table1Fst}).
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Figure~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
 
@@
 
-\subsection*{Evidence of introgression}
+\subsection*{Evidence of introgression} %JRI HERE
 
@@
 A) Ejutla A; B) Ejutla B; C) La Mesa; D).
-San Lorenzo}
+San Lorenzo}  \jri{this needs description of what upper and lower triangle are}
 \label{Fig2LD}

### 2014-05-20 manuscript.tex
Edits to abstract

@@
 The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
-Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances its expression, causing much of the reduction in tillering observed in domesticated maize.
-Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the \emph{Hopscotch} element, led us to investigate the prevalence and potential role of \emph{tb1} in teosinte.
+Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize.
+Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the \emph{Hopscotch} element, led us to investigate its prevalence and potential role in teosinte.
 Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
@@
 \emph{mexicana}.
-Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte, and we find no evidence of environmental correlations that might suggest recent selection.
+Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte.
+Population genetic signatures, however, are consistent with selection on this locus and suggest the \emph{Hopscotch} insertion at \emph{tb1} may play an important ecological role in teosinte.
 Finally, two greenhouse experiments with \emph{Zea mays} ssp.
-\emph{parviglumis} do not suggest an important role for \emph{tb1} in controlling tillering in natural populations of this subspecies.
-Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte is not as straightforward as is in domesticated maize, and that other loci may play a role in observed variation in this trait.
+\emph{parviglumis} do not suggest \emph{tb1} controls tillering in natural populations of this subspecies.
+Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte differs from domesticated maize, and that other loci may play a role in observed variation in this trait.
 

### 2014-05-20 manuscript.tex
Edits to abstract

@@
 Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize.
-Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of the age of insertion of the \emph{Hopscotch} element, led us to investigate its prevalence and potential role in teosinte.
+Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that pre-dates domestication, led us to investigate its prevalence and potential role in teosinte.
 Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
@@
 Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte.
-Population genetic signatures, however, are consistent with selection on this locus and suggest the \emph{Hopscotch} insertion at \emph{tb1} may play an important ecological role in teosinte.
+Population genetic signatures, however, are consistent with selection on this locus and suggest the \emph{Hopscotch} insertion at \emph{tb1} may play an ecological role in teosinte.
 Finally, two greenhouse experiments with \emph{Zea mays} ssp.

### 2014-05-20 manuscript.tex
Edits to abstract/introduction

@@
 Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize.
-Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that pre-dates domestication, led us to investigate its prevalence and potential role in teosinte.
+Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication, led us to investigate its prevalence and potential role in teosinte.
 Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
@@
 Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citet{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 

### 2014-05-20 manuscript.tex
Edits to methods, vendor information added

@@
 
-\mbh{from AJB formatting instructions: add name, city, spelled-out state (if in USA), and country of manufacturers/suppliers after brand names}
-
 \subsection*{Sampling and genotyping}
@@
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-Hop R, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
+HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and XX \mbh{LV, please add the size of the second band}), whereas heterozygotes are three-banded (Fig.~S2).
 When only one PCR resolved well, we scored one allele for the individual.
-We used Phusion High Fidelity Enzyme (Finnzymes, Inc.) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
+We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
 PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.
@@
 For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (Region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} (66,169bp upstream from the start of the \emph{tb1} ORF; Region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 50kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Fig.~S1).
-We used Taq polymerase (New England Biolabs) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
-Annealing temperatures for sequenced region 1 and sequenced region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Fig.~S1).
+We used Taq polymerase (New England Biolabs Inc., Ipswich, Massachusetts, USA) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
+Annealing temperatures for Region 1 and Region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
 To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 $\mu$L of amplification product.
@@
 
-We cloned cleaned fragments into a TOPO-TA vector (Invitrogen, Carlsbad) using OneShot TOP10 chemically competent \emph{E.
+We cloned cleaned fragments into a TOPO-TA vector (Life Technologies, Grand Island, New York, USA) using OneShot TOP10 chemically competent \emph{E.
 coli} cells, with an extended ligation time of 30 min for a complex target fragment.
@@
 For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrapping (100 repetitions) to examine the support of nodes in our trees.
+To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}), creating neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
 For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.
@@
 
-We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and the same phased 55K SNP data from \citep{Pyhajarvi2013} that we used for LD analysis, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
+We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
@@
 We treated seeds with fungicide and germinated them in petri dishes with filter paper.
-Following germination, 206 successful germinations were then planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
+Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
 Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer.
@@
 \jri{what is this url?} We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
-Based on these initial data, we conducted a post hoc power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
+Based on these initial data, we conducted a \emph{post hoc} power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
 
@@
 Plants were watered twice daily, alternating between fertilized and non-fertilized water.
-We began phenotyping successful germinations (302) for tillering index on day 15 post fruit-case removal, and phenotyped every five days until day 50.
+We began phenotyping successful germinations (302 plants) for tillering index on day 15 post fruit-case removal, and phenotyped every five days until day 50.
 At day 50 we measured culm diameter between the third and fourth nodes.
@@
 
-Resulting tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$) \mbh{what is the alpha value here?} implemented in the car package in R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
+Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).

### 2014-05-20 manuscript.tex
Edits to Discussion
Edits to Discussion cont.

@@
 
- % Table 4 Assigment to maize and teosinte in tb1 region and chr 1 
+ % Table 4 Assignment to maize and teosinte in tb1 region and chr 1 
 \begin{table}[htbp]
@@
 
-Adaptation occurs either due to selection on standing variation or on \emph{de novo} mutations.
-Adaptation as a result of selection on standing variation has been well-described in a number of systems; for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}, variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}, and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
-Although the role of standing variation with respect to adaptation has been described in many systems, its importance to domestication is not as well studied.
-
-In maize, alleles at important domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) have been shown to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in the domestication of maize.
-The \emph{teosinte branched1} gene has long been a central focus of research concerning maize domestication, and, while previous studies have suggested that differences in plant architecture between domesticated maize and teosinte are a result of selection on standing variation, little is known about variation at this locus in teosinte \citep{Clark2006, Studer2011}.
+Adaptation occurs due to selection on standing variation or \emph{de novo} mutations.
+Adaptation from standing variation has been well-described in a number of systems; for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}, variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}, and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
+Although the adaptive role of standing variation has been described in many systems, its importance in domestication is not as well studied.
+
+In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication.
+The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus, little is known about natural variation at this locus and its ecological role in teosinte \citep{Clark2006, Studer2011}.
 \citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte \citep{Studer2011}.
@@
 Given that the \emph{Hopscotch} insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
-However, in sampling numerous individuals from many teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
-While our findings are consistent with \citet{Studer2011} in that we identified the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, possibly suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
+However, by widely sampling across teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
+While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
+Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
 Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from Jalisco include gene flow from maize, genetic drift, and natural selection.
 
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte \citep{Hufford2013}.
-Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize; \citealt{Hufford2013}).
+While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte.
+Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).
 We find no evidence of recent introgression in our analyses.
@@
 \ref{Fig3Structure}, Table \ref{Table4Q}).
-Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in some of  our \emph{parviglumis} populations.
+Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations.
 
 Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the \emph{Hopscotch} in these populations.
-If the \emph{Hopscotch} allele was introgressed in the distant past, recombination may have broken up LD, a pattern that would be consistent with our data.
+If the \emph{Hopscotch} allele was introgressed in the distant past, recombination may have broken up LD, a process that would be consistent with our data.
 We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
 In fact, early generation maize-teosinte hybrids are easy to find in these populations today (MB Hufford, pers.
-observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{Zea mays} ssp.
-\emph{mexicana} and \emph{Zea mays} ssp.
-\emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011}.
+observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{mexicana} and \emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011}.
 

### 2014-05-20 manuscript.tex
Edits to Discussion cont.

@@
 Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
-Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from Jalisco include gene flow from maize, genetic drift, and natural selection.
+Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from the Jalisco cluster include gene flow from maize, genetic drift, and natural selection.
 
@@
 Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-Moreover, analysis of linkage in the \emph{tb1} region does not reveal patterns of high LD relative to the rest of chromosome 1, and assignment to maize in this region in our STRUCTURE analysis is lower than the average across chromosome 1 (Fig.
+Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is lower than the chromosome-wide average (Fig.
 \ref{Fig3Structure}, Table \ref{Table4Q}).
@@
 We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
-In fact, early generation maize-teosinte hybrids are easy to find in these populations today (MB Hufford, pers.
+In fact, early generation maize-teosinte hybrids are common in these populations today (MB Hufford, pers.
 observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{mexicana} and \emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011}.
@@
 Drift may have played a role in the San Lorenzo \emph{parviglumis} population.
-Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings.
+Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings suggesting a bottleneck.
 \jri{deviations from HWE may be consistent too if we see excess of homozygotes.
-do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele at \emph{tb1} could have risen to high frequency entirely by chance.
-This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} allele in multiple populations in the Jalisco group.
-Moreover, available information on diversity and population structure among additional Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that might predict a high frequency of the allele across populations.
-Values of the Tajima's D statistic in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural population of \emph{parviglumis}.
+do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance.
+This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster.
+Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations.
+Finally, values of the Tajima's D statistic in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural population of \emph{parviglumis}.
 Whereas the genome-wide average of Tajima's D in genic regions of \emph{parviglumis} is 0.45 \citep{Hufford2012b}, the statistic is quite negative in the 5' UTR of \emph{tb1} (Table \ref{Table2Diversity}).
-This result is consistent with repeated selective sweeps near \emph{tb1} and an important ecological role for the gene in \emph{parviglumis}.
+This result is consistent with repeated selective sweeps near \emph{tb1} and a putative ecological role for the gene in \emph{parviglumis}.
 

### 2014-05-20 manuscript.tex
Edits to Discussion cont.

@@
 Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
-\citet{Weber2007} have described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length within a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
+\citet{Weber2007} have described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
 Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte.
@@
 One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
-In fact, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.
+Consistent with this interpretation, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.
 A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.
 \citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated phenotypic effects.
-Phenotypes were shown to cluster by taxon, indicating \emph{tb1} potentially played a role in the morphological diversification of \emph{Zea}.
-Additional analysis suggested tillering index was controlled both by \emph{tb1} and loci elsewhere in the genome.
-Clues to the identity of these loci may be found in QTL studies that identified loci controlling branching architecture \citep{DoebleyStec1991, DoebleyStec1993}.
+Phenotypes were shown to cluster by taxon, indicating \emph{tb1} may underlie morphological diversification of \emph{Zea}.
+Additional analysis in \citet{StuderDoebley2012} suggested tillering index was controlled both by \emph{tb1} and loci elsewhere in the genome.
+Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).
 Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
@@
 
-In summary, our findings demonstrate that the \emph{Hopscotch} allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
+In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
 Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
-The \emph{Hopscotch} does not appear to reduce tillering in \emph{parviglumis} as it does in maize.
+We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus plays an ecological role in teosinte.
+In contrast to domesticated maize, the \emph{Hopscotch} insertion in \emph{parviglumis} does not appear to reduce tillering.
 Other loci involved in branching architecture may regulate tillering in teosinte.
-Finally, we find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus plays an ecological role in teosinte.
-In the future, additional experiments will be needed to examine expression levels of \emph{tb1} and additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{ter1}) in conjunction with a more exhaustive phenotyping and genotyping assay.
+Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
+\emph{gt1}, \emph{tru1}, and \emph{ter1}).
+These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
 \jri{why not Phyb and phya?

### 2014-05-20 manuscript.tex
Formatting
Formatted title and footnote pages, see space for acknowledgements

@@
 \usepackage{lineno}
+\usepackage{lipsum}
 \usepackage{color}
@@
 
+\makeatletter
+\renewcommand{\maketitle}{\bgroup\vspace*{1cm}\setlength{\parindent}{0pt}
+\begin{center}
+  \textbf{\@title}
+
+  \@author
+\end{center}\egroup
+}
+\makeatother
+
+\title{\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark}
+
+\author[2]{Laura Vann}
+\author[2,3]{Thomas Kono}
+\author[2,4]{Tanja Pyh\"aj\"arvi}
+\author[2,5,7]{Matthew B. Hufford}
+\author[2,6,7]{Jeffrey Ross-Ibarra}
+\affil[2]{Department of Plant Sciences, University of California Davis}
+\affil[3]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
+\affil[4]{Department of Biology, University of Oulu}
+\affil[5]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
+\affil[6]{Center for Population Biology and Genome Center, University of California Davis}
+\affil[7]{Author for Correspondence}
+\date{}
+
 \begin{document}
 
-\title{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}
-
-\author[1]{Laura Vann}
-\author[1,2]{Thomas Kono}
-\author[1,3]{Tanja Pyh\"aj\"arvi}
-\author[1,4]{Matthew B. Hufford\thanks{mhufford@iastate.edu}}
-\author[1,5]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
-\affil[1]{Department of Plant Sciences, University of California Davis}
-\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
-\affil[3]{Department of Biology, University of Oulu}
-\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
-\affil[5]{Center for Population Biology and Genome Center, University of California Davis}
+\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}
+
+{\let\newpage\relax\maketitle}
 
@@
 
-\maketitle
-
+\clearpage
+\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted\line(1,0){120}.}
+\vspace*{1cm}
+\subsection*{Acknowledgements}
+Place brief acknowledgments, if desired, as a separate paragraph, using the following style: “The author(s) thank(s)…”. For brevity, do not use first names. Include grant acknowledgments here.
+}
 \clearpage
@@
 
+Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
+
 %Figure 1 Map 
@@
 %--------------------------------------------
-
-Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
-Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
 

### 2014-05-20 manuscript.tex
Formatting

@@
 
-\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}
+\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
 
@@
 
-\parindent=20pt
-
 \subsection*{Abstract}
 
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
+\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
 Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize.
 Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication, led us to investigate its prevalence and potential role in teosinte.
-Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
+
+\vspace*{0.2cm}
+
+\emph{Methods:}
+
+\vspace*{0.2cm}
+
+\emph{Key results:} Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
 \emph{parviglumis} and \emph{Zea mays} ssp.
@@
 \emph{parviglumis} do not suggest \emph{tb1} controls tillering in natural populations of this subspecies.
-Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte differs from domesticated maize, and that other loci may play a role in observed variation in this trait.
+
+\vspace*{0.2cm}
+
+\emph{Conclusions:} Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte differs from domesticated maize, and that other loci may play a role in observed variation in this trait.
+
+\vspace*{0.2cm}
 
 \clearpage
+
+\parindent=20pt
 

### 2014-05-20 manuscript.tex
New Abstract

@@
 
-\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of lateral organ growth, is a major QTL involved in branching differences between maize and its wild progenitor, teosinte.
-Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in domesticated maize.
-Observations of the maize \emph{tb1} allele in teosinte individuals, coupled with estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication, led us to investigate its prevalence and potential role in teosinte.
+\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene is a major QTL controlling branching differences between maize and its wild progenitor, teosinte.
+Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in maize.
+Observations of the maize \emph{tb1} allele in teosinte and estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication led us to investigate its prevalence and potential role in teosinte.
 
@@
 
-\emph{Methods:}
+\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 1110 maize and teosinte individuals using a co-dominant PCR assay.
+Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
+Phenotypic data were also collected from a single teosinte population where \emph{Hopscotch} was found segregating.
 
@@
 
-\emph{Key results:} Results from genotyping across many natural populations suggest that the \emph{Hopscotch} element is segregating at a higher than expected frequency in a number of populations of two subspecies of teosinte, \emph{Zea mays} ssp.
-\emph{parviglumis} and \emph{Zea mays} ssp.
-\emph{mexicana}.
-Analysis of linkage disequilibrium between the \emph{Hopscotch} element and variation in surrounding regions does not support a hypothesis of recent introgression from maize into teosinte.
-Population genetic signatures, however, are consistent with selection on this locus and suggest the \emph{Hopscotch} insertion at \emph{tb1} may play an ecological role in teosinte.
-Finally, two greenhouse experiments with \emph{Zea mays} ssp.
-\emph{parviglumis} do not suggest \emph{tb1} controls tillering in natural populations of this subspecies.
+\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte.
+Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.
+Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte.
+Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
 
@@
 
-\emph{Conclusions:} Our findings suggest that the role of the \emph{Hopscotch} in tillering in teosinte differs from domesticated maize, and that other loci may play a role in observed variation in this trait.
+\emph{Conclusions:} Our findings suggest the role of \emph{Hopscotch} differs between maize and teosinte.
+Future work should assess \emph{tb1} expression levels in teosinte with and without the \emph{Hopscotch} and more comprehensively phenotype teosinte to assess the ecological significance of the \emph{Hopscotch} insertion and, more broadly, the \emph{tb1} locus in teosinte.
 
 \vspace*{0.2cm}
+
+\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}
 

### 2014-05-20 manuscript.tex
AJB format

@@
 \usepackage{lipsum}
+\usepackage{indentfirst}
 \usepackage{color}
@@
 \bibliographystyle{AJB}
+\renewcommand{\refname}{\centering{LITERATURE CITED}}
 
@@
 
-\subsection*{Sampling and genotyping}
-
-We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1 and Table~S2).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1 and Table~S2).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 
-\subsection*{Sequencing}
-
-In addition to genotyping, we chose a subset of \emph{parviglumis} individuals for sequencing.
+\textbf{Sequencing---}In addition to genotyping, we chose a subset of \emph{parviglumis} individuals for sequencing.
 We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
@@
 
-\subsection*{Genotyping analysis}
-
-We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).
+\textbf{Genotyping analysis---}We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).
 To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
@@
 
-\subsection*{Sequence analysis}
-
-For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
+\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
 To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}), creating neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
@@
 
-\subsection*{Introgression analysis}
-
-In order to assess patterns of linkage disequilibrium (LD) around the \emph{Hopscotch} element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
+\textbf{Introgression analysis---}In order to assess patterns of linkage disequilibrium (LD) around the \emph{Hopscotch} element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
 We chose these populations because we had both genotyping data for the \emph{Hopscotch} as well as chromosome-wide SNP data for chromosome 1.
@@
 
-\subsection*{Phenotyping of \emph{parviglumis}}
-
-To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
+\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
 We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
@@
 
-\subsection*{Genotyping}
-
-Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
+\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2).
@@
 
-\subsection*{Sequencing}
-
-To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
+\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
 After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
@@
 \end{table}
-
-\subsection*{Evidence of introgression} %JRI HERE
-
-The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
+%JRI HERE
+
+\textbf{Evidence of introgression---}The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
@@
 
-\subsection*{Phenotyping}
-
-To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
+\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
 We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.

### 2014-05-22 manuscript.tex
edits
edits

@@
 %Code for SAS and command line stuff on Git
-%Phenotyping and genotyping data on Git (it's on Figshare?) <- http://dx.doi.org/10.6084/m9.figshare.776926
+%Phenotyping on Figshare http://dx.doi.org/10.6084/m9.figshare.776926 -- do we need to cite this in the manuscript somewhere?
 
@@
 
-\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}
+\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; QTL \lev{??}
 
@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and XX \mbh{LV, please add the size of the second band}), whereas heterozygotes are three-banded (Fig.~S2).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $/sim$1.1kb), whereas heterozygotes are three-banded (Fig.~S2).
 When only one PCR resolved well, we scored one allele for the individual.

### 2014-05-22 manuscript.tex
edits
edits

@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes.
+We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code?
+Matt is was your 'chunkfy.pl script} that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 We chose between 10-13 seeds from each of 23 sampling sites.
-We treated seeds with fungicide and germinated them in petri dishes with filter paper.
+We treated seeds with Captan fungicide (Southern Agricultural Insecticides, Inc.) \lev{do we need to say which?} and germinated them in petri dishes with filter paper.
 Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.

### 2014-05-22 manuscript.tex
edits
edits

@@
 
-Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
+Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).

### 2014-05-22 manuscript.tex
edits
edits

@@
 \mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
+\lev{so are you wanting to remove the above blip about hew stuff?}
 
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} \lev{should be calculated among all pops...I believe that is what HierFstat does} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).

### 2014-05-22 manuscript.tex
edits
edits

@@
 Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}).
-\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
+\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} \lev{the frequencies for the Hopscotch in these populations are in the supplement.} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
 Tajima's D is positive in all populations except San Lorenzo, \mbh{is the table wrong?
@@
 I'd had them in before in a paragraph but had been voted out} \jri{I'd ditch gt1 tru1 ter1 and maybe just cite some people including phyb etc.}
+
+\lev{ I had phya and phyb in before and you two voted it out in favor of other loci. please let me know which you would like in}
 

### 2014-05-22 manuscript.tex
comments, minor edits

@@
 
-\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; QTL \lev{??}
+\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; QTL \lev{??}  \mbh{not sure about QTL...might make people think we've done some QTL mapping which we haven't}
 
@@
 We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code?
-Matt is was your 'chunkfy.pl script} that grouped SNPs separated by less than 5kb into haplotypes.
+Matt is was your 'chunkfy.pl script} \mbh{we should make this script available; my vote would be to get this submitted and during review get all scripts, etc...
+in GitHub or some other public repository} that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 We chose between 10-13 seeds from each of 23 sampling sites.
-We treated seeds with Captan fungicide (Southern Agricultural Insecticides, Inc.) \lev{do we need to say which?} and germinated them in petri dishes with filter paper.
+We treated seeds with Captan fungicide (Southern Agricultural Insecticides Inc., Palmetto, Florida, USA) and germinated them in petri dishes with filter paper.
 Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
@@
 
-Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
+Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} \mbh{the box-cox transformation function in the car package uses a lambda vector, but I do not see mention of alpha} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
@@
 \lev{so are you wanting to remove the above blip about hew stuff?}
+\mbh{I say ditch it, we never come back to it}
 

### 2014-05-22 manuscript.tex
edits
edits

@@
 }
+
+\lev{Grants as far as I know were Matt's UC Mexus money, whatever money paid for the 55K data since we used that, and I am not sure what paid for greenhouse space}
 \clearpage
@@
 Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
+\lev{do you think that we could change this at all to say something to make it sound less like hopscotch has absolutely no effect?
+I think it's likely that it was something with our experiments, or that it is the effect of other loci}
 
@@
 
-\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; QTL \lev{??}  \mbh{not sure about QTL...might make people think we've done some QTL mapping which we haven't}
+\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1};
 

### 2014-05-22 manuscript.tex
edits
edits

@@
 
-\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1};
+\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; transposable element
+
+\lev{does Hopscotch need to be italicized? Don't believe we had ever italicized it before.}
 
@@
 In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
+An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering\lev{we saw a bunch of early flowering in phenotyping 2} \citep{KebromBrutnell2007}.
 The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.

### 2014-05-23 manuscript.tex
minor edits

@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $/sim$1.1kb), whereas heterozygotes are three-banded (Fig.~S2).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Fig.~S2).
 When only one PCR resolved well, we scored one allele for the individual.

### 2014-05-23 manuscript.tex
edits
edits
edits
supplement confusion/new file

@@
 \ref{Fig1Map}).
-%We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
-Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05).
-\mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
-\lev{so are you wanting to remove the above blip about hew stuff?}
-\mbh{I say ditch it, we never come back to it}
 
@@
   \begin{center}
-   \includegraphics[width=150mm]{FigS1LocusCartoon.pdf}
-   % \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+   \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
+   \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
 Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}

### 2014-05-23 manuscript.tex
JRI done through results

--- 2014-05-23 manuscript.tex
@@
 
-\lev{Grants as far as I know were Matt's UC Mexus money, whatever money paid for the 55K data since we used that, and I am not sure what paid for greenhouse space}
+\lev{Grants as far as I know were Matt's UC Mexus money, whatever money paid for the 55K data since we used that, and I am not sure what paid for greenhouse space} \jri{55K data was public, no prob.
+acknowledge support from the Dept.
+of Plant Sciences for LV's GSR and for "research funds" or something vague, as this came from startup.
+acknowledge graham coop for helpful discussion}
 \clearpage
@@
 \lev{do you think that we could change this at all to say something to make it sound less like hopscotch has absolutely no effect?
-I think it's likely that it was something with our experiments, or that it is the effect of other loci}
+I think it's likely that it was something with our experiments, or that it is the effect of other loci} \jri{other loci means tb1 doesn't solely control.
+if we have no faith in our experimental results, we shouldn't be writing the paper!
+it may be daylength, crowding whatever in our experiment, but as long as we don't think anything was wrong with data collection and analysis, I say keep as is.}
 
@@
 In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering\lev{we saw a bunch of early flowering in phenotyping 2} \citep{KebromBrutnell2007}.
+An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering\lev{we saw a bunch of early flowering in phenotyping 2} \jri{were plants taller in pheno2?
+were they more crowded?} \citep{KebromBrutnell2007}.
 The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.
@@
 Matt is was your 'chunkfy.pl script} \mbh{we should make this script available; my vote would be to get this submitted and during review get all scripts, etc...
-in GitHub or some other public repository} that grouped SNPs separated by less than 5kb into haplotypes.
+in GitHub or some other public repository} \jri{i'm ok with wiating to do after submit; I'm not OK with publication until everything is on github} that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05).
-\mbh{Jeff mentioned this needs to be fully reported in a supplementary table; alternatively we can just delete the result since we do not interpret or use it later in the manuscript.}
-\lev{so are you wanting to remove the above blip about hew stuff?}
-\mbh{I say ditch it, we never come back to it}
 
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} \lev{should be calculated among all pops...I believe that is what HierFstat does} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} \lev{should be calculated among all pops...I believe that is what HierFstat does} \jri{please check to be sure!} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
@@
 Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}).
-\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} \lev{the frequencies for the Hopscotch in these populations are in the supplement.} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
+\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} \lev{the frequencies for the Hopscotch in these populations are in the supplement.} \jri{please cite appropriate table.
+is this the last table?
+looks like it's missing column headings.} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
 Tajima's D is positive in all populations except San Lorenzo, \mbh{is the table wrong?
@@
 \end{table}
-%JRI HERE
 
@@
 We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.
-\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} genotype.
+\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
@@
   \centering
-  \caption{$r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
+  \caption{mean $r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
     \begin{tabular}{ccccc}\\\toprule
@@
 
-The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
+The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
 However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}.
-The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations.
+The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants which would arise as the result of recent admixture between populations.
 If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
@@
 The dotted lines mark the beginning of the sequenced region 50kb upstream (Sequenced region 2) and the end of the \emph{tb1} ORF.
+} \jri{ we use 50kb sometimes and 66kb elsehwere.
+is this typo, or did we really only consider 50kb?
 }
@@
 We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find a weak positive correlation between presence of the \emph{Hopscotch} and tillering index on day 40 (p=0.0848), a result indicating the \emph{Hopscotch} may actually increase tillering in \emph{parviglumis} in contrast to its phenotypic effect in maize.
-We find no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
-Additionally we find no significant correlation between tiller number and \emph{Hopscotch} genotype, or culm diameter and \emph{Hopscotch} genotype in Phenotyping 1.
+After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no correlation between genotype at the \emph{Hopscotch}  insertion and tillering index (\ref{Fig4Boxplots}), tiller number, or culm diameter.
 
@@
 For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
-We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
-We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p$<$0.0611), with no significant correlation on any day.
-Similarly, relationships between \emph{Hopscotch} genotype and tiller number and \emph{Hopscotch} genotype and culm diameter were not significant.
+We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?
+my guess is no!} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.
 
@@
 \end{centering}
-
+%JRI
 Adaptation occurs due to selection on standing variation or \emph{de novo} mutations.
@@
 I agree its a good idea and could end up being really interesting.
-Perhaps something we could look at after submission and incorporate during revisions?}
+Perhaps something we could look at after submission and incorporate during revisions?} \jri{ sure, we can look after submission, but i want to do that before publication.
+to some degree i'd feel better having stuff done before submission, but if you guys want to push to submit, this and github stuff can wait until revision}
 
@@
 
-\lev{ I had phya and phyb in before and you two voted it out in favor of other loci. please let me know which you would like in}
+\lev{ I had phya and phyb in before and you two voted it out in favor of other loci. please let me know which you would like in} \jri{don't have strong feelings here.  matt?}
 

### 2014-05-23 Supplement.tex

--- 2014-05-23 manuscript.tex
@@
-\documentclass[11pt]{article}
-\usepackage{geometry} 
-\usepackage{hyperref}
-\usepackage[leftcaption]{sidecap}
-\sidecaptionvpos{figure}{c} %position side caption
-\usepackage{lineno}
-\usepackage{lipsum}
-\usepackage{indentfirst}
-\usepackage{color}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage[pdftex]{graphicx}  
-%\usepackage{caption}
-\usepackage{natbib}
-\usepackage{authblk}
-\usepackage[table]{xcolor} %for table line color alteration
-\usepackage{subscript}
-\renewcommand{\baselinestretch}{1.5}
-\bibliographystyle{AJB}
-\renewcommand{\refname}{\centering{LITERATURE CITED}}
-
-\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
-\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
-\newcommand{\lev}[1]{\textcolor{green}{\emph{\scriptsize #1}} }
-
-%----------------------------------------
-%AUTHORS
-%----------------------------------------
-
-\makeatletter
-\renewcommand{\maketitle}{\bgroup\vspace*{1cm}\setlength{\parindent}{0pt}
-\begin{center}
-  \textbf{\@title}
-
-  \@author
-\end{center}\egroup
-}
-\makeatother
-
-\title{\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark}
-
-\author[2]{Laura Vann}
-\author[2,3]{Thomas Kono}
-\author[2,4]{Tanja Pyh\"aj\"arvi}
-\author[2,5,7]{Matthew B. Hufford}
-\author[2,6,7]{Jeffrey Ross-Ibarra}
-\affil[2]{Department of Plant Sciences, University of California Davis}
-\affil[3]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
-\affil[4]{Department of Biology, University of Oulu}
-\affil[5]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
-\affil[6]{Center for Population Biology and Genome Center, University of California Davis}
-\affil[7]{Author for Correspondence}
-\date{}
-
-\begin{document}
-
-\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
-
-{\let\newpage\relax\maketitle}
-
-\begin{linenumbers}
-	\pagewiselinenumbers
-	\renewcommand\linenumberfont{\normalfont\bfseries\small}
-
-\clearpage
-\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted\line(1,0){120}.}
-\vspace*{1cm}
-\subsection*{Acknowledgements}
-Place brief acknowledgments, if desired, as a separate paragraph, using the following style: “The author(s) thank(s)…”. For brevity, do not use first names. Include grant acknowledgments here.
-}
-
-\lev{Grants as far as I know were Matt's UC Mexus money, whatever money paid for the 55K data since we used that, and I am not sure what paid for greenhouse space}
-\clearpage
-
-%TO DO
-%check citations in last paragraph for ter1
-%reformat for AJB
-%preprint somewhere
-%finish fixing paragraph that jeff doesn't like
-%check map details
-%Code for SAS and command line stuff on Git
-%Phenotyping on Figshare http://dx.doi.org/10.6084/m9.figshare.776926 -- do we need to cite this in the manuscript somewhere?
-
-\begin{flushleft}
-
-\subsection*{Abstract}
-
-\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene is a major QTL controlling branching differences between maize and its wild progenitor, teosinte.
-Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in maize.
-Observations of the maize \emph{tb1} allele in teosinte and estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication led us to investigate its prevalence and potential role in teosinte.
-
-\vspace*{0.2cm}
-
-\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 1110 maize and teosinte individuals using a co-dominant PCR assay.
-Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
-Phenotypic data were also collected from a single teosinte population where \emph{Hopscotch} was found segregating.
-
-\vspace*{0.2cm}
-
-\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte.
-Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.
-Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte.
-Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
-\lev{do you think that we could change this at all to say something to make it sound less like hopscotch has absolutely no effect?
-I think it's likely that it was something with our experiments, or that it is the effect of other loci}
-
-\vspace*{0.2cm}
-
-\emph{Conclusions:} Our findings suggest the role of \emph{Hopscotch} differs between maize and teosinte.
-Future work should assess \emph{tb1} expression levels in teosinte with and without the \emph{Hopscotch} and more comprehensively phenotype teosinte to assess the ecological significance of the \emph{Hopscotch} insertion and, more broadly, the \emph{tb1} locus in teosinte.
-
-\vspace*{0.2cm}
-
-\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; transposable element
-
-\lev{does Hopscotch need to be italicized? Don't believe we had ever italicized it before.}
-
-\clearpage
-
-\parindent=20pt
-
-\begin{centering}
-\section*{INTRODUCTION}
-\end{centering}
-
-Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
-Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
-Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
-
-Maize (\emph{Zea mays} ssp.
-\emph{mays}) was domesticated from the teosinte \emph{Zea mays} ssp.
-\emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 9,000 B.P.
-in southwest Mexico \citep{Piperno2009, Matsuoka2002}.
-Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-Additionally, large naturally occurring populations of both \emph{Zea mays} ssp.
-\emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp.
-\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
-
-Many morphological changes are associated with maize domestication, and understanding the genetic basis of these changes has been a focus of maize research for a number of years \citep{Doebley2004}.
-One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches.
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in  branching \citep{DoebleyStecGustus1995} and tillering \citep{DoebleyStec1991} differences between maize and teosinte.
-A 4.9 kb retrotransposon (\emph{Hopscotch}) insertion into the upstream control region of \emph{tb1} in maize acts to enhance expression of \emph{tb1}, thus repressing lateral organ growth \citep{Doebley1997, Studer2011}.
-Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
-
-In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering\lev{we saw a bunch of early flowering in phenotyping 2} \citep{KebromBrutnell2007}.
-The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.
-In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
-We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus.
-Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.
-
-\begin{centering}
-\section*{MATERIALS AND METHODS}
-\end{centering}
-
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1 and Table~S2).
-DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
-We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion.
-Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
-Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Fig.~S2).
-When only one PCR resolved well, we scored one allele for the individual.
-We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
-PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.
-
-\textbf{Sequencing---}In addition to genotyping, we chose a subset of \emph{parviglumis} individuals for sequencing.
-We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (Region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} (66,169bp upstream from the start of the \emph{tb1} ORF; Region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Fig.~S1).
-We used Taq polymerase (New England Biolabs Inc., Ipswich, Massachusetts, USA) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
-Annealing temperatures for Region 1 and Region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
-To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 $\mu$L of amplification product.
-This mix was placed on a thermal cycler with the following program: 37$^{\circ}$C for 30 min, 80$^{\circ}$C for 15 min, and a final cool-down step to 4$^{\circ}$C.
-
-We cloned cleaned fragments into a TOPO-TA vector (Life Technologies, Grand Island, New York, USA) using OneShot TOP10 chemically competent \emph{E.
-coli} cells, with an extended ligation time of 30 min for a complex target fragment.
-We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96$^{\circ}$C for 5 min; then 35 cycles at 96$^{\circ}$C for 30 s, 53$^{\circ}$C for 30 s, 72$^{\circ}$C for 2 min; and a final extension at 72$^{\circ}$C for 4 min.
-We visualized amplification products for incorporation of our insert on a 1\% agarose TAE gel.
-
-Amplification products with successful incorporation of our insert were cleaned using Exonuclease 1 and Antarctic Phosphatase following the procedures detailed above, and sequenced with vector primers M13 Forward and M13 Reverse using Sanger sequencing at the College of Agriculture and Environmental Sciences (CAES) sequencing center at UC Davis.
-We aligned and trimmed primer sequences from resulting sequences using the software Geneious \citep{Kearse2012}.
-Following alignment, we verified singleton SNPs by sequencing an additional one to four colonies from each clone.
-If the singleton was not present in these additional sequences it was considered an amplification or cloning error, and we replaced the base with the base of the additional sequences.
-If the singleton appeared in at least one of the additional sequences we considered it a real variant and kept it for further analyses.
-
-\textbf{Genotyping analysis---}We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).
-To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations in which 8 or more individuals were sampled.
-To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
-BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
-We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (Table~S3).
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
-
-\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}), creating neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
-For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.
-
-\textbf{Introgression analysis---}In order to assess patterns of linkage disequilibrium (LD) around the \emph{Hopscotch} element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
-We chose these populations because we had both genotyping data for the \emph{Hopscotch} as well as chromosome-wide SNP data for chromosome 1.
-For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for SLO, EjuB, EjuA, and MSA, respectively.
-We then used Tassel \citep{Bradbury2007} to calculate linkage disequilibrium ($r^{2}$) across chromosome 1 for each population.
-
-We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
-SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code?
-Matt is was your 'chunkfy.pl script} \mbh{we should make this script available; my vote would be to get this submitted and during review get all scripts, etc...
-in GitHub or some other public repository} that grouped SNPs separated by less than 5kb into haplotypes.
-We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
-
-\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
-In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
-We chose between 10-13 seeds from each of 23 sampling sites.
-We treated seeds with Captan fungicide (Southern Agricultural Insecticides Inc., Palmetto, Florida, USA) and germinated them in petri dishes with filter paper.
-Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
-Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer.
-
-Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
-Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
-On day 65 we measured culm diameter between the third and fourth nodes of each plant.
-Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1}.
-Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (\url{http://www.ars.usda.gov}).
-\jri{what is this url?} We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
-Based on these initial data, we conducted a \emph{post hoc} power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
-
-We performed a second phenotyping experiment (Phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the \emph{Hopscotch} insertion.
-Seeds were germinated and planted on day 7 post fruit-case removal into two gallon pots.
-Plants were watered twice daily, alternating between fertilized and non-fertilized water.
-We began phenotyping successful germinations (302 plants) for tillering index on day 15 post fruit-case removal, and phenotyped every five days until day 50.
-At day 50 we measured culm diameter between the third and fourth nodes.
-We extracted DNA and genotyped plants following the same guidelines as in Phenotyping 1.
-
-Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} \mbh{the box-cox transformation function in the car package uses a lambda vector, but I do not see mention of alpha} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
-We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
-Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
-
-\begin{centering}
-\section*{RESULTS}
-\end{centering}
-
-\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
-Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).
-
-Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} \lev{should be calculated among all pops...I believe that is what HierFstat does} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
-Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
-
-%Figure 1 Map 
-%--------------------------------------------
-\begin{SCfigure}[][t!]
-  \centering \label{Fig1Map}
-   \includegraphics[width=0.7\textwidth]{Fig1Map.pdf}
-    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
-Size of circles reflects number of alleles sampled.
-}
-\end{SCfigure}
-%--------------------------------------------
-
-% Table 1
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
-
-    \begin{tabular}{cccc}\\\toprule
-    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
-    EjuA \& EjuB & 0     & 0     & 0 \\
-    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
-    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
-    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
-    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
-    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
-    \end{tabular}
-  \label{Table1Fst}
-\end{table}
-
-\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
-After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
-For Region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}).
-\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} \lev{the frequencies for the Hopscotch in these populations are in the supplement.} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive in all populations except San Lorenzo, \mbh{is the table wrong?
-MSA is the only negative value in the table} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
-Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch} \jri{table 1 shows 0.016 for hopscotch, not 0.
-which is right?}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Figure~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
-
-% Table 2 Diversity Statistics
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
-    \begin{tabular}{ccccc}\\\toprule
-    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
-    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
-    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
-    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
-    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
-    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
-    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
-    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
-    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
-    \end{tabular}
-  \label{Table2Diversity}
-\end{table}
-%JRI HERE
-
-\textbf{Evidence of introgression---}The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
-Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
-To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one and specifically in the \emph{tb1} region.
-If the \emph{Hopscotch} is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.
-\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} genotype.
-In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
-\mbh{LV, please go through and make sure the data entered into all the tables is correct.
-In Table3, both sequenced regions were labeled as "Region 1".
-I changed the second to Region 2 but don't know if the data in this column are really from Region 2}
-
-% Table 3 $r^{2}$ values
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{$r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
-    \begin{tabular}{ccccc}\\\toprule
-    \textbf{Population} & \textbf{Chr. 1} & \textbf{\emph{tb1} region} & \textbf{Region 1} & \textbf{Region 2} \\\midrule
-    Ejutla A & 0.095 & 0.050 & 0.747 & 0.215 \\
-    Ejutla B & 0.069 & 0.051 & 0.660 & 0.186 \\
-    La Mesa & 0.070 & 0.053 & 0.914 & 0.766 \\
-    San Lorenzo & 0.101 & 0.067 & 0.912 & 0.636 \\
-    \bottomrule
-    \end{tabular}
-  \label{Table3R2}
-\end{table}
-
-% Figure 2. LD plots
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=140mm]{Fig2LDPlots.pdf}
-    \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
-The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
-A) Ejutla A; B) Ejutla B; C) La Mesa; D).
-San Lorenzo}  \jri{this needs description of what upper and lower triangle are}
-\label{Fig2LD}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
-However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}.
-The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations.
-If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig.
-\ref{Fig3Structure}).
-
- % Table 4 Assignment to maize and teosinte in tb1 region and chr 1 
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
-    \begin{tabular}{ccccc} \
-          & \multicolumn{2}{c}{\textbf{\emph{tb1} region}} & \multicolumn{2}{c}{\textbf{Chr 1}} \\\toprule
-    \bf{Population} & \bf{Maize} & \bf{Teosinte} & \bf{Maize} & \bf{Teosinte} \\\midrule
-    Ejutla A & 0.022 & 0.978 & 0.203 & 0.797 \\
-    Ejutla B & 0.019 & 0.981 & 0.187 & 0.813 \\
-    La Mesa & 0.012 & 0.988 & 0.193 & 0.807 \\
-    San Lorenzo & 0.016 & 0.984 & 0.205 & 0.795 \\\bottomrule
-    \end{tabular}
-  \label{Table4Q}
-\end{table}
-
-% Figure 3 STRUCTURE Plot 4 Pops
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=90mm]{Fig3Structure.pdf}
-    \caption{STRUCTURE assignment to maize across a section of chromosome 1.
-The dotted lines mark the beginning of the sequenced region 50kb upstream (Sequenced region 2) and the end of the \emph{tb1} ORF.
-}
-\label{Fig3Structure}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
-We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
-We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find a weak positive correlation between presence of the \emph{Hopscotch} and tillering index on day 40 (p=0.0848), a result indicating the \emph{Hopscotch} may actually increase tillering in \emph{parviglumis} in contrast to its phenotypic effect in maize.
-We find no correlation between tillering index and genotype on any other day (\ref{Fig4Boxplots}).
-Additionally we find no significant correlation between tiller number and \emph{Hopscotch} genotype, or culm diameter and \emph{Hopscotch} genotype in Phenotyping 1.
-
-% Figure 4 Phenotyping BoxPlots
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=95mm]{Fig4Boxplots.pdf}
-    \caption{Box-plots showing tillering index in greenhouse grow-outs for Phenotyping 1 (A) and Phenotyping 2 (B).
-White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.
-Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
-\jri{please explain whiskers and dots on figure too.}}
-\label{Fig4Boxplots}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.
-For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
-We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
-We found similar patterns, with a weak positive correlation between tillering index and \emph{Hopscotch} genotype at day 40 (p$<$0.0611), with no significant correlation on any day.
-Similarly, relationships between \emph{Hopscotch} genotype and tiller number and \emph{Hopscotch} genotype and culm diameter were not significant.
-
-\begin{centering}
-\section*{DISCUSSION}
-\end{centering}
-
-Adaptation occurs due to selection on standing variation or \emph{de novo} mutations.
-Adaptation from standing variation has been well-described in a number of systems; for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}, variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}, and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
-Although the adaptive role of standing variation has been described in many systems, its importance in domestication is not as well studied.
-
-In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication.
-The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus, little is known about natural variation at this locus and its ecological role in teosinte \citep{Clark2006, Studer2011}.
-\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte \citep{Studer2011}.
-
-Given that the \emph{Hopscotch} insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
-However, by widely sampling across teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
-While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
-Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from the Jalisco cluster include gene flow from maize, genetic drift, and natural selection.
-
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte.
-Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).
-We find no evidence of recent introgression in our analyses.
-Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is lower than the chromosome-wide average (Fig.
-\ref{Fig3Structure}, Table \ref{Table4Q}).
-Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations.
-
-Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the \emph{Hopscotch} in these populations.
-If the \emph{Hopscotch} allele was introgressed in the distant past, recombination may have broken up LD, a process that would be consistent with our data.
-We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
-In fact, early generation maize-teosinte hybrids are common in these populations today (MB Hufford, pers.
-observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{mexicana} and \emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011}.
-
-Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.
-Drift may have played a role in the San Lorenzo \emph{parviglumis} population.
-Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings suggesting a bottleneck.
-\jri{deviations from HWE may be consistent too if we see excess of homozygotes.
-do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance.
-This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster.
-Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations.
-Finally, values of the Tajima's D statistic in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural population of \emph{parviglumis}.
-Whereas the genome-wide average of Tajima's D in genic regions of \emph{parviglumis} is 0.45 \citep{Hufford2012b}, the statistic is quite negative in the 5' UTR of \emph{tb1} (Table \ref{Table2Diversity}).
-This result is consistent with repeated selective sweeps near \emph{tb1} and a putative ecological role for the gene in \emph{parviglumis}.
-
-\jri{do we know the Hop genotype for sequenced lines?
-can we separate the sequences into hop/no hop and look for differences?
-it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check} \mbh{I've added a few sentences on selection.
-Do we still want to compare sequences with and without \emph{Hopscotch}?
-I agree its a good idea and could end up being really interesting.
-Perhaps something we could look at after submission and incorporate during revisions?}
-
-Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
-\citet{Weber2007} have described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
-Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte.
-We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
-One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
-Consistent with this interpretation, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.
-A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.
-\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated phenotypic effects.
-Phenotypes were shown to cluster by taxon, indicating \emph{tb1} may underlie morphological diversification of \emph{Zea}.
-Additional analysis in \citet{StuderDoebley2012} suggested tillering index was controlled both by \emph{tb1} and loci elsewhere in the genome.
-Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
-\emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}).
-Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
-
-In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
-Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
-We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus plays an ecological role in teosinte.
-In contrast to domesticated maize, the \emph{Hopscotch} insertion in \emph{parviglumis} does not appear to reduce tillering.
-Other loci involved in branching architecture may regulate tillering in teosinte.
-Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{ter1}).
-These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
-\jri{why not Phyb and phya?
-} \lev{Are they necessary to include?
-I'd had them in before in a paragraph but had been voted out} \jri{I'd ditch gt1 tru1 ter1 and maybe just cite some people including phyb etc.}
-
-\lev{ I had phya and phyb in before and you two voted it out in favor of other loci. please let me know which you would like in}
-
-\jri{please check format of supp figs and tables; some are running off the page.
-you can use "longtable" to fix that (ask Paul for example).
-check fig/table references, bibliography, etc.
-what does "rotation" mean in supp.
-table 3?
-it isn't mentioned in methods.
-please check that all the tables and figs (including supplement) are referenced in the text.}
-
-%\jri{seems like we should end with at least a hypothesis.
-as is we are saying we ruled out drift, introgression, and selection.
-what else is there?
-perhaps we should think about selection or introgression scenarios that could explain our data.}
-
-%\mbh{I'd lean toward saying that 1) \emph{Hopscotch} does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, \emph{Hopscotch} may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the \emph{Hopscotch}.}
-
-%\lev{I went with Matt's suggestions.
-I am happy to include such scenarios, but everything I have thought of seems rather unlikely; i.e.
-only introgression or selection in Jalisco.
-Jeff I can mention that maybe parv is more of a weed there than other locations, but i thought that maize pollen spread rather substantial distances.
-If so wouldn't a number of our parv and mex sites would be within pollen dispersal distance?}
-
-\clearpage
-\bibliography{refs,bob}
-
-%-------------------------------------------------
-Supplementary Materials
-%-------------------------------------------------
-\clearpage
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\renewcommand{\figurename}{Sup.\Fig.}
-\renewcommand{\tablename}{Sup.\Table}
-
 % Figure S1. Cartoon representation of tb1 locus
@@
    \includegraphics[width=150mm]{FigS2Gel.png}
-   % \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+   \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
 \label{FigS2Gel}
@@
    \includegraphics[width=150mm]{FigS3NJtrees.pdf}
-    %\caption{Neighbor-Joining tree of the sequenced region in the 5?
+    \caption{Neighbor-Joining tree of the sequenced region in the 5?
 UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}.
@@
   \centering
-  %\caption{Accessions of \emph{Zea mays} ssp.
+  \caption{Accessions of \emph{Zea mays} ssp.
 \emph{mexicana} (RIMME) and \emph{Zea mays} ssp.
@@
   \centering
- % \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
+ \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
     \begin{tabular}{ccc}
@@
   \centering
-  %\caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}
+  \caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}
     \begin{tabular}{rrrrrrrrrrrr}
@@
   \centering
-  %\caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
+  \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
     \begin{tabular}{ccccc}
     \toprule
-    %\textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
+    \textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
     \midrule
@@
 \end{table}%
-
-\end{flushleft}
-
-\end{linenumbers}
-
-\end{document}

### 2014-05-23 manuscript.tex
JRI done. peace out.

@@
 \end{centering}
-%JRI
 Adaptation occurs due to selection on standing variation or \emph{de novo} mutations.
@@
 In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication.
-The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus, little is known about natural variation at this locus and its ecological role in teosinte \citep{Clark2006, Studer2011}.
+The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus \citep{Clark2006, Studer2011}, little is known about natural variation at this locus and its ecological role in teosinte.
+\jri{can we say this with the weber paper and studer allelic series papers?
+perhaps rephrase?}
+
 \citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte \citep{Studer2011}. 
-
 Given that the \emph{Hopscotch} insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
 However, by widely sampling across teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
-While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested \citep{Studer2011}.
-Many of our populations with high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
+While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
+Many of our populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, perhaps suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
 Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from the Jalisco cluster include gene flow from maize, genetic drift, and natural selection.
 
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte.
-Furthermore, \citet{Hufford2013} showed that domestication loci, such as \emph{tb1}, are particularly resistant to introgression in both directions of gene flow (i.e., maize to teosinte and teosinte to maize).
+While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci \citet{Hufford2013}.
 We find no evidence of recent introgression in our analyses.
-Clustering patterns in our NJ trees do not reflect a pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is lower than the chromosome-wide average (Fig.
+Clustering in our NJ trees do not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
+Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig.
 \ref{Fig3Structure}, Table \ref{Table4Q}).
 Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations.
+\jri{ not worth includind in ms.
+but interesting the structure plot shows an uptick of introgression right around the hopscotch }
 
@@
 Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.
-Drift may have played a role in the San Lorenzo \emph{parviglumis} population.
-Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings suggesting a bottleneck.
-\jri{deviations from HWE may be consistent too if we see excess of homozygotes.
-do we?} Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance.
-This bottleneck, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster.
+Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings.
+Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance.
+A bottleneck in San Lorenzo, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster.
 Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations.
-Finally, values of the Tajima's D statistic in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural population of \emph{parviglumis}.
-Whereas the genome-wide average of Tajima's D in genic regions of \emph{parviglumis} is 0.45 \citep{Hufford2012b}, the statistic is quite negative in the 5' UTR of \emph{tb1} (Table \ref{Table2Diversity}).
-This result is consistent with repeated selective sweeps near \emph{tb1} and a putative ecological role for the gene in \emph{parviglumis}.
-
-\jri{do we know the Hop genotype for sequenced lines?
+Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural populations of \emph{parviglumis}.
+Overall nucleotide diversity is \jri{X\%} less than seen in the sequences from the 66kb upstream regions, and Tajima's D is considerably lower and consistently negative.
+Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in \emph{parviglumis}.
+\jri{whoa there Dr.
+Gillespie!
+i'm not so sure of this.
+first illumina data are wonky and the TajD from Hufford2012b drops all singletons.
+if we are to compare to other data it should be to Tiffin and Moeller.
+do we know the Hop genotype for sequenced lines?
 can we separate the sequences into hop/no hop and look for differences?
@@
 Perhaps something we could look at after submission and incorporate during revisions?} \jri{ sure, we can look after submission, but i want to do that before publication.
-to some degree i'd feel better having stuff done before submission, but if you guys want to push to submit, this and github stuff can wait until revision}
+to some degree i'd feel better having stuff done before submission especially if we want to argue selection, but if you guys want to push to submit, this and github stuff can wait until revision}
 
 Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
-\citet{Weber2007} have described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
+\citet{Weber2007}  described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
 Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte.
@@
 A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.
-\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated phenotypic effects.
+\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated their phenotypic effects.
 Phenotypes were shown to cluster by taxon, indicating \emph{tb1} may underlie morphological diversification of \emph{Zea}.
@@
 Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
-\emph{tassel-replaces-upper-ears1} (\emph{tru1}), for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}).
+\emph{tru1}, for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}).
 Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
+\jri{worth also mentioning possibly crowding or other non-normal factors of our greenhouse?}
 
@@
 Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
-We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus plays an ecological role in teosinte.
-In contrast to domesticated maize, the \emph{Hopscotch} insertion in \emph{parviglumis} does not appear to reduce tillering.
-Other loci involved in branching architecture may regulate tillering in teosinte.
+We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus may play an ecological role in teosinte.
+In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in \emph{parviglumis}.
 Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
@@
 These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
-\jri{why not Phyb and phya?
-} \lev{Are they necessary to include?
-I'd had them in before in a paragraph but had been voted out} \jri{I'd ditch gt1 tru1 ter1 and maybe just cite some people including phyb etc.}
-
-\lev{ I had phya and phyb in before and you two voted it out in favor of other loci. please let me know which you would like in} \jri{don't have strong feelings here.  matt?}
 
@@
 please check that all the tables and figs (including supplement) are referenced in the text.}
-
-%\jri{seems like we should end with at least a hypothesis.
-as is we are saying we ruled out drift, introgression, and selection.
-what else is there?
-perhaps we should think about selection or introgression scenarios that could explain our data.}
-
-%\mbh{I'd lean toward saying that 1) \emph{Hopscotch} does not appear to uniformly reduce tillering in teosinte as it seems to in maize; 2) other loci are likely involved in regulating tillering; 3) given its high frequency in several teosinte populations, \emph{Hopscotch} may play additional ecological roles not obvious here based on what we chose to phenotype and, in the future,  additional experiments are needed to look at 1) expression level differences at various loci shown to affect tillering in teosinte; and 2) to more completely phenotype teosinte plants with and without the \emph{Hopscotch}.}
-
-%\lev{I went with Matt's suggestions.
-I am happy to include such scenarios, but everything I have thought of seems rather unlikely; i.e.
-only introgression or selection in Jalisco.
-Jeff I can mention that maybe parv is more of a weed there than other locations, but i thought that maize pollen spread rather substantial distances.
-If so wouldn't a number of our parv and mex sites would be within pollen dispersal distance?}
 

### 2014-05-23 manuscript.tex
edits
edits

@@
 Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).
-%We examined Hardy-Weinberg equilibrium in a total of 14 populations (10 \emph{parviglumis} and 4 \emph{mexicana}) with more than 8 individuals sampled per population.
-Three populations (RIMPA0073, RIMPA0093, and RIMPA0158) show evidence of deviations from expected genotype frequencies under the assumptions of HWE (p$<$0.05).
-
-Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+\ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} \lev{should be calculated among all pops...I believe that is what HierFstat does} \jri{please check to be sure!} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).

### 2014-05-23 Supplement.tex
edits
edits
supp
supp

@@
+%Online Supplemental Materials
+%Authors may wish to augment their manuscripts with online supplemental materials (e.g., large data sets, three-dimensional reconstructions, simulations, real-time movies, color photographs).
+Upload these appendices as separate files with the initial manuscript submission.
+Include a header on each file using this format: Smith et al.?American Journal of Botany 99(#): ###-###.
+2012.
+?
+Data Supplement S1 ?
+Page 1?.
+Name online supplements Appendix S1, Appendix S2, etc, in the order in which they appear in the text, regardless of whether they are tables, figures, text, other media, or a combination thereof.
+In the manuscript, after the mention of an online appendix, include the following: ?(see Supplemental Data with the online version of this article)?.
+
+%Note that if authors wish to submit long DNA sequence appendices as supplemental material, they should select the "DNA sequences (online-only supplemental)" option on Editorial Manager.
+This ensures that lengthy appendices are not built into the reviewers' PDF, but are still accessible to the reviewers.
+
+\begin{document}
+
+\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
+
 % Figure S1. Cartoon representation of tb1 locus
@@
 
+% Table 1
+\begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
+
+    \begin{tabular}{cccc}\\\toprule
+    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
+    EjuA \& EjuB & 0     & 0     & 0 \\
+    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
+    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
+    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
+    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
+    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
+    \end{tabular}
+  \label{Table1Fst}
+\end{table}
+
 % Table S1. Parv and mex samples 
@@
 }
-    \begin{tabular}{rrrrrr}
-%    \toprule
-    Accession & USDA Accession ID & Locality & Number alleles sampled & \emph{Hopscotch} Frequency & No \emph{Hopscotch} Frequency \\
- %   \midrule
+    \begin{tabular}{cccc}\\\toprule
+    \textbf{Accession} & \textbf{USDA Accession ID} & \textbf{Locality} & \textbf{Number Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\
     RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
@@
     RIMPA0159 & N/A   & Jalisco, Mexico & 26    & 0     & 1 \\
-    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
- %   \bottomrule
+    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\\bottomrule
     \end{tabular}

### 2014-05-23 manuscript.tex
incorporating edits

@@
 \usepackage[pdftex]{graphicx}  
-%\usepackage{caption}
+\usepackage{caption}
 \usepackage{natbib}
@@
 Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
-\lev{do you think that we could change this at all to say something to make it sound less like hopscotch has absolutely no effect?
-I think it's likely that it was something with our experiments, or that it is the effect of other loci} \jri{other loci means tb1 doesn't solely control.
-if we have no faith in our experimental results, we shouldn't be writing the paper!
-it may be daylength, crowding whatever in our experiment, but as long as we don't think anything was wrong with data collection and analysis, I say keep as is.}
-
 \vspace*{0.2cm}
@@
 \textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; transposable element
-
-\lev{does Hopscotch need to be italicized? Don't believe we had ever italicized it before.}
 
@@
 In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering\lev{we saw a bunch of early flowering in phenotyping 2} \jri{were plants taller in pheno2?
-were they more crowded?} \citep{KebromBrutnell2007}.
+An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
 The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.
@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-We created haplotype blocks using a custom Perl script \lev{is this where I should cite Matt's paper, or his code?
-Matt is was your 'chunkfy.pl script} \mbh{we should make this script available; my vote would be to get this submitted and during review get all scripts, etc...
-in GitHub or some other public repository} \jri{i'm ok with wiating to do after submit; I'm not OK with publication until everything is on github} that grouped SNPs separated by less than 5kb into haplotypes.
+We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
    \includegraphics[width=90mm]{Fig3Structure.pdf}
-    \caption{STRUCTURE assignment to maize across a section of chromosome 1.
-The dotted lines mark the beginning of the sequenced region 50kb upstream (Sequenced region 2) and the end of the \emph{tb1} ORF.
-} \jri{ we use 50kb sometimes and 66kb elsehwere.
-is this typo, or did we really only consider 50kb?
-}
+    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF. }
 \label{Fig3Structure}
@@
 In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication.
-The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus \citep{Clark2006, Studer2011}, little is known about natural variation at this locus and its ecological role in teosinte.
-\jri{can we say this with the weber paper and studer allelic series papers?
-perhaps rephrase?}
+The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus \citep{Clark2006, Studer2011}, much remains to be discovered regarding natural variation at this locus and its ecological role in teosinte.
 
@@
 
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci \citet{Hufford2013}.
+While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci.
 We find no evidence of recent introgression in our analyses.
@@
 Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations.
-\jri{ not worth includind in ms.
-but interesting the structure plot shows an uptick of introgression right around the hopscotch }
 
@@
    \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
-   \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+  %\caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
 Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}

### 2014-05-23 manuscript.tex
edits
edits
continued final edits

@@
 Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural populations of \emph{parviglumis}.
-Overall nucleotide diversity is \jri{X\%} less than seen in the sequences from the 66kb upstream regions, and Tajima's D is considerably lower and consistently negative.
+Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative.
+In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}.
 Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in \emph{parviglumis}.
-\jri{whoa there Dr.
-Gillespie!
-i'm not so sure of this.
-first illumina data are wonky and the TajD from Hufford2012b drops all singletons.
-if we are to compare to other data it should be to Tiffin and Moeller.
-do we know the Hop genotype for sequenced lines?
-can we separate the sequences into hop/no hop and look for differences?
-it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check} \mbh{I've added a few sentences on selection.
-Do we still want to compare sequences with and without \emph{Hopscotch}?
-I agree its a good idea and could end up being really interesting.
-Perhaps something we could look at after submission and incorporate during revisions?} \jri{ sure, we can look after submission, but i want to do that before publication.
-to some degree i'd feel better having stuff done before submission especially if we want to argue selection, but if you guys want to push to submit, this and github stuff can wait until revision}
 
@@
 Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
-\jri{worth also mentioning possibly crowding or other non-normal factors of our greenhouse?}
+Finally, although photoperiod for Phenotyping 2 reasonably approximated that of the normal \emph{parviglumis} growing season, greenhouse-specific environmental conditions (plant density, light regime, etc...) may have contributed to tillering responses different from those found in nature, obscuring the effect of the \emph{Hopscotch} insertion on variation.
 

### 2014-05-25 Supplement.tex
supplement
trying to get tables formatted…
edits
edits
edits
edits. having trouble with supplement, trying to figure out issues
edits
fixed so in pdf!
edits
figure out long table stuff
edits
trying to figure out long table formatting

@@
 This ensures that lengthy appendices are not built into the reviewers' PDF, but are still accessible to the reviewers.
-
+\documentclass[11pt]{article}
+\usepackage{geometry} 
+\usepackage{hyperref}
+\usepackage[leftcaption]{sidecap}
+\sidecaptionvpos{figure}{c} %position side caption
+\usepackage{lineno}
+\usepackage{lipsum}
+\usepackage{indentfirst}
+\usepackage{color}
+\usepackage{multirow}
+\usepackage{booktabs}
+\usepackage[pdftex]{graphicx}  
+\usepackage{caption}
+\usepackage{natbib}
+\usepackage{authblk}
+\usepackage[table]{xcolor} %for table line color alteration
+\usepackage{subscript}
+\usepackage{longtable}
 \begin{document}
@@
   \begin{center}
-   \includegraphics[width=150mm]{FigS2Gel.png}
+   \includegraphics[width=140mm]{FigS2Gel.png}
    \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
@@
   \begin{center}
-   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
+   \includegraphics[width=140mm]{FigS3NJtrees.pdf}
     \caption{Neighbor-Joining tree of the sequenced region in the 5?
@@
 
-% Table 1
+% Table S1
 \begin{table}[htbp]
@@
 % Table S1. Parv and mex samples 
-\begin{table}[htbp]
-  \centering
+\setlength\LTleft{-30pt}            
+\setlength\LTright{-30pt}
+\begin{longtable}{ccp{5cm}cp{5cm}ccc}
 \caption{Accessions of \emph{Zea mays} ssp.
@@
 }
-    \begin{tabular}{cccc}\\\toprule
-    \textbf{Accession} & \textbf{USDA Accession ID} & \textbf{Locality} & \textbf{Number Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\
+\hline
+ \textbf{Accession} & \textbf{USDA ID} & \textbf{Locality} & \textbf{Number Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\\toprule
+\endhead
+\hline
     RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
-    RIMME0006 & 566673 & Durango, Mexico & 2     & 0     & 1 \\
-    RIMME0007 & 566680 & Guanajuato, Mexico & 2     & 0     & 1 \\
-    RIMME0008 & 566681 & Michoacan, Mexico & 2     & 0     & 1 \\
-    RIMME0009 & 566682 & Distrito Federal, Mexico & 2     & 0     & 1 \\
-    RIMME0011 & 566685 & Mexico, Mexico & 2     & 0     & 1 \\
+    RIMME0006 & 566673 & Durango & 2	& 0 & 1 \\
+    RIMME0007 & 566680 & Guanajuato & 2 & 0 & 1 \\
+    RIMME0008 & 566681 & Michoacan & 2 & 0 & 1 \\
+    RIMME0009 & 566682 & Distrito Federal & 2 & 0 & 1 \\
+    RIMME0011 & 566685 & Mexico & 2     & 0     & 1 \\
     RIMME0014 & 714151 & Breeders line; Puga: 11066 & 6     & 0     & 1 \\
-    RIMME0017 & 699874 & Ayotlan, Mexico & 8     & 0     & 1 \\
-    RIMME0021 & N/A   & El Porvenir, Mexico & 69    & 0.173913043 & 0.826086957 \\
-    RIMME0026 & N/A   & Opopeo, Mexico & 42    & 0.071428571 & 0.928571429 \\
-    RIMME0028 & N/A   & Puruandiro, Mexico & 28    & 0.035714286 & 0.964285714 \\
-    RIMME0029 & N/A   & Ixtlan, Mexico & 35    & 0     & 1 \\
-    RIMME0030 & N/A   & San Pedro, Mexico & 27    & 0     & 1 \\
-    RIMME0031 & N/A   & Tenango del Aire, Mexico & 25    & 0.08  & 0.92 \\
-    RIMME0032 & N/A   & Nabogame, Mexico & 24    & 0     & 1 \\
-    RIMME0033 & N/A   & Puerta Encantada, Mexico & 25    & 0     & 1 \\
-    RIMME0034 & N/A   & Santa Clara, Mexico & 23    & 0     & 1 \\
-    RIMME0035 & N/A   & Xochimilco, Mexico & 25    & 0     & 1 \\
-    RIMPA0001 & 87168 & El Salado, Mexico & 4     & 0     & 1 \\
-    RIMPA0003 & 87171 & Mazatlan, Mexico & 8     & 0.125 & 0.875 \\
+    RIMME0017 & 699874 & Ayotlan & 8     & 0     & 1 \\
+    RIMME0021 & N/A   & El Porvenir & 69    & 0.17 & 0.83 \\
+    RIMME0026 & N/A   & Opopeo & 42    & 0.07 & 0.93 \\
+    RIMME0028 & N/A   & Puruandiro & 28    & 0.04 & 0.96 \\
+    RIMME0029 & N/A   & Ixtlan & 35    & 0     & 1 \\
+    RIMME0030 & N/A   & San Pedro & 27    & 0     & 1 \\
+    RIMME0031 & N/A   & Tenango del Aire & 25    & 0.08  & 0.92 \\
+    RIMME0032 & N/A   & Nabogame & 24    & 0     & 1 \\
+    RIMME0033 & N/A   & Puerta Encantada & 25    & 0     & 1 \\
+    RIMME0034 & N/A   & Santa Clara & 23    & 0     & 1 \\
+    RIMME0035 & N/A   & Xochimilco & 25    & 0     & 1 \\
+    RIMPA0001 & 87168 & El Salado & 4     & 0     & 1 \\
+    RIMPA0003 & 87171 & Mazatlan & 8     & 0.13 & 0.87 \\
     RIMPA0017 & 87200 & N/A   & 4     & 0     & 1 \\
-    RIMPA0019 & 87213 & El Salado, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0029 & 87244 & N/A   & 2     & 0.5   & 0.5 \\
+    RIMPA0019 & 87213 & El Salado & 2     & 0.50   & 0.50 \\
+    RIMPA0029 & 87244 & N/A   & 2     & 0.50   & 0.50 \\
     RIMPA0031 & 87249 & N/A   & 2     & 0.5   & 0.5 \\
-    RIMPA0035 & 87288 & Jalisco, Mexico & 4     & 0     & 1 \\
-    RIMPA0040 & 288185 & Mexico, Mexico & 4     & 0     & 1 \\
-    RIMPA0042 & 288187 & Guerrero, Mexico & 4     & 0.25  & 0.75 \\
-    RIMPA0043 & 288188 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0045 & 288193 & Guerrero, Mexico & 4     & 0     & 1 \\
+    RIMPA0035 & 87288 & Jalisco & 4     & 0     & 1 \\
+    RIMPA0040 & 288185 & Mexico & 4     & 0     & 1 \\
+    RIMPA0042 & 288187 & Guerrero & 4     & 0.25  & 0.75 \\
+    RIMPA0043 & 288188 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0045 & 288193 & Guerrero & 4     & 0     & 1 \\
     RIMPA0055 & 714152 & Breeders line & 2     & 0     & 1 \\
-    RIMPA0056 & 714153 & Breeders line & 2     & 0.5   & 0.5 \\
-    RIMPA0057 & 714154 & Breeders line & 2     & 0.5   & 0.5 \\
-    RIMPA0058 & N/A   & N/A   & 4     & 0.5   & 0.5 \\
+    RIMPA0056 & 714153 & Breeders line & 2     & 0.50   & 0.50 \\
+    RIMPA0057 & 714154 & Breeders line & 2     & 0.50   & 0.50 \\
+    RIMPA0058 & N/A   & N/A   & 4     & 0.50   & 0.50 \\
     RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
@@
     RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
-    RIMPA0069 & 699862 & Ixtlan, Mexico & 14    & 0.142857143 & 0.857142857 \\
-    RIMPA0070 & 699863 & Benito Jaurez, Mexico & 16    & 0     & 1 \\
-    RIMPA0071 & 699864 & Tuzantla, Mexico & 28    & 0     & 1 \\
-    RIMPA0072 & 699865 & Tiquicheo, Mexico & 16    & 0     & 1 \\
-    RIMPA0073 & 699866 & Tiquicheo, Mexico & 16    & 0.125 & 0.875 \\
-    RIMPA0074 & 699867 & Huetamo, Mexico & 12    & 0     & 1 \\
-    RIMPA0075 & 699868 & Huetamo, Mexico & 2     & 0     & 1 \\
-    RIMPA0076 & 699869 & Huetamo, Mexico & 4     & 0     & 1 \\
-    RIMPA0077 & 699870 & Caracuaro, Mexico & 2     & 0     & 1 \\
-    RIMPA0078 & 699871 & Caracuaro, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0079 & 699872 & Villa Madero, Mexico & 14    & 0     & 1 \\
-    RIMPA0080 & 699873 & Guachinango, Mexico & 12    & 0     & 1 \\
-    RIMPA0081 & 699875 & Ameca, Mexico & 16    & 0     & 1 \\
-    RIMPA0083 & 699877 & Tepoztlan, Mexico & 14    & 0     & 1 \\
-    RIMPA0084 & 699878 & Tepoztlan, Mexico & 16    & 0     & 1 \\
-    RIMPA0085 & 699879 & Miahuatlan, Mexico & 16    & 0     & 1 \\
-    RIMPA0086 & 699880 & Miahuatlan, Mexico & 16    & 0.0625 & 0.9375 \\
-    RIMPA0087 & 699881 & Tecoanapa, Mexico & 24    & 0     & 1 \\
-    RIMPA0089 & 699883 & Guerrero, Mexico & 12    & 0     & 1 \\
-    RIMPA0090 & 699884 & Guerrero, Mexico & 10    & 0     & 1 \\
-    RIMPA0091 & 699885 & Guerrero, Mexico & 16    & 0     & 1 \\
-    RIMPA0092 & 699886 & Guerrero, Mexico & 10    & 0     & 1 \\
-    RIMPA0093 & 699887 & Guerrero, Mexico & 26    & 0.076923077 & 0.923076923 \\
-    RIMPA0094 & 699888 & Guerrero, Mexico & 2     & 0     & 1 \\
-    RIMPA0095 & 699889 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0096 & 699890 & Guerrero, Mexico & 26    & 0.038461538 & 0.961538462 \\
-    RIMPA0097 & 699891 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0098 & 699892 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0099 & 699893 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0100 & 699894 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0101 & 699895 & Guerrero, Mexico & 2     & 0     & 1 \\
-    RIMPA0103 & 699897 & Guerrero, Mexico & 2     & 0     & 1 \\
-    RIMPA0104 & 699898 & Guerrero, Mexico & 22    & 0.090909091 & 0.909090909 \\
-    RIMPA0105 & 699899 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0106 & 699900 & Guerrero, Mexico & 6     & 0.333333333 & 0.666666667 \\
-    RIMPA0107 & 699901 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0108 & 699902 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0109 & 699903 & Michoacan, Mexico & 4     & 0.25  & 0.75 \\
-    RIMPA0110 & 699904 & Michoacan, Mexico & 2     & 0     & 1 \\
-    RIMPA0111 & 699905 & Michoacan, Mexico & 4     & 0     & 1 \\
-    RIMPA0112 & 699906 & Michoacan, Mexico & 4     & 0.25  & 0.75 \\
-    RIMPA0114 & 699908 & Michoacan, Mexico & 6     & 0.166666667 & 0.833333333 \\
-    RIMPA0116 & 699910 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMPA0117 & 699911 & Mexico, Mexico & 4     & 0     & 1 \\
-    RIMPA0118 & 699912 & Mexico, Mexico & 6     & 0.166666667 & 0.833333333 \\
-    RIMPA0119 & 699913 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMPA0120 & 699914 & Mexico, Mexico & 1     & 1     & 0 \\
-    RIMPA0121 & 699915 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMPA0128 & 699922 & Mexico, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0129 & 699923 & Michoacan, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0135 & 699929 & Nayarit, Mexico & 24    & 0     & 1 \\
-    RIMPA0138 & 699932 & Jalisco, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0139 & 699933 & Jalisco, Mexico & 1     & 1     & 0 \\
-    RIMPA0142 & 699936 & Colima, Mexico & 18    & 0.444444444 & 0.555555556 \\
-    RIMPA0144 & 699938 & Jalisco, Mexico & 2     & 1     & 0 \\
-    RIMPA0145 & 699939 & Michoacan, Mexico & 1     & 1     & 0 \\
-    RIMPA0147 & 699941 & Jalisco, Mexico & 1     & 1     & 0 \\
-    RIMPA0155 & N/A   & Jalisco, Mexico & 73    & 0.01369863 & 0.98630137 \\
-    RIMPA0156 & N/A   & Jalisco, Mexico & 20    & 0     & 1 \\
-    RIMPA0157 & N/A   & Jalisco, Mexico & 58    & 0.344827586 & 0.655172414 \\
-    RIMPA0158 & N/A   & Jalisco, Mexico & 64    & 0.53125 & 0.46875 \\
-    RIMPA0159 & N/A   & Jalisco, Mexico & 26    & 0     & 1 \\
-    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\\bottomrule
-    \end{tabular}
+    RIMPA0069 & 699862 & Ixtlan & 14    & 0.142857143 & 0.857142857 \\
+    RIMPA0070 & 699863 & Benito Jaurez & 16    & 0     & 1 \\
+    RIMPA0071 & 699864 & Tuzantla & 28    & 0     & 1 \\
+    RIMPA0072 & 699865 & Tiquicheo & 16    & 0     & 1 \\
+    RIMPA0073 & 699866 & Tiquicheo & 16    & 0.125 & 0.875 \\
+    RIMPA0074 & 699867 & Huetamo & 12    & 0     & 1 \\
+    RIMPA0075 & 699868 & Huetamo & 2     & 0     & 1 \\
+    RIMPA0076 & 699869 & Huetamo & 4     & 0     & 1 \\
+    RIMPA0077 & 699870 & Caracuaro & 2     & 0     & 1 \\
+    RIMPA0078 & 699871 & Caracuaro & 2     & 0.5   & 0.5 \\
+    RIMPA0079 & 699872 & Villa Madero & 14    & 0     & 1 \\
+    RIMPA0080 & 699873 & Guachinango & 12    & 0     & 1 \\
+    RIMPA0081 & 699875 & Ameca & 16    & 0     & 1 \\
+    RIMPA0083 & 699877 & Tepoztlan & 14    & 0     & 1 \\
+    RIMPA0084 & 699878 & Tepoztlan & 16    & 0     & 1 \\
+    RIMPA0085 & 699879 & Miahuatlan & 16    & 0     & 1 \\
+    RIMPA0086 & 699880 & Miahuatlan & 16    & 0.0625 & 0.9375 \\
+    RIMPA0087 & 699881 & Tecoanapa & 24    & 0     & 1 \\
+    RIMPA0089 & 699883 & Guerrero & 12    & 0     & 1 \\
+    RIMPA0090 & 699884 & Guerrero & 10    & 0     & 1 \\
+    RIMPA0091 & 699885 & Guerrero & 16    & 0     & 1 \\
+    RIMPA0092 & 699886 & Guerrero & 10    & 0     & 1 \\
+    RIMPA0093 & 699887 & Guerrero & 26    & 0.076923077 & 0.923076923 \\
+    RIMPA0094 & 699888 & Guerrero & 2     & 0     & 1 \\
+    RIMPA0095 & 699889 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0096 & 699890 & Guerrero & 26    & 0.038461538 & 0.961538462 \\
+    RIMPA0097 & 699891 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0098 & 699892 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0099 & 699893 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0100 & 699894 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0101 & 699895 & Guerrero & 2     & 0     & 1 \\
+    RIMPA0103 & 699897 & Guerrero & 2     & 0     & 1 \\
+    RIMPA0104 & 699898 & Guerrero & 22    & 0.090909091 & 0.909090909 \\
+    RIMPA0105 & 699899 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0106 & 699900 & Guerrero & 6     & 0.333333333 & 0.666666667 \\
+    RIMPA0107 & 699901 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0108 & 699902 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0109 & 699903 & Michoacan & 4     & 0.25  & 0.75 \\
+    RIMPA0110 & 699904 & Michoacan & 2     & 0     & 1 \\
+    RIMPA0111 & 699905 & Michoacan & 4     & 0     & 1 \\
+    RIMPA0112 & 699906 & Michoacan & 4     & 0.25  & 0.75 \\
+    RIMPA0114 & 699908 & Michoacan & 6     & 0.166666667 & 0.833333333 \\
+    RIMPA0116 & 699910 & Mexico & 2     & 0     & 1 \\
+    RIMPA0117 & 699911 & Mexico & 4     & 0     & 1 \\
+    RIMPA0118 & 699912 & Mexico & 6     & 0.166666667 & 0.833333333 \\
+    RIMPA0119 & 699913 & Mexico & 2     & 0     & 1 \\
+    RIMPA0120 & 699914 & Mexico & 1     & 1     & 0 \\
+    RIMPA0121 & 699915 & Mexico & 2     & 0     & 1 \\
+    RIMPA0128 & 699922 & Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0129 & 699923 & Michoacan & 2     & 0.5   & 0.5 \\
+    RIMPA0135 & 699929 & Nayarit & 24    & 0     & 1 \\
+    RIMPA0138 & 699932 & Jalisco & 2     & 0.5   & 0.5 \\
+    RIMPA0139 & 699933 & Jalisco & 1     & 1     & 0 \\
+    RIMPA0142 & 699936 & Colima & 18    & 0.444444444 & 0.555555556 \\
+    RIMPA0144 & 699938 & Jalisco & 2     & 1     & 0 \\
+    RIMPA0145 & 699939 & Michoacan & 1     & 1     & 0 \\
+    RIMPA0147 & 699941 & Jalisco & 1     & 1     & 0 \\
+    RIMPA0155 & N/A   & Jalisco & 73    & 0.01369863 & 0.98630137 \\
+    RIMPA0156 & N/A   & Jalisco & 20    & 0     & 1 \\
+    RIMPA0157 & N/A   & Jalisco & 58    & 0.344827586 & 0.655172414 \\
+    RIMPA0158 & N/A   & Jalisco & 64    & 0.53125 & 0.46875 \\
+    RIMPA0159 & N/A   & Jalisco & 26    & 0     & 1 \\
+    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
+\end{longtable}
  \label{TableS1}
-\end{table}
 
@@
 \begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+
   \centering
@@
 \end{table}%
+
+\end{document}

### 2014-05-26 manuscript.tex
edits
Citations are showing up as ?’s and not sure why

@@
 \subsection*{Acknowledgements}
-Place brief acknowledgments, if desired, as a separate paragraph, using the following style: “The author(s) thank(s)…”. For brevity, do not use first names. Include grant acknowledgments here.
-}
+We would like to thank the Department of Plant Sciences at UC Davis for Graduate Student Research funding to LEV, \lev{Matt UC Mexus}, and G. Coop for helpful discussion.
 
@@
 %check citations in last paragraph for ter1
-%reformat for AJB
+
 %preprint somewhere
@@
 
+\lev{So the citations are showing up as ?'s now}
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.

### 2014-05-26 manuscript.tex
edits
edits. weird things happening with references not working any more and
with lots of strange temp files in the repository. don’t know how to
fix yet.

@@
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1}.
-Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol (\url{http://www.ars.usda.gov}).
-\jri{what is this url?} We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
+Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol.
+We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
 Based on these initial data, we conducted a \emph{post hoc} power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
@@
 
-Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\alpha=0$ \mbh{what is the alpha value here?}; \lev{alpha is a parameter used in Box-Cox transformations} \mbh{the box-cox transformation function in the car package uses a lambda vector, but I do not see mention of alpha} Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
+Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\lambda=0$) Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
@@
 \ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) \jri{are these an average of pairwise or is this calculated among all pops?} \lev{should be calculated among all pops...I believe that is what HierFstat does} \jri{please check to be sure!} are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
@@
 For Region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}).
-\mbh{need to reference \emph{Hopscotch} frequencies in supplemental table somewhere} \lev{the frequencies for the Hopscotch in these populations are in the supplement.} \jri{please cite appropriate table.
-is this the last table?
-looks like it's missing column headings.} For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive in all populations except San Lorenzo, \mbh{is the table wrong?
-MSA is the only negative value in the table} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}, \ref{TableS2}).
+For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
+\lev{Do you want me to edit table here to have Hop freq as a column for the 4 pops?.
+I didn't know if it would be confusing because the frequencing is not representative of the frequencing of Hop in the sequenced sample set} Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch} \jri{table 1 shows 0.016 for hopscotch, not 0.

### 2014-05-26 manuscript.tex

@@
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
-\mbh{LV, please go through and make sure the data entered into all the tables is correct.
-In Table3, both sequenced regions were labeled as "Region 1".
-I changed the second to Region 2 but don't know if the data in this column are really from Region 2}
 
@@
 A) Ejutla A; B) Ejutla B; C) La Mesa; D).
-San Lorenzo}  \jri{this needs description of what upper and lower triangle are}
+San Lorenzo}.
+The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.
 \label{Fig2LD}
@@
 Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
-\jri{please explain whiskers and dots on figure too.}}
+Outliers are displayed as dots, the maximum value exluding outliers is shown with the top whisker, whicle the minimum exlucing outliers is shown with the bottom whisker.
 \label{Fig4Boxplots}
@@
 
-\jri{please check format of supp figs and tables; some are running off the page.
-you can use "longtable" to fix that (ask Paul for example).
-check fig/table references, bibliography, etc.
-what does "rotation" mean in supp.
-table 3?
-it isn't mentioned in methods.
-please check that all the tables and figs (including supplement) are referenced in the text.}
+\lev{I think we're done with text.
+What I am having trouble with is the pdf converstion -- was working fine up until a few hours ago, and I don't think I changed any of the '\usepackage' stuff.
+I've tried in both LaTexian and TexShop and neither works.
+Also the references have started showing up as question marks.}
 
@@
 \bibliography{refs,bob}
-
-%-------------------------------------------------
-Supplementary Materials
-%-------------------------------------------------
-\clearpage
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\renewcommand{\figurename}{Sup.\Fig.}
-\renewcommand{\tablename}{Sup.\Table}
-
-% Figure S1. Cartoon representation of tb1 locus
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
-  %\caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
-Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
-\label{FigS1Locus}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-% Figure S2. Gel image
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=150mm]{FigS2Gel.png}
-   % \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
-\label{FigS2Gel}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-% Figure S3. NJ Trees
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
-    %\caption{Neighbor-Joining tree of the sequenced region in the 5?
-UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}.
-Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).
-TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion.
-Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
-\label{FigS3NJ}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-% Table S1. Parv and mex samples
-\begin{table}[htbp]
-  \centering
-  %\caption{Accessions of \emph{Zea mays} ssp.
-\emph{mexicana} (RIMME) and \emph{Zea mays} ssp.
-\emph{parviglumis} (RIMPA) sampled.
-RIHY is a \emph{Z.
-mays} ssp.
-\emph{parviglumis} and \emph{Zea mays} ssp.
-\emph{mays} hybrid.
-}
-    \begin{tabular}{rrrrrr}
-%    \toprule
-    Accession & USDA Accession ID & Locality & Number alleles sampled & \emph{Hopscotch} Frequency & No \emph{Hopscotch} Frequency \\
- %   \midrule
-    RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
-    RIMME0006 & 566673 & Durango, Mexico & 2     & 0     & 1 \\
-    RIMME0007 & 566680 & Guanajuato, Mexico & 2     & 0     & 1 \\
-    RIMME0008 & 566681 & Michoacan, Mexico & 2     & 0     & 1 \\
-    RIMME0009 & 566682 & Distrito Federal, Mexico & 2     & 0     & 1 \\
-    RIMME0011 & 566685 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMME0014 & 714151 & Breeders line; Puga: 11066 & 6     & 0     & 1 \\
-    RIMME0017 & 699874 & Ayotlan, Mexico & 8     & 0     & 1 \\
-    RIMME0021 & N/A   & El Porvenir, Mexico & 69    & 0.173913043 & 0.826086957 \\
-    RIMME0026 & N/A   & Opopeo, Mexico & 42    & 0.071428571 & 0.928571429 \\
-    RIMME0028 & N/A   & Puruandiro, Mexico & 28    & 0.035714286 & 0.964285714 \\
-    RIMME0029 & N/A   & Ixtlan, Mexico & 35    & 0     & 1 \\
-    RIMME0030 & N/A   & San Pedro, Mexico & 27    & 0     & 1 \\
-    RIMME0031 & N/A   & Tenango del Aire, Mexico & 25    & 0.08  & 0.92 \\
-    RIMME0032 & N/A   & Nabogame, Mexico & 24    & 0     & 1 \\
-    RIMME0033 & N/A   & Puerta Encantada, Mexico & 25    & 0     & 1 \\
-    RIMME0034 & N/A   & Santa Clara, Mexico & 23    & 0     & 1 \\
-    RIMME0035 & N/A   & Xochimilco, Mexico & 25    & 0     & 1 \\
-    RIMPA0001 & 87168 & El Salado, Mexico & 4     & 0     & 1 \\
-    RIMPA0003 & 87171 & Mazatlan, Mexico & 8     & 0.125 & 0.875 \\
-    RIMPA0017 & 87200 & N/A   & 4     & 0     & 1 \\
-    RIMPA0019 & 87213 & El Salado, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0029 & 87244 & N/A   & 2     & 0.5   & 0.5 \\
-    RIMPA0031 & 87249 & N/A   & 2     & 0.5   & 0.5 \\
-    RIMPA0035 & 87288 & Jalisco, Mexico & 4     & 0     & 1 \\
-    RIMPA0040 & 288185 & Mexico, Mexico & 4     & 0     & 1 \\
-    RIMPA0042 & 288187 & Guerrero, Mexico & 4     & 0.25  & 0.75 \\
-    RIMPA0043 & 288188 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0045 & 288193 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0055 & 714152 & Breeders line & 2     & 0     & 1 \\
-    RIMPA0056 & 714153 & Breeders line & 2     & 0.5   & 0.5 \\
-    RIMPA0057 & 714154 & Breeders line & 2     & 0.5   & 0.5 \\
-    RIMPA0058 & N/A   & N/A   & 4     & 0.5   & 0.5 \\
-    RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
-    RIMPA0060 & 714157 & Breeders line: CIMMYT 11355 & 2     & 0     & 1 \\
-    RIMPA0061 & 714158 & Breeders line: USDA PI566686 & 4     & 0.5   & 0.5 \\
-    RIMPA0062 & 714159 & Breeders line & 4     & 0.5   & 0.5 \\
-    RIMPA0063 & 714160 & Breeders line & 4     & 0     & 1 \\
-    RIMPA0064 & 714161 & Breeders line & 3     & 0     & 1 \\
-    RIMPA0065 & 714162 & Breeders line & 4     & 0.25  & 0.75 \\
-    RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
-    RIMPA0069 & 699862 & Ixtlan, Mexico & 14    & 0.142857143 & 0.857142857 \\
-    RIMPA0070 & 699863 & Benito Jaurez, Mexico & 16    & 0     & 1 \\
-    RIMPA0071 & 699864 & Tuzantla, Mexico & 28    & 0     & 1 \\
-    RIMPA0072 & 699865 & Tiquicheo, Mexico & 16    & 0     & 1 \\
-    RIMPA0073 & 699866 & Tiquicheo, Mexico & 16    & 0.125 & 0.875 \\
-    RIMPA0074 & 699867 & Huetamo, Mexico & 12    & 0     & 1 \\
-    RIMPA0075 & 699868 & Huetamo, Mexico & 2     & 0     & 1 \\
-    RIMPA0076 & 699869 & Huetamo, Mexico & 4     & 0     & 1 \\
-    RIMPA0077 & 699870 & Caracuaro, Mexico & 2     & 0     & 1 \\
-    RIMPA0078 & 699871 & Caracuaro, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0079 & 699872 & Villa Madero, Mexico & 14    & 0     & 1 \\
-    RIMPA0080 & 699873 & Guachinango, Mexico & 12    & 0     & 1 \\
-    RIMPA0081 & 699875 & Ameca, Mexico & 16    & 0     & 1 \\
-    RIMPA0083 & 699877 & Tepoztlan, Mexico & 14    & 0     & 1 \\
-    RIMPA0084 & 699878 & Tepoztlan, Mexico & 16    & 0     & 1 \\
-    RIMPA0085 & 699879 & Miahuatlan, Mexico & 16    & 0     & 1 \\
-    RIMPA0086 & 699880 & Miahuatlan, Mexico & 16    & 0.0625 & 0.9375 \\
-    RIMPA0087 & 699881 & Tecoanapa, Mexico & 24    & 0     & 1 \\
-    RIMPA0089 & 699883 & Guerrero, Mexico & 12    & 0     & 1 \\
-    RIMPA0090 & 699884 & Guerrero, Mexico & 10    & 0     & 1 \\
-    RIMPA0091 & 699885 & Guerrero, Mexico & 16    & 0     & 1 \\
-    RIMPA0092 & 699886 & Guerrero, Mexico & 10    & 0     & 1 \\
-    RIMPA0093 & 699887 & Guerrero, Mexico & 26    & 0.076923077 & 0.923076923 \\
-    RIMPA0094 & 699888 & Guerrero, Mexico & 2     & 0     & 1 \\
-    RIMPA0095 & 699889 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0096 & 699890 & Guerrero, Mexico & 26    & 0.038461538 & 0.961538462 \\
-    RIMPA0097 & 699891 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0098 & 699892 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0099 & 699893 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0100 & 699894 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0101 & 699895 & Guerrero, Mexico & 2     & 0     & 1 \\
-    RIMPA0103 & 699897 & Guerrero, Mexico & 2     & 0     & 1 \\
-    RIMPA0104 & 699898 & Guerrero, Mexico & 22    & 0.090909091 & 0.909090909 \\
-    RIMPA0105 & 699899 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0106 & 699900 & Guerrero, Mexico & 6     & 0.333333333 & 0.666666667 \\
-    RIMPA0107 & 699901 & Guerrero, Mexico & 4     & 0     & 1 \\
-    RIMPA0108 & 699902 & Guerrero, Mexico & 6     & 0     & 1 \\
-    RIMPA0109 & 699903 & Michoacan, Mexico & 4     & 0.25  & 0.75 \\
-    RIMPA0110 & 699904 & Michoacan, Mexico & 2     & 0     & 1 \\
-    RIMPA0111 & 699905 & Michoacan, Mexico & 4     & 0     & 1 \\
-    RIMPA0112 & 699906 & Michoacan, Mexico & 4     & 0.25  & 0.75 \\
-    RIMPA0114 & 699908 & Michoacan, Mexico & 6     & 0.166666667 & 0.833333333 \\
-    RIMPA0116 & 699910 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMPA0117 & 699911 & Mexico, Mexico & 4     & 0     & 1 \\
-    RIMPA0118 & 699912 & Mexico, Mexico & 6     & 0.166666667 & 0.833333333 \\
-    RIMPA0119 & 699913 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMPA0120 & 699914 & Mexico, Mexico & 1     & 1     & 0 \\
-    RIMPA0121 & 699915 & Mexico, Mexico & 2     & 0     & 1 \\
-    RIMPA0128 & 699922 & Mexico, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0129 & 699923 & Michoacan, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0135 & 699929 & Nayarit, Mexico & 24    & 0     & 1 \\
-    RIMPA0138 & 699932 & Jalisco, Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0139 & 699933 & Jalisco, Mexico & 1     & 1     & 0 \\
-    RIMPA0142 & 699936 & Colima, Mexico & 18    & 0.444444444 & 0.555555556 \\
-    RIMPA0144 & 699938 & Jalisco, Mexico & 2     & 1     & 0 \\
-    RIMPA0145 & 699939 & Michoacan, Mexico & 1     & 1     & 0 \\
-    RIMPA0147 & 699941 & Jalisco, Mexico & 1     & 1     & 0 \\
-    RIMPA0155 & N/A   & Jalisco, Mexico & 73    & 0.01369863 & 0.98630137 \\
-    RIMPA0156 & N/A   & Jalisco, Mexico & 20    & 0     & 1 \\
-    RIMPA0157 & N/A   & Jalisco, Mexico & 58    & 0.344827586 & 0.655172414 \\
-    RIMPA0158 & N/A   & Jalisco, Mexico & 64    & 0.53125 & 0.46875 \\
-    RIMPA0159 & N/A   & Jalisco, Mexico & 26    & 0     & 1 \\
-    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
- %   \bottomrule
-    \end{tabular}
-  \label{TableS1}
-\end{table}
-
-% Table S2. Mays sampling
-\begin{table}[htbp]
-  \centering
- % \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
-    \begin{tabular}{ccc}
- %   \toprule
-    Accession & Number of alleles sampled & \emph{Hopscotch} Frequency \\
- %   \midrule
-    RIMMA0066 & 2     & 1 \\
-    RIMMA0075 & 2     & 1 \\
-    RIMMA0077 & 2     & 1 \\
-    RIMMA0079 & 2     & 1 \\
-    RIMMA0081 & 2     & 1 \\
-    RIMMA0084 & 2     & 1 \\
-    RIMMA0086 & 2     & 1 \\
-    RIMMA0088 & 2     & 1 \\
-    RIMMA0089 & 2     & 1 \\
-    RIMMA0090 & 2     & 1 \\
-    RIMMA0092 & 4     & 1 \\
-    RIMMA0094 & 4     & 1 \\
-    RIMMA0097 & 2     & 1 \\
-    RIMMA0099 & 2     & 1 \\
-    RIMMA0100 & 2     & 1 \\
-    RIMMA0101 & 2     & 1 \\
-    RIMMA0104 & 2     & 1 \\
-    RIMMA0108 & 2     & 1 \\
-    RIMMA0111 & 6     & 1 \\
-    RIMMA0115 & 2     & 1 \\
-    RIMMA0117 & 2     & 1 \\
-    RIMMA0130 & 2     & 1 \\
-    RIMMA0133 & 2     & 1 \\
-    RIMMA0134 & 2     & 1 \\
-    RIMMA0135 & 2     & 1 \\
-    RIMMA0142 & 2     & 0.5 \\
-    RIMMA0143 & 4     & 1 \\
-    RIMMA0146 & 4     & 1 \\
-    RIMMA0149 & 2     & 1 \\
-    RIMMA0152 & 2     & 1 \\
-    RIMMA0153 & 2     & 1 \\
-    RIMMA0154 & 2     & 1 \\
-    RIMMA0155 & 2     & 1 \\
-    RIMMA0156 & 2     & 1 \\
-    RIMMA0157 & 2     & 1 \\
-    RIMMA0158 & 2     & 1 \\
-    RIMMA0159 & 2     & 1 \\
-    RIMMA0160 & 2     & 1 \\
-    RIMMA0162 & 2     & 1 \\
-    RIMMA0166 & 2     & 1 \\
-    RIMMA0167 & 2     & 1 \\
-    RIMMA0168 & 2     & 1 \\
-    RIMMA0169 & 2     & 1 \\
-    RIMMA0172 & 2     & 1 \\
-    RIMMA0174 & 4     & 1 \\
-    RIMMA0177 & 2     & 1 \\
-    RIMMA0178 & 2     & 1 \\
-    RIMMA0179 & 2     & 1 \\
-    RIMMA0181 & 2     & 1 \\
-    RIMMA0183 & 2     & 1 \\
-    RIMMA0184 & 2     & 1 \\
-    RIMMA0186 & 2     & 1 \\
-    RIMMA0187 & 1     & 1 \\
-    RIMMA0188 & 2     & 1 \\
-    RIMMA0195 & 2     & 1 \\
-    RIMMA0196 & 2     & 1 \\
-    RIMMA0197 & 2     & 1 \\
-    RIMMA0198 & 2     & 1 \\
-    RIMMA0199 & 2     & 1 \\
-    RIMMA0200 & 2     & 1 \\
-    RIMMA0202 & 2     & 1 \\
-    RIMMA0203 & 2     & 1 \\
-    RIMMA0206 & 2     & 1 \\
-    RIMMA0208 & 2     & 1 \\
-    RIMMA0209 & 2     & 1 \\
-    RIMMA0210 & 2     & 1 \\
-    RIMMA0212 & 2     & 1 \\
-    RIMMA0213 & 2     & 1 \\
-    RIMMA0214 & 2     & 1 \\
-    RIMMA0217 & 2     & 1 \\
-    RIMMA0218 & 2     & 1 \\
-    RIMMA0220 & 2     & 1 \\
-    RIMMA0221 & 2     & 1 \\
-    RIMMA0222 & 2     & 1 \\
-    RIMMA0223 & 2     & 1 \\
-    RIMMA0226 & 2     & 1 \\
-    RIMMA0227 & 2     & 1 \\
-    RIMMA0228 & 2     & 1 \\
-    RIMMA0229 & 2     & 1 \\
-    RIMMA0230 & 2     & 1 \\
-    RIMMA0232 & 2     & 1 \\
-    RIMMA0233 & 2     & 1 \\
-    RIMMA0235 & 2     & 0.5 \\
-    RIMMA0242 & 2     & 1 \\
-    RIMMA0243 & 2     & 1 \\
-    RIMMA0247 & 4     & 1 \\
-    RIMMA0248 & 2     & 1 \\
-    RIMMA0249 & 2     & 1 \\
-    RIMMA0252 & 2     & 1 \\
-    RIMMA0253 & 2     & 1 \\
-    RIMMA0254 & 2     & 1 \\
-    RIMMA0256 & 2     & 1 \\
-    RIMMA0257 & 2     & 1 \\
-    RIMMA0258 & 2     & 1 \\
-    RIMMA0259 & 2     & 1 \\
-    RIMMA0260 & 2     & 1 \\
-    RIMMA0262 & 2     & 1 \\
-    RIMMA0263 & 2     & 1 \\
-    RIMMA0264 & 2     & 1 \\
-    RIMMA0265 & 2     & 1 \\
-    RIMMA0268 & 2     & 1 \\
-    RIMMA0269 & 2     & 1 \\
-    RIMMA0270 & 2     & 1 \\
-    RIMMA0272 & 2     & 1 \\
-    RIMMA0275 & 2     & 1 \\
-    RIMMA0276 & 2     & 1 \\
-    RIMMA0277 & 2     & 1 \\
-    RIMMA0279 & 2     & 1 \\
-    RIMMA0280 & 2     & 1 \\
-    RIMMA0283 & 2     & 1 \\
-    RIMMA0285 & 2     & 1 \\
-    RIMMA0288 & 2     & 1 \\
-    RIMMA0290 & 2     & 1 \\
-    RIMMA0291 & 2     & 1 \\
-    RIMMA0292 & 2     & 1 \\
-    RIMMA0293 & 2     & 1 \\
-    RIMMA0298 & 2     & 1 \\
-    RIMMA0302 & 2     & 1 \\
-    RIMMA0305 & 2     & 1 \\
-    RIMMA0310 & 2     & 1 \\
-    RIMMA0312 & 2     & 1 \\
-    RIMMA0320 & 2     & 1 \\
-    RIMMA0322 & 2     & 1 \\
-    RIMMA0324 & 2     & 1 \\
-    RIMMA0334 & 2     & 1 \\
-    RIMMA0336 & 2     & 1 \\
-    RIMMA0337 & 2     & 1 \\
-    RIMMA0338 & 2     & 1 \\
-    RIMMA0339 & 2     & 1 \\
-    RIMMA0340 & 2     & 1 \\
-    RIMMA0341 & 2     & 1 \\
-    RIMMA0342 & 2     & 1 \\
-    RIMMA0344 & 2     & 1 \\
-    RIMMA0346 & 2     & 1 \\
-    RIMMA0347 & 2     & 1 \\
-    RIMMA0350 & 2     & 1 \\
-    RIMMA0351 & 2     & 1 \\
-    RIMMA0352 & 2     & 1 \\
-    RIMMA0357 & 2     & 1 \\
-    RIMMA0360 & 18    & 1 \\
-    RIMMA0362 & 8     & 1 \\
-    RIMMA0363 & 16    & 1 \\
-    RIMMA0370 & 17    & 1 \\
-    RIMMA0372 & 24    & 1 \\
-    RIMMA0373 & 22    & 1 \\
-    RIMMA0408 & 2     & 1 \\
-    RIMMA0411 & 1     & 1 \\
-    RIMMA0413 & 1     & 1 \\
-    RIMMA0414 & 1     & 1 \\
-    RIMMA0424 & 8     & 0.9 \\
-    RIMMA0432 & 1     & 1 \\
-    RIMMA0434 & 2     & 0.5 \\
-    RIMMA0435 & 1     & 1 \\
-    RIMMA0442 & 2     & 1 \\
-    RIMMA0443 & 2     & 1 \\
-    RIMMA0444 & 2     & 1 \\
-    RIMMA0445 & 2     & 1 \\
-    RIMMA0446 & 2     & 1 \\
-    RIMMA0447 & 2     & 1 \\
-    RIMMA0448 & 2     & 1 \\
-    RIMMA0449 & 2     & 1 \\
-    RIMMA0450 & 2     & 1 \\
-    RIMMA0451 & 2     & 1 \\
-    RIMMA0452 & 2     & 1 \\
-    RIMMA0453 & 2     & 1 \\
-    RIMMA0454 & 2     & 1 \\
-    RIMMA0455 & 2     & 1 \\
-    RIMMA0456 & 2     & 1 \\
-    RIMMA0457 & 2     & 1 \\
-    RIMMA0458 & 2     & 1 \\
-    RIMMA0459 & 2     & 1 \\
-    RIMMA0483 & 2     & 1 \\
-    RIMMA0490 & 2     & 1 \\
-    RIMMA0515 & 2     & 1 \\
-    RIMMA0537 & 2     & 1 \\
-    RIMMA0550 & 2     & 1 \\
-    RIMMA0553 & 2     & 1 \\
-    RIMMA0559 & 2     & 1 \\
-    RIMMA0561 & 2     & 1 \\
-    RIMMA0562 & 2     & 1 \\
-    RIMMA0571 & 2     & 1 \\
-    RIMMA0572 & 2     & 1 \\
-    RIMMA0577 & 2     & 1 \\
-    RIMMA0579 & 2     & 1 \\
-    RIMMA0590 & 2     & 1 \\
-    RIMMA0591 & 2     & 1 \\
-    RIMMA0592 & 2     & 1 \\
-    RIMMA0593 & 2     & 1 \\
-    RIMMA0594 & 2     & 1 \\
-    RIMMA0595 & 2     & 1 \\
-    RIMMA0596 & 2     & 1 \\
-    RIMMA0597 & 2     & 1 \\
-    RIMMA0598 & 2     & 1 \\
-    RIMMA0599 & 2     & 1 \\
-    RIMMA0600 & 2     & 1 \\
-    RIMMA0601 & 2     & 1 \\
-    RIMMA0602 & 2     & 1 \\
-    RIMMA0603 & 2     & 1 \\
-    RIMMA0604 & 2     & 1 \\
-    RIMMA0605 & 2     & 1 \\
-    RIMMA0606 & 2     & 1 \\
-    RIMMA0607 & 2     & 1 \\
-    RIMMA0608 & 2     & 1 \\
-    RIMMA0609 & 2     & 1 \\
-    RIMMA0610 & 2     & 1 \\
-    RIMMA0611 & 2     & 1 \\
-    RIMMA0612 & 2     & 1 \\
-    RIMMA0613 & 2     & 1 \\
-    RIMMA0622 & 2     & 0.5 \\
-    RIMMA0624 & 1     & 1 \\
-    RIMMA0629 & 2     & 0.5 \\
-    RIMMA0631 & 2     & 0.5 \\
-    RIMMA0659 & 2     & 1 \\
-    RIMMA0660 & 2     & 1 \\
-    RIMMA0678 & 1     & 1 \\
-    RIMMA0681 & 2     & 1 \\
-    RIMMA0683 & 2     & 1 \\
-    RIMMA0684 & 2     & 1 \\
-    RIMMA0685 & 4     & 1 \\
-    RIMMA0693 & 2     & 1 \\
-    RIMMA0694 & 2     & 1 \\
-    RIMMA0695 & 1     & 1 \\
-    RIMMA0699 & 2     & 1 \\
-    RIMMA0704 & 2     & 1 \\
-    RIMMA0706 & 2     & 1 \\
-    RIMMA0711 & 1     & 1 \\
-    RIMMA0713 & 2     & 1 \\
-    RIMMA0715 & 2     & 1 \\
-    RIMMA0717 & 2     & 1 \\
-    RIMMA0718 & 2     & 1 \\
-    RIMMA0723 & 2     & 1 \\
-    RIMMA0724 & 2     & 1 \\
-    RIMMA0725 & 1     & 1 \\
-    RIMMA0728 & 1     & 1 \\
-    RIMMA0732 & 2     & 1 \\
-    RIMMA0734 & 2     & 1 \\
-    RIMMA0738 & 2     & 1 \\
-    RIMMA0739 & 2     & 1 \\
-    RIMMA0744 & 2     & 1 \\
-    RIMMA0747 & 1     & 1 \\
-    RIMMA0748 & 2     & 1 \\
-    RIMMA0749 & 2     & 1 \\
-    RIMMA0750 & 2     & 0.5 \\
-    RIMMA0751 & 1     & 1 \\
-    RIMMA0752 & 1     & 1 \\
-    RIMMA0753 & 2     & 1 \\
-    RIMMA0755 & 2     & 1 \\
-    \bottomrule
-    \end{tabular}
-  \label{TableS2Mays}
-\end{table}
-
-% Table S3. PCs for BayEnv
-\begin{table}[htbp]
-  \centering
-  %\caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}
-    \begin{tabular}{rrrrrrrrrrrr}
-    \toprule
-    \textbf{PC1} & \textbf{} & \textbf{PC2} & \textbf{} & \textbf{PC3} & \textbf{} & \textbf{PC4} & \textbf{} & \textbf{PC5} & \textbf{} & \textbf{PC6} & \textbf{} \\
-    \midrule
-    \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} \\
-    bio1  & 0.146 & bio4  & 0.244 & prec7 & 0.287 & ts\_clay & 0.41  & bio2  & 0.38  & bio2  & 0.412 \\
-    tmean11 & 0.146 & bio3  & 0.241 & prec8 & 0.276 & v\_mod & 0.359 & sq4   & 0.328 & x\_mod & 0.365 \\
-    tmean12 & 0.145 & bio7  & 0.241 & prec11 & 0.262 & ts\_sand & 0.329 & ts\_loam & 0.289 & sq7   & 0.332 \\
-    bio11 & 0.145 & prec6 & 0.237 & bio13 & 0.247 & bio15 & 0.272 & ts\_sand & 0.266 & bio7  & 0.312 \\
-    tmax12 & 0.145 & sq7   & 0.218 & prec1 & 0.246 & prec4 & 0.259 & sq7   & 0.231 & v\_mod & 0.307 \\
-    tmin5 & 0.145 & prec9 & 0.217 & bio16 & 0.242 & x\_mod & 0.244 & bio18 & 0.213 & prec11 & 0.23 \\
-    tmean1 & 0.145 & sq3   & 0.207 & prec12 & 0.24  & prec3 & 0.226 & bio13 & 0.207 & bio18 & 0.22 \\
-    tmean2 & 0.145 & prec12 & 0.207 & bio19 & 0.238 & sq3   & 0.21  & prec11 & 0.183 & sq3   & 0.176 \\
-    tmin4 & 0.145 & bio12 & 0.204 & bio12 & 0.231 & prec5 & 0.21  & bio7  & 0.17  & sq4   & 0.153 \\
-    tmax1 & 0.145 & bio19 & 0.196 & prec2 & 0.222 & prec7 & 0.19  & bio16 & 0.163 & ts\_sand & 0.153 \\
-    tmean4 & 0.145 & prec2 & 0.188 & bio18 & 0.221 & sq4   & 0.186 & bio4  & 0.157 & bio4  & 0.148 \\
-    tmin11 & 0.144 & prec1 & 0.185 & sq4   & 0.2   & bio3  & 0.185 & bio12 & 0.156 & prec7 & 0.139 \\
-    tmax11 & 0.144 & prec10 & 0.184 & prec9 & 0.18  & bio18 & 0.178 & bio3  & 0.155 & tmax3 & 0.127 \\
-    tmin12 & 0.144 & bio16 & 0.183 & prec10 & 0.171 & sq7   & 0.132 & prec6 & 0.154 & tmax4 & 0.126 \\
-    tmin2 & 0.144 & prec8 & 0.17  & prec5 & 0.161 & bio14 & 0.116 & x\_mod & 0.152 & bio5  & 0.118 \\
-    tmean5 & 0.144 & prec5 & 0.165 & prec4 & 0.154 & bio13 & 0.099 & prec9 & 0.144 & tmax5 & 0.115 \\
-    tmean10 & 0.144 & bio14 & 0.158 & sq3   & 0.147 & bio16 & 0.095 & prec8 & 0.143 & bio3  & 0.107 \\
-    bio6  & 0.144 & bio13 & 0.151 & bio2  & 0.143 & prec8 & 0.09  & v\_mod & 0.142 & ts\_loam & 0.095 \\
-    tmax2 & 0.144 & bio17 & 0.149 & bio17 & 0.129 & bio7  & 0.077 & bio15 & 0.136 & ts\_clay & 0.078 \\
-    tmean3 & 0.144 & prec3 & 0.144 & ts\_loam & 0.127 & bio4  & 0.075 & prec7 & 0.112 & tmin9 & 0.072 \\
-    tmin1 & 0.143 & ts\_clay & 0.141 & v\_mod & 0.123 & bio2  & 0.074 & prec4 & 0.108 & tmin8 & 0.071 \\
-    tmin10 & 0.143 & bio2  & 0.129 & prec3 & 0.113 & prec2 & 0.074 & bio14 & 0.096 & prec9 & 0.07 \\
-    Altitude & 0.143 & prec7 & 0.108 & x\_mod & 0.111 & bio19 & 0.068 & tmax7 & 0.093 & tmin10 & 0.069 \\
-    bio9  & 0.143 & tmax6 & 0.107 & bio14 & 0.099 & prec12 & 0.056 & tmax8 & 0.092 & tmin12 & 0.069 \\
-    tmin3 & 0.143 & x\_mod & 0.106 & bio4  & 0.07  & ts\_loam & 0.053 & prec1 & 0.091 & tmin7 & 0.066 \\
-    bio10 & 0.142 & bio15 & 0.098 & tmax3 & 0.067 & tmax12 & 0.047 & prec2 & 0.086 & tmean4 & 0.066 \\
-    tmax10 & 0.142 & ts\_loam & 0.088 & ts\_clay & 0.065 & bio17 & 0.047 & tmin11 & 0.086 & tmax2 & 0.066 \\
-    tmax3 & 0.142 & tmean6 & 0.085 & bio15 & 0.056 & bio9  & 0.043 & prec5 & 0.082 & tmax6 & 0.064 \\
-    tmax4 & 0.142 & tmin7 & 0.082 & tmax2 & 0.055 & tmax8 & 0.042 & bio17 & 0.082 & tmean3 & 0.064 \\
-    tmin6 & 0.142 & bio5  & 0.082 & tmean3 & 0.052 & tmax1 & 0.041 & tmin12 & 0.08  & bio9  & 0.058 \\
-    tmean9 & 0.141 & tmean7 & 0.081 & ts\_sand & 0.05  & tmax5 & 0.039 & prec3 & 0.078 & tmin11 & 0.058 \\
-    tmin9 & 0.141 & prec4 & 0.08  & prec6 & 0.048 & tmax7 & 0.039 & tmax9 & 0.078 & prec3 & 0.056 \\
-    tmean8 & 0.141 & tmax7 & 0.079 & sq7   & 0.048 & prec10 & 0.038 & tmin1 & 0.077 & tmean5 & 0.056 \\
-    bio8  & 0.14  & bio8  & 0.079 & tmin7 & 0.046 & Altitude & 0.037 & tmin10 & 0.074 & bio10 & 0.051 \\
-    tmean6 & 0.14  & tmax9 & 0.077 & bio3  & 0.044 & tmax10 & 0.035 & bio6  & 0.071 & bio15 & 0.047 \\
-    tmean7 & 0.14  & tmean8 & 0.076 & tmax4 & 0.043 & tmax2 & 0.033 & prec12 & 0.067 & tmin1 & 0.045 \\
-    tmin8 & 0.14  & tmin8 & 0.076 & bio7  & 0.042 & tmax9 & 0.03  & tmin2 & 0.061 & prec12 & 0.045 \\
-    tmax5 & 0.14  & tmax5 & 0.074 & tmax1 & 0.036 & tmean12 & 0.029 & tmin6 & 0.061 & bio17 & 0.045 \\
-    tmax9 & 0.139 & tmax8 & 0.074 & tmin3 & 0.035 & tmax11 & 0.027 & tmax6 & 0.059 & tmean10 & 0.043 \\
-    tmax8 & 0.139 & tmean9 & 0.072 & bio9  & 0.035 & tmean1 & 0.027 & tmin3 & 0.052 & prec4 & 0.043 \\
-    bio5  & 0.139 & bio18 & 0.07  & tmin8 & 0.034 & bio5  & 0.026 & bio8  & 0.046 & Altitude & 0.041 \\
-    tmax7 & 0.139 & v\_mod & 0.066 & tmean4 & 0.031 & tmean5 & 0.026 & tmean11 & 0.041 & bio6  & 0.039 \\
-    tmin7 & 0.138 & tmin9 & 0.066 & tmean2 & 0.031 & bio11 & 0.025 & tmean1 & 0.04  & tmin6 & 0.038 \\
-    tmax6 & 0.135 & bio10 & 0.065 & tmax12 & 0.03  & tmean2 & 0.022 & tmin9 & 0.04  & tmean12 & 0.035 \\
-    bio17 & 0.107 & tmax10 & 0.063 & tmean7 & 0.028 & tmin6 & 0.022 & tmean12 & 0.04  & tmean9 & 0.035 \\
-    bio14 & 0.095 & tmin6 & 0.061 & tmin9 & 0.023 & prec1 & 0.022 & tmax10 & 0.038 & prec5 & 0.033 \\
-    prec3 & 0.094 & prec11 & 0.056 & Altitude & 0.021 & tmean7 & 0.02  & tmin5 & 0.038 & tmean8 & 0.03 \\
-    bio15 & 0.092 & ts\_sand & 0.054 & tmax11 & 0.019 & tmean8 & 0.02  & bio11 & 0.037 & tmin2 & 0.03 \\
-    prec4 & 0.088 & sq4   & 0.049 & tmin4 & 0.018 & tmean9 & 0.019 & tmean8 & 0.035 & tmax1 & 0.03 \\
-    prec10 & 0.084 & tmin3 & 0.048 & tmax5 & 0.017 & prec11 & 0.019 & tmean7 & 0.033 & prec1 & 0.025 \\
-    ts\_sand & 0.073 & tmean10 & 0.047 & tmin10 & 0.017 & tmean11 & 0.018 & tmean2 & 0.03  & bio13 & 0.024 \\
-    ts\_loam & 0.072 & tmean5 & 0.044 & tmin6 & 0.017 & tmax4 & 0.016 & bio19 & 0.029 & tmean11 & 0.023 \\
-    prec6 & 0.069 & tmax4 & 0.042 & tmean6 & 0.017 & tmean10 & 0.016 & bio9  & 0.027 & tmax7 & 0.023 \\
-    prec9 & 0.063 & tmax11 & 0.041 & tmax6 & 0.017 & tmin5 & 0.015 & tmin7 & 0.026 & tmean7 & 0.023 \\
-    prec2 & 0.063 & bio6  & 0.038 & tmin12 & 0.017 & tmean6 & 0.015 & tmin4 & 0.025 & prec10 & 0.022 \\
-    bio4  & 0.06  & tmin1 & 0.038 & bio11 & 0.016 & prec6 & 0.014 & tmean5 & 0.023 & prec2 & 0.02 \\
-    prec5 & 0.058 & tmin2 & 0.037 & tmin11 & 0.015 & bio6  & 0.012 & tmean3 & 0.022 & bio8  & 0.018 \\
-    bio13 & 0.055 & Altitude & 0.036 & tmin1 & 0.013 & tmin1 & 0.012 & tmin8 & 0.022 & tmean2 & 0.017 \\
-    x\_mod & 0.053 & tmin10 & 0.034 & tmean5 & 0.013 & tmin12 & 0.011 & tmean10 & 0.021 & bio16 & 0.016 \\
-    prec1 & 0.05  & bio9  & 0.032 & tmean8 & 0.013 & tmin2 & 0.011 & tmean9 & 0.019 & tmean6 & 0.014 \\
-    bio16 & 0.049 & tmean1 & 0.029 & bio6  & 0.012 & bio1  & 0.011 & tmean4 & 0.019 & bio19 & 0.012 \\
-    bio7  & 0.043 & bio1  & 0.028 & tmean9 & 0.01  & tmin4 & 0.009 & ts\_clay & 0.018 & tmax11 & 0.012 \\
-    bio19 & 0.042 & tmean3 & 0.026 & tmean1 & 0.01  & tmax6 & 0.008 & prec10 & 0.015 & tmax10 & 0.011 \\
-    bio12 & 0.042 & tmean2 & 0.024 & bio10 & 0.01  & tmin11 & 0.007 & bio1  & 0.014 & tmax8 & 0.01 \\
-    bio3  & 0.04  & bio11 & 0.022 & tmax8 & 0.009 & bio8  & 0.007 & tmax4 & 0.014 & tmean1 & 0.009 \\
-    prec7 & 0.034 & tmean11 & 0.019 & tmax7 & 0.009 & bio10 & 0.007 & bio10 & 0.011 & bio14 & 0.009 \\
-    prec12 & 0.033 & tmin12 & 0.019 & bio5  & 0.008 & tmin9 & 0.006 & tmax5 & 0.009 & prec8 & 0.007 \\
-    bio2  & 0.033 & tmax1 & 0.017 & tmin5 & 0.008 & prec9 & 0.005 & bio5  & 0.008 & bio11 & 0.007 \\
-    prec8 & 0.032 & tmean4 & 0.016 & tmin2 & 0.007 & bio12 & 0.004 & tmax11 & 0.008 & tmin5 & 0.006 \\
-    sq4   & 0.026 & tmin5 & 0.013 & bio1  & 0.007 & tmax3 & 0.003 & tmax12 & 0.008 & tmax9 & 0.005 \\
-    sq3   & 0.024 & tmean12 & 0.012 & tmax10 & 0.006 & tmean4 & 0.003 & tmax3 & 0.007 & bio12 & 0.004 \\
-    v\_mod & 0.015 & tmax2 & 0.011 & tmean10 & 0.006 & tmin3 & 0.002 & sq3   & 0.007 & tmin3 & 0.004 \\
-    ts\_clay & 0.012 & tmin4 & 0.01  & tmean12 & 0.005 & tmin7 & 0.002 & tmax2 & 0.006 & prec6 & 0.004 \\
-    sq7   & 0.004 & tmax3 & 0.003 & bio8  & 0.004 & tmean3 & 0.002 & Altitude & 0.004 & tmin4 & 0.003 \\
-    prec11 & 0.004 & tmax12 & 0.002 & tmax9 & 0.003 & tmin10 & 0.002 & tmean6 & 0.002 & bio1  & 0.002 \\
-    bio18 & 0.001 & tmin11 & 0.001 & tmean11 & 0.001 & tmin8 & 0     & tmax1 & 0.001 & tmax12 & 0.002 \\
-    \textbf{Bayes Factors} &       &       &       &       &       &       &       &       &       &       &  \\
-    \multicolumn{2}{c}{0.77165} & \multicolumn{2}{c}{0.51652} & \multicolumn{2}{c}{0.21120} & \multicolumn{2}{c}{0.24063} & \multicolumn{2}{c}{1.96010} & \multicolumn{2}{c}{0.24594} \\
-    \bottomrule
-    \end{tabular}%
-  \label{TableS3BayEnv}%
-\end{table}%
-
-% Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
-\begin{table}[htbp]
-  \centering
-  %\caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
-    \begin{tabular}{ccccc}
-    \toprule
-    %\textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
-    \midrule
-    Ejutla A & 4     & 0.15217 & 0.11902 & 0.76191 \\
-    Ejutla B & 5     & 0.15258 & 0.14877 & 0.07412 \\
-    La Mesa & 3     & 0.12802 & 0.08926 & 1.09209 \\
-    San Lorenzo & 3     & 0.09098 & 0.08926 & 0.04845 \\
-    \bottomrule
-    \end{tabular}%
-  \label{TableS4DiversitySNP}%
-\end{table}%
 

### 2014-05-26 manuscript.tex
fixed compiling errors

@@
 \usepackage{natbib}
+\bibliographystyle{AJB}
 \usepackage{authblk}
@@
 \renewcommand{\baselinestretch}{1.5}
-\bibliographystyle{AJB}
 \renewcommand{\refname}{\centering{LITERATURE CITED}}
@@
 \clearpage
-\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted\line(1,0){120}.}
+\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted \line(1,0){120}.}
 \vspace*{1cm}
@@
 acknowledge graham coop for helpful discussion}
+}
 \clearpage
@@
 
-\lev{So the citations are showing up as ?'s now}
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
@@
 \begin{SCfigure}[][t!]
-  \centering \label{Fig1Map}
-   \includegraphics[width=0.7\textwidth]{Fig1Map.pdf}
+   \includegraphics[width=0.6\textwidth]{Fig1Map.pdf}
     \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
@@
 }
+     \label{Fig1Map}
 \end{SCfigure}
@@
 For Region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}, \ref{TableS2}).
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}, Table~S2).
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
@@
 \ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
-In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$) (\ref{Table3R2}).
+In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).
 
@@
    \includegraphics[width=140mm]{Fig2LDPlots.pdf}
+   \end{center}
     \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
 The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
-A) Ejutla A; B) Ejutla B; C) La Mesa; D).
-San Lorenzo}.
-The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.
+A) Ejutla A; B) Ejutla B; C) La Mesa; D) San Lorenzo.
+The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.}
 \label{Fig2LD}
-  \end{center}
 \end{figure*}
@@
    \includegraphics[width=90mm]{Fig3Structure.pdf}
-    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF. }
+     \end{center}
+    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF.}
   \label{Fig3Structure}
-  \end{center}
 \end{figure*}
@@
 We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no correlation between genotype at the \emph{Hopscotch}  insertion and tillering index (\ref{Fig4Boxplots}), tiller number, or culm diameter.
+After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no correlation between genotype at the \emph{Hopscotch}  insertion and tillering index (Fig.
+\ref{Fig4Boxplots}), tiller number, or culm diameter.
 
@@
   \begin{center}
-   \includegraphics[width=95mm]{Fig4Boxplots.pdf}
+   \includegraphics[width=90mm]{Fig4Boxplots.pdf}
+   \end{center}
     \caption{Box-plots showing tillering index in greenhouse grow-outs for Phenotyping 1 (A) and Phenotyping 2 (B).
@@
 Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
-Outliers are displayed as dots, the maximum value exluding outliers is shown with the top whisker, whicle the minimum exlucing outliers is shown with the bottom whisker.
+Outliers are displayed as dots, the maximum value excluding outliers is shown with the top whisker, while the minimum excluding outliers is shown with the bottom whisker.}
 \label{Fig4Boxplots}
-  \end{center}
 \end{figure*}
@@
 
-\lev{I think we're done with text.
-What I am having trouble with is the pdf converstion -- was working fine up until a few hours ago, and I don't think I changed any of the '\usepackage' stuff.
-I've tried in both LaTexian and TexShop and neither works.
-Also the references have started showing up as question marks.}
+\clearpage
 
-\clearpage
-\bibliography{refs,bob}
+\bibliography{refs}
 

### 2014-05-27 Supplement.tex
edits
reformatting tables for \longtable
edits
working on table formatting

@@
 
-% Table S1. Parv and mex samples 
-\setlength\LTleft{-30pt}            
-\setlength\LTright{-30pt}
-\begin{longtable}{ccp{5cm}cp{5cm}ccc}
+% Table S2. Parv and mex samples 
+\footnotesize  % Switch from 12pt to 11pt; otherwise, table won't fit
+\setlength\LTleft{-30pt}            % default: \fill
+\setlength\LTright{-30pt}           % default: \fill
+\begin{longtable}{@{\extracolsep{\fill}}llllllllll@{}}
+  \hline\hline %inserts double horizontal lines
 \caption{Accessions of \emph{Zea mays} ssp.
@@
 \hline
- \textbf{Accession} & \textbf{USDA ID} & \textbf{Locality} & \textbf{Number Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\\toprule
+ \textbf{Accession} & \textbf{USDA ID} & \textbf{Locality} & \textbf{\# Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\\toprule
 \endhead
@@
     RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
-    RIMPA0069 & 699862 & Ixtlan & 14    & 0.142857143 & 0.857142857 \\
+    RIMPA0069 & 699862 & Ixtlan & 14    & 0.14 & 0.86 \\
     RIMPA0070 & 699863 & Benito Jaurez & 16    & 0     & 1 \\
@@
     RIMPA0072 & 699865 & Tiquicheo & 16    & 0     & 1 \\
-    RIMPA0073 & 699866 & Tiquicheo & 16    & 0.125 & 0.875 \\
+    RIMPA0073 & 699866 & Tiquicheo & 16    & 0.12 & 0.88 \\
     RIMPA0074 & 699867 & Huetamo & 12    & 0     & 1 \\
@@
     RIMPA0085 & 699879 & Miahuatlan & 16    & 0     & 1 \\
-    RIMPA0086 & 699880 & Miahuatlan & 16    & 0.0625 & 0.9375 \\
+    RIMPA0086 & 699880 & Miahuatlan & 16    & 0.06 & 0.94 \\
     RIMPA0087 & 699881 & Tecoanapa & 24    & 0     & 1 \\
@@
     RIMPA0092 & 699886 & Guerrero & 10    & 0     & 1 \\
-    RIMPA0093 & 699887 & Guerrero & 26    & 0.076923077 & 0.923076923 \\
+    RIMPA0093 & 699887 & Guerrero & 26    & 0.08 & 0.92 \\
     RIMPA0094 & 699888 & Guerrero & 2     & 0     & 1 \\
     RIMPA0095 & 699889 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0096 & 699890 & Guerrero & 26    & 0.038461538 & 0.961538462 \\
+    RIMPA0096 & 699890 & Guerrero & 26    & 0.04 & 0.96 \\
     RIMPA0097 & 699891 & Guerrero & 6     & 0     & 1 \\
@@
     RIMPA0103 & 699897 & Guerrero & 2     & 0     & 1 \\
-    RIMPA0104 & 699898 & Guerrero & 22    & 0.090909091 & 0.909090909 \\
+    RIMPA0104 & 699898 & Guerrero & 22    & 0.09 & 0.91 \\
     RIMPA0105 & 699899 & Guerrero & 6     & 0     & 1 \\
-    RIMPA0106 & 699900 & Guerrero & 6     & 0.333333333 & 0.666666667 \\
+    RIMPA0106 & 699900 & Guerrero & 6     & 0.33 & 0.67 \\
     RIMPA0107 & 699901 & Guerrero & 4     & 0     & 1 \\
@@
     RIMPA0112 & 699906 & Michoacan & 4     & 0.25  & 0.75 \\
-    RIMPA0114 & 699908 & Michoacan & 6     & 0.166666667 & 0.833333333 \\
+    RIMPA0114 & 699908 & Michoacan & 6     & 0.17 & 0.83 \\
     RIMPA0116 & 699910 & Mexico & 2     & 0     & 1 \\
     RIMPA0117 & 699911 & Mexico & 4     & 0     & 1 \\
-    RIMPA0118 & 699912 & Mexico & 6     & 0.166666667 & 0.833333333 \\
+    RIMPA0118 & 699912 & Mexico & 6     & 0.17 & 0.83 \\
     RIMPA0119 & 699913 & Mexico & 2     & 0     & 1 \\
@@
     RIMPA0139 & 699933 & Jalisco & 1     & 1     & 0 \\
-    RIMPA0142 & 699936 & Colima & 18    & 0.444444444 & 0.555555556 \\
+    RIMPA0142 & 699936 & Colima & 18    & 0.44 & 0.56 \\
     RIMPA0144 & 699938 & Jalisco & 2     & 1     & 0 \\
@@
     RIMPA0147 & 699941 & Jalisco & 1     & 1     & 0 \\
-    RIMPA0155 & N/A   & Jalisco & 73    & 0.01369863 & 0.98630137 \\
+    RIMPA0155 & N/A   & Jalisco & 73    & 0.01 & 0.99 \\
     RIMPA0156 & N/A   & Jalisco & 20    & 0     & 1 \\
-    RIMPA0157 & N/A   & Jalisco & 58    & 0.344827586 & 0.655172414 \\
-    RIMPA0158 & N/A   & Jalisco & 64    & 0.53125 & 0.46875 \\
+    RIMPA0157 & N/A   & Jalisco & 58    & 0.34 & 0.66 \\
+    RIMPA0158 & N/A   & Jalisco & 64    & 0.53 & 0.47 \\
     RIMPA0159 & N/A   & Jalisco & 26    & 0     & 1 \\
@@
 
-% Table S3. PCs for BayEnv
+% Table S4. PCs for BayEnv
 \begin{table}[htbp]

### 2014-05-28 manuscript.tex
edits
fixing table and figure labels

@@
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-\lev{Do you want me to edit table here to have Hop freq as a column for the 4 pops?.
-I didn't know if it would be confusing because the frequencing is not representative of the frequencing of Hop in the sequenced sample set} Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch} \jri{table 1 shows 0.016 for hopscotch, not 0.
@@
 If the \emph{Hopscotch} is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.
-\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
+We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Figure~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).

### 2014-05-28 Supplement.tex
edits
edits
edits
table formatting

@@
 \begin{document}
-
-\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
 
@@
 % Table S2. Parv and mex samples 
+\clearpage
 \footnotesize  % Switch from 12pt to 11pt; otherwise, table won't fit
-\setlength\LTleft{-30pt}            % default: \fill
-\setlength\LTright{-30pt}           % default: \fill
-\begin{longtable}{@{\extracolsep{\fill}}llllllllll@{}}
-  \hline\hline %inserts double horizontal lines
+%\setlength\LTleft{-30pt}            % default: \fill
+%\setlength\LTright{-30pt}           % default: \fill
+\begin{longtable}{llllll}
 \caption{Accessions of \emph{Zea mays} ssp.
@@
 \hline
- \textbf{Accession} & \textbf{USDA ID} & \textbf{Locality} & \textbf{\# Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\\toprule
+& \textbf{Accession} & \textbf{USDA ID} & \textbf{Locality} & \textbf{\# Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\\toprule
 \endhead
-\hline
     RIHY0009 & N/A & N/A & 2 & 0.5   & 0.5 \\
@@
 % Table S2. Mays sampling
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-
-  \centering
+\footnotesize  % Switch from 12pt to 11pt; otherwise, table won't fit
+\setlength\LTleft{-30pt}            % default: \fill
+\setlength\LTright{-30pt}           % default: \fill
+\begin{longtable}{lcc}
  \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
-    \begin{tabular}{ccc}
- %   \toprule
-    Accession & Number of alleles sampled & \emph{Hopscotch} Frequency \\
- %   \midrule
+& \textbf{Accession} & \textbf{\# Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} \\\toprule
+\endhead
     RIMMA0066 & 2     & 1 \\
@@
     \bottomrule
-    \end{tabular}
+    \end{longtable}
   \label{TableS2Mays}
-\end{table}
 

### 2014-05-29 Supplement_mbh.tex
New supplement file added

--- 2014-05-27 Supplement.tex
@@
 \usepackage{multirow}
+\usepackage[utf8]{inputenc}
+\usepackage{array}
+	\newcounter{rowno}
+	\setcounter{rowno}{0}
 \usepackage{booktabs}
@@
 \usepackage{longtable}
+
 \begin{document}
-
-\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
 
@@
 Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
-\label{FigS1Locus}
   \end{center}
@@
 %-------------------------------------------------------------------
+
+\clearpage
 
@@
   \begin{center}
-   \includegraphics[width=140mm]{FigS2Gel.png}
+   \includegraphics[width=100mm]{FigS2Gel.png}
    \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
-\label{FigS2Gel}
   \end{center}
@@
   \begin{center}
-   \includegraphics[width=140mm]{FigS3NJtrees.pdf}
+   \includegraphics[width=100mm]{FigS3NJtrees.pdf}
     \caption{Neighbor-Joining tree of the sequenced region in the 5?
@@
 Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
-\label{FigS3NJ}
   \end{center}
@@
     \end{tabular}
-  \label{Table1Fst}
 \end{table}
 
+\clearpage
+
 % Table S2. Parv and mex samples 
-\footnotesize  % Switch from 12pt to 11pt; otherwise, table won't fit
-\setlength\LTleft{-30pt}            % default: \fill
-\setlength\LTright{-30pt}           % default: \fill
-\begin{longtable}{@{\extracolsep{\fill}}llllllllll@{}}
-  \hline\hline %inserts double horizontal lines
+\begin{scriptsize}  % Switch from 12pt to 11pt; otherwise, table won't fit
+\rowcolors{1}{gray!25}{white}
+\begin{longtable}{>{\stepcounter{rowno}}cccccc}
+\hiderowcolors
 \caption{Accessions of \emph{Zea mays} ssp.
@@
 	\emph{parviglumis} and \emph{Zea mays} ssp.
-\emph{mays} hybrid.
-}
+\emph{mays} hybrid.}\\
 \hline
- \textbf{Accession} & \textbf{USDA ID} & \textbf{Locality} & \textbf{\# Alleles Sampled} & \textbf{\emph{Hopscotch} Frequency} & \textbf{No \emph{Hopscotch} Frequency} \\\toprule
+\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{USDA ID}} & \multicolumn{1}{c}{\textbf{Locality}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} & \multicolumn{1}{c}{\textbf{No \emph{Hopscotch} Freq.}} \\
+\hline
 \endhead
-\hline
+\showrowcolors
+
     RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
@@
     RIMME0011 & 566685 & Mexico & 2     & 0     & 1 \\
-    RIMME0014 & 714151 & Breeders line; Puga: 11066 & 6     & 0     & 1 \\
+    RIMME0014 & 714151 & Breeders line & 6     & 0     & 1 \\
     RIMME0017 & 699874 & Ayotlan & 8     & 0     & 1 \\
@@
     RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
-    RIMPA0060 & 714157 & Breeders line: CIMMYT 11355 & 2     & 0     & 1 \\
-    RIMPA0061 & 714158 & Breeders line: USDA PI566686 & 4     & 0.5   & 0.5 \\
+    RIMPA0060 & 714157 & Breeders line & 2     & 0     & 1 \\
+    RIMPA0061 & 714158 & Breeders line & 4     & 0.5   & 0.5 \\
     RIMPA0062 & 714159 & Breeders line & 4     & 0.5   & 0.5 \\
@@
 \end{longtable}
- \label{TableS1}
+\end{scriptsize}
+
+\clearpage
 
@@
 \begin{table}[htbp]
+\tiny
 \rowcolors{2}{gray!25}{white}
-  \centering
-
   \centering
@@
  %   \toprule
-    Accession & Number of alleles sampled & \emph{Hopscotch} Frequency \\
+    Accession & Alleles Sampled & \emph{Hopscotch} Freq. \\
  %   \midrule
@@
     \end{tabular}
-  \label{TableS2Mays}
 \end{table}
+
+\clearpage
 
@@
     \end{tabular}%
-  \label{TableS3BayEnv}%
 \end{table}%
+
+\clearpage
 
@@
   \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
-    \begin{tabular}{ccccc}
-    \toprule
-    \textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
+    \begin{tabular}{ccccc}\\
+    \textbf{Population} & \textbf{\# Seg Sites} & \textbf{$\hat\theta_\pi)$ / bp} & \textbf{$\hat\theta_W$ / bp} & \textbf{Tajima's D} \\
     \midrule
@@
     \end{tabular}%
-  \label{TableS4DiversitySNP}%
 \end{table}%
-
 \end{document}

### 2014-05-29 Supplement_mbh.tex
New supplement edited
final formatting for supplement

@@
 \documentclass[11pt]{article}
-\usepackage{geometry} 
+\usepackage[letterpaper, margin=1in]{geometry} 
 \usepackage{hyperref}
@@
 \usepackage{indentfirst}
+\usepackage{pdflscape}
 \usepackage{color}
@@
 \usepackage{longtable}
+\usepackage{fancyhdr}
+	\renewcommand{\headrulewidth}{0pt}
+	\rfoot{\thepage}
+\fancyhf{}
+\renewcommand{\headrulewidth}{0pt}
+\cfoot{\thepage}
+
+\pagestyle{fancy}
 
@@
 
-% Figure S1. Cartoon representation of tb1 locus
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=140mm]{FigS1LocusCartoon.pdf}
-   \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
-Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
+\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}
 
-\clearpage
-
-% Figure S2. Gel image
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=100mm]{FigS2Gel.png}
-   \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-\clearpage
-
-% Figure S3. NJ Trees
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=100mm]{FigS3NJtrees.pdf}
-    \caption{Neighbor-Joining tree of the sequenced region in the 5?
-UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}.
-Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).
-TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion.
-Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
-  \end{center}
-\end{figure*}
-%-------------------------------------------------------------------
-
-\clearpage
-
-% Table S1
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
-
-    \begin{tabular}{cccc}\\\toprule
-    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
-    EjuA \& EjuB & 0     & 0     & 0 \\
-    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
-    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
-    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
-    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
-    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
-    \end{tabular}
-\end{table}
-
-\clearpage
-
-% Table S2. Parv and mex samples 
-\begin{scriptsize}  % Switch from 12pt to 11pt; otherwise, table won't fit
-\rowcolors{1}{gray!25}{white}
-\begin{longtable}{>{\stepcounter{rowno}}cccccc}
-\hiderowcolors
-\caption{Accessions of \emph{Zea mays} ssp.
+\centering{Appendix S1: Accessions of \emph{Zea mays} ssp.
 \emph{mexicana} (RIMME) and \emph{Zea mays} ssp.
@@
 \emph{parviglumis} and \emph{Zea mays} ssp.
-\emph{mays} hybrid.}\\
+\emph{mays} hybrid.}
+
+% Appendix S1. Parv and mex samples 
+\begin{scriptsize}  % Switch from 12pt to 11pt; otherwise, table won't fit
+\rowcolors{1}{gray!25}{white}
+\begin{longtable}{>{\stepcounter{rowno}}cccccc}
+\hiderowcolors
 \hline
@@
 \showrowcolors
-
     RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
@@
     RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
+    \hline
 \end{longtable}
 \end{scriptsize}
-
 \clearpage
 
-% Table S2. Mays sampling
-\begin{table}[htbp]
-\tiny
-\rowcolors{2}{gray!25}{white}
-  \centering
- \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}
-    \begin{tabular}{ccc}
- %   \toprule
-    Accession & Alleles Sampled & \emph{Hopscotch} Freq. \\
- %   \midrule
+\pagenumbering{arabic}
+\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
+% Appendix S2. Mays sampling
+
+\centering{Appendix S2: \emph{Hopscotch} frequency in sampled \emph{Zea mays} ssp. \emph{mays} (RIMMA).}
+
+\begin{scriptsize}  % Switch from 12pt to 11pt; otherwise, table won't fit
+\rowcolors{1}{gray!25}{white}
+\begin{longtable}{>{\stepcounter{rowno}}ccc}
+\hiderowcolors
+\hline
+\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} \\
+\hline
+\endhead
+\showrowcolors
     RIMMA0066 & 2     & 1 \\
@@
     RIMMA0755 & 2     & 1 \\
-    \bottomrule
-    \end{tabular}
-\end{table}
+    \hline
+\end{longtable}
+\end{scriptsize}
 
 \clearpage
-
-% Table S4. PCs for BayEnv
-\begin{table}[htbp]
-  \centering
-  \caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}
-    \begin{tabular}{rrrrrrrrrrrr}
-    \toprule
-    \textbf{PC1} & \textbf{} & \textbf{PC2} & \textbf{} & \textbf{PC3} & \textbf{} & \textbf{PC4} & \textbf{} & \textbf{PC5} & \textbf{} & \textbf{PC6} & \textbf{} \\
-    \midrule
-    \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} & \textbf{Var} & \textbf{Rot} \\
-    bio1  & 0.146 & bio4  & 0.244 & prec7 & 0.287 & ts\_clay & 0.41  & bio2  & 0.38  & bio2  & 0.412 \\
-    tmean11 & 0.146 & bio3  & 0.241 & prec8 & 0.276 & v\_mod & 0.359 & sq4   & 0.328 & x\_mod & 0.365 \\
-    tmean12 & 0.145 & bio7  & 0.241 & prec11 & 0.262 & ts\_sand & 0.329 & ts\_loam & 0.289 & sq7   & 0.332 \\
-    bio11 & 0.145 & prec6 & 0.237 & bio13 & 0.247 & bio15 & 0.272 & ts\_sand & 0.266 & bio7  & 0.312 \\
-    tmax12 & 0.145 & sq7   & 0.218 & prec1 & 0.246 & prec4 & 0.259 & sq7   & 0.231 & v\_mod & 0.307 \\
-    tmin5 & 0.145 & prec9 & 0.217 & bio16 & 0.242 & x\_mod & 0.244 & bio18 & 0.213 & prec11 & 0.23 \\
-    tmean1 & 0.145 & sq3   & 0.207 & prec12 & 0.24  & prec3 & 0.226 & bio13 & 0.207 & bio18 & 0.22 \\
-    tmean2 & 0.145 & prec12 & 0.207 & bio19 & 0.238 & sq3   & 0.21  & prec11 & 0.183 & sq3   & 0.176 \\
-    tmin4 & 0.145 & bio12 & 0.204 & bio12 & 0.231 & prec5 & 0.21  & bio7  & 0.17  & sq4   & 0.153 \\
-    tmax1 & 0.145 & bio19 & 0.196 & prec2 & 0.222 & prec7 & 0.19  & bio16 & 0.163 & ts\_sand & 0.153 \\
-    tmean4 & 0.145 & prec2 & 0.188 & bio18 & 0.221 & sq4   & 0.186 & bio4  & 0.157 & bio4  & 0.148 \\
-    tmin11 & 0.144 & prec1 & 0.185 & sq4   & 0.2   & bio3  & 0.185 & bio12 & 0.156 & prec7 & 0.139 \\
-    tmax11 & 0.144 & prec10 & 0.184 & prec9 & 0.18  & bio18 & 0.178 & bio3  & 0.155 & tmax3 & 0.127 \\
-    tmin12 & 0.144 & bio16 & 0.183 & prec10 & 0.171 & sq7   & 0.132 & prec6 & 0.154 & tmax4 & 0.126 \\
-    tmin2 & 0.144 & prec8 & 0.17  & prec5 & 0.161 & bio14 & 0.116 & x\_mod & 0.152 & bio5  & 0.118 \\
-    tmean5 & 0.144 & prec5 & 0.165 & prec4 & 0.154 & bio13 & 0.099 & prec9 & 0.144 & tmax5 & 0.115 \\
-    tmean10 & 0.144 & bio14 & 0.158 & sq3   & 0.147 & bio16 & 0.095 & prec8 & 0.143 & bio3  & 0.107 \\
-    bio6  & 0.144 & bio13 & 0.151 & bio2  & 0.143 & prec8 & 0.09  & v\_mod & 0.142 & ts\_loam & 0.095 \\
-    tmax2 & 0.144 & bio17 & 0.149 & bio17 & 0.129 & bio7  & 0.077 & bio15 & 0.136 & ts\_clay & 0.078 \\
-    tmean3 & 0.144 & prec3 & 0.144 & ts\_loam & 0.127 & bio4  & 0.075 & prec7 & 0.112 & tmin9 & 0.072 \\
-    tmin1 & 0.143 & ts\_clay & 0.141 & v\_mod & 0.123 & bio2  & 0.074 & prec4 & 0.108 & tmin8 & 0.071 \\
-    tmin10 & 0.143 & bio2  & 0.129 & prec3 & 0.113 & prec2 & 0.074 & bio14 & 0.096 & prec9 & 0.07 \\
-    Altitude & 0.143 & prec7 & 0.108 & x\_mod & 0.111 & bio19 & 0.068 & tmax7 & 0.093 & tmin10 & 0.069 \\
-    bio9  & 0.143 & tmax6 & 0.107 & bio14 & 0.099 & prec12 & 0.056 & tmax8 & 0.092 & tmin12 & 0.069 \\
-    tmin3 & 0.143 & x\_mod & 0.106 & bio4  & 0.07  & ts\_loam & 0.053 & prec1 & 0.091 & tmin7 & 0.066 \\
-    bio10 & 0.142 & bio15 & 0.098 & tmax3 & 0.067 & tmax12 & 0.047 & prec2 & 0.086 & tmean4 & 0.066 \\
-    tmax10 & 0.142 & ts\_loam & 0.088 & ts\_clay & 0.065 & bio17 & 0.047 & tmin11 & 0.086 & tmax2 & 0.066 \\
-    tmax3 & 0.142 & tmean6 & 0.085 & bio15 & 0.056 & bio9  & 0.043 & prec5 & 0.082 & tmax6 & 0.064 \\
-    tmax4 & 0.142 & tmin7 & 0.082 & tmax2 & 0.055 & tmax8 & 0.042 & bio17 & 0.082 & tmean3 & 0.064 \\
-    tmin6 & 0.142 & bio5  & 0.082 & tmean3 & 0.052 & tmax1 & 0.041 & tmin12 & 0.08  & bio9  & 0.058 \\
-    tmean9 & 0.141 & tmean7 & 0.081 & ts\_sand & 0.05  & tmax5 & 0.039 & prec3 & 0.078 & tmin11 & 0.058 \\
-    tmin9 & 0.141 & prec4 & 0.08  & prec6 & 0.048 & tmax7 & 0.039 & tmax9 & 0.078 & prec3 & 0.056 \\
-    tmean8 & 0.141 & tmax7 & 0.079 & sq7   & 0.048 & prec10 & 0.038 & tmin1 & 0.077 & tmean5 & 0.056 \\
-    bio8  & 0.14  & bio8  & 0.079 & tmin7 & 0.046 & Altitude & 0.037 & tmin10 & 0.074 & bio10 & 0.051 \\
-    tmean6 & 0.14  & tmax9 & 0.077 & bio3  & 0.044 & tmax10 & 0.035 & bio6  & 0.071 & bio15 & 0.047 \\
-    tmean7 & 0.14  & tmean8 & 0.076 & tmax4 & 0.043 & tmax2 & 0.033 & prec12 & 0.067 & tmin1 & 0.045 \\
-    tmin8 & 0.14  & tmin8 & 0.076 & bio7  & 0.042 & tmax9 & 0.03  & tmin2 & 0.061 & prec12 & 0.045 \\
-    tmax5 & 0.14  & tmax5 & 0.074 & tmax1 & 0.036 & tmean12 & 0.029 & tmin6 & 0.061 & bio17 & 0.045 \\
-    tmax9 & 0.139 & tmax8 & 0.074 & tmin3 & 0.035 & tmax11 & 0.027 & tmax6 & 0.059 & tmean10 & 0.043 \\
-    tmax8 & 0.139 & tmean9 & 0.072 & bio9  & 0.035 & tmean1 & 0.027 & tmin3 & 0.052 & prec4 & 0.043 \\
-    bio5  & 0.139 & bio18 & 0.07  & tmin8 & 0.034 & bio5  & 0.026 & bio8  & 0.046 & Altitude & 0.041 \\
-    tmax7 & 0.139 & v\_mod & 0.066 & tmean4 & 0.031 & tmean5 & 0.026 & tmean11 & 0.041 & bio6  & 0.039 \\
-    tmin7 & 0.138 & tmin9 & 0.066 & tmean2 & 0.031 & bio11 & 0.025 & tmean1 & 0.04  & tmin6 & 0.038 \\
-    tmax6 & 0.135 & bio10 & 0.065 & tmax12 & 0.03  & tmean2 & 0.022 & tmin9 & 0.04  & tmean12 & 0.035 \\
-    bio17 & 0.107 & tmax10 & 0.063 & tmean7 & 0.028 & tmin6 & 0.022 & tmean12 & 0.04  & tmean9 & 0.035 \\
-    bio14 & 0.095 & tmin6 & 0.061 & tmin9 & 0.023 & prec1 & 0.022 & tmax10 & 0.038 & prec5 & 0.033 \\
-    prec3 & 0.094 & prec11 & 0.056 & Altitude & 0.021 & tmean7 & 0.02  & tmin5 & 0.038 & tmean8 & 0.03 \\
-    bio15 & 0.092 & ts\_sand & 0.054 & tmax11 & 0.019 & tmean8 & 0.02  & bio11 & 0.037 & tmin2 & 0.03 \\
-    prec4 & 0.088 & sq4   & 0.049 & tmin4 & 0.018 & tmean9 & 0.019 & tmean8 & 0.035 & tmax1 & 0.03 \\
-    prec10 & 0.084 & tmin3 & 0.048 & tmax5 & 0.017 & prec11 & 0.019 & tmean7 & 0.033 & prec1 & 0.025 \\
-    ts\_sand & 0.073 & tmean10 & 0.047 & tmin10 & 0.017 & tmean11 & 0.018 & tmean2 & 0.03  & bio13 & 0.024 \\
-    ts\_loam & 0.072 & tmean5 & 0.044 & tmin6 & 0.017 & tmax4 & 0.016 & bio19 & 0.029 & tmean11 & 0.023 \\
-    prec6 & 0.069 & tmax4 & 0.042 & tmean6 & 0.017 & tmean10 & 0.016 & bio9  & 0.027 & tmax7 & 0.023 \\
-    prec9 & 0.063 & tmax11 & 0.041 & tmax6 & 0.017 & tmin5 & 0.015 & tmin7 & 0.026 & tmean7 & 0.023 \\
-    prec2 & 0.063 & bio6  & 0.038 & tmin12 & 0.017 & tmean6 & 0.015 & tmin4 & 0.025 & prec10 & 0.022 \\
-    bio4  & 0.06  & tmin1 & 0.038 & bio11 & 0.016 & prec6 & 0.014 & tmean5 & 0.023 & prec2 & 0.02 \\
-    prec5 & 0.058 & tmin2 & 0.037 & tmin11 & 0.015 & bio6  & 0.012 & tmean3 & 0.022 & bio8  & 0.018 \\
-    bio13 & 0.055 & Altitude & 0.036 & tmin1 & 0.013 & tmin1 & 0.012 & tmin8 & 0.022 & tmean2 & 0.017 \\
-    x\_mod & 0.053 & tmin10 & 0.034 & tmean5 & 0.013 & tmin12 & 0.011 & tmean10 & 0.021 & bio16 & 0.016 \\
-    prec1 & 0.05  & bio9  & 0.032 & tmean8 & 0.013 & tmin2 & 0.011 & tmean9 & 0.019 & tmean6 & 0.014 \\
-    bio16 & 0.049 & tmean1 & 0.029 & bio6  & 0.012 & bio1  & 0.011 & tmean4 & 0.019 & bio19 & 0.012 \\
-    bio7  & 0.043 & bio1  & 0.028 & tmean9 & 0.01  & tmin4 & 0.009 & ts\_clay & 0.018 & tmax11 & 0.012 \\
-    bio19 & 0.042 & tmean3 & 0.026 & tmean1 & 0.01  & tmax6 & 0.008 & prec10 & 0.015 & tmax10 & 0.011 \\
-    bio12 & 0.042 & tmean2 & 0.024 & bio10 & 0.01  & tmin11 & 0.007 & bio1  & 0.014 & tmax8 & 0.01 \\
-    bio3  & 0.04  & bio11 & 0.022 & tmax8 & 0.009 & bio8  & 0.007 & tmax4 & 0.014 & tmean1 & 0.009 \\
-    prec7 & 0.034 & tmean11 & 0.019 & tmax7 & 0.009 & bio10 & 0.007 & bio10 & 0.011 & bio14 & 0.009 \\
-    prec12 & 0.033 & tmin12 & 0.019 & bio5  & 0.008 & tmin9 & 0.006 & tmax5 & 0.009 & prec8 & 0.007 \\
-    bio2  & 0.033 & tmax1 & 0.017 & tmin5 & 0.008 & prec9 & 0.005 & bio5  & 0.008 & bio11 & 0.007 \\
-    prec8 & 0.032 & tmean4 & 0.016 & tmin2 & 0.007 & bio12 & 0.004 & tmax11 & 0.008 & tmin5 & 0.006 \\
-    sq4   & 0.026 & tmin5 & 0.013 & bio1  & 0.007 & tmax3 & 0.003 & tmax12 & 0.008 & tmax9 & 0.005 \\
-    sq3   & 0.024 & tmean12 & 0.012 & tmax10 & 0.006 & tmean4 & 0.003 & tmax3 & 0.007 & bio12 & 0.004 \\
-    v\_mod & 0.015 & tmax2 & 0.011 & tmean10 & 0.006 & tmin3 & 0.002 & sq3   & 0.007 & tmin3 & 0.004 \\
-    ts\_clay & 0.012 & tmin4 & 0.01  & tmean12 & 0.005 & tmin7 & 0.002 & tmax2 & 0.006 & prec6 & 0.004 \\
-    sq7   & 0.004 & tmax3 & 0.003 & bio8  & 0.004 & tmean3 & 0.002 & Altitude & 0.004 & tmin4 & 0.003 \\
-    prec11 & 0.004 & tmax12 & 0.002 & tmax9 & 0.003 & tmin10 & 0.002 & tmean6 & 0.002 & bio1  & 0.002 \\
-    bio18 & 0.001 & tmin11 & 0.001 & tmean11 & 0.001 & tmin8 & 0     & tmax1 & 0.001 & tmax12 & 0.002 \\
-    \textbf{Bayes Factors} &       &       &       &       &       &       &       &       &       &       &  \\
-    \multicolumn{2}{c}{0.77165} & \multicolumn{2}{c}{0.51652} & \multicolumn{2}{c}{0.21120} & \multicolumn{2}{c}{0.24063} & \multicolumn{2}{c}{1.96010} & \multicolumn{2}{c}{0.24594} \\
-    \bottomrule
-    \end{tabular}%
-\end{table}%
+\pagenumbering{arabic}
+\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
+% Appendix S3. Cartoon representation of tb1 locus
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=180mm]{FigS1LocusCartoon.pdf}
+   \flushleft{Appendix S3: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
 
 \clearpage
+\pagenumbering{arabic}
+\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}
 
-% Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
-\begin{table}[htbp]
-  \centering
-  \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
-    \begin{tabular}{ccccc}\\
-    \textbf{Population} & \textbf{\# Seg Sites} & \textbf{$\hat\theta_\pi)$ / bp} & \textbf{$\hat\theta_W$ / bp} & \textbf{Tajima's D} \\
-    \midrule
-    Ejutla A & 4     & 0.15217 & 0.11902 & 0.76191 \\
-    Ejutla B & 5     & 0.15258 & 0.14877 & 0.07412 \\
-    La Mesa & 3     & 0.12802 & 0.08926 & 1.09209 \\
-    San Lorenzo & 3     & 0.09098 & 0.08926 & 0.04845 \\
-    \bottomrule
-    \end{tabular}%
-\end{table}%
+% Appendix S4. Gel image
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=100mm]{FigS2Gel.png}
+   \flushleft{Appendix S4: Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
+
+\clearpage
+\pagenumbering{arabic}
+\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}
+
+% Appendix S5. NJ Trees
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
+    \flushleft{Appendix S5: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}.
+Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).
+TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion.
+Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
+  \end{center}
+\end{figure*}
+%-------------------------------------------------------------------
 \end{document}

### 2014-05-29 manuscript.tex
final formatting for supplement

@@
 BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
-We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP (Table~S3).
+We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
 Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
@@
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
-Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1; Table~S3).
+Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).
 
@@
 which is right?}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Figure~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Fig.~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
 

### 2014-05-29 manuscript.tex
final formatting for manuscript

@@
 \usepackage[pdftex]{graphicx}  
-\usepackage{caption}
+\usepackage[labelsep=period]{caption}
 \usepackage{natbib}
@@
 \subsection*{Acknowledgements}
-We would like to thank the Department of Plant Sciences at UC Davis for Graduate Student Research funding to LEV, \lev{Matt UC Mexus}, and G. Coop for helpful discussion.
-
-\lev{Grants as far as I know were Matt's UC Mexus money, whatever money paid for the 55K data since we used that, and I am not sure what paid for greenhouse space} \jri{55K data was public, no prob.
-acknowledge support from the Dept.
-of Plant Sciences for LV's GSR and for "research funds" or something vague, as this came from startup.
-acknowledge graham coop for helpful discussion}
+The authors thank the Department of Plant Sciences at UC Davis for Graduate Student Research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
+Coop for helpful discussion.
 }
@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table~S1 and Table~S2).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.~S1, Fig.~S2) \mbh{Primers in Fig.~S1 should be labeled}.
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Fig.~S2).
+HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article).
 When only one PCR resolved well, we scored one allele for the individual.
@@
 For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (Region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} (66,169bp upstream from the start of the \emph{tb1} ORF; Region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Fig.~S1).
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Appendix 1, See Supplemental Materials with the online version of this article).
 We used Taq polymerase (New England Biolabs Inc., Ipswich, Massachusetts, USA) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
@@
 
-\textbf{Genotyping analysis---}We examined discrepancies between observed and expected genotype frequencies by calculating Hardy-Weinberg Equilibrium (HWE).
-To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations in which 8 or more individuals were sampled.
+\textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
+These analyses only included populations in which eight or more individuals were sampled.
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
@@
 \textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
-Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table~S1 and Table~S2).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and 4 populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled 8 or more alleles.
+Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and four populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more alleles.
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
@@
 
-%Figure 1 Map 
-%--------------------------------------------
-\begin{SCfigure}[][t!]
-   \includegraphics[width=0.6\textwidth]{Fig1Map.pdf}
-    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
-Size of circles reflects number of alleles sampled.
-}
-     \label{Fig1Map}
-\end{SCfigure}
-%--------------------------------------------
-
-% Table 1
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
-
-    \begin{tabular}{cccc}\\\toprule
-    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
-    EjuA \& EjuB & 0     & 0     & 0 \\
-    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
-    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
-    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
-    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
-    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
-    \end{tabular}
-  \label{Table1Fst}
-\end{table}
-
 \textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
 After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
-For Region 1, Ejutla A has the highest values of haplotype diversity, and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}, Table~S2).
+For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}, Appendix 2, See Supplemental Materials with the online version of this article).
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
 Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
-Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are 0 for both sequenced regions as well as for the \emph{Hopscotch} \jri{table 1 shows 0.016 for hopscotch, not 0.
-which is right?}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Fig.~S3); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
-
-% Table 2 Diversity Statistics
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
-    \begin{tabular}{ccccc}\\\toprule
-    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
-    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
-    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
-    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
-    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
-    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
-    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
-    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
-    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
-    \end{tabular}
-  \label{Table2Diversity}
-\end{table}
+Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Appendix 5, See Supplemental Materials with the online version of this article); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
 
@@
 
-% Table 3 $r^{2}$ values
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{mean $r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
-    \begin{tabular}{ccccc}\\\toprule
-    \textbf{Population} & \textbf{Chr. 1} & \textbf{\emph{tb1} region} & \textbf{Region 1} & \textbf{Region 2} \\\midrule
-    Ejutla A & 0.095 & 0.050 & 0.747 & 0.215 \\
-    Ejutla B & 0.069 & 0.051 & 0.660 & 0.186 \\
-    La Mesa & 0.070 & 0.053 & 0.914 & 0.766 \\
-    San Lorenzo & 0.101 & 0.067 & 0.912 & 0.636 \\
-    \bottomrule
-    \end{tabular}
-  \label{Table3R2}
-\end{table}
-
-% Figure 2. LD plots
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=140mm]{Fig2LDPlots.pdf}
-   \end{center}
-    \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
-The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
-A) Ejutla A; B) Ejutla B; C) La Mesa; D) San Lorenzo.
-The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.}
-\label{Fig2LD}
-\end{figure*}
-%-------------------------------------------------------------------
-
 The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
@@
 
- % Table 4 Assignment to maize and teosinte in tb1 region and chr 1 
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
-    \begin{tabular}{ccccc} \
-          & \multicolumn{2}{c}{\textbf{\emph{tb1} region}} & \multicolumn{2}{c}{\textbf{Chr 1}} \\\toprule
-    \bf{Population} & \bf{Maize} & \bf{Teosinte} & \bf{Maize} & \bf{Teosinte} \\\midrule
-    Ejutla A & 0.022 & 0.978 & 0.203 & 0.797 \\
-    Ejutla B & 0.019 & 0.981 & 0.187 & 0.813 \\
-    La Mesa & 0.012 & 0.988 & 0.193 & 0.807 \\
-    San Lorenzo & 0.016 & 0.984 & 0.205 & 0.795 \\\bottomrule
-    \end{tabular}
-  \label{Table4Q}
-\end{table}
-
-% Figure 3 STRUCTURE Plot 4 Pops
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=90mm]{Fig3Structure.pdf}
-     \end{center}
-    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF.}
-\label{Fig3Structure}
-\end{figure*}
-%-------------------------------------------------------------------
-
 \textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
@@
 
-% Figure 4 Phenotyping BoxPlots
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=90mm]{Fig4Boxplots.pdf}
-   \end{center}
-    \caption{Box-plots showing tillering index in greenhouse grow-outs for Phenotyping 1 (A) and Phenotyping 2 (B).
-White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.
-Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
-Outliers are displayed as dots, the maximum value excluding outliers is shown with the top whisker, while the minimum excluding outliers is shown with the bottom whisker.}
-\label{Fig4Boxplots}
-\end{figure*}
-%-------------------------------------------------------------------
-
 We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.
 For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
-We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, \jri{is it in HWE in this pop?
-my guess is no!} with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
 Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.
@@
 
+% Table 1
+\begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
+
+    \begin{tabular}{cccc}\\\toprule
+    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
+    EjuA \& EjuB & 0     & 0     & 0 \\
+    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
+    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
+    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
+    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
+    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
+    \end{tabular}
+  \label{Table1Fst}
+\end{table}
+
+% Table 2 Diversity Statistics
+\begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
+    \begin{tabular}{ccccc}\\\toprule
+    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
+    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
+    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
+    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
+    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
+    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
+    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
+    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
+    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
+    \end{tabular}
+  \label{Table2Diversity}
+\end{table}
+
+% Table 3 $r^{2}$ values
+\begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+  \caption{mean $r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
+    \begin{tabular}{ccccc}\\\toprule
+    \textbf{Population} & \textbf{Chr. 1} & \textbf{\emph{tb1} region} & \textbf{Region 1} & \textbf{Region 2} \\\midrule
+    Ejutla A & 0.095 & 0.050 & 0.747 & 0.215 \\
+    Ejutla B & 0.069 & 0.051 & 0.660 & 0.186 \\
+    La Mesa & 0.070 & 0.053 & 0.914 & 0.766 \\
+    San Lorenzo & 0.101 & 0.067 & 0.912 & 0.636 \\
+    \bottomrule
+    \end{tabular}
+  \label{Table3R2}
+\end{table}
+
+ % Table 4 Assignment to maize and teosinte in tb1 region and chr 1 
+\begin{table}[htbp]
+\rowcolors{2}{gray!25}{white}
+  \centering
+  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
+    \begin{tabular}{ccccc} \
+          & \multicolumn{2}{c}{\textbf{\emph{tb1} region}} & \multicolumn{2}{c}{\textbf{Chr 1}} \\\toprule
+    \bf{Population} & \bf{Maize} & \bf{Teosinte} & \bf{Maize} & \bf{Teosinte} \\\midrule
+    Ejutla A & 0.022 & 0.978 & 0.203 & 0.797 \\
+    Ejutla B & 0.019 & 0.981 & 0.187 & 0.813 \\
+    La Mesa & 0.012 & 0.988 & 0.193 & 0.807 \\
+    San Lorenzo & 0.016 & 0.984 & 0.205 & 0.795 \\\bottomrule
+    \end{tabular}
+  \label{Table4Q}
+\end{table}
+
+\clearpage
+
+%Figure 1 Map 
+%--------------------------------------------
+\begin{figure*}[!t]
+\begin{center}
+   %\includegraphics[width=0.6\textwidth]{Fig1Map.pdf}
+   \end{center}
+    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
+Size of circles reflects number of alleles sampled.}
+\label{Fig1Map}
+\end{figure*}
+%--------------------------------------------
+
+% Figure 2. LD plots
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   %\includegraphics[width=140mm]{Fig2LDPlots.pdf}
+   \end{center}
+    \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
+The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
+A) Ejutla A; B) Ejutla B; C) La Mesa; D) San Lorenzo.
+The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.}
+\label{Fig2LD}
+\end{figure*}
+%-------------------------------------------------------------------
+
+% Figure 3 STRUCTURE Plot 4 Pops
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   %\includegraphics[width=90mm]{Fig3Structure.pdf}
+     \end{center}
+    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF.}
+\label{Fig3Structure}
+\end{figure*}
+%-------------------------------------------------------------------
+
+% Figure 4 Phenotyping BoxPlots
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   %\includegraphics[width=90mm]{Fig4Boxplots.pdf}
+   \end{center}
+    \caption{Box-plots showing tillering index in greenhouse grow-outs for Phenotyping 1 (A) and Phenotyping 2 (B).
+White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.
+Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
+Outliers are displayed as dots, the maximum value excluding outliers is shown with the top whisker, while the minimum excluding outliers is shown with the bottom whisker.}
+\label{Fig4Boxplots}
+\end{figure*}
+%-------------------------------------------------------------------
+
 \end{flushleft}

### 2014-05-30 manuscript.tex
final document edits prior to submission

@@
 \subsection*{Acknowledgements}
-The authors thank the Department of Plant Sciences at UC Davis for Graduate Student Research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
+The authors thank the Department of Plant Sciences at UC Davis for graduate student research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
 Coop for helpful discussion.
@@
 %TO DO
-%check citations in last paragraph for ter1
+%check citations in last paragraph for te1
 
@@
 \emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene is a major QTL controlling branching differences between maize and its wild progenitor, teosinte.
-Previous work has shown that the insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} enhances the gene's expression, causing much of the reduction in tillering observed in maize.
-Observations of the maize \emph{tb1} allele in teosinte and estimates of an age of insertion of the \emph{Hopscotch} element that predates domestication led us to investigate its prevalence and potential role in teosinte.
+The insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} is known to enhance the gene's expression, causing reduced tillering in maize.
+Observations of the maize \emph{tb1} allele in teosinte and estimates of an insertion age of the \emph{Hopscotch} that predates domestication led us to investigate its prevalence and potential role in teosinte.
 
@@
 Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{ter1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{ter1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
+Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{te1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{te1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
 \emph{tru1}, for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}).
@@
 Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{ter1}).
+\emph{gt1}, \emph{tru1}, and \emph{te1}).
 These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.

### 2014-06-04 manuscript.tex
adding comments to manuscript file

@@
-\documentclass[11pt]{article}
-\usepackage[letterpaper, margin=1in]{geometry} 
+\documentclass[11pt]{article} %indicating type of document, font size
+\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
 \usepackage{hyperref}
-\usepackage[leftcaption]{sidecap}
+\usepackage[leftcaption]{sidecap} %package for adding side captions
 \sidecaptionvpos{figure}{c} %position side caption
-\usepackage{lineno}
+\usepackage{lineno} %package for adding line numbers
 \usepackage{lipsum}
@@
 \usepackage[pdftex]{graphicx}  
-\usepackage[labelsep=period]{caption}
-\usepackage{natbib}
-\bibliographystyle{AJB}
+\usepackage[labelsep=period]{caption} %package for formatting figure captions, separation between figure number and caption is period per AJB formatting requirements
+\usepackage{natbib} % package for formatting citations
+\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for formatting...journals often have these available but these can also be customized
 \usepackage{authblk}
@@
 \renewcommand{\baselinestretch}{1.5}
-\renewcommand{\refname}{\centering{LITERATURE CITED}}
+\renewcommand{\refname}{\centering{LITERATURE CITED}} %changing title of references to Literature Cited as stipulated by AJB
 
-\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
-\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
-\newcommand{\lev}[1]{\textcolor{green}{\emph{\scriptsize #1}} }
+\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %introducting new command "\jri" for jeff ross-ibarra to make comments in blue italics
+\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %same for matt hufford in red
+\newcommand{\lev}[1]{\textcolor{green}{\emph{\scriptsize #1}} } %same for laura vann in green
 
@@
 \makeatletter
-\renewcommand{\maketitle}{\bgroup\vspace*{1cm}\setlength{\parindent}{0pt}
+\renewcommand{\maketitle}{\bgroup\vspace*{1cm}\setlength{\parindent}{0pt}%special formatting for AJB for title page
 \begin{center}
@@
 
-\author[2]{Laura Vann}
+\author[2]{Laura Vann}%author information
 \author[2,3]{Thomas Kono}
@@
 
-\clearpage
+\clearpage%inserts page break
 \footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted \line(1,0){120}.}
 \vspace*{1cm}
-\subsection*{Acknowledgements}
+\subsection*{Acknowledgements}%subsection header
 The authors thank the Department of Plant Sciences at UC Davis for graduate student research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
@@
 
-\vspace*{0.2cm}
+\vspace*{0.2cm}%introduces vertical space between abstract sections
 
@@
 
-\begin{centering}
+\begin{centering}%centers section title
 \section*{INTRODUCTION}
@@
 
-\bibliography{refs}
+\bibliography{refs}%print bibliography
 
@@
 \begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
+\rowcolors{2}{gray!25}{white}%alternating row colors in table
   \centering
-  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}
+  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}%table caption
 
-    \begin{tabular}{cccc}\\\toprule
+    \begin{tabular}{cccc}\\\toprule%number of columns and line across top of table
     \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
@@
     \end{tabular}
-  \label{Table1Fst}
+  \label{Table1Fst}%indicates link between table and mention in text
 \end{table}

### 2014-06-27 ResponseToReview.tex
Response to Review
Starting response to review.

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\usepackage{lipsum}
\usepackage{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}

\begin{document}

We would like to thank the Editor in Chief and the two reviewers for their helpful comments and reviews. We have addressed the comments to the text, and while we feel that the suggested new experiments would provide interesting results, we feel that they are not a reasonable request to include in our paper.  We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses below.\\\

\textbf{Response to Associate Editor:}

For my own part, I agree with both reviewers that the study would be of interest to readers of AJB.  However, I took a more negative view towards several aspects of the Methods, Results, and Discussion.  Three examples:

1) I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don’t provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.

2) Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken? Second, why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.  Third, it was not clear to me what “to detect the observed effect” means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.  Fourth, what was the purpose of the Phenotyping 2
experiment?  You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.

3) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.  It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.

Because of what I perceive to be substantial problems both with the methods and the presentation and discussion of results, and because Reviewer 1 raises some important issues with the phenotyping experiments, I am recommending that the manuscript be rejected.

Yours sincerely,
Andrew Schnabel
Associate Editor

Reviewer #1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1. The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis. The authors go on to sequence regions up and downstream of Hopscotch in subset of their initial sample. In analyzing their sequence data, they find no evidence for recent introgression of the maize Hopscotch locus into wild teosintes. They also find evidence for selection acting on tb1 in parviglumis. In greenhouse experiments, no difference in tiller index or tiller number was observed between teosinte lines with or without Hopscotch.

They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate. However, I feel like they need to make some further effort to understand what is happening mechanistically. What is Hopscotch doing in maize vs these sampled teosintes? I feel like the expression analyses and phenotyping experiments they suggest should be part of a later study should be part of this study. At a minimum, qRT-PCR experiments should be performed assessing tb1 expression in their with/without Hopscotch lines. Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.

Another potential avenue, that the authors also suggest, lies in more directed and exhaustive greenhouse experiments. Variation in red and far-red light has been shown to be important in regulating both tb1 (in Sorghum) and gt1 in maize. Growth chamber experiments (with or without supplemental far red light, with or without Hopscotch), coupled to expression analyses in seedlings, seem like they should be possible.  In the greenhouse experiments that were done, phenotyping was incomplete. Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture and internode length. Was there any evidence for variation in these other traits? In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.

Reviewer #2: Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.  It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana. Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.
Methods and statistics were appropriately used as far as my expertise allows me to judge.
Specific suggestions follow:
Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples?
Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…"
Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte?
Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions?
Materials and Methods:
Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.
Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.
Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".  Sorry, what does this mean? That you have one band (one assumes the short one?). If you are talking presence/absence, how do you know the allele?
Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include?
Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis? Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?
Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte?
Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb?
Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically? Or you turned on the drip by hand? In which case, it was not automatic…
Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1. F" So, why did you do it? Because of the last sentence in the materials and methods? Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.
Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols?
Results:
Page 10 lines 5 - 6: "Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total? So we can see if this comes out to a very different ratio than in maize.
Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles? I thought you only had presence or absence; how are alleles differentiated with your PCR?
Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011). Add Warburton, et al., 2011.  Gene flow among different teosinte taxa and into the domesticated maize gene pool.  Genet Resour Crop Evol 58:1243-1261.
Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32.
\end{document}

### 2014-06-28 ResponseToReview.tex
Response to review
short comments
Reviewer 1
notes for reviewer 1

@@
 \usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
+\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
 \usepackage{lipsum}
@@
 
-We would like to thank the Editor in Chief and the two reviewers for their helpful comments and reviews.
-We have addressed the comments to the text, and while we feel that the suggested new experiments would provide interesting results, we feel that they are not a reasonable request to include in our paper.
-We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses below.\\\
+\noindent Dear Dr. Andrew Schnabel and anonymous reviewers, \\
 
-\textbf{Response to Associate Editor:}
+We would like to thank the Editor in Chief and the two anonymous reviewers for their helpful comments and suggestions.
+We have addressed the comments to the text, and while we feel that the suggested new experiments would provide interesting results, we do not think that they are a reasonable request to include in resubmission of our paper.
+We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses in blue below.
+We hope that you are willing to consider our revised manuscript for resubmission.\\\
+
+\noindent Sincerely,\\
+
+\res{would this be jeff, matt, me? or everyone?}
+
+\section*{Response to Associate Editor:}
 
@@
 
+\res{I vote we take out trees}
+
 2) Phenotyping experiments - These are curious experiments.
@@
 Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?
-How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?
+\res{we didn't know genotypes ahead of time...they were seeds collected from natural population} How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?
 Second, why are sample sizes not reported?
-You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.
+You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class \res{I will check this.
+I thought it was listed in the text.
+maybe older version}.
 Third, it was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
@@
 Thus, it is not clear what effect you are referring to.
-Fourth, what was the purpose of the Phenotyping 2
-experiment?
-You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.
+Fourth, what was the purpose of the Phenotyping 2 \res{we did phenotyping 2 because we had such an odd distribution of genotypes in Phenotyping 1 that we had no power to make any inferences regarding effect of Hop on TI} experiment?
+You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations {We didn't sample an additional population aside from SLO for phenotyping -- is that what he is inferring?} with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.
 
@@
 It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
+\res{should we restructure this?
+I thought it read clearly.
+Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
 
@@
 Associate Editor
+
+\section*{Reviewer 1:}
 

### 2014-06-28 ResponseToReview.tex
Response
Finished shorthand notes in response to comments. Let me know what you
guys think.

@@
 In greenhouse experiments, no difference in tiller index or tiller number was observed between teosinte lines with or without Hopscotch.
+\res{yes sounds right}
 
@@
 Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.
+\res{sure this would all be cool and interesting, but not going to happen for this paper.
+it is enough work on it's own to be a separate paper}.
 
@@
 In the greenhouse experiments that were done, phenotyping was incomplete.
-Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture and internode length.
+Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture \res{we can't do this unless plants all go to seed} and internode length.
 Was there any evidence for variation in these other traits?
 In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.
+\res{no more phenotypine experiments}.
 
-One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
+One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown. \res{is this necessary?}
 
@@
 It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.
-Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
-This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.
+\res{sure we knew it existed in some individuals, but not to the extent that we found in our study} Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
+This paper is fairly maize-centric.
+A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract.
+This could be easily remedied.
+\res{okay.
+but it is a paper about maize...}
 Methods and statistics were appropriately used as far as my expertise allows me to judge.
 Specific suggestions follow:
+
 Introduction:
@@
 Can you give some examples?
-Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…"
+\res{okay can do this.}
+
+Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…" \res{doable}
+
 Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte?
@@
 If the second, was it actually studied, or merely hypothesized in teosinte?
+\res{doable}
+
 Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
@@
 Does the analysis of this group of teosintes add anything to your conclusions?
+\res{no mexicana phenotyping.
+seed limited.
+there were a couple populations of mex that had hop at intermediate frequency.}
+
 Materials and Methods:
-Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.
-Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.
+Page 5, lines 22 - 23: how many individuals per accession?
+A fixed number, or variable, and depending on what?
+This can be mentioned so the reader does not have to go to the appendix or supplements.
+\res{no way we can list all those numbers in the text.
+varied anywhere from 2 to 30 is my guess.
+can say that?}
+
+Page 5, Line 28: mention how big the hopscotch element is.
+Can it be amplified in one amplicon?
+Again, this would be useful to know without having to look in appendix or supplements.
+\res{i think this is obvious in the methods when we discuss primers sets}
+
 Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".
@@
 If you are talking presence/absence, how do you know the allele?
-Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include?
+\res{PCRs for different alleles.
+maybe should say that in the text to clarify?}
+
+Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include? \res{can put this in text}
+
 Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis?
 Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?
-Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte?
+\res{i have no idea.
+do either of you know if this is in tanja's paper?
+or should i look at the website she got data from?}
+
+Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte? \res{can explain dataset in paper}
+
 Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb?
+\res{hmm.
+Matt why did you use 5kb when you did this for your study?}
+
 Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?
 Or you turned on the drip by hand?
-In which case, it was not automatic…
+In which case, it was not automatic… \res{i'll just take out the by hand.
+it's too confusing to explain what happened}.
+
 Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1.
@@
 Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.
-Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols?
+\res{we wanted an independent trait to compare patterns of TI to}
+
+Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols? \res{PCR}
+
 Results:
@@
 So we can see if this comes out to a very different ratio than in maize.
+\res{okay can put in ratios or something}
+
 Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
 I thought you only had presence or absence; how are alleles differentiated with your PCR?
+\res{can clarify.
+basically only used populations that had aboive a certain sample size.
+alleles based on size difference in PCR}.
+
 Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
@@
 Genet Resour Crop Evol 58:1243-1261.
-Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32.
+\res{okay good suggestion}.
+
+Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32. \res{okay}.
+
 \end{document}

### 2014-07-03 ResponseToReview.tex
response edits

@@
 \newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
+\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
 \usepackage{lipsum}
@@
 We would like to thank the Editor in Chief and the two anonymous reviewers for their helpful comments and suggestions.
-We have addressed the comments to the text, and while we feel that the suggested new experiments would provide interesting results, we do not think that they are a reasonable request to include in resubmission of our paper.
+We have addressed the comments to the text, but while we feel that the suggested new experiments would provide interesting results, we worry that they would require considerable extra time (many months) and are outside of the scope of this current work.
 We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses in blue below.
@@
 
-\res{would this be jeff, matt, me? or everyone?}
+Jeffrey Ross-Ibarra and Matthew Hufford
 
@@
 
-For my own part, I agree with both reviewers that the study would be of interest to readers of AJB.
+%For my own part, I agree with both reviewers that the study would be of interest to readers of AJB.
 However, I took a more negative view towards several aspects of the Methods, Results, and Discussion.
@@
 As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.
-The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.
+The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\
 
+\jri{please add two backslashes after every reviewer comment for spacing.
+happy to remove trees, but perhaps better we redo trees collapsing identical haplotypes.
+or  make a haplotype network with pie charts or something.}
 \res{I vote we take out trees}
@@
 Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?
-\res{we didn't know genotypes ahead of time...they were seeds collected from natural population} How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?
+How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?
+
+\res{we didn't know genotypes ahead of time...they were seeds collected from natural population} 
+\jri{write appropriate response, no or minor change to text for explanation}
+
 Second, why are sample sizes not reported?
-You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class \res{I will check this.
-I thought it was listed in the text.
-maybe older version}.
+You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class
+
+\res{I will check this. I thought it was listed in the text. maybe older version}.  
+\jri{fix, respond}
+
 Third, it was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
@@
 Thus, it is not clear what effect you are referring to.
-Fourth, what was the purpose of the Phenotyping 2 \res{we did phenotyping 2 because we had such an odd distribution of genotypes in Phenotyping 1 that we had no power to make any inferences regarding effect of Hop on TI} experiment?
-You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations {We didn't sample an additional population aside from SLO for phenotyping -- is that what he is inferring?} with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.
+
+\jri{explain, clarify in text, give new text here}
+
+Fourth, what was the purpose of the Phenotyping 2  experiment?
+You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.
+
+\res{we did phenotyping 2 because we had such an odd distribution of genotypes in Phenotyping 1 that we had no power to make any inferences regarding effect of Hop on TI.
+We didn't sample an additional population aside from SLO for phenotyping -- is that what he is inferring?}
+\jri{explain, respond, add some clarifying text to document, provide new text}
 
@@
 It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
-\res{should we restructure this?
-I thought it read clearly.
-Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
 
-Because of what I perceive to be substantial problems both with the methods and the presentation and discussion of results, and because Reviewer 1 raises some important issues with the phenotyping experiments, I am recommending that the manuscript be rejected.
+\res{should we restructure this? I thought it read clearly. Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
+\jri{clean up, restructure, thank editor, explain changes}
 
-Yours sincerely,
-Andrew Schnabel
-Associate Editor
+%Because of what I perceive to be substantial problems both with the methods and the presentation and discussion of results, and because Reviewer 1 raises some important issues with the phenotyping experiments, I am recommending that the manuscript be rejected.
 
-\section*{Reviewer 1:}
+%Yours sincerely,
+%Andrew Schnabel
+%Associate Editor
 
-Reviewer #1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1.
+\section*{Reviewer 1}
+
+%Reviewer \#1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1.
 The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis.
@@
 In greenhouse experiments, no difference in tiller index or tiller number was observed between teosinte lines with or without Hopscotch.
-\res{yes sounds right}
+
+%\res{yes sounds right}
 
@@
 Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.
-\res{sure this would all be cool and interesting, but not going to happen for this paper.
-it is enough work on it's own to be a separate paper}.
+
+\res{sure this would all be cool and interesting, but not going to happen for this paper. it is enough work on it's own to be a separate paper}. 
+\jri{see below, same here}
 
@@
 In the greenhouse experiments that were done, phenotyping was incomplete.
-Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture \res{we can't do this unless plants all go to seed} and internode length.
+Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture and internode length.
 Was there any evidence for variation in these other traits?
 In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.
+
 \res{no more phenotypine experiments}. 
+\jri{explain this is lots of work, beyond scope, decided not to do. add text to conclusions discussing future experiments, provide text here.}
 
-One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown. \res{is this necessary?}
+One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
+
+\res{is this necessary?}
+\jri{add to figure}
 
@@
 It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.
-\res{sure we knew it existed in some individuals, but not to the extent that we found in our study} Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
-This paper is fairly maize-centric.
-A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract.
-This could be easily remedied.
-\res{okay.
-but it is a paper about maize...}
+
+ \res{sure we knew it existed in some individuals, but not to the extent that we found in our study}
+\jri{explain point. add minor text to paper clarifying distinction, provide text}
+
+Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
+
+\jri{thank reviewer}
+
+This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.
+
+\res{okay. but it is a paper about maize...}
+\jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
+
 Methods and statistics were appropriately used as far as my expertise allows me to judge.
+
+\jri{thank reviewer}
+
 Specific suggestions follow:
@@
 Can you give some examples?
+
 \res{okay can do this.}
+\jri{change, provide new text}
 
-Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…" \res{doable}
+Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…"
+
+\res{doable}
+\jri{fix}
 
@@
 If the second, was it actually studied, or merely hypothesized in teosinte?
+
 \res{doable}
+\jri{explain in manuscript, provide new text}
 
@@
 Does the analysis of this group of teosintes add anything to your conclusions?
-\res{no mexicana phenotyping.
-seed limited.
-there were a couple populations of mex that had hop at intermediate frequency.}
+
+\res{no mexicana phenotyping. seed limited. there were a couple populations of mex that had hop at intermediate frequency.}
+\jri{clarify/tweak introduction, provide new text. explain here utility. mention other teosintes in discussion if not currently there.}
 
 Materials and Methods:
-Page 5, lines 22 - 23: how many individuals per accession?
-A fixed number, or variable, and depending on what?
-This can be mentioned so the reader does not have to go to the appendix or supplements.
-\res{no way we can list all those numbers in the text.
-varied anywhere from 2 to 30 is my guess.
-can say that?}
+Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.
 
-Page 5, Line 28: mention how big the hopscotch element is.
-Can it be amplified in one amplicon?
-Again, this would be useful to know without having to look in appendix or supplements.
+\res{no way we can list all those numbers in the text. varied anywhere from 2 to 30 is my guess. can say that?}
+\jri{add range and mean or median to text}
+
+Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.
+
 \res{i think this is obvious in the methods when we discuss primers sets}
+\jri{explicitly give size of hopscotch, note change to text}
 
@@
 maybe should say that in the text to clarify?}
+\jri{clarify in manuscript, provide new text}
 
-Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include? \res{can put this in text}
+Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include? 
+\res{can put this in text}
+\jri{add to text}
 
@@
 Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?
-\res{i have no idea.
-do either of you know if this is in tanja's paper?
-or should i look at the website she got data from?}
+\res{i have no idea. do either of you know if this is in tanja's paper? or should i look at the website she got data from?}
+\jri{1 year. cite huff paper that conditions didn't change that much over last 10K years. explain why we think using extant data best idea. clarify 1 year in text, mention change.}
 
-Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte? \res{can explain dataset in paper}
+Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte?
+
+\res{can explain dataset in paper}
+\jri{clarify in manuscript, provide new text}
 
 Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb?
-\res{hmm.
-Matt why did you use 5kb when you did this for your study?}
+
+\res{hmm. Matt why did you use 5kb when you did this for your study?}
+\jri{read Matt's paper, explain to reviewer}
 
@@
 Or you turned on the drip by hand?
-In which case, it was not automatic… \res{i'll just take out the by hand.
-it's too confusing to explain what happened}.
+In which case, it was not automatic…
+
+\res{i'll just take out the by hand. it's too confusing to explain what happened}. 
+\jri{no. clarify in manuscript, provide new text. some readers might care about watering regime.}
 
@@
 Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.
+
 \res{we wanted an independent trait to compare patterns of TI to}
+\jri{rewrite as suggested in manuscript, provide new text}
 
 Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols? \res{PCR}
+\jri{clarify in manuscript}
 
@@
 So we can see if this comes out to a very different ratio than in maize.
+
 \res{okay can put in ratios or something}
+\jri{clarify in manuscript}
 
@@
 alleles based on size difference in PCR}.
+\jri{explain (alleles in pop, not alleles per locus), clarify in manuscript}
 
@@
 \res{okay good suggestion}.
+\jri{no, bad suggestion. but do it anyway.}
 
-Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32. \res{okay}.
+Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32.
+
+\res{okay}
+\jri{do}
 

### 2014-07-07 ResponseToReview.tex
Edits
Edits to response

@@
 \newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
+\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
 \usepackage{lipsum}
@@
 \res{I vote we take out trees}
+\lev{I'd rather take them out. I don't think that they add a huge amount that isn't elsewhere in paper}
 
@@
 
-\res{we didn't know genotypes ahead of time...they were seeds collected from natural population} 
-\jri{write appropriate response, no or minor change to text for explanation}
+\res{We agree that a planned grow-out with previously determined genotypes would be useful, but were unable to do so due to our collections and available seed, as samples used for phenotyping were were collected as seed from a naturally occuring population in Jalisco, Mexico.}
 

### 2014-07-07 ResponseToReview.tex
Edits
Edits to text and response letter

@@
 
-%For my own part, I agree with both reviewers that the study would be of interest to readers of AJB.
-However, I took a more negative view towards several aspects of the Methods, Results, and Discussion.
-Three examples:
-
 1) I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.
@@
 Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?
-How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?
+How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\
 
-\res{We agree that a planned grow-out with previously determined genotypes would be useful, but were unable to do so due to our collections and available seed, as samples used for phenotyping were were collected as seed from a naturally occuring population in Jalisco, Mexico.}
+\res{We agree that a planned grow-out with previously determined genotypes would be useful, but were unable to do so due to our collections and available seed since samples used for phenotyping were were collected as seed from a naturally occuring population in Jalisco, Mexico.}
 
-Second, why are sample sizes not reported?
+3) Why are sample sizes not reported?
 You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class
 
-\res{I will check this. I thought it was listed in the text. maybe older version}.  
-\jri{fix, respond}
+\res{This has been addressed. We report accurate sample sizes for both phenotypings that are consistent throughout the text.}
 
-Third, it was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
+4) It was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
 9.

### 2014-07-07 manuscript.tex
Edits
Edits to text and response letter

@@
 
-\author[2]{Laura Vann}%author information
-\author[2,3]{Thomas Kono}
-\author[2,4]{Tanja Pyh\"aj\"arvi}
-\author[2,5,7]{Matthew B. Hufford}
-\author[2,6,7]{Jeffrey Ross-Ibarra}
-\affil[2]{Department of Plant Sciences, University of California Davis}
-\affil[3]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
-\affil[4]{Department of Biology, University of Oulu}
-\affil[5]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
-\affil[6]{Center for Population Biology and Genome Center, University of California Davis}
-\affil[7]{Author for Correspondence}
+\author[1]{Laura Vann}%author information
+\author[1,2]{Thomas Kono}
+\author[1,3]{Tanja Pyh\"aj\"arvi}
+\author[1,4,6]{Matthew B. Hufford}
+\author[1,5,6]{Jeffrey Ross-Ibarra}
+\affil[1]{Department of Plant Sciences, University of California Davis}
+\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
+\affil[3]{Department of Biology, University of Oulu}
+\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
+\affil[5]{Center for Population Biology and Genome Center, University of California Davis}
+\affil[6]{Author for Correspondence}
 \date{}
@@
 \textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
-We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
+We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 206 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
+Our plantings produced 20 homozygotes for the teosinte (no Hopscotch) allele, 104 heterozygotes, and 82 homozygotes for the maize (Hopscotch) allele.
 We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
@@
 We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.
-For the second grow-out we measured tillering index every five days through day 50 for 302 plants.
-We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele.
+For the second grow-out we measured tillering index every five days through day 50 for 247 plants.
+We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, with a 0.56 frequency of \emph{Hopscotch} homozygotes (138 individuals), a 0.2 frequency of homozygotes for the teosinte allele (49 individuals) and a 0.24 frequency of heterozygotes (60 individuals).
 Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.

### 2014-07-07 ResponseToReview.tex
edits
to methods and response

@@
 3) Why are sample sizes not reported?
-You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class
+You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\
 
@@
 Your Phenotyping 1 results show no effect of genotype on tillering.
-Thus, it is not clear what effect you are referring to.
+Thus, it is not clear what effect you are referring to.\\
 
@@
 
-Fourth, what was the purpose of the Phenotyping 2  experiment?
+5) What was the purpose of the Phenotyping 2  experiment?
 You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.

### 2014-07-07 manuscript.tex
edits
to methods and response

@@
 We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
-Based on these initial data, we conducted a \emph{post hoc} power analysis using data from day 40 of Phenotyping 1, indicating that a minimum of 71 individuals in each genotypic class would be needed to detect the observed effect of the \emph{Hopscotch} on tillering index.
+Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from \citep{Briggs2007}.
 

### 2014-07-07 ResponseToReview.tex
edits
edits

@@
 
-\jri{explain, clarify in text, give new text here}
+\res{We have clarified this in the text: "Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
 
 5) What was the purpose of the Phenotyping 2  experiment?
-You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.
+You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.\\
 
-\res{we did phenotyping 2 because we had such an odd distribution of genotypes in Phenotyping 1 that we had no power to make any inferences regarding effect of Hop on TI.
-We didn't sample an additional population aside from SLO for phenotyping -- is that what he is inferring?}
-\jri{explain, respond, add some clarifying text to document, provide new text}
+\res{You are correct that our Phenotyping 1 did not show an effect of the Hopscotch on tillering index; however, the distribution of genotype classes was so heavily skewed to individuals with the Hopscotch insertion (both heterozygotes and homozygotes) that we felt we may not have had enough individuals without the insertion to have been able to detect an effect.}
 
-3) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
+\res{Our sampling for phenotyping experiments was limited to a population from which we had a large number of seed collected, and we were unable to survey more than one population for these experiments.
+For our more broad genotyping of both landrace maize, ssp.
+\emph{parviglumis}, and ssp.
+\emph{mexicana} we sampled as many population as we could.
+For most of these populations the prevalence of the Hopscotch was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout these locations.}
+
+6) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
 It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
@@
 \jri{clean up, restructure, thank editor, explain changes}
-
-%Because of what I perceive to be substantial problems both with the methods and the presentation and discussion of results, and because Reviewer 1 raises some important issues with the phenotyping experiments, I am recommending that the manuscript be rejected.
-
-%Yours sincerely,
-%Andrew Schnabel
-%Associate Editor
 

### 2014-07-07 ResponseToReview.tex
edits
Edits for reviewer 1

@@
 
-They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate.
+1) They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate.
 However, I feel like they need to make some further effort to understand what is happening mechanistically.
@@
 Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.
-
-\res{sure this would all be cool and interesting, but not going to happen for this paper. it is enough work on it's own to be a separate paper}. 
-\jri{see below, same here}
 
@@
 Was there any evidence for variation in these other traits?
-In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.
+In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.\\
 
-\res{no more phenotypine experiments}. 
-\jri{explain this is lots of work, beyond scope, decided not to do. add text to conclusions discussing future experiments, provide text here.}
+\res{Expression analyses of \emph{tb1} and additional loci coupled with a more exhaustive phenotyping would be interesting and useful in understanding the role of \emph{tb1} in branching architecture in teosinte, but is beyond the scope of our study.
+The suggested experiments would in themselves be enough for an additional paper.
+While we are unable to include these experiments in this paper we have added text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte.}
 
-One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
+2) One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
 

### 2014-07-07 ResponseToReview.tex
edits
additional experiments discussed

@@
 The suggested experiments would in themselves be enough for an additional paper.
-While we are unable to include these experiments in this paper we have added text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte.}
+While we are unable to include these experiments in this paper we have text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: "Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
+\emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
+These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}
 

### 2014-07-08 ResponseToReview.tex
Edits for reviewer 2
Edits for reviewer 2

@@
 
-Reviewer #2: Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
-It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.
+3) Reviewer 2: Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
+It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.\\
 
- \res{sure we knew it existed in some individuals, but not to the extent that we found in our study}
 \jri{explain point. add minor text to paper clarifying distinction, provide text}
+
+\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some ssp.
+\emph{parviglumis}, ssp.
+\emph{mexicana}, and some landrace maize; however, previous studies did not sample as many populations or individuals per population as the study we present.
+Our study is not only concerned with whether or not the Hopscotch is present or absent in these populations but is also interested in what its frequency is so that we can better understand the evolutionary history of this insertion in teosinte.}
 

### 2014-07-09 ResponseToReview.tex
Reviewer 2 edits
Addressing reviewer 2 edits

@@
 
-\jri{explain point. add minor text to paper clarifying distinction, provide text}
-
 \res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some ssp.
 \emph{parviglumis}, ssp.
-\emph{mexicana}, and some landrace maize; however, previous studies did not sample as many populations or individuals per population as the study we present.
-Our study is not only concerned with whether or not the Hopscotch is present or absent in these populations but is also interested in what its frequency is so that we can better understand the evolutionary history of this insertion in teosinte.}
+\emph{mexicana}, and some landrace maize; however, previous studies did not sample as many populations or individuals per population as the study we present here.
+Our study is not only concerned with whether or not the \emph{Hopscotch} is present or absent in these populations but is also interested in what its frequency is so that we can better understand the evolutionary history of this insertion in teosinte.
+We have added clarifying text to the introduction: "The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.
+\emph{parviglumis}, ssp.
+\emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations."}
 
-Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
+4) Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.\\
 
-\jri{thank reviewer}
+\res{Thank you; we hope that our paper in conjunction with future research and experiments can help to further the understanding of the role of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}
 
-This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.
+5) This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
 
-\res{okay. but it is a paper about maize...}
 \jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
-
+\res{We have added sentences to the abstract and discussion in order to make 
 Methods and statistics were appropriately used as far as my expertise allows me to judge.

### 2014-07-09 manuscript.tex
edits
additional experiments discussed
Reviewer 2 edits
Addressing reviewer 2 edits

@@
 
-%TO DO
-%check citations in last paragraph for te1
-
-%preprint somewhere
-%finish fixing paragraph that jeff doesn't like
-%check map details
 %Code for SAS and command line stuff on Git
@@
 The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.
+\emph{parviglumis}, ssp.
+\emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
 
@@
 Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{te1}).
+\emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
 These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.

### 2014-07-09 ResponseToReview.tex
Reviewer 2 too maize like
There is already a sentence in the intro and discussion about tb1 in
other grasses. I am not adding anything else about this unless you both
feel it is absolutely necessary.

@@
 \jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
-\res{We have added sentences to the abstract and discussion in order to make 
-Methods and statistics were appropriately used as far as my expertise allows me to judge.
+\lev{Jeff & Matt, this is already in the intro.
+I don't see what more we can say that would be appropriate.
+"The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
 
-\jri{thank reviewer}
+6) Methods and statistics were appropriately used as far as my expertise allows me to judge.
+
+\res{Thank you.}.
 
@@
 
-Introduction:
+7) Introduction:
 Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".

### 2014-07-09 ResponseToReview.tex
Reviewer 2
More detail about wild alleles in domesticates; rufipogon traits into
sativa

@@
 Can you tell us why an understanding is useful?
-Can you give some examples?
+Can you give some examples?\\
 
-\res{okay can do this.}
-\jri{change, provide new text}
-
-Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…"
-
-\res{doable}
-\jri{fix}
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: "For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008).}
 

### 2014-07-09 manuscript.tex
Reviewer 2
More detail about wild alleles in domesticates; rufipogon traits into
sativa

@@
 Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-But while some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice {KovackMcCouch2008}.
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
 

### 2014-07-09 manuscript.tex
edits
edits

@@
 Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice {KovackMcCouch2008}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice {KovackMcCouch2008}.
 While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.

### 2014-07-09 ResponseToReview.tex
edits
reviewer 2

@@
 
-Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte?
+8) Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte?
 Which of the two distinct things have actually been studied?
@@
 
-\res{doable}
-\jri{explain in manuscript, provide new text}
+\res{We have clarified issues regarding the references for this section of the paper.}
 
-Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
+9) Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
 Did you phenotype mexicana?
 Can you present the data here, if so?
-Does the analysis of this group of teosintes add anything to your conclusions?
+Does the analysis of this group of teosintes add anything to your conclusions?\\
 
-\res{no mexicana phenotyping. seed limited. there were a couple populations of mex that had hop at intermediate frequency.}
-\jri{clarify/tweak introduction, provide new text. explain here utility. mention other teosintes in discussion if not currently there.}
+\res{Reviewer 2 is correct in noting that ssp.
+\emph{mexicana} is not as prevalent in our paper as ssp.
+\emph{parviglumis}.
+This fact is due to limitations in access to accessions of ssp.
+\emph{mexicana} as field collections for ssp.
+\emph{mexicana} are not nearly as extensive as those of ssp.
+\emph{parviglumis}.
+Additionally ssp.
+\emph{mexicana} are more difficult to cultivate in a greenhouse setting.
+We felt that it would be important to include ssp.
+\emph{mexicana} in our study, especially since it is known that gene flow occurs between ssp.
+\emph{mexicana} and landrace maize.
+We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.}
 
-Materials and Methods:
+10) Materials and Methods:
 Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.

### 2014-07-09 manuscript.tex
edits
reviewer 2

@@
 An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
-The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.
+The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} and may therefore be crucial to the ecology of teosinte.
 In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.

### 2014-07-10 ResponseToReview.tex
edits
edits

@@
 I don't see what more we can say that would be appropriate.
-"The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
+``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
 
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: "For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008).}
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008).}
 
@@
 If the first, could you tell us how tb1 has been shown to work in maize under shade conditions?
-If the second, was it actually studied, or merely hypothesized in teosinte?
+If the second, was it actually studied, or merely hypothesized in teosinte?\\
 
@@
 
-9) Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
+9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
 Did you phenotype mexicana?
@@
 10) Materials and Methods:
-Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.
+Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.\\
 
-\res{no way we can list all those numbers in the text. varied anywhere from 2 to 30 is my guess. can say that?}
-\jri{add range and mean or median to text}
+\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
 
-Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.
+11) Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.\\
 

### 2014-07-10 manuscript.tex
edits
edits

@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.

### 2014-07-10 ResponseToReview.tex
Reviewer 2 comments
Mostly materials and methods edits

@@
 
-\res{i think this is obvious in the methods when we discuss primers sets}
-\jri{explicitly give size of hopscotch, note change to text}
+\res{The size of the amplicon is listed in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}
 
-Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".
+12) Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".
 Sorry, what does this mean?
 That you have one band (one assumes the short one?).
-If you are talking presence/absence, how do you know the allele?
-\res{PCRs for different alleles.
-maybe should say that in the text to clarify?}
-\jri{clarify in manuscript, provide new text}
+If you are talking presence/absence, how do you know the allele?\\
 
-Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include? 
-\res{can put this in text}
-\jri{add to text}
+\res{We have clarified in the text.
+The 2 PCRs are for the two different alleles.
+If only one PCR resolved we only included one allele for that individual instead of trying to impute a diploid genotype.
+We have included the following sentence in the text: ``Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual."}
 
-Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis?
-Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?
+13) Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include?\\
+
+\res{We included 32 populations for these analyses, and have edited the text to reflect this number: ``These analyses only included populations in which eight or more individuals were sampled, totally 32 populations."}
+
+14) Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis?
+Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?\\
+
 \res{i have no idea. do either of you know if this is in tanja's paper? or should i look at the website she got data from?}

### 2014-07-10 manuscript.tex
Reviewer 2 comments
Mostly materials and methods edits

@@
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article).
-When only one PCR resolved well, we scored one allele for the individual.
+Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual.
 We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
@@
 \textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations in which eight or more individuals were sampled.
+These analyses only included populations in which eight or more individuals were sampled, totally 32 populations.
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.

### 2014-07-10 ResponseToReview.tex
Reviewer 2
Reviewer 2; addition of suggested references to refs.bib file

@@
 \jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
-\lev{Jeff & Matt, this is already in the intro.
+\lev{Jeff and Matt, this is already in the intro.
 I don't see what more we can say that would be appropriate.
@@
 
-\res{i have no idea. do either of you know if this is in tanja's paper? or should i look at the website she got data from?}
-\jri{1 year. cite huff paper that conditions didn't change that much over last 10K years. explain why we think using extant data best idea. clarify 1 year in text, mention change.}
+\res{The environmental daya used from {Pyh{\"a}j{\"a}rvi} \emph{et al.} (2013) was one year's worth of environmental data.
+A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.
+We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following {Pyh{\"a}j{\"a}rvi} \emph{et al.} (2013)."}
 
-Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte?
+15) Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte?\\
 
@@
 
-Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb?
+16) Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb?\\
 
@@
 
-Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?
+17) Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?
 Or you turned on the drip by hand?
-In which case, it was not automatic…
+In which case, it was not automatic… \\
 
@@
 
-Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1.
+18) Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1.
 F" So, why did you do it?
 Because of the last sentence in the materials and methods?
-Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.
+Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.\\
 
@@
 
-Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols? \res{PCR}
+19) Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols?\\
+
+\res{PCR}
 \jri{clarify in manuscript}
 
-Results:
+20) Results:
 Page 10 lines 5 - 6: "Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total?
-So we can see if this comes out to a very different ratio than in maize.
+So we can see if this comes out to a very different ratio than in maize.\\
 
@@
 
-Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
-I thought you only had presence or absence; how are alleles differentiated with your PCR?
-\res{can clarify.
-basically only used populations that had aboive a certain sample size.
-alleles based on size difference in PCR}.
+21) Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
+I thought you only had presence or absence; how are alleles differentiated with your PCR?\\
+
+\res{can clarify. basically only used populations that had aboive a certain sample size. alleles based on size difference in PCR}. 
 \jri{explain (alleles in pop, not alleles per locus), clarify in manuscript}
 
-Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
+22) Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
 Add Warburton, et al., 2011.
 Gene flow among different teosinte taxa and into the domesticated maize gene pool.
-Genet Resour Crop Evol 58:1243-1261.
-\res{okay good suggestion}.
-\jri{no, bad suggestion. but do it anyway.}
+Genet Resour Crop Evol 58:1243-1261.\\
 
-Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32.
+\res{We have added this reference as suggested.}
 
-\res{okay}
-\jri{do}
+23) Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32.\\
 
+\res{We have split the paragraph into two.}
 \end{document}

### 2014-07-10 manuscript.tex
Reviewer 2
Reviewer 2; addition of suggested references to refs.bib file

@@
 We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
+Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
 
@@
 In fact, early generation maize-teosinte hybrids are common in these populations today (MB Hufford, pers.
-observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{mexicana} and \emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011}.
+observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{mexicana} and \emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011, Warburton2011}.
 

### 2014-07-10 ResponseToReview.tex
Edits
Methods for BayEnv

@@
 
-\res{The environmental daya used from {Pyh{\"a}j{\"a}rvi} \emph{et al.} (2013) was one year's worth of environmental data.
+\res{The environmental daya used from Pyh{\"a}j{\"a}rvi \emph{et al.} (2013) was one year's worth of environmental data.
 A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.
-We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following {Pyh{\"a}j{\"a}rvi} \emph{et al.} (2013)."}
+We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh{\"a}j{\"a}rvi \emph{et al.} (2013)."}
 
@@
 
-\res{can explain dataset in paper}
-\jri{clarify in manuscript, provide new text}
+\res{We have clarified the dataset in the paper: ``We used teosinte (ssp.
+\emph{parviglumis} and ssp.
+\emph{mexicana} genotyping and covariance data from Pyh{\"a}j{\"a}rvi \emph{et al.} (2013) for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+SNP data from Pyh{\"a}j{\"a}rvi \emph{et al.} (2013) were obtained using the MaizeSNP50 BeadChip and Infinium® HD Assay (Illumina, San Diego, CA, USA)."}
 
@@
 
-\res{can clarify. basically only used populations that had aboive a certain sample size. alleles based on size difference in PCR}. 
-\jri{explain (alleles in pop, not alleles per locus), clarify in manuscript}
+\res{We believe we addressed this with your previous comment concerning uncertainty of the PCRs we were using.
+As edited into the Materials and Methods section, there is a PCR that produces a specific size for each allele.
+We have edited subsequent text accordingly.}
 

### 2014-07-10 manuscript.tex
Edits
Methods for BayEnv

@@
 BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
-We used genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+We used teosinte (ssp.
+\emph{parviglumis} and ssp.
+\emph{mexicana} genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium¨ HD Assay (Illumina, San Diego, CA, USA).
 Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
@@
 Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and four populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more alleles.
+\ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
@@
 Consistent with this interpretation, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.
+
 A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.

### 2014-07-10 ResponseToReview.tex
edits to phenotyping
phenotyping edits from reviewer 2

@@
 
-\res{i'll just take out the by hand. it's too confusing to explain what happened}. 
-\jri{no. clarify in manuscript, provide new text. some readers might care about watering regime.}
+\res{We have edited the text to clarify: ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
 
@@
 
-\res{we wanted an independent trait to compare patterns of TI to}
-\jri{rewrite as suggested in manuscript, provide new text}
+\res{Thank you for the suggestion, we have edited it in the text and added it to the end of the Materials and Methods section: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}
 

### 2014-07-10 manuscript.tex
edits to phenotyping
phenotyping edits from reviewer 2

@@
 Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
-Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer.
+Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days.
 
@@
 On day 65 we measured culm diameter between the third and fourth nodes of each plant.
-Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1}.
 Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol.
@@
 Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses.
 

### 2014-07-11 ResponseToReview.tex
Edits
working on reviewer 2 still. Edited .bib file — need to figure out how
to make it recompile.
edits from reviewer 2
edits to sampling details

@@
 
-\res{The size of the amplicon is listed in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{The size of the amplicon is listed in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}
 
@@
 
-\res{PCR}
-\jri{clarify in manuscript}
+\res{We have clarified this in the manuscript.}
 
@@
 
-\res{okay can put in ratios or something}
-\jri{clarify in manuscript}
+\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico.}"
 
@@
 
-\res{We believe we addressed this with your previous comment concerning uncertainty of the PCRs we were using.
+\res{We believe we addressed this with editing for your previous comment concerning uncertainty of the PCRs we were using.
 As edited into the Materials and Methods section, there is a PCR that produces a specific size for each allele.

### 2014-07-11 manuscript.tex
Edits
working on reviewer 2 still. Edited .bib file — need to figure out how
to make it recompile.
edits from reviewer 2
edits to sampling details

@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article).
 Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual.
@@
 Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol.
-We genotyped individuals for the \emph{Hopscotch} insertion following the protocols listed above.
+We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above.
 Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from \citep{Briggs2007}.
@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 and four populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.

### 2014-07-16 ResponseToReview.tex
Edits
Almost done!

@@
 
-\jri{please add two backslashes after every reviewer comment for spacing.
-happy to remove trees, but perhaps better we redo trees collapsing identical haplotypes.
-or  make a haplotype network with pie charts or something.}
-\res{I vote we take out trees}
 \lev{I'd rather take them out. I don't think that they add a huge amount that isn't elsewhere in paper}
@@
 The suggested experiments would in themselves be enough for an additional paper.
-While we are unable to include these experiments in this paper we have text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: "Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
+While we are unable to include these experiments in this paper we have text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
 \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
@@
 \jri{add to figure}
+\lev{Matt, do you still have ArcGIS on your computer?}
 
@@
 Our study is not only concerned with whether or not the \emph{Hopscotch} is present or absent in these populations but is also interested in what its frequency is so that we can better understand the evolutionary history of this insertion in teosinte.
-We have added clarifying text to the introduction: "The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+We have added clarifying text to the introduction: ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.
@@
 
-\res{Thank you.}.
+\res{Thank you.}
 
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008).}
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008)."}
 
@@
 Additionally ssp.
-\emph{mexicana} are more difficult to cultivate in a greenhouse setting.
+\emph{mexicana} is more difficult to cultivate in a greenhouse setting.
 We felt that it would be important to include ssp.
@@
 
-\res{The size of the amplicon is listed in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{The size of the amplicon is listed in now clarified in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}
 
@@
 
-\res{We included 32 populations for these analyses, and have edited the text to reflect this number: ``These analyses only included populations in which eight or more individuals were sampled, totally 32 populations."}
+\res{We included 32 populations for these analyses, and have edited the text to reflect this number: ``These analyses only included populations in which eight or more individuals were sampled, totaling 32 populations."}
 

### 2014-07-28 manuscript.tex
Edits
Almost done!
edits

@@
 \textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations in which eight or more individuals were sampled, totally 32 populations.
+These analyses only included populations in which eight or more individuals were sampled, totaling 32 populations.
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
@@
 We find no evidence of recent introgression in our analyses.
-Clustering in our NJ trees do not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
+Clustering in our NJ trees does not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
 Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig.

### 2014-07-30 manuscript.tex
edits
discussion

@@
 \end{centering}
+
+6) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
+It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
+
+\res{should we restructure this? I thought it read clearly. Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
+\jri{clean up, restructure, thank editor, explain changes}
+
 Adaptation occurs due to selection on standing variation or \emph{de novo} mutations.
@@
 
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci.
-We find no evidence of recent introgression in our analyses.
+While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results do not suggest introgression from maize into teosinte at the \emph{tb1} locus, and are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci.
 Clustering in our NJ trees does not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.

### 2014-08-13 ResponseToReview.tex

@@
 \renewcommand{\baselinestretch}{1.5}
+\usepackage{natbib} % package for formatting citations
+\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for
 
@@
 We would like to thank the Editor in Chief and the two anonymous reviewers for their helpful comments and suggestions.
-We have addressed the comments to the text, but while we feel that the suggested new experiments would provide interesting results, we worry that they would require considerable extra time (many months) and are outside of the scope of this current work.
+We have addressed the comments to the text, but while we feel that the suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
 We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses in blue below.
@@
 
-1) I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.
+I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.
 Why would you expect individuals with similar Hopscotch genotypes to cluster together?
@@
 
-\lev{I'd rather take them out. I don't think that they add a huge amount that isn't elsewhere in paper}
+\res{Our intention with the NJ trees was to show that there is no clear signal of the Hopscotch genotype at neighboring sequence.
+If the Hopscotch element were under selection in toesinte populations we might expect to see such a signal due to Hill-Robertson effects -- in fact the original discovery of Hopscotch was made using sequence that delineated sequences with and without the insertion but was almost 60Kb away from the Hopscotch itself.
+We agree with the reviewer that the trees are a bit confusing and difficult to interpret, and the large polytomies don't provide much information (beyond the lack of association with Hopscotch genotype).
+We have opted to keep the NJ trees in the supplement, where interested readers can view them but they don't unnecessarily complicate the main story.
+We have added explanation of the accession names as well.}
 
-2) Phenotyping experiments - These are curious experiments.
+\jri{need to add explanation of accession names in separate table or in figure legend.}
+
+Phenotyping experiments - These are curious experiments.
 First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?
@@
 
-\res{We agree that a planned grow-out with previously determined genotypes would be useful, but were unable to do so due to our collections and available seed since samples used for phenotyping were were collected as seed from a naturally occuring population in Jalisco, Mexico.}
+\res{We agree that a planned grow-out with previously determined genotypes would be useful, but our samples were collected as seed from a naturally occuring population.
+We know the Hopscotch has a phenotypic effect in a controlled isogenic background (and a preliminary experiment with a small sample of teosinte inbred lines also saw an effect), but we were interested in knowing whether the Hopsctoch had an effect in the segregating genetic backgrounds present in a natural population.
+This prevents us from using crosses.
+As leaf tissue can only be sampled from reasonably large seedligns and genotyping takes some time, we initially decided against trying to grow up many extra seedlings and genotype.
+This is, however, the approach we used for our second-growout to even sample sizes.
+\jri{true?} Note that because plants were situated in the greenhouse before genotyping, genotypes were effectively randomize with respect to position in the greenhouse.}
 
-3) Why are sample sizes not reported?
+Why are sample sizes not reported?
 You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\
 
-\res{This has been addressed. We report accurate sample sizes for both phenotypings that are consistent throughout the text.}
+\res{We have fixed the eror in reporting sample size, and have added sample size information for all three genotypes.} \jri{I didn't see sample sizes for each of the genotpyes.
+are they there?
+if not, please add.}
 
-4) It was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
+It was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
 9.
@@
 
-5) What was the purpose of the Phenotyping 2  experiment?
+What was the purpose of the Phenotyping 2  experiment?
 You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.\\
 
-\res{You are correct that our Phenotyping 1 did not show an effect of the Hopscotch on tillering index; however, the distribution of genotype classes was so heavily skewed to individuals with the Hopscotch insertion (both heterozygotes and homozygotes) that we felt we may not have had enough individuals without the insertion to have been able to detect an effect.}
+\res{Although Phenotyping 1 did not show an effect of the Hopscotch on tillering index, the distribution of genotype classes was  heavily skewed to individuals with the Hopscotch insertion (this relates to the reviewer's point \#2).
+We thus performed a second growout from the same population, including individuals from sample sites within the population where the Hopscotch was at low frequency to to increase the sample size of individuals without the insertion in hopes of better estimating effects.}
 
-\res{Our sampling for phenotyping experiments was limited to a population from which we had a large number of seed collected, and we were unable to survey more than one population for these experiments.
+\res{Note that all the sampling for both phenotyping experiments was limited to a single large population from which we had a large number of seed collected.
+Seed was collected from various "sampling sites" within the contiguous population at which we had estimated Hopscotch allele frequency.
 For our more broad genotyping of both landrace maize, ssp.
 \emph{parviglumis}, and ssp.
-\emph{mexicana} we sampled as many population as we could.
+\emph{mexicana} we sampled as many populations as we could.
 For most of these populations the prevalence of the Hopscotch was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout these locations.}
 
-6) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
+The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
 It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
 
-\res{should we restructure this? I thought it read clearly. Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
-\jri{clean up, restructure, thank editor, explain changes}
+\res{We have attempted to refocus this section.
+Our main points are that the Hopscotch allele is more widespread than previously thought, and that there is circumstantial evidence to suggest it may be under selection in natural teosinte populations.}
 
@@
 
-%Reviewer \#1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1.
+Reviewer \#1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1.
 The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis.
@@
 
-%\res{yes sounds right}
-
-1) They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate.
+They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate.
 However, I feel like they need to make some further effort to understand what is happening mechanistically.
@@
 Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.
+
+\res{We attempted to measure expression of \emph{tb1} following the protocol of \citep{hubbard2002expression} but were unable to get reasonable results.
+Subsequent discussion with colleagues revealed teosinte expression of \emph{tb1} is very difficult to detect except in early ear tissue.
+Extensive expression analysis of \emph{tb1} and a number of other genes is a good suggestion for follow-up work but seems to us to be a substantial undertaking and indeed more appropriate for another paper.
+}
 
@@
 
-\res{Expression analyses of \emph{tb1} and additional loci coupled with a more exhaustive phenotyping would be interesting and useful in understanding the role of \emph{tb1} in branching architecture in teosinte, but is beyond the scope of our study.
-The suggested experiments would in themselves be enough for an additional paper.
-While we are unable to include these experiments in this paper we have text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
+\res{We did not measure ear architecture or internode length.
+While \emph{tb1} does impact these traits, our primary interest was in the effect of  \emph{tb1} on tillering as a potentially ecologically relevant phenotype.}
+
+\res{Both additional greenhouse and expression analyses would indeed be useful avenuse of further investigation.
+Noentheless, these are extensive additonal experiments which we are unable to include at this time.
+We have however added text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
 \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
@@
 
-2) One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
+One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
 
-\res{is this necessary?}
-\jri{add to figure}
-\lev{Matt, do you still have ArcGIS on your computer?}
+\res{We have added this to the figure.} \jri{LV please add, don't wait for Matt.}
 
-3) Reviewer 2: Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
-It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.\\
+\section*{Reviewer 2} 
+Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
+It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.
+\jri{Is this all the intro Reviewer 2 gave?
+please copy/paste review verbatim if not}
 
@@
 \emph{parviglumis}, ssp.
-\emph{mexicana}, and some landrace maize; however, previous studies did not sample as many populations or individuals per population as the study we present here.
+\emph{mexicana}, and some landrace maize; however, previous studies did not sample as many populations or individuals per population as the study we present here and thus did not provide data of the frequency of the Hopscotch within populations.
 Our study is not only concerned with whether or not the \emph{Hopscotch} is present or absent in these populations but is also interested in what its frequency is so that we can better understand the evolutionary history of this insertion in teosinte.
@@
 \emph{parviglumis}, ssp.
-\emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations."}
+\emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations."} \jri{check for consistency in using emph with Hopscotch throughout}
 
-4) Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.\\
+Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.\\
 
@@
 
-5) This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
+This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
 
@@
 ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
+\jri{add some discussion about domesitcaiton traits in other species, weedy rice, etc. }
 
-6) Methods and statistics were appropriately used as far as my expertise allows me to judge.
-
-\res{Thank you.}
-
+Methods and statistics were appropriately used as far as my expertise allows me to judge.
 Specific suggestions follow:
 
-7) Introduction:
+Introduction:
 Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008)."}
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008)."} \jri{read matt's teosinte TIG paper, there are good cites there about use of teosinte specifically.}
 
@@
 \emph{mexicana} and landrace maize.
-We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.}
+We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.} \jri{just say we don't have as much seed from mexicana, which is true.
+it's just as easy to grow in the greenhouse.
+also I'd think of adding a sentence to the discussion mentioning mexicana.
+if it is in mexicana too, doesn't that suggest it's pretty old, given the divergence time between mexicana and parviglumis?
+or do we think it's in mexicana because of gene flow with parv or maize?
+this could be worth speculating on.}
 
@@
 
-\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)." \jri{I'd also add a median or mean, or note how many populations had >8 individuals or somet other way for them to get a sense of the data as a range by itself is hard to think about}}
+
+%JRI STOPPED HERE
 
@@
 
-\res{The environmental daya used from Pyh{\"a}j{\"a}rvi \emph{et al.} (2013) was one year's worth of environmental data.
+\res{The environmental daya used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
 A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.
-We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh{\"a}j{\"a}rvi \emph{et al.} (2013)."}
+We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh\"aj\"arvi  \emph{et al.} (2013)."}
 
@@
 \res{We have split the paragraph into two.}
+
+\bibliography{refs}
+
 \end{document}

### 2014-08-13 manuscript.tex

@@
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article).
-Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual.
+If only one of the two PCR reactions resolved well, we scored one allele for the individual.
 We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
@@
 SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium¨ HD Assay (Illumina, San Diego, CA, USA).
-Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
+Environmental data from a single year were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
 
@@
 We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above.
-Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from \citep{Briggs2007}.
+Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index.
 
@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (out of 86) and four (out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
@@
 Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.
+\jri{why did these numbers change?}
 

### 2014-08-13 ResponseToReview.tex
edits to response

@@
 \usepackage{lipsum}
+\usepackage{url}
 \usepackage{color}
@@
 
-\noindent Dear Dr. Andrew Schnabel and anonymous reviewers, \\
+\noindent Dear Dr. Schnabel and additional editorial staff at the \emph{American Journal of Botany}, \\
 
-We would like to thank the Editor in Chief and the two anonymous reviewers for their helpful comments and suggestions.
-We have addressed the comments to the text, but while we feel that the suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
-We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses in blue below.
-We hope that you are willing to consider our revised manuscript for resubmission.\\\
+We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript.
+We have addressed the majority of the issues raised in a revised version of our manuscript, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
+We have attached a revised version of our manuscript, including the suggested changes, and have detailed our responses in blue below.
+We hope that you are willing to consider our revised manuscript as a resubmission.\\\
 
@@
 
-Jeffrey Ross-Ibarra and Matthew Hufford
+\noindent Jeffrey Ross-Ibarra and Matthew Hufford
 
@@
 Why would you expect individuals with similar Hopscotch genotypes to cluster together?
-Also, although you don’t provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.
+Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.
 As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.
@@
 
-\section*{Reviewer 1}
+\section*{Response to Reviewer 1}
 
@@
 
-\section*{Reviewer 2} 
+\section*{Response to Reviewer 2} 
 Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
@@
 
-\res{The environmental daya used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
+\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
 A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.

### 2014-08-13 ResponseToReview.tex
edits to response

@@
 \newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
+\newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
 \usepackage{lipsum}
 \usepackage{url}
-\usepackage{color}
+\usepackage[usenames,dvipsnames]{color}
 \usepackage{multirow}
@@
 We have addressed the majority of the issues raised in a revised version of our manuscript, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
-We have attached a revised version of our manuscript, including the suggested changes, and have detailed our responses in blue below.
+We have attached a revised version of our manuscript, including the suggested changes, and have detailed our responses in blue text below.
 We hope that you are willing to consider our revised manuscript as a resubmission.\\\
@@
 
-\res{Our intention with the NJ trees was to show that there is no clear signal of the Hopscotch genotype at neighboring sequence.
-If the Hopscotch element were under selection in toesinte populations we might expect to see such a signal due to Hill-Robertson effects -- in fact the original discovery of Hopscotch was made using sequence that delineated sequences with and without the insertion but was almost 60Kb away from the Hopscotch itself.
-We agree with the reviewer that the trees are a bit confusing and difficult to interpret, and the large polytomies don't provide much information (beyond the lack of association with Hopscotch genotype).
-We have opted to keep the NJ trees in the supplement, where interested readers can view them but they don't unnecessarily complicate the main story.
-We have added explanation of the accession names as well.}
+\res{We agree entirely with this assessment of the Neighbor Joining trees. 
+Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
+Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
+In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}.
+\mbh{is this true?
+What's the reference here?}
+We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure across the \emph{tb1} genomic region that correlates with \emph{Hopscotch} genotype).
+We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
+In addition, we have added explanation of the accession names.}
 

### 2014-08-13 ResponseToReview.tex
edits to response

@@
 In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}.
-\mbh{is this true?
-What's the reference here?}
+\mbh{CITE}
 We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure across the \emph{tb1} genomic region that correlates with \emph{Hopscotch} genotype).
@@
 
-\jri{need to add explanation of accession names in separate table or in figure legend.}
+\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd say add this to the legend}
 
@@
 Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?
-How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\
+How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\
 
-\res{We agree that a planned grow-out with previously determined genotypes would be useful, but our samples were collected as seed from a naturally occuring population.
-We know the Hopscotch has a phenotypic effect in a controlled isogenic background (and a preliminary experiment with a small sample of teosinte inbred lines also saw an effect), but we were interested in knowing whether the Hopsctoch had an effect in the segregating genetic backgrounds present in a natural population.
-This prevents us from using crosses.
+\res{We agree that a planned grow-out with previously determined genotypes could be useful, but our interests lie in the effects of the \emph{Hopscotch} in the genetic background of individuals sampled from a naturally-occurring population.
+It has previously been shown that \emph{Hopscotch} has a phenotypic effect in a controlled isogenic background, \mbh{CITE} and our pilot studies with a small sample of teosinte inbred lines also revealed a minor effect.
 As leaf tissue can only be sampled from reasonably large seedligns and genotyping takes some time, we initially decided against trying to grow up many extra seedlings and genotype.

### 2014-08-14 ResponseToReview.tex
edits to response

@@
 We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript.
-We have addressed the majority of the issues raised in a revised version of our manuscript, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
-We have attached a revised version of our manuscript, including the suggested changes, and have detailed our responses in blue text below.
+In the attached revision, we have addressed the majority of issues raised, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
+In addition to our revision, we include a detailed response to reviewers below.
 We hope that you are willing to consider our revised manuscript as a resubmission.\\\
@@
 \mbh{CITE}
-We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure across the \emph{tb1} genomic region that correlates with \emph{Hopscotch} genotype).
+We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
 We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
@@
 
-\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd say add this to the legend}
+\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}
 
@@
 
-\res{We agree that a planned grow-out with previously determined genotypes could be useful, but our interests lie in the effects of the \emph{Hopscotch} in the genetic background of individuals sampled from a naturally-occurring population.
-It has previously been shown that \emph{Hopscotch} has a phenotypic effect in a controlled isogenic background, \mbh{CITE} and our pilot studies with a small sample of teosinte inbred lines also revealed a minor effect.
-As leaf tissue can only be sampled from reasonably large seedligns and genotyping takes some time, we initially decided against trying to grow up many extra seedlings and genotype.
-This is, however, the approach we used for our second-growout to even sample sizes.
-\jri{true?} Note that because plants were situated in the greenhouse before genotyping, genotypes were effectively randomize with respect to position in the greenhouse.}
+\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+Moreover, our sampling of individuals from this population for our grow out was not random.  
+We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include a mixture of individuals from both \emph{Hopscotch} positive and negative sites.
+This procedure resulted in a high frequency of both genotypes.
+We have clarified this methodology on page \mbh{X}, lines \mbh{XX}.
+Finally, plants were situated randomly in the greenhouse prior to genotyping.
+Regardless of whether genotype eventually became known, this still represents a randomized experimental design.}
 
@@
 
-\res{We have fixed the eror in reporting sample size, and have added sample size information for all three genotypes.} \jri{I didn't see sample sizes for each of the genotpyes.
-are they there?
-if not, please add.}
+\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}
 
-It was not clear to me what “to detect the observed effect” means in lines 12-13 of p.
+It was not clear to me what ``to detect the observed effect" means in lines 12-13 of p.
 9.
@@
 
-\res{We have clarified this in the text: "Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
+\res{We have clarified this in the text: ``Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1}-associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
+\mbh{I am concerned that originally our manuscript stated we conducted a power analysis based on our day 40 results (which is how I remembered the power analysis being done) and now it has been revised to say it was based on effect sizes in Briggs.
+These approaches seem mutually exclusive.
+Which did we use and why was this changed?}
 

### 2014-08-14 ResponseToReview.tex
edits to response

@@
 Finally, plants were situated randomly in the greenhouse prior to genotyping.
-Regardless of whether genotype eventually became known, this still represents a randomized experimental design.}
+Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}
 
@@
 \res{We have clarified this in the text: ``Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1}-associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
-\mbh{I am concerned that originally our manuscript stated we conducted a power analysis based on our day 40 results (which is how I remembered the power analysis being done) and now it has been revised to say it was based on effect sizes in Briggs.
-These approaches seem mutually exclusive.
-Which did we use and why was this changed?}
+\mbh{The previous edits here introduced a fair amount of confusion.
+Our submitted version indicated we used Day 40 data from phenotyping 1 for the power analysis to plan phenotyping 2, not Briggs.
+Jeff and I chatted for a long while and went through previous drafts and our understanding is that Day 40 data were used to plan phenotyping 2 and then a post hoc analysis was done with the Briggs data.
+I will edit the text to reflect this.
+Please be very careful not to make edits that inaccurately describe our methods.
+This could have caused us problems during the next round of review.}
 

### 2014-08-14 ResponseToReview.tex
edits to response

@@
 Moreover, our sampling of individuals from this population for our grow out was not random.  
-We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include a mixture of individuals from both \emph{Hopscotch} positive and negative sites.
-This procedure resulted in a high frequency of both genotypes.
-We have clarified this methodology on page \mbh{X}, lines \mbh{XX}.
+We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
+Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele.
+We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
+For the Phenotyping 2 experiment we therefore selected individuals from a mixture of \emph{Hopscotch} positive and negative sites.
+Despite this modified sampling approach, we still ended up with a higher frequency of \emph{Hopscotch} positive individuals, perhaps due to the fact that this allele appears to be under positive directional selection within the San Lorenzo population.
+We have clarified this sampling methodology on page \mbh{X}, lines \mbh{XX}.
 Finally, plants were situated randomly in the greenhouse prior to genotyping.
@@
 
-\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}
+\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX} (Phenotyping 1) and on page  \mbh{X} lines \mbh{XX} (Phenotyping 2).
+We have also corrected our typo for sample size in Phenotyping 1 and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes.
+are they there?
+if not, please add.}
 
@@
 \res{We have clarified this in the text: ``Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1}-associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
-\mbh{The previous edits here introduced a fair amount of confusion.
+\mbh{The previous edits here were problematic.
 Our submitted version indicated we used Day 40 data from phenotyping 1 for the power analysis to plan phenotyping 2, not Briggs.

### 2014-08-14 ResponseToReview.tex
edits to response

@@
 
-\res{We have clarified this in the text: ``Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1}-associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
-\mbh{The previous edits here were problematic.
-Our submitted version indicated we used Day 40 data from phenotyping 1 for the power analysis to plan phenotyping 2, not Briggs.
-Jeff and I chatted for a long while and went through previous drafts and our understanding is that Day 40 data were used to plan phenotyping 2 and then a post hoc analysis was done with the Briggs data.
-I will edit the text to reflect this.
-Please be very careful not to make edits that inaccurately describe our methods.
-This could have caused us problems during the next round of review.}
-
 What was the purpose of the Phenotyping 2  experiment?
@@
 
-\res{Although Phenotyping 1 did not show an effect of the Hopscotch on tillering index, the distribution of genotype classes was  heavily skewed to individuals with the Hopscotch insertion (this relates to the reviewer's point \#2).
-We thus performed a second growout from the same population, including individuals from sample sites within the population where the Hopscotch was at low frequency to to increase the sample size of individuals without the insertion in hopes of better estimating effects.}
+\res{On Day 40 of the Phenotyping 1 experiment we saw a weak positive correlation between the \emph{Hopscotch} insertion and tillering index (p=0.0848).  
+Using data from Phenotyping 1, we conducted a power analysis to determine the number of individuals we would need in both homozygous genotypic classes to observe a significant result.
+This analysis indicated we would need 71 individuals in both classes to detect a significant difference.
+We have clarified this on page \mbh{X}, lines \mbh{X}.}
+
+\mbh{I don't think the next paragraph is necessary}
 
@@
 
-\res{We have attempted to refocus this section.
-Our main points are that the Hopscotch allele is more widespread than previously thought, and that there is circumstantial evidence to suggest it may be under selection in natural teosinte populations.}
+\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte}
 

### 2014-08-15 ResponseToReview.tex
edits to response

@@
 
-\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte}
+\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype of reduced tiller number, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte.
+Surprisingly, tiller number is not a major phenotypic difference between teosinte plants with and without the \emph{Hopscotch} insertion, perhaps indicating the insertion affects multiple phenotypes in teosinte}
 
@@
 
-Reviewer \#1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1.
+Reviewer \#1: In their paper `Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1.
 The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis.
@@
 Subsequent discussion with colleagues revealed teosinte expression of \emph{tb1} is very difficult to detect except in early ear tissue.
-Extensive expression analysis of \emph{tb1} and a number of other genes is a good suggestion for follow-up work but seems to us to be a substantial undertaking and indeed more appropriate for another paper.
+Extensive expression analysis of \emph{tb1} and a number of other genes is a good suggestion for follow-up work but would be a substantial undertaking and indeed more appropriate for another paper.
 }
@@
 
-\res{Both additional greenhouse and expression analyses would indeed be useful avenuse of further investigation.
-Noentheless, these are extensive additonal experiments which we are unable to include at this time.
-We have however added text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
+\res{Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.
+Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
+We have, however, added text to the conclusion (\mbh{Page X, Lines X}) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
+``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
 \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
@@
 
-\res{We have added this to the figure.} \jri{LV please add, don't wait for Matt.}
+\res{We have added the Balsas River Basin in a revised version of Figure 1.} \jri{LV please add, don't wait for Matt.} \mbh{see the R tutorial I sent, LV}
 
@@
 It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.
-\jri{Is this all the intro Reviewer 2 gave?
-please copy/paste review verbatim if not}
 
-\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some ssp.
-\emph{parviglumis}, ssp.
-\emph{mexicana}, and some landrace maize; however, previous studies did not sample as many populations or individuals per population as the study we present here and thus did not provide data of the frequency of the Hopscotch within populations.
-Our study is not only concerned with whether or not the \emph{Hopscotch} is present or absent in these populations but is also interested in what its frequency is so that we can better understand the evolutionary history of this insertion in teosinte.
+\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
+Our study is not only concerned with presence or absence of the \emph{Hopscotch} allele in these populations but also the frequency at which it segregates.  
+This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion in teosinte. 
 We have added clarifying text to the introduction: ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
@@
 \emph{parviglumis}, ssp.
-\emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations."} \jri{check for consistency in using emph with Hopscotch throughout}
+\emph{mexicana}, and landrace maize; yet little is known about the frequency at which the \emph{Hopscotch} is segregating in natural populations."}
 
@@
 
-\res{Thank you; we hope that our paper in conjunction with future research and experiments can help to further the understanding of the role of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}
+\res{Thank you; we hope that our paper in conjunction with future research and experiments can help to further the understanding of ecological significance of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}
 

### 2014-08-18 ResponseToReview.tex
Comments in response
Comments regarding mex

@@
 this could be worth speculating on.}
+\lev{Hmm.
+Well didn't Matt's paper on introgression support a significant lack of gene flow from maize to mex around domestication loci?
+So how likely would it be that it is in mex because of gene flow?
+So would it make more sense to argue that it's there because it is old and hasn't been lost?
+Would it make sense too because it was estimated to predate the domestication of maize?
+Maybe it is in parv at a higher frequency because it's 'doing' something, even though we didn't pick up an effect on tillering, and maybe it's just drifting in maize, which could explain why it is at a much lower frequency in mex?}
 

### 2014-08-18 ResponseToReview.tex
Response
minor changes

@@
 20) Results:
-Page 10 lines 5 - 6: "Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total?
-So we can see if this comes out to a very different ratio than in maize.\\
+Page 10 lines 5 - 6: ``Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total?
+So we can see if this comes out to a very different ratio than in maize."\\
 
-\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico.}"
+\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
 
-21) Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
+21) Page 10 lines 7 - 9: ``Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
 I thought you only had presence or absence; how are alleles differentiated with your PCR?\\
@@
 
-22) Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
+22) Page 13 lines 25 - 27: ``and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
 Add Warburton, et al., 2011.

### 2014-08-20 manuscript.tex
edits to response
Removing phenotyping 2
Edits to manuscript to reflect removal of phenotyping 2 for
resubmission to AJB

@@
 Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte.
-Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
+Finally, a greenhouse experiment with teosinte does not suggest \emph{tb1} controls tillering in natural populations.
 \vspace*{0.2cm}
@@
 
-\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1).
-We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
+\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
 In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
@@
 We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above.
-Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index.
-
-We performed a second phenotyping experiment (Phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the \emph{Hopscotch} insertion.
-Seeds were germinated and planted on day 7 post fruit-case removal into two gallon pots.
-Plants were watered twice daily, alternating between fertilized and non-fertilized water.
-We began phenotyping successful germinations (302 plants) for tillering index on day 15 post fruit-case removal, and phenotyped every five days until day 50.
-At day 50 we measured culm diameter between the third and fourth nodes.
-We extracted DNA and genotyped plants following the same guidelines as in Phenotyping 1.
+\lev{should we take this bit about power analysis out?} Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index.
 
@@
 
-\res{should we restructure this? I thought it read clearly. Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
+\lev{should we restructure this? I thought it read clearly. Key points were hop more widely spread than previously and evidence of selection on Hop in two}
+
 \jri{clean up, restructure, thank editor, explain changes}

### 2014-08-20 manuscript.tex
Phenotyping 2 removal
Edits for removing phenotyping 2. Redone Figure 4 with box plots for
phenotyping 1 only.

@@
 \ref{Fig4Boxplots}), tiller number, or culm diameter.
-
-We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.
-For the second grow-out we measured tillering index every five days through day 50 for 247 plants.
-We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, with a 0.56 frequency of \emph{Hopscotch} homozygotes (138 individuals), a 0.2 frequency of homozygotes for the teosinte allele (49 individuals) and a 0.24 frequency of heterozygotes (60 individuals).
-Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.
-\jri{why did these numbers change?}
+However, we do observe a near significant (p<xxx) correlation between tillering index and genotype on day 40, though in the opposite direction of what is expected based on how the element acts phenotypically in maize.
 
@@
 \end{centering}
-
-6) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
-It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
-
-\lev{should we restructure this? I thought it read clearly. Key points were hop more widely spread than previously and evidence of selection on Hop in two}
-
-\jri{clean up, restructure, thank editor, explain changes}
 
@@
   \begin{center}
-   %\includegraphics[width=90mm]{Fig4Boxplots.pdf}
+   %\includegraphics[width=90mm]{Fig4.BoxplotPheno1.pdf}
    \end{center}
-    \caption{Box-plots showing tillering index in greenhouse grow-outs for Phenotyping 1 (A) and Phenotyping 2 (B).
+    \caption{Box-plots showing tillering index in greenhouse grow-outs for phenotyping.
 White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.

### 2014-08-20 manuscript.tex
Removal of phenotyping 2
Edits to figure 1 (map), figure legends, and text
Edits
Edits concerning removal of phenotyping 2 dataset

@@
 \ref{Fig4Boxplots}), tiller number, or culm diameter.
-However, we do observe a near significant (p<xxx) correlation between tillering index and genotype on day 40, though in the opposite direction of what is expected based on how the element acts phenotypically in maize.
+However, we do observe a near significant ($p<0.08$) correlation between tillering index and genotype on day 40, though in the opposite direction of what is expected based on how the element acts phenotypically in maize.
 
@@
 \begin{center}
-   %\includegraphics[width=0.6\textwidth]{Fig1Map.pdf}
+   %\includegraphics[width=0.6\textwidth]{Fig1.MapBalsas.pdf}
    \end{center}
     \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
-Size of circles reflects number of alleles sampled.}
+Size of circles reflects number of alleles sampled.
+The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.}
 \label{Fig1Map}

### 2014-08-21 ResponseToReview.tex
edits to response

@@
 \res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
-Our study is not only concerned with presence or absence of the \emph{Hopscotch} allele in these populations but also the frequency at which it segregates.  
+Our study is not only concerned with the presence or absence of the \emph{Hopscotch} allele in these populations but also the frequency at which it segregates.  
 This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion in teosinte. 

### 2014-08-21 manuscript.tex

@@
 Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
-Phenotypic data were also collected from a single teosinte population where \emph{Hopscotch} was found segregating.
+Phenotypic data were also collected using seed from a single teosinte population where \emph{Hopscotch} was found segregating.
 
@@
 Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte.
-Finally, a greenhouse experiment with teosinte does not suggest \emph{tb1} controls tillering in natural populations.
+Finally, a greenhouse experiment with teosinte does not suggest \emph{tb1} controls tillering in teosinte.
 \vspace*{0.2cm}
@@
 In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
-We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus.
+We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} combined with a larger SNP dataset in a subset of teosinte populations to explore patterns of genetic variation at this locus.
 Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.
@@
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
-We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion.
+We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb).
 Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
 If only one of the two PCR reactions resolved well, we scored one allele for the individual.

### 2014-08-22 ResponseToReview.tex
edits to the response

@@
 
-\noindent Dear Dr. Schnabel and additional editorial staff at the \emph{American Journal of Botany}, \\
+\noindent Dear Dr. Schnabel and editorial staff at the \emph{American Journal of Botany}, \\
 
@@
 Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
-Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
+Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
 In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}.
@@
 We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
-In addition, we have added explanation of the accession names.}
+In addition, we have added explanation of the accession names in the supplemental figure legend.}
 
@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-For the Phenotyping 2 experiment we therefore selected individuals from a mixture of \emph{Hopscotch} positive and negative sites.
-Despite this modified sampling approach, we still ended up with a higher frequency of \emph{Hopscotch} positive individuals, perhaps due to the fact that this allele appears to be under positive directional selection within the San Lorenzo population.
-We have clarified this sampling methodology on page \mbh{X}, lines \mbh{XX}.
+We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}.
 Finally, plants were situated randomly in the greenhouse prior to genotyping.
@@
 
-\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX} (Phenotyping 1) and on page  \mbh{X} lines \mbh{XX} (Phenotyping 2).  
-We have also corrected our typo for sample size in Phenotyping 1 and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes.
+\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.
+We have also corrected our typo for sample size and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes.
 are they there?
@@
 Using data from Phenotyping 1, we conducted a power analysis to determine the number of individuals we would need in both homozygous genotypic classes to observe a significant result.
-This analysis indicated we would need 71 individuals in both classes to detect a significant difference.
-We have clarified this on page \mbh{X}, lines \mbh{X}.}
+This analysis indicated we would need 71 individuals in both classes to detect a significant difference in tillering index.  
+We attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) where we had already shown \emph{Hopscotch} to be segregating at high frequency in order to detect a significant difference between homozygous genotypes.
+Our Phenotyping 2 experiment detects the same trend reported in Phenotyping 1, but the trend is still not significant.
+We agree with the editor that this second, non-significant result adds little to the manuscript and we have thus removed Phenotyping 2 and refer to a single grow-out in our revised version of the manuscript.}
 
@@
 \res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype of reduced tiller number, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte.
-Surprisingly, tiller number is not a major phenotypic difference between teosinte plants with and without the \emph{Hopscotch} insertion, perhaps indicating the insertion affects multiple phenotypes in teosinte}
+Surprisingly, based on our grow-out of individuals from the San Lorenzo population, tiller number is not a major phenotypic difference between teosinte plants with and without the \emph{Hopscotch} insertion, perhaps indicating the insertion affects multiple phenotypes in teosinte}
 
@@
 \res{We attempted to measure expression of \emph{tb1} following the protocol of \citep{hubbard2002expression} but were unable to get reasonable results. 
-Subsequent discussion with colleagues revealed teosinte expression of \emph{tb1} is very difficult to detect except in early ear tissue.
-Extensive expression analysis of \emph{tb1} and a number of other genes is a good suggestion for follow-up work but would be a substantial undertaking and indeed more appropriate for another paper.
+Subsequent discussion with colleagues revealed teosinte expression of \emph{tb1} is very difficult to detect except in early ear tissue.}
+
+\mbh{This reads as a pretty weak response...we tried to measure expression, we got weird results that we don't report, and our colleagues tells us this is hard.
+I think I'd lean toward leaving this out and just reemphasizing that expression experiments are beyond the scope of the current work}
+
+\res{Extensive expression analysis of individuals with and without the \emph{Hopscotch} insertion at \emph{tb1} and a number of other genes is a good suggestion for follow-up work but would be a substantial undertaking and indeed more appropriate for another paper.
 }
@@
 \res{We did not measure ear architecture or internode length. 
-While \emph{tb1} does impact these traits, our primary interest was in the effect of \emph{tb1} on tillering as a potentially ecologically relevant phenotype.}
-
-\res{Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.
+While \emph{tb1} does impact these traits, our primary interest was in the effect of \emph{tb1} on tillering as a potentially ecologically relevant phenotype given its known substantial effects on this trait in a maize background.
+Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.
 Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
@@
 
-\res{We have added the Balsas River Basin in a revised version of Figure 1.} \jri{LV please add, don't wait for Matt.} \mbh{see the R tutorial I sent, LV}
+\res{We have added the Balsas River in a revised version of Figure 1.}
 

### 2014-08-22 ResponseToReview.tex
edits
edits to NJ tree stuff and citations

@@
 Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
-In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}.
-\mbh{CITE}
+In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} (Studer \emph{et al.} 2011).
 We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.

### 2014-08-22 ResponseToReview.tex
edits to response

--- 2014-08-22 ResponseToReview.tex
@@
 
-\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
-Our study is not only concerned with the presence or absence of the \emph{Hopscotch} allele in these populations but also the frequency at which it segregates.  
-This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion in teosinte. 
+\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals \mbh{CITE}; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
+Our study is not only concerned with the presence or absence of the \emph{Hopscotch} allele in teosinte but also the frequency at which it segregates in a range-wide sample of teosinte.  
+This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion (\emph{i.e.}, the domestication allele at \emph{tb1}) in teosinte. 
 We have added clarifying text to the introduction: ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
@@
 
-\res{Thank you; we hope that our paper in conjunction with future research and experiments can help to further the understanding of ecological significance of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}
+\res{Thank you; we hope that our paper in conjunction with future research and experiments can help shed light on the ecological significance of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}
 
@@
 \jri{add some discussion about domesitcaiton traits in other species, weedy rice, etc. }
+
+\mbh{Yes, I think we could include more text on traits commonly under selection during domestication (\emph{i.e.}, the domestication syndrome) and perhaps highlight how these are relevant to the fitness of wild relatives and could have been the targets of natural selection prior to domestication.
+In my opinion there is a false dichotomy out there regarding "natural" vs.
+"artificial" selection.
+Give this a shot, LV, and I'll edit with my POV.}
 
@@
 
+\mbh{lots of fodder in this manuscript as well, LV: \url{http://bioscience.oxfordjournals.org/content/61/7/559.full}}
+
 8) Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte?
@@
 
-\res{We have clarified issues regarding the references for this section of the paper.}
+\res{We have moved the references cited here to more clearly reflect the papers' findings.
+We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte}
 
-9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
+\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness}
+
+9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic effects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
 Did you phenotype mexicana?

### 2014-08-22 ResponseToReview.tex
Associate editor edits
Edits in response to the associate editor

@@
 
-\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}
-
 Phenotyping experiments - These are curious experiments.
@@
 \res{We agree that a planned grow-out with previously determined genotypes could be useful.  
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
 Moreover, our sampling of individuals from this population for our grow out was not random.  
@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}.
-Finally, plants were situated randomly in the greenhouse prior to genotyping.
+We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}. Finally, plants were situated randomly in the greenhouse prior to genotyping.
 Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

### 2014-08-22 ResponseToReview.tex
Associate editor
edits to AE response

@@
 \res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.
+\lev{will fix the line numbers when we've made sure the manuscript text is okay to go}.
 We have also corrected our typo for sample size and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes.
 are they there?
-if not, please add.}
+if not, please add.} \lev{they should be in the results section.}
 
@@
 
-\mbh{I don't think the next paragraph is necessary}
+\mbh{I don't think the next paragraph is necessary} \lev{which next one? the one that starts 'Note that all the sampling for both...}
 

### 2014-08-25 ResponseToReview.tex
Methods, AE
Associate editor edits to methods
Associate editor response edits
Edits to response to associate editor

--- 2014-08-22 ResponseToReview.tex
@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}. Finally, plants were situated randomly in the greenhouse prior to genotyping.
+We have clarified our sampling methodology on pages 8-9, lines 174-185. Finally, plants were situated randomly in the greenhouse prior to genotyping.
 Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}
@@
 
-\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.
-We have also corrected our typo for sample size and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes.
-are they there?
-if not, please add.}
+\res{We have reported sample size information for all three genotypes on page 11 lines 266-267.}
 
@@
 We agree with the editor that this second, non-significant result adds little to the manuscript and we have thus removed Phenotyping 2 and refer to a single grow-out in our revised version of the manuscript.}
-
-\mbh{I don't think the next paragraph is necessary}
 

### 2014-08-25 ResponseToReview.tex
Response to reviewer 1 edits
Edits to response to reviewer 1. removed comments & updated page/line
numbers to reflect manuscript edits

@@
 The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.
-It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.
+It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.\\
 
@@
 At a minimum, qRT-PCR experiments should be performed assessing tb1 expression in their with/without Hopscotch lines.
-Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.
-
-\res{We attempted to measure expression of \emph{tb1} following the protocol of \citep{hubbard2002expression} but were unable to get reasonable results. 
-Subsequent discussion with colleagues revealed teosinte expression of \emph{tb1} is very difficult to detect except in early ear tissue.}
-
-\mbh{This reads as a pretty weak response...we tried to measure expression, we got weird results that we don't report, and our colleagues tells us this is hard.
-I think I'd lean toward leaving this out and just reemphasizing that expression experiments are beyond the scope of the current work}
+Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.\\
 
@@
 Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
-We have, however, added text to the conclusion (\mbh{Page X, Lines X}) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
+We have, however, added text to the conclusion (Page 15, 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
 ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.

### 2014-08-25 ResponseToReview.tex
Edits to Response to review
Edits to response to reviewer 2

@@
 Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
-It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.
+It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.\\
 
-\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals \mbh{CITE}; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
+\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals (Studer \emph{et al.} 2011); however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
 Our study is not only concerned with the presence or absence of the \emph{Hopscotch} allele in teosinte but also the frequency at which it segregates in a range-wide sample of teosinte.  
 This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion (\emph{i.e.}, the domestication allele at \emph{tb1}) in teosinte. 
-We have added clarifying text to the introduction: ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+We have added clarifying text to the introduction (Page 5, lines 55-61): ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.

### 2014-08-25 ResponseToReview.tex
edits to response

--- 2014-08-25 ResponseToReview.tex
@@
 Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
-In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} (Studer \emph{et al.} 2011).
+In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}.
+\mbh{CITE}
 We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
@@
 In addition, we have added explanation of the accession names in the supplemental figure legend.}
+
+\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}
 
@@
 \res{We agree that a planned grow-out with previously determined genotypes could be useful.  
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
 Moreover, our sampling of individuals from this population for our grow out was not random.  
@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-We have clarified our sampling methodology on pages 8-9, lines 174-185. Finally, plants were situated randomly in the greenhouse prior to genotyping.
+We have clarified our sampling methodology on page \mbh{X}, lines \mbh{XX}.
+Finally, plants were situated randomly in the greenhouse prior to genotyping.
 Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}
@@
 
-\res{We have reported sample size information for all three genotypes on page 11 lines 266-267.}
+\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.
+We have also corrected our typo for sample size and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes.
+are they there?
+if not, please add.}
 
@@
 We agree with the editor that this second, non-significant result adds little to the manuscript and we have thus removed Phenotyping 2 and refer to a single grow-out in our revised version of the manuscript.}
+
+\mbh{I don't think the next paragraph is necessary}
 
@@
 
+\mbh{I'll take a stab at this...pretty sure it was at low frequency in all \emph{mexicana} for which we had population sampling.
+It would not have made sense to use this for phenotyping because we would have had to sample a ton of individuals in order to compare \emph{Hopscotch} phenotype to no-\emph{Hopscotch} phenotype.}
+
 10) Materials and Methods:
@@
 \res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)." \jri{I'd also add a median or mean, or note how many populations had >8 individuals or somet other way for them to get a sense of the data as a range by itself is hard to think about}}
+
+\mbh{I agree with Jeff here}
 

### 2014-08-25 Supplement_mbh.tex
adding comments to manuscript file

@@
 
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}
+%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}
 
@@
 
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
+%\pagenumbering{arabic}
+%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
 % Appendix S2. Mays sampling
@@
 \clearpage
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
+%\pagenumbering{arabic}
+%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
 % Appendix S3. Cartoon representation of tb1 locus
@@
 \clearpage
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}
+%\pagenumbering{arabic}
+%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}
 
@@
 \clearpage
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}
+%\pagenumbering{arabic}
+%\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}
 
@@
 TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion.
-Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
+Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.
+EjuA refers to individuals from population Ejutla A, EjuB from Ejutla B, SLO from San Lorenzo, and MSA from La Mesa.
+Remaining individuals are lines of maize (\emph{Zea mays} ssp.
+\emph{mays}).
   \end{center}

### 2014-08-26 ResponseToReview.tex
response to review edits
edits to response to reviewer 2; introduction mostly

@@
 
-\jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
-\lev{Jeff and Matt, this is already in the intro.
-I don't see what more we can say that would be appropriate.
-``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
-\jri{add some discussion about domesitcaiton traits in other species, weedy rice, etc. }
-
-\mbh{Yes, I think we could include more text on traits commonly under selection during domestication (\emph{i.e.}, the domestication syndrome) and perhaps highlight how these are relevant to the fitness of wild relatives and could have been the targets of natural selection prior to domestication.
-In my opinion there is a false dichotomy out there regarding "natural" vs.
-"artificial" selection.
-Give this a shot, LV, and I'll edit with my POV.}
-
 Methods and statistics were appropriately used as far as my expertise allows me to judge.

### 2014-08-26 manuscript.tex
response to review edits
edits to response to reviewer 2; introduction mostly

@@
 Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
+Plant domestication usually involves a suite of phenotypic changes, such as increase yield, loss of seed shattering, and increase fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}.
 Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice {KovackMcCouch2008}.
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+\lev{teosinte example here} While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
 

### 2014-08-26 ResponseToReview.tex
edits to response

--- 2014-08-25 ResponseToReview.tex
@@
 
-\res{The size of the amplicon is listed in now clarified in the Materials and Methods following the primer sequences: ``Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page \mbh{X}, lines \mbh{XX}) following primer sequences: 
+``Homozygotes show a single band for absence of the \emph{Hopscotch} element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."
+The 5kb amplicon including the entire \emph{Hopscotch} could be amplified in a single reaction.}
 
@@
 
-\res{We have clarified in the text.
-The 2 PCRs are for the two different alleles.
-If only one PCR resolved we only included one allele for that individual instead of trying to impute a diploid genotype.
-We have included the following sentence in the text: ``Since we had a PCR for each allele, if only one PCR resolved well, we scored one allele for the individual."}
+\res{We have clarified this in the text on page \mbh{X}, lines \mbh{XX}.
+We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
+no \emph{Hopscotch}).
+If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
+We have included the following sentence in the text: \mbh{update the following in the text} ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
 
@@
 
-\res{We included 32 populations for these analyses, and have edited the text to reflect this number: ``These analyses only included populations in which eight or more individuals were sampled, totaling 32 populations."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page \mbh{X}, line \mbh{XX}: \mbh{Update text in manuscript}``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
 
@@
 
-\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
+%\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
 A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.
 We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh\"aj\"arvi  \emph{et al.} (2013)."}
+
+\res{Our climatic data consisted of the 19 bioclim layers commonly used in ecological niche analyses.
+Full methods for generation of these data can be found at the following url: \url{http://www.worldclim.org/methods}.
+Briefly, temperature and precipitation were averaged for each global location over a minimum of ten years during the period from 1950-2000.
+Soil data were downloaded from the Harmonized World Soil Database (FAOHWSD; \url{http://webarchive.iiasa.ac.at/Research/LUC/External-World-soil-database/HWSD_Documentation.pdf}).
+This recently published data set drew together global soil maps generated by a world-wide network of scientists over the last several decades and summarized them in GIS data layers that are available as polygon coverages.
+Data for our sampling sites were extracted from these data sets based on latitude and longitude using the DIVA-GIS platform.
+Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
+We have further clarified this on page \mbh{X}, lines \mbh{XX} of the text.}
 

### 2014-08-26 ResponseToReview.tex

@@
 This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
-
-\jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
-\lev{Jeff and Matt, this is already in the intro.
-I don't see what more we can say that would be appropriate.
-``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}
-\jri{add some discussion about domesitcaiton traits in other species, weedy rice, etc. }
-
-\mbh{Yes, I think we could include more text on traits commonly under selection during domestication (\emph{i.e.}, the domestication syndrome) and perhaps highlight how these are relevant to the fitness of wild relatives and could have been the targets of natural selection prior to domestication.
-In my opinion there is a false dichotomy out there regarding "natural" vs.
-"artificial" selection.
-Give this a shot, LV, and I'll edit with my POV.}
 

### 2014-08-26 ResponseToReview.tex
Edits to introduction
Response to reviewer 2, additional teosinte examples in introduction

@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008)."} \jri{read matt's teosinte TIG paper, there are good cites there about use of teosinte specifically.}
-
-\mbh{lots of fodder in this manuscript as well, LV: \url{http://bioscience.oxfordjournals.org/content/61/7/559.full}}
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008).
+Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010).
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.} Bai and Lindhout 2007), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors (Whitton \emph{et al.} 1997).}
 

### 2014-08-26 manuscript.tex
Edits to introduction
Response to reviewer 2, additional teosinte examples in introduction

@@
 For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
-\lev{teosinte example here} While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
+Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
 

### 2014-08-27 ResponseToReview.tex
edits
edits to response to reviewer 2; introduction & SAS sections

@@
 
-\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness}
+\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness} \lev{the refs are in the sentences discussing the SAS in teo/grasses}
 
@@
 Would it make sense too because it was estimated to predate the domestication of maize?
-Maybe it is in parv at a higher frequency because it's 'doing' something, even though we didn't pick up an effect on tillering, and maybe it's just drifting in maize, which could explain why it is at a much lower frequency in mex?}
+Maybe it is in parv at a higher frequency because it's 'doing' something, even though we didn't pick up an effect on tillering, and maybe it's just drifting in mex, which could explain why it is at a much lower frequency in mex?}
 

### 2014-08-27 manuscript.tex
edits
edits to response to reviewer 2; introduction & SAS sections

@@
 An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
-The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} and may therefore be crucial to the ecology of teosinte.
+The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading, and may therefore be crucial to the ecology of teosinte.
 In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.

### 2014-08-27 manuscript.tex
Edits to response to reviewer 2
Additional details concerning tb1 in the shade avoidance pathway in
teosinte, discussion of importance to ecology of teosinte

--- 2014-08-26 manuscript.tex
@@
 An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
-The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} and may therefore be crucial to the ecology of teosinte.
+The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
+Lukens and Doebley (1999) introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller \citep{LukensDoebley1999}.
+Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density.
 In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.

### 2014-08-27 ResponseToReview.tex
Edits to response to reviewer 2
Additional details concerning tb1 in the shade avoidance pathway in
teosinte, discussion of importance to ecology of teosinte
Edits to response to reviewer 2
changes to sampling details — added average per population/accession

@@
 
-\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness} \lev{the refs are in the sentences discussing the SAS in teo/grasses}
-
 9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic effects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
@@
 
-\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)." \jri{I'd also add a median or mean, or note how many populations had >8 individuals or somet other way for them to get a sense of the data as a range by itself is hard to think about}}
-
-\mbh{I agree with Jeff here}
+\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
 

### 2014-08-27 manuscript.tex
Edits to response to reviewer 2
changes to sampling details — added average per population/accession

@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.

### 2014-08-27 ResponseToReview.tex
Response to reviewer 2
Edits to BayEnv data description and analyses

@@
 
-\res{We have clarified this in the text on page \mbh{X}, lines \mbh{XX}.
+\res{We have clarified this in the text on page 6, lines 99-103.
 We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
@@
 If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
-We have included the following sentence in the text: \mbh{update the following in the text} ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
+We have included the following sentence in the text: ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
 
@@
 
-\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page \mbh{X}, line \mbh{XX}: \mbh{Update text in manuscript}``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 7, line 143-144: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
 
@@
 Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?\\
-
-%\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
-A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.
-We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh\"aj\"arvi  \emph{et al.} (2013)."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page \mbh{X}, lines \mbh{XX} of the text.}
+We have further clarified this on page 8, lines154-157 of the text.}
 

### 2014-08-27 manuscript.tex
Response to reviewer 2
Edits to BayEnv data description and analyses

@@
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
-If only one of the two PCR reactions resolved well, we scored one allele for the individual.
+Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
 We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
@@
 \textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations in which eight or more individuals were sampled, totaling 32 populations.
+These analyses only included populations ($n = 32$) in which eight or more individuals were sampled.
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
@@
 SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium¨ HD Assay (Illumina, San Diego, CA, USA).
-Environmental data from a single year were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and summarized by principle component analysis following \citet{Pyhajarvi2013}.
+Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}.
 

### 2014-08-27 ResponseToReview.tex
Edits to response to reviewer 2
Edits to section regarding LD and STRUCTURE analyses

@@
 \jri{read Matt's paper, explain to reviewer}
+\lev{I cannot find it explicitly stated in the paper.
+My assumption is that since STRUCTURE doesn't account for physical linkage you grouped things into chunks of 5kb chunks to try to elimiate the effects of physical linkage since LD decays in maize in ~1kb if I remember correctly.
+Is this correct?
+I've written it in this way, so let me know if this is wrong}
 

### 2014-08-27 ResponseToReview.tex
Edits to response to reviewer 2, text, refs
Edits concerning STRUCTURE analyses; including response to review,
methods, and references

--- 2014-08-27 ResponseToReview.tex
@@
 
+\res{Previous work has shown that LD decays in maize in 2000bp, on average.
+We wanted to avoid picking up signals of linkage due to physical linkage in our analyses  (STRUCTURE does not account for linkage due to physical distance), thus we grouped SNPs that were separated by less than a certain distance (5kb) into haplotype blocks in hopes that we would be grouping SNPs that would otherwise give signals of linkage due to physical linkage.
+We have clarified this point in the text, page 9, lines 182-186: ``Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from Hufford \emph{et al.} (2013) that grouped SNPs separated by less than 5kb into haplotypes (Hufford \emph{et al.} 2011).
+LD decays on average in 2000bp in maize (Remington \emph{et al.} 2001), thus we feel comfortable using haplotype blocks of 5kb in order to circumvent picking up evidence of physical linkage in our STRUCTURE analyses."}
+
 \res{hmm. Matt why did you use 5kb when you did this for your study?}
 \jri{read Matt's paper, explain to reviewer}
+\lev{I cannot find it explicitly stated in the paper.
+My assumption is that since STRUCTURE doesn't account for physical linkage you grouped things into chunks of 5kb chunks to try to elimiate the effects of physical linkage since LD decays in maize in ~2kb if I remember correctly.
+Is this correct?
+I've written it in this way, so let me know if this is wrong}
 

### 2014-08-27 manuscript.tex
Edits to response to reviewer 2, text, refs
Edits concerning STRUCTURE analyses; including response to review,
methods, and references

@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes.
+Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from Hufford \emph{et al.} (2013) that grouped SNPs separated by less than 5kb into haplotypes \citep{Hufford2013}.
+LD decays on average in 2000bp in maize \citep{Remington2001}, thus we feel comfortable using haplotype blocks of 5kb in order to circumvent picking up evidence of physical linkage in our STRUCTURE analyses.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.

### 2014-08-27 ResponseToReview.tex
Edits to response to reviewer 2
Edits concerning materials and methods, greenhouse, sampling and
analyses

@@
 
-\res{We have edited the text to clarify: ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
+\res{We have edited the text to clarify, page 9, lines 199-200: ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
 
@@
 
-\res{Thank you for the suggestion, we have edited it in the text and added it to the end of the Materials and Methods section: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+\res{Thank you for the suggestion, we have edited it in the text and added it to the end of the Materials and Methods section page 10, lines 215-219: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}
@@
 
-\res{We have clarified this in the manuscript.}
+\res{We have clarified this in the manuscript, page 9 lines 205-206: ``We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above."}
 
@@
 
-\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
+\res{We have added ratios to the text, page 10, lines 224-227:``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
 
@@
 As edited into the Materials and Methods section, there is a PCR that produces a specific size for each allele.
-We have edited subsequent text accordingly.}
+We have edited subsequent text accordingly, page 6, lines 94-104:  ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
+Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
+HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
+Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
 

### 2014-08-27 ResponseToReview.tex
edits to response

--- 2014-08-27 ResponseToReview.tex
@@
 
+\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness} \lev{the refs are in the sentences discussing the SAS in teo/grasses}
+
 9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic effects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed.
@@
 
-\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)." \jri{I'd also add a median or mean, or note how many populations had >8 individuals or somet other way for them to get a sense of the data as a range by itself is hard to think about}}
+
+\mbh{I agree with Jeff here}
 
@@
 
-\res{We have clarified this in the text on page 6, lines 99-103.
+\res{We have clarified this in the text on page \mbh{X}, lines \mbh{XX}.
 We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
@@
 If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
-We have included the following sentence in the text: ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
+We have included the following sentence in the text: \mbh{update the following in the text} ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
 
@@
 
-\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 7, line 143-144: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page \mbh{X}, line \mbh{XX}: \mbh{Update text in manuscript}``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
 
@@
 Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?\\
+
+%\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data.
+A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years.
+We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh\"aj\"arvi  \emph{et al.} (2013)."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page 8, lines154-157 of the text.}
+We have further clarified this on page \mbh{X}, lines \mbh{XX} of the text.}
 
@@
 
-\res{hmm. Matt why did you use 5kb when you did this for your study?}
-\jri{read Matt's paper, explain to reviewer}
-\lev{I cannot find it explicitly stated in the paper.
-My assumption is that since STRUCTURE doesn't account for physical linkage you grouped things into chunks of 5kb chunks to try to elimiate the effects of physical linkage since LD decays in maize in ~1kb if I remember correctly.
-Is this correct?
-I've written it in this way, so let me know if this is wrong}
+\res{While the recombination landscape will vary to some extent across individuals of teosinte and maize, broad-scale patterns are conserved.
+Previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{remington2001}}.
 

### 2014-08-27 ResponseToReview.tex
edits to response

@@
 
-\res{While the recombination landscape will vary to some extent across individuals of teosinte and maize, broad-scale patterns are conserved.
-Previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{remington2001}}.
+\res{While the recombination landscape will vary to some extent across individuals of teosinte and maize, broad-scale patterns are conserved \citep{bauer2013}.
+We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize that suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{remington2001}.
+Using this measure, we chose 5kb as a conservative estimate of the extent of linkage disequilibrium in our sample and the physical distance over which SNPs would have to be separated in order to be in weak linkage and not violate assumptions of the STRUCTURE model.}
 
-17) Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?
+17) Page 9 lines 2 - 3: ``Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?
 Or you turned on the drip by hand?
-In which case, it was not automatic… \\
+In which case, it was not automatic\\
 
@@
 
-18) Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1.
-F" So, why did you do it?
+18) Page 9 lines 7 - 8: ``Culm diameter is not believed to be correlated with tillering index or variation at tb1." So, why did you do it?
 Because of the last sentence in the materials and methods?
@@
 
-\res{Thank you for the suggestion, we have edited it in the text and added it to the end of the Materials and Methods section: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}
@@
 
-\res{We have clarified this in the manuscript.}
+\res{We have clarified in the manuscript that genotypes were obtained using our PCR protocol.}
 
@@
 
-\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (37 out of 86) and four (4 out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
+\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) \mbh{update manuscript to $n=X$ format} populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
 
@@
 
-\res{We believe we addressed this with editing for your previous comment concerning uncertainty of the PCRs we were using.
-As edited into the Materials and Methods section, there is a PCR that produces a specific size for each allele.
-We have edited subsequent text accordingly.}
+\res{Please see our response to point 12.
+We developed a co-dominant assay that was capable of detecting whether an individual was homozygous or heterozygous for the \emph{Hopscotch} insertion or lacked the insertion entirely.
+We have also edited this to read "eight or more chromosomes" which is a more precise description of our methodology.}
+\mbh{update the text in the manuscript}
 
-22) Page 13 lines 25 - 27: ``and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
+22) Page 13 lines 25 - 27: ``and genetic data support ongoing gene flow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Hufford et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011).
 Add Warburton, et al., 2011.

### 2014-08-28 ResponseToReview.tex
edits to response to review
removed notes to update text that had been dealt with previously, added
additional page/line #s

@@
 Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
-We have, however, added text to the conclusion (Page 15, 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
+We have, however, added text to the conclusion (Page 15, lines 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
 ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008).
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example, page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008).
 Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010).
@@
 \res{We have moved the references cited here to more clearly reflect the papers' findings.
-We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte}
-
-\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness} \lev{the refs are in the sentences discussing the SAS in teo/grasses}
+We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte.
+Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} (Lukens and Doebley 1999) and other grasses (Kebrom and Brutnell 2007) via changes in expression levels in response to shading.
+Lukens and Doebley (1999) introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller (Lukens and Doebley 1999).
+Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density."}
 
@@
 
-\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)." \jri{I'd also add a median or mean, or note how many populations had >8 individuals or somet other way for them to get a sense of the data as a range by itself is hard to think about}}
-
-\mbh{I agree with Jeff here}
+\res{We have added a range of values (1-38 individuals) per population in the methods, pages 5-6, lines 85-90: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
 
@@
 
-\res{We have clarified this in the text on page \mbh{X}, lines \mbh{XX}.
-We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
+\res{We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
 no \emph{Hopscotch}).
 If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
-We have included the following sentence in the text: \mbh{update the following in the text} ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
+We have clarified this in the text on page 6, lines 92-104.
+: ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
+Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
+HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
+Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
 

### 2014-08-28 ResponseToReview.tex
Edits to response to reviewer 2
updated line numbers and removed notes

@@
 
-\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page \mbh{X}, line \mbh{XX}: \mbh{Update text in manuscript}``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 7, lines 143-144: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page \mbh{X}, lines \mbh{XX} of the text.}
+We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following (Pyh{\"a}j{\"a}rvi \emph{et al.} 2013)."}
 
@@
 
-\res{We have clarified the dataset in the paper: ``We used teosinte (ssp.
+\res{We have clarified the dataset in the paper on page 8, lines 151-154: ``We used teosinte (ssp.
 \emph{parviglumis} and ssp.
@@
 \res{While the recombination landscape will vary to some extent across individuals of teosinte and maize, broad-scale patterns are conserved \citep{bauer2013}.
-We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize that suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{remington2001}.
+We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize that suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{Remington2001}.
 Using this measure, we chose 5kb as a conservative estimate of the extent of linkage disequilibrium in our sample and the physical distance over which SNPs would have to be separated in order to be in weak linkage and not violate assumptions of the STRUCTURE model.}
@@
 
-\res{We have edited the text to clarify: ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
+\res{We have edited the text to clarify, page 9, lines 199-200 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
 

### 2014-08-28 ResponseToReview.tex
edits to response to reviewer 2
updated line numbers, issues with sampling methodology for analyses

@@
 
-\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) \mbh{update manuscript to $n=X$ format} populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
+\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
 
@@
 We have also edited this to read "eight or more chromosomes" which is a more precise description of our methodology.}
-\mbh{update the text in the manuscript}
 

### 2014-08-28 ResponseToReview.tex
Edited response to reviewer 2
updated line numbers

@@
 
-\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page \mbh{X}, lines \mbh{XX}) following primer sequences: 
+\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page 6, lines 99-104) following primer sequences: 
 ``Homozygotes show a single band for absence of the \emph{Hopscotch} element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."
@@
 We developed a co-dominant assay that was capable of detecting whether an individual was homozygous or heterozygous for the \emph{Hopscotch} insertion or lacked the insertion entirely.
-We have also edited this to read "eight or more chromosomes" which is a more precise description of our methodology.}
+We have also edited this to read ``eight or more chromosomes", which is a more precise description of our methodology.}
 

### 2014-08-28 ResponseToReview.tex
edits to response

--- 2014-08-28 ResponseToReview.tex
@@
 Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
-We have, however, added text to the conclusion (Page 15, lines 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
+We have, however, added text to the conclusion (Page 15, 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
 ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example, page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008).
+\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008).
 Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010).
@@
 \res{We have moved the references cited here to more clearly reflect the papers' findings.
-We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte.
-Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} (Lukens and Doebley 1999) and other grasses (Kebrom and Brutnell 2007) via changes in expression levels in response to shading.
-Lukens and Doebley (1999) introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller (Lukens and Doebley 1999).
-Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density."}
+We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte}
+
+\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness} \lev{the refs are in the sentences discussing the SAS in teo/grasses}
 
@@
 
-\res{We have added a range of values (1-38 individuals) per population in the methods, pages 5-6, lines 85-90: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)." \jri{I'd also add a median or mean, or note how many populations had >8 individuals or somet other way for them to get a sense of the data as a range by itself is hard to think about}}
+
+\mbh{I agree with Jeff here}
 
@@
 
-\res{We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
+\res{We have clarified this in the text on page \mbh{X}, lines \mbh{XX}.
+We developed primers for two PCR reactions that would amplify products with size polymorphism characteristic of the two different alleles (\emph{i.e.}, \emph{Hopscotch} vs.
 no \emph{Hopscotch}).
 If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
-We have clarified this in the text on page 6, lines 92-104.
-: ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
-Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
-Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
+We have included the following sentence in the text: \mbh{update the following in the text} ``Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype."}
 
@@
 
-\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 7, lines 143-144: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page \mbh{X}, line \mbh{XX}: \mbh{Update text in manuscript}``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following (Pyh{\"a}j{\"a}rvi \emph{et al.} 2013)."}
+We have further clarified this on page \mbh{X}, lines \mbh{XX} of the text.}
 
@@
 
-\res{We have clarified the dataset in the paper on page 8, lines 151-154: ``We used teosinte (ssp.
+\res{We have clarified the dataset in the paper: ``We used teosinte (ssp.
 \emph{parviglumis} and ssp.
@@
 \res{While the recombination landscape will vary to some extent across individuals of teosinte and maize, broad-scale patterns are conserved \citep{bauer2013}.
-We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize that suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{Remington2001}.
+We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize that suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{remington2001}.
 Using this measure, we chose 5kb as a conservative estimate of the extent of linkage disequilibrium in our sample and the physical distance over which SNPs would have to be separated in order to be in weak linkage and not violate assumptions of the STRUCTURE model.}
@@
 
-\res{We have edited the text to clarify, page 9, lines 199-200 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
+\res{We have edited the text to clarify: ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
 
@@
 We developed a co-dominant assay that was capable of detecting whether an individual was homozygous or heterozygous for the \emph{Hopscotch} insertion or lacked the insertion entirely.
-We have also edited this to read "eight or more chromosomes" which is a more precise description of our methodology.}
+We have also edited this to read ``eight or more chromosomes" which is a more precise description of our methodology.}
 \mbh{update the text in the manuscript}

### 2014-08-28 manuscript.tex

@@
 \textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations ($n = 32$) in which eight or more individuals were sampled.
+These analyses only included populations ($n = 32$) in which eight or more ichromosomes were sampled.
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 (out of 86) and four (out of 17) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.

### 2014-08-28 ResponseToReview.tex
edits to response

--- 2014-08-28 ResponseToReview.tex
@@
 We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript.
-In the attached revision, we have addressed the majority of issues raised, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
-In addition to our revision, we include a detailed response to reviewers below.
-We hope that you are willing to consider our revised manuscript as a resubmission.\\\
+In the attached revision, we have addressed the majority of issues raised, but, while we feel that the new experiments suggested by Reviewer 1 would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
+In our submitted manuscript, we report results of a genotyping assay surveying several hundred wild and domesticated maize individuals for presence of the domesticated maize allele at \emph{tb1}, Sanger sequencing of a targeted subset of these samples to assess evidence of selection on the domesticated allele in wild populations, a phenotyping experiment to gauge the effects of the domesticated allele in a wild population, and reanalysis of existing genome-wide data to place the signature of selection at this locus in a wider context.
+This manuscript reports a substantial amount of novel work and explores, in a meaningful way, the concept of domestication from standing variation and the potential ecological relevance of domesticated alleles in wild populations.
+In addition to our revised manuscript, we include a detailed response to reviewers below.\\\
 

### 2014-08-28 ResponseToReview.tex
edits to response

@@
 However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
-Moreover, our sampling of individuals from this population for our grow out was not random.  
-We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
+Moreover, our sampling of individuals from this population for our grow-out was not random.  
+We initially genotyped a single individual from each sampling site within the population and then targeted our sample to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
 Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele.
@@
 
-\res{Note that all the sampling for both phenotyping experiments was limited to a single large population from which we had a large number of seed collected.
-Seed was collected from various "sampling sites" within the contiguous population at which we had estimated Hopscotch allele frequency.
-For our more broad genotyping of both landrace maize, ssp.
+\res{Note that sampling for both phenotyping experiments was limited to a single population from which we had a large number of seed collected.
+Seed was collected from various ``sampling sites" within the contiguous population at which we had estimated \emph{Hopscotch} allele frequency.
+For our broader genotyping of landrace maize, ssp.
 \emph{parviglumis}, and ssp.
-\emph{mexicana} we sampled as many populations as we could.
-For most of these populations the prevalence of the Hopscotch was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout these locations.}
+\emph{mexicana} we included samples from all populations to which we had access.
+For most of these populations the prevalence of the \emph{Hopscotch} was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout the natural ranges of these taxa.}
 

### 2014-08-28 ResponseToReview.tex
Edits to response to review
citations reformatted

--- 2014-08-28 ResponseToReview.tex
@@
 Presumably, if the \emph{Hopscotch} element were under strong selection in teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
-In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} (Studer \emph{et al.} 2011).
+In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch} \citep{Studer2011}.
 We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
@@
 \res{We agree that a planned grow-out with previously determined genotypes could be useful.  
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds \citep{Studer2011} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
 Moreover, our sampling of individuals from this population for our grow out was not random.  
@@
 
-\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals (Studer \emph{et al.} 2011); however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
+\res{Reviewer 2 is correct in stating that it was known that the Hopscotch was present in some teosinte individuals \citep{Studer2011}; however, previous studies sampled a small fraction of the populations and individuals per population that we present here and thus did not provide comprehensive estimates.
 Our study is not only concerned with the presence or absence of the \emph{Hopscotch} allele in teosinte but also the frequency at which it segregates in a range-wide sample of teosinte.  
 This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion (\emph{i.e.}, the domestication allele at \emph{tb1}) in teosinte. 
-We have added clarifying text to the introduction (Page 5, lines 55-61): ``The effects of the \emph{Hopscotch} insertion have been studied in maize (Studer \emph{et al.} 2011), and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes (Studer and Doebley 2012), but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+We have added clarifying text to the introduction (Page 5, lines 55-61): ``The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.
@@
 This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
+
+\lev{need to add something here. I had added something I thought in the text...}
 
@@
 Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010).
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.} Bai and Lindhout 2007), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors (Whitton \emph{et al.} 1997).}
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}\citep{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}
 
@@
 Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} (Lukens and Doebley 1999) and other grasses (Kebrom and Brutnell 2007) via changes in expression levels in response to shading.
-Lukens and Doebley (1999) introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller (Lukens and Doebley 1999).
+\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller \citet{LukensDoebley1999}.
 Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density."}
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following (Pyh{\"a}j{\"a}rvi \emph{et al.} 2013)."}
+We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}."}
 
@@
 \emph{parviglumis} and ssp.
-\emph{mexicana} genotyping and covariance data from Pyh{\"a}j{\"a}rvi \emph{et al.} (2013) for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
-SNP data from Pyh{\"a}j{\"a}rvi \emph{et al.} (2013) were obtained using the MaizeSNP50 BeadChip and Infinium® HD Assay (Illumina, San Diego, CA, USA)."}
+\emph{mexicana} genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium® HD Assay (Illumina, San Diego, CA, USA)."}
 

### 2014-08-28 ResponseToReview.tex

@@
 We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+Moreover, our sampling of individuals from this population for our grow-out was not random.  
+We initially genotyped a single individual from each sampling site within the population and then targeted our sample to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
 Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele.

### 2014-08-28 ResponseToReview.tex
edits to response

--- 2014-08-28 ResponseToReview.tex
@@
 \res{We did not measure ear architecture or internode length. 
-While \emph{tb1} does impact these traits, our primary interest was in the effect of \emph{tb1} on tillering as a potentially ecologically relevant phenotype given its known substantial effects on this trait in a maize background.
+While \emph{tb1} does impact these traits, our primary interest was in the effect of the \emph{Hopscotch} insertion at \emph{tb1} on tillering as a potentially ecologically relevant phenotype given its known substantial effects on this trait in a maize background.
 Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.

### 2014-08-28 ResponseToReview.tex
edits to response

@@
 
+\res{Comments from the Associate Editor and Reviewers are in black text and our responses are in blue text}
+
 I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.
@@
 However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds \citep{Studer2011} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
-Moreover, our sampling of individuals from this population for our grow out was not random.  
+Moreover, our sampling of individuals from this population for our grow-out was not random.  
 We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
@@
 \emph{mexicana} we included samples from all populations to which we had access.
-For most of these populations the prevalence of the \emph{Hopscotch} was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout the natural ranges of these taxa.}
+For most of these populations the prevalence of the \emph{Hopscotch} was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the \emph{Hopscotch} throughout the natural ranges of these taxa.}
 
@@
 ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
+\emph{gt1}, \emph{tru1}, and \emph{te1}).
 These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}
@@
 Introduction:
-Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".
+Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".
 Can you tell us why an understanding is useful?
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example, page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008).
+\mbh{there needs to be a clearer explanation of how wild diversity can be important for crop improvement}
+
+\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice, \emph{Oryza sativa} (Kovach and McCouch 2008).
 Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010).
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}\citep{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}, \citealt{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}
 
@@
 We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte.
-Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} (Lukens and Doebley 1999) and other grasses (Kebrom and Brutnell 2007) via changes in expression levels in response to shading.
-\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller \citet{LukensDoebley1999}.
-Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density."}
+Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} (Lukens and Doebley 1999) and other grasses (Kebrom and Brutnell 2007), exhibiting changes in expression levels in response to shading.
+\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, whereas under high density, plants showed significantly reduced tillers and grew taller \citep{LukensDoebley1999}.
+Based on these results we hypothesized that \emph{tb1} may play a role in the ecology of teosinte."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}."}
+We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
+These data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}."}
 
@@
 \emph{parviglumis} and ssp.
-\emph{mexicana} genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
-SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium® HD Assay (Illumina, San Diego, CA, USA)."}
+\emph{mexicana}) SNP genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA)."}
 

### 2014-08-28 ResponseToReview.tex
edits to response

@@
 In the attached revision, we have addressed the majority of issues raised, but, while we feel that the new experiments suggested by Reviewer 1 would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
-In our submitted manuscript, we report results of a genotyping assay surveying several hundred wild and domesticated maize individuals for presence of the domesticated maize allele at \emph{tb1}, Sanger sequencing of a targeted subset of these samples to assess evidence of selection on the domesticated allele in wild populations, a phenotyping experiment to gauge the effects of the domesticated allele in a wild population, and reanalysis of existing genome-wide data to place the signature of selection at this locus in a wider context.
+In our submitted manuscript, we report results of a genotyping assay surveying several hundred wild and domesticated maize individuals for presence of the domesticated maize allele at \emph{tb1}, Sanger sequencing of a targeted subset of these samples to assess evidence of selection on the domesticated allele in wild populations, a phenotyping experiment to gauge the effects of the domesticated allele in a wild population, and reanalysis of existing genome-wide SNP data to place the signature of selection at this locus in a wider context.
 This manuscript reports a substantial amount of novel work and explores, in a meaningful way, the concept of domestication from standing variation and the potential ecological relevance of domesticated alleles in wild populations.
@@
 \res{We agree that a planned grow-out with previously determined genotypes could be useful.  
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds \citep{Studer2011} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
-Moreover, our sampling of individuals from this population for our grow-out was not random.  
-We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, (Studer \emph{et al.} 2011) and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \citep{Studer2011} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
 Moreover, our sampling of individuals from this population for our grow-out was not random.  

### 2014-08-28 ResponseToReview.tex
edits to Response

@@
 This analysis indicated we would need 71 individuals in both classes to detect a significant difference in tillering index.  
-We attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) where we had already shown \emph{Hopscotch} to be segregating at high frequency in order to detect a significant difference between homozygous genotypes.
+For the Phenotyping 2 experiment, we attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) where we had already shown \emph{Hopscotch} to be segregating at high frequency in order to detect a significant difference between homozygous genotypes.
 Our Phenotyping 2 experiment detects the same trend reported in Phenotyping 1, but the trend is still not significant.

### 2014-08-29 ResponseToReview.tex
edits to response to review
fixed citations

@@
 Introduction:
-Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".
+Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovackMcCouch2008}.".
 Can you tell us why an understanding is useful?
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice, \emph{Oryza sativa} (Kovach and McCouch 2008).
-Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010).
+\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice, \emph{Oryza sativa} \citep{KovachMcCouch2008}.
+Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004; SigmonVollbrecht2010}.
 While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}, \citealt{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}
@@
 We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte.
-Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} (Lukens and Doebley 1999) and other grasses (Kebrom and Brutnell 2007), exhibiting changes in expression levels in response to shading.
+Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007}, exhibiting changes in expression levels in response to shading.
 \citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, whereas under high density, plants showed significantly reduced tillers and grew taller \citep{LukensDoebley1999}.

### 2014-08-29 ResponseToReview.tex
edits to response

@@
 \res{We agree that a planned grow-out with previously determined genotypes could be useful.  
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \citep{Studer2011} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds \citep{Studer2011}, and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
 Moreover, our sampling of individuals from this population for our grow-out was not random.  
@@
 This analysis indicated we would need 71 individuals in both classes to detect a significant difference in tillering index.  
-For the Phenotyping 2 experiment, we attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) where we had already shown \emph{Hopscotch} to be segregating at high frequency in order to detect a significant difference between homozygous genotypes.
+For the Phenotyping 2 experiment, we attempted to sample enough individuals from the San Lorenzo population (the same population sampled for Phenotyping 1) in order to detect a significant difference between homozygous genotypes.
 Our Phenotyping 2 experiment detects the same trend reported in Phenotyping 1, but the trend is still not significant.

### 2014-08-29 ResponseToReview.tex
edits to response

@@
 \res{We did not measure ear architecture or internode length. 
-While \emph{tb1} does impact these traits, our primary interest was in the effect of the \emph{Hopscotch} insertion at \emph{tb1} on tillering as a potentially ecologically relevant phenotype given its known substantial effects on this trait in a maize background.
+While \emph{tb1} does impact these traits, our primary interest was in the effect of the \emph{Hopscotch} insertion at \emph{tb1} on tillering given that this is an ecologically relevant phenotype that is known to be substantially affected by \emph{Hopscotch} in a maize background.
 Both additional greenhouse and expression analyses would indeed be useful avenues of further investigation.
@@
 \emph{gt1}, \emph{tru1}, and \emph{te1}).
-These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}
+These data, in conjunction with more exhaustive phenotyping, should help to further clarify the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}
 

### 2014-08-29 ResponseToReview.tex
edits to response

@@
 Introduction:
-Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovackMcCouch2008}.".
+Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.".
 Can you tell us why an understanding is useful?
@@
 \res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice, \emph{Oryza sativa} \citep{KovachMcCouch2008}.
-Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004; SigmonVollbrecht2010}.
+Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrecht2010}.
 While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}, \citealt{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}

### 2014-08-29 ResponseToReview.tex
Edits to response to review
Matt, a note to you about the post-hoc power analyses. Not sure if we
want to leave it in the manuscript or not. If we take it out we can
just say that we removed phenotyping 2 so their initial comment no
longer needs a long response.
Edits to response to review
edits to response to review, spacing

@@
 
+\lev{Matt, should we remove this? I left in the bit about the post-hoc power analyses because I didn't know if we wanted to leave it in even though we removed phenotyping 2}
+
 What was the purpose of the Phenotyping 2  experiment?
@@
 
-One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.
+One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.\\
 

### 2014-08-29 ResponseToReview.tex
jri edits to response

@@
 
-\noindent Jeffrey Ross-Ibarra and Matthew Hufford
+\noindent Matthew Hufford and Jeffrey Ross-Ibarra
 
@@
 \res{We agree that a planned grow-out with previously determined genotypes could be useful.  
-However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds \citep{Studer2011}, and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
+However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds \citep{Studer2011}, and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring in nature.
 Moreover, our sampling of individuals from this population for our grow-out was not random.  
@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-We have clarified our sampling methodology on pages 8-9, lines 174-185. Finally, plants were situated randomly in the greenhouse prior to genotyping.
-Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}
+We have clarified our sampling methodology on pages 8-9, lines 174-185.
+Finally, plants were situated randomly in the greenhouse prior to genotyping, which effectively represents a randomized experimental design.}
 
@@
 Your Phenotyping 1 results show no effect of genotype on tillering.
-Thus, it is not clear what effect you are referring to.\\
-
-\lev{Matt, should we remove this? I left in the bit about the post-hoc power analyses because I didn't know if we wanted to leave it in even though we removed phenotyping 2}
-
+Thus, it is not clear what effect you are referring to.
 What was the purpose of the Phenotyping 2  experiment?
@@
 
-%JRI STOPPED HERE
-
 11) Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.\\
@@
 
+\jri{please make sure the perl script is on github or online elsewhere and we cite it as such.}
+
 17) Page 9 lines 2 - 3: ``Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?

### 2014-08-31 ResponseToReview.tex
edits to response
Edits to response to review
edits
edits to response to review
edits to comment 7 regarding domestication

@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice, \emph{Oryza sativa} \citep{KovachMcCouch2008}.
-Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrecht2010}.
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}, \citealt{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}
+\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 31-43: ``Through the process of domestication and improvement, genetic diversity in many domesticates was reduced relative to their wild progenitors \citep{KovachMcCouch2008}.
+Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}."}
 
@@
 
-\res{Reviewer 2 is correct in noting that ssp.
+\res{Our goal in screening for the \emph{Hopscotch} insertion, was to screen as widely as possible and assess both its taxonomic and geographic distribution.
+TheReviewer 2 is correct in noting that ssp.
 \emph{mexicana} is not as prevalent in our paper as ssp.
@@
 \jri{please make sure the perl script is on github or online elsewhere and we cite it as such.}
+\lev{it's script from matt's introgression paper}
 

### 2014-08-31 ResponseToReview.tex
Edits to response to review
edits to comment 7
edits to response

--- 2014-08-29 ResponseToReview.tex
@@
 
-\res{Reviewer 2 is correct in noting that ssp.
+\res{Our goal in screening for the \emph{Hopscotch} insertion was to assess its taxonomic and geographic distribution to the extent possible given the samples to which we had access.
+Reviewer 2 is correct in noting that ssp.
 \emph{mexicana} is not as prevalent in our paper as ssp.
 \emph{parviglumis}.
-This fact is due to limitations in access to accessions of ssp.
-\emph{mexicana} as field collections for ssp.
-\emph{mexicana} are not nearly as extensive as those of ssp.
-\emph{parviglumis}.
-Additionally ssp.
-\emph{mexicana} is more difficult to cultivate in a greenhouse setting.
-We felt that it would be important to include ssp.
-\emph{mexicana} in our study, especially since it is known that gene flow occurs between ssp.
-\emph{mexicana} and landrace maize.
-We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.} \jri{just say we don't have as much seed from mexicana, which is true.
-it's just as easy to grow in the greenhouse.
-also I'd think of adding a sentence to the discussion mentioning mexicana.
-if it is in mexicana too, doesn't that suggest it's pretty old, given the divergence time between mexicana and parviglumis?
-or do we think it's in mexicana because of gene flow with parv or maize?
-this could be worth speculating on.}
-\lev{Hmm.
-Well didn't Matt's paper on introgression support a significant lack of gene flow from maize to mex around domestication loci?
-So how likely would it be that it is in mex because of gene flow?
-So would it make more sense to argue that it's there because it is old and hasn't been lost?
-Would it make sense too because it was estimated to predate the domestication of maize?
-Maybe it is in parv at a higher frequency because it's 'doing' something, even though we didn't pick up an effect on tillering, and maybe it's just drifting in mex, which could explain why it is at a much lower frequency in mex?}
-
-\mbh{I'll take a stab at this...pretty sure it was at low frequency in all \emph{mexicana} for which we had population sampling.
-It would not have made sense to use this for phenotyping because we would have had to sample a ton of individuals in order to compare \emph{Hopscotch} phenotype to no-\emph{Hopscotch} phenotype.}
+Though our sample of \emph{mexicana} is range-wide, we only had access to seeds from 17 populations.
+Frequency of the \emph{Hopscotch} insertion was low in these populations relative to \emph{parviglumis}, an interesting point that we have discussed further on page \mbh{X}, lines \mbh{XX} of the text.
+The low frequency of \emph{Hopscotch} in \emph{mexicana} precluded a phenotyping experiment with this subspecies since a very high number of seeds would have to be germinated in order to obtain enough individuals with the \emph{Hopscotch} allele for a well-powered experiment.}
 

### 2014-08-31 ResponseToReview.tex

@@
 
-\mbh{there needs to be a clearer explanation of how wild diversity can be important for crop improvement}
-
-\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 33-41: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice, \emph{Oryza sativa} \citep{KovachMcCouch2008}.
-Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrecht2010}.
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.}, \citealt{Bai2007}), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.}
+\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 31-43: ``Through the process of domestication and improvement, genetic diversity in many domesticates was reduced relative to their wild progenitors \citep{KovachMcCouch2008}.
+Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
+While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}."}
 

### 2014-08-31 manuscript.tex
Edits to manuscript
Introduction, discussion of utility of wild alleles with respect to
crop improvement

@@
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
-Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
 Plant domestication usually involves a suite of phenotypic changes, such as increase yield, loss of seed shattering, and increase fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}.
-Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
+Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying these agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
+Through the process of domestication and improvement, genetic diversity in many domesticates was reduced relative to their wild progenitors \citep{KovachMcCouch2008}.
+Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
 For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.

### 2014-09-01 ResponseToReview.tex
minor edits to response

@@
 Frequency of the \emph{Hopscotch} insertion was low in these populations relative to \emph{parviglumis}, an interesting point that we have discussed further on page \mbh{X}, lines \mbh{XX} of the text.
-The low frequency of \emph{Hopscotch} in \emph{mexicana} precluded a phenotyping experiment with this subspecies since a very high number of seeds would have to be germinated in order to obtain enough individuals with the \emph{Hopscotch} allele for a well-powered experiment.}
+The low frequency of \emph{Hopscotch} in \emph{mexicana} precluded a phenotyping experiment with this subspecies since a very high number of seeds would have to be germinated in order to obtain enough individuals with the \emph{Hopscotch} allele.}
 
@@
 
-\jri{please make sure the perl script is on github or online elsewhere and we cite it as such.}
-
 17) Page 9 lines 2 - 3: ``Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically?

### 2014-09-01 manuscript.tex
new first paragraph intro. this will require changing a response to review -- I don't think the whole agronomically useful angle worked since we are really asking an ecological question.

--- 2014-08-28 manuscript.tex
@@
 \emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte.
-Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.
-Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte.
-Finally, a greenhouse experiment with teosinte does not suggest \emph{tb1} controls tillering in teosinte.
+Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize.
+Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte, but a greenhouse experiment fails to find \emph{tb1} a strong association with tillering in plants from a single natural population.
 \vspace*{0.2cm}
@@
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
-Perhaps the central focus of the study of domestication has been the identification of genetic variation underlying agronomically important traits such as fruit size and plant architecture \citep{OlsenGross2010}.
-Plant domestication usually involves a suite of phenotypic changes, such as increase yield, loss of seed shattering, and increase fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}.
-Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
-Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
+Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
+Because most domesticates show reduced genetic diversity relative to their wild counterparts, a number of efforts have been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide} \jri{add some more}.
+Researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops} \jri{add more}. 
+But even though domesticated alleles are often found segregating in crop wild relatives \citep{}, we know almost nothing about the ecological role of this variation in natural populations.
+\jri{add some examples, can include teosinte ones mentioned below but also a few other crops}
+In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize. 
+%some effort has been invested into identifying agronomically useful loci in wild relatives \citep{}understanding how alleles from wild relatives behave in domesticated crops \citep[e.g.][]{Bai2007}, or identifying agronomically usefulvery little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
+%In addition to the loss of diversity due to selection on loci important for domestication, demography and the effects of linked selection have acted to decrease diversity genome-wide, leading to lower genetic diversity in many domesticates relative to their wild progenitors \citep{KovachMcCouch2008}.
+ \jri{@article{flint2009wide,
+  title={Wide variability in kernel composition, seed characteristics, and zein profiles among diverse maize inbreds, landraces, and teosinte},
+  author={Flint-Garcia, Sherry A and Bodnar, Anastasia L and Scott, M Paul},
+  journal={Theoretical and applied genetics},
+  volume={119},
+  number={6},
+  pages={1129--1142},
+  year={2009},
+  publisher={Springer-Verlag}
+}
+
+@article{ellstrand2010crops,
+  title={Crops gone wild: evolution of weeds and invasives from domesticated ancestors},
+  author={Ellstrand, Norman C and Heredia, Sylvia M and Leak-Garcia, Janet A and Heraty, Joanne M and Burger, Jutta C and Yao, Li and Nohzadeh-Malakshah, Sahar and Ridley, Caroline E},
+  journal={Evolutionary Applications},
+  volume={3},
+  number={5-6},
+  pages={494--504},
+  year={2010},
+  publisher={Wiley Online Library}
+}
+
+}
+%Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
+%For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+%\jri{why not include some specific teosinte examples like those cited in Hufford 2012 TIG?} Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
 
@@
 Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-Additionally, large naturally occurring populations of both \emph{Zea mays} ssp.
-\emph{parviglumis} (the wild progenitor of maize) and \emph{Zea mays} ssp.
-\emph{mexicana} (highland teosinte; hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
+Additionally, large naturally occurring populations of both \emph{parviglumis and the highland teosinte \emph{Zea mays} ssp.
+\emph{mexicana} (hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
 
@@
 A 4.9 kb retrotransposon (\emph{Hopscotch}) insertion into the upstream control region of \emph{tb1} in maize acts to enhance expression of \emph{tb1}, thus repressing lateral organ growth \citep{Doebley1997, Studer2011}.
-Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in ancient populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
+Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
 The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
-Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.
-\emph{parviglumis}, ssp.
-\emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
+Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
 
@@
 The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
-Lukens and Doebley (1999) introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller \citep{LukensDoebley1999}.
+\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller.
 Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density.
@@
 We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} combined with a larger SNP dataset in a subset of teosinte populations to explore patterns of genetic variation at this locus.
-Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.
+Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in samples from a natural population of \emph{parviglumis}.
 
@@
 \textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations ($n = 32$) in which eight or more ichromosomes were sampled.
+These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled.
 To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
@@
 SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium¨ HD Assay (Illumina, San Diego, CA, USA).
-Environmental data were were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}.
+Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}.
 
@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from Hufford \emph{et al.} (2013) that grouped SNPs separated by less than 5kb into haplotypes \citep{Hufford2013}.
-LD decays on average in 2000bp in maize \citep{Remington2001}, thus we feel comfortable using haplotype blocks of 5kb in order to circumvent picking up evidence of physical linkage in our STRUCTURE analyses.
+Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \url{XXX}]{Hufford2013}.
+In maize LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 
-Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data using a Box-Cox transformation ($\lambda=0$) Car Package for R, \citealt{FoxWeisberg2011}) to improve the normality and homogeneity of variance among genotype classes.
+Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data with a Box-Cox transformation ($\lambda=0$) in the Car Package for R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
 We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
@@
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses.
+SAS code used for analysis is available at \url{XXXX}.
 
@@
 However, we do observe a near significant ($p<0.08$) correlation between tillering index and genotype on day 40, though in the opposite direction of what is expected based on how the element acts phenotypically in maize.
+Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}.
 

### 2014-09-01 manuscript.tex
edits to main text

@@
 
+\textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
+These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled.
+To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
+BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
+We used teosinte (ssp.
+\emph{parviglumis} and ssp.
+\emph{mexicana} genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
+Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}.
+
 \textbf{Sequencing---}In addition to genotyping, we chose a subset of \emph{parviglumis} individuals for sequencing.
@@
 
-\textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled.
-To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
-BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
-We used teosinte (ssp.
-\emph{parviglumis} and ssp.
-\emph{mexicana} genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
-SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium¨ HD Assay (Illumina, San Diego, CA, USA).
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}.
-
 \textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}), creating neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
+To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html} to create neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
 For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.
@@
 
-\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set.
-In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion.
+\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} insertion is segregating at highest frequency (0.44) in our initial genotyping sample set.
+In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites within the population where genotyped individuals were homozygous or heterozygous for the insertion.
 We chose between 10-13 seeds from each of 23 sampling sites.
@@
 We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above.
-\lev{should we take this bit about power analysis out?} Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index.
 
@@
 
-\textbf{Genotyping---}Genotype of the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals. \jri{does this conflict with the 1100 mentioned in the summary?}
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
@@
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+Tajima's D is positive in all populations except La Mesa, \jri{positive D indicates too few rare variants.
+the discussion also says D is negative everywhere.
+is D positive or is the text wrong?
+please check that discussion and results are consistent and both match what is in the table.} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
@@
 The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
-However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}.
+However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) \citep{Pyhajarvi2013} and the maize 282 diversity panel \citep{Cook2012}.
 The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants which would arise as the result of recent admixture between populations.
@@
 \textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
-We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 206 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
-Our plantings produced 20 homozygotes for the teosinte (no Hopscotch) allele, 104 heterozygotes, and 82 homozygotes for the maize (Hopscotch) allele.
-We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig.
+We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
+We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{XXX} \jri(CITE) but is not thought to be affected by the \emph{Hopscotch} insertion. 
+Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
+Our plantings produced 82 homozygotes for the \emph{Hopscotch} insertion at \emph{tb1}, 104 heterozygotes, and 20 homozygotes lacking the insertion; these numbers do not deviate from expectations of Hardy-Weinberg equilibrium.
+After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig.
 \ref{Fig4Boxplots}), tiller number, or culm diameter.
-However, we do observe a near significant ($p<0.08$) correlation between tillering index and genotype on day 40, though in the opposite direction of what is expected based on how the element acts phenotypically in maize.
-Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}.
+Only on day 40 did we observe a weak ($r=XXX$) but statistically insignificant correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
+\jri{is this correct?}
 
@@
 While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
+
 Many of our populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, perhaps suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
@@
 
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results do not suggest introgression from maize into teosinte at the \emph{tb1} locus, and are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into teosinte around domestication loci.
+While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results do not suggest introgression from maize at the \emph{tb1} locus, and are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into \emph{mexicana} around domestication loci.
 Clustering in our NJ trees does not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
@@
 Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural populations of \emph{parviglumis}.
-Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative.
+Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative (Table XXX) \jri{please add table}.
 In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}.
@@
 
-Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
+\jri{matt, do we inclue mention of weber 2007 in the introduction as part of the motivation?
+might be good?} Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
 \citet{Weber2007} described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
@@
 Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
-Finally, although photoperiod for Phenotyping 2 reasonably approximated that of the normal \emph{parviglumis} growing season, greenhouse-specific environmental conditions (plant density, light regime, etc...) may have contributed to tillering responses different from those found in nature, obscuring the effect of the \emph{Hopscotch} insertion on variation.
+\jri{do we mention photoperiod as a possible explanation for phenotyping 1 results?}
+%Finally, although photoperiod for Phenotyping 2 reasonably approximated that of the normal \emph{parviglumis} growing season, greenhouse-specific environmental conditions (plant density, light regime, etc...) may have contributed to tillering responses different from those found in nature, obscuring the effect of the \emph{Hopscotch} insertion on variation.
 
-In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is more widespread in populations of \emph{parviglumis} and \emph{mexicana} than previously thought.
-Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression; however, it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
-We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}; this coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations suggests that the locus may play an ecological role in teosinte.
-In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in \emph{parviglumis}.
+In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is widespread in populations of \emph{parviglumis} and \emph{mexicana} and occasionally at high allele frequencies. 
+Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression, and it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
+We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}.
+Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus may play an ecological role in teosinte.
+In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in a diverse sample of \emph{parviglumis} from a natural population and the phenotypic consequences of variation at \emph{tb1} thus remain unclear.
 Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
 \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
-These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
+nce of the domesticated \emph{tb1} allele in natural populations of teosinte.
 

### 2014-09-02 manuscript.tex
edits to introduction

@@
 Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
-Phenotypic data were also collected using seed from a single teosinte population where \emph{Hopscotch} was found segregating.
+Phenotypic data were also collected using seed from a single teosinte population where \emph{Hopscotch} was found segregating at high frequency.
 
@@
 
-\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte.
-Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize.
-Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte, but a greenhouse experiment fails to find \emph{tb1} a strong association with tillering in plants from a single natural population.
+\emph{Key results:} Genotyping results indicate the \emph{Hopscotch} element is found in a number of teosinte populations and linkage disequilibrium near \emph{tb1} does not support recent introgression from maize.
+Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte, but a greenhouse experiment does not detect a strong association between \emph{tb1} and tillering in teosinte.
 \vspace*{0.2cm}
@@
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
-Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as 'domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
-Because most domesticates show reduced genetic diversity relative to their wild counterparts, a number of efforts have been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide} \jri{add some more}.
+Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
+Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide} \jri{add some more}.
+\mbh{this reference should be a gold mine: \url{http://bioscience.oxfordjournals.org/content/61/7/559.full}}
 Researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops} \jri{add more}.
-But even though domesticated alleles are often found segregating in crop wild relatives \citep{}, we know almost nothing about the ecological role of this variation in natural populations.
-\jri{add some examples, can include teosinte ones mentioned below but also a few other crops}
+\mbh{It looks like the Ellstrand paper is a review; if so and there are a number of good examples of this point, I would simply say ``(reviewed in \citealt{ellstrand2010crops})"}But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
+\jri{add some examples, can include teosinte ones mentioned below but also a few other crops} \mbh{I added some maize/teosinte examples...agree with Jeff that examples from other species would be good...here's a very recent example from a rabbit domestication paper: \url{http://www.sciencemag.org/content/345/6200/1074.full}}
 In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize. 
@@
 %In addition to the loss of diversity due to selection on loci important for domestication, demography and the effects of linked selection have acted to decrease diversity genome-wide, leading to lower genetic diversity in many domesticates relative to their wild progenitors \citep{KovachMcCouch2008}.
- \jri{@article{flint2009wide,
-  title={Wide variability in kernel composition, seed characteristics, and zein profiles among diverse maize inbreds, landraces, and teosinte},
-  author={Flint-Garcia, Sherry A and Bodnar, Anastasia L and Scott, M Paul},
-  journal={Theoretical and applied genetics},
-  volume={119},
-  number={6},
-  pages={1129--1142},
-  year={2009},
-  publisher={Springer-Verlag}
-}
 
-@article{ellstrand2010crops,
-  title={Crops gone wild: evolution of weeds and invasives from domesticated ancestors},
-  author={Ellstrand, Norman C and Heredia, Sylvia M and Leak-Garcia, Janet A and Heraty, Joanne M and Burger, Jutta C and Yao, Li and Nohzadeh-Malakshah, Sahar and Ridley, Caroline E},
-  journal={Evolutionary Applications},
-  volume={3},
-  number={5-6},
-  pages={494--504},
-  year={2010},
-  publisher={Wiley Online Library}
-}
-
-}
 %Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
@@
 Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-Additionally, large naturally occurring populations of both \emph{parviglumis and the highland teosinte \emph{Zea mays} ssp.
+Additionally, large naturally occurring populations of both \emph{parviglumis} and the highland teosinte \emph{Zea mays} ssp.
 \emph{mexicana} (hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.

### 2014-09-02 ResponseToReview.tex
edits through methods

@@
 \res{While the recombination landscape will vary to some extent across individuals of teosinte and maize, broad-scale patterns are conserved \citep{bauer2013}.
-We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from 102 lines of maize that suggest LD (as measured by $r^2$) will decline to 0.1 within 2,000 bp \citep{Remington2001}.
-Using this measure, we chose 5kb as a conservative estimate of the extent of linkage disequilibrium in our sample and the physical distance over which SNPs would have to be separated in order to be in weak linkage and not violate assumptions of the STRUCTURE model.}
+We therefore relied on previous estimates of the decay of linkage disequilibrium based on data from HapMap2 of maize that suggest LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb as an estimate of the extent of linkage disequilibrium in our sample and the physical distance over which SNPs would have to be separated in order to be in weak linkage and not violate assumptions of the STRUCTURE model.}
 
@@
 
-21) Page 10 lines 7 - 9: ``Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
+21) Page 10 lines 7 - 9: ``Using our Hopscotch genotyping, we calculated differentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles?
 I thought you only had presence or absence; how are alleles differentiated with your PCR?\\

### 2014-09-02 manuscript.tex
edits through methods

@@
 Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-Additionally, large naturally occurring populations of both \emph{parviglumis} and the highland teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
+Additionally, large, naturally-occurring populations of both \emph{parviglumis} and the highland teosinte \emph{Zea mays} ssp.
+\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
 
@@
 These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled.
-To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between the \emph{Hopscotch} frequency and environmental variables using BayEnv \citep{Coop2010}.
+To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between \emph{Hopscotch} frequency and environmental variables using the software BayEnv \citep{Coop2010}.
 BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
@@
 SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
-Environmental data were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database \citep{FAOHWSD} and \url{www.harvestchoice.org} and were either compiled over the last several decades (climatic data) or were from a single year (soil) and summarized by principle component analysis following \citet{Pyhajarvi2013}.
+Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
+These data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.
 
@@
 \textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html} to create neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
+To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
 For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.
@@
 Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \url{XXX}]{Hufford2013}.
-In maize LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
+In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
 We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals. \jri{does this conflict with the 1100 mentioned in the summary?}
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
+\jri{does this conflict with the 1100 mentioned in the summary?} \mbh{So for the rest of the samples, did we have only one PCR that worked or no amplification?
+We should clarify here the number for which we obtained \emph{Hopscotch} data and also use this number in the abstract.
+In the methods we can keep the total number we screened (1100).}
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
 Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
+\ref{Fig1Map}).
+Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).

### 2014-09-02 manuscript.tex
edits through results

@@
 is D positive or is the text wrong?
-please check that discussion and results are consistent and both match what is in the table.} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+please check that discussion and results are consistent and both match what is in the table.} \mbh{based on my read of the manuscript, the text is correct here (and it agrees with Table 2 and the Discussion)...if D is negative in La Mesa, that indicates an excess of rare variants in this population; also, the consistently negative D comment in the Discussion is referring to the 5' region of tb1 which shows negative D across populations so this is correct; check me on this, though, LV} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
@@
 The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants which would arise as the result of recent admixture between populations.
-If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster.
+If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the ``maize" cluster in our STRUCTURE runs, not the ``teosinte" cluster.
 In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig.
@@
 We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
-We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{XXX} \jri(CITE) but is not thought to be affected by the \emph{Hopscotch} insertion. 
+We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{Briggs2007} but is not thought to be affected by the \emph{Hopscotch} insertion. 
 Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
@@
 \ref{Fig4Boxplots}), tiller number, or culm diameter.
-Only on day 40 did we observe a weak ($r=XXX$) but statistically insignificant correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
-\jri{is this correct?}
+Only on day 40 did we observe a weak but statistically insignificant ($p=0.0848$) \mbh{would be good to add the $r^2$ value as well if we have it} correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
+\jri{is this correct?} \mbh{Looks correct to me; good idea to double-check, LV}
 

### 2014-09-02 manuscript.tex
edits to discussion

@@
 
-\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte \citep{Studer2011}. 
+\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte. 
 Given that the \emph{Hopscotch} insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
@@
 Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations.
-Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in natural populations of \emph{parviglumis}.
-Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative (Table XXX) \jri{please add table}.
+Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in populations of \emph{parviglumis}.
+Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative across populations (Table \ref{Table2Diversity}).
 In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}.
@@
 \jri{matt, do we inclue mention of weber 2007 in the introduction as part of the motivation?
-might be good?} Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
+might be good?} \mbh{Hmm...the significant results from Weber in \emph{parviglumis} were for traits we did not measure.
+If we mention this as motivation in the intro., would reviewers wonder why we measured different traits?} Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
 \citet{Weber2007} described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
@@
 Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
-\jri{do we mention photoperiod as a possible explanation for phenotyping 1 results?}
+\jri{do we mention photoperiod as a possible explanation for phenotyping 1 results?} \mbh{I don't think so...since phenotyping 1 and phenotyping 2 occurred at different times in the year and showed similar results}
 %Finally, although photoperiod for Phenotyping 2 reasonably approximated that of the normal \emph{parviglumis} growing season, greenhouse-specific environmental conditions (plant density, light regime, etc...) may have contributed to tillering responses different from those found in nature, obscuring the effect of the \emph{Hopscotch} insertion on variation.
@@
 Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
-nce of the domesticated \emph{tb1} allele in natural populations of teosinte.
+\emph{gt1}, \emph{tru1}, and \emph{te1}).
+These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
 

### 2014-09-04 ResponseToReview.tex
edits
Edits to response to review, references, introduction rewrite

@@
 \section*{Response to Reviewer 2} 
-Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
+5) Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.
 It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana.\\
@@
 
-This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
+6) This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\
 
@@
 
-Introduction:
+7) Introduction:
 Page 4, Lines 7 - 9: You say ``Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.".
@@
 
-\res{We have expanded on this concept in the introduction using \emph{Oryza} as an example on page 4, lines 31-43: ``Through the process of domestication and improvement, genetic diversity in many domesticates was reduced relative to their wild progenitors \citep{KovachMcCouch2008}.
-Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
+\res{We have expanded on this concept in the introduction on page \lev{x} lines \lev{x-x} ``Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
+Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
+Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
+Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
 For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
-Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
-While some effort has been invested into understanding how wild alleles behave in their domesticated relatives \citep[e.g.][]{Bai2007}, very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}."}
+In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
+A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acting on standing variation \citep{Rabbit}.
+There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}).
+But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
+In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize."}
 
@@
 
+\lev{am not seeing where this is in the text}
+
 10) Materials and Methods:
@@
 
+\lev{check numbers}
 \res{We have added a range of values (1-38 individuals) per population in the methods, pages 5-6, lines 85-90: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
@@
 
-\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section, \lev{page x, linesx} : ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}

### 2014-09-04 manuscript.tex
edits
Edits to response to review, references, introduction rewrite

--- 2014-09-02 manuscript.tex
@@
 Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
-Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide} \jri{add some more}.
-\mbh{this reference should be a gold mine: \url{http://bioscience.oxfordjournals.org/content/61/7/559.full}}
-Researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops} \jri{add more}.
-\mbh{It looks like the Ellstrand paper is a review; if so and there are a number of good examples of this point, I would simply say ``(reviewed in \citealt{ellstrand2010crops})"}But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
-\jri{add some examples, can include teosinte ones mentioned below but also a few other crops} \mbh{I added some maize/teosinte examples...agree with Jeff that examples from other species would be good...here's a very recent example from a rabbit domestication paper: \url{http://www.sciencemag.org/content/345/6200/1074.full}}
+Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
+Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
+A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acting on standing variation \citep{Rabbit}.
+There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}).
+But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
 In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize.
@@
 Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.
-nce of the domesticated \emph{tb1} allele in natural populations of teosinte.
+\emph{gt1}, \emph{tru1}, and \emph{te1}).
+These data, in conjunction with more exhaustive phenotyping, should help to further clarify the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
 

### 2014-09-04 ResponseToReview.tex
Edits
line numbers, text

@@
 
-\lev{need to add something here. I had added something I thought in the text...}
+\lev{Do you guys think what we tweaked in the abstract and the discussion good enough to have addressed this point?}
 
@@
 
-\res{We have expanded on this concept in the introduction on page \lev{x} lines \lev{x-x} ``Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
+\res{We have expanded on this concept in the introduction on page 4, lines 24-45: ``Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
 Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.

### 2014-09-07 manuscript.tex
edits
to response, text, line/page #s

@@
 
-\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 1110 maize and teosinte individuals using a co-dominant PCR assay.
+\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 837 maize and teosinte individuals using a co-dominant PCR assay.
 Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
@@
 In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize.
-%some effort has been invested into identifying agronomically useful loci in wild relatives \citep{}understanding how alleles from wild relatives behave in domesticated crops \citep[e.g.][]{Bai2007}, or identifying agronomically usefulvery little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors \citep{Whitton1997}.
-%In addition to the loss of diversity due to selection on loci important for domestication, demography and the effects of linked selection have acted to decrease diversity genome-wide, leading to lower genetic diversity in many domesticates relative to their wild progenitors \citep{KovachMcCouch2008}.
-
-%Many of these wild alleles no longer found in their domestic counterparts often harbor beneficial traits such as disease resistance or drought tolerance, and an understanding of the distribution of this diversity in the wild and its phenotypic effects has become increasingly useful to crop improvement \citep{KovachMcCouch2008}.
-%For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
-%\jri{why not include some specific teosinte examples like those cited in Hufford 2012 TIG?} Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize \citep{Gallavotti2004, SigmonVollbrect2010}.
 
@@
 \emph{parviglumis} and ssp.
-\emph{mexicana} genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
 SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
@@
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses.
-SAS code used for analysis is available at \url{XXXX}.
+SAS code used for analysis is available at \lev{\url{XXXX}}.
 
@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals.
-\jri{does this conflict with the 1100 mentioned in the summary?} \mbh{So for the rest of the samples, did we have only one PCR that worked or no amplification?
-We should clarify here the number for which we obtained \emph{Hopscotch} data and also use this number in the abstract.
-In the methods we can keep the total number we screened (1100).}
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1100 screened.
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
@@
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive in all populations except La Mesa, \jri{positive D indicates too few rare variants.
-the discussion also says D is negative everywhere.
-is D positive or is the text wrong?
-please check that discussion and results are consistent and both match what is in the table.} \mbh{based on my read of the manuscript, the text is correct here (and it agrees with Table 2 and the Discussion)...if D is negative in La Mesa, that indicates an excess of rare variants in this population; also, the consistently negative D comment in the Discussion is referring to the 5' region of tb1 which shows negative D across populations so this is correct; check me on this, though, LV} indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
@@
 
-\jri{matt, do we inclue mention of weber 2007 in the introduction as part of the motivation?
-might be good?} Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
+Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
 \citet{Weber2007} described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
@@
 Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
-\jri{do we mention photoperiod as a possible explanation for phenotyping 1 results?}
-%Finally, although photoperiod for Phenotyping 2 reasonably approximated that of the normal \emph{parviglumis} growing season, greenhouse-specific environmental conditions (plant density, light regime, etc...) may have contributed to tillering responses different from those found in nature, obscuring the effect of the \emph{Hopscotch} insertion on variation.
 
@@
 Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus may play an ecological role in teosinte.
+
 In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in a diverse sample of \emph{parviglumis} from a natural population and the phenotypic consequences of variation at \emph{tb1} thus remain unclear.

### 2014-09-07 ResponseToReview.tex
edits
to response, text, line/page #s
Edits
manuscript, réponse

@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-We have clarified our sampling methodology on pages 8-9, lines 174-185.
+We have clarified our sampling methodology on page 9, lines 193-205.
 Finally, plants were situated randomly in the greenhouse prior to genotyping, which effectively represents a randomized experimental design.}
@@
 
-\res{We have reported sample size information for all three genotypes on page 11 lines 266-267.}
+\res{We have reported sample size information for all three genotypes on page 12 lines 286-289.}
 
@@
 
-\lev{am not seeing where this is in the text}
+\lev{Matt, I am not seeing where this is in the text. Do you need me to add something to this point?}
 

### 2014-09-07 manuscript.tex
Edits
manuscript, réponse

@@
 \ref{Fig4Boxplots}), tiller number, or culm diameter.
-Only on day 40 did we observe a weak but statistically insignificant ($p=0.0848$) \mbh{would be good to add the $r^2$ value as well if we have it} correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
-\jri{is this correct?} \mbh{Looks correct to me; good idea to double-check, LV}
+Only on day 40 did we observe a weak but statistically insignificant ($p=0.0848$, $r^2$ = 0.02) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
 

### 2014-09-07 ResponseToReview.tex
edits
line and page numbers in response

@@
 Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
-We have, however, added text to the conclusion (Page 15, lines 372-377) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
+We have, however, added text to the conclusion (Page 15, lines 389-395) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
 ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
@@
 This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion (\emph{i.e.}, the domestication allele at \emph{tb1}) in teosinte. 
-We have added clarifying text to the introduction (Page 5, lines 55-61): ``The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+We have added clarifying text to the introduction (Page 5, lines 64-69): ``The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.

### 2014-09-07 ResponseToReview.tex
edits
edits to response to review. all page and line numbers should be
accurate now

@@
 We have added clarifying text to the introduction (Page 5, lines 64-69): ``The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
-Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of ssp.
-\emph{parviglumis}, ssp.
-\emph{mexicana}, and landrace maize; yet little is known about the frequency at which the \emph{Hopscotch} is segregating in natural populations."}
+Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations."}
 
@@
 We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte.
-Changes can be seen on page 5, lines 71-77: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007}, exhibiting changes in expression levels in response to shading.
-\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, whereas under high density, plants showed significantly reduced tillers and grew taller \citep{LukensDoebley1999}.
-Based on these results we hypothesized that \emph{tb1} may play a role in the ecology of teosinte."}
+Changes can be seen on page 5, lines 74-79: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
+\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller.
+Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density."}
 
@@
 
-\lev{check numbers}
-\res{We have added a range of values (1-38 individuals) per population in the methods, pages 5-6, lines 85-90: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{We have added a range of values (1-38 individuals) per population in the methods, pages 6, lines 88-93: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
 
@@
 
-\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page 6, lines 99-104) following primer sequences: 
+\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page 6, lines 102-105) following primer sequences: 
 ``Homozygotes show a single band for absence of the \emph{Hopscotch} element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."
@@
 If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
-We have clarified this in the text on page 6, lines 92-104.
+We have clarified this in the text on page 6, lines 97-107.
 : ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
@@
 
-\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 7, lines 143-144: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 6, lines 112-114: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page 8, lines 154-157 of the text : ``Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
+We have further clarified this on page 7, lines 124-129 of the text : ``Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
 These data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
@@
 
-\res{We have clarified the dataset in the paper on page 8, lines 151-154: ``We used teosinte (ssp.
+\res{We have clarified the dataset in the paper on page 7, lines 121-124: ``We used teosinte (ssp.
 \emph{parviglumis} and ssp.
-\emph{mexicana}) SNP genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
-SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA)."}
+\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA)."}
 
@@
 
-\res{We have edited the text to clarify, page 9, lines 199-200 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
+\res{We have edited the text to clarify, page 9, lines 203-205 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
 
@@
 
-\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section, \lev{page x, linesx} : ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section, page 10, lines 217-221: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}
@@
 
-\res{We have added ratios to the text: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
+\res{We have added ratios to the text, page 10, lines 227-230: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
 

### 2014-09-07 manuscript.tex
Edits
minor edits

@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \url{XXX}]{Hufford2013}.
+Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \lev{Matt is your code available through your github?} \url{XXX}]{Hufford2013}.
 In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1100 screened.
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).

### 2014-09-09 ResponseToReview.tex
Edits to response to review
Edits to point 6 in the response to review; corresponding to changes in
introduction concerning domestication and shade avoidance syndrome in
grasses

@@
 
-\lev{Do you guys think what we tweaked in the abstract and the discussion good enough to have addressed this point?}
+\res{We have editied the introduction to include a more broad explanation of domestication and shade avoidance responses in plants. Edits can be seen on page 2, lines 22-45 and page 5, lines 70-78.}
 

### 2014-09-09 manuscript.tex
Edits
added URL for chunking script

@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \lev{Matt is your code available through your github?} \url{XXX}]{Hufford2013}.
+Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \url{http://dx.doi.org/10.6084/m9.figshare.1165577}]{Hufford2013}.
 In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.

### 2014-09-09 manuscript.tex

@@
 Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovachMcCouch2008}.
 In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.

### 2014-09-09 manuscript.tex
minor edits

@@
 SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \url{http://dx.doi.org/10.6084/m9.figshare.1165577}]{Hufford2013}.
+Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \lev{Matt is your code available through your github?} \url{XXX}]{Hufford2013}.
 In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.

### 2014-09-10 manuscript.tex
minor

@@
 In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
-A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acting on standing variation \citep{Rabbit}.
+A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acted on standing variation \citep{Rabbit}.
 There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}).

### 2014-09-10 manuscript.tex
minor edits

--- 2014-09-09 manuscript.tex
@@
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
-Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
+Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as the `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
 Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
 Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovachMcCouch2008}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovachMcCouch2008}.
 In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.

### 2014-09-10 manuscript.tex
edits to introduction

--- 2014-09-10 manuscript.tex
@@
 Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
-Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
+Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as the `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
 Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
@@
 In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
-A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acted on standing variation \citep{Rabbit}.
-There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}).
 But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
-In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize.
+In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1}, and specifically the domesticated haplotype at \emph{tb1}, in natural populations of the wild ancestor of domesticated maize.
 
@@
 in southwest Mexico \citep{Piperno2009, Matsuoka2002}.
-Domesticated maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
+Maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
 Additionally, large, naturally-occurring populations of both \emph{parviglumis} and the highland teosinte \emph{Zea mays} ssp.
@@
 In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
+An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome, a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
 The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
-\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller.
-Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density.
+\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered but that under high density, plants showed significantly reduced tillers and grew taller.
+Based on these results we hypothesize that the \emph{Hopscotch} (\emph{i.e.}, the domesticated allele) at \emph{tb1} may play a role in the ecology of teosinte, especially in high-density populations.
 In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.

### 2014-09-10 manuscript.tex
edits
sas code included
edits through results

@@
 \emph{parviglumis} and ssp.
-\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
-SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
+\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
+SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
 Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
-These data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
 Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.
@@
 
-\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} insertion is segregating at highest frequency (0.44) in our initial genotyping sample set.
+\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo; \citealt{Hufford2010}) where the \emph{Hopscotch} insertion is segregating at highest frequency (0.44) in our initial genotyping sample set.
 In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites within the population where genotyped individuals were homozygous or heterozygous for the insertion.
@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to those reported genome-wide in previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).
@@
 For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is less negative in San Lorenzo (Table \ref{Table2Diversity}, Appendix 2, See Supplemental Materials with the online version of this article).
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}, Appendix 2, See Supplemental Materials with the online version of this article).
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
@@
 \ref{Fig4Boxplots}), tiller number, or culm diameter.
-Only on day 40 did we observe a weak but statistically insignificant ($p=0.0848$, $r^2$ = 0.02) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
+Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
 

### 2014-09-10 manuscript.tex
edits through results

@@
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses.
-SAS code used for analysis is available at \lev{\url{XXXX}}.
+SAS code used for analysis is available at \url{ http://dx.doi.org/10.6084/m9.figshare.1166630}.
 
@@
 While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
-
-Many of our populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, perhaps suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
+Many of our populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster of \emph{parviglumis} identified by \citet{Fukunaga2005}, perhaps suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
 Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from the Jalisco cluster include gene flow from maize, genetic drift, and natural selection.

### 2014-09-10 manuscript.tex
final edits to manuscript, minor response edits

@@
 While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
-Many of our populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster of \emph{parviglumis} identified by \citet{Fukunaga2005}, perhaps suggesting a different history of the \emph{tb1} locus in this region than in the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
+The \emph{Hopscotch} allele is more prevalent in \emph{parviglumis} than in \emph{mexicana} in our sample (Appendix 1, See Supplemental Materials with the online version of this article), suggesting a different history of the allele amongst teosinte subspecies.
+Moreover, many of our \emph{parviglumis} populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, and further distinguish this region from the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
 Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from the Jalisco cluster include gene flow from maize, genetic drift, and natural selection.
@@
 \ref{Fig3Structure}, Table \ref{Table4Q}).
-Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in a subset of our \emph{parviglumis} populations.
+Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in the Jalisco cluster of our \emph{parviglumis} populations.
 
@@
 In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}.
-Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in \emph{parviglumis}.
+Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in Jaliscan populations of \emph{parviglumis}.
 
@@
 
+\clearpage
 % Table 1
@@
   \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}%table caption
-
     \begin{tabular}{cccc}\\\toprule%number of columns and line across top of table

### 2014-09-10 ResponseToReview.tex
final edits to manuscript, minor response edits
edits to response

@@
 
-We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript.
+We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our previously-submitted manuscript (AJB-D-14-00246).
 In the attached revision, we have addressed the majority of issues raised, but, while we feel that the new experiments suggested by Reviewer 1 would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.
@@
 
-\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype of reduced tiller number, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte.
+\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype of reduced tiller number, is at higher than expected frequency in a subset of teosinte populations (the Jaliscan cluster of \emph{parviglumis}) and may play an adaptive role in certain environments encountered by teosinte.
 Surprisingly, based on our grow-out of individuals from the San Lorenzo population, tiller number is not a major phenotypic difference between teosinte plants with and without the \emph{Hopscotch} insertion, perhaps indicating the insertion affects multiple phenotypes in teosinte}
@@
 Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovackMcCouch2008}.
+For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovachMcCouch2008}.
 In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
-A recent paper examining the genomic changes associated with rabbit domestication found that many domestication phenotypes are due to mutations in many small effect loci, and that many of these mutations are still found segregating in natural populations of wild rabbit, suggesting that selection acting on standing variation \citep{Rabbit}.
-There are numerous examples of weedy plants evolving to mimic their domestic counterpart, and even some instances where weedy plants have evolved domesticates (reviewed in \citealt{ellstrand2010crops}).
 But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
-In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize."}
+In this paper, we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize."}
 
@@
 \citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller.
-Based on these results we hypothesize that \emph{tb1} may play a role in the ecology of teosinte, especially in populations where it grows at a very high density."}
+Based on these results we hypothesize that the \emph{Hopscotch} (\emph{i.e.}, the domesticated allele) allele at \emph{tb1} may play a role in the ecology of teosinte,  especially in high-density populations."}
 
@@
 The low frequency of \emph{Hopscotch} in \emph{mexicana} precluded a phenotyping experiment with this subspecies since a very high number of seeds would have to be germinated in order to obtain enough individuals with the \emph{Hopscotch} allele.}
-
-\lev{Matt, I am not seeing where this is in the text. Do you need me to add something to this point?}
 
@@
 We have further clarified this on page 7, lines 124-129 of the text : ``Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
-These data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
 Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}."}
@@
 \emph{parviglumis} and ssp.
-\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional SNP.
+\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
 SNP data from \citep{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA)."}

### 2014-09-10 ResponseToReview.tex
edits to response

@@
 Though our sample of \emph{mexicana} is range-wide, we only had access to seeds from 17 populations.
-Frequency of the \emph{Hopscotch} insertion was low in these populations relative to \emph{parviglumis}, an interesting point that we have discussed further on page \mbh{X}, lines \mbh{XX} of the text.
+Frequency of the \emph{Hopscotch} insertion was low in these populations relative to \emph{parviglumis}, an interesting point that we have discussed further on page 13, lines 17-19 of the text.
 The low frequency of \emph{Hopscotch} in \emph{mexicana} precluded a phenotyping experiment with this subspecies since a very high number of seeds would have to be germinated in order to obtain enough individuals with the \emph{Hopscotch} allele.}

### 2014-09-10 Supplement_mbh.tex
edits to supplement

--- 2014-05-29 Supplement_mbh.tex
@@
 TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the \emph{Hopscotch} insertion.
-Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.}
+Black indicates individuals not genotyped for the \emph{Hopscotch} insertion.
+EjuA refers to individuals from population Ejutla A, EjuB from Ejutla B, SLO from San Lorenzo, and MSA from La Mesa.
+Remaining individuals are lines of maize (\emph{Zea mays} ssp.
+\emph{mays}).}
+
 \end{center}

### 2014-09-10 ResponseToReview.tex

@@
 We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
-We have clarified our sampling methodology on page 9, lines 193-205.
+We have clarified our sampling methodology on page 9, lines 188-206.
 Finally, plants were situated randomly in the greenhouse prior to genotyping, which effectively represents a randomized experimental design.}
@@
 
-\res{We have reported sample size information for all three genotypes on page 12 lines 286-289.}
+\res{We have reported sample size information for all three genotypes on page 12 lines 281-284.}
 
@@
 Nonetheless, these are extensive additional experiments which we are unable to include at this time. 
-We have, however, added text to the conclusion (Page 15, lines 389-395) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
+We have, however, added text to the conclusion (Page 15, lines 389-393) of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte:
 ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
@@
 This finer level of detail sheds light on the evolutionary history of the \emph{Hopscotch} insertion (\emph{i.e.}, the domestication allele at \emph{tb1}) in teosinte. 
-We have added clarifying text to the introduction (Page 5, lines 64-69): ``The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+We have added clarifying text to the introduction (Page 5, lines 59-64): ``TThe effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations."}
@@
 
-\res{We have editied the introduction to include a more broad explanation of domestication and shade avoidance responses in plants. Edits can be seen on page 2, lines 22-45 and page 5, lines 70-78.}
+\res{We have editied the introduction to include a more broad explanation of domestication and shade avoidance responses in plants. Edits can be seen on page 4, lines 24-38 and page 5, lines 65-70.}
 
@@
 
-\res{We have expanded on this concept in the introduction on page 4, lines 24-45: ``Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
-Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
+\res{We have expanded on this concept in the introduction on page 4, lines 24-40: ``Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}.
+Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as the `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
 Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
@@
 But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
-In this paper, we present an ecological genetic analysis of the domestication locus \emph{tb1} in natural populations of the wild ancestor of domesticated maize."}
+In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1}, and specifically the domesticated haplotype at \emph{tb1}, in natural populations of the wild ancestor of domesticated maize."}
 
@@
 We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte.
-Changes can be seen on page 5, lines 74-79: ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
-\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered, but that under high density, plants showed significantly reduced tillers and grew taller.
-Based on these results we hypothesize that the \emph{Hopscotch} (\emph{i.e.}, the domesticated allele) allele at \emph{tb1} may play a role in the ecology of teosinte,  especially in high-density populations."}
+Changes can be seen on page 5, lines 65-75: ``In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
+An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome, a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
+The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
+\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered but that under high density, plants showed significantly reduced tillers and grew taller.
+Based on these results we hypothesize that the \emph{Hopscotch} (\emph{i.e.}, the domesticated allele) at \emph{tb1} may play a role in the ecology of teosinte, especially in high-density populations.}
 
@@
 Though our sample of \emph{mexicana} is range-wide, we only had access to seeds from 17 populations.
-Frequency of the \emph{Hopscotch} insertion was low in these populations relative to \emph{parviglumis}, an interesting point that we have discussed further on page 13, lines 17-19 of the text.
+Frequency of the \emph{Hopscotch} insertion was low in these populations relative to \emph{parviglumis}, an interesting point that we have discussed further on page 13, lines 310-313 of the text.
 The low frequency of \emph{Hopscotch} in \emph{mexicana} precluded a phenotyping experiment with this subspecies since a very high number of seeds would have to be germinated in order to obtain enough individuals with the \emph{Hopscotch} allele.}
@@
 
-\res{We have added a range of values (1-38 individuals) per population in the methods, pages 6, lines 88-93: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
+\res{We have added a range of values (1-38 individuals) per population in the methods, pages 5-6, lines 83-88: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}
 
@@
 
-\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page 6, lines 102-105) following primer sequences: 
+\res{The sizes of amplicons are now mentioned explicitly in the Materials and Methods (page 5, lines 97-102) following primer sequences: 
 ``Homozygotes show a single band for absence of the \emph{Hopscotch} element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article)."
@@
 If PCR products were clearly resolved for only one of these reactions, we included one allele for that individual in our data set rather than infer the diploid genotype.
-We have clarified this in the text on page 6, lines 97-107.
+We have clarified this in the text on page 6, lines 92-102.
 : ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
@@
 
-\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 6, lines 112-114: ``These analyses only included populations ($n = 32$) in which eight or more individuals were sampled."}
+\res{Thirty-two populations were included in these analyses, and we have edited the text to reflect this number on page 6, lines 108-109: ``These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled."}
 
@@
 Since data for our analyses were either summarized over the last several decades (climatic data) or are likely to be stable over time (soil data), they have potentially influenced current frequencies of the \emph{Hopscotch} insertion in our sampled populations.
-We have further clarified this on page 7, lines 124-129 of the text : ``Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
+We have further clarified this on page 7, lines 119-124 of the text : ``Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
 Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
@@
 
-\res{We have clarified the dataset in the paper on page 7, lines 121-124: ``We used teosinte (ssp.
+\res{We have clarified the dataset in the paper on pages 6-7, lines 116-119: ``We used teosinte (ssp.
 \emph{parviglumis} and ssp.
@@
 
-\res{We have edited the text to clarify, page 9, lines 203-205 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
+\res{We have edited the text to clarify, page 9, lines 198-200 : ``Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days."}
 
@@
 
-\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section, page 10, lines 217-221: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
+\res{Thank you for the suggestion, we have edited the text accordingly and have added our explanation of the inclusion of culm diameter at the end of the Materials and Methods section, pages 9-10, lines 212-216: ``Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
 Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses."}
@@
 
-\res{We have added ratios to the text, page 10, lines 227-230: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
+\res{We have added ratios to the text, page 10, lines 222-225: ``Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico."}
 

### 2014-10-24 AJB_Res2.tex
Response to review, second submission
Notes to response; what we can/not change

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
\newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
\usepackage{lipsum}
\usepackage{url}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}

\begin{document}

Associate Editor and Reviewer comments:

Dear Dr. Hufford and coauthors - 
    Thank you for your thorough response to previous criticisms of your manuscript and your attention to improving the manuscript.  I sent the current manuscript to one of the previous reviewers (Reviewer 2 here) and to a new reviewer (Reviewer 1 here) to get a more independent view of the work.  Reviewer 2 has recommended acceptance of the manuscript, albeit a bit begrudgingly, whereas Reviewer 1 recommends rejection and voices some of the same objections that arose in the original rounds of review - the lack of data on tb1 expression in the presence of Hopscotch and the confusion over explanations related to Hopscotch and explanations about evolution of the tb1 gene as a whole (eg, see Reviewer 2’s statements, “the paper is written explicitly assuming a connection between Hopscotch and tb1 expression in teosinte, and often conflates variation at the tb1 locus with presence or absence of Hopscotch” and “veering in focus between the purported object of this study, hopscotch
elements, and the tb1 gene itself”).  Your Abstract provides an example of how a reader could be confused about the focus of the paper: “Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for Hopscotch in teosinte, but a greenhouse experiment does not detect a strong association between tb1 and tillering in teosinte.”  From this sentence, the reader would infer that you tested for selection on Hopscotch and did the greenhouse experiment on different tb1 genotypes, when the reverse is actually true.
     Because these reviewers reach such different opinions, I have spent a considerable amount of time giving the manuscript a very careful reading and also reviewing some of the most important literature you cite.  Based on this review, I’ve concluded that much of the criticism from Reviewer 1 arises because you have not clearly stated your hypotheses.  You voice some unknowns, you make a couple of hypotheses, and you have some aims, but none of this is organized in a way that gives the reader a clear picture of the logic of your thinking or the structure of the study.  For example, you have a hypothesis in your final paragraph of the Introduction (about density effects on tillering), but it is not actually something you tested, because you did not test different densities.  You would be well served, I think, by discussing amongst yourselves exactly what hypotheses the various pieces of your study were designed to test and then stating those clearly at the end of the
Introduction.  If you state those hypotheses clearly, then you are much less likely to be accused of mix-and-match arguments that switch between Hopscotch itself and the tb1 locus as a whole.  Based on my understanding of your paper and some of the major literature you cite, you seem to be testing hypotheses about (1) the frequency of the Hopscotch insertion in natural population (genotyping survey) based on some preliminary studies; (2) the effect of Hopscotch on tillering (greenhouse experiment) based on what is known about the effect of Hopscotch in maize and the evidence that Hopscotch was present in teosinte populations prior to domestication; (3) the possibility of explaining the high frequency of Hopscotch in teosinte by introgression from maize (LD tests using sequences); and (4) the possibility of past or present selection on the entire locus in natural teosinte populations (Tajima D test).  Of these, the weakest part is the selection analysis, and I think you might
be better off dropping that part and just focusing on the first three hypotheses.  Once you have some of the expression data that all three reviewers have called for, you could return to the question of the ecological role for Hopscotch and tb1 in natural populations.
    In addition to these concerns, I found several oversights in the paper.  For example, you state that there are 1110 individuals sampled, but based on the numbers of populations and landraces you give and the average sample size of each, the total number of individuals should be 1627 (ave. 2 accessions/landrace x 247 landraces + ave. 11 ind/population x 103 teosinte populations).  The problem appears to lie, at least in part, in a confusion about alleles vs individuals, since most of the landraces shown in Appendix S2 show 2 alleles sampled, not 2 individuals.  Even if the number is actually 1110, then I found it troubling that nearly 25% of those could not be successfully genotyped.  Second, your figure showing the PCR assay (Appendix S4, not S2 as stated in the text) has an incorrect heterozygote genotype (assuming “no Hop/Pif” is the heterozygote) that does not show the three bands you describe in the text.  Third, you state that you calculate Watterson’s estimator of
population mutation rate, but you never mention it again after the Methods.
    In summary, despite the significant improvements to the current manuscript, I cannot recommend publication, due to the consistent call for expression data from three reviewers, continued confusion about the focus of the paper, and problems relating to sampling and genotyping.  I agree with Reviewer 1 that there are “excellent data in this manuscript that deserves to be published in some form”, and I hope you can find a way to revise the manuscript to be acceptable in a different venue.

Yours sincerely,
Andrew Schnabel
Associate Editor

Reviewer #1: Vann review AJB 2014

Summary
This manuscript is a study of the phenotypic effect in a wide sample of teosinte populations of the Hopscotch retrotransposon known to regulate expression of teosinte branched1 (tb1) in maize.  It also contributes to our understanding of the population genetic history of teosinte and of the Hopscotch element.  The basic argument of the manuscript is as follows: 1) Hopscotch regulation of tb1 has been shown to be important in controlling plant architecture in domesticated maize, 2) Hopscotch segregates in natural populations of teosinte (ancestor of maize), therefore 3) what is the  effect of Hopscotch presence or absence on tb1 expression and phenotype in teosinte.  The authors focus on tillering and do not explore other known effects of tb1 on internode elongation or sexuality.

The main problem for me about this argument is that there is an implicit link between the presence or absence of Hopscotch and levels of tb1 expression in teosinte plants, and the effect this has on tillering.  There is excellent data in this manuscript that deserves to be published in some form, but I do not believe that the argument that is constructed is valid without showing that Hopscotch actually affects tb1 expression in teosinte. It would be a different matter if the paper were merely focused on the effect of the presence or absence of Hopscotch, but the paper is written explicitly assuming a connection between Hopscotch and tb1 expression in teosinte, and often conflates variation at the tb1 locus with presence or absence of Hopscotch.  This is a problem, especially when the data show little effect of Hopscotch presence or absence on the measured phenotypes.  Below I expand on this summary and also present some other queries.

Introduction
page 5, lines 21-28.  The logic here is unclear.  The Lukens and Doebley experiment showed that maize lines with an introgressed teosinte tb1 allele were affected by shading.  It does not follow that the domesticated allele (Hopscotch containing allele) should play a role in teosinte ecology, especially since maize with the Hopscotch allele is insensitive to shading.  This argument would make more sense if the authors had shown that Hopscotch alleles affect tb1 expression in teosinte.

Page 5, lines 28-30:  Absolutely nothing wrong with this aim, as far as it goes.  However, the paper tries to do more than just this, which is a problem because of the flaws in the argument summarized above.
["In this study we aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic effects of the insertion in parviglumis."]

Materials and Methods
Page 6, line 30: Minor point, and I am not an expert on population genetics, but why say that you only include populations in which 8 or more chromosomes were sampled, rather than 8 or more individuals.  If I am missing something, it could be that others will too, so please explain.

Page 6, line 30- page 7, line 2: Testing the adaptive hypothesis.  I understand that in the end there was no evidence for Hopscotch allele frequency being correlated with environmental variables.  Was there evidence for any of the other SNPs being so correlated?  How robust is this analysis against false positives?  The adaptive hypothesis is attractive, but it is unclear why you would make it.  You do not make the claim in these lines, but it seems to me that it is the three classes of teosinte alleles that might be more likely to correlate with environment (since Studer et al. 2012 showed that they correlated with taxonomy) than the presence or absence of the Hopscotch element.

Page 9, line 15: In the STRUCTURE analysis, why was K=2 chosen?  Had you already done exploratory runs with a range of K values or is there some theoretical reason for assuming K=2.  Either way, please explain.

Page 10, line 12: Why is culm diameter not believed to correlated with tillering index?  Is there a reference for this? What do your results show?  More importantly, if you want to prove that Hopscotch has more effect on tillering than on other independent traits you really need to do more comprehensive phenotyping.  Please explain more fully what this particular analysis is meant to achieve.
["Culm diameter is not believed to be correlated with tillering index or variation at tb1 and is used as our independent trait for phenotyping analyses."]

Discussion
Page 14, 3rd para: This paragraph starts off by examining explanations of genetic drift and natural selection for different frequencies of Hopscotch alleles amongst teosinte populations.  After discussing evidence for bottlenecks, the argument shifts to the lack of diversity in the 5' UTR of the tb1 gene, concluding that the results are consistent with the action of selection on the 5'UTR, and that there might be an ecological role for the gene in some populations.  This is a classic example of the veering in focus between the purported object of this study, hopscotch elements, and the tb1 gene itself.  In the end there is no explanation given for the hopscotch frequencies, instead the focus is switched back to the gene itself.  To the reader there is an implication (perhaps unintentional) that there is a causative link between Hopscotch and variation in the tb1 alleles, whereas I would say this has simply not been shown in teosinte, and the evidence in this manuscript would
argue against such a link.

Page 15, line 10: the argument at this point appears to be that we don't see the effect of Hopscotch in teosinte because of variation in other loci.  I would like more explanation of how that might occur, given that the Hopscotch element in maize upregulates tb1 expression so much that it explains most of the variation.

Page 15, line 14, discussion of Studer et al. (2012).  Studer et al. showed that there is an allelic series of tb1 in teosinte, and that this series is correlated with taxonomy of maize subspecies.  They also showed that differences between the three classes of tb1 alleles were significantly correlated with mean internode length of the uppermost lateral branch as well as two inflorescence traits, but that the classes did NOT correlate with variation in tillering.  Studer et al. concluded that there was evidence for allelic series only for the internode length and inflorescence characters, not for tillering.  Their explanation was NOT that tillering index was controlled by both tb1 and other loci, but that there was no evidence for variation in tb1 alleles affecting tillering.  Changes in other loci were invoked by Studer et al. (along with environmental effects on seed development) to explain why separate families differed from each other irrespective of the tb1 allele they
contained.  Again, at the end of this paragraph (page 15, lines 27-28), there is confusion over what is being explained, as the last sentence refers to a lack of correlation between Hopscotch genotype and tillering, but the discussion of the Studer et al. (2012) paper does not concern Hopscotch, but only an allelic series at the tb1 locus.  In fact, the Studer paper does not mention Hopscotch elements at all and the allelic series analyzed are only the coding region of the tb1 gene.

Reviewer #2: This paper presents some truly tantalizing results: a higher frequency of Hopscotch  in teosinte than expected, some evidence for positive selection acting on tb1 in teosinte, and no evidence for recent introgression of Hopscotch into teosinte from maize. The methods used and statistical analyses are all appropriate and carefully conducted, and the authors don't overstate their conclusions. The paper, as it stands, is well written and easy to follow. I found this to be true before it had been revised, and the revisions have only improved it.

However, it is extremely unlucky that the authors found no convincing connection between Hopscotch genotype and tillering, or between genotype and climatic variables. It is precisely because the results are so interesting that I was hoping for further functional characterization - either in the form of expression analyses (even of just tb1), or in the form of more complete phenotypic characterization. I do appreciate that this would require a significant input of time, at some risk of obtaining yet more negative data. Although the authors have already spent a lot of time and effort trying to detect an effect linked to Hopscotch, the ecological role of Hopscotch in teosinte (if any) remains elusive.

\end{document}

### 2014-10-24 AJB_Res2.tex
Edits to AJB Res 2
Comments in the response as to what can be addressed.

@@
 Once you have some of the expression data that all three reviewers have called for, you could return to the question of the ecological role for Hopscotch and tb1 in natural populations.
+\lev{No on expression data and additional experiments. I thought our hypotheses/aims were rather clearly stated, but perhaps we can reword if you guys think that they need clarification.}
     In addition to these concerns, I found several oversights in the paper.
@@
 population mutation rate, but you never mention it again after the Methods.
+\lev{Will check numbers. I think the confusion is in the individuals vs. alleles.}
     In summary, despite the significant improvements to the current manuscript, I cannot recommend publication, due to the consistent call for expression data from three reviewers, continued confusion about the focus of the paper, and problems relating to sampling and genotyping.
@@
 Below I expand on this summary and also present some other queries.
-
+\lev{no expression analyses.
+should we downplay any mention of expression in the paper?
+or somehow make it more clear that it will be a nice follow up in another study?
+If all the reviewers we have had so far feel expression is crucial, I do think that we'll have to find a way to address the lack of expression in the paper}
 Introduction
@@
 This argument would make more sense if the authors had shown that Hopscotch alleles affect tb1 expression in teosinte.
+\lev{I don't understand.
+If Hopscotch = taller plant with fewer tillers then why couldn't it play a role in the ecology of teosinte?
+Our phenotyping obviously doesn't support this pattern in teosinte, but I don't understand how the hypothesis is a problem.}
 
@@
 
+\lev{I don't think there is a big problem with this.
+I think it's clearly stated.
+Is the problem he is pointing out that we have explicit hypotheses about the phenotypic effects and that it may play a role in the ecology that we don't include in this statement?}
 Materials and Methods
@@
 If I am missing something, it could be that others will too, so please explain.
+\lev{This was just our sampling cut-off for analyses...}
 
@@
 2012 showed that they correlated with taxonomy) than the presence or absence of the Hopscotch element.
+\lev{I only used the Hopscotch, coded as a SNP.
+Tanja's paper would have included the other SNPs in the panel.
+I don't remember off the top of my head if she found any that were significant on Chromosome 1, but that is easily checked.
+Should we mention SNPs she identified in her paper her?
+Yes, the three classes of teosinte likely correlate with environment, but that's not something we address here.
+Should we take out BayEnv stuff all together?
+I like it in, but it doesn't add much I guess.}
 
@@
 Either way, please explain.
+\lev{I think we chose K=2 because we expected things to group either into a teosinte cluster or a maize cluster}
 
@@
 ["Culm diameter is not believed to be correlated with tillering index or variation at tb1 and is used as our independent trait for phenotyping analyses."]
-
+\lev{I think we'd cite studer for this. we wanted a trait that would be independent so that we could compare genotype + culm patterns to genotype + tillering patterns.}
 Discussion
@@
 argue against such a link.
+\lev{I don't mind this paragraph.
+It ties in findings from previous studies that identified evidence of bottleneck in some of these pops with our findings, which i think is useful.
+We state that a population bottleneck could explain the pattern that we see in SLO, but that it cannot account for this pattern across multiple populations in the Jalisco region.}
 
@@
 I would like more explanation of how that might occur, given that the Hopscotch element in maize upregulates tb1 expression so much that it explains most of the variation.
+\lev{There were other loci in the initial QTL paper that are involved in the difference in branching architecture.
+Plus we now know of a lot of other loci (mentioned in paper) that act in the SAS and could affect plant architecture.}
 
@@
 In fact, the Studer paper does not mention Hopscotch elements at all and the allelic series analyzed are only the coding region of the tb1 gene.
+\lev{I think we should reread this section of the studer paper and check.
+If I remember correctly they found differences between their maize isogenic lines that were comparable to the differences they found among maize with introgressed segments of tb1}
 
@@
 
+\lev{nothing to change for reviewer 2}
+
 \end{document}

### 2014-11-05 manuscript.tex
changed fig names, added figs back in for biorxiv submission
Edits
Edits to manuscript for future submission to PeerJ

@@
 \newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %same for matt hufford in red
-\newcommand{\lev}[1]{\textcolor{green}{\emph{\scriptsize #1}} } %same for laura vann in green
+\newcommand{\lev}[1]{\textcolor{magenta}{\emph{\normalsize #1}} } %same for laura vann in green
 
@@
 
-\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 837 maize and teosinte individuals using a co-dominant PCR assay.
+\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of \lev{837} maize and teosinte individuals using a co-dominant PCR assay.
 Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}\lev{We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations}; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
-We ran STRUCTURE at K=2 under the linkage model, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
+We ran STRUCTURE at K=2 under the linkage model, with the assumption being that individuals fall into either a maize or teosinte cluster, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
 
@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for \lev{837 individuals of the 1,100 screened}.
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
@@
 \begin{center}
-   %\includegraphics[width=0.6\textwidth]{Fig1.MapBalsas.pdf}
+   \includegraphics[width=0.6\textwidth]{Fig1MapBalsas.pdf}
    \end{center}
@@
   \begin{center}
-   %\includegraphics[width=140mm]{Fig2LDPlots.pdf}
+   \includegraphics[width=140mm]{Fig2LDPlots.pdf}
    \end{center}
@@
   \begin{center}
-   %\includegraphics[width=90mm]{Fig3Structure.pdf}
+   \includegraphics[width=90mm]{Fig3Structure.pdf}
      \end{center}
@@
   \begin{center}
-   %\includegraphics[width=90mm]{Fig4.BoxplotPheno1.pdf}
+   \includegraphics[width=90mm]{Figure4BoxplotPheno1.pdf}
    \end{center}

### 2014-11-05 AJB_Res2.tex
Edits
Edits to response noting what has been completed and edits to
manuscript for submission to PeerJ

@@
 \newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
-\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
+\newcommand{\lev}[1]{\noindent \textcolor{magenta}{{#1}} \\}
 \newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
@@
 
-Page 9, line 15: In the STRUCTURE analysis, why was K=2 chosen?
+\lev{\bf{DONE}}Page 9, line 15: In the STRUCTURE analysis, why was K=2 chosen?
 Had you already done exploratory runs with a range of K values or is there some theoretical reason for assuming K=2.
@@
 
-Page 10, line 12: Why is culm diameter not believed to correlated with tillering index?
+\lev{\bf{DONE}} Page 10, line 12: Why is culm diameter not believed to correlated with tillering index?
 Is there a reference for this?
@@
 ["Culm diameter is not believed to be correlated with tillering index or variation at tb1 and is used as our independent trait for phenotyping analyses."]
-\lev{I think we'd cite studer for this. we wanted a trait that would be independent so that we could compare genotype + culm patterns to genotype + tillering patterns.}
+\lev{I think we'd cite studer for this. we wanted a trait that would be independent so that we could compare genotype + culm patterns to genotype + tillering patterns.}\\
 Discussion

### 2014-11-05 manuscript.tex
Edits
Edits to response noting what has been completed and edits to
manuscript for submission to PeerJ

@@
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\lev{Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations}, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).
-Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+\lev{Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.} We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).
@@
 We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
-We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{Briggs2007} but is not thought to be affected by the \emph{Hopscotch} insertion. 
+We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{Briggs2007} but has not been shown to be affected by the \emph{Hopscotch} insertion.
+Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data.
+If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype.
 Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 

### 2014-11-05 AJB_Res2.tex
Edits
Edits to manuscript and response notes

@@
 ["Culm diameter is not believed to be correlated with tillering index or variation at tb1 and is used as our independent trait for phenotyping analyses."]
-\lev{I think we'd cite studer for this. we wanted a trait that would be independent so that we could compare genotype + culm patterns to genotype + tillering patterns.}\\
+\lev{This is cited as Briggs et al. 2007 in the text}\\
 Discussion

### 2014-11-05 manuscript.tex
Edits
Edits to manuscript and response notes

@@
 We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
-We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{Briggs2007} but has not been shown to be affected by the \emph{Hopscotch} insertion.
+We also measured culm diameter, a phenotype that differs between maize and teosinte but has not been shown to be affected by the \emph{Hopscotch} insertion \citep{Briggs2007}.
 Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data.

### 2014-11-17 manuscript.tex
Cover letter
Updates to text concerning allelic series info.

@@
 Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes \citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes of the coding domain of \emph{tb1}\citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.

### 2014-11-20 AJB_Res2.tex
Updates to manuscript
Rewrite of allelic series section, same conclusions but rewritten. I
think we disagree with the reviewer on this point.

@@
 If I remember correctly they found differences between their maize isogenic lines that were comparable to the differences they found among maize with introgressed segments of tb1}
+\lev{I have reread his thesis a bunch since getting the review back, and either i am misunderstanding or the reviewer is.
+He is correct that the chapter/paper does not explicitly mention the Hopscotch, but if I am reading the figure correctly all of tb1 ends up in one or more of the introgressed segments from the long arm of chr1.
+The methods also discuss amplifying both the coding region and the control region.
+I don't think we are misinterpreting.
+The paper explicitly says that they find evidence that "different teosinte tb1 chromosomal segments possess distinct genetic factors that confer different amounts of tillering." It goes on to say that they found significant variation in tillering in their control line W22 and their conclusions for tb1 chromosomal segments with respect to tillering is "These results suggest that there are factors other than tb1 affecting phenotype.
+Therefore we cannot conclude that the observed differences in tillering among the teosinte introgressions are caused by tb1 alone".}
+
+\lev{What do you guys think about changing this or leaving it as is?}
 

### 2014-11-20 manuscript.tex
Updates to manuscript
Rewrite of allelic series section, same conclusions but rewritten. I
think we disagree with the reviewer on this point.

@@
 \citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated their phenotypic effects.
-Phenotypes were shown to cluster by taxon, indicating \emph{tb1} may underlie morphological diversification of \emph{Zea}.
-Additional analysis in \citet{StuderDoebley2012} suggested tillering index was controlled both by \emph{tb1} and loci elsewhere in the genome.
+Their findings suggest that different teosinte \emph{tb1} segments contain distinct genetic factors that differentially contribute to tillering in teosinte.
+Furthermore, \citet{StuderDoebley2012} found significant variation in tillering within their control maize lines, concluding that variation in \emph{tb1} alone cannot account for the differences in tillering, and that other genetic factors must also be contributing to differences in tillering in teosinte.
 Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).

### 2014-11-20 manuscript.tex
Edits
Few more edits to allelic series

@@
 Their findings suggest that different teosinte \emph{tb1} segments contain distinct genetic factors that differentially contribute to tillering in teosinte.
-Furthermore, \citet{StuderDoebley2012} found significant variation in tillering within their control maize lines, concluding that variation in \emph{tb1} alone cannot account for the differences in tillering, and that other genetic factors must also be contributing to differences in tillering in teosinte.
+Furthermore, \citet{StuderDoebley2012} found significant variation in tillering within their control maize lines, concluding that variation in \emph{tb1} alone cannot account for the differences in tillering, and that other genetic factors must also contribute to differences in tillering in teosinte.
 Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).

### 2014-11-21 manuscript.tex
Final edits
Edits to manuscript, supplement, and additional file showing counts of
individuals/population. Ready to submit to PeerJ

@@
 
-\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of \lev{837} maize and teosinte individuals using a co-dominant PCR assay.
+\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 837 maize and teosinte individuals using a co-dominant PCR assay.
 Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
@@
 Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes of the coding domain of \emph{tb1}\citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
+The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes of \emph{tb1}\citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
 Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
@@
 
-\textbf{Sampling and genotyping---}\lev{We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations}; ranging from 1-38 individuals per population, with an average of 11 individuals per population for \emph{parviglumis} and \emph{mexicana} and 2 individuals per landrace accession) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for \lev{837 individuals of the 1,100 screened}.
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
 Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-\lev{Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations}, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).
-\lev{Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.} We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
+We used eight as a cut-off for the minimum number of individuals sampled from a population.
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).

### 2014-11-22 manuscript.tex
edits
Fixed individuals vs chromosomes in genotyping analysis section

@@
 \ref{Fig1Map}).
-Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled eight or more individuals.
-We used eight as a cut-off for the minimum number of individuals sampled from a population.
+Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes.
+We used 16 as a cut-off for the minimum number of chromosomes sampled from a population.
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).

### 2014-11-24 AbstractPeerJ.tex
Edits to manuscript and abstract document
Abstract for PeerJ, edits to manuscript for formatting

--- 2014-11-22 manuscript.tex
@@
-\documentclass[11pt]{article} %indicating type of document, font size
-\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
-\usepackage{hyperref}
-\usepackage[leftcaption]{sidecap} %package for adding side captions
-\sidecaptionvpos{figure}{c} %position side caption
-\usepackage{lineno} %package for adding line numbers
-\usepackage{lipsum}
-\usepackage{indentfirst}
-\usepackage{color}
-\usepackage{multirow}
-\usepackage{booktabs}
-\usepackage[pdftex]{graphicx}  
-\usepackage[labelsep=period]{caption} %package for formatting figure captions, separation between figure number and caption is period per AJB formatting requirements
-\usepackage{natbib} % package for formatting citations
-\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for formatting...journals often have these available but these can also be customized
-\usepackage{authblk}
-\usepackage[table]{xcolor} %for table line color alteration
-\usepackage{subscript}
-\renewcommand{\baselinestretch}{1.5}
-\renewcommand{\refname}{\centering{LITERATURE CITED}} %changing title of references to Literature Cited as stipulated by AJB
-
-\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %introducting new command "\jri" for jeff ross-ibarra to make comments in blue italics
-\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %same for matt hufford in red
-\newcommand{\lev}[1]{\textcolor{magenta}{\emph{\normalsize #1}} } %same for laura vann in green
-
-%----------------------------------------
-%AUTHORS
-%----------------------------------------
-
-\makeatletter
-\renewcommand{\maketitle}{\bgroup\vspace*{1cm}\setlength{\parindent}{0pt}%special formatting for AJB for title page
-\begin{center}
-  \textbf{\@title}
-
-  \@author
-\end{center}\egroup
-}
-\makeatother
-
-\title{\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark}
-
-\author[1]{Laura Vann}%author information
-\author[1,2]{Thomas Kono}
-\author[1,3]{Tanja Pyh\"aj\"arvi}
-\author[1,4,6]{Matthew B. Hufford}
-\author[1,5,6]{Jeffrey Ross-Ibarra}
-\affil[1]{Department of Plant Sciences, University of California Davis}
-\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
-\affil[3]{Department of Biology, University of Oulu}
-\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
-\affil[5]{Center for Population Biology and Genome Center, University of California Davis}
-\affil[6]{Author for Correspondence}
-\date{}
+\documentclass[]{article}
 
@@
 
-\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
-
-{\let\newpage\relax\maketitle}
-
-\begin{linenumbers}
-	\pagewiselinenumbers
-	\renewcommand\linenumberfont{\normalfont\bfseries\small}
-
-\clearpage%inserts page break
-\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted \line(1,0){120}.}
-\vspace*{1cm}
-\subsection*{Acknowledgements}%subsection header
-The authors thank the Department of Plant Sciences at UC Davis for graduate student research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
-Coop for helpful discussion.
-}
-\clearpage
-
-%Code for SAS and command line stuff on Git
-%Phenotyping on Figshare http://dx.doi.org/10.6084/m9.figshare.776926 -- do we need to cite this in the manuscript somewhere?
-
-\begin{flushleft}
-
 \subsection*{Abstract}
 
-\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene is a major QTL controlling branching differences between maize and its wild progenitor, teosinte.
+The \emph{teosinte branched1} (\emph{tb1}) gene is a major QTL controlling branching differences between maize and its wild progenitor, teosinte.
 The insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} is known to enhance the gene's expression, causing reduced tillering in maize.
 Observations of the maize \emph{tb1} allele in teosinte and estimates of an insertion age of the \emph{Hopscotch} that predates domestication led us to investigate its prevalence and potential role in teosinte.
-
-\vspace*{0.2cm}%introduces vertical space between abstract sections
-
-\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 837 maize and teosinte individuals using a co-dominant PCR assay.
-Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
-Phenotypic data were also collected using seed from a single teosinte population where \emph{Hopscotch} was found segregating at high frequency.
-
-\vspace*{0.2cm}
-
-\emph{Key results:} Genotyping results indicate the \emph{Hopscotch} element is found in a number of teosinte populations and linkage disequilibrium near \emph{tb1} does not support recent introgression from maize.
+We assessed the prevalence of the \emph{Hopscotch} element across an Americas-wide sample of 837 maize and teosinte individuals using a co-dominant PCR assay.
+Additionally we calculated population genetic summaries using sequence data from a subset of individuals from four teosinte populations and collected phenotypic data using seed from a single teosinte population where \emph{Hopscotch} was found segregating at high frequency.
+Genotyping results indicate the \emph{Hopscotch} element is found in a number of teosinte populations and linkage disequilibrium near \emph{tb1} does not support recent introgression from maize.
 Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte, but a greenhouse experiment does not detect a strong association between \emph{tb1} and tillering in teosinte.
-\vspace*{0.2cm}
 
@@
 
-\parindent=20pt
-
-\begin{centering}%centers section title
-\section*{INTRODUCTION}
-\end{centering}
-
-Domesticated crops and their wild progenitors provide an excellent system in which to study adaptation and genomic changes associated with human-mediated selection \citep{Ross-Ibarra2007}. 
-Plant domestication usually involves a suite of phenotypic changes such as loss of seed shattering and increased fruit or grain size, which are commonly referred to as the `domestication syndrome' \citep{OlsenWendel2013}, and much of the study of domestication has focused on understanding the genetic variation underlying these traits \citep{OlsenGross2010}.
-Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
-Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
-For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovachMcCouch2008}.
-In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
-But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
-In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1}, and specifically the domesticated haplotype at \emph{tb1}, in natural populations of the wild ancestor of domesticated maize.
-
-Maize (\emph{Zea mays} ssp.
-\emph{mays}) was domesticated from the teosinte \emph{Zea mays} ssp.
-\emph{parviglumis} (hereafter, \emph{parviglumis}) roughly 9,000 B.P.
-in southwest Mexico \citep{Piperno2009, Matsuoka2002}.
-Maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-Additionally, large, naturally-occurring populations of both \emph{parviglumis} and the highland teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
-
-Many morphological changes are associated with maize domestication, and understanding the genetic basis of these changes has been a focus of maize research for a number of years \citep{Doebley2004}.
-One of the most dramatic changes is found in plant architecture: domesticated maize is characterized by a central stalk with few tillers and lateral branches terminating in a female inflorescence, while teosinte is highly tillered and bears tassels (male inflorescences) at the end of its lateral branches.
-The \emph{teosinte branched1} (\emph{tb1}) gene, a repressor of organ growth, was identified as a major QTL involved in  branching \citep{DoebleyStecGustus1995} and tillering \citep{DoebleyStec1991} differences between maize and teosinte.
-A 4.9 kb retrotransposon (\emph{Hopscotch}) insertion into the upstream control region of \emph{tb1} in maize acts to enhance expression of \emph{tb1}, thus repressing lateral organ growth \citep{Doebley1997, Studer2011}.
-Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes of \emph{tb1}\citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
-Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
-
-In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.
-An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome, a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}.
-The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading.
-\citet{LukensDoebley1999} introgressed the teosinte \emph{tb1} allele into a maize inbred background and noted that under low density conditions plants were highly tillered but that under high density, plants showed significantly reduced tillers and grew taller.
-Based on these results we hypothesize that the \emph{Hopscotch} (\emph{i.e.}, the domesticated allele) at \emph{tb1} may play a role in the ecology of teosinte, especially in high-density populations.
-In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}.
-We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} combined with a larger SNP dataset in a subset of teosinte populations to explore patterns of genetic variation at this locus.
-Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in samples from a natural population of \emph{parviglumis}.
-
-\begin{centering}
-\section*{MATERIALS AND METHODS}
-\end{centering}
-
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
-DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
-We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb).
-Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
-Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
-Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
-We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
-PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.
-
-\textbf{Genotyping analysis---}To calculate differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) we used HierFstat \citep{Goudet2005}.
-These analyses only included populations ($n = 32$) in which eight or more chromosomes were sampled.
-To test the hypothesis that the \emph{Hopscotch} insertion may be adaptive under certain environmental conditions, we looked for significant associations between \emph{Hopscotch} frequency and environmental variables using the software BayEnv \citep{Coop2010}.
-BayEnv creates a covariance matrix of relatedness between populations and then tests a null model that allele frequencies in populations are determined by the covariance matrix of relatedness alone against the alternative model that allele frequencies are determined by a combination of the covariance matrix and an environmental variable, producing a posterior probability (\emph{i.e.,} Bayes Factor; \citealt{Coop2010}).
-We used teosinte (ssp.
-\emph{parviglumis} and ssp.
-\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
-SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
-Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
-Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
-Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.
-
-\textbf{Sequencing---}In addition to genotyping, we chose a subset of \emph{parviglumis} individuals for sequencing.
-We chose twelve individuals from each of four populations from Jalisco state, Mexico (San Lorenzo, La Mesa, Ejutla A, and Ejutla B).
-For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (Region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} (66,169bp upstream from the start of the \emph{tb1} ORF; Region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Appendix 1, See Supplemental Materials with the online version of this article).
-We used Taq polymerase (New England Biolabs Inc., Ipswich, Massachusetts, USA) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
-Annealing temperatures for Region 1 and Region 2 were 59.7$^{\circ}$C and 58.8$^{\circ}$C, respectively.
-To clean excess primer and dNTPs we added two units of Exonuclease1 and 2.5 units of Antarctic Phosphatase to 8.0 $\mu$L of amplification product.
-This mix was placed on a thermal cycler with the following program: 37$^{\circ}$C for 30 min, 80$^{\circ}$C for 15 min, and a final cool-down step to 4$^{\circ}$C.
-
-We cloned cleaned fragments into a TOPO-TA vector (Life Technologies, Grand Island, New York, USA) using OneShot TOP10 chemically competent \emph{E.
-coli} cells, with an extended ligation time of 30 min for a complex target fragment.
-We plated cells on LB agar plates containing kanamycin, and screened colonies using vector primers M13 Forward and M13 Reverse under the following conditions: 96$^{\circ}$C for 5 min; then 35 cycles at 96$^{\circ}$C for 30 s, 53$^{\circ}$C for 30 s, 72$^{\circ}$C for 2 min; and a final extension at 72$^{\circ}$C for 4 min.
-We visualized amplification products for incorporation of our insert on a 1\% agarose TAE gel.
-
-Amplification products with successful incorporation of our insert were cleaned using Exonuclease 1 and Antarctic Phosphatase following the procedures detailed above, and sequenced with vector primers M13 Forward and M13 Reverse using Sanger sequencing at the College of Agriculture and Environmental Sciences (CAES) sequencing center at UC Davis.
-We aligned and trimmed primer sequences from resulting sequences using the software Geneious \citep{Kearse2012}.
-Following alignment, we verified singleton SNPs by sequencing an additional one to four colonies from each clone.
-If the singleton was not present in these additional sequences it was considered an amplification or cloning error, and we replaced the base with the base of the additional sequences.
-If the singleton appeared in at least one of the additional sequences we considered it a real variant and kept it for further analyses.
-
-\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
-For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.
-
-\textbf{Introgression analysis---}In order to assess patterns of linkage disequilibrium (LD) around the \emph{Hopscotch} element in the context of chromosomal patterns of LD we used Tassel \citep{Bradbury2007} and calculated LD between SNPs across chromosome 1 using previously published data from twelve plants each of the Ejutla A (EjuA), Ejutla B (EjuB), San Lorenzo (SLO), and La Mesa (MSA) populations \citep{Pyhajarvi2013}.
-We chose these populations because we had both genotyping data for the \emph{Hopscotch} as well as chromosome-wide SNP data for chromosome 1.
-For each population we filtered the initial set of 5,897 SNPs on chromosome 1 to accept only SNPs with a minor allele frequency of at least 0.1, resulting in 1,671, 3,023, 3,122, and 2,167 SNPs for SLO, EjuB, EjuA, and MSA, respectively.
-We then used Tassel \citep{Bradbury2007} to calculate linkage disequilibrium ($r^{2}$) across chromosome 1 for each population.
-
-We examined evidence of introgression on chromosome 1 in these same four populations (EjuA, EjuB, MSA, SLO) using STRUCTURE \citep{Falush2003} and phased data from \citet{Pyhajarvi2013}, combined with the corresponding SNP data from a diverse panel of 282 maize lines \citep{Cook2012}.
-SNPs were anchored in a modified version of the IBM genetic map \citep{Gerke2013}.
-Since STRUCTURE does not account for LD due to physical linkage we created haplotype blocks using a custom Perl script from \citet[][code available at \url{http://dx.doi.org/10.6084/m9.figshare.1165577}]{Hufford2013}.
-In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
-We ran STRUCTURE at K=2 under the linkage model, with the assumption being that individuals fall into either a maize or teosinte cluster, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
-
-\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo; \citealt{Hufford2010}) where the \emph{Hopscotch} insertion is segregating at highest frequency (0.44) in our initial genotyping sample set.
-In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites within the population where genotyped individuals were homozygous or heterozygous for the insertion.
-We chose between 10-13 seeds from each of 23 sampling sites.
-We treated seeds with Captan fungicide (Southern Agricultural Insecticides Inc., Palmetto, Florida, USA) and germinated them in petri dishes with filter paper.
-Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches.
-Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days.
-
-Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}.
-Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60.
-On day 65 we measured culm diameter between the third and fourth nodes of each plant.
-Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol. 
-We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above.
-
-Tillering index data for each genotypic class did not meet the criteria for a repeated measures ANOVA, so we transformed the data with a Box-Cox transformation ($\lambda=0$) in the Car Package for R \citep{FoxWeisberg2011} to improve the normality and homogeneity of variance among genotype classes.
-We analyzed relationships between genotype and tillering index and tiller number using a repeated measures ANOVA through a general linear model function implemented in SAS v.9.3 (SAS Institute Inc., Cary, NC, USA).
-Additionally, in order to compare any association between \emph{Hopscotch} genotype and tillering and associations at other presumably unrelated traits, we performed an ANOVA between culm diameter and genotype using the same general linear model in SAS.
-Culm diameter is not believed to be correlated with tillering index or variation at \emph{tb1} and is used as our independent trait for phenotyping analyses.
-SAS code used for analysis is available at \url{ http://dx.doi.org/10.6084/m9.figshare.1166630}.
-
-\begin{centering}
-\section*{RESULTS}
-\end{centering}
-
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
-Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).
-Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes.
-We used 16 as a cut-off for the minimum number of chromosomes sampled from a population.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
-Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).
-
-\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
-After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
-For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}, Appendix 2, See Supplemental Materials with the online version of this article).
-For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
-Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Appendix 5, See Supplemental Materials with the online version of this article); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
-
-\textbf{Evidence of introgression---}The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
-Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
-To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one and specifically in the \emph{tb1} region.
-If the \emph{Hopscotch} is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Figure~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
-In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).
-
-The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
-However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) \citep{Pyhajarvi2013} and the maize 282 diversity panel \citep{Cook2012}.
-The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants which would arise as the result of recent admixture between populations.
-If the \emph{Hopscotch} insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the ``maize" cluster in our STRUCTURE runs, not the ``teosinte" cluster.
-In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig.
-\ref{Fig3Structure}).
-
-\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
-We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.
-We also measured culm diameter, a phenotype that differs between maize and teosinte but has not been shown to be affected by the \emph{Hopscotch} insertion \citep{Briggs2007}.
-Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data.
-If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype.
-Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
-Our plantings produced 82 homozygotes for the \emph{Hopscotch} insertion at \emph{tb1}, 104 heterozygotes, and 20 homozygotes lacking the insertion; these numbers do not deviate from expectations of Hardy-Weinberg equilibrium.
-After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig.
-\ref{Fig4Boxplots}), tiller number, or culm diameter.
-Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
-
-\begin{centering}
-\section*{DISCUSSION}
-\end{centering}
-
-Adaptation occurs due to selection on standing variation or \emph{de novo} mutations.
-Adaptation from standing variation has been well-described in a number of systems; for example, selection for lactose tolerance in humans \citep{Plantinga2012, Tishkoff2007}, variation at the \emph{Eda} locus in three-spined stickleback \citep{Kitano2008, Colosimo2005}, and pupal diapause in the Apple Maggot fly \citep{Feder2003}.
-Although the adaptive role of standing variation has been described in many systems, its importance in domestication is not as well studied.
-
-In maize, alleles at domestication loci (\emph{RAMOSA1}, \citealt{SigmonVollbrecht2010}; \emph{barren stalk1}, \citealt{Gallavotti2004}; and \emph{grassy tillers1}, \citealt{Whipple2011}) are thought to have been selected from standing variation, suggesting that diversity already present in teosinte may have played an important role in maize domestication.
-The \emph{teosinte branched1} gene is one of the best characterized domestication loci, and, while previous studies have suggested that differences in plant architecture between maize and teosinte are a result of selection on standing variation at this locus \citep{Clark2006, Studer2011}, much remains to be discovered regarding natural variation at this locus and its ecological role in teosinte.
-
-\citet{Studer2011} genotyped 90 accessions of teosinte (inbred and outbred), providing the first evidence that the \emph{Hopscotch} insertion is segregating in teosinte. 
-Given that the \emph{Hopscotch} insertion has been estimated to predate the domestication of maize, it is not surprising that it can be found segregating in populations of teosinte.
-However, by widely sampling across teosinte populations our study provides greater insight into the distribution and prevalence of the \emph{Hopscotch} in teosinte.
-While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
-The \emph{Hopscotch} allele is more prevalent in \emph{parviglumis} than in \emph{mexicana} in our sample (Appendix 1, See Supplemental Materials with the online version of this article), suggesting a different history of the allele amongst teosinte subspecies.
-Moreover, many of our \emph{parviglumis} populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, and further distinguish this region from the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
-Potential explanations for the high frequency of the \emph{Hopscotch} element in \emph{parviglumis} from the Jalisco cluster include gene flow from maize, genetic drift, and natural selection.
-
-While gene flow from crops into their wild relatives is well-known, \citep{Ellstrand1999, Zhang2009, Thurber2010, Baack2008, Hubner2012, Wilkes1977, VanHeerwaarden2011, Barrett1983}, our results do not suggest introgression from maize at the \emph{tb1} locus, and are more consistent with \citet{Hufford2013} who found resistance to introgression from maize into \emph{mexicana} around domestication loci.
-Clustering in our NJ trees does not reflect the pattern expected if maize alleles at the \emph{tb1} locus had introgressed into populations of teosinte.
-Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig.
-\ref{Fig3Structure}, Table \ref{Table4Q}).
-Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in the Jalisco cluster of our \emph{parviglumis} populations.
-
-Although recent introgression seems unlikely, we cannot rule out ancient introgression as an explanation for the presence of the \emph{Hopscotch} in these populations.
-If the \emph{Hopscotch} allele was introgressed in the distant past, recombination may have broken up LD, a process that would be consistent with our data.
-We find this scenario less plausible, however, as there is no reason why gene flow should have been high in the past but absent in present-day sympatric populations.
-In fact, early generation maize-teosinte hybrids are common in these populations today (MB Hufford, pers.
-observation), and genetic data support ongoing gene flow between domesticated maize and both \emph{mexicana} and \emph{parviglumis} in a number of sympatric populations \citep{Hufford2013, Ellstrand2007, VanHeerwaarden2011, Warburton2011}.
-
-Remaining explanations for differential frequencies of the \emph{Hopscotch} among teosinte populations include both genetic drift and natural selection.
-Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in this population in the 5' UTR (Region 1) coupled with more positive values of Tajima's D are consistent with these earlier findings.
-Such population bottlenecks can exaggerate the effects of genetic drift through which the \emph{Hopscotch} allele may have risen to high frequency entirely by chance.
-A bottleneck in San Lorenzo, however, does not explain the high frequency of the \emph{Hopscotch} in multiple populations in the Jalisco cluster.
-Moreover, available information on diversity and population structure among Jaliscan populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of recent colonization or other demographic events that would predict a high frequency of the allele across populations.
-Finally, diversity values in the 5' UTR of \emph{tb1} are suggestive of natural selection acting upon the gene in populations of \emph{parviglumis}.
-Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative across populations (Table \ref{Table2Diversity}).
-In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}.
-Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in Jaliscan populations of \emph{parviglumis}.
-
-Significant effects of the \emph{Hopscotch} insertion on lateral branch length, number of cupules, and tillering index in domesticated maize have been well documented \citep{Studer2011}.
-\citet{Weber2007} described significant phenotypic associations between markers in and around \emph{tb1} and lateral branch length and female ear length in a sample from 74 natural populations of \emph{parviglumis} \citep{Weber2007}; however, these data did not include markers from the \emph{Hopscotch} region 66kb upstream of \emph{tb1}.
-Our study is the first to explicitly examine the phenotypic effects of the \emph{Hopscotch} insertion across a wide collection of individuals sampled from natural populations of teosinte.
-We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
-One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
-Consistent with this interpretation, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.
-
-A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.
-\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated their phenotypic effects.
-Their findings suggest that different teosinte \emph{tb1} segments contain distinct genetic factors that differentially contribute to tillering in teosinte.
-Furthermore, \citet{StuderDoebley2012} found significant variation in tillering within their control maize lines, concluding that variation in \emph{tb1} alone cannot account for the differences in tillering, and that other genetic factors must also contribute to differences in tillering in teosinte.
-Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).
-Many of these loci (\emph{grassy tillers}, \emph{gt1}; \emph{tassel-replaces-upper-ears1}, \emph{tru1}; \emph{terminal ear1}, \emph{te1}) have been shown to interact with \emph{tb1} \citep{Whipple2011, Li2012},  and both \emph{tru1} and \emph{te1} affect the same phenotypic traits as \emph{tb1} \citep{DoebleyStecGustus1995}.
-\emph{tru1}, for example, has been shown to act either epistatically or downstream of \emph{tb1}, affecting both branching architecture (decreased apical dominance) and tassel phenotypes (shortened tassel and shank length and reduced tassel number; \citealt{Li2012}).
-Variation in these additional loci may have affected tillering in our collections and contributed to the lack of correlation we see between \emph{Hopscotch} genotype and tillering.
-
-In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is widespread in populations of \emph{parviglumis} and \emph{mexicana} and occasionally at high allele frequencies. 
-Analysis of linkage using SNPs from across chromosome 1 does not suggest that the \emph{Hopscotch} allele is present in these populations due to recent introgression, and it seems unlikely that the insertion would have drifted to high frequency in multiple populations.
-We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}.
-Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus may play an ecological role in teosinte.
-
-In contrast to domesticated maize, the \emph{Hopscotch} insertion does not appear to have a large effect on tillering in a diverse sample of \emph{parviglumis} from a natural population and the phenotypic consequences of variation at \emph{tb1} thus remain unclear.
-Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g.
-\emph{gt1}, \emph{tru1}, and \emph{te1}).
-These data, in conjunction with more exhaustive phenotyping, should help to further clarify the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte.
-
-\clearpage
-
-\bibliography{refs}%print bibliography
-
-\clearpage
-% Table 1
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}%alternating row colors in table
-  \centering
-  \caption{Pairwise F\textsubscript{ST} values from sequence and \emph{Hopscotch} genotyping data}%table caption
-    \begin{tabular}{cccc}\\\toprule%number of columns and line across top of table
-    \textbf{Comparison} & \textbf{Region 1} & \textbf{Region 2} & \textbf{\emph{Hopscotch}} \\\midrule
-    EjuA \& EjuB & 0     & 0     & 0 \\
-    EjuA \& MSA & 0.326 & 0.328 & 0.186 \\
-    EjuA \& SLO & 0.416 & 0.258 & 0.280 \\
-    EjuB \& MSA & 0.397 & 0.365 & 0.188 \\
-    EjuB \& SLO & 0.512  & 0.290 & 0.280 \\
-    MSA \& SLO & 0.007  & 0 & 0.016 \\\bottomrule
-    \end{tabular}
-  \label{Table1Fst}%indicates link between table and mention in text
-\end{table}
-
-% Table 2 Diversity Statistics
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
-    \begin{tabular}{ccccc}\\\toprule
-    \textbf{Population} & \# \textbf{Haplotypes} & \textbf{Hap. Diversity} & \textbf{$\hat\theta_\pi$}    & \textbf{Tajima's D} \\\midrule
-    \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
-    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
-    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
-    \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\
-    \multicolumn{5}{c}{\textit{Region 2 (66kb upstream)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.894 & 0.018 & 0.623 \\
-    \multicolumn{1}{c}{EJUB} & 8     & 0.894 & 0.016 & 0.295 \\
-    \multicolumn{1}{c}{MSA} & 3     & 0.682 & 0.011 & -0.222 \\
-    \multicolumn{1}{c}{SLO} & 4     & 0.742 & 0.014 & 0.932 \\\bottomrule
-    \end{tabular}
-  \label{Table2Diversity}
-\end{table}
-
-% Table 3 $r^{2}$ values
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{mean $r^{2}$ values between SNPs on chromosome 1, in the broad \emph{tb1} region, within the 5' UTR of \emph{tb1} (Region 1), and 66kb upstream of \emph{tb1} (Region 2).}
-    \begin{tabular}{ccccc}\\\toprule
-    \textbf{Population} & \textbf{Chr. 1} & \textbf{\emph{tb1} region} & \textbf{Region 1} & \textbf{Region 2} \\\midrule
-    Ejutla A & 0.095 & 0.050 & 0.747 & 0.215 \\
-    Ejutla B & 0.069 & 0.051 & 0.660 & 0.186 \\
-    La Mesa & 0.070 & 0.053 & 0.914 & 0.766 \\
-    San Lorenzo & 0.101 & 0.067 & 0.912 & 0.636 \\
-    \bottomrule
-    \end{tabular}
-  \label{Table3R2}
-\end{table}
-
- % Table 4 Assignment to maize and teosinte in tb1 region and chr 1 
-\begin{table}[htbp]
-\rowcolors{2}{gray!25}{white}
-  \centering
-  \caption{Assignments to maize and teosinte in the \emph{tb1} and chromosome 1 regions from STRUCTURE}
-    \begin{tabular}{ccccc} \
-          & \multicolumn{2}{c}{\textbf{\emph{tb1} region}} & \multicolumn{2}{c}{\textbf{Chr 1}} \\\toprule
-    \bf{Population} & \bf{Maize} & \bf{Teosinte} & \bf{Maize} & \bf{Teosinte} \\\midrule
-    Ejutla A & 0.022 & 0.978 & 0.203 & 0.797 \\
-    Ejutla B & 0.019 & 0.981 & 0.187 & 0.813 \\
-    La Mesa & 0.012 & 0.988 & 0.193 & 0.807 \\
-    San Lorenzo & 0.016 & 0.984 & 0.205 & 0.795 \\\bottomrule
-    \end{tabular}
-  \label{Table4Q}
-\end{table}
-
-\clearpage
-
-%Figure 1 Map 
-%--------------------------------------------
-\begin{figure*}[!t]
-\begin{center}
-   \includegraphics[width=0.6\textwidth]{Fig1MapBalsas.pdf}
-   \end{center}
-    \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
-Size of circles reflects number of alleles sampled.
-The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.}
-\label{Fig1Map}
-\end{figure*}
-%--------------------------------------------
-
-% Figure 2. LD plots
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=140mm]{Fig2LDPlots.pdf}
-   \end{center}
-    \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
-The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
-A) Ejutla A; B) Ejutla B; C) La Mesa; D) San Lorenzo.
-The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.}
-\label{Fig2LD}
-\end{figure*}
-%-------------------------------------------------------------------
-
-% Figure 3 STRUCTURE Plot 4 Pops
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=90mm]{Fig3Structure.pdf}
-     \end{center}
-    \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF.}
-\label{Fig3Structure}
-\end{figure*}
-%-------------------------------------------------------------------
-
-% Figure 4 Phenotyping BoxPlots
-%-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=90mm]{Figure4BoxplotPheno1.pdf}
-   \end{center}
-    \caption{Box-plots showing tillering index in greenhouse grow-outs for phenotyping.
-White indicates individuals homozygous for the \emph{Hopscotch}, light grey represents heterozygotes, and dark grey represents homozygotes for the teosinte (No \emph{Hopscotch}) allele.
-Within boxes, dark black lines represent the median, and the edges of the boxes are the first and third quartiles.
-Outliers are displayed as dots, the maximum value excluding outliers is shown with the top whisker, while the minimum excluding outliers is shown with the bottom whisker.}
-\label{Fig4Boxplots}
-\end{figure*}
-%-------------------------------------------------------------------
-
-\end{flushleft}
-
-\end{linenumbers}
-
 \end{document}

### 2014-11-24 AbstractPeerJ.tex
abstract edits
edits to abstract for peerj

@@
 Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte, but a greenhouse experiment does not detect a strong association between \emph{tb1} and tillering in teosinte.
-
-\emph{Conclusions:} Our findings suggest the role of \emph{Hopscotch} differs between maize and teosinte.
-Future work should assess \emph{tb1} expression levels in teosinte with and without the \emph{Hopscotch} and more comprehensively phenotype teosinte to assess the ecological significance of the \emph{Hopscotch} insertion and, more broadly, the \emph{tb1} locus in teosinte.
-
-\vspace*{0.2cm}
+Our findings suggest the role of \emph{Hopscotch} differs between maize and teosinte.
+Future work should assess \emph{tb1} expression levels in teosinte with and without the \emph{Hopscotch} and more comprehensively phenotype teosinte to assess the ecological significance of the \emph{Hopscotch} insertion and, more broadly, the \emph{tb1} locus in teosinte.\\
 

### 2014-12-08 manuscript.tex

@@
 \makeatletter
-\renewcommand{\maketitle}{\bgroup\vspace*{1cm}\setlength{\parindent}{0pt}%special formatting for AJB for title page
+\renewcommand{\maketitle}{\bgroup\vspace*{1cm}%\setlength{\parindent}{0pt}%special formatting for AJB for title page
 \begin{center}
@@
 
-\title{\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark}
+%\title{\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark}
 
@@
 \author[1,5,6]{Jeffrey Ross-Ibarra}
-\affil[1]{Department of Plant Sciences, University of California Davis}
-\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
-\affil[3]{Department of Biology, University of Oulu}
-\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
-\affil[5]{Center for Population Biology and Genome Center, University of California Davis}
+\affil[1]{Department of Plant Sciences, University of California Davis, Davis, CA, USA}
+\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities, Minneapolis, MN, USA}
+\affil[3]{Department of Biology, University of Oulu, Oulu, Finland}
+\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University, Ames, Iowa, USA}
+\affil[5]{Center for Population Biology and Genome Center, University of California Davis, Davis, CA, USA}
 \affil[6]{Author for Correspondence}
@@
 \begin{flushleft}
-
-\subsection*{Abstract}
-
-\emph{Premise of the study:} The \emph{teosinte branched1} (\emph{tb1}) gene is a major QTL controlling branching differences between maize and its wild progenitor, teosinte.
-The insertion of a transposable element (\emph{Hopscotch}) upstream of \emph{tb1} is known to enhance the gene's expression, causing reduced tillering in maize.
-Observations of the maize \emph{tb1} allele in teosinte and estimates of an insertion age of the \emph{Hopscotch} that predates domestication led us to investigate its prevalence and potential role in teosinte.
-
-\vspace*{0.2cm}%introduces vertical space between abstract sections
-
-\emph{Methods:} Prevalence of the \emph{Hopscotch} element was assessed across an Americas-wide sample of 837 maize and teosinte individuals using a co-dominant PCR assay.
-Population genetic summaries were calculated for a subset of individuals from four teosinte populations in central Mexico.
-Phenotypic data were also collected using seed from a single teosinte population where \emph{Hopscotch} was found segregating at high frequency.
-
-\vspace*{0.2cm}
-
-\emph{Key results:} Genotyping results indicate the \emph{Hopscotch} element is found in a number of teosinte populations and linkage disequilibrium near \emph{tb1} does not support recent introgression from maize.
-Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte, but a greenhouse experiment does not detect a strong association between \emph{tb1} and tillering in teosinte.
-\vspace*{0.2cm}
-
-\emph{Conclusions:} Our findings suggest the role of \emph{Hopscotch} differs between maize and teosinte.
-Future work should assess \emph{tb1} expression levels in teosinte with and without the \emph{Hopscotch} and more comprehensively phenotype teosinte to assess the ecological significance of the \emph{Hopscotch} insertion and, more broadly, the \emph{tb1} locus in teosinte.
-
-\vspace*{0.2cm}
-
-\textbf{Key words:} domestication; maize; teosinte; \emph{teosinte branched1}; transposable element
-
-\clearpage
 

### 2014-12-08 manuscript.tex
Edits for PeerJ
Formatting tables and figures, spacing

@@
 \clearpage%inserts page break
-\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted \line(1,0){120}.}
+%\footnotetext{\flushleft{Manuscript received \line(1,0){120};   revision accepted \line(1,0){120}.}
 \vspace*{1cm}
-\subsection*{Acknowledgements}%subsection header
+%\subsection*{Acknowledgements}%subsection header
 The authors thank the Department of Plant Sciences at UC Davis for graduate student research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
@@
 
+\clearpage
+
 % Table 3 $r^{2}$ values
@@
 \end{table}
+
+\clearpage
 
@@
 %--------------------------------------------
+\clearpage
 
@@
 %-------------------------------------------------------------------
+\clearpage
 

### 2014-12-09 manuscript.tex
Edits
reformatting
edits throughout for PeerJ

@@
 
-\author[1]{Laura Vann}%author information
+\author[1]{Authors: Laura Vann}%author information
 \author[1,2]{Thomas Kono}
@@
 \author[1,4,6]{Matthew B. Hufford}
-\author[1,5,6]{Jeffrey Ross-Ibarra}
+\author[1,5,7]{Jeffrey Ross-Ibarra}
 \affil[1]{Department of Plant Sciences, University of California Davis, Davis, CA, USA}
@@
 \affil[5]{Center for Population Biology and Genome Center, University of California Davis, Davis, CA, USA}
-\affil[6]{Author for Correspondence}
+\affil[6]{Corresponding Author: Matthew B. Hufford; 339A Bessey Hall, Iowa State University, Ames, IA, USA; phone: 1-515-294-8511; email: mhufford@iastate.edu}
+\affil[7]{Corresponding Author: Jeffrey Ross-Ibarra; 262 Robbins Hall, Mail Stop 4, University of California, Davis, CA, USA; phone: 1-530-752-1152; email: rossibarra@ucdavis.edu}
 \date{}
@@
 
-\flushleft{\large{Vann \emph{et al.}: Variation in teosinte at the \emph{tb1} locus}}
-
 {\let\newpage\relax\maketitle}
@@
 \begin{linenumbers}
-	\pagewiselinenumbers
 	\renewcommand\linenumberfont{\normalfont\bfseries\small}
@@
 \vspace*{1cm}
-%\subsection*{Acknowledgements}%subsection header
-The authors thank the Department of Plant Sciences at UC Davis for graduate student research funding to LEV and for research funds supporting the project, UC Mexus for a postdoctoral scholar grant to MBH and JR-I, and G.
-Coop for helpful discussion.
-}
-\clearpage
 
@@
 Because most domesticates show reduced genetic diversity relative to their wild counterparts, effort has been made to identify agronomically useful variation in crop wild relatives \citep{flint2009wide}.
-Often, after identification, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
+In some instances, the alleles conferring these beneficial traits are bred into domesticates for crop improvement.
 For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice \citep{KovachMcCouch2008}.
 In addition to researching the role of wild alleles in domesticates, researchers have also investigated the role of variation in domesticated taxa in the evolution of feral and weedy populations \citep{ellstrand2010crops}.
-But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, we know almost nothing about the ecological role of this variation in natural populations.
-In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1}, and specifically the domesticated haplotype at \emph{tb1}, in natural populations of the wild ancestor of domesticated maize.
+But even though domesticated alleles are often found segregating in wild relatives \citep{Gallavotti2004, SigmonVollbrecht2010}, little is known about the ecological role of this variation in natural populations.
+In this paper we present an ecological genetic analysis of the domestication locus \emph{tb1}---specifically the domesticated haplotype at \emph{tb1}---in natural populations of the wild ancestor of domesticated maize.
 
@@
 Dating of the \emph{Hopscotch} retrotransposon suggests that its insertion predates the domestication of maize, leading to the hypothesis that it was segregating as standing variation in populations of teosinte and increased to high frequency in maize due to selection during domestication \citep{Studer2011}.
-The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes of \emph{tb1}\citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion in natural populations of teosinte.
-Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis}, \emph{mexicana}, and landrace maize; however little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
+The effects of the \emph{Hopscotch} insertion have been studied in maize \citep{Studer2011}, and analysis of teosinte alleles at \emph{tb1} has identified functionally distinct allelic classes of \emph{tb1}\citep{StuderDoebley2012}, but little is known about the role of \emph{tb1} or the \emph{Hopscotch} insertion at this locus in natural populations of teosinte.
+Previous studies have confirmed the presence of the \emph{Hopscotch} in samples of \emph{parviglumis} and landrace maize \citep{Studer2011}; however, little is known about the frequency with which the \emph{Hopscotch} is segregating in natural populations.
 
@@
 \begin{centering}
-\section*{MATERIALS AND METHODS}
+\section*{MATERIALS \& METHODS}
 \end{centering}
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
@@
 In maize, LD decays over an average distance of 5500bp \citep{Chia2012}; because LD decay is even more rapid in teosinte \citep{Chia2012} we used a conservative haplotype block size of 5kb.
-We ran STRUCTURE at K=2 under the linkage model, with the assumption being that individuals fall into either a maize or teosinte cluster, performing 3 replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
+We ran STRUCTURE at K=2 under the linkage model, with the assumption being that individuals fall into either a maize or teosinte cluster, performing three replicates with an MCMC burn-in of 10,000 steps and 50,000 steps post burn-in.
 
@@
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes.
-We used 16 as a cut-off for the minimum number of chromosomes sampled from a population.
 We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
@@
 
-\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
+\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
 After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
@@
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
-To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome one and specifically in the \emph{tb1} region.
-If the \emph{Hopscotch} is found in these populations due to recent introgression we would expect to find large blocks of linked markers near this element.
+To investigate this possibility we examined overall patterns of linkage disequilibrium across chromosome 1 and specifically in the \emph{tb1} region.
+If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element.
 We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Figure~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
@@
 Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data.
-If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype.
+If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion, the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype.
 Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
@@
 Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in Jaliscan populations of \emph{parviglumis}.
+Finally, while these results are consistent with selection at the \emph{tb1} locus in teosinte, they do not confirm selection specifically on the \emph{Hopscotch} insertion at this locus.
 
@@
 
+\begin{centering}
+\section*{CONCLUSIONS}
+\end{centering}
+
 In conclusion, our findings demonstrate that the \emph{Hopscotch} allele is widespread in populations of \emph{parviglumis} and \emph{mexicana} and occasionally at high allele frequencies. 
@@
 We do, however, find preliminary evidence of selection on the \emph{tb1} locus in \emph{parviglumis}.
-Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus may play an ecological role in teosinte.
+Coupled with our observation of high frequency of the \emph{Hopscotch} insertion in a number of populations, this suggests that the locus---and potentially the domestication allele at this locus---may play an ecological role in teosinte.
 
@@
 
+\subsection*{Acknowledgements}%subsection header
+The authors thank Graham Coop for helpful discussion.
+
+\clearpage
 \bibliography{refs}%print bibliography
@@
 \begin{center}
-   \includegraphics[width=0.6\textwidth]{Fig1MapBalsas.pdf}
+   \includegraphics[width=0.8\textwidth]{Fig1MapBalsas.png}
    \end{center}
     \caption{Map showing the frequency of the \emph{Hopscotch} allele in populations of \emph{parviglumis} where we sampled more than 6 individuals.
-Size of circles reflects number of alleles sampled.
+Size of circles reflects number of individuals sampled.
 The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.}
@@
   \begin{center}
-   \includegraphics[width=140mm]{Fig2LDPlots.pdf}
+   \includegraphics[width=180mm]{Fig2LDPlots.png}
    \end{center}
     \caption{Linkage disequilibrium for SNPs in Mb 261-268 on chromosome 1.
-The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green represents the \emph{tb1} ORF.
+The yellow rectangle indicates the location of the \emph{Hopscotch} insertion and the green rectangle represents the \emph{tb1} ORF.
 A) Ejutla A; B) Ejutla B; C) La Mesa; D) San Lorenzo.
@@
   \begin{center}
-   \includegraphics[width=90mm]{Fig3Structure.pdf}
+   \includegraphics[width=160mm]{Fig3Structure.png}
      \end{center}
@@
   \begin{center}
-   \includegraphics[width=90mm]{Figure4BoxplotPheno1.pdf}
+   \includegraphics[width=160mm]{Figure4BoxplotPheno1.png}
    \end{center}

### 2014-12-09 Supplement_mbh.tex
Final edits
Edits to manuscript, supplement, and additional file showing counts of
individuals/population. Ready to submit to PeerJ
PeerJ Edits
Edits to supplement, figures
formatting supplement

@@
-%Online Supplemental Materials
-%Authors may wish to augment their manuscripts with online supplemental materials (e.g., large data sets, three-dimensional reconstructions, simulations, real-time movies, color photographs).
-Upload these appendices as separate files with the initial manuscript submission.
-Include a header on each file using this format: Smith et al.?American Journal of Botany 99(#): ###-###.
-2012.
-?
-Data Supplement S1 ?
-Page 1?.
-Name online supplements Appendix S1, Appendix S2, etc, in the order in which they appear in the text, regardless of whether they are tables, figures, text, other media, or a combination thereof.
-In the manuscript, after the mention of an online appendix, include the following: ?(see Supplemental Data with the online version of this article)?.
-
-%Note that if authors wish to submit long DNA sequence appendices as supplemental material, they should select the "DNA sequences (online-only supplemental)" option on Editorial Manager.
-This ensures that lengthy appendices are not built into the reviewers' PDF, but are still accessible to the reviewers.
 \documentclass[11pt]{article}
@@
 
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S1 -- Page \thepage}}
-
-\centering{Appendix S1: Accessions of \emph{Zea mays} ssp.
+\centering{Table S1: Accessions of \emph{Zea mays} ssp.
 \emph{mexicana} (RIMME) and \emph{Zea mays} ssp.
@@
 
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S2 -- Page \thepage}}
 % Appendix S2. Mays sampling
 
-\centering{Appendix S2: \emph{Hopscotch} frequency in sampled \emph{Zea mays} ssp. \emph{mays} (RIMMA).}
+\centering{Table S2: \emph{Hopscotch} frequency in sampled \emph{Zea mays} ssp. \emph{mays} (RIMMA).}
 
@@
 \clearpage
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S3 -- Page \thepage}}
+
 % Appendix S3. Cartoon representation of tb1 locus
@@
   \begin{center}
-   \includegraphics[width=180mm]{FigS1LocusCartoon.pdf}
-   \flushleft{Appendix S3: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+   \includegraphics[width=180mm]{FigS1LocusCartoon.png}
+   \flushleft{Figure S1: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
 Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
@@
 \clearpage
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S4 -- Page \thepage}}
 
@@
   \begin{center}
-   \includegraphics[width=100mm]{FigS2Gel.png}
-   \flushleft{Appendix S4: Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+   \includegraphics[width=100mm]{FigS2GelPeerJ.png}
+   \flushleft{Figure S2: Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
   \end{center}
@@
 \clearpage
-\pagenumbering{arabic}
-\chead{\small{Vann et al.---American Journal of Botany (\#\#)(\#): \#\#\#-\#\#\#. 2014. -- Appendix S5 -- Page \thepage}}
 
@@
   \begin{center}
-   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
-    \flushleft{Appendix S5: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}.
+   \includegraphics[width=150mm]{FigS3NJtrees.png}
+    \flushleft{Figure S3: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}.
 Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).

### 2014-12-09 manuscript.tex
formatting supplement

@@
 \renewcommand{\baselinestretch}{1.5}
-\renewcommand{\refname}{\centering{LITERATURE CITED}} %changing title of references to Literature Cited as stipulated by AJB
+\renewcommand{\refname}{\centering{REFERENCES}} %changing title of references to Literature Cited as stipulated by AJB
 
@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
@@
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
-HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article).
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Appendix 2, See Supplemental Materials with the online version of this article).
+HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.
+S1 and Fig.
+S2, See Supplemental Materials with the online version of this article).
+Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Table S2, See Supplemental Materials with the online version of this article).
 Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
@@
 For amplification and sequencing, we selected two regions approximately 600bp in size from within the 5' UTR of \emph{tb1} (Region 1) and from 1,235bp upstream of the start of the \emph{Hopscotch} (66,169bp upstream from the start of the \emph{tb1} ORF; Region 2).
-We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Appendix 1, See Supplemental Materials with the online version of this article).
+We designed the following primers using PRIMER3 \citep{RozenSkaletsky2000}: for the 5' UTR, {\small 5-'GGATAATGTGCACCAGGTGT-3'} and {\small 5'-GCGTGCTAGAGACACYTGTTGCT-3'}; for the 66kb upstream region, {\small 5'-TGTCCTCGCCGCAACTC-3'} and {\small 5'-TGTACGCCCGCCCCTCATCA-3'} (Table S1, See Supplemental Materials with the online version of this article).
 We used Taq polymerase (New England Biolabs Inc., Ipswich, Massachusetts, USA) and the following thermal cycler conditions to amplify fragments: 94$^{\circ}$C for 3 min, 30 cycles of 92$^{\circ}$C for 40 s, annealing for 1 min, 72$^{\circ}$C for 40 s, and a final 10 min extension at 72$^{\circ}$C.
@@
 \textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
-Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article).
+Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
@@
 For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}, Appendix 2, See Supplemental Materials with the online version of this article).
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}, Table S2, See Supplemental Materials with the online version of this article).
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
@@
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Appendix 5, See Supplemental Materials with the online version of this article); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Fig.
+S3, See Supplemental Materials with the online version of this article); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
 
@@
 If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Figure~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
+We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).
@@
 While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
-The \emph{Hopscotch} allele is more prevalent in \emph{parviglumis} than in \emph{mexicana} in our sample (Appendix 1, See Supplemental Materials with the online version of this article), suggesting a different history of the allele amongst teosinte subspecies.
+The \emph{Hopscotch} allele is more prevalent in \emph{parviglumis} than in \emph{mexicana} in our sample (Table S1, See Supplemental Materials with the online version of this article), suggesting a different history of the allele amongst teosinte subspecies.
 Moreover, many of our \emph{parviglumis} populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, and further distinguish this region from the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.

### 2014-12-10 manuscript.tex
final edits before submission to PeerJ

@@
 %----------------------------------------
-
-\makeatletter
-\renewcommand{\maketitle}{\bgroup\vspace*{1cm}%\setlength{\parindent}{0pt}%special formatting for AJB for title page
-\begin{center}
-  \textbf{\@title}
-
-  \@author
-\end{center}\egroup
-}
-\makeatother
 
@@
 
-A recent study by \citet{StuderDoebley2012} examined variation across traits in a three-taxa allelic series at the \emph{tb1} locus.
-\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize background and investigated their phenotypic effects.
-Their findings suggest that different teosinte \emph{tb1} segments contain distinct genetic factors that differentially contribute to tillering in teosinte.
-Furthermore, \citet{StuderDoebley2012} found significant variation in tillering within their control maize lines, concluding that variation in \emph{tb1} alone cannot account for the differences in tillering, and that other genetic factors must also contribute to differences in tillering in teosinte.
+A recent study by \citet{StuderDoebley2012} examined variation across traits in an allelic series study of the \emph{tb1} locus.
+\citet{StuderDoebley2012} introgressed nine unique teosinte \emph{tb1} segments (one from \emph{Zea diploperennis}, and four each from \emph{mexicana} and \emph{parviglumis}) into an inbred maize (W22) background and investigated their phenotypic effects.
+Their findings suggest that different teosinte \emph{tb1} segments produce equivalent effects on tillering and that variation in tillering observed across these taxa is not due to a \emph{tb1} allelic series but potentially due to variation at other, unlinked loci.
 Clues to the identity of these loci may be found in QTL studies that have identified loci controlling branching architecture (\emph{e.g.,} \citealt{DoebleyStec1991, DoebleyStec1993}).
@@
 \subsection*{Acknowledgements}%subsection header
-The authors thank Graham Coop for helpful discussion.
+The authors thank Graham Coop for helpful discussion and Lauryn Brown, Joshua Hegarty, Pui Yan Ho, and Garry Pearson for assistance with the phenotyping portion of this study.
 

### 2015-02-02 PeerJComments.tex
Comments from PeerJ Reviewers
PeerJ reviewer comments detailed out and numbered

\documentclass[]{article}

\begin{document}

\title{Reviewer comments from PeerJ}
\textbf{Reviewer 1 (Anonymous)}\\

\underline{Comments for the author}\\

Populations of crop wild relatives often carry alleles associated with domestication traits at low to moderate frequencies. The presence of these alleles could potentially reflect post-domestication crop-to-wild gene flow or standing genetic variation that predates the domestication process. This study examines the frequency and phenotypic consequences of one of the best characterized crop domestication alleles, maize tb1, in populations of its wild ancestor, teosinte. The tb1 gene functions as a shade avoidance-mediated repressor of organ growth. In domesticated maize, a Hopscotch TE insertion in the tb1 promoter leads to enhanced gene expression and the loss of tillering and branching that characterizes domesticated maize. In this paper, the authors use PCR assays to survey for the presence of the TE insertion in a large sample of accessions, including maize, its wild ancestor teosinte, and highland teosinte (a separate subspecies less closely related to the domesticate). The study examines the potential ecological significance of the TE insertion allele through a Bayenv analysis, and population genetic analyses are used to confirm previous findings that the TE insertion predates maize domestication.\\

The paper is generally well written, and the conclusions are basically sound. Addressing the points below would improve the paper. \\

1) The terminology on “accessions,” “individuals,” and “populations” is confusing (e.g., 67-71). Please clarify the difference between individuals and accessions (e.g., line 70 — is this referring to the number of individuals per accession? How are individuals of one accession related? Are these full sibs from the same maternal plant?..) \\

2) Is the “Locality” designation in Table S1 the same as “population” as used in the text? Please clarify and/or be consistent with this terminology.\\

3) Fig. 1 should indicate the boundaries of the Mexican states that are used in describing population locations in the text (Jalisco, Colima, etc.). The four focal populations from Jalisco state should also be labeled in this figure (i.e., San Lorenzo, La Mesa, Ejutla A, Ejutla B).\\

4) It would be helpful to include Fig. S1 in the main text of the paper and to label the two sequenced regions (Region 1 and 2) in the figure itself (not just in the caption). \\

5) Fig. S2. Annotate this figure to show the size (bp) of each band in the 1 kb ladder. The caption should also define the abbreviations used in the labels. \\

6) What does “No Hop/Pif” indicate in Fig. S2? It doesn’t have the three predicted bands of a TE heterozygote, and lane 6 seems to be empty altogether. Please clarify. \\

7) Line 118. Figure S1 should be cited here, not Table S1. \\

8) Line 150. Why were only 100 bootstrap replicates used for the NJ analysis? 1000 is more typical.\\

9) Lines 164-175. It would be helpful to provide some more detail on the methods used in defining haplotypes in the STRUCTURE introgression analysis. If I understand correctly, SNPs within 5 kb windows were used to define haplotypes for that window. Were the haplotypes then used to define diploid genotypes for each individual in the STRUCTURE input file? If so, how was phasing performed in defining the haplotypes? Is the output in Fig. 3 showing maize population assignment (i.e., membership coefficient values) for adjacent 5 kb windows? Was the STRUCTURE analysis performed only for the 8 Mb region shown in Fig. 3, or do the results presented in Table 4 reflect the entire chromosome? \\

11) Line 186. How does the 1-foot spacing of plants in the phenotyping experiment compare to conditions used in the experiments of Lukens and Doebley that documented density-dependent tillering? It could be worth bringing this up in the discussion of the negative phenotyping results (i.e., Fig 4 and Discussion, p. 13). \\

12) Line 223. Indicate here that these four populations are parviglumis and are from Jalisco state. \\

13) Line 228. Why is Table S2 cited here? That table doesn’t include any information on nucleotide diversity estimators.\\

14) Table 2 and the corresponding text (p. 9) should indicate which Tajima’s D values, if any, are statistically significant. If none are statistically significant, then this should be pointed out in the text. \\

15) Lines 231-232. Suggested rewording: “in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants…”\\

16) Lines 249-250. “average r2 is slightly lower in the tb1 region..” Also, include these mean values as a row in Table 3. \\

17) Fig. 3 could be moved to online supplementary data, since the key information with respect to the tb1 region is already presented in Table 4. \\

18) Discussion, lines 306-8. “The Hopscotch allele is more prevalent in parviglumis than in mexicana in our sample, suggesting a different history of the allele amongst teosinte subspecies.” This inference is contradicted by the complete lack of differentiation between the subspecies for the TE (Fct = 0; Results, lines 213-217). This should be addressed. I don’t see any evidence presented that the TE differences between the subspecies are statistically significant. Given the unbalanced geographical sampling of the two subspecies (see Table S1), I would also question whether these samples are appropriate for comparing TE frequencies in this way. \\

19) There are some sloppy inferences about selection in the Discussion and Conclusion — e.g., lines 345-354, 388-389: there’s no evidence that the Tajima’s D values cited here are actually statistically significant. The statement and citations in lines 349-351 suggests that they likely are, but this should be backed up with tests of statistical significance. See comment 14 above.\\

\textbf{Reviewer 2 (Anonymous)}

\underline{Basic reporting}\\

1) In the Introduction, the authors give some background about teosinte, but do so mainly as the precursor of maize, not so much as a wild species. For instance, they do not give much information about the ecological differences between parviglumis and mexicana, or about what is known for this species in terms of population structure, population density or ecological behavior. This should be added. In particular, the authors state that the tb1 gene is suspected to play a role in shade avoidance, so information on population density should be highlighted: Is it similar in all teosinte populations? Does it vary with altitude? With other ecological data? Does it vary between parviglumis and mexicana?\\

2) The authors refer several times to a previous study they performed: Pyhäjärvi et al., 2013. This is fine, but it is not always easy to understand which results are new and which come from this previous study. For instance, on line 158, it is stated “we had” while whole genome SNP data come from Pyhäjärvi et al., 2013. Similarly, “Table 1” on line 217 should be moved right after “populations (0.23)” on line 216 to avoid confusion.\\

3) In the Results section, some titles like “Genotyping”, “Sequencing” or “Phenotyping” are too vague, and sound more like Materials and Methods than to Results. They should be modified in order to highlight what question is addressed.
Paragraph “Neighbor joining trees… insertion”” on lines 235-239 should be moved to the section “Evidence of introgression”.\\

\underline{Experimental design}\\

4) The biological question is to characterize whether the Hopscotch transposable element at tb1 (which enhances expression of tb1) plays a role in the ecology of teosinte, especially in high-density populations. To do so, they characterize the distribution of this element in parviglumis, mexicana and maize landraces and they examine the phenotypic effects of the insertion in parviglumis.
The authors state that they sampled 1,110 individuals from 350 accessions, with between 1-18 and 1-43 individuals per population. However, they do not provide explanation on (i) how these populations were chosen and (ii) why a different number of individual was sampled per accession (visible also in suppl. Tables S1 and S2). This should be explained.\\

5) In Suppl. Table S1, some of the parviglumis accessions are listed as “Breeders line”. Can these be considered as natural populations? The authors should explain where they were originally collected. In Suppl. Table S2, the USDA (or other provider) ID corresponding to each accession should be provided, as it is done in suppl. Table S1.\\

6) The authors write that the PCR amplification to investigate presence of the Hopscotch element leads to two amplification products: one for the entire element, (5kb) and one for amplification of part of the element only (1.1kb). Why is this? How do the authors explain the origin of the 1.1kb band? The sentence with “and” on line 83, suggests that the two bands are produced in one single homozygous plant. If some accessions really amplify two bands, is the primer located in the LTR? If it is a typo and the “and” on line 83 should be replaced by a “or”, is the element truncated in some accessions, leading to a 1.1kb band instead of the 5kb one? This should be clarified.
From Figure S1, the middle primer seems to be right in the middle of the element, which suggests it was not designed in the LTR region. Is this true? In which part of the retrotransposon was the primer designed? The authors should add position of the retrotransposon LTRs and coding regions on this figure. 
In suppl. Figure S1, the names “HopF/HopR/HopIntR” should be added, as well as corresponding expected amplification sizes.
In suppl. Figure S2, band size of several bands (close in size to these amplified) should be indicated next to the ladder. Legend should explain better what the figure shows. It states “Genotypes are indicated at the top of the gel.” What does this refer to? Numbers? “Hop/Hop” code? The figure should be completely relabeled, so that primer pair names and type of genotype (with presence of absence of Hopscotch) are clearly identified. What is “no Hop/Pif”? Why are there two bands on lane 5? Lane 6 has a weird smear. No primers are visible on lanes 5, 6 and 7.
Bottom gel resolution is poor. On the right for low molecular weight bands, it is difficult to assess that there is a single band (lane 7). A 1% gel and 1kb ladder are clearly not adequate for 300bp band detection.
It is not stated whether PCR products were sequenced to check for correct amplification. Were some of the PCR products sequenced? If not, this should be done.

The authors write: “Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data)”. Does this mean that soil data was not averaged over the last decades? How was it estimated? The fact that soil parameters are “likely” stable over time actually quite depends on what we consider as “soil”. Does this include micro-organisms? A bit more detail should be given.
Figure 1: This figure should show 37 populations of parviglumis and 4 populations of mexicana (see page 8, lines 211 to 213). But the figure legends indicates only parviglumis. It should include mexicana populations, and they should be differentiated by an appropriate color code. In some cases, several circles seem to derive from one circle, suggesting they correspond to several populations from the same location. On the left hand side, it is unexpected to globally have a null frequency of “No Hopscotch” while several populations have a large fraction of “Hopscotch”. This means that, within 25 miles, the frequency of the “Hopscotch” allele varies greatly. The authors should discuss this point. Names of the parviglumis populations used for the rest of the study (La Mesa, San Lorenzo, Ejutla1 and Ejutla2) should be indicated.

Validity of the findings

Results of the PCR amplifications are not well enough described and of high enough quality to be able to conclude whether the retrotransposon detection is valid or not (see “Experimental Design”). The authors should improve labelling of suppl. Figure S2 and, if needed, run the PCR amplicons on a more concentrated gel and with appropriate ladder.
Based on the association genetics study performed, the authors find an absence of association between the Hopscotch insertion and tillering index. At 40 days, they even find a weak but significant correlation, but in the unexpected direction (homozygotes for the Hopscotch insertion have a higher tillering). The authors discuss that this could be due to variation at other unlinked loci. However, they do not discuss on the effect of the environment on tillering. Considering the experiment was performed in greenhouse conditions while teosintes grow in much different environmental conditions, the genotype x environment interaction could differ greatly to this obtained in teosinte natural environmental conditions. This should be discussed.

Comments for the author

No additional comments.

\end{document}

### 2015-02-02 PeerJComments.tex
Complete comments from PeerJ reviewers
Comments for both authors listed out and numbered

@@
 If it is a typo and the “and” on line 83 should be replaced by a “or”, is the element truncated in some accessions, leading to a 1.1kb band instead of the 5kb one?
-This should be clarified.
-From Figure S1, the middle primer seems to be right in the middle of the element, which suggests it was not designed in the LTR region.
+This should be clarified.\\
+
+7) From Figure S1, the middle primer seems to be right in the middle of the element, which suggests it was not designed in the LTR region.
 Is this true?
@@
 The authors should add position of the retrotransposon LTRs and coding regions on this figure.
-In suppl. Figure S1, the names “HopF/HopR/HopIntR” should be added, as well as corresponding expected amplification sizes.
-In suppl.
+\\
+
+8) In suppl. Figure S1, the names “HopF/HopR/HopIntR” should be added, as well as corresponding expected amplification sizes.\\
+
+9) In suppl.
 Figure S2, band size of several bands (close in size to these amplified) should be indicated next to the ladder.
@@
 Lane 6 has a weird smear.
-No primers are visible on lanes 5, 6 and 7.
-Bottom gel resolution is poor.
+No primers are visible on lanes 5, 6 and 7.\\
+
+10) Bottom gel resolution is poor.
 On the right for low molecular weight bands, it is difficult to assess that there is a single band (lane 7).
-A 1% gel and 1kb ladder are clearly not adequate for 300bp band detection.
-It is not stated whether PCR products were sequenced to check for correct amplification. Were some of the PCR products sequenced? If not, this should be done.
+A 1\% gel and 1kb ladder are clearly not adequate for 300bp band detection.\\
 
-The authors write: “Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data)”.
+11) It is not stated whether PCR products were sequenced to check for correct amplification. Were some of the PCR products sequenced? If not, this should be done.\\
+
+12) The authors write: “Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data)”.
 Does this mean that soil data was not averaged over the last decades?
@@
 Does this include micro-organisms?
-A bit more detail should be given.
-Figure 1: This figure should show 37 populations of parviglumis and 4 populations of mexicana (see page 8, lines 211 to 213).
+A bit more detail should be given.\\
+
+13) Figure 1: This figure should show 37 populations of parviglumis and 4 populations of mexicana (see page 8, lines 211 to 213).
 But the figure legends indicates only parviglumis.
@@
 The authors should discuss this point.
-Names of the parviglumis populations used for the rest of the study (La Mesa, San Lorenzo, Ejutla1 and Ejutla2) should be indicated.
+Names of the parviglumis populations used for the rest of the study (La Mesa, San Lorenzo, Ejutla1 and Ejutla2) should be indicated.\\
 
-Validity of the findings
+\underline{Validity of the findings}\\
 
-Results of the PCR amplifications are not well enough described and of high enough quality to be able to conclude whether the retrotransposon detection is valid or not (see “Experimental Design”).
+14) Results of the PCR amplifications are not well enough described and of high enough quality to be able to conclude whether the retrotransposon detection is valid or not (see “Experimental Design”).
 The authors should improve labelling of suppl.
-Figure S2 and, if needed, run the PCR amplicons on a more concentrated gel and with appropriate ladder.
-Based on the association genetics study performed, the authors find an absence of association between the Hopscotch insertion and tillering index.
+Figure S2 and, if needed, run the PCR amplicons on a more concentrated gel and with appropriate ladder.\\
+
+15) Based on the association genetics study performed, the authors find an absence of association between the Hopscotch insertion and tillering index.
 At 40 days, they even find a weak but significant correlation, but in the unexpected direction (homozygotes for the Hopscotch insertion have a higher tillering).
@@
 Considering the experiment was performed in greenhouse conditions while teosintes grow in much different environmental conditions, the genotype x environment interaction could differ greatly to this obtained in teosinte natural environmental conditions.
-This should be discussed.
-
-Comments for the author
-
-No additional comments.
+This should be discussed.\\
 

### 2015-02-02 PeerJComments.tex
PeerJ Reviewer comments
Comments to reviewer 1 points

@@
-\documentclass[]{article}
+\documentclass[11pt]{article}
+\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
+\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
+\newcommand{\lev}[1]{\noindent \textcolor{magenta}{{#1}} \\}
+\newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
+\usepackage{lipsum}
+\usepackage{url}
+\usepackage[usenames,dvipsnames]{color}
+\usepackage{multirow}
+\renewcommand{\baselinestretch}{1.5}
 
@@
 Are these full sibs from the same maternal plant?..) \\
+\lev{okay we can do this...}
 
 2) Is the “Locality” designation in Table S1 the same as “population” as used in the text? Please clarify and/or be consistent with this terminology.\\
+\lev{Yes, can clarify}
 
@@
 The four focal populations from Jalisco state should also be labeled in this figure (i.e., San Lorenzo, La Mesa, Ejutla A, Ejutla B).\\
+\lev{I think this would make figure super busy and it's already verging on super busy. Maybe compromise on labelling the four pops within their pie charts.}
 
 4) It would be helpful to include Fig. S1 in the main text of the paper and to label the two sequenced regions (Region 1 and 2) in the figure itself (not just in the caption). \\
+\lev{Okay can do this}.
 
 5) Fig. S2. Annotate this figure to show the size (bp) of each band in the 1 kb ladder. The caption should also define the abbreviations used in the labels. \\
+\lev{this will be too much. I can label the brightest band (1kb) and the top and bottom bands}.
 
@@
 
+\lev{Maybe in the text I can clarify total bands for individuals.
+I.e.
+if an individual is Hop/Hop, it will have a total of two bands that are split between the two PCRs, whereas if you are a heterozygote you will have 3 split between two PCRs.
+I wonder if we should ditch this figure entirely and I can draw up a cartoon one?}
+
 7) Line 118. Figure S1 should be cited here, not Table S1. \\
+\lev{Okay can fix}.
 
 8) Line 150. Why were only 100 bootstrap replicates used for the NJ analysis? 1000 is more typical.\\
+\lev{I do not know the answer to this. I don't think that 1000 bootstraps would give us really a difference.}
 

### 2015-02-02 PeerJComments.tex
PeerJ Reviewer comments
Comments in response to points from reviewer 1 and reviewer 2

@@
 4) It would be helpful to include Fig. S1 in the main text of the paper and to label the two sequenced regions (Region 1 and 2) in the figure itself (not just in the caption). \\
-\lev{Okay can do this}.
+\lev{Okay can do this}
 
 5) Fig. S2. Annotate this figure to show the size (bp) of each band in the 1 kb ladder. The caption should also define the abbreviations used in the labels. \\
-\lev{this will be too much. I can label the brightest band (1kb) and the top and bottom bands}.
+\lev{this will be too much. I can label the brightest band (1kb) and the top and bottom bands}
 
@@
 7) Line 118. Figure S1 should be cited here, not Table S1. \\
-\lev{Okay can fix}.
+\lev{Okay can fix}
 
@@
 
+\lev{yes SNPs within 5KB window are a haplotype.
+I don't know about the diploid/phasing stuff.
+Yes to maize assignment.
+Point of figure is that low assignment to maize in populations with high frequency of Hopscotch.
+Am plotting the assigment value across a region of chromosome 1.
+I showed a region and not entire chromosome to be able to highlight the position of tb1 and so that you can see the difference in assignment values.}
+
 11) Line 186.
@@
 
+\lev{No idea off the top of my head. Good point, can look up and mention}
+
 12) Line 223. Indicate here that these four populations are parviglumis and are from Jalisco state. \\
 
+\lev{Okay can mention}
+
 13) Line 228. Why is Table S2 cited here? That table doesn’t include any information on nucleotide diversity estimators.\\
+
+\lev{will check}
 
@@
 
+\lev{Okay. I am not sure that we have significance values for these.}
+
 15) Lines 231-232. Suggested rewording: “in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants…”\\
+
+\lev{Okay reword.}
 
@@
 
+\lev{Do you guys want to do this?}
+
 17) Fig. 3 could be moved to online supplementary data, since the key information with respect to the tb1 region is already presented in Table 4. \\
+
+\lev{I have no strong feelings about this.}
 
@@
 
+\lev{Okay we can maybe take these statements out?
+It would seem that none of the reviewers have ever been happy with what we have had to say with respect to mex vs parv, or unequal sampling.
+we have discussed why sampling is unequal and why there was no phenotyping for mex.}
+
 19) There are some sloppy inferences about selection in the Discussion and Conclusion — e.g., lines 345-354, 388-389: there’s no evidence that the Tajima’s D values cited here are actually statistically significant.
@@
 See comment 14 above.\\
+
+\lev{Okay. What do you guys think of this? I don't think we have tests of significance for tajd}
 
@@
 
+\lev{I don't think this is necessary.
+Again, everyone has seemed to want something different in the intro.
+Last person wanted more maize/shade avoidance/tb1.
+Perhaps we can add a couple of sentences in about mex vs parv or the population stuff.
+I would like to not have to rewrite the entire intro again.}
+
 2) The authors refer several times to a previous study they performed: Pyhäjärvi et al., 2013.
@@
 
+\lev{Okay I can check to make sure citations and data stuff from tanja are clear}
+
 3) In the Results section, some titles like “Genotyping”, “Sequencing” or “Phenotyping” are too vague, and sound more like Materials and Methods than to Results.
@@
 Paragraph “Neighbor joining trees… insertion”” on lines 235-239 should be moved to the section “Evidence of introgression”.\\
+ \lev{I like our sections...}
 
@@
 
+\lev{Simply put it is what we had available at the time. There was no specific choosing or decision about number of individuals per accession}
+
 5) In Suppl.
@@
 Table S1.\\
+
+\lev{Okay. Is there somewhere we can get all of this info.? I don't believe I have access to the database any more}
 
@@
 
+\lev{Why what?
+Why two products?
+or why did we do it that way?
+I think it is clearly stated that we can test for presence/absence/hets and pcr bias with sometimes a large fragment being difficult to amplify.
+The location of the primers is all shown in the cartoon figure and stated in the text.
+two primers outside of LTR one primer within.}
+
 7) From Figure S1, the middle primer seems to be right in the middle of the element, which suggests it was not designed in the LTR region.
@@
 
+\lev{What? It is in the LTR!!!! Gah! That's the whole point!}
+
 8) In suppl. Figure S1, the names “HopF/HopR/HopIntR” should be added, as well as corresponding expected amplification sizes.\\
+
+\lev{Okay can do this. maybe will clear up their misunderstanding detailed above}
 
@@
 
+\lev{I can relabel or clarify in the figure legend. I don't think that redoing a gel is possible at this point.}
+
 10) Bottom gel resolution is poor.
@@
 
+\lev{Redoing gel not possible at this point}
+
 11) It is not stated whether PCR products were sequenced to check for correct amplification. Were some of the PCR products sequenced? If not, this should be done.\\
+
+\lev{I think that we did initially to check.
+But all of the products match the appropriate and expected size.
+The only issue we had was suddenly the different size with the Pif, which we did check with sequencing.}
 
@@
 A bit more detail should be given.\\
+
+\lev{My guess is that we can look this up on WorldClim if you guys think it is necessary.
+My guess is that microorganisms were not included in analysis of soil characteristics, that it was things like what compounds are in soil, etc.}
 
@@
 
+\lev{i will check this but i believe mexicana not included because we didn't discuss them with respect to the high frequency stuff.
+I do not have arcgis any more and would prefer to not completely redo this figure.
+I am happy to put names on pops though.
+I don't understand comments about 'left hand side' stuff.
+Yes, some of the populations are close together when you are that zoomed out on a map.
+There is no way to have them not all coming from the same place in some cases unless the map was so gigantic to take up an entire page.}
+
 \underline{Validity of the findings}\\
@@
 Figure S2 and, if needed, run the PCR amplicons on a more concentrated gel and with appropriate ladder.\\
+
+\lev{We can relabel but not redo gels at this point}
 
@@
 
+\lev{Sure greenhouse is different than outside, but we can't control for variation outside whereas we can in a greenhouse.
+I don't understand their point.
+Greenhouse is best we could do here since they wouldn't all fit in a growth chamber.}
+
 \end{document}

### 2015-02-19 PeerJResponse.tex
Edits to comments, additional files
Track changes document that is required for resubmission; Response
letter, edits to comments, list of things to address

--- 2015-02-02 PeerJComments.tex
@@
 \usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
+\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
 \newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
-\newcommand{\lev}[1]{\noindent \textcolor{magenta}{{#1}} \\}
+\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
 \newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
@@
 \renewcommand{\baselinestretch}{1.5}
+\usepackage{natbib} % package for formatting citations
+%\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for
 
@@
 
-\title{Reviewer comments from PeerJ}
-\textbf{Reviewer 1 (Anonymous)}\\
+\noindent Dear Dr.Vision and editorial staff at the \emph{PeerJ}, \\
+
+We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our previously-submitted manuscript (#2014:12:3442:0:0:REVIEW).
+In the attached revision, we have addressed the majority of issues raised, and have included a detailed response to reviewers below.
+In our submitted manuscript, we report results of a genotyping assay surveying several hundred wild and domesticated maize individuals for presence of the domesticated maize allele at \emph{tb1}, Sanger sequencing of a targeted subset of these samples to assess evidence of selection on the domesticated allele in wild populations, a phenotyping experiment to gauge the effects of the domesticated allele in a wild population, and reanalysis of existing genome-wide SNP data to place the signature of selection at this locus in a wider context.
+This manuscript reports a substantial amount of novel work and explores, in a meaningful way, the concept of domestication from standing variation and the potential ecological relevance of domesticated alleles in wild populations.\\\
+
+\noindent Sincerely,\\
+
+\noindent Matthew Hufford and Jeffrey Ross-Ibarra
+
+\section*{Response to Editor:}
+
+1) The reviewers comments seems to be largely in agreement and self-explanatory. Please be sure to address the comment regarding the validity of the PCR assays raised by Reviewer 2.
+
+\res{We have addressed the comments from Reviewer 2 regarding the validity of the PCR assays.
+Sequencing of PCR products was never performed as primer combinations for each of the 2 PCRs gave us the expected sized bands.
+Primers were designed off of previously sequenced teosinte and maize sequences \lev{these were from John Doebley, do we need to say this?
+I don' think they were published ever} and were initially tested in both maize and teosinte for validity.
+We have clarified this point in text \lev{text, line numbers here} as well added additional labelling to the gel image (Figure 2 \lev{check to make sure figure number doesn't change during edits}).
+
+\section*{Response to Reviewer 1}
 
@@
 
-1) The terminology on “accessions,” “individuals,” and “populations” is confusing (e.g., 67-71).
-Please clarify the difference between individuals and accessions (e.g., line 70 — is this referring to the number of individuals per accession?
+1) The terminology on accessions, individuals, and populations is confusing (e.g., 67-71).
+Please clarify the difference between individuals and accessions (e.g., line 70‚ is this referring to the number of individuals per accession?
 How are individuals of one accession related?
@@
 \lev{okay we can do this...}
+\mbh{yep, pretty straight-forward}
 
-2) Is the “Locality” designation in Table S1 the same as “population” as used in the text? Please clarify and/or be consistent with this terminology.\\
+2) Is the Locality designation in Table S1 the same as population as used in the text? Please clarify and/or be consistent with this terminology.\\
 \lev{Yes, can clarify}
@@
 \lev{I think this would make figure super busy and it's already verging on super busy. Maybe compromise on labelling the four pops within their pie charts.} 
+\mbh{if you still have access to the figure in ArcGIS, should be easy to add state boundaries and you can make the boundaries very light to limit busyness; if not, I'd just label the Jalisco pies}
+\jri{raster in R is not too hard either.}
+\lev{I do not have access to the either the figure in ArcGIS or ArcGIS, so will just label the pies in something else and call it a day}
 
@@
 \lev{Okay can do this}
+\mbh{yep, easy to do and this is an online publication without figure charges so why not!}
 
@@
 \lev{this will be too much. I can label the brightest band (1kb) and the top and bottom bands}
+\mbh{I'd label as many as it takes so that it's obvious what intervening band sizes are.  We'll want to define the abbreviations too}
+\lev{usually the bottom, middle brightest and top are labelled so I will do that}
 
-6) What does “No Hop/Pif” indicate in Fig. S2? It doesn’t have the three predicted bands of a TE heterozygote, and lane 6 seems to be empty altogether. Please clarify. \\
+6) What does No Hop/Pif indicate in Fig. S2? It doesn't have the three predicted bands of a TE heterozygote, and lane 6 seems to be empty altogether. Please clarify. \\
 
@@
 I wonder if we should ditch this figure entirely and I can draw up a cartoon one?}
+\mbh{since reviewer 2 is also concerned about the PCR and this is the only thing the editor really mentioned, I think we need to be pretty thorough here.
+Ideally we would have a Hop/Hop homozygote, a Hop/No Hop heterozygote, and a No Hop/No Hop homozygote in the figure.
+We don't need the Pif since this really isn't mentioned in the manuscript.
+Do you have gel figures showing both homozygotes and the heterozygote banding patterns?
+I'm afraid a cartoon would send the message that we don't have clear PCR results}
+\jri{agree with the huff}
+\lev{I do not have gel figures.
+I am guessing the ones that had been saved online in Jeff's lab have been tossed.
+I can check my lab notebook and see if there is something to scan, otherwise there will not be a figure with all of those genotypes that we can give to them.}
 
@@
 \lev{I do not know the answer to this. I don't think that 1000 bootstraps would give us really a difference.}
+\mbh{if you still have the input file for this analysis, should be a quick fix and show we're doing everything we can to address reviewer concerns}
+\jri{agreed, should be easy to redo. while agree this is silly and i don't think 1000 will change anything, but since it's easy to do it's hard to argue against doing it.}
+\lev{I do not have the files as this is something that Tom did with the sequence data that he gathered}
 
@@
 
+\mbh{methods say that you used phased data from Tanja's manuscript.
+Did she give you phased data or did you phase this yourself with fastPHASE?
+Either way, probably best to mention the data were phased with fastPHASE and add a citation to the methods paper for this}
+
+\lev{I believe that she phased it since it was the exact thing that she used for her BayEnv stuff}
+
 11) Line 186.
@@
 
+\mbh{yep, good point to check on...I'd add some discussion on this, particularly if they saw very low tillering at this density}
+
 12) Line 223. Indicate here that these four populations are parviglumis and are from Jalisco state. \\
@@
 
-13) Line 228. Why is Table S2 cited here? That table doesn’t include any information on nucleotide diversity estimators.\\
+13) Line 228. Why is Table S2 cited here? That table doesn't include any information on nucleotide diversity estimators.\\
 
@@
 14) Table 2 and the corresponding text (p.
-9) should indicate which Tajima’s D values, if any, are statistically significant.
+9) should indicate which Tajima's D values, if any, are statistically significant.
 If none are statistically significant, then this should be pointed out in the text.
@@
 
-15) Lines 231-232. Suggested rewording: “in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants…”\\
+\mbh{We don't and it's kind of a strange/silly request...Jeff and I will deal with this}
+\jri{ i can do the simple but dumb calculation if you guys want for response to reviewers, but disagree we should include in manuscript.
+we mention they are at the tail of the empirical distribution from wright et al don't we?}
+
+15) Lines 231-232. Suggested rewording: 'in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants'\\
 
@@
 
-16) Lines 249-250. “average r2 is slightly lower in the tb1 region..” Also, include these mean values as a row in Table 3. \\
+16) Lines 249-250. 'average r2 is slightly lower in the tb1 region..' Also, include these mean values as a row in Table 3. \\
 
 \lev{Do you guys want to do this?}
+
+\mbh{simple enough to make the reviewer happy here and just add one more row to the table}
 
@@
 \lev{I have no strong feelings about this.}
+\mbh{fine with me to move to supplement}
 
 18) Discussion, lines 306-8.
-“The Hopscotch allele is more prevalent in parviglumis than in mexicana in our sample, suggesting a different history of the allele amongst teosinte subspecies.” This inference is contradicted by the complete lack of differentiation between the subspecies for the TE (Fct = 0; Results, lines 213-217).
+'The Hopscotch allele is more prevalent in parviglumis than in mexicana in our sample, suggesting a different history of the allele amongst teosinte subspecies.' This inference is contradicted by the complete lack of differentiation between the subspecies for the TE (Fct = 0; Results, lines 213-217).
 This should be addressed.
-I don’t see any evidence presented that the TE differences between the subspecies are statistically significant.
+I don't see any evidence presented that the TE differences between the subspecies are statistically significant.
 Given the unbalanced geographical sampling of the two subspecies (see Table S1), I would also question whether these samples are appropriate for comparing TE frequencies in this way.
@@
 
-19) There are some sloppy inferences about selection in the Discussion and Conclusion — e.g., lines 345-354, 388-389: there’s no evidence that the Tajima’s D values cited here are actually statistically significant.
+\mbh{my vote would be to remove the comparisons between \emph{parv} and \emph{mex} but to still note the high frequency in Jalisco}
+\jri{ok either way here}
+\lev{I vote remove since everyone has seemed unhappy with this}
+
+19) There are some sloppy inferences about selection in the Discussion and Conclusion 'e.g., lines 345-354, 388-389: there's no evidence that the Tajima's D values cited here are actually statistically significant.
 The statement and citations in lines 349-351 suggests that they likely are, but this should be backed up with tests of statistical significance.
@@
 
-\textbf{Reviewer 2 (Anonymous)}
+\mbh{Jeff and I will address this}
+\jri{sloppy my ass. <censored> reviewer!}
+\lev{ :) }
+
+\section*{Response to Reviewer 2}
 
@@
 
-2) The authors refer several times to a previous study they performed: Pyhäjärvi et al., 2013.
+\mbh{I might mention something briefly about \emph{parviglumis} being a ruderal species that can occur at high density (cite my \emph{Molecular Ecology} manuscript), but, no, I would not add a whole paragraph on the ecologies of the teosintes.}
+\jri{this sounds OK. a wee bit of self-citation there hufford?}
+
+2) The authors refer several times to a previous study they performed: Pyh√§j√§rvi et al., 2013.
 This is fine, but it is not always easy to understand which results are new and which come from this previous study.
-For instance, on line 158, it is stated “we had” while whole genome SNP data come from Pyhäjärvi et al., 2013.
-Similarly, “Table 1” on line 217 should be moved right after “populations (0.23)” on line 216 to avoid confusion.\\
+For instance, on line 158, it is stated 'we had whole genome SNP data come from Pyh√§j√§rvi et al., 2013.
+Similarly, Table 1 on line 217 should be moved right after 'populations (0.23) on line 216 to avoid confusion.\\
 
@@
 
-3) In the Results section, some titles like “Genotyping”, “Sequencing” or “Phenotyping” are too vague, and sound more like Materials and Methods than to Results.
+3) In the Results section, some titles like "Genotyping" "Sequencing" or "Phenotyping" are too vague, and sound more like Materials and Methods than to Results.
 They should be modified in order to highlight what question is addressed.
-Paragraph “Neighbor joining trees… insertion”” on lines 235-239 should be moved to the section “Evidence of introgression”.\\
  \lev{I like our sections...}
+ \mbh{easy enough to make these more hypothesis oriented; good idea to do as many of these simple things as possible to show we're making a strong effort toward revisions}
+
+Paragraph "Neighbor joining trees" on lines 235-239 should be moved to the section "Evidence of introgression"\\
 
@@
 \lev{Simply put it is what we had available at the time. There was no specific choosing or decision about number of individuals per accession}
+\mbh{we'll have to diplomatically state this...we sampled all teosinte available to us and while this didn't allow for even sample sizes across populations, it did allow us to calculate both frequency in a subset of populations and geographic breadth of the \emph{Hopscotch} across a large number of independent sampling locations}
 
 5) In Suppl.
-Table S1, some of the parviglumis accessions are listed as “Breeders line”.
+Table S1, some of the parviglumis accessions are listed as "Breeders line".
 Can these be considered as natural populations?
@@
 \lev{Okay. Is there somewhere we can get all of this info.? I don't believe I have access to the database any more}
+\mbh{Yes, we need to include this info.; that's a fair request}
+\lev{Do either of you know where I can find this info?}
 
@@
 How do the authors explain the origin of the 1.1kb band?
-The sentence with “and” on line 83, suggests that the two bands are produced in one single homozygous plant.
+The sentence with "and" on line 83, suggests that the two bands are produced in one single homozygous plant.
 If some accessions really amplify two bands, is the primer located in the LTR?
-If it is a typo and the “and” on line 83 should be replaced by a “or”, is the element truncated in some accessions, leading to a 1.1kb band instead of the 5kb one?
+If it is a typo and the "and" on line 83 should be replaced by a "or" is the element truncated in some accessions, leading to a 1.1kb band instead of the 5kb one?
 This should be clarified.\\
@@
 
+\mbh{if there's any way to be even slightly more clear, I'd edit just so you can say here that you've edited for clarity}
+
 7) From Figure S1, the middle primer seems to be right in the middle of the element, which suggests it was not designed in the LTR region.
@@
 
-8) In suppl. Figure S1, the names “HopF/HopR/HopIntR” should be added, as well as corresponding expected amplification sizes.\\
+8) In suppl. Figure S1, the names "HopF/HopR/HopIntR" should be added, as well as corresponding expected amplification sizes.\\
 
@@
 Legend should explain better what the figure shows.
-It states “Genotypes are indicated at the top of the gel.” What does this refer to?
+It states "Genotypes are indicated at the top of the gel" What does this refer to?
 Numbers?
-“Hop/Hop” code?
+"Hop/Hop" code?
 The figure should be completely relabeled, so that primer pair names and type of genotype (with presence of absence of Hopscotch) are clearly identified.
-What is “no Hop/Pif”?
+What is "no Hop/Pif"?
 Why are there two bands on lane 5?
@@
 
+\mbh{even if you have a better resolution image to include, that might help.  I'd also remove the pif lanes as possible as we never mention this}
+
 11) It is not stated whether PCR products were sequenced to check for correct amplification. Were some of the PCR products sequenced? If not, this should be done.\\
@@
 
-12) The authors write: “Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data)”.
+\mbh{we need to check this.  can you go back through emails or to your lab notebook to find whether this was sequenced?}
+\jri{i don't remember ever checking this. we checked the pif by PCRing stuff John had already sequenced in which i identified the pif}
+\lev{I agree with Jeff. I believe bands matched with what was expected based on where we designed primers so we did not sequence them}
+
+12) The authors write: "Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data)".
 Does this mean that soil data was not averaged over the last decades?
 How was it estimated?
-The fact that soil parameters are “likely” stable over time actually quite depends on what we consider as “soil”.
+The fact that soil parameters are "likely" stable over time actually quite depends on what we consider as "soil".
 Does this include micro-organisms?
@@
 
+\mbh{soil data are from the Harmonized world soil database that we cite and that Tanja also used in her manuscript.
+There will be further details on data collection there which we can briefly describe in our response and point the reviewer/editor to}
+
 13) Figure 1: This figure should show 37 populations of parviglumis and 4 populations of mexicana (see page 8, lines 211 to 213).
@@
 In some cases, several circles seem to derive from one circle, suggesting they correspond to several populations from the same location.
-On the left hand side, it is unexpected to globally have a null frequency of “No Hopscotch” while several populations have a large fraction of “Hopscotch”.
-This means that, within 25 miles, the frequency of the “Hopscotch” allele varies greatly.
+On the left hand side, it is unexpected to globally have a null frequency of "No Hopscotch" while several populations have a large fraction of "Hopscotch".
+This means that, within 25 miles, the frequency of the "Hopscotch" allele varies greatly.
 The authors should discuss this point.
@@
 
+\mbh{the reviewer is indicating they are surprised that there are such marked differences in Hop frequency in such a small area in the populations on the lefthand side of the map.
+You can cite my dissertation and the Tiffin and Moeller paper as finding that populations in this area were highly differentiated and that populations with Hop were much more closely related to each other than they were to pops without Hop...just some very different histories in these populations and little gene flow between them}
+
 \underline{Validity of the findings}\\
 
-14) Results of the PCR amplifications are not well enough described and of high enough quality to be able to conclude whether the retrotransposon detection is valid or not (see “Experimental Design”).
+14) Results of the PCR amplifications are not well enough described and of high enough quality to be able to conclude whether the retrotransposon detection is valid or not (see "Experimental Design").
 The authors should improve labelling of suppl.
@@
 \lev{We can relabel but not redo gels at this point}
+\mbh{the PCR results were conclusive, but I'd make an effort to make the figure more clear with labeling or a higher resolution image and to clarify the description of the banding patterns of the genotypes in the text; my guess is we'll need to do a good job of this in particular if we want this manuscript in \emph{PeerJ}}
 
@@
 
+\mbh{I'd point out that the effect of \emph{tb1} on tillering has been verified a number of times in a number of environments (probably in a greenhouse setting) and cite studies like Briggs, etc...}
+
+\bibliography{refs}
+
 \end{document}

### 2015-02-19 PeerJResponse.tex
Edits
working through comments in response, cleaned up manuscript, and
required track changes document

@@
 
-We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our previously-submitted manuscript (#2014:12:3442:0:0:REVIEW).
+We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our previously-submitted manuscript (\#2014:12:3442:0:0:REVIEW).
 In the attached revision, we have addressed the majority of issues raised, and have included a detailed response to reviewers below.
@@
 I don' think they were published ever} and were initially tested in both maize and teosinte for validity.
-We have clarified this point in text \lev{text, line numbers here} as well added additional labelling to the gel image (Figure 2 \lev{check to make sure figure number doesn't change during edits}).
+We have clarified this point in text \lev{text, line numbers here} as well added additional labelling to the gel image (Figure 2 \lev{check to make sure figure number doesn't change during edits})}.
 
@@
 Are these full sibs from the same maternal plant?..) \\
-\lev{okay we can do this...}
-\mbh{yep, pretty straight-forward}
+
+\res{We have clarified this; accession refers to population; page 2, lines 62-65: ``We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article)."}
 

### 2015-02-19 PeerJ_trackchanges.tex
Edits
working through comments in response, cleaned up manuscript, and
required track changes document

--- 2014-12-10 manuscript.tex
@@
 \newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} } %introducting new command "\jri" for jeff ross-ibarra to make comments in blue italics
-\newcommand{\mbh}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} } %same for matt hufford in red
+\newcommand{\mbh}[1]{\textcolor{dandelion}{ \emph{\scriptsize  #1}} } %same for matt hufford in red
 \newcommand{\lev}[1]{\textcolor{magenta}{\emph{\normalsize #1}} } %same for laura vann in green
+\newcommand{\track}[1]{\textcolor{red}{\emph{\normalsize #1}} }
 
@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.

### 2015-02-19 manuscript.tex
Edits
working through comments in response, cleaned up manuscript, and
required track changes document

@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.

### 2015-02-19 PeerJResponse.tex
Edits
Edits to Figure 1 (Map), response, text, supplementary tables

@@
 
-\res{We have clarified this; accession refers to population; page 2, lines 62-65: ``We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article)."}
+\res{We have clarified this; accession refers to population; page 2, lines 62-66: ``We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations."}
 
 2) Is the Locality designation in Table S1 the same as population as used in the text? Please clarify and/or be consistent with this terminology.\\
-\lev{Yes, can clarify}
+
+\res{Yes, we have changed the terminology in Table S1 to be consistent with the terminology used in the text.}
 
@@
 \lev{I do not have access to the either the figure in ArcGIS or ArcGIS, so will just label the pies in something else and call it a day}
+
+\res{We feel that the addition of state lines makes the figure too busy and difficult to read.
+However, we have added labels to the 4 Jalisco populations (La Mesa (MSA), San Lorenzo (SLO), Ejutla A (EjuA), and Ejutla B (EjuB)).}
 

### 2015-02-19 PeerJ_trackchanges.tex
Edits
Edits to Figure 1 (Map), response, text, supplementary tables

@@
 \textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
-Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
+Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.

### 2015-02-19 Supplement_mbh.tex
final edits before submission to PeerJ
Edits
Edits to Figure 1 (Map), response, text, supplementary tables

@@
 \renewcommand{\headrulewidth}{0pt}
-\cfoot{\thepage}
 
@@
 \hline
-\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{USDA ID}} & \multicolumn{1}{c}{\textbf{Locality}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} & \multicolumn{1}{c}{\textbf{No \emph{Hopscotch} Freq.}} \\
+\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{USDA ID}} & \multicolumn{1}{c}{\textbf{Population}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} & \multicolumn{1}{c}{\textbf{No \emph{Hopscotch} Freq.}} \\
 \hline

### 2015-02-19 PeerJResponse.tex
Edits
Edits to cartoon figure; text; response

@@
 The four focal populations from Jalisco state should also be labeled in this figure (i.e., San Lorenzo, La Mesa, Ejutla A, Ejutla B).\\
-\lev{I think this would make figure super busy and it's already verging on super busy. Maybe compromise on labelling the four pops within their pie charts.} 
-\mbh{if you still have access to the figure in ArcGIS, should be easy to add state boundaries and you can make the boundaries very light to limit busyness; if not, I'd just label the Jalisco pies}
-\jri{raster in R is not too hard either.}
-\lev{I do not have access to the either the figure in ArcGIS or ArcGIS, so will just label the pies in something else and call it a day}
 
 \res{We feel that the addition of state lines makes the figure too busy and difficult to read.
-However, we have added labels to the 4 Jalisco populations (La Mesa (MSA), San Lorenzo (SLO), Ejutla A (EjuA), and Ejutla B (EjuB)).}
+However, we have added labels to the 4 focal Jalisco populations (La Mesa (MSA), San Lorenzo (SLO), Ejutla A (EjuA), and Ejutla B (EjuB)).}
 

### 2015-02-19 manuscript.tex
Edits
Edits to Figure 1 (Map), response, text, supplementary tables
Edits
Edits to cartoon figure; text; response

@@
 \textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
-Numbers of individuals sampled ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
+Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
 DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
@@
 Size of circles reflects number of individuals sampled.
+The four Jalisco populations are labeled: San Lorenzo (SLO), La Mesa (MSA), Ejutla A (EjuA), and Ejutla B (EjuB0.
 The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.}

### 2015-02-19 PeerJ_trackchanges.tex
Figure edits
Renumbered figures; issues with them counting properly in the legends
still

@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.
-S1 and Fig.
-S2, See Supplemental Materials with the online version of this article).
+S1 and \track{Fig.
+1}, See Supplemental Materials with the online version of this article).
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Table S2, See Supplemental Materials with the online version of this article).
@@
 
-%Figure 1 Map 
+%Figure 1 Locus Cartoon
+%-------------------------------------------------------------------
+\begin{figure*}[!t]
+  \begin{center}
+   \includegraphics[width=180mm]{Fig1LocusCartoon.png}
+   \flushleft{Figure 1.
+Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
+  \end{center}
+\label{Fig1Cartoon}
+\end{figure*}
+%-------------------------------------------------------------------
+
+\clearpage
+
+%Figure 2 Map 
 %--------------------------------------------
@@
 \begin{center}
-   \includegraphics[width=0.8\textwidth]{Fig1MapBalsas.png}
+   \includegraphics[width=0.8\textwidth]{Fig2MapBalsas.png}
    \end{center}
@@
 
-% Figure 2. LD plots
+% Figure 3. LD plots
 %-------------------------------------------------------------------
@@
   \begin{center}
-   \includegraphics[width=180mm]{Fig2LDPlots.png}
+   \includegraphics[width=180mm]{Fig3LDPlots.png}
    \end{center}
@@
 The upper triangle above the black diagonal is colored based on the $r^{2}$ value between SNPs while the bottom triangle is colored based on p-value for the corresponding  $r^{2}$ value.}
-\label{Fig2LD}
+\label{Fig3LD}
 \end{figure*}
@@
 
-% Figure 3 STRUCTURE Plot 4 Pops
+% Figure 4 STRUCTURE Plot 4 Pops
 %-------------------------------------------------------------------
@@
   \begin{center}
-   \includegraphics[width=160mm]{Fig3Structure.png}
+   \includegraphics[width=160mm]{Fig4Structure.png}
      \end{center}
     \caption{STRUCTURE assignment to maize across a section of chromosome 1. The dotted lines mark the beginning of the sequenced region 66kb upstream (Region 2) and the end of the \emph{tb1} ORF.}
-\label{Fig3Structure}
+\label{Fig4Structure}
 \end{figure*}
@@
 
-% Figure 4 Phenotyping BoxPlots
+% Figure 5 Phenotyping BoxPlots
 %-------------------------------------------------------------------
@@
   \begin{center}
-   \includegraphics[width=160mm]{Figure4BoxplotPheno1.png}
+   \includegraphics[width=160mm]{Figure5BoxplotPheno1.png}
    \end{center}
@@
 Outliers are displayed as dots, the maximum value excluding outliers is shown with the top whisker, while the minimum excluding outliers is shown with the bottom whisker.}
-\label{Fig4Boxplots}
+\label{Fig5Boxplots}
 \end{figure*}

### 2015-02-19 PeerJ_trackchanges.tex
Edits
figure numbers

@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.
-S1 and \track{Fig.
-1}, See Supplemental Materials with the online version of this article).
+S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article).
 Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Table S2, See Supplemental Materials with the online version of this article).
@@
 Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
-\ref{Fig1Map}).
+\ref{Fig2Map}).
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes.
@@
 If the \emph{Hopscotch} is found in these populations due to recent introgression from maize we would expect to find large blocks of linked markers near this element.
-We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.~\ref{Fig2LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
+We find no evidence of elevated linkage disequilibrium between the \emph{Hopscotch} and SNPs surrounding the \emph{tb1} region in our resequenced populations (Fig.\ref{Fig3LD}), and $r^{2}$ in the \emph{tb1} region  does not differ significantly between populations with (average $r^{2}$ of 0.085) and without (average $r^{2}=0.082$) the \emph{Hopscotch} insertion.
 In fact, average $r^{2}$ is lower in the \emph{tb1} region ($r^{2}=0.056$) than across the rest of chromosome 1 ($r^{2}=0.083$; Table \ref{Table3R2}).
@@
 In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) and much below the chromosome-wide average (0.20; Table \ref{Table4Q}; Fig.
-\ref{Fig3Structure}).
+\ref{Fig4Structure}).
 
@@
 After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig.
-\ref{Fig4Boxplots}), tiller number, or culm diameter.
+\ref{Fig5Boxplots}), tiller number, or culm diameter.
 Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
@@
 Moreover, there is no signature of elevated LD in the \emph{tb1} region relative to the rest of chromosome 1, and Bayesian assignment to a maize cluster in this region is both low and below the chromosome-wide average (Fig.
-\ref{Fig3Structure}, Table \ref{Table4Q}).
+\ref{Fig4Structure}, Table \ref{Table4Q}).
 Together, these data point to an explanation other than recent introgression for the high observed frequency of \emph{Hopscotch} in the Jalisco cluster of our \emph{parviglumis} populations.
@@
 The Balsas River is shown, as the Balsas River Basin is believed to be the center of domestication of maize.}
-\label{Fig1Map}
+\label{Fig2Map}
 \end{figure*}

### 2015-02-19 PeerJResponse.tex
Edit to response letter
Edits to response

@@
 4) It would be helpful to include Fig. S1 in the main text of the paper and to label the two sequenced regions (Region 1 and 2) in the figure itself (not just in the caption). \\
-\lev{Okay can do this}
-\mbh{yep, easy to do and this is an online publication without figure charges so why not!}
+
+\res{Thank you for the suggestion. We have included Fig. S1 in the main manuscript, now labelled Fig. 1, and have added labels for the two sequenced regions, Region 1 and Region 2.}
+
+\lev{Hi guys, I'm having trouble getting the figures to renumber in order and haven't been able to figure out what's going on, if either of you knows how to fix or has a suggestion.}
 

### 2015-02-19 PeerJResponse.tex
Edits
Edits to methods

@@
 7) Line 118. Figure S1 should be cited here, not Table S1. \\
-\lev{Okay can fix}
+
+\res{We are referring to Table S1 and Table S2.
+We have relocated the table references to be more clear, page 8, lines 198-200: ``The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article).}
 

### 2015-02-19 PeerJ_trackchanges.tex
Edits
Edits to methods

@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
-Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened \track{(Table S1 and Table S2, See Supplemental Materials with the online version of this article).}Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig2Map}).

### 2015-02-19 manuscript.tex
Edits
Edits to methods

@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened.
-Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
-Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article).Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig1Map}).

### 2015-02-20 PeerJResponse.tex
Edits to response
Edits to response, reviewer 1

@@
 \lev{I believe that she phased it since it was the exact thing that she used for her BayEnv stuff}
+
+\res{Data were previously phased using fastPHASE (\lef{cite tanja and fastPHASE paper}).
+We have added clarifying statements to the text of this section, pages X, lines X-X.
+Figure 4 is showing maize population assignment for adjacent 5 kb windows.
+The STRUCTURE analysis was performed for the entirety of chromosome 1; however, the figure only shows a zoomed in portion to highlight the section of chromosome 1 that contains the \emph{tb1} region.
+The results in Table 4 reflect the entire chromosome.}
 

### 2015-02-20 PeerJResponse.tex
Edits
Edits to text and response

@@
 
-\lev{Okay can mention}
+\res{We have editied the text to provide further clarification, page 8, lines 211-213: ``o investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
 
@@
 
-\lev{will check}
+\res{We have removed this from the citation.}
 

### 2015-02-20 PeerJ_trackchanges.tex
Edits
Edits to text and response

@@
 
-\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations.
+\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations \track{from the Jalisco region}.
 After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
 For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}, Table S2, See Supplemental Materials with the online version of this article).
+Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}).
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).

### 2015-02-20 PeerJ_trackchanges.tex
Edits for reviewer 1
Edits to response and track changes text

@@
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive in all populations except La Mesa, indicating an excess of low frequency variants in this population (Table \ref{Table2Diversity}).
+Tajima's D is positive \track{in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants} (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).

### 2015-02-20 PeerJResponse.tex
Edits for reviewer 1
Edits to response and track changes text
Edits reviewer 1
Edits to response

@@
 
-\lev{Okay reword.}
+\res{Thank you for the suggestion.
+We have reworded this sentence, page 9, lines 221-221: ``in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants."}
 
-16) Lines 249-250. 'average r2 is slightly lower in the tb1 region..' Also, include these mean values as a row in Table 3. \\
+16) Lines 249-250. 'average r2 is slightly lower in the tb1 region..' Also, include these mean values as a row in Table 3.\\
 
-\lev{Do you guys want to do this?}
-
-\mbh{simple enough to make the reviewer happy here and just add one more row to the table}
+\res{The numbers reflected in Table 3 are averages.}
 

### 2015-02-20 PeerJResponse.tex
Edits
questions for you two with respect to the density stuff from Doebley
and Lukens as it compares to our design

@@
 
+\lev{So I have checked this.
+Their experimental design was with maize (don't know if that matters) but in low density plants were placed 36'' apart within rows and rows were 30'' apart.
+In the high density design plants were 12'' apart within rows and 30'' apart between rows.
+On each side of the row for the high density Helianthus annus was palnted 12'' away from seedlings to provide extra canopy cover.}
+
+\lev{My guess is that later on our plants were crowded, but early on in the measurements they should not have been shading each other, and tillering then would not have been affected in the early measurements.
+What do you guys think?}
+
 12) Line 223. Indicate here that these four populations are parviglumis and are from Jalisco state. \\
@@
 
-\lev{I have no strong feelings about this.}
-\mbh{fine with me to move to supplement}
+\res{We appreciate your suggestion but have decided to leave Figure 3 in the main text, as we feel it provides a nice visual representation for the numbers and take home message presented from the STRUCTURE analysis.}
 
@@
 
-\lev{Okay we can maybe take these statements out?
-It would seem that none of the reviewers have ever been happy with what we have had to say with respect to mex vs parv, or unequal sampling.
-we have discussed why sampling is unequal and why there was no phenotyping for mex.}
-
-\mbh{my vote would be to remove the comparisons between \emph{parv} and \emph{mex} but to still note the high frequency in Jalisco}
-\jri{ok either way here}
-\lev{I vote remove since everyone has seemed unhappy with this}
+\res{We have removed the comparisons made about the difference in frequencies of the \emph{Hopscotch} allele between ssp. \emph{parviglumis} and ssp. \emph{mexicana}.}
 

### 2015-02-20 PeerJResponse.tex
Edits for reviewer 1
Edits to include discussion of Doebley & Lukens 1999, with respect to
plant density for our phenotyping

@@
 
-\lev{No idea off the top of my head. Good point, can look up and mention}
-
-\mbh{yep, good point to check on...I'd add some discussion on this, particularly if they saw very low tillering at this density}
-
 \lev{So I have checked this.
@@
 On each side of the row for the high density Helianthus annus was palnted 12'' away from seedlings to provide extra canopy cover.}
+
+\res{We have edited our discussion to reflect a brief discussion of the work by \citet{DoebleyLukens1999}, page 12, lines 344-346: ``We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
+It is possible that the planting density of our seedlings (plants spaced 12" apart) was too high and led to an overall increase in tillering as previously seen in \citet{LukensDoebley1999}, which would have affected our measurements of tillering index."}
 

### 2015-02-20 PeerJ_trackchanges.tex
Edits
questions for you two with respect to the density stuff from Doebley
and Lukens as it compares to our design
Edits for reviewer 1
Edits to include discussion of Doebley & Lukens 1999, with respect to
plant density for our phenotyping

@@
 While our findings are consistent with \citet{Studer2011} in that we identify the \emph{Hopscotch} allele segregating in teosinte, we find it at higher frequency than previously suggested.
-The \emph{Hopscotch} allele is more prevalent in \emph{parviglumis} than in \emph{mexicana} in our sample (Table S1, See Supplemental Materials with the online version of this article), suggesting a different history of the allele amongst teosinte subspecies.
 Moreover, many of our \emph{parviglumis} populations with a high frequency of the \emph{Hopscotch} allele fall in the Jalisco cluster identified by \citet{Fukunaga2005}, and further distinguish this region from the Balsas River Basin where maize was domesticated \citep{Matsuoka2002}.
@@
 We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
-One interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
+\track{It is possible that the planting density of our seedlings (plants spaced 12" apart) was too high and led to an overall increase in tillering as previously seen in \citet{LukensDoebley1999}, which would have affected our measurements of tillering index.}
+
+An alternative interpretation of this result would be that the \emph{Hopscotch} controls tillering in maize \citep{Studer2011}, but tillering in teosinte is affected by variation at other loci.
 Consistent with this interpretation, \emph{tb1} is thought to be part of a complex pathway controlling branching, tillering and other phenotypic traits \citep{KebromBrutnell2007, Clark2006}.

### 2015-02-25 PeerJResponse.tex
Edits for PeerJ resubmission
Edits to introduction and response

@@
 
-\lev{So I have checked this.
-Their experimental design was with maize (don't know if that matters) but in low density plants were placed 36'' apart within rows and rows were 30'' apart.
-In the high density design plants were 12'' apart within rows and 30'' apart between rows.
-On each side of the row for the high density Helianthus annus was palnted 12'' away from seedlings to provide extra canopy cover.}
-
 \res{We have edited our discussion to reflect a brief discussion of the work by \citet{DoebleyLukens1999}, page 12, lines 344-346: ``We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
@@
 
-\lev{I don't think this is necessary.
-Again, everyone has seemed to want something different in the intro.
-Last person wanted more maize/shade avoidance/tb1.
-Perhaps we can add a couple of sentences in about mex vs parv or the population stuff.
-I would like to not have to rewrite the entire intro again.}
-
-\mbh{I might mention something briefly about \emph{parviglumis} being a ruderal species that can occur at high density (cite my \emph{Molecular Ecology} manuscript), but, no, I would not add a whole paragraph on the ecologies of the teosintes.}
-\jri{this sounds OK. a wee bit of self-citation there hufford?}
+\res{We have added a few sentences discussing population density of distribution of both \emph{Zea mays} ssp.
+\emph{parviglumis} and \emph{Zea mays} ssp.
+\emph{mexicana} to the introduction.
+This addition is reflected in the text on page 2, lines 22-26: ``Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp.
+\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}.
+\emph{parviglumis}, with \emph{parviglumis} abundant in the lowlands of Mexico and \emph{mexicana} abundant in the highlands.
+Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}."}
 

### 2015-02-25 PeerJ_trackchanges.tex
Edits for PeerJ resubmission
Edits to introduction and response

@@
 Maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-Additionally, large, naturally-occurring populations of both \emph{parviglumis} and the highland teosinte \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, and genetic diversity of these taxa is estimated to be high \citep{Ross-Ibarra2009}.
+\track{Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp.
+\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}.
+\emph{parviglumis}, with \emph{parviglumis} abundant in the lowlands of Mexico and \emph{mexicana} abundant in the highlands.
+Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}}.
 

### 2015-03-01 PeerJResponse.tex
Edits
Edits for reviewer 2, response, text, refs

@@
 
-\lev{Okay I can check to make sure citations and data stuff from tanja are clear}
+\res{We have edited the reference to Table 1 and have edited the text to be more clear about which data were used where and how they were obtained: page 5, lines 97-100: ``We used teosinte (ssp.
+\emph{parviglumis} and ssp.
+\emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
+\track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using fastPHASE \citep{Scheet2006}.
+Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}."} and page 9, lines 230-234: ``We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}."}
 

### 2015-03-01 PeerJ_trackchanges.tex
Edits
Edits for reviewer 2, response, text, refs

@@
 \emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
-SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA).
-Environmental data were obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
-Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+\track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using fastPHASE \citep{Scheet2006}.
+Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
 Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.
@@
 Using our \emph{Hopscotch} genotyping, we calculated differentiation between populations (F\textsubscript{ST}) and subspecies (F\textsubscript{CT}) for populations in which we sampled sixteen or more chromosomes.
-We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) are similar to genome-wide estimates from previous studies (\citealt{Pyhajarvi2013}; Table \ref{Table1Fst}).
+We found that F\textsubscript{CT} = 0, and levels of F\textsubscript{ST} among populations within each subspecies (0.22) and among all populations (0.23) (Table \ref{Table1Fst}) are similar to genome-wide estimates from previous studies \citealt{Pyhajarvi2013}.
 Although we found large variation in \emph{Hopscotch} allele frequency among our populations, BayEnv analysis did not indicate a correlation between the \emph{Hopscotch} insertion and environmental variables (all Bayes Factors $<$ 1).
@@
 
-\textbf{Evidence of introgression---}The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
+\textbf{Evidence of introgression---} \track{We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}.} The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.

### 2015-03-02 PeerJResponse.tex
Edits
Edits for reviewer 2; text, response

@@
 3) In the Results section, some titles like "Genotyping" "Sequencing" or "Phenotyping" are too vague, and sound more like Materials and Methods than to Results.
-They should be modified in order to highlight what question is addressed.
- \lev{I like our sections...}
- \mbh{easy enough to make these more hypothesis oriented; good idea to do as many of these simple things as possible to show we're making a strong effort toward revisions}
+They should be modified in order to highlight what question is addressed.\\
 
-Paragraph "Neighbor joining trees" on lines 235-239 should be moved to the section "Evidence of introgression"\\
+\res{Thank you for the suggestion. We have edited our headings to be more detailed.}
+
+5) Paragraph "Neighbor joining trees" on lines 235-239 should be moved to the section "Evidence of introgression"\\
+
+\res{We have moved this to the ``Evidence of Introgression" section.}
 
@@
 
-4) The biological question is to characterize whether the Hopscotch transposable element at tb1 (which enhances expression of tb1) plays a role in the ecology of teosinte, especially in high-density populations.
+6) The biological question is to characterize whether the Hopscotch transposable element at tb1 (which enhances expression of tb1) plays a role in the ecology of teosinte, especially in high-density populations.
 To do so, they characterize the distribution of this element in parviglumis, mexicana and maize landraces and they examine the phenotypic effects of the insertion in parviglumis.

### 2015-03-02 PeerJ_trackchanges.tex
Edits
Edits for reviewer 2; text, response

@@
 
-\textbf{Genotyping---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened \track{(Table S1 and Table S2, See Supplemental Materials with the online version of this article).}Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\textbf{Genotyping for the \emph{Hopscotch} insertion---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened \track{(Table S1 and Table S2, See Supplemental Materials with the online version of this article).}Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig2Map}).
@@
 
-\textbf{Sequencing---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations \track{from the Jalisco region}.
+\textbf{Sequencing upstream regions of the \emph{tb1} ORF---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations \track{from the Jalisco region}.
 After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
@@
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
-Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Fig.
-S3, See Supplemental Materials with the online version of this article); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
 
-\textbf{Evidence of introgression---} \track{We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}.} The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
+\textbf{Evidence of introgression around the \emph{tb1} region---} \track{We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}.} The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
@@
 
+Neighbor joining trees of our sequence data and data from the teosinte inbred lines (TILs; data from Maize HapMapV2, \citealt{Chia2012}) do not reveal any clear clustering pattern with respect to population or \emph{Hopscotch} genotype (Fig.
+S3, See Supplemental Materials with the online version of this article); individuals within our sample that have the \emph{Hopscotch} insertion do not group with the teosinte inbred lines or domesticated maize that have the \emph{Hopscotch} insertion.
 The lack of clustering of \emph{Hopscotch} genotypes in our NJ tree as well as the lack of LD around \emph{tb1} do not support the hypothesis that the \emph{Hopscotch} insertion in these populations of \emph{parviglumis} is the result of recent introgression.
@@
 
-\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
+\textbf{Phenotyping of \emph{Zea mays} ssp.
+\emph{parviglumis}---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping.
 We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion.

### 2015-03-02 PeerJResponse.tex
Edits
edits to text and response for Reviewer 2

@@
 
-\lev{Simply put it is what we had available at the time. There was no specific choosing or decision about number of individuals per accession}
-\mbh{we'll have to diplomatically state this...we sampled all teosinte available to us and while this didn't allow for even sample sizes across populations, it did allow us to calculate both frequency in a subset of populations and geographic breadth of the \emph{Hopscotch} across a large number of independent sampling locations}
+\res{We sampled all individuals and populations that were available to us, which while it did not allow for even sample sizes across populations, did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as the geographic distribution of the \emph{Hopscotch} across many independent sampling locations.
+We have added a clarifying statement to the text regarding our sampling: page 4, lines 63-68: ``We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
+Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites."}
 
@@
 
-\lev{Okay. Is there somewhere we can get all of this info.? I don't believe I have access to the database any more}
-\mbh{Yes, we need to include this info.; that's a fair request}
-\lev{Do either of you know where I can find this info?}
+\lev{I don't have access to this info. Can one of you guys help?}
 
@@
 
-\lev{Why what?
-Why two products?
-or why did we do it that way?
-I think it is clearly stated that we can test for presence/absence/hets and pcr bias with sometimes a large fragment being difficult to amplify.
-The location of the primers is all shown in the cartoon figure and stated in the text.
-two primers outside of LTR one primer within.}
+\res{There are three primers that make up the two primer sets used to genotype individuals for the \emph{Hopscotch} element.
+The primers are HopF, located upstream of the element; IntHopR, located within the LTR; and HopR, located downstream of the element.
+HopF and HopR are used as one reaction, resulting in either a 5kb product (the whole element) or a small band ~300bp (sequence on eitheir end of the element).
+HopF and IntHopR are used to amplify a portion of th element (~1.1kb).
+Smaller bands preferentially amplify over larger bands when in the same reaction, so we chose to add the HopF/IntHopR PCR reaction as a way to get around PCR bias when amplifying both a larger (5kb) and smaller (300bp) band at the same time during the HopF/HopR reaction.
+A plant homozygous for the \emph{Hopscotch} insertion will produce one 5kb band in the HopF/HopR reaction and one 1.1kb band in the HopF/HopIntR reaction.
+A heterozygous individuals will have two bands for the HopF/HopR reaction (one 5kb and one 300bp) as well as a 1.1kb band for the HopF/HopIntRR reaction, totaling 3 bands.
+The two reactions were run in separate wells on an agarose gel to allow for scoring.}
 
-\mbh{if there's any way to be even slightly more clear, I'd edit just so you can say here that you've edited for clarity}
+\res{We have added some clarifying text to the manuscript: page 4, lines 75-86: ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR).
+Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
+HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.
+S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article).
+Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is.
+Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR.
+Heterozygotes for the \emph{Hopscotch} allele show three bands total; both a 300bp band and a 5kb band for the HopF/HopR PCR and a 1.1Kb band for the HopF/HopIntR PCR (Table S2, See Supplemental Materials with the online version of this article)."}
 
@@
 
-\lev{What? It is in the LTR!!!! Gah! That's the whole point!}
+\res{The HopIntR primer is within the LTR element and was designed off of previously sequence data provided by John Doebley.}
 

### 2015-03-02 PeerJ_trackchanges.tex
Edits
edits to text and response for Reviewer 2

@@
 
-\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+\textbf{Sampling and genotyping---}\track{We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
-DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
+Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites.} DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
 We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb).
-Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR).
+\track{Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR).
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
@@
 S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article).
-Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb, amplification of the entire element, and $\sim$1.1kb, amplification of part of the element), whereas heterozygotes show all three bands (Table S2, See Supplemental Materials with the online version of this article).
-Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
+Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is.
+Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR.
+Heterozygotes for the \emph{Hopscotch} allele show three bands total; both a 300bp band and a 5kb band for the HopF/HopR PCR and a 1.1Kb band for the HopF/HopIntR PCR (Table S2, See Supplemental Materials with the online version of this article).} Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
 We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.

### 2015-03-02 PeerJResponse.tex
Edits to responpse
Reviewer 2, comments

@@
 
+\lev{changed to be within manuscript. sizes added to figure legend}
+
 9) In suppl.
@@
 
-\lev{I think that we did initially to check.
-But all of the products match the appropriate and expected size.
-The only issue we had was suddenly the different size with the Pif, which we did check with sequencing.}
+\lev{The only sequence data we have was when we sequenced a bit of the Pif to make sure that was what it was.}
 
-\mbh{we need to check this.  can you go back through emails or to your lab notebook to find whether this was sequenced?}
-\jri{i don't remember ever checking this. we checked the pif by PCRing stuff John had already sequenced in which i identified the pif}
-\lev{I agree with Jeff. I believe bands matched with what was expected based on where we designed primers so we did not sequence them}
+\res{We did not sequence PCR products.
+We felt this was unnecessary based on the fact that primers were designed off of previous maize and teosinte sequence data.
+Furthermore, primers are unique to their location in the genome and do not BLAST with any confidence to any other location in the maize genome.
+PCRs produced bands of the expected sizes, which we felt was adequate.}
 

### 2015-03-02 PeerJResponse.tex
Edits
edits to text, response

@@
 
-\lev{My guess is that we can look this up on WorldClim if you guys think it is necessary.
-My guess is that microorganisms were not included in analysis of soil characteristics, that it was things like what compounds are in soil, etc.}
-
-\mbh{soil data are from the Harmonized world soil database that we cite and that Tanja also used in her manuscript.
-There will be further details on data collection there which we can briefly describe in our response and point the reviewer/editor to}
+\res{Soil data is from the Harmonized World Soil Database \url{http://webarchive.iiasa.ac.at/Research/LUC/External-World-soil-database/HTML/} and include measurements for ``organic Carbon, pH, water storage capacity, soil depth, cation exchange capacity of the soil and the clay fraction, total exchangeable nutrients, lime and gypsum contents, sodium exchange percentage, salinity, textural class and granulometry."}
 

### 2015-03-02 PeerJ_trackchanges.tex
Edits
edits to text, response

@@
 \track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using fastPHASE \citep{Scheet2006}.
-Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org} for \citet{Pyhajarvi2013}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
 Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.

### 2015-03-02 PeerJResponse.tex
Edits
edits to response, text, for reviewer 2

@@
 
-\lev{i will check this but i believe mexicana not included because we didn't discuss them with respect to the high frequency stuff.
-I do not have arcgis any more and would prefer to not completely redo this figure.
-I am happy to put names on pops though.
-I don't understand comments about 'left hand side' stuff.
-Yes, some of the populations are close together when you are that zoomed out on a map.
-There is no way to have them not all coming from the same place in some cases unless the map was so gigantic to take up an entire page.}
-
 \mbh{the reviewer is indicating they are surprised that there are such marked differences in Hop frequency in such a small area in the populations on the lefthand side of the map.
 You can cite my dissertation and the Tiffin and Moeller paper as finding that populations in this area were highly differentiated and that populations with Hop were much more closely related to each other than they were to pops without Hop...just some very different histories in these populations and little gene flow between them}
+
+\res{We only included \emph{parviglumis} in this graph because we wanted to highlight the difference in frequency within the Jalisco region, even between populations that are geographically very close to each other.
+Reviewer 2 is correct in stating that there is a large difference in the frequency of the \emph{Hopscotch} allele within 25 miles.
+Previous studies by \citet{Hufford2010; Moeller01072007} have shown that populations in this area are highly differentiated and likely have different evolutionary histories.}
+
+\res{We have edited the figure to include the names of the four populations in the Jalisco region that we intend to highlight: La Mesa, San Lorenzo, Ejulta A, and Ejutla B.}
 

### 2015-03-05 PeerJResponse.tex
Edits
response to review; reviewer 2
Edits
Edits to last of reviewer 2 comments regarding greenhouse vs. field

@@
 
-\lev{Sure greenhouse is different than outside, but we can't control for variation outside whereas we can in a greenhouse.
-I don't understand their point.
-Greenhouse is best we could do here since they wouldn't all fit in a growth chamber.}
-
-\mbh{I'd point out that the effect of \emph{tb1} on tillering has been verified a number of times in a number of environments (probably in a greenhouse setting) and cite studies like Briggs, etc...}
+\res{The effects of \emph{tb1} on tillering in maize have been verified a number of times in different environments \citet{Briggs2007, Clark2006, StuderDoebley2012, Studer2011}.
+We feel that any effect of growing teosinte in a greenhouse setting as opposed to as a field trial would not be so large as to mask the effects of \emph{tb1} on tillering.}
 

### 2015-03-06 Supplement_mbh.tex
Edits
Edits to supplement regarding gel image.

@@
   \begin{center}
-   \includegraphics[width=100mm]{FigS2GelPeerJ.png}
-   \flushleft{Figure S2: Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+   \includegraphics[width=100mm]{FigS2GelPeerJ_pretty.png}
+   \flushleft{Figure S2: Agarose gel image of amplification products for genotyping of the \emph{Hopscotch} element.
+Lanes 1 (HopF/HopR; 5kb band) and 2 (HopF/HopIntR; 1.1kb) are the products for one individual that is homozygous for the element; 3 (HopF/HopR; 5kb band) & 4 (HopF/HopIntR; 1.1kb) are also the products of an individual that is homozygous for the element; and 5 (HopF/HopR; 300bp) & 6 (HopF/HopIntR; N/A)  are the products of an individual that is homozygous for the teosinte (lacking the \emph{Hopscotch}) allele.}
   \end{center}

### 2015-03-06 PeerJResponse.tex
Edits
Edits to response concerning phasing of SNPs

@@
 Sequencing of PCR products was never performed as primer combinations for each of the 2 PCRs gave us the expected sized bands.
-Primers were designed off of previously sequenced teosinte and maize sequences \lev{these were from John Doebley, do we need to say this?
-I don' think they were published ever} and were initially tested in both maize and teosinte for validity.
-We have clarified this point in text \lev{text, line numbers here} as well added additional labelling to the gel image (Figure 2 \lev{check to make sure figure number doesn't change during edits})}.
+Primers were designed off of previously sequenced teosinte and maize sequences provided by John Doebley, and were initially tested in both maize and teosinte for validity.
+We have clarified this point in text \lev{text, line numbers here} as well added additional labelling to the gel image (Figure 2).}
 
@@
 5) Fig. S2. Annotate this figure to show the size (bp) of each band in the 1 kb ladder. The caption should also define the abbreviations used in the labels. \\
-\lev{this will be too much. I can label the brightest band (1kb) and the top and bottom bands}
-\mbh{I'd label as many as it takes so that it's obvious what intervening band sizes are.  We'll want to define the abbreviations too}
-\lev{usually the bottom, middle brightest and top are labelled so I will do that}
+
+\res{We have edited this figure to be more clear.}
 
@@
 
-\lev{Maybe in the text I can clarify total bands for individuals.
-I.e.
-if an individual is Hop/Hop, it will have a total of two bands that are split between the two PCRs, whereas if you are a heterozygote you will have 3 split between two PCRs.
-I wonder if we should ditch this figure entirely and I can draw up a cartoon one?}
-\mbh{since reviewer 2 is also concerned about the PCR and this is the only thing the editor really mentioned, I think we need to be pretty thorough here.
-Ideally we would have a Hop/Hop homozygote, a Hop/No Hop heterozygote, and a No Hop/No Hop homozygote in the figure.
-We don't need the Pif since this really isn't mentioned in the manuscript.
-Do you have gel figures showing both homozygotes and the heterozygote banding patterns?
-I'm afraid a cartoon would send the message that we don't have clear PCR results}
-\jri{agree with the huff}
-\lev{I do not have gel figures.
-I am guessing the ones that had been saved online in Jeff's lab have been tossed.
-I can check my lab notebook and see if there is something to scan, otherwise there will not be a figure with all of those genotypes that we can give to them.}
+\res{We have edited Fig. S2 to be more clear and include appropriate labels and explanations.}
 
@@
 
-\lev{yes SNPs within 5KB window are a haplotype.
-I don't know about the diploid/phasing stuff.
-Yes to maize assignment.
-Point of figure is that low assignment to maize in populations with high frequency of Hopscotch.
-Am plotting the assigment value across a region of chromosome 1.
-I showed a region and not entire chromosome to be able to highlight the position of tb1 and so that you can see the difference in assignment values.}
-
-\mbh{methods say that you used phased data from Tanja's manuscript.
-Did she give you phased data or did you phase this yourself with fastPHASE?
-Either way, probably best to mention the data were phased with fastPHASE and add a citation to the methods paper for this}
-
-\lev{I believe that she phased it since it was the exact thing that she used for her BayEnv stuff}
-
-\res{Data were previously phased using fastPHASE (\lef{cite tanja and fastPHASE paper}).
-We have added clarifying statements to the text of this section, pages X, lines X-X.
+\res{Data were previously phased using fastPHASE (\citet{Pyhajarvi2013, Scheet2006}).
+We have added clarifying statements to the text of this section, pages 5, lines 104-105.
 Figure 4 is showing maize population assignment for adjacent 5 kb windows.

### 2015-03-06 PeerJ_trackchanges.tex
Comments and text, figures
Edits to locus figure. Weird box around it that we need to get rid of

@@
   \begin{center}
-   \includegraphics[width=180mm]{Fig1LocusCartoon.png}
+   \includegraphics[width=180mm]{Fig1Cartoon.png}
    \flushleft{Figure 1.
 Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
-Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
+Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.
+The amplification product for the HopF/Hop R is either a 5kb band (an allele that includes the \emph{Hopscotch} insertion, or a 300bp band (an allele that does not include the \emph{Hopscotch} insertion.
+The HopF/HopIntR primer combination produces a 1.1kb band in individuals that have the \emph{Hopscotch} allele, and no band for individuals that lack the insertion, since the HopIntR primer sits within the LTR.}
   \end{center}

### 2015-03-06 PeerJResponse.tex
Comments and text, figures
Edits to locus figure. Weird box around it that we need to get rid of
Response edits
Edits to response concerning gel; unable to redo but have edited image
and legend & corresponding text to be more clear

@@
 
-\lev{Okay can do this. maybe will clear up their misunderstanding detailed above}
-
-\lev{changed to be within manuscript. sizes added to figure legend}
+\res{We have edited this figure to include primer names and added expected amplification sizes to the figure legend.}
 
@@
 
-\lev{I can relabel or clarify in the figure legend. I don't think that redoing a gel is possible at this point.}
+\res{We have edited the figure and the figure legend to be more clear.}
 
@@
 
-\lev{Redoing gel not possible at this point}
-
-\mbh{even if you have a better resolution image to include, that might help.  I'd also remove the pif lanes as possible as we never mention this}
+\res{It is not feasible for us to redo the gel at this point in time; however, we have editied the figure and the figure legend to be more clear.}
 
@@
 
-\lev{We can relabel but not redo gels at this point}
-\mbh{the PCR results were conclusive, but I'd make an effort to make the figure more clear with labeling or a higher resolution image and to clarify the description of the banding patterns of the genotypes in the text; my guess is we'll need to do a good job of this in particular if we want this manuscript in \emph{PeerJ}}
+\res{We have editied the figure and the legend and corresponding text reference to be more clear about the expected banding patterns for the 3 genotypes.}
 

### 2015-03-20 PeerJResponse.tex
Matt's edits to PeerJ Revision

@@
 \usepackage{multirow}
+\usepackage{hyperref}
+\usepackage[leftcaption]{sidecap} %package for adding side captions
+\sidecaptionvpos{figure}{c} %position side caption
+\usepackage{lineno} %package for adding line numbers
+\usepackage{lipsum}
+\usepackage{indentfirst}
+\usepackage{color}
+\usepackage{booktabs}
+\usepackage[pdftex]{graphicx}  
+\usepackage[labelsep=period]{caption} %package for formatting figure captions, separation between figure number and caption is period
+\usepackage{natbib} % package for formatting citations
+\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for formatting...journals often have these available but these can also be customized
+\usepackage{authblk}
+\usepackage[table]{xcolor} %for table line color alteration
+\usepackage{subscript}
 \renewcommand{\baselinestretch}{1.5}
-\usepackage{natbib} % package for formatting citations
-%\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for
 
@@
 
-\noindent Dear Dr.Vision and editorial staff at the \emph{PeerJ}, \\
+\noindent Dear Dr.Vision and editorial staff at \emph{PeerJ}, \\
 
@@
 
-\res{We have addressed the comments from Reviewer 2 regarding the validity of the PCR assays.
-Sequencing of PCR products was never performed as primer combinations for each of the 2 PCRs gave us the expected sized bands.
-Primers were designed off of previously sequenced teosinte and maize sequences provided by John Doebley, and were initially tested in both maize and teosinte for validity.
-We have clarified this point in text \lev{text, line numbers here} as well added additional labelling to the gel image (Figure 2).}
+\res{We have addressed comments from Reviewer 2 regarding the validity of the PCR assays.
+Sequencing of PCR products were not sequenced, but primer combinations for each of the 2 PCRs produced the expected sized bands.
+Primers were designed using previously sequenced teosinte and maize sequences provided by John Doebley, and were initially tested in both maize and teosinte for validity.
+We have clarified this point in the text \lev{text, line numbers here} and have added additional labeling to the gel image (Figure 2).}
 
@@
 
+\mbh{I don't think state lines would make the figure too busy.
+Is the issue that you no longer have access to ArcGIS, LV?
+Jeff, do you think it's worth trying to redo this in R or do you think Todd will be fine with this response?}
+
 4) It would be helpful to include Fig. S1 in the main text of the paper and to label the two sequenced regions (Region 1 and 2) in the figure itself (not just in the caption). \\
 
-\res{Thank you for the suggestion. We have included Fig. S1 in the main manuscript, now labelled Fig. 1, and have added labels for the two sequenced regions, Region 1 and Region 2.}
+\res{Thank you for the suggestion. We have included Fig. S1 in the main manuscript, now labeled Fig. 1, and have added labels for the two sequenced regions, Region 1 and Region 2.}
 
 \lev{Hi guys, I'm having trouble getting the figures to renumber in order and haven't been able to figure out what's going on, if either of you knows how to fix or has a suggestion.}
+
+\mbh{I fixed this}
 
@@
 
-\res{We have edited Fig. S2 to be more clear and include appropriate labels and explanations.}
+\res{We have edited Fig. S2 (now Fig. S1) to be more clear and to include appropriate labels and explanations.}
 
@@
 \res{We are referring to Table S1 and Table S2.
-We have relocated the table references to be more clear, page 8, lines 198-200: ``The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article).}
+We have edited our reference to the tables to be more clear, page 8, lines 198-200: ``The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article)."}
 
@@
 \lev{I do not have the files as this is something that Tom did with the sequence data that he gathered}
+\mbh{o.k., should we ask Tom for this? The only argument we can make for not doing this is that 100 bootstraps are commonly used, but this makes us sound pretty unaccommodating}
 
@@
 
-\res{Data were previously phased using fastPHASE (\citet{Pyhajarvi2013, Scheet2006}).
+\res{Data were previously phased using the program fastPHASE (\citet{Pyhajarvi2013, Scheet2006}).
 We have added clarifying statements to the text of this section, pages 5, lines 104-105.
-Figure 4 is showing maize population assignment for adjacent 5 kb windows.
+\mbh{please go through and recheck these line numbers just prior to submitting revisions} Figure 4 is showing maize population assignment for adjacent 5 kb windows.
 The STRUCTURE analysis was performed for the entirety of chromosome 1; however, the figure only shows a zoomed in portion to highlight the section of chromosome 1 that contains the \emph{tb1} region.
@@
 
-\res{We have edited our discussion to reflect a brief discussion of the work by \citet{DoebleyLukens1999}, page 12, lines 344-346: ``We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
-It is possible that the planting density of our seedlings (plants spaced 12" apart) was too high and led to an overall increase in tillering as previously seen in \citet{LukensDoebley1999}, which would have affected our measurements of tillering index."}
+\res{We have edited our discussion to include a brief mention of the work by \citet{LukensDoebley1999}, page 12, lines 344-346: ``We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
+It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}.
+This factor may have limited our capacity to observe variation in tillering index."}
 
@@
 
+\mbh{I read the Lukens and Doebley paper and I think what you had previously was backwards...higher density = less tillering, not more.
+If we saw less overall tillering due to high density, this could have compromised our ability to detect significant differences between genotypes.}
+
 12) Line 223. Indicate here that these four populations are parviglumis and are from Jalisco state. \\
 
-\res{We have editied the text to provide further clarification, page 8, lines 211-213: ``o investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
+\res{We have editied the text to provide further clarification, page 8, lines 211-213: ``To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
 
@@
 
-\res{We have removed this from the citation.}
+\res{We have removed reference to Table S2 from the citation.}
 
@@
 we mention they are at the tail of the empirical distribution from wright et al don't we?}
+\mbh{I agree this is a silly request and that we shouldn't include in the manuscript; yes, we indicate Taj.D for our pops is in the extreme tail based on comparison to both Wright et al.
+and Tiffin and Moeller}
 
@@
 
-\res{We appreciate your suggestion but have decided to leave Figure 3 in the main text, as we feel it provides a nice visual representation for the numbers and take home message presented from the STRUCTURE analysis.}
+\res{We appreciate this suggestion but have decided to leave Figure 3 in the main text, as we feel it provides a nice visual representation for the numbers and take home message presented from the STRUCTURE analysis.}
 
@@
 
-\res{We have added a few sentences discussing population density of distribution of both \emph{Zea mays} ssp.
+\res{We have added a few sentences discussing population density and distribution of both \emph{Zea mays} ssp.
 \emph{parviglumis} and \emph{Zea mays} ssp.
@@
 This addition is reflected in the text on page 2, lines 22-26: ``Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}.
-\emph{parviglumis}, with \emph{parviglumis} abundant in the lowlands of Mexico and \emph{mexicana} abundant in the highlands.
-Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}."}
+\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, with \emph{parviglumis} distributed in the lowlands of Mexico and \emph{mexicana} in the highlands.
+Furthermore, both \emph{parviglumis} and \emph{mexicana} can occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}."}
 
@@
 \emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
-\track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using fastPHASE \citep{Scheet2006}.
-Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}."} and page 9, lines 230-234: ``We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}."}
+SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}.
+Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}." and page 9, lines 230-234: ``We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}."}
 
@@
 \res{We sampled all individuals and populations that were available to us, which while it did not allow for even sample sizes across populations, did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as the geographic distribution of the \emph{Hopscotch} across many independent sampling locations.
-We have added a clarifying statement to the text regarding our sampling: page 4, lines 63-68: ``We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
-Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
+We have added a clarifying statement to the text regarding our sampling: page 4, lines 63-68: ``We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
 Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites."}
@@
 
+\mbh{Jeff, can you help here?}
+
 6) The authors write that the PCR amplification to investigate presence of the Hopscotch element leads to two amplification products: one for the entire element, (5kb) and one for amplification of part of the element only (1.1kb).
@@
 The primers are HopF, located upstream of the element; IntHopR, located within the LTR; and HopR, located downstream of the element.
-HopF and HopR are used as one reaction, resulting in either a 5kb product (the whole element) or a small band ~300bp (sequence on eitheir end of the element).
-HopF and IntHopR are used to amplify a portion of th element (~1.1kb).
+HopF and HopR are used as one reaction, resulting in either a 5kb product (the whole element) or a small band ~300bp (sequence on either end of the element).
+HopF and IntHopR are used to amplify a portion of the element (~1.1kb).
 Smaller bands preferentially amplify over larger bands when in the same reaction, so we chose to add the HopF/IntHopR PCR reaction as a way to get around PCR bias when amplifying both a larger (5kb) and smaller (300bp) band at the same time during the HopF/HopR reaction.
@@
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.
-S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article).
-Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is.
+1 and Fig.
+S1, See Supplemental Materials with the online version of this article).
+Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is located.
 Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR.
@@
 
-\res{The HopIntR primer is within the LTR element and was designed off of previously sequence data provided by John Doebley.}
+\res{The HopIntR primer is within the LTR element and was designed off of published sequence data provided by John Doebley.}
 
@@
 
-\res{It is not feasible for us to redo the gel at this point in time; however, we have editied the figure and the figure legend to be more clear.}
+\res{We have edited the figure and the figure legend to be more clear.  While the 300bp band is somewhat diffuse, we feel it is clear that there is a single band of the appropriate, expected size.}
 
@@
 
-\res{Soil data is from the Harmonized World Soil Database \url{http://webarchive.iiasa.ac.at/Research/LUC/External-World-soil-database/HTML/} and include measurements for ``organic Carbon, pH, water storage capacity, soil depth, cation exchange capacity of the soil and the clay fraction, total exchangeable nutrients, lime and gypsum contents, sodium exchange percentage, salinity, textural class and granulometry."}
+\res{Soil data are from the Harmonized World Soil Database \url{http://webarchive.iiasa.ac.at/Research/LUC/External-World-soil-database/HTML/} and include measurements for ``organic Carbon, pH, water storage capacity, soil depth, cation exchange capacity of the soil and the clay fraction, total exchangeable nutrients, lime and gypsum contents, sodium exchange percentage, salinity, textural class and granulometry."}
 
@@
 
-\mbh{the reviewer is indicating they are surprised that there are such marked differences in Hop frequency in such a small area in the populations on the lefthand side of the map.
-You can cite my dissertation and the Tiffin and Moeller paper as finding that populations in this area were highly differentiated and that populations with Hop were much more closely related to each other than they were to pops without Hop...just some very different histories in these populations and little gene flow between them}
-
 \res{We only included \emph{parviglumis} in this graph because we wanted to highlight the difference in frequency within the Jalisco region, even between populations that are geographically very close to each other.
 Reviewer 2 is correct in stating that there is a large difference in the frequency of the \emph{Hopscotch} allele within 25 miles.
-Previous studies by \citet{Hufford2010; Moeller01072007} have shown that populations in this area are highly differentiated and likely have different evolutionary histories.}
+Previous studies by \citet{Hufford2010, Moeller01072007} have shown that populations in this area are highly differentiated and likely have different evolutionary histories.}
 
@@
 
-\res{We have editied the figure and the legend and corresponding text reference to be more clear about the expected banding patterns for the 3 genotypes.}
+\res{We have edited the figure and the legend and corresponding text reference to be more clear about the expected banding patterns for the 3 genotypes.}
 
@@
 \res{The effects of \emph{tb1} on tillering in maize have been verified a number of times in different environments \citet{Briggs2007, Clark2006, StuderDoebley2012, Studer2011}.
-We feel that any effect of growing teosinte in a greenhouse setting as opposed to as a field trial would not be so large as to mask the effects of \emph{tb1} on tillering.}
+We feel that any effect of growing teosinte in a greenhouse setting as opposed to a a field trial would not be so large as to mask the effects of \emph{tb1} on tillering.}
 

### 2015-03-20 PeerJ_trackchanges.tex
Matt's edits to PeerJ Revision

@@
 \usepackage[pdftex]{graphicx}  
-\usepackage[labelsep=period]{caption} %package for formatting figure captions, separation between figure number and caption is period per AJB formatting requirements
+\usepackage[labelsep=period]{caption} %package for formatting figure captions, separation between figure number and caption is period
 \usepackage{natbib} % package for formatting citations
@@
 \renewcommand{\baselinestretch}{1.5}
-\renewcommand{\refname}{\centering{REFERENCES}} %changing title of references to Literature Cited as stipulated by AJB
+\renewcommand{\refname}{\centering{REFERENCES}}
 
@@
 \newcommand{\lev}[1]{\textcolor{magenta}{\emph{\normalsize #1}} } %same for laura vann in green
-\newcommand{\track}[1]{\textcolor{red}{\emph{\normalsize #1}} }
+\newcommand{\track}[1]{\textcolor{red}{\normalsize #1}}
 
@@
 \track{Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp.
-\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}.
-\emph{parviglumis}, with \emph{parviglumis} abundant in the lowlands of Mexico and \emph{mexicana} abundant in the highlands.
+\emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, with \emph{parviglumis} distributed in the lowlands of Mexico and \emph{mexicana} in the highlands.
 Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}}.
@@
 S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article).
-Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is.
+Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is located.
 Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR.
@@
 \emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
-\track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using fastPHASE \citep{Scheet2006}.
-Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org} for \citet{Pyhajarvi2013}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+\track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}.
+Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
 Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.
@@
 We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
-\track{It is possible that the planting density of our seedlings (plants spaced 12" apart) was too high and led to an overall increase in tillering as previously seen in \citet{LukensDoebley1999}, which would have affected our measurements of tillering index.}
+\track{It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}.
+This factor may have limited our capacity to observe variation in tillering index.}
 
@@
   \begin{center}
-   \includegraphics[width=180mm]{Fig1Cartoon.png}
-   \flushleft{Figure 1.
+   \includegraphics[width=150mm]{Fig1Cartoon.png}
+   \end{center}
+   \caption{Figure 1.
 Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
@@
 The HopF/HopIntR primer combination produces a 1.1kb band in individuals that have the \emph{Hopscotch} allele, and no band for individuals that lack the insertion, since the HopIntR primer sits within the LTR.}
-  \end{center}
 \label{Fig1Cartoon}

### 2015-03-23 PeerJResponse.tex
Edits
LV edits to response and text

@@
 Primers were designed using previously sequenced teosinte and maize sequences provided by John Doebley, and were initially tested in both maize and teosinte for validity.
-We have clarified this point in the text \lev{text, line numbers here} and have added additional labeling to the gel image (Figure 2).}
+We have clarified this point in the text page 4, lines 75-87 and have added additional labeling to the gel image (Figure 2).}
 
@@
 
-\res{We feel that the addition of state lines makes the figure too busy and difficult to read.
-However, we have added labels to the 4 focal Jalisco populations (La Mesa (MSA), San Lorenzo (SLO), Ejutla A (EjuA), and Ejutla B (EjuB)).}
-
-\mbh{I don't think state lines would make the figure too busy.
-Is the issue that you no longer have access to ArcGIS, LV?
-Jeff, do you think it's worth trying to redo this in R or do you think Todd will be fine with this response?}
+\res{We have edited this figure (now Figure 2) to include state lines and have added labels to the 4 focal Jalisco populations (La Mesa (MSA), San Lorenzo (SLO), Ejutla A (EjuA), and Ejutla B (EjuB)).}
 

### 2015-03-23 PeerJResponse.tex
Edits
LV edits to page & line numbers

@@
 
-\lev{Hi guys, I'm having trouble getting the figures to renumber in order and haven't been able to figure out what's going on, if either of you knows how to fix or has a suggestion.}
-
-\mbh{I fixed this}
-
 5) Fig. S2. Annotate this figure to show the size (bp) of each band in the 1 kb ladder. The caption should also define the abbreviations used in the labels. \\
@@
 \res{Data were previously phased using the program fastPHASE (\citet{Pyhajarvi2013, Scheet2006}).
-We have added clarifying statements to the text of this section, pages 5, lines 104-105.
-\mbh{please go through and recheck these line numbers just prior to submitting revisions} Figure 4 is showing maize population assignment for adjacent 5 kb windows.
+We have added clarifying statements to the text of this section, pages 5, lines 105-108.
+Figure 4 is showing maize population assignment for adjacent 5 kb windows.
 The STRUCTURE analysis was performed for the entirety of chromosome 1; however, the figure only shows a zoomed in portion to highlight the section of chromosome 1 that contains the \emph{tb1} region.
@@
 
-\res{We have edited our discussion to include a brief mention of the work by \citet{LukensDoebley1999}, page 12, lines 344-346: ``We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
+\res{We have edited our discussion to include a brief mention of the work by \citet{LukensDoebley1999}, page 13, lines 358-360: ``We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
 It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}.
@@
 
-\lev{My guess is that later on our plants were crowded, but early on in the measurements they should not have been shading each other, and tillering then would not have been affected in the early measurements.
-What do you guys think?}
-
-\mbh{I read the Lukens and Doebley paper and I think what you had previously was backwards...higher density = less tillering, not more.
-If we saw less overall tillering due to high density, this could have compromised our ability to detect significant differences between genotypes.}
-
 12) Line 223. Indicate here that these four populations are parviglumis and are from Jalisco state. \\
 
-\res{We have editied the text to provide further clarification, page 8, lines 211-213: ``To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
+\res{We have editied the text to provide further clarification, page 8, lines 222-225: ``To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
 

### 2015-03-23 PeerJResponse.tex
Edits
LV Edits; command line for BayEnv and STRUCTURE; edits to response

@@
 
-1) The reviewers comments seems to be largely in agreement and self-explanatory. Please be sure to address the comment regarding the validity of the PCR assays raised by Reviewer 2.
+1) The reviewers comments seems to be largely in agreement and self-explanatory. Please be sure to address the comment regarding the validity of the PCR assays raised by Reviewer 2.\\
 
@@
 \res{Thank you for the suggestion.
-We have reworded this sentence, page 9, lines 221-221: ``in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants."}
+We have reworded this sentence, page 9, lines 232-233: ``in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants."}
 
@@
 
-\res{We have edited the reference to Table 1 and have edited the text to be more clear about which data were used where and how they were obtained: page 5, lines 97-100: ``We used teosinte (ssp.
+\res{We have edited the reference to Table 1 and have edited the text to be more clear about which data were used where and how they were obtained: page 5, lines 105-108: ``We used teosinte (ssp.
 \emph{parviglumis} and ssp.
@@
 SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}.
-Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}." and page 9, lines 230-234: ``We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}."}
+Environmental data were previously obtained for \citet{Pyhajarvi2013} from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}." and page 9, lines 237-239: ``We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}."}
 
@@
 \res{We sampled all individuals and populations that were available to us, which while it did not allow for even sample sizes across populations, did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as the geographic distribution of the \emph{Hopscotch} across many independent sampling locations.
-We have added a clarifying statement to the text regarding our sampling: page 4, lines 63-68: ``We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+We have added a clarifying statement to the text regarding our sampling: page 4, lines 63-71: ``We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
@@
 
-\res{We have added some clarifying text to the manuscript: page 4, lines 75-86: ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR).
+\res{We have added some clarifying text to the manuscript: page 4, lines 75-87: ``Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR).
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
@@
 11) It is not stated whether PCR products were sequenced to check for correct amplification. Were some of the PCR products sequenced? If not, this should be done.\\
-
-\lev{The only sequence data we have was when we sequenced a bit of the Pif to make sure that was what it was.}
 

### 2015-03-23 chunkLV.pl
Edits
LV edits, scripts for STRUCTURE — modified from MBH

@@ -1,0 +1,55 @@
+#!/usr/bin/perl
+
+use strict;
+use warnings;
+
+my @link_set;  # A two-dimensional array
+my $output = ''; # Our final output string - you can print this for redirection 
+                 # on the command line, or do something later in the program 
+                 # to print it to a file
+
+open FILE, 'parv.SLO.trans.pos'; # change to my filename
+while(<FILE>) {
+    chomp;
+    my @split_line = split(/\t/);
+    # If @link_set is empty (i.e. processing first line) or if the position 
+    # for the current line is less than 5000 from the start of the current 
+    # chunk ($link_set[0][0]), add the array on to the end of our array-of-arrays
+    if ( (scalar(@link_set) == 0) or (($split_line[0] - 5000) < $link_set[-1][0])) {
+        push(@link_set, [@split_line]);
+    } else {
+        # The current line is the first line of the next chunk
+        # Add the position from the previous line to our output
+        # Note I think this is the behavior we want if this happens to be the 
+        # last line (i.e. it will 
+        $output .= "$link_set[-1][0]\t";
+        foreach my $cindex (1..$#{$link_set[0]}) {
+            foreach my $rindex (0..$#link_set) {
+                # For each column (starting at index 1), iterate through all 
+                # the rows and add that to output
+                $output .= $link_set[$rindex][$cindex]
+            }
+            # After iterating through the rows for each column, add a tab
+            $output .= "\t";
+        }
+        $output .= "\n";
+        # Add the current line to @link_set
+        @link_set = [@split_line];
+    }
+}
+# And add the last chunk to the $output
+$output .= "$link_set[-1][0]\t";
+foreach my $cindex (1..$#{$link_set[0]}) {
+    foreach my $rindex (0..$#link_set) {
+        # For each column (starting at index 1), iterate through all 
+        # the rows and add that to output
+        $output .= $link_set[$rindex][$cindex]
+    }
+    # After iterating through the rows for each column, add a tab
+    $output .= "\t";
+}
+
+$output .= "\n";
+
+print $output;
+

### 2015-03-23 PeerJ_trackchanges.tex
Edits
LV edits to response and text
minor changes

@@
 
-\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the  Libsequence package \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
+\textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the Analysis package from the  Libsequence library \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
 To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
@@
 \begin{center}
-   \includegraphics[width=0.8\textwidth]{Fig2MapBalsas.png}
+   \includegraphics[width=0.8\textwidth]{Fig2MapStates.png}
    \end{center}

### 2015-03-23 Supplement_mbh.tex
edits to supplement

--- 2015-02-19 Supplement_mbh.tex
@@
 %-------------------------------------------------------------------
-\begin{figure*}[!t]
-  \begin{center}
-   \includegraphics[width=180mm]{FigS1LocusCartoon.png}
-   \flushleft{Figure S1: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
+%\begin{figure*}[!t]
+  %\begin{center}
+   %\includegraphics[width=180mm]{FigS1LocusCartoon.png}
+   %\flushleft{Figure S1: Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the \emph{Hopscotch} insertion (red).
 Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Region 1; within the 5' UTR, and Region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the \emph{Hopscotch} insertion.}
-  \end{center}
-\end{figure*}
+  %\end{center}
+%\end{figure*}
 %-------------------------------------------------------------------
@@
   \begin{center}
-   \includegraphics[width=100mm]{FigS2GelPeerJ.png}
-   \flushleft{Figure S2: Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
+   \includegraphics[width=100mm]{FigS2GelPeerJ_pretty.png}
+   \flushleft{Figure S1: Agarose gel image of amplification products for genotyping of the \emph{Hopscotch} element.
+Lanes 1 (HopF/HopR; 5kb band) and 2 (HopF/HopIntR; 1.1kb) are the products for one individual that is homozygous for the element; Lanes 3 (HopF/HopR; 5kb band) and 4 (HopF/HopIntR; 1.1kb) are also the products of an individual that is homozygous for the element; and Lanes 5 (HopF/HopR; 300bp) and 6 (HopF/HopIntR; N/A) are the products of an individual that is homozygous for the teosinte (lacking the \emph{Hopscotch}) allele.}
   \end{center}
@@
    \includegraphics[width=150mm]{FigS3NJtrees.png}
-    \flushleft{Figure S3: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}.
+    \flushleft{Figure S2: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}.
 Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).

### 2015-03-23 PeerJResponse.tex
edits to response

@@
 
-\noindent Dear Dr.Vision and editorial staff at \emph{PeerJ}, \\
+\noindent Dear Dr. Vision and editorial staff at \emph{PeerJ}, \\
 
@@
 \res{We have addressed comments from Reviewer 2 regarding the validity of the PCR assays.
-Sequencing of PCR products were not sequenced, but primer combinations for each of the 2 PCRs produced the expected sized bands.
-Primers were designed using previously sequenced teosinte and maize sequences provided by John Doebley, and were initially tested in both maize and teosinte for validity.
+PCR products were not sequenced, but primer combinations for both PCR assays produced only the expected sized bands.
+Primers were designed using previously sequenced teosinte and maize sequences provided by John Doebley and were initially tested in both maize and teosinte for validity.
 We have clarified this point in the text page 4, lines 75-87 and have added additional labeling to the gel image (Figure 2).}

### 2015-03-23 PeerJResponse.tex
edits to response

@@
 8) Line 150. Why were only 100 bootstrap replicates used for the NJ analysis? 1000 is more typical.\\
-\lev{I do not know the answer to this. I don't think that 1000 bootstraps would give us really a difference.}
-\mbh{if you still have the input file for this analysis, should be a quick fix and show we're doing everything we can to address reviewer concerns}
-\jri{agreed, should be easy to redo. while agree this is silly and i don't think 1000 will change anything, but since it's easy to do it's hard to argue against doing it.}
-\lev{I do not have the files as this is something that Tom did with the sequence data that he gathered}
-\mbh{o.k., should we ask Tom for this? The only argument we can make for not doing this is that 100 bootstraps are commonly used, but this makes us sound pretty unaccommodating}
+
+\res{We have conducted our NJ analysis with 1000 rather than 100 bootstraps and results are qualitatively very similar.}
 

### 2015-03-24 PeerJResponse.tex
edits to response

@@
 
-\res{We have editied the text to provide further clarification, page 8, lines 222-225: ``To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
+\res{We have edited the text to provide further clarification, page 8, lines 222-225: ``To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region."}
 
@@
 
-\res{We have removed the comparisons made about the difference in frequencies of the \emph{Hopscotch} allele between ssp. \emph{parviglumis} and ssp. \emph{mexicana}.}
+\res{The reviewer has made a valid point and we have removed the comparisons made about the difference in frequencies of the \emph{Hopscotch} allele between ssp.
+\emph{parviglumis} and ssp.
+\emph{mexicana}.}
 
@@
 
-\res{We sampled all individuals and populations that were available to us, which while it did not allow for even sample sizes across populations, did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as the geographic distribution of the \emph{Hopscotch} across many independent sampling locations.
+\res{We sampled all individuals and populations that were available to us, which, while it did not allow for even sample sizes across populations, did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as the geographic distribution of the \emph{Hopscotch} across many independent sampling locations.
 We have added a clarifying statement to the text regarding our sampling: page 4, lines 63-71: ``We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
@@
 
-\lev{I don't have access to this info. Can one of you guys help?}
-
-\mbh{Jeff, can you help here?}
+\res{We have added this additional information to Supplemental Tables 1 and 2}
 

### 2015-03-24 PeerJ_trackchanges.tex
Edits
LV edits to text and supplement regarding NJ tree

@@
 \textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the Analysis package from the  Libsequence library \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
-To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (100 repetitions).
+To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (10,000 repetitions).
 For creation of trees we also included homologous sequence data from Maize HapMapV2 \citep{Chia2012} for teosinte inbred lines (TILs), some of which are known to be homozygous for the \emph{Hopscotch} insertion (TIL03, TIL17, TIL09), as well as 59 lines of domesticated maize.

### 2015-03-24 Supplement_mbh.tex
edits
LV final edits to table s2
Edits
LV edits to text and supplement regarding NJ tree

@@
 \rowcolors{1}{gray!25}{white}
-\begin{longtable}{>{\stepcounter{rowno}}ccc}
+\begin{longtable}{>{\stepcounter{rowno}}cccc}
 \hiderowcolors
 \hline
-\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} \\
+\multicolumn{1}{c}{\textbf{Accession}} & \multicolumn{1}{c}{\textbf{USDA ID}} & \multicolumn{1}{c}{\textbf{Alleles Sampled}} & \multicolumn{1}{c}{\textbf{\emph{Hopscotch} Freq.}} \\
 \hline
@@
 \showrowcolors
-    RIMMA0066 & 2     & 1 \\
-    RIMMA0075 & 2     & 1 \\
-    RIMMA0077 & 2     & 1 \\
-    RIMMA0079 & 2     & 1 \\
-    RIMMA0081 & 2     & 1 \\
-    RIMMA0084 & 2     & 1 \\
-    RIMMA0086 & 2     & 1 \\
-    RIMMA0088 & 2     & 1 \\
-    RIMMA0089 & 2     & 1 \\
-    RIMMA0090 & 2     & 1 \\
-    RIMMA0092 & 4     & 1 \\
-    RIMMA0094 & 4     & 1 \\
-    RIMMA0097 & 2     & 1 \\
-    RIMMA0099 & 2     & 1 \\
-    RIMMA0100 & 2     & 1 \\
-    RIMMA0101 & 2     & 1 \\
-    RIMMA0104 & 2     & 1 \\
-    RIMMA0108 & 2     & 1 \\
-    RIMMA0111 & 6     & 1 \\
-    RIMMA0115 & 2     & 1 \\
-    RIMMA0117 & 2     & 1 \\
-    RIMMA0130 & 2     & 1 \\
-    RIMMA0133 & 2     & 1 \\
-    RIMMA0134 & 2     & 1 \\
-    RIMMA0135 & 2     & 1 \\
-    RIMMA0142 & 2     & 0.5 \\
-    RIMMA0143 & 4     & 1 \\
-    RIMMA0146 & 4     & 1 \\
-    RIMMA0149 & 2     & 1 \\
-    RIMMA0152 & 2     & 1 \\
-    RIMMA0153 & 2     & 1 \\
-    RIMMA0154 & 2     & 1 \\
-    RIMMA0155 & 2     & 1 \\
-    RIMMA0156 & 2     & 1 \\
-    RIMMA0157 & 2     & 1 \\
-    RIMMA0158 & 2     & 1 \\
-    RIMMA0159 & 2     & 1 \\
-    RIMMA0160 & 2     & 1 \\
-    RIMMA0162 & 2     & 1 \\
-    RIMMA0166 & 2     & 1 \\
-    RIMMA0167 & 2     & 1 \\
-    RIMMA0168 & 2     & 1 \\
-    RIMMA0169 & 2     & 1 \\
-    RIMMA0172 & 2     & 1 \\
-    RIMMA0174 & 4     & 1 \\
-    RIMMA0177 & 2     & 1 \\
-    RIMMA0178 & 2     & 1 \\
-    RIMMA0179 & 2     & 1 \\
-    RIMMA0181 & 2     & 1 \\
-    RIMMA0183 & 2     & 1 \\
-    RIMMA0184 & 2     & 1 \\
-    RIMMA0186 & 2     & 1 \\
-    RIMMA0187 & 1     & 1 \\
-    RIMMA0188 & 2     & 1 \\
-    RIMMA0195 & 2     & 1 \\
-    RIMMA0196 & 2     & 1 \\
-    RIMMA0197 & 2     & 1 \\
-    RIMMA0198 & 2     & 1 \\
-    RIMMA0199 & 2     & 1 \\
-    RIMMA0200 & 2     & 1 \\
-    RIMMA0202 & 2     & 1 \\
-    RIMMA0203 & 2     & 1 \\
-    RIMMA0206 & 2     & 1 \\
-    RIMMA0208 & 2     & 1 \\
-    RIMMA0209 & 2     & 1 \\
-    RIMMA0210 & 2     & 1 \\
-    RIMMA0212 & 2     & 1 \\
-    RIMMA0213 & 2     & 1 \\
-    RIMMA0214 & 2     & 1 \\
-    RIMMA0217 & 2     & 1 \\
-    RIMMA0218 & 2     & 1 \\
-    RIMMA0220 & 2     & 1 \\
-    RIMMA0221 & 2     & 1 \\
-    RIMMA0222 & 2     & 1 \\
-    RIMMA0223 & 2     & 1 \\
-    RIMMA0226 & 2     & 1 \\
-    RIMMA0227 & 2     & 1 \\
-    RIMMA0228 & 2     & 1 \\
-    RIMMA0229 & 2     & 1 \\
-    RIMMA0230 & 2     & 1 \\
-    RIMMA0232 & 2     & 1 \\
-    RIMMA0233 & 2     & 1 \\
-    RIMMA0235 & 2     & 0.5 \\
-    RIMMA0242 & 2     & 1 \\
-    RIMMA0243 & 2     & 1 \\
-    RIMMA0247 & 4     & 1 \\
-    RIMMA0248 & 2     & 1 \\
-    RIMMA0249 & 2     & 1 \\
-    RIMMA0252 & 2     & 1 \\
-    RIMMA0253 & 2     & 1 \\
-    RIMMA0254 & 2     & 1 \\
-    RIMMA0256 & 2     & 1 \\
-    RIMMA0257 & 2     & 1 \\
-    RIMMA0258 & 2     & 1 \\
-    RIMMA0259 & 2     & 1 \\
-    RIMMA0260 & 2     & 1 \\
-    RIMMA0262 & 2     & 1 \\
-    RIMMA0263 & 2     & 1 \\
-    RIMMA0264 & 2     & 1 \\
-    RIMMA0265 & 2     & 1 \\
-    RIMMA0268 & 2     & 1 \\
-    RIMMA0269 & 2     & 1 \\
-    RIMMA0270 & 2     & 1 \\
-    RIMMA0272 & 2     & 1 \\
-    RIMMA0275 & 2     & 1 \\
-    RIMMA0276 & 2     & 1 \\
-    RIMMA0277 & 2     & 1 \\
-    RIMMA0279 & 2     & 1 \\
-    RIMMA0280 & 2     & 1 \\
-    RIMMA0283 & 2     & 1 \\
-    RIMMA0285 & 2     & 1 \\
-    RIMMA0288 & 2     & 1 \\
-    RIMMA0290 & 2     & 1 \\
-    RIMMA0291 & 2     & 1 \\
-    RIMMA0292 & 2     & 1 \\
-    RIMMA0293 & 2     & 1 \\
-    RIMMA0298 & 2     & 1 \\
-    RIMMA0302 & 2     & 1 \\
-    RIMMA0305 & 2     & 1 \\
-    RIMMA0310 & 2     & 1 \\
-    RIMMA0312 & 2     & 1 \\
-    RIMMA0320 & 2     & 1 \\
-    RIMMA0322 & 2     & 1 \\
-    RIMMA0324 & 2     & 1 \\
-    RIMMA0334 & 2     & 1 \\
-    RIMMA0336 & 2     & 1 \\
-    RIMMA0337 & 2     & 1 \\
-    RIMMA0338 & 2     & 1 \\
-    RIMMA0339 & 2     & 1 \\
-    RIMMA0340 & 2     & 1 \\
-    RIMMA0341 & 2     & 1 \\
-    RIMMA0342 & 2     & 1 \\
-    RIMMA0344 & 2     & 1 \\
-    RIMMA0346 & 2     & 1 \\
-    RIMMA0347 & 2     & 1 \\
-    RIMMA0350 & 2     & 1 \\
-    RIMMA0351 & 2     & 1 \\
-    RIMMA0352 & 2     & 1 \\
-    RIMMA0357 & 2     & 1 \\
-    RIMMA0360 & 18    & 1 \\
-    RIMMA0362 & 8     & 1 \\
-    RIMMA0363 & 16    & 1 \\
-    RIMMA0370 & 17    & 1 \\
-    RIMMA0372 & 24    & 1 \\
-    RIMMA0373 & 22    & 1 \\
-    RIMMA0408 & 2     & 1 \\
-    RIMMA0411 & 1     & 1 \\
-    RIMMA0413 & 1     & 1 \\
-    RIMMA0414 & 1     & 1 \\
-    RIMMA0424 & 8     & 0.9 \\
-    RIMMA0432 & 1     & 1 \\
-    RIMMA0434 & 2     & 0.5 \\
-    RIMMA0435 & 1     & 1 \\
-    RIMMA0442 & 2     & 1 \\
-    RIMMA0443 & 2     & 1 \\
-    RIMMA0444 & 2     & 1 \\
-    RIMMA0445 & 2     & 1 \\
-    RIMMA0446 & 2     & 1 \\
-    RIMMA0447 & 2     & 1 \\
-    RIMMA0448 & 2     & 1 \\
-    RIMMA0449 & 2     & 1 \\
-    RIMMA0450 & 2     & 1 \\
-    RIMMA0451 & 2     & 1 \\
-    RIMMA0452 & 2     & 1 \\
-    RIMMA0453 & 2     & 1 \\
-    RIMMA0454 & 2     & 1 \\
-    RIMMA0455 & 2     & 1 \\
-    RIMMA0456 & 2     & 1 \\
-    RIMMA0457 & 2     & 1 \\
-    RIMMA0458 & 2     & 1 \\
-    RIMMA0459 & 2     & 1 \\
-    RIMMA0483 & 2     & 1 \\
-    RIMMA0490 & 2     & 1 \\
-    RIMMA0515 & 2     & 1 \\
-    RIMMA0537 & 2     & 1 \\
-    RIMMA0550 & 2     & 1 \\
-    RIMMA0553 & 2     & 1 \\
-    RIMMA0559 & 2     & 1 \\
-    RIMMA0561 & 2     & 1 \\
-    RIMMA0562 & 2     & 1 \\
-    RIMMA0571 & 2     & 1 \\
-    RIMMA0572 & 2     & 1 \\
-    RIMMA0577 & 2     & 1 \\
-    RIMMA0579 & 2     & 1 \\
-    RIMMA0590 & 2     & 1 \\
-    RIMMA0591 & 2     & 1 \\
-    RIMMA0592 & 2     & 1 \\
-    RIMMA0593 & 2     & 1 \\
-    RIMMA0594 & 2     & 1 \\
-    RIMMA0595 & 2     & 1 \\
-    RIMMA0596 & 2     & 1 \\
-    RIMMA0597 & 2     & 1 \\
-    RIMMA0598 & 2     & 1 \\
-    RIMMA0599 & 2     & 1 \\
-    RIMMA0600 & 2     & 1 \\
-    RIMMA0601 & 2     & 1 \\
-    RIMMA0602 & 2     & 1 \\
-    RIMMA0603 & 2     & 1 \\
-    RIMMA0604 & 2     & 1 \\
-    RIMMA0605 & 2     & 1 \\
-    RIMMA0606 & 2     & 1 \\
-    RIMMA0607 & 2     & 1 \\
-    RIMMA0608 & 2     & 1 \\
-    RIMMA0609 & 2     & 1 \\
-    RIMMA0610 & 2     & 1 \\
-    RIMMA0611 & 2     & 1 \\
-    RIMMA0612 & 2     & 1 \\
-    RIMMA0613 & 2     & 1 \\
-    RIMMA0622 & 2     & 0.5 \\
-    RIMMA0624 & 1     & 1 \\
-    RIMMA0629 & 2     & 0.5 \\
-    RIMMA0631 & 2     & 0.5 \\
-    RIMMA0659 & 2     & 1 \\
-    RIMMA0660 & 2     & 1 \\
-    RIMMA0678 & 1     & 1 \\
-    RIMMA0681 & 2     & 1 \\
-    RIMMA0683 & 2     & 1 \\
-    RIMMA0684 & 2     & 1 \\
-    RIMMA0685 & 4     & 1 \\
-    RIMMA0693 & 2     & 1 \\
-    RIMMA0694 & 2     & 1 \\
-    RIMMA0695 & 1     & 1 \\
-    RIMMA0699 & 2     & 1 \\
-    RIMMA0704 & 2     & 1 \\
-    RIMMA0706 & 2     & 1 \\
-    RIMMA0711 & 1     & 1 \\
-    RIMMA0713 & 2     & 1 \\
-    RIMMA0715 & 2     & 1 \\
-    RIMMA0717 & 2     & 1 \\
-    RIMMA0718 & 2     & 1 \\
-    RIMMA0723 & 2     & 1 \\
-    RIMMA0724 & 2     & 1 \\
-    RIMMA0725 & 1     & 1 \\
-    RIMMA0728 & 1     & 1 \\
-    RIMMA0732 & 2     & 1 \\
-    RIMMA0734 & 2     & 1 \\
-    RIMMA0738 & 2     & 1 \\
-    RIMMA0739 & 2     & 1 \\
-    RIMMA0744 & 2     & 1 \\
-    RIMMA0747 & 1     & 1 \\
-    RIMMA0748 & 2     & 1 \\
-    RIMMA0749 & 2     & 1 \\
-    RIMMA0750 & 2     & 0.5 \\
-    RIMMA0751 & 1     & 1 \\
-    RIMMA0752 & 1     & 1 \\
-    RIMMA0753 & 2     & 1 \\
-    RIMMA0755 & 2     & 1 \\
+    RIMMA0066 & 27096 & 2     & 1 \\
+    RIMMA0075 & 19309 & 2     & 1 \\
+    RIMMA0077 & 19313 & 2     & 1 \\
+    RIMMA0079 & 19321 & 2     & 1 \\
+    RIMMA0081 & 20190 & 2     & 1 \\
+    RIMMA0084 & 22439 & 2     & 1 \\
+    RIMMA0086 & 22748 & 2     & 1 \\
+    RIMMA0088 & 22756 & 2     & 1 \\
+    RIMMA0089 & 23392 & 2     & 1 \\
+    RIMMA0090 & 23423 & 2     & 1 \\
+    RIMMA0092 & 23427 & 4     & 1 \\
+    RIMMA0094 & 23447 & 4     & 1 \\
+    RIMMA0097 & 23488 & 2     & 1 \\
+    RIMMA0099 & 23509 & 2     & 1 \\
+    RIMMA0100 & 23518 & 2     & 1 \\
+    RIMMA0101 & 23519 & 2     & 1 \\
+    RIMMA0104 & 24590 & 2     & 1 \\
+    RIMMA0108 & 26022 & 2     & 1 \\
+    RIMMA0111 & 26770 & 6     & 1 \\
+    RIMMA0115 & 26798 & 2     & 1 \\
+    RIMMA0117 & 27045 & 2     & 1 \\
+    RIMMA0130 & 28336 & 2     & 1 \\
+    RIMMA0133 & 8581 & 2     & 1 \\
+    RIMMA0134 & 22632 & 2     & 1 \\
+    RIMMA0135 & 22634 & 2     & 1 \\
+    RIMMA0142 & 30060 & 2     & 0.5 \\
+    RIMMA0143 & 30065 & 4     & 1 \\
+    RIMMA0146 & 30863 & 4     & 1 \\
+    RIMMA0149 & 30869 & 2     & 1 \\
+    RIMMA0152 & 32722 & 2     & 1 \\
+    RIMMA0153 & 32728 & 2     & 1 \\
+    RIMMA0154 & 32732 & 2     & 1 \\
+    RIMMA0155 & 42809 & 2     & 1 \\
+    RIMMA0156 & 42872 & 2     & 1 \\
+    RIMMA0157 & 42874 & 2     & 1 \\
+    RIMMA0158 & 42875 & 2     & 1 \\
+    RIMMA0159 & 42878 & 2     & 1 \\
+    RIMMA0160 & 65864 & 2     & 1 \\
+    RIMMA0162 & 65866 & 2     & 1 \\
+    RIMMA0166 & 213696 & 2     & 1 \\
+    RIMMA0167 & 213697 & 2     & 1 \\
+    RIMMA0168 & 213698 & 2     & 1 \\
+    RIMMA0169 & 213699 & 2     & 1 \\
+    RIMMA0172 & 213705 & 2     & 1 \\
+    RIMMA0174 & 213708 & 4     & 1 \\
+    RIMMA0177 & 213715 & 2     & 1 \\
+    RIMMA0178 & 213716 & 2     & 1 \\
+    RIMMA0179 & 213717 & 2     & 1 \\
+    RIMMA0181 & 213719 & 2     & 1 \\
+    RIMMA0183 & 213721 & 2     & 1 \\
+    RIMMA0184 & 213722 & 2     & 1 \\
+    RIMMA0186 & 213724 & 2     & 1 \\
+    RIMMA0187 & 213725 & 1     & 1 \\
+    RIMMA0188 & 213727 & 2     & 1 \\
+    RIMMA0195 & 213781 & 2     & 1 \\
+    RIMMA0196 & 213793 & 2     & 1 \\
+    RIMMA0197 & 214197 & 2     & 1 \\
+    RIMMA0198 & 214293 & 2     & 1 \\
+    RIMMA0199 & 214294 & 2     & 1 \\
+    RIMMA0200 & 214295 & 2     & 1 \\
+    RIMMA0202 & 217407 & 2     & 1 \\
+    RIMMA0203 & 217408 & 2     & 1 \\
+    RIMMA0206 & 217473 & 2     & 1 \\
+    RIMMA0208 & 218005 & 2     & 1 \\
+    RIMMA0209 & 218195 & 2     & 1 \\
+    RIMMA0210 & 219871 & 2     & 1 \\
+    RIMMA0212 & 219882 & 2     & 1 \\
+    RIMMA0213 & 219883 & 2     & 1 \\
+    RIMMA0214 & 219884 & 2     & 1 \\
+    RIMMA0217 & 221872 & 2     & 1 \\
+    RIMMA0218 & 221873 & 2     & 1 \\
+    RIMMA0220 & 221876 & 2     & 1 \\
+    RIMMA0221 & 221877 & 2     & 1 \\
+    RIMMA0222 & 221878 & 2     & 1 \\
+    RIMMA0223 & 221880 & 2     & 1 \\
+    RIMMA0226 & 221889 & 2     & 1 \\
+    RIMMA0227 & 222314 & 2     & 1 \\
+    RIMMA0228 & 222315 & 2     & 1 \\
+    RIMMA0229 & 222316 & 2     & 1 \\
+    RIMMA0230 & 222317 & 2     & 1 \\
+    RIMMA0232 & 222469 &  2     & 1 \\
+    RIMMA0233 & 222470 & 2     & 1 \\
+    RIMMA0235 & 222474 & 2     & 0.5 \\
+    RIMMA0242 & 222614 & 2     & 1 \\
+    RIMMA0243 & 222615 & 2     & 1 \\
+    RIMMA0247 & 222639 & 4     & 1 \\
+    RIMMA0248 & 222640 & 2     & 1 \\
+    RIMMA0249 & 222641 & 2     & 1 \\
+    RIMMA0252 & 233002 & 2     & 1 \\
+    RIMMA0253 & 233006 & 2     & 1 \\
+    RIMMA0254 & 233008 & 2     & 1 \\
+    RIMMA0256 & 237000 & 2     & 1 \\
+    RIMMA0257 & 257506 & 2     & 1 \\
+    RIMMA0258 & 267179 & 2     & 1 \\
+    RIMMA0259 & 269743 & 2     & 1 \\
+    RIMMA0260 & 269744 & 2     & 1 \\
+    RIMMA0262 & 270297 & 2     & 1 \\
+    RIMMA0263 & 278713 & 2     & 1 \\
+    RIMMA0264 & 278716 & 2     & 1 \\
+    RIMMA0265 & 278717 & 2     & 1 \\
+    RIMMA0268 & 278721 & 2     & 1 \\
+    RIMMA0269 & 278724 & 2     & 1 \\
+    RIMMA0270 & 280061 & 2     & 1 \\
+    RIMMA0272 & 280853 & 2     & 1 \\
+    RIMMA0275 & 311232 & 2     & 1 \\
+    RIMMA0276 & 311235 & 2     & 1 \\
+    RIMMA0277 & 311236 & 2     & 1 \\
+    RIMMA0279 & 311241 & 2     & 1 \\
+    RIMMA0280 & 311243 & 2     & 1 \\
+    RIMMA0283 & 401757 & 2     & 1 \\
+    RIMMA0285 & 414176 & 2     & 1 \\
+    RIMMA0288 & 452032 & 2     & 1 \\
+    RIMMA0290 & 452045 & 2     & 1 \\
+    RIMMA0291 & 452046 & 2     & 1 \\
+    RIMMA0292 & 452047 & 2     & 1 \\
+    RIMMA0293 & 452048 & 2     & 1 \\
+    RIMMA0298 & 483273 & 2     & 1 \\
+    RIMMA0302 & 540784 & 2     & 1 \\
+    RIMMA0305 & 548792 & 2     & 1 \\
+    RIMMA0310 & 550462 & 2     & 1 \\
+    RIMMA0312 & 550466 & 2     & 1 \\
+    RIMMA0320 & 583896 & 2     & 1 \\
+    RIMMA0322 & 583898 & 2     & 1 \\
+    RIMMA0324 & 587125 & 2     & 1 \\
+    RIMMA0334 & 587153 & 2     & 1 \\
+    RIMMA0336 & 607519 & 2     & 1 \\
+    RIMMA0337 & 607522 & 2     & 1 \\
+    RIMMA0338 & 607593 & 2     & 1 \\
+    RIMMA0339 & 607599 & 2     & 1 \\
+    RIMMA0340 & 607602 & 2     & 1 \\
+    RIMMA0341 & 608465 & 2     & 1 \\
+    RIMMA0342 & 608502 & 2     & 1 \\
+    RIMMA0344 & 608539 & 2     & 1 \\
+    RIMMA0346 & 608618 & 2     & 1 \\
+    RIMMA0347 & 608619 & 2     & 1 \\
+    RIMMA0350 & 608646 & 2     & 1 \\
+    RIMMA0351 & 608652 & 2     & 1 \\
+    RIMMA0352 & 608764 & 2     & 1 \\
+    RIMMA0357 & 608777 & 2     & 1 \\
+    RIMMA0360 & N/A & 18    & 1 \\
+    RIMMA0362 & N/A & 8     & 1 \\
+    RIMMA0363 & N/A & 16    & 1 \\
+    RIMMA0370 & N/A & 17    & 1 \\
+    RIMMA0372 & N/A & 24    & 1 \\
+    RIMMA0373 & N/A & 22    & 1 \\
+    RIMMA0408 & 478967 & 2     & 1 \\
+    RIMMA0411 & 479044 & 1     & 1 \\
+    RIMMA0413 & 484413 & 1     & 1 \\
+    RIMMA0414 & 484421 & 1     & 1 \\
+    RIMMA0424 & 485120 & 8     & 0.9 \\
+    RIMMA0432 & 489126 & 1     & 1 \\
+    RIMMA0434 & 490921 & 2     & 0.5 \\
+    RIMMA0435 & 490973 & 1     & 1 \\
+    RIMMA0442 & 537097 & 2     & 1 \\
+    RIMMA0443 & 537099 & 2     & 1 \\
+    RIMMA0444 & 538009 & 2     & 1 \\
+    RIMMA0445 & 538011 & 2     & 1 \\
+    RIMMA0446 & 539920 & 2     & 1 \\
+    RIMMA0447 & 539921 & 2     & 1 \\
+    RIMMA0448 & 539922 & 2     & 1 \\
+    RIMMA0449 & 539923 & 2     & 1 \\
+    RIMMA0450 & 539924 & 2     & 1 \\
+    RIMMA0451 & 539926 & 2     & 1 \\
+    RIMMA0452 & 539927 & 2     & 1 \\
+    RIMMA0453 & 543843 & 2     & 1 \\
+    RIMMA0454 & 543846 & 2     & 1 \\
+    RIMMA0455 & 543847 & 2     & 1 \\
+    RIMMA0456 & 543849 & 2     & 1 \\
+    RIMMA0457 & 543850 & 2     & 1 \\
+    RIMMA0458 & 547086 & 2     & 1 \\
+    RIMMA0459 & 547088 & 2     & 1 \\
+    RIMMA0483 & 600954 & 2     & 1 \\
+    RIMMA0490 & 601003 & 2     & 1 \\
+    RIMMA0515 & 601270 & 2     & 1 \\
+    RIMMA0537 & 601416 & 2     & 1 \\
+    RIMMA0550 & 601468 & 2     & 1 \\
+    RIMMA0553 & 601489 & 2     & 1 \\
+    RIMMA0559 & 601495 & 2     & 1 \\
+    RIMMA0561 & 601497 & 2     & 1 \\
+    RIMMA0562 & 601498 & 2     & 1 \\
+    RIMMA0571 & 601562 & 2     & 1 \\
+    RIMMA0572 & 601563 & 2     & 1 \\
+    RIMMA0577 & 601568 & 2     & 1 \\
+    RIMMA0579 & 601570 & 2     & 1 \\
+    RIMMA0590 & 601724 & 2     & 1 \\
+    RIMMA0591 & 601725 & 2     & 1 \\
+    RIMMA0592 & 601729 & 2     & 1 \\
+    RIMMA0593 & 601773 & 2     & 1 \\
+    RIMMA0594 & 601774 & 2     & 1 \\
+    RIMMA0595 & 601775 & 2     & 1 \\
+    RIMMA0596 & 601776 & 2     & 1 \\
+    RIMMA0597 & 601777 & 2     & 1 \\
+    RIMMA0598 & 601778 & 2     & 1 \\
+    RIMMA0599 & 601779 & 2     & 1 \\
+    RIMMA0600 & 601780 & 2     & 1 \\
+    RIMMA0601 & 601781 & 2     & 1 \\
+    RIMMA0602 & 601782 & 2     & 1 \\
+    RIMMA0603 & 601783 & 2     & 1 \\
+    RIMMA0604 & 601784 & 2     & 1 \\
+    RIMMA0605 & 601785 & 2     & 1 \\
+    RIMMA0606 & 601786 & 2     & 1 \\
+    RIMMA0607 & 601787 & 2     & 1 \\
+    RIMMA0608 & 601788 & 2     & 1 \\
+    RIMMA0609 & 601789 & 2     & 1 \\
+    RIMMA0610 & 601790 & 2     & 1 \\
+    RIMMA0611 & 601791 & 2     & 1 \\
+    RIMMA0612 & 601792 & 2     & 1 \\
+    RIMMA0613 & 601808 & 2     & 1 \\
+    RIMMA0622 & 645837 & 2     & 0.5 \\
+    RIMMA0624 & 645930 & 1     & 1 \\
+    RIMMA0629 & 646057 & 2     & 0.5 \\
+    RIMMA0631 & 646072 & 2     & 0.5 \\
+    RIMMA0659 & 286729 & 2     & 1 \\
+    RIMMA0660 & 485451 & 2     & 1 \\
+    RIMMA0678 & 285833 & 1     & 1 \\
+    RIMMA0681 & 628446 & 2     & 1 \\
+    RIMMA0683 & 483568 & 2     & 1 \\
+    RIMMA0684 & 28482 & 2     & 1 \\
+    RIMMA0685 & 287623 & 4     & 1 \\
+    RIMMA0693 & 487959 & 2     & 1 \\
+    RIMMA0694 & 28487 & 2     & 1 \\
+    RIMMA0695 & 28567 & 1     & 1 \\
+    RIMMA0699 & 28465 & 2     & 1 \\
+    RIMMA0704 & 28623 & 2     & 1 \\
+    RIMMA0706 & 28615 & 2     & 1 \\
+    RIMMA0711 & 29075 & 1     & 1 \\
+    RIMMA0713 & 489399 & 2     & 1 \\
+    RIMMA0715 & 28454 & 2     & 1 \\
+    RIMMA0717 & 483926 & 2     & 1 \\
+    RIMMA0718 & 484805 & 2     & 1 \\
+    RIMMA0723 & 28565 & 2     & 1 \\
+    RIMMA0724 & 490826 & 2     & 1 \\
+    RIMMA0725 & 490827 & 1     & 1 \\
+    RIMMA0728 & 484969 & 1     & 1 \\
+    RIMMA0732 & 28456 & 2     & 1 \\
+    RIMMA0734 & 26028 & 2     & 1 \\
+    RIMMA0738 & N/A & 2     & 1 \\
+    RIMMA0739 & N/A & 2     & 1 \\
+    RIMMA0744 & N/A & 2     & 1 \\
+    RIMMA0747 & N/A & 1     & 1 \\
+    RIMMA0748 & N/A & 2     & 1 \\
+    RIMMA0749 & N/A & 2     & 1 \\
+    RIMMA0750 & N/A & 2     & 0.5 \\
+    RIMMA0751 & N/A & 1     & 1 \\
+    RIMMA0752 & N/A & 1     & 1 \\
+    RIMMA0753 & N/A & 2     & 1 \\
+    RIMMA0755 & N/A & 2     & 1 \\
     \hline
@@
    \includegraphics[width=150mm]{FigS3NJtrees.png}
-    \flushleft{Figure S2: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1}.
+    \flushleft{Figure S2: Neighbor-joining tree of the sequenced region in the 5' UTR (right; Region 1) and the 66,169 bp upstream region (left; Region 2) of \emph{tb1} using 10,000 bootstraps.
 Individuals with genotype data are colored: Homozygous for the teosinte (no \emph{Hopscotch}) allele (red), homozygous for the maize (\emph{Hopscotch}) allele (blue), heterozygotes (purple).

### 2015-03-24 PeerJResponse.tex
edits to response

@@
 
-\res{We have conducted our NJ analysis with 1000 rather than 100 bootstraps and results are qualitatively very similar.}
+\res{We actually conducted our NJ analysis with 10,000 bootstrap replicates and mistakenly reported 100 replicates.
+This is clear based on our bootstrap support values which are reported to two decimal places.
+We have corrected this typo.}
 

### 2015-03-24 PeerJResponse.tex
edits
LV edits to page/line numbers

@@
 
-\res{We have clarified this; accession refers to population; page 2, lines 62-66: ``We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+\res{We have clarified this; accession refers to population; page 4, lines 63-71: ``We sampled 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
 Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations."}
@@
 \res{We are referring to Table S1 and Table S2.
-We have edited our reference to the tables to be more clear, page 8, lines 198-200: ``The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article)."}
+We have edited our reference to the tables to be more clear, page 8, lines 209-210: ``The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article)."}
 

### 2015-03-25 PeerJResponse.tex
final edits revision

--- 2015-03-24 PeerJResponse.tex
@@
 
-\lev{Okay. I am not sure that we have significance values for these.}
-
-\mbh{We don't and it's kind of a strange/silly request...Jeff and I will deal with this}
-\jri{ i can do the simple but dumb calculation if you guys want for response to reviewers, but disagree we should include in manuscript.
-we mention they are at the tail of the empirical distribution from wright et al don't we?}
-\mbh{I agree this is a silly request and that we shouldn't include in the manuscript; yes, we indicate Taj.D for our pops is in the extreme tail based on comparison to both Wright et al.
-and Tiffin and Moeller}
+\res{We have determined significance of the Tajima's D results by comparing our empirical values to coalescent simulations generated using the program ms \citep{Hudson2002}.
+We describe these methods and results on page 7, lines 157-161 and page 13, lines 368-369.
+Significant values are now marked with an asterisk in Table 2}
 
@@
 
-\lev{Okay. What do you guys think of this? I don't think we have tests of significance for tajd}
-
-\mbh{Jeff and I will address this}
-\jri{sloppy my ass. <censored> reviewer!}
-\lev{ :) }
+\res{Please see our response to item 14.}
 

### 2015-03-25 Vann_et_al._Manuscript.tex
final edits revision

--- 2015-03-24 PeerJ_trackchanges.tex
@@
 Maize and the teosintes are an attractive system in which to study domestication due to the abundance of genetic tools developed for maize and well-characterized domestication loci \citep{Hufford2012a, Doebley2004, Hufford2012b}.
-\track{Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp.
+Additionally, large, naturally-occurring populations of both \emph{parviglumis} and \emph{Zea mays} ssp.
 \emph{mexicana} (hereafter, \emph{mexicana}) can be found throughout Mexico \citep{Wilkes1977, Hufford2013}, with \emph{parviglumis} distributed in the lowlands of Mexico and \emph{mexicana} in the highlands.
-Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}}.
+Furthermore, both \emph{parviglumis} and \emph{mexicana} occur at high densities and genetic diversity of these taxa is estimated to be high \citep{Hufford2012a, Ross-Ibarra2009}.
 
@@
 
-\textbf{Sampling and genotyping---}\track{We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 \track{populations} (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
-Numbers of individuals sampled \track{per population} ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
-Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites.} DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
+\textbf{Sampling and genotyping---}We sampled all individuals and populations that were available to us, consisting of 1,110 individuals from 350 populations (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+Numbers of individuals sampled per population ranged from 1-43 for \emph{parviglumis}, 1-35 for \emph{mexicana}, and 1-18 for the maize landrace populations.
+Available samples did not allow us to sample evenly from populations, but did allow us to calculate \emph{Hopscotch} frequency in a subset of populations, as well as elucidate the geographic distribution of the \emph{Hopscotch} across multiple independent sampling sites.
+DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}.
 We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion due to its large size ($\sim$5kb).
-\track{Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR).
+Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer (HopF) and an internal primer (HopIntR).
 Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'};
 HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Fig.
-S1 and \track{\ref{Fig1Cartoon}}, See Supplemental Materials with the online version of this article).
+\ref{Fig1Cartoon} and Fig.
+S1; See Supplemental Materials with the online version of this article).
 Homozygotes for the no-\emph{Hopscotch} allele show a single band for absence of the element ($\sim$300bp) produced by the HopF/HopR primer set, and 0 bands for the HopF/HopIntR primer set since they lack the LTR where the internal primer sequence is located.
 Homozygotes for the \emph{Hopscotch} allele also show one band at 5kb for the HopF/HopR PCR product as well as one band at 1.1kb for the HopF/HopIntR PCR.
-Heterozygotes for the \emph{Hopscotch} allele show three bands total; both a 300bp band and a 5kb band for the HopF/HopR PCR and a 1.1Kb band for the HopF/HopIntR PCR (Table S2, See Supplemental Materials with the online version of this article).} Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
+Heterozygotes for the \emph{Hopscotch} allele show three bands total; both a 300bp band and a 5kb band for the HopF/HopR PCR and a 1.1Kb band for the HopF/HopIntR PCR (Table S2; See Supplemental Materials with the online version of this article).
+Since we developed a PCR protocol for each allele, if only one PCR resolved well, we scored one allele for that individual rather than infer the diploid genotype.
 We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min.
@@
 \emph{mexicana}) genotyping and covariance data from \citet{Pyhajarvi2013} for BayEnv, with the \emph{Hopscotch} insertion coded as an additional biallelic marker.
-\track{SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}.
-Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.}  Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
+SNP data from \citet{Pyhajarvi2013} were obtained using the MaizeSNP50 BeadChip and Infinium HD Assay (Illumina, San Diego, CA, USA) and phased using the program fastPHASE \citep{Scheet2006}.
+Environmental data were previously obtained from \url{www.worldclim.org} and soil data were downloaded from the Harmonized World Soil Database \citep{FAOHWSD} at \url{www.harvestchoice.org}.
+Environmental data represent average values for the last several decades (climatic data) or are likely stable over time (soil data) and therefore represent conditions important for local adaptation of our samples.
 Information from these data sets was summarized by principle component analysis following \citet{Pyhajarvi2013}.
@@
 \textbf{Sequence analysis---}For population genetic analyses of sequenced Region 1 and sequenced Region 2 we used the Analysis package from the  Libsequence library \citep{Thornton2003} to calculate pairwise F\textsubscript{ST} between populations and to calculate standard diversity statistics (number of haplotypes, haplotype diversity, Watterson's estimator $\hat\theta_W$, pairwise nucleotide diversity $\hat\theta_\pi$, and Tajima's D).
+Significance of Tajima's D results was gauged by comparing empirical data to 10,000 coalescent simulations conducted using the program ms \citep{Hudson2002} under a standard neutral model based on observed estimates of the population mutation rate theta and assuming an identical value for the population recombination rate rho.
+Empirical results falling outside the 95\% confidence interval of our simulated data were deemed significant.
 To produce a visual representation of differentiation between sequences and examine patterns in sequence clustering by \emph{Hopscotch} genotype we used Phylip (\url{http://evolution.genetics.washington.edu/phylip.html}) to create neighbor-joining trees with bootstrap-supported nodes (10,000 repetitions).
@@
 
-\textbf{Genotyping for the \emph{Hopscotch} insertion---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened \track{(Table S1 and Table S2, See Supplemental Materials with the online version of this article).}Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
+\textbf{Genotyping for the \emph{Hopscotch} insertion---}The genotype at the \emph{Hopscotch} insertion was confirmed with two PCRs for 837 individuals of the 1,100 screened (Table S1 and Table S2, See Supplemental Materials with the online version of this article).
+Among the 247 maize landrace accessions genotyped, all but eight were homozygous for the presence of the insertion Within our \emph{parviglumis} and \emph{mexicana} samples we found the \emph{Hopscotch} insertion segregating in 37 ($n=86$) and four ($n=17$) populations, respectively, and at highest frequency within populations in the states of Jalisco, Colima, and Michoac\'{a}n in central-western Mexico (Fig.
 \ref{Fig2Map}).
@@
 
-\textbf{Sequencing upstream regions of the \emph{tb1} ORF---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations \track{from the Jalisco region}.
+\textbf{Sequencing upstream regions of the \emph{tb1} ORF---}To investigate patterns of sequence diversity and linkage disequilibrium (LD) in the \emph{tb1} region and any evidence of selection on this locus, we sequenced two small ($<$1kb) regions upstream of the \emph{tb1} ORF in four populations from the Jalisco region.
 After alignment and singleton checking we recovered 48 and 40 segregating sites for the 5' UTR region (Region 1) and the 66kb upstream region (Region 2), respectively.
 For Region 1, Ejutla A has the highest values of haplotype diversity and $\theta_\pi$, while Ejutla B and La Mesa have comparable values of these summary statistics, and San Lorenzo has much lower values.
-Additionally, Tajima's D is strongly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}).
+Additionally, Tajima's D is significantly negative in the two Ejutla populations and La Mesa, but is closer to zero in San Lorenzo (Table \ref{Table2Diversity}).
 For Region 2, haplotype diversity and $\theta_\pi$, are similar for Ejutla A and Ejutla B, while La Mesa and San Lorenzo have slightly lower values for these statistics (Table \ref{Table2Diversity}).
-Tajima's D is positive \track{in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants} (Table \ref{Table2Diversity}).
+Tajima's D is positive in all populations except La Mesa, where a slightly negative value suggests a slight excess of low frequency variants (Table \ref{Table2Diversity}).
 Pairwise values of F\textsubscript{ST} within population pairs Ejutla A/Ejutla B and San Lorenzo/La Mesa are close to zero for both sequenced regions as well as for the \emph{Hopscotch}, while they are high for other population pairs (Table \ref{Table1Fst}).
 
-\textbf{Evidence of introgression around the \emph{tb1} region---} \track{We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citep{Pyhajarvi2013}.} The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
+\textbf{Evidence of introgression around the \emph{tb1} region---} We investigated the possibility of introgression as an explanation for the frequency of the \emph{Hopscotch} allele in populations of teosinte using previously collected SNP data from \citet{Pyhajarvi2013}.
+The highest frequency of the \emph{Hopscotch} insertion in teosinte was found in \emph{parviglumis} sympatric with cultivated maize.
 Our initial hypothesis was that the high frequency of the \emph{Hopscotch} element in these populations could be attributed to introgression from maize into teosinte.
@@
 Overall nucleotide diversity is 76\% less than seen in the sequences from the 66kb upstream region, and Tajima's D is considerably lower and consistently negative across populations (Table \ref{Table2Diversity}).
-In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007}.
+In fact, values of Tajima's D in the 5' UTR are toward the extreme negative end of the distribution of this statistic previously calculated across loci sequenced in \emph{parviglumis} \citep{Wright27052005, Moeller01072007} and significantly negative in three of our surveyed populations (EjuA, EjuB, MSA) based on coalescent simulations under a standard neutral model.
 Though not definitive, these results are consistent with the action of selection on the upstream region of \emph{tb1}, perhaps suggesting an ecological role for the gene in Jaliscan populations of \emph{parviglumis}.
@@
 We have found no significant effect of the \emph{Hopscotch} insertion on tillering index or tiller number, a result that is discordant with its clear phenotypic effects in maize.
-\track{It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}.
-This factor may have limited our capacity to observe variation in tillering index.}
+It is possible that the planting density of our seedlings (plants spaced 12 inches apart) was too high, leading to an overall decrease in tillering as previously seen in \citet{LukensDoebley1999}.
+This factor may have limited our capacity to observe variation in tillering index.
 
@@
   \centering
-  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus}
+  \caption{Population genetic statistics from resequenced regions near the \emph{tb1} locus. Significant values are marked with an asterisk}.
     \begin{tabular}{ccccc}\\\toprule
@@
     \multicolumn{5}{c}{\textit{Region 1(5' UTR)}} \\
-    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650 \\
-    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831 \\
-    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755 \\
+    \multicolumn{1}{c}{EJUA} & 8     & 0.859 & 0.005 & -1.650* \\
+    \multicolumn{1}{c}{EJUB} & 5     & 0.709 & 0.004 & -1.831* \\
+    \multicolumn{1}{c}{MSA} & 6     & 0.682 & 0.004 & -1.755* \\
     \multicolumn{1}{c}{SLO} & 3     & 0.318 & 0.001 & -0.729 \\

### 2015-03-25 Supplement_mbh.tex
Edits
LV Edits to Table S1

--- 2015-03-23 Supplement_mbh.tex
@@
     RIHY0009 & N/A   & N/A   & 2     & 0.5   & 0.5 \\
-    RIMME0006 & 566673 & Durango & 2	& 0 & 1 \\
-    RIMME0007 & 566680 & Guanajuato & 2 & 0 & 1 \\
-    RIMME0008 & 566681 & Michoacan & 2 & 0 & 1 \\
-    RIMME0009 & 566682 & Distrito Federal & 2 & 0 & 1 \\
-    RIMME0011 & 566685 & Mexico & 2     & 0     & 1 \\
-    RIMME0014 & 714151 & Breeders line & 6     & 0     & 1 \\
-    RIMME0017 & 699874 & Ayotlan & 8     & 0     & 1 \\
+    RIMME0006 & PI 566673 & Durango & 2	& 0 & 1 \\
+    RIMME0007 & PI 566680 & Guanajuato & 2 & 0 & 1 \\
+    RIMME0008 & PI 566681 & Michoacan & 2 & 0 & 1 \\
+    RIMME0009 & PI 566682 & Distrito Federal & 2 & 0 & 1 \\
+    RIMME0011 & PI 566685 & Mexico & 2     & 0     & 1 \\
+    RIMME0014 & Ames 28398 & TIL25 Jalisco  & 6     & 0     & 1 \\
+    RIMME0017 & Ames 699874 & Ayotlan & 8     & 0     & 1 \\
     RIMME0021 & N/A   & El Porvenir & 69    & 0.17 & 0.83 \\
@@
     RIMME0035 & N/A   & Xochimilco & 25    & 0     & 1 \\
-    RIMPA0001 & 87168 & El Salado & 4     & 0     & 1 \\
-    RIMPA0003 & 87171 & Mazatlan & 8 & 0.13 & 0.87 \\
-    RIMPA0017 & 87200 & N/A   & 4     & 0     & 1 \\
-    RIMPA0019 & 87213 & El Salado & 2     & 0.50   & 0.50 \\
-    RIMPA0029 & 87244 & N/A   & 2     & 0.50   & 0.50 \\
-    RIMPA0031 & 87249 & N/A   & 2     & 0.5   & 0.5 \\
-    RIMPA0035 & 87288 & Jalisco & 4     & 0     & 1 \\
-    RIMPA0040 & 288185 & Mexico & 4     & 0     & 1 \\
-    RIMPA0042 & 288187 & Guerrero & 4     & 0.25  & 0.75 \\
-    RIMPA0043 & 288188 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0045 & 288193 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0055 & 714152 & Breeders line & 2     & 0     & 1 \\
-    RIMPA0056 & 714153 & Breeders line & 2     & 0.50   & 0.50 \\
-    RIMPA0057 & 714154 & Breeders line & 2     & 0.50   & 0.50 \\
+    RIMPA0001 & Ames 21786 & El Salado & 4     & 0     & 1 \\
+    RIMPA0003 & Ames 21789 & Mazatlan & 8 & 0.13 & 0.87 \\
+    RIMPA0017 & Ames 21814  & N/A   & 4     & 0     & 1 \\
+    RIMPA0019 & Ames 21826 & El Salado & 2     & 0.50   & 0.50 \\
+    RIMPA0029 & Ames 21853 & N/A   & 2     & 0.50   & 0.50 \\
+    RIMPA0031 & Ames 21856 & N/A   & 2     & 0.5   & 0.5 \\
+    RIMPA0035 & Ames 21889 & Jalisco & 4     & 0     & 1 \\
+    RIMPA0040 & PI 384063 & Mexico & 4     & 0     & 1 \\
+    RIMPA0042 & PI 384065 & Guerrero & 4     & 0.25  & 0.75 \\
+    RIMPA0043 & PI 384066 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0045 & PI 384071 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0055 & Ames 28399 & TIL01, Michoacan & 2     & 0     & 1 \\
+    RIMPA0056 & Ames 28400 & TIL03 Jalisco & 2     & 0.50   & 0.50 \\
+    RIMPA0057 & Ames 28401 & TIL06 Guerrero & 2     & 0.50   & 0.50 \\
     RIMPA0058 & N/A   & N/A   & 4     & 0.50   & 0.50 \\
     RIMPA0059 & N/A   & N/A   & 4     & 1     & 0 \\
-    RIMPA0060 & 714157 & Breeders line & 2     & 0     & 1 \\
-    RIMPA0061 & 714158 & Breeders line & 4     & 0.5   & 0.5 \\
-    RIMPA0062 & 714159 & Breeders line & 4     & 0.5   & 0.5 \\
-    RIMPA0063 & 714160 & Breeders line & 4     & 0     & 1 \\
-    RIMPA0064 & 714161 & Breeders line & 3     & 0     & 1 \\
-    RIMPA0065 & 714162 & Breeders line & 4     & 0.25  & 0.75 \\
-    RIMPA0068 & 699861 & Jalisco, Mexico & 16    & 0     & 1 \\
-    RIMPA0069 & 699862 & Ixtlan & 14    & 0.14 & 0.86 \\
-    RIMPA0070 & 699863 & Benito Jaurez & 16    & 0     & 1 \\
-    RIMPA0071 & 699864 & Tuzantla & 28    & 0     & 1 \\
-    RIMPA0072 & 699865 & Tiquicheo & 16    & 0     & 1 \\
-    RIMPA0073 & 699866 & Tiquicheo & 16    & 0.12 & 0.88 \\
-    RIMPA0074 & 699867 & Huetamo & 12    & 0     & 1 \\
-    RIMPA0075 & 699868 & Huetamo & 2     & 0     & 1 \\
-    RIMPA0076 & 699869 & Huetamo & 4     & 0     & 1 \\
-    RIMPA0077 & 699870 & Caracuaro & 2     & 0     & 1 \\
-    RIMPA0078 & 699871 & Caracuaro & 2     & 0.5   & 0.5 \\
-    RIMPA0079 & 699872 & Villa Madero & 14    & 0     & 1 \\
-    RIMPA0080 & 699873 & Guachinango & 12    & 0     & 1 \\
-    RIMPA0081 & 699875 & Ameca & 16    & 0     & 1 \\
-    RIMPA0083 & 699877 & Tepoztlan & 14    & 0     & 1 \\
-    RIMPA0084 & 699878 & Tepoztlan & 16    & 0     & 1 \\
-    RIMPA0085 & 699879 & Miahuatlan & 16    & 0     & 1 \\
-    RIMPA0086 & 699880 & Miahuatlan & 16    & 0.06 & 0.94 \\
-    RIMPA0087 & 699881 & Tecoanapa & 24    & 0     & 1 \\
-    RIMPA0089 & 699883 & Guerrero & 12    & 0     & 1 \\
-    RIMPA0090 & 699884 & Guerrero & 10    & 0     & 1 \\
-    RIMPA0091 & 699885 & Guerrero & 16    & 0     & 1 \\
-    RIMPA0092 & 699886 & Guerrero & 10    & 0     & 1 \\
-    RIMPA0093 & 699887 & Guerrero & 26    & 0.08 & 0.92 \\
-    RIMPA0094 & 699888 & Guerrero & 2     & 0     & 1 \\
-    RIMPA0095 & 699889 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0096 & 699890 & Guerrero & 26    & 0.04 & 0.96 \\
-    RIMPA0097 & 699891 & Guerrero & 6     & 0     & 1 \\
-    RIMPA0098 & 699892 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0099 & 699893 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0100 & 699894 & Guerrero & 6     & 0     & 1 \\
-    RIMPA0101 & 699895 & Guerrero & 2     & 0     & 1 \\
-    RIMPA0103 & 699897 & Guerrero & 2     & 0     & 1 \\
-    RIMPA0104 & 699898 & Guerrero & 22    & 0.09 & 0.91 \\
-    RIMPA0105 & 699899 & Guerrero & 6     & 0     & 1 \\
-    RIMPA0106 & 699900 & Guerrero & 6     & 0.33 & 0.67 \\
-    RIMPA0107 & 699901 & Guerrero & 4     & 0     & 1 \\
-    RIMPA0108 & 699902 & Guerrero & 6     & 0     & 1 \\
-    RIMPA0109 & 699903 & Michoacan & 4     & 0.25  & 0.75 \\
-    RIMPA0110 & 699904 & Michoacan & 2     & 0     & 1 \\
-    RIMPA0111 & 699905 & Michoacan & 4     & 0     & 1 \\
-    RIMPA0112 & 699906 & Michoacan & 4     & 0.25  & 0.75 \\
-    RIMPA0114 & 699908 & Michoacan & 6     & 0.17 & 0.83 \\
-    RIMPA0116 & 699910 & Mexico & 2     & 0     & 1 \\
-    RIMPA0117 & 699911 & Mexico & 4     & 0     & 1 \\
-    RIMPA0118 & 699912 & Mexico & 6     & 0.17 & 0.83 \\
-    RIMPA0119 & 699913 & Mexico & 2     & 0     & 1 \\
-    RIMPA0120 & 699914 & Mexico & 1     & 1     & 0 \\
-    RIMPA0121 & 699915 & Mexico & 2     & 0     & 1 \\
-    RIMPA0128 & 699922 & Mexico & 2     & 0.5   & 0.5 \\
-    RIMPA0129 & 699923 & Michoacan & 2     & 0.5   & 0.5 \\
-    RIMPA0135 & 699929 & Nayarit & 24    & 0     & 1 \\
-    RIMPA0138 & 699932 & Jalisco & 2     & 0.5   & 0.5 \\
-    RIMPA0139 & 699933 & Jalisco & 1     & 1     & 0 \\
-    RIMPA0142 & 699936 & Colima & 18    & 0.44 & 0.56 \\
-    RIMPA0144 & 699938 & Jalisco & 2     & 1     & 0 \\
-    RIMPA0145 & 699939 & Michoacan & 1     & 1     & 0 \\
-    RIMPA0147 & 699941 & Jalisco & 1     & 1     & 0 \\
+    RIMPA0060 & Ames 28404 & TIL10 Guerrero & 2     & 0     & 1 \\
+    RIMPA0061 & Ames 28405 & TIL11 Nayarit & 4     & 0.5   & 0.5 \\
+    RIMPA0062 & Ames 28406 & TIL14 Jalisco & 4     & 0.5   & 0.5 \\
+    RIMPA0063 & Ames 28407 & TIL15 Guerrero & 4     & 0     & 1 \\
+    RIMPA0064 & Ames 28408 & TIL16 Guerrero & 3     & 0     & 1 \\
+    RIMPA0065 & Ames 28409 & TIL17 Guerrero & 4     & 0.25  & 0.75 \\
+    RIMPA0068 & Ames 28066 & Jalisco, Mexico & 16    & 0     & 1 \\
+    RIMPA0069 & Ames 28067 & Ixtlan & 14    & 0.14 & 0.86 \\
+    RIMPA0070 & Ames 28068 & Benito Jaurez & 16    & 0     & 1 \\
+    RIMPA0071 & Ames 28069 & Tuzantla & 28    & 0     & 1 \\
+    RIMPA0072 & Ames 28070 & Tiquicheo & 16    & 0     & 1 \\
+    RIMPA0073 & Ames 28071 & Tiquicheo & 16    & 0.12 & 0.88 \\
+    RIMPA0074 & Ames 28072 & Huetamo & 12    & 0     & 1 \\
+    RIMPA0075 & Ames 28073 & Huetamo & 2     & 0     & 1 \\
+    RIMPA0076 & Ames 28074 & Huetamo & 4     & 0     & 1 \\
+    RIMPA0077 & Ames 28075 & Caracuaro & 2     & 0     & 1 \\
+    RIMPA0078 & Ames 28076 & Caracuaro & 2     & 0.5   & 0.5 \\
+    RIMPA0079 & Ames 28077 & Villa Madero & 14    & 0     & 1 \\
+    RIMPA0080 & Ames 28078 & Guachinango & 12    & 0     & 1 \\
+    RIMPA0081 & Ames 28080 & Ameca & 16    & 0     & 1 \\
+    RIMPA0083 & Ames 28082 & Tepoztlan & 14    & 0     & 1 \\
+    RIMPA0084 & Ames 28083 & Tepoztlan & 16    & 0     & 1 \\
+    RIMPA0085 & Ames 28084 & Miahuatlan & 16    & 0     & 1 \\
+    RIMPA0086 & Ames 28085& Miahuatlan & 16    & 0.06 & 0.94 \\
+    RIMPA0087 & Ames 28086 & Tecoanapa & 24    & 0     & 1 \\
+    RIMPA0089 & Ames 28088 & Guerrero & 12    & 0     & 1 \\
+    RIMPA0090 & Ames 28089 & Guerrero & 10    & 0     & 1 \\
+    RIMPA0091 & Ames 28090 & Guerrero & 16    & 0     & 1 \\
+    RIMPA0092 & Ames 28091 & Guerrero & 10    & 0     & 1 \\
+    RIMPA0093 & Ames 28092 & Guerrero & 26    & 0.08 & 0.92 \\
+    RIMPA0094 & Ames 28093 & Guerrero & 2     & 0     & 1 \\
+    RIMPA0095 & Ames	28094 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0096 & Ames	28095 & Guerrero & 26    & 0.04 & 0.96 \\
+    RIMPA0097 & Ames	28096 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0098 & Ames	28097 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0099 & Ames	28098 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0100 & Ames	28099 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0101 & Ames	28100 & Guerrero & 2     & 0     & 1 \\
+    RIMPA0103 & Ames	28102 & Guerrero & 2     & 0     & 1 \\
+    RIMPA0104 & Ames	28103 & Guerrero & 22    & 0.09 & 0.91 \\
+    RIMPA0105 & Ames	28104 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0106 & Ames	28105 & Guerrero & 6     & 0.33 & 0.67 \\
+    RIMPA0107 & Ames	28106 & Guerrero & 4     & 0     & 1 \\
+    RIMPA0108 & Ames	28107 & Guerrero & 6     & 0     & 1 \\
+    RIMPA0109 & Ames	28108 & Michoacan & 4     & 0.25  & 0.75 \\
+    RIMPA0110 & Ames	28109 & Michoacan & 2     & 0     & 1 \\
+    RIMPA0111 & Ames	28110 & Michoacan & 4     & 0     & 1 \\
+    RIMPA0112 & Ames	28111 & Michoacan & 4     & 0.25  & 0.75 \\
+    RIMPA0114 & Ames	28113 & Michoacan & 6     & 0.17 & 0.83 \\
+    RIMPA0116 & Ames	28115 & Mexico & 2     & 0     & 1 \\
+    RIMPA0117 & Ames	28116 & Mexico & 4     & 0     & 1 \\
+    RIMPA0118 & Ames	28117 & Mexico & 6     & 0.17 & 0.83 \\
+    RIMPA0119 & Ames	28118 & Mexico & 2     & 0     & 1 \\
+    RIMPA0120 & Ames	28119 & Mexico & 1     & 1     & 0 \\
+    RIMPA0121 & Ames	28120 & Mexico & 2     & 0     & 1 \\
+    RIMPA0128 & Ames	28127 & Mexico & 2     & 0.5   & 0.5 \\
+    RIMPA0129 & Ames	28128 & Michoacan & 2     & 0.5   & 0.5 \\
+    RIMPA0135 & Ames	28134 & Nayarit & 24    & 0     & 1 \\
+    RIMPA0138 & Ames	28137 & Jalisco & 2     & 0.5   & 0.5 \\
+    RIMPA0139 & Ames	28138 & Jalisco & 1     & 1     & 0 \\
+    RIMPA0142 & Ames	28141 & Colima & 18    & 0.44 & 0.56 \\
+    RIMPA0144 & Ames	28143 & Jalisco & 2     & 1     & 0 \\
+    RIMPA0145 & Ames	28144 & Michoacan & 1     & 1     & 0 \\
+    RIMPA0147 & Ames	28146 & Jalisco & 1     & 1     & 0 \\
     RIMPA0155 & N/A   & Jalisco & 73    & 0.01 & 0.99 \\
@@
     RIMPA0159 & N/A   & Jalisco & 26    & 0     & 1 \\
-    RIMPA0162 & 21785 & N/A   & 4     & 0     & 1 \\
+    RIMPA0162 & Ames	21785 & N/A   & 4     & 0     & 1 \\
     \hline
@@
 \showrowcolors
-    RIMMA0066 & 2     & 1 \\
-    RIMMA0075 & 2     & 1 \\
-    RIMMA0077 & 2     & 1 \\
-    RIMMA0079 & 2     & 1 \\
-    RIMMA0081 & 2     & 1 \\
-    RIMMA0084 & 2     & 1 \\
-    RIMMA0086 & 2     & 1 \\
-    RIMMA0088 & 2     & 1 \\
-    RIMMA0089 & 2     & 1 \\
-    RIMMA0090 & 2     & 1 \\
-    RIMMA0092 & 4     & 1 \\
-    RIMMA0094 & 4     & 1 \\
-    RIMMA0097 & 2     & 1 \\
-    RIMMA0099 & 2     & 1 \\
-    RIMMA0100 & 2     & 1 \\
-    RIMMA0101 & 2     & 1 \\
-    RIMMA0104 & 2     & 1 \\
-    RIMMA0108 & 2     & 1 \\
-    RIMMA0111 & 6     & 1 \\
-    RIMMA0115 & 2     & 1 \\
-    RIMMA0117 & 2     & 1 \\
-    RIMMA0130 & 2     & 1 \\
-    RIMMA0133 & 2     & 1 \\
-    RIMMA0134 & 2     & 1 \\
-    RIMMA0135 & 2     & 1 \\
-    RIMMA0142 & 2     & 0.5 \\
-    RIMMA0143 & 4     & 1 \\
-    RIMMA0146 & 4     & 1 \\
-    RIMMA0149 & 2     & 1 \\
-    RIMMA0152 & 2     & 1 \\
-    RIMMA0153 & 2     & 1 \\
-    RIMMA0154 & 2     & 1 \\
-    RIMMA0155 & 2     & 1 \\
-    RIMMA0156 & 2     & 1 \\
-    RIMMA0157 & 2     & 1 \\
-    RIMMA0158 & 2     & 1 \\
-    RIMMA0159 & 2     & 1 \\
-    RIMMA0160 & 2     & 1 \\
-    RIMMA0162 & 2     & 1 \\
-    RIMMA0166 & 2     & 1 \\
-    RIMMA0167 & 2     & 1 \\
-    RIMMA0168 & 2     & 1 \\
-    RIMMA0169 & 2     & 1 \\
-    RIMMA0172 & 2     & 1 \\
-    RIMMA0174 & 4     & 1 \\
-    RIMMA0177 & 2     & 1 \\
-    RIMMA0178 & 2     & 1 \\
-    RIMMA0179 & 2     & 1 \\
-    RIMMA0181 & 2     & 1 \\
-    RIMMA0183 & 2     & 1 \\
-    RIMMA0184 & 2     & 1 \\
-    RIMMA0186 & 2     & 1 \\
-    RIMMA0187 & 1     & 1 \\
-    RIMMA0188 & 2     & 1 \\
-    RIMMA0195 & 2     & 1 \\
-    RIMMA0196 & 2     & 1 \\
-    RIMMA0197 & 2     & 1 \\
-    RIMMA0198 & 2     & 1 \\
-    RIMMA0199 & 2     & 1 \\
-    RIMMA0200 & 2     & 1 \\
-    RIMMA0202 & 2     & 1 \\
-    RIMMA0203 & 2     & 1 \\
-    RIMMA0206 & 2     & 1 \\
-    RIMMA0208 & 2     & 1 \\
-    RIMMA0209 & 2     & 1 \\
-    RIMMA0210 & 2     & 1 \\
-    RIMMA0212 & 2     & 1 \\
-    RIMMA0213 & 2     & 1 \\
-    RIMMA0214 & 2     & 1 \\
-    RIMMA0217 & 2     & 1 \\
-    RIMMA0218 & 2     & 1 \\
-    RIMMA0220 & 2     & 1 \\
-    RIMMA0221 & 2     & 1 \\
-    RIMMA0222 & 2     & 1 \\
-    RIMMA0223 & 2     & 1 \\
-    RIMMA0226 & 2     & 1 \\
-    RIMMA0227 & 2     & 1 \\
-    RIMMA0228 & 2     & 1 \\
-    RIMMA0229 & 2     & 1 \\
-    RIMMA0230 & 2     & 1 \\
-    RIMMA0232 & 2     & 1 \\
-    RIMMA0233 & 2     & 1 \\
-    RIMMA0235 & 2     & 0.5 \\
-    RIMMA0242 & 2     & 1 \\
-    RIMMA0243 & 2     & 1 \\
-    RIMMA0247 & 4     & 1 \\
-    RIMMA0248 & 2     & 1 \\
-    RIMMA0249 & 2     & 1 \\
-    RIMMA0252 & 2     & 1 \\
-    RIMMA0253 & 2     & 1 \\
-    RIMMA0254 & 2     & 1 \\
-    RIMMA0256 & 2     & 1 \\
-    RIMMA0257 & 2     & 1 \\
-    RIMMA0258 & 2     & 1 \\
-    RIMMA0259 & 2     & 1 \\
-    RIMMA0260 & 2     & 1 \\
-    RIMMA0262 & 2     & 1 \\
-    RIMMA0263 & 2     & 1 \\
-    RIMMA0264 & 2     & 1 \\
-    RIMMA0265 & 2     & 1 \\
-    RIMMA0268 & 2     & 1 \\
-    RIMMA0269 & 2     & 1 \\
-    RIMMA0270 & 2     & 1 \\
-    RIMMA0272 & 2     & 1 \\
-    RIMMA0275 & 2     & 1 \\
-    RIMMA0276 & 2     & 1 \\
-    RIMMA0277 & 2     & 1 \\
-    RIMMA0279 & 2     & 1 \\
-    RIMMA0280 & 2     & 1 \\
-    RIMMA0283 & 2     & 1 \\
-    RIMMA0285 & 2     & 1 \\
-    RIMMA0288 & 2     & 1 \\
-    RIMMA0290 & 2     & 1 \\
-    RIMMA0291 & 2     & 1 \\
-    RIMMA0292 & 2     & 1 \\
-    RIMMA0293 & 2     & 1 \\
-    RIMMA0298 & 2     & 1 \\
-    RIMMA0302 & 2     & 1 \\
-    RIMMA0305 & 2     & 1 \\
-    RIMMA0310 & 2     & 1 \\
-    RIMMA0312 & 2     & 1 \\
-    RIMMA0320 & 2     & 1 \\
-    RIMMA0322 & 2     & 1 \\
-    RIMMA0324 & 2     & 1 \\
-    RIMMA0334 & 2     & 1 \\
-    RIMMA0336 & 2     & 1 \\
-    RIMMA0337 & 2     & 1 \\
-    RIMMA0338 & 2     & 1 \\
-    RIMMA0339 & 2     & 1 \\
-    RIMMA0340 & 2     & 1 \\
-    RIMMA0341 & 2     & 1 \\
-    RIMMA0342 & 2     & 1 \\
-    RIMMA0344 & 2     & 1 \\
-    RIMMA0346 & 2     & 1 \\
-    RIMMA0347 & 2     & 1 \\
-    RIMMA0350 & 2     & 1 \\
-    RIMMA0351 & 2     & 1 \\
-    RIMMA0352 & 2     & 1 \\
-    RIMMA0357 & 2     & 1 \\
-    RIMMA0360 & 18    & 1 \\
-    RIMMA0362 & 8     & 1 \\
-    RIMMA0363 & 16    & 1 \\
-    RIMMA0370 & 17    & 1 \\
-    RIMMA0372 & 24    & 1 \\
-    RIMMA0373 & 22    & 1 \\
-    RIMMA0408 & 2     & 1 \\
-    RIMMA0411 & 1     & 1 \\
-    RIMMA0413 & 1     & 1 \\
-    RIMMA0414 & 1     & 1 \\
-    RIMMA0424 & 8     & 0.9 \\
-    RIMMA0432 & 1     & 1 \\
-    RIMMA0434 & 2     & 0.5 \\
-    RIMMA0435 & 1     & 1 \\
-    RIMMA0442 & 2     & 1 \\
-    RIMMA0443 & 2     & 1 \\
-    RIMMA0444 & 2     & 1 \\
-    RIMMA0445 & 2     & 1 \\
-    RIMMA0446 & 2     & 1 \\
-    RIMMA0447 & 2     & 1 \\
-    RIMMA0448 & 2     & 1 \\
-    RIMMA0449 & 2     & 1 \\
-    RIMMA0450 & 2     & 1 \\
-    RIMMA0451 & 2     & 1 \\
-    RIMMA0452 & 2     & 1 \\
-    RIMMA0453 & 2     & 1 \\
-    RIMMA0454 & 2     & 1 \\
-    RIMMA0455 & 2     & 1 \\
-    RIMMA0456 & 2     & 1 \\
-    RIMMA0457 & 2     & 1 \\
-    RIMMA0458 & 2     & 1 \\
-    RIMMA0459 & 2     & 1 \\
-    RIMMA0483 & 2     & 1 \\
-    RIMMA0490 & 2     & 1 \\
-    RIMMA0515 & 2     & 1 \\
-    RIMMA0537 & 2     & 1 \\
-    RIMMA0550 & 2     & 1 \\
-    RIMMA0553 & 2     & 1 \\
-    RIMMA0559 & 2     & 1 \\
-    RIMMA0561 & 2     & 1 \\
-    RIMMA0562 & 2     & 1 \\
-    RIMMA0571 & 2     & 1 \\
-    RIMMA0572 & 2     & 1 \\
-    RIMMA0577 & 2     & 1 \\
-    RIMMA0579 & 2     & 1 \\
-    RIMMA0590 & 2     & 1 \\
-    RIMMA0591 & 2     & 1 \\
-    RIMMA0592 & 2     & 1 \\
-    RIMMA0593 & 2     & 1 \\
-    RIMMA0594 & 2     & 1 \\
-    RIMMA0595 & 2     & 1 \\
-    RIMMA0596 & 2     & 1 \\
-    RIMMA0597 & 2     & 1 \\
-    RIMMA0598 & 2     & 1 \\
-    RIMMA0599 & 2     & 1 \\
-    RIMMA0600 & 2     & 1 \\
-    RIMMA0601 & 2     & 1 \\
-    RIMMA0602 & 2     & 1 \\
-    RIMMA0603 & 2     & 1 \\
-    RIMMA0604 & 2     & 1 \\
-    RIMMA0605 & 2     & 1 \\
-    RIMMA0606 & 2     & 1 \\
-    RIMMA0607 & 2     & 1 \\
-    RIMMA0608 & 2     & 1 \\
-    RIMMA0609 & 2     & 1 \\
-    RIMMA0610 & 2     & 1 \\
-    RIMMA0611 & 2     & 1 \\
-    RIMMA0612 & 2     & 1 \\
-    RIMMA0613 & 2     & 1 \\
-    RIMMA0622 & 2     & 0.5 \\
-    RIMMA0624 & 1     & 1 \\
-    RIMMA0629 & 2     & 0.5 \\
-    RIMMA0631 & 2     & 0.5 \\
-    RIMMA0659 & 2     & 1 \\
-    RIMMA0660 & 2     & 1 \\
-    RIMMA0678 & 1     & 1 \\
-    RIMMA0681 & 2     & 1 \\
-    RIMMA0683 & 2     & 1 \\
-    RIMMA0684 & 2     & 1 \\
-    RIMMA0685 & 4     & 1 \\
-    RIMMA0693 & 2     & 1 \\
-    RIMMA0694 & 2     & 1 \\
-    RIMMA0695 & 1     & 1 \\
-    RIMMA0699 & 2     & 1 \\
-    RIMMA0704 & 2     & 1 \\
-    RIMMA0706 & 2     & 1 \\
-    RIMMA0711 & 1     & 1 \\
-    RIMMA0713 & 2     & 1 \\
-    RIMMA0715 & 2     & 1 \\
-    RIMMA0717 & 2     & 1 \\
-    RIMMA0718 & 2     & 1 \\
-    RIMMA0723 & 2     & 1 \\
-    RIMMA0724 & 2     & 1 \\
-    RIMMA0725 & 1     & 1 \\
-    RIMMA0728 & 1     & 1 \\
-    RIMMA0732 & 2     & 1 \\
-    RIMMA0734 & 2     & 1 \\
-    RIMMA0738 & 2     & 1 \\
-    RIMMA0739 & 2     & 1 \\
-    RIMMA0744 & 2     & 1 \\
-    RIMMA0747 & 1     & 1 \\
-    RIMMA0748 & 2     & 1 \\
-    RIMMA0749 & 2     & 1 \\
-    RIMMA0750 & 2     & 0.5 \\
-    RIMMA0751 & 1     & 1 \\
-    RIMMA0752 & 1     & 1 \\
-    RIMMA0753 & 2     & 1 \\
-    RIMMA0755 & 2     & 1 \\
+    RIMMA0066 & Ames	4327 & 2     & 1 \\
+    RIMMA0075 & Ames	19309 & 2     & 1 \\
+    RIMMA0077 & Ames 19313 & 2     & 1 \\
+    RIMMA0079 & Ames 19321 & 2     & 1 \\
+    RIMMA0081 & Ames 20190 & 2     & 1 \\
+    RIMMA0084 & Ames 22439 & 2     & 1 \\
+    RIMMA0086 & Ames 22748 & 2     & 1 \\
+    RIMMA0088 & Ames 22756 & 2     & 1 \\
+    RIMMA0089 & Ames 23392 & 2     & 1 \\
+    RIMMA0090 & Ames 23423 & 2     & 1 \\
+    RIMMA0092 & Ames 23427 & 4     & 1 \\
+    RIMMA0094 & Ames 23447 & 4     & 1 \\
+    RIMMA0097 & Ames 23488 & 2     & 1 \\
+    RIMMA0099 & Ames 23509 & 2     & 1 \\
+    RIMMA0100 & Ames 23518 & 2     & 1 \\
+    RIMMA0101 & Ames 23519 & 2     & 1 \\
+    RIMMA0104 & Ames 24590 & 2     & 1 \\
+    RIMMA0108 & Ames 26022 & 2     & 1 \\
+    RIMMA0111 & Ames 26770 & 6     & 1 \\
+    RIMMA0115 & Ames 26798 & 2     & 1 \\
+    RIMMA0117 & Ames 27045 & 2     & 1 \\
+    RIMMA0130 & Ames 28336 & 2     & 1 \\
+    RIMMA0133 & NSL	8581 & 2     & 1 \\
+    RIMMA0134 & NSL	22632 & 2     & 1 \\
+    RIMMA0135 & NSL	22634 & 2     & 1 \\
+    RIMMA0142 & NSL	30060 & 2     & 0.5 \\
+    RIMMA0143 & NSL	30065 & 4     & 1 \\
+    RIMMA0146 & NSL	30863 & 4     & 1 \\
+    RIMMA0149 & NSL	30869 & 2     & 1 \\
+    RIMMA0152 & NSL	32722 & 2     & 1 \\
+    RIMMA0153 & NSL	32728 & 2     & 1 \\
+    RIMMA0154 & NSL	32732 & 2     & 1 \\
+    RIMMA0155 & NSL	42809 & 2     & 1 \\
+    RIMMA0156 & NSL	42872 & 2     & 1 \\
+    RIMMA0157 & NSL	42874 & 2     & 1 \\
+    RIMMA0158 & NSL	42875 & 2     & 1 \\
+    RIMMA0159 & NSL	42878 & 2     & 1 \\
+    RIMMA0160 & NSL	65864 & 2     & 1 \\
+    RIMMA0162 & NSL	65866 & 2     & 1 \\
+    RIMMA0166 & PI	213696 & 2     & 1 \\
+    RIMMA0167 & PI	213697 & 2     & 1 \\
+    RIMMA0168 & PI	213698 & 2     & 1 \\
+    RIMMA0169 & PI	213699 & 2     & 1 \\
+    RIMMA0172 & PI	213705 & 2     & 1 \\
+    RIMMA0174 & PI	213708 & 4     & 1 \\
+    RIMMA0177 & PI	213715 & 2     & 1 \\
+    RIMMA0178 & PI	213716 & 2     & 1 \\
+    RIMMA0179 & PI	213717 & 2     & 1 \\
+    RIMMA0181 & PI	213719 & 2     & 1 \\
+    RIMMA0183 & PI	213721 & 2     & 1 \\
+    RIMMA0184 & PI	213722 & 2     & 1 \\
+    RIMMA0186 & PI	213724 & 2     & 1 \\
+    RIMMA0187 & PI	213725 & 1     & 1 \\
+    RIMMA0188 & PI	213727 & 2     & 1 \\
+    RIMMA0195 & PI	213781 & 2     & 1 \\
+    RIMMA0196 & PI	213793 & 2     & 1 \\
+    RIMMA0197 & PI	214197 & 2     & 1 \\
+    RIMMA0198 & PI	214293 & 2     & 1 \\
+    RIMMA0199 & PI	214294 & 2     & 1 \\
+    RIMMA0200 & PI	214295 & 2     & 1 \\
+    RIMMA0202 & PI	217407 & 2     & 1 \\
+    RIMMA0203 & PI	217408 & 2     & 1 \\
+    RIMMA0206 & PI	217473 & 2     & 1 \\
+    RIMMA0208 & PI	218005 & 2     & 1 \\
+    RIMMA0209 & PI	218195 & 2     & 1 \\
+    RIMMA0210 & PI 219871 & 2     & 1 \\
+    RIMMA0212 & PI 219882 & 2     & 1 \\
+    RIMMA0213 & PI 219883 & 2     & 1 \\
+    RIMMA0214 & PI 219884 & 2     & 1 \\
+    RIMMA0217 & PI 221872 & 2     & 1 \\
+    RIMMA0218 & PI 221873 & 2     & 1 \\
+    RIMMA0220 & PI 221876 & 2     & 1 \\
+    RIMMA0221 & PI 221877 & 2     & 1 \\
+    RIMMA0222 & PI 221878 & 2     & 1 \\
+    RIMMA0223 & PI	221880 & 2     & 1 \\
+    RIMMA0226 & PI	221889 & 2     & 1 \\
+    RIMMA0227 & PI	222314 & 2     & 1 \\
+    RIMMA0228 & PI	222315 & 2     & 1 \\
+    RIMMA0229 & PI	222316 & 2     & 1 \\
+    RIMMA0230 & PI	222317 & 2     & 1 \\
+    RIMMA0232 & PI	222469 &  2     & 1 \\
+    RIMMA0233 & PI	222470 & 2     & 1 \\
+    RIMMA0235 & PI	222474 & 2     & 0.5 \\
+    RIMMA0242 & PI	222614 & 2     & 1 \\
+    RIMMA0243 & PI	222615 & 2     & 1 \\
+    RIMMA0247 & PI	222639 & 4     & 1 \\
+    RIMMA0248 & PI	222640 & 2     & 1 \\
+    RIMMA0249 & PI	222641 & 2     & 1 \\
+    RIMMA0252 & PI	233002 & 2     & 1 \\
+    RIMMA0253 & PI	233006 & 2     & 1 \\
+    RIMMA0254 & PI	233008 & 2     & 1 \\
+    RIMMA0256 & PI	237000 & 2     & 1 \\
+    RIMMA0257 & PI	257506 & 2     & 1 \\
+    RIMMA0258 & PI 267179 & 2     & 1 \\
+    RIMMA0259 & PI 269743 & 2     & 1 \\
+    RIMMA0260 & PI 269744 & 2     & 1 \\
+    RIMMA0262 & PI 270297 & 2     & 1 \\
+    RIMMA0263 & PI 278713 & 2     & 1 \\
+    RIMMA0264 & PI 278716 & 2     & 1 \\
+    RIMMA0265 & PI 278717 & 2     & 1 \\
+    RIMMA0268 & PI 278721 & 2     & 1 \\
+    RIMMA0269 & PI 278724 & 2     & 1 \\
+    RIMMA0270 & PI 280061 & 2     & 1 \\
+    RIMMA0272 & PI 280853 & 2     & 1 \\
+    RIMMA0275 & PI 311232 & 2     & 1 \\
+    RIMMA0276 & PI 311235 & 2     & 1 \\
+    RIMMA0277 & PI 311236 & 2     & 1 \\
+    RIMMA0279 & PI 311241 & 2     & 1 \\
+    RIMMA0280 & PI 311243 & 2     & 1 \\
+    RIMMA0283 & PI 401757 & 2     & 1 \\
+    RIMMA0285 & PI 414176 & 2     & 1 \\
+    RIMMA0288 & PI 452032 & 2     & 1 \\
+    RIMMA0290 & PI 452045 & 2     & 1 \\
+    RIMMA0291 & PI 452046 & 2     & 1 \\
+    RIMMA0292 & PI 452047 & 2     & 1 \\
+    RIMMA0293 & PI 452048 & 2     & 1 \\
+    RIMMA0298 & PI 483273 & 2     & 1 \\
+    RIMMA0302 & PI 540784 & 2     & 1 \\
+    RIMMA0305 & PI 548792 & 2     & 1 \\
+    RIMMA0310 & PI 550462 & 2     & 1 \\
+    RIMMA0312 & PI 550466 & 2     & 1 \\
+    RIMMA0320 & PI 583896 & 2     & 1 \\
+    RIMMA0322 & PI 583898 & 2     & 1 \\
+    RIMMA0324 & PI 587125 & 2     & 1 \\
+    RIMMA0334 & PI 587153 & 2     & 1 \\
+    RIMMA0336 & PI 607519 & 2     & 1 \\
+    RIMMA0337 & PI 607522 & 2     & 1 \\
+    RIMMA0338 & PI 607593 & 2     & 1 \\
+    RIMMA0339 & PI 607599 & 2     & 1 \\
+    RIMMA0340 & PI 607602 & 2     & 1 \\
+    RIMMA0341 & PI 608465 & 2     & 1 \\
+    RIMMA0342 & PI 608502 & 2     & 1 \\
+    RIMMA0344 & PI 608539 & 2     & 1 \\
+    RIMMA0346 & PI 608618 & 2     & 1 \\
+    RIMMA0347 & PI 608619 & 2     & 1 \\
+    RIMMA0350 & PI 608646 & 2     & 1 \\
+    RIMMA0351 & PI 608652 & 2     & 1 \\
+    RIMMA0352 & PI 608764 & 2     & 1 \\
+    RIMMA0357 & PI 608777 & 2     & 1 \\
+    RIMMA0360 & N/A & 18    & 1 \\
+    RIMMA0362 & N/A & 8     & 1 \\
+    RIMMA0363 & N/A & 16    & 1 \\
+    RIMMA0370 & N/A & 17    & 1 \\
+    RIMMA0372 & N/A & 24    & 1 \\
+    RIMMA0373 & N/A & 22    & 1 \\
+    RIMMA0408 & PI	478967 & 2     & 1 \\
+    RIMMA0411 & PI	479044 & 1     & 1 \\
+    RIMMA0413 & PI	484413 & 1     & 1 \\
+    RIMMA0414 & PI	484421 & 1     & 1 \\
+    RIMMA0424 & PI	485120 & 8     & 0.9 \\
+    RIMMA0432 & PI	489126 & 1     & 1 \\
+    RIMMA0434 & PI	490921 & 2     & 0.5 \\
+    RIMMA0435 & PI	490973 & 1     & 1 \\
+    RIMMA0442 & PI	537097 & 2     & 1 \\
+    RIMMA0443 & PI	537099 & 2     & 1 \\
+    RIMMA0444 & PI	538009 & 2     & 1 \\
+    RIMMA0445 & PI	538011 & 2     & 1 \\
+    RIMMA0446 & PI	539920 & 2     & 1 \\
+    RIMMA0447 & PI	539921 & 2     & 1 \\
+    RIMMA0448 & PI	539922 & 2     & 1 \\
+    RIMMA0449 & PI	539923 & 2     & 1 \\
+    RIMMA0450 & PI	539924 & 2     & 1 \\
+    RIMMA0451 & PI	539926 & 2     & 1 \\
+    RIMMA0452 & PI	539927 & 2     & 1 \\
+    RIMMA0453 & PI	543843 & 2     & 1 \\
+    RIMMA0454 & PI	543846 & 2     & 1 \\
+    RIMMA0455 & PI	543847 & 2     & 1 \\
+    RIMMA0456 & PI	543849 & 2     & 1 \\
+    RIMMA0457 & PI	543850 & 2     & 1 \\
+    RIMMA0458 & PI	547086 & 2     & 1 \\
+    RIMMA0459 & PI	547088 & 2     & 1 \\
+    RIMMA0483 & PI	600954 & 2     & 1 \\
+    RIMMA0490 & PI	601003 & 2     & 1 \\
+    RIMMA0515 & PI	601270 & 2     & 1 \\
+    RIMMA0537 & PI	601416 & 2     & 1 \\
+    RIMMA0550 & PI	601468 & 2     & 1 \\
+    RIMMA0553 & PI	601489 & 2     & 1 \\
+    RIMMA0559 & PI	601495 & 2     & 1 \\
+    RIMMA0561 & PI	601497 & 2     & 1 \\
+    RIMMA0562 & PI	601498 & 2     & 1 \\
+    RIMMA0571 & PI	601562 & 2     & 1 \\
+    RIMMA0572 & PI	601563 & 2     & 1 \\
+    RIMMA0577 & PI	601568 & 2     & 1 \\
+    RIMMA0579 & PI	601570 & 2     & 1 \\
+    RIMMA0590 & PI	601724 & 2     & 1 \\
+    RIMMA0591 & PI	601725 & 2     & 1 \\
+    RIMMA0592 & PI	601729 & 2     & 1 \\
+    RIMMA0593 & PI	601773 & 2     & 1 \\
+    RIMMA0594 & PI	601774 & 2     & 1 \\
+    RIMMA0595 & PI	601775 & 2     & 1 \\
+    RIMMA0596 & PI	601776 & 2     & 1 \\
+    RIMMA0597 & PI	601777 & 2     & 1 \\
+    RIMMA0598 & PI	601778 & 2     & 1 \\
+    RIMMA0599 & PI	601779 & 2     & 1 \\
+    RIMMA0600 & PI	601780 & 2     & 1 \\
+    RIMMA0601 & PI	601781 & 2     & 1 \\
+    RIMMA0602 & PI	601782 & 2     & 1 \\
+    RIMMA0603 & PI	601783 & 2     & 1 \\
+    RIMMA0604 & PI	601784 & 2     & 1 \\
+    RIMMA0605 & PI	601785 & 2     & 1 \\
+    RIMMA0606 & PI	601786 & 2     & 1 \\
+    RIMMA0607 & PI	601787 & 2     & 1 \\
+    RIMMA0608 & PI	601788 & 2     & 1 \\
+    RIMMA0609 & PI	601789 & 2     & 1 \\
+    RIMMA0610 & PI	601790 & 2     & 1 \\
+    RIMMA0611 & PI	601791 & 2     & 1 \\
+    RIMMA0612 & PI	601792 & 2     & 1 \\
+    RIMMA0613 & PI	601808 & 2     & 1 \\
+    RIMMA0622 & PI	645837 & 2     & 0.5 \\
+    RIMMA0624 & PI	645930 & 1     & 1 \\
+    RIMMA0629 & PI	646057 & 2     & 0.5 \\
+    RIMMA0631 & PI	646072 & 2     & 0.5 \\
+    RIMMA0659 & NSL	286729 & 2     & 1 \\
+    RIMMA0660 & PI	485451 & 2     & 1 \\
+    RIMMA0678 & NSL	285833 & 1     & 1 \\
+    RIMMA0681 & PI	628446 & 2     & 1 \\
+    RIMMA0683 & PI	483568 & 2     & 1 \\
+    RIMMA0684 & Ames	28482 & 2     & 1 \\
+    RIMMA0685 & NSL	287623 & 4     & 1 \\
+    RIMMA0693 & PI	487959 & 2     & 1 \\
+    RIMMA0694 & Ames	28487 & 2     & 1 \\
+    RIMMA0695 & Ames	28567 & 1     & 1 \\
+    RIMMA0699 & Ames	28465 & 2     & 1 \\
+    RIMMA0704 & Ames	28623 & 2     & 1 \\
+    RIMMA0706 & Ames	28615 & 2     & 1 \\
+    RIMMA0711 & Ames	29075 & 1     & 1 \\
+    RIMMA0713 & PI	489399 & 2     & 1 \\
+    RIMMA0715 & Ames	28454 & 2     & 1 \\
+    RIMMA0717 & PI	483926 & 2     & 1 \\
+    RIMMA0718 & PI	484805 & 2     & 1 \\
+    RIMMA0723 & Ames	28565 & 2     & 1 \\
+    RIMMA0724 & PI	490826& 2     & 1 \\
+    RIMMA0725 & PI	490827 & 1     & 1 \\
+    RIMMA0728 & PI	484969 & 1     & 1 \\
+    RIMMA0732 & Ames 28456 & 2     & 1 \\
+    RIMMA0734 & Ames 26028 & 2     & 1 \\
+    RIMMA0738 & N/A & 2     & 1 \\
+    RIMMA0739 & N/A & 2     & 1 \\
+    RIMMA0744 & N/A & 2     & 1 \\
+    RIMMA0747 & N/A & 1     & 1 \\
+    RIMMA0748 & N/A & 2     & 1 \\
+    RIMMA0749 & N/A & 2     & 1 \\
+    RIMMA0750 & N/A & 2     & 0.5 \\
+    RIMMA0751 & N/A & 1     & 1 \\
+    RIMMA0752 & N/A & 1     & 1 \\
+    RIMMA0753 & N/A & 2     & 1 \\
+    RIMMA0755 & N/A & 2     & 1 \\
     \hline
